### 2008-12-15 pkg/diseasemapping/R/New Folder/SpatialFun/SpatialFun.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@18 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{article}
\usepackage{graphicx}
\textwidth 6.0 true in
\textheight 9 true in
\topmargin -1.0 true in
\headsep 0.375 true in
\oddsidemargin 0.0 true in
\evensidemargin 0.5 true in

\title{The Spatial Model Functions}
\def\ffmc{\mbox{ffmc}}
\def\wind{\mbox{wind}}
\def\ROS{\mbox{ROS}}
\def\ISI{\mbox{ISI}}
\def\BUI{\mbox{BUI}}
\begin{document}
\maketitle

\large
\section{Introduction}

\section{Method}
\subsection{The Poisson Model}
The followings are the theoretical functions:

Let $Y_k$ denotes the number of cancer cases in group k.
$P_k$ denotes the number of people (population size).
$\lambda_k$ denotes the rate of getting cancer.

We have
     \[Y_k \sim Poisson (P_k\lambda_k),\]  we won't put time under our consideration at the beginning.  %or the function should be Y_k ~ Poisson(\lambda_i T)
     \[\log(\lambda_k) = \mu +  x_k\beta + U_k + V_k,\]
where      $\mu$ is the intercept, it could be $\log(P_k)$;

           $x_k$ is the covariates i.e income variable;

           $\beta$ is the log relative risk;

           $U_k$ is the spatial random variable and \[U_k \sim N(0, \sigma^2_u)\]

           $V_k$ is the non-spatial random variable and \[V_k \sim N(0, \sigma^2_v)\]

Therefore, we will fit a $glm$ (generalized linear model) as:
    \[model = glm( cases \sim offset(logpop) + factor(age) factor(sex) )\]

\subsection{The R Functions}
Currently the diseasemapping package contains $6$ functions:

\begin{enumerate}
\item

\begin{verbatim}
''formatCases'' <- function(casedata)
\end{verbatim}

We get the data sets from different kinds of sources. Therefore, we need to clean and format the data set
for the future manipulation. The function dealing with two scenarios: when there is  \texttt{group} column in the data set
or when there is  \small \texttt{AGE\_SEX\_GROUP} column in data set. Essentially, we want to add in two
columns : sex and age.
\large

Group is usually a column looks like:
\begin{table}[ht]
\begin{center}
\begin{tabular}{|c|c|}
\hline
 $M0\_5$ \\
 $M5\_9$ \\
 $M10\_14$ \\
  \vdots  \\
 $M85PLUS$ \\
\hline
\end{tabular}
\end{center}
\end{table}

AGE\_SEX\_Group is usually a column of numbers composed with $3$ digits:
Male are the numbers begin with $1$, the rest two digits are corresponding to the age.
i.e:

\begin{table}[ht]
\begin{center}
\begin{tabular}{|c|c|}
\hline
$100$  &  $M0\_4$ \\
$101$  &  $M5\_9$ \\
$102$  &  $M10\_14$ \\
\vdots  &  \vdots  \\
$117$  &  $M85PLUS$ \\
\hline
\end{tabular}
\end{center}
\end{table}

Female is the same, but starting at $200$.

\item
\begin{verbatim}
''formatPopulation'' <- function(popdata)
\end{verbatim}

This function is similar to the formatCases function.
In some of the population data set, there will be a group of the people who are
age $85$ and plus, marked as: \small $M85PLUS$ or $F85PLUS$ \large for male and female respectively.
The first part of the formatPopulation function is to change them to \small $M85\_89$ or $F85\_89$,
\large so that they will have the same format as the other sex and age group, and that will
simplifies our manipulation later.
The second part of the function is to reshape the population data set using the function
\texttt{reshape}. So that it changes the population data set from the wide format
to the long format, and it will have the same format as the case data set. This step make
the merge of case data set and population data set easier in the future calculations.

\item 
\begin{verbatim}
''getRates'' <-
function(casedata, popdata, formula, family=poisson,
minimumAge=0, maximumAge=100, S=c("M", "F"))
\end{verbatim}

This is the function we constructed to fit a Generalized Linear Model (\texttt{glm} function) to
the case data set and the population data set. Then we get out estimated coefficients for the future
prediction.
The argument \texttt{casedata} and \texttt{popdata} are the places where we put case data set
and population data set correspondingly. \texttt{formula} is the place we fit a formula function
in the \texttt{glm} model. i.e : $~age*sex $. \texttt{family} is the place we
choose what distribution we want to fit in the \texttt{glm} model. By default we set it to be Poisson model, it
can be binomial as well.
\texttt{$minimumAge=0$} and \texttt{$maximumAge=100$} are the arguments to control the sample's age range. 
By default it is $age \in ($0, 100$)$ , which includes all the cases. You can change the range to ($0$, $10$) 
for chose all the kids' data set.  \texttt{S} is in charge of control what kind of sex group we want to choose. 
By default it choose both the male and female.

\item 
\begin{verbatim}
''getSMR'' <-
function(model, population, cases=NULL, regionCode="CSDUID",
  regionCodeCases="CSD2006")
\end{verbatim}

Function \texttt{getSMR} is used to do the prediction using the model we get from the
\texttt{getRates} function. Eventually, it returns a shape file (if the population data set
we insert the function is a shape file) contains everything of the original population
data set and several columns : expected, logExpected, cases, SMR. expected are the expected
case numbers we estimated using the model we get from \texttt{getRates} function. logExpected
are the log values of the expected values. cases are the number of cases we got from the
case data set (if we fit a poisson distribution it will be the aggregate number of cases by
different Decent Areas. SMR is the rate of observed number of cases over expected number of cases.

The argument \texttt{model} is the glm model estimated coefficients we got from the \texttt{getRates} function.
\texttt{population} is the population data set, it can be a shape file. \texttt{cases} is the number of
case data set. \texttt{regionCodeCases} and \texttt{regionCode} are used to indicate the region for case data and
population data sets respectively.

\item 
\begin{verbatim}
''area'' <- function(sp)
\end{verbatim}
Function \texttt{area} is used to calculate the area of each region, return a vector of areas. The argument \texttt{sp} is a single 
spatial polygon object.
\item 
\begin{verbatim}
''mergeBugsData'' <- function(x, bugsSummary, by.x = NULL, newcol="mean", ...) 
\end{verbatim}
Function \texttt{mergeBugsData} is used to merge the results from the \texttt{bugs} function to the other data set i.e population data set.

The argument \texttt{x} is the data set we want the result from the \texttt{bugs} to merge to. \texttt{bugsSummary} is the simulation result
get from the \texttt{bugs} function. \texttt{by.x} is the common column in both the {x} and {bugsSummary} data set. \texttt{newcol} is the summary statistic that to be merged back to the data frame.   
\end{enumerate}

\subsection{The Examples}
We will take the ontario long cancer data set as an example. Due to confidential needs, we simulated the case (cancer) data set.  
For example, we have a case data set as:

> data(casedata)
> head(casedata)

library
\begin{table}[ht]
\begin{center}
\begin{tabular}{cccccc}
Year& $CSD2006$ & $CD2006$ & $PR2006$ & AGE\_SEX\_GROUP & Cases \\
$1999$&$3501005$&$3501$&$35$&$210$&$3$ \\
$1999$&$3501005$&$3501$&$35$&$213$&$1$ \\
$1999$&$3501005$&$3501$&$35$&$214$&$6$ \\
$1999$&$3501005$&$3501$&$35$&$215$&$1$ \\
$1999$&$3501005$&$3501$&$35$&$216$&$3$ \\
\end{tabular}
\end{center}
\end{table}

> formatCases(casedata)
> head(casedata)

After applying the \texttt{formatCases} function, it will be:
\begin{table}[ht]
\begin{center}
\begin{tabular}{cccccccc}
 Year& $CSD2006$ & $CD2006$ & $PR2006$ & AGE\_SEX\_GROUP &Cases &sex &age    \\
 $1999$& $3501005 $ & $3501 $  &  $35 $     &     $210$   &  $3$ &  F  &$50$ \\
 $1999$& $3501005 $ & $3501 $  &  $35 $     &     $213$   &  $1$ &  F  &$65$ \\
 $1999$& $3501005 $ & $3501 $  &  $35 $     &     $214$   &  $6$ &  F  &$70$ \\
 $1999$& $3501005 $ & $3501 $  &  $35 $     &     $215$   &  $1$ &  F  &$75$ \\
 $1999$& $3501005 $ & $3501 $  &  $35 $     &     $216$   &  $3$ &  F  &$80$ \\
 \end{tabular}
\end{center}
\end{table}

So that we can see, there are two new columns: sex and age.

Now let's look at the \texttt{formatPopulation} function:

For example, we have a population data set as:

> data(popdata)
> head(popdata@data)

\begin{table}[ht]
\begin{center}
\begin{tabular}{ccccccc}

  CSDUID   &     CSDNAME      & $M0\_4$& $M5\_9$ &\dots & $F0\_4$& $F5\_9$ \\
 $3501005$ & South Glengarry  & $295$ & $355$  &\dots & $285$ & $320$  \\
 $3501011$ & South Stormont   & $310$ & $380$  &\dots & $265$ & $345$  \\
 $3501012$ &      Cornwall    & $1180$& $1295$ &\dots & $1080$& $1210$ \\
 $3501020$ &  South Dundas    & $230$ & $285$  &\dots & $240$ & $300$  \\
 $3501030$ &  North Dundas    & $295$ & $315$  &\dots & $310$ & $285$  \\
 $3501042$ &  North Stormont  & $200$ & $220$  &\dots & $180$ & $220$  \\

\end{tabular}
\end{center}
\end{table}

After applying the \texttt{formatPopulation} function, it will be:

> formatPopulation(popdata)

\begin{table}[ht]
\begin{center}
\begin{tabular}{cccccccccc}
  CSDUID   & CSDNAME  & $M0\_4$    &\dots &$F0\_4$  & \ldots  &  GROUP &POPULATION& AGE &SEX   \\
 $3501005$ & South Glengarry  & $295$ &\dots & $285$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\
 $3501011$ & South Stormont   & $310$ &\dots & $265$ &\ldots & $M0\_4$& $310$ & $0\_4$ &  M \\ 
 $3501012$ &      Cornwall    & $1180$&\dots & $1080$&\ldots & $M0\_4$& $1180$& $0\_4$ &  M \\ 
 $3501020$ &  South Dundas    & $230$ &\dots & $240$ &\ldots & $M0\_4$& $230$ & $0\_4$ &  M \\ 
 $3501030$ &  North Dundas    & $295$ &\dots & $310$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\ 
 $3501042$ &  North Stormont  & $200$ &\dots & $180$ &\ldots & $M0\_4$& $200$ & $0\_4$ &  M \\ 
\end{tabular}
\end{center}
\end{table}

There are four new columns added in: \texttt{Group}, \texttt{POPULATION}, \texttt{AGE} and \texttt{SEX}.

If we ran the following code:

cancerRates = getRates(casedata, popdata, ~age*sex)

The result of model will be the predicted coefficients of different age and sex groups:

cancerRates

\begin{verbatim}
Call:  glm(formula = formula1, family = family, data = newdata)

Coefficients:
  (Intercept)       age70_74       age60_64       age75_79       age80_84       age55_59       age85_89       age50_54  
     -4.72155        0.18543       -0.38010        0.07425        0.44084       -0.86524        0.30266       -1.86916  
     age45_49       age40_44       age35_39       age30_34       age25_29       age20_24       age15_19         age5_9  
     -1.99423       -2.76184       -3.64470       -5.13532       -4.74062       -6.79155       -7.14367       -6.99895  
     age10_14         age0_4           sexF  age70_74:sexF  age60_64:sexF  age75_79:sexF  age80_84:sexF  age55_59:sexF  
     -7.16576       -8.02353       -0.31172       -0.17597        0.01024       -0.02530       -0.67770             NA  
age85_89:sexF  age50_54:sexF  age45_49:sexF  age40_44:sexF  age35_39:sexF  age30_34:sexF  age25_29:sexF  age20_24:sexF  
     -0.66813        0.42881        0.05375        0.81430        0.72865        1.25281       -0.30888        1.01391  
age15_19:sexF    age5_9:sexF  age10_14:sexF    age0_4:sexF  
           NA             NA             NA             NA

Degrees of Freedom: 30 Total (i.e. Null);  0 Residual
Null Deviance:      38660 
Residual Deviance: 3.535e-13    AIC: 278.9 
\end{verbatim}
The above is the typical result from running the \texttt{glm} model. The $0.18543$, $-0.38010$ etc are the \log of relative risk i.e \beta's.
The base line is %%%%%%%%%

After getting those estimated log relative risk, we can predict the expected number of cancer cases and calculate SMR using function 
\texttt{getSMR}.

ontario = getSMR(cancerRates, popdata, casedata) 
head(ontario@data)

\begin{table}[ht]
\begin{center}
\begin{tabular}{cccccccccc}
  CSDUID   & CSDNAME  & $M0\_4$    &\dots &$F0\_4$  & \ldots  &  GROUP &POPULATION& AGE &SEX   \\
 $3501005$ & South Glengarry  & $295$ &\dots & $285$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\
 $3501011$ & South Stormont   & $310$ &\dots & $265$ &\ldots & $M0\_4$& $310$ & $0\_4$ &  M \\ 
 $3501012$ &      Cornwall    & $1180$&\dots & $1080$&\ldots & $M0\_4$& $1180$& $0\_4$ &  M \\ 
 $3501020$ &  South Dundas    & $230$ &\dots & $240$ &\ldots & $M0\_4$& $230$ & $0\_4$ &  M \\ 
 $3501030$ &  North Dundas    & $295$ &\dots & $310$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\ 
 $3501042$ &  North Stormont  & $200$ &\dots & $180$ &\ldots & $M0\_4$& $200$ & $0\_4$ &  M \\ 
\end{tabular}
\end{center}
\end{table}

\section{Summary and Future Development}

\section{Bibliography}

\end{document}

### 2008-12-15 pkg/diseasemapping/R/New Folder/SpatialFun/SpatialFun.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@18 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2008-12-15 pkg/diseasemapping/R/New Folder/SpatialFun/SpatialFun.Rnw
@@
 $P_k$ denotes the number of people (population size).
-$\lambda_k$ denotes the rate of getting cancer.
 
@@
      \[\log(\lambda_k) = \mu +  x_k\beta + U_k + V_k,\]
-where      $\mu$ is the intercept, it could be $\log(P_k)$;
+where      $\mu$ is the intercept, it could be \log(P_k);
 
-           $x_k$ is the covariates i.e income variable;
+           $x_k$ is the covariates i.e income variables;
 
@@
 
-           $U_k$ is the spatial random variable and \[U_k \sim N(0, \sigma^2_u)\]
+           $U_k$ is the spatial random variable and $U_k$ \sim $N(0, \sigma_u^2)$
 
-           $V_k$ is the non-spatial random variable and \[V_k \sim N(0, \sigma^2_v)\]
+           $V_k$ is the non-spatial random variable and $V_k$ \sim $N(0, \sigma_v^2)$
 
@@
 
-> data(casedata)
-> head(casedata)
-
 library
@@
 
-> formatCases(casedata)
-> head(casedata)
-
 After applying the \texttt{formatCases} function, it will be:
@@
  Year& $CSD2006$ & $CD2006$ & $PR2006$ & AGE\_SEX\_GROUP &Cases &sex &age    \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $210$   &  $3$ &  F  &$50$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $213$   &  $1$ &  F  &$65$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $214$   &  $6$ &  F  &$70$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $215$   &  $1$ &  F  &$75$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $216$   &  $3$ &  F  &$80$ \\
+ $1999$& $3501005 $ & $3501 $  &  $35 $     &     $109$   &  $1$ &  M  &$45$ \\
+ $1999$& $3501005 $ & $3501 $  &  $35 $     &     $110$   &  $1$ &  M  &$50$ \\
+ $1999$& $3501005 $ & $3501 $  &  $35 $     &     $111$   &  $3$ &  M  &$55$ \\
+ $1999$& $3501005 $ & $3501 $  &  $35 $     &     $112$   &  $1$ &  M  &$60$ \\
+ $1999$& $3501005 $ & $3501 $  &  $35 $     &     $113$   &  $2$ &  M  &$65$ \\
  \end{tabular}
@@
 
-> data(popdata)
-> head(popdata@data)
-
 \begin{table}[ht]
 \begin{center}
-\begin{tabular}{ccccccc}
+\begin{tabular}{cccccc}
 
-  CSDUID   &     CSDNAME      & $M0\_4$& $M5\_9$ &\dots & $F0\_4$& $F5\_9$ \\
- $3501005$ & South Glengarry  & $295$ & $355$  &\dots & $285$ & $320$  \\
- $3501011$ & South Stormont   & $310$ & $380$  &\dots & $265$ & $345$  \\
- $3501012$ &      Cornwall    & $1180$& $1295$ &\dots & $1080$& $1210$ \\
- $3501020$ &  South Dundas    & $230$ & $285$  &\dots & $240$ & $300$  \\
- $3501030$ &  North Dundas    & $295$ & $315$  &\dots & $310$ & $285$  \\
- $3501042$ &  North Stormont  & $200$ & $220$  &\dots & $180$ & $220$  \\
+  CSDUID   &     CSDNAME  & CSDTYPE & PRUID  & PRNAME & CDUID \\
+ $3501005$ & South Glengarry  &    TP  & $35$ &Ontario & $3501$ \\
+ $3501011$ & South Stormont   &   TP   & $35$ &Ontario & $3501$  \\
+ $3501012$ &      Cornwall    &  CY    &$35$  &Ontario & $3501$  \\
+ $3501020$ &  South Dundas    &  TP    &$35$  &Ontario & $3501$  \\
+ $3501030$ &  North Dundas    &  TP    &$35$  &Ontario & $3501$  \\
+ $3501042$ &  North Stormont  &    TP  &  $35$& Ontario&  $3501$ \\
 
@@
 After applying the \texttt{formatPopulation} function, it will be:
-
-> formatPopulation(popdata)
-
 \begin{table}[ht]
 \begin{center}
-\begin{tabular}{cccccccccc}
-  CSDUID   & CSDNAME  & $M0\_4$    &\dots &$F0\_4$  & \ldots  &  GROUP &POPULATION& AGE &SEX   \\
- $3501005$ & South Glengarry  & $295$ &\dots & $285$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\
- $3501011$ & South Stormont   & $310$ &\dots & $265$ &\ldots & $M0\_4$& $310$ & $0\_4$ &  M \\ 
- $3501012$ &      Cornwall    & $1180$&\dots & $1080$&\ldots & $M0\_4$& $1180$& $0\_4$ &  M \\ 
- $3501020$ &  South Dundas    & $230$ &\dots & $240$ &\ldots & $M0\_4$& $230$ & $0\_4$ &  M \\ 
- $3501030$ &  North Dundas    & $295$ &\dots & $310$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\ 
- $3501042$ &  North Stormont  & $200$ &\dots & $180$ &\ldots & $M0\_4$& $200$ & $0\_4$ &  M \\ 
+\begin{tabular}{cccccccc}
+   CSDUID &    cases&\ldots & CDUID & \ldots &expected    &logExpected  &     SMR    \\
+ $3501005$&    $28$& \ldots & $3501$& \ldots &$8.433922$  &  $2.132262$ &$3.319926 $ \\
+ $3501007$&    $ 0$& \ldots & $3501$& \ldots &$      NA$  &  $      NA$ &$      NA $ \\
+ $3501011$&    $24$& \ldots & $3501$& \ldots &$7.143624$  &  $1.966220$ &$3.359639 $ \\
+ $3501012$&    $52$& \ldots & $3501$& \ldots &$28.673675$ &  $ 3.355979$&$ 1.813510$ \\
+ $3501020$&    $22$& \ldots & $3501$& \ldots &$6.654331$  &  $1.895268$ &$3.306117 $ \\
+ $3501030$&    $19$& \ldots & $3501$& \ldots &$5.855935$  &  $1.767456$ &$3.244572 $ \\
  \end{tabular}
@@
 
-There are four new columns added in: \texttt{Group}, \texttt{POPULATION}, \texttt{AGE} and \texttt{SEX}.
+There are four new columns added in: cases, this is the number we got from the case data set;
+expected, this is the predicted number of cases we calculated using the glm function;
+logexpected are the \mbox{log} values of the expected numbers; SMR is the rate of \[\frac {oberved(cases)}{expected(expected)}.\]
 
-If we ran the following code:
-
-cancerRates = getRates(casedata, popdata, ~age*sex)
+if we ran the following code:
+\begin{verbatim}
+> model <- getRates( casedata, popdata, ~age*sex, S=c("M","F"))
+\end{verbatim}
 
 The result of model will be the predicted coefficients of different age and sex groups:
+\begin{verbatim}
+> model
 
-cancerRates
-
-\begin{verbatim}
-Call:  glm(formula = formula1, family = family, data = newdata)
+Call:  glm(formula = formula, family = family, data = newdata)
 
 Coefficients:
-  (Intercept)       age70_74       age60_64       age75_79       age80_84       age55_59       age85_89       age50_54  
-     -4.72155        0.18543       -0.38010        0.07425        0.44084       -0.86524        0.30266       -1.86916  
-     age45_49       age40_44       age35_39       age30_34       age25_29       age20_24       age15_19         age5_9  
-     -1.99423       -2.76184       -3.64470       -5.13532       -4.74062       -6.79155       -7.14367       -6.99895  
-     age10_14         age0_4           sexF  age70_74:sexF  age60_64:sexF  age75_79:sexF  age80_84:sexF  age55_59:sexF  
-     -7.16576       -8.02353       -0.31172       -0.17597        0.01024       -0.02530       -0.67770             NA  
-age85_89:sexF  age50_54:sexF  age45_49:sexF  age40_44:sexF  age35_39:sexF  age30_34:sexF  age25_29:sexF  age20_24:sexF  
-     -0.66813        0.42881        0.05375        0.81430        0.72865        1.25281       -0.30888        1.01391  
-age15_19:sexF    age5_9:sexF  age10_14:sexF    age0_4:sexF  
-           NA             NA             NA             NA
+(Intercept)         age5        age10        age15        age20        age25
+  -13.41885     -0.07403      0.52666      0.87985      1.62710      2.35525
+      age30        age35        age40        age45        age50        age55
+    3.49147      4.32618      5.12292      5.49001      5.96051      5.79087
+      age60        age65        age70        age75        age80         sexM
+    7.03365      7.11180      7.22133      7.24461      7.10451      0.67377
+  age5:sexM   age10:sexM   age15:sexM   age20:sexM   age25:sexM   age30:sexM
+         NA           NA           NA     -1.08827     -0.20732     -1.65308
+ age35:sexM   age40:sexM   age45:sexM   age50:sexM   age55:sexM   age60:sexM
+   -0.81485     -1.05938     -0.60803     -0.59874           NA     -0.73841
+ age65:sexM   age70:sexM   age75:sexM   age80:sexM
+   -0.38252     -0.25474     -0.20245           NA
 
-Degrees of Freedom: 30 Total (i.e. Null);  0 Residual
-Null Deviance:      38660 
-Residual Deviance: 3.535e-13    AIC: 278.9 
+Degrees of Freedom: 28 Total (i.e. Null);  0 Residual
+Null Deviance:      10680
+Residual Deviance: -5.529e-14   AIC: 229.2
+
 \end{verbatim}
-The above is the typical result from running the \texttt{glm} model. The $0.18543$, $-0.38010$ etc are the \log of relative risk i.e \beta's.
-The base line is %%%%%%%%%
-
-After getting those estimated log relative risk, we can predict the expected number of cancer cases and calculate SMR using function 
-\texttt{getSMR}.
-
-ontario = getSMR(cancerRates, popdata, casedata) 
-head(ontario@data)
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{cccccccccc}
-  CSDUID   & CSDNAME  & $M0\_4$    &\dots &$F0\_4$  & \ldots  &  GROUP &POPULATION& AGE &SEX   \\
- $3501005$ & South Glengarry  & $295$ &\dots & $285$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\
- $3501011$ & South Stormont   & $310$ &\dots & $265$ &\ldots & $M0\_4$& $310$ & $0\_4$ &  M \\ 
- $3501012$ &      Cornwall    & $1180$&\dots & $1080$&\ldots & $M0\_4$& $1180$& $0\_4$ &  M \\ 
- $3501020$ &  South Dundas    & $230$ &\dots & $240$ &\ldots & $M0\_4$& $230$ & $0\_4$ &  M \\ 
- $3501030$ &  North Dundas    & $295$ &\dots & $310$ &\ldots & $M0\_4$& $295$ & $0\_4$ &  M \\ 
- $3501042$ &  North Stormont  & $200$ &\dots & $180$ &\ldots & $M0\_4$& $200$ & $0\_4$ &  M \\ 
-\end{tabular}
-\end{center}
-\end{table}
 

### 2008-12-15 pkg/diseasemapping/R/New Folder/SpatialFun/code.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@18 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

library(diseasemapping)
data(casedata)
data(popdata)
head(casedata)
head(popdata@data)
cancerRates = getRates(casedata, popdata, ~age*sex)
summary(cancerRates)
ontario = getSMR(cancerRates, popdata, casedata) 
head(ontario@data)
spplot(ontario, "SMR")

library(glmmBUGS)
data(popDataAdjMat)

forBugs = glmmBUGS(formula = Observed + logExpected ~ 1, effects= "CSDUID", 
                   family="poisson", spatial =popDataAdjMat, data=ontario@data)
names(forBugs$ragged)
startingValues=forBugs$startingValues
source("getInits.R")
library(R2WinBUGS)
ontarioResult = bugs(forBugs$ragged, getInits, parameters.to.save=names(getInits()), 
                     model.file="model.bug", n.chain=3, n.iter=1000, n.burnin=100, 
                     n.thin=10, program="winbugs")

ontarioParams=restoreParams(ontarioResult, forBugs$ragged)
ontarioSummary = summaryChain(ontarioParams)

ontario = mergeBugsData(ontario, ontarioSummary)
head(ontario@data)
spplot(ontario, "RCSDUIDSpatial.mean")

Sweave("code.Rnw")

### 2009-01-08 pkg/docs/SpatialFun/code.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@19 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2008-12-15 pkg/diseasemapping/R/New Folder/SpatialFun/code.tex
@@
 data(popDataAdjMat)
+library(spdep)
+popDataAdjMat = poly2nb(ontario, row.names=as.character(ontario$CSDUID))
 
@@
 head(ontario@data)
-spplot(ontario, "RCSDUIDSpatial.mean")
+spplot(ontario, "RCSDUIDSpatial")
 

### 2009-03-26 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@145 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{article}
\usepackage{amsmath}
\begin{document}
\section{The model}
Person $i$ lives at location $s_i$, and is susceptible to cancer if $Y_i=1$.  If they are susceptible, they have a cancer incidence time $T_i$.

\begin{align*}
Y_i & \sim \text{Bernoulli}(\rho)\\
[T_i | Y_i = 1] &\sim \text{Weibull}(\mu_i, \nu)\\
[T_i | Y_i = 0] & = \infty \\
\log(\mu_i) &= X_i \beta + U(s_i)\\
\text{cov}[U(s_0 + s), U(s_0)] &= \sigma^2 M(s; \theta, \kappa) 
\end{align*}

\section{The Data}
For each individual we observe an enrolment date $D_i$, an event date $E_i$, and an indicator variable $Z_i = 1$ if $E_i$ is a cancer incidence and 0 otherwise.

\[
pr(E_i, Z_i | D_i) = 
\begin{cases}
pr(T_i > E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) + pr(Y_i =1) & Z_i = 0 \\
pr(T_i = E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) & Z_i = 1
\end{cases}
\]

\end{document}

### 2009-03-26 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@146 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{amsmath}
+\usepackage[margin=1in]{geometry}
+
 \begin{document}
 \section{The model}
-Person $i$ lives at location $s_i$, and is susceptible to cancer if $Y_i=1$.  If they are susceptible, they have a cancer incidence time $T_i$.
-
+Person $i, i=1\ldots N$ lives at location $s_i$, and is not susceptible to cancer (cured) if $Y_i=1$.  If they are susceptible, they have a cancer incidence time $T_i$.
 \begin{align*}
 Y_i & \sim \text{Bernoulli}(\rho)\\
-[T_i | Y_i = 1] &\sim \text{Weibull}(\mu_i, \nu)\\
-[T_i | Y_i = 0] & = \infty \\
+[T_i | Y_i = 0] &\sim \text{Weibull}(\mu_i, \nu)\\
+[T_i | Y_i = 1] & = \infty \\
 \log(\mu_i) &= X_i \beta + U(s_i)\\
@@
 \end{align*}
+$\mu_i$ and $\nu$ are the scale and shape parameters of the Weibull respectively, and $\rho$ is the cure fraction.
+$U(s)$ is a Gaussian random field with a Matern correlation function with scale parameter $\theta$ and roughness $\kappa$.
+In practice $U(s)$ will be approximated by a Markov random field $\bar U$ on a grid $g_1 \ldots g_M$ with $U(g_k) \approx \bar U_k$.
 
-\section{The Data}
-For each individual we observe an enrolment date $D_i$, an event date $E_i$, and an indicator variable $Z_i = 1$ if $E_i$ is a cancer incidence and 0 otherwise.
+\section{The Likelihood}
+For each individual we observe an enrolment date $L_i$, an event (or death) date $R_i$, and an indicator variable $Z_i = 1$ if $R_i$ is a cancer incidence and 0 if it's a censoring time.
+The density of a Weibull with scale parameter $\mu$ and shape parameter $\nu$ is
+\[
+f(x;\mu,\nu) = \frac{\nu}{\mu} \left(\frac{x}{\mu}\right)^{\nu-1} \exp\left[-\left(\frac{x}{\mu}\right)^\nu\right],
+\]
+and
+\[
+\int_0^x f(u;\mu,\nu)du = 1-\exp\left[-\left(\frac{x}{\mu}\right)^\nu\right].
+\]
+using Wikipedia's parametrization.
 
-\[
-pr(E_i, Z_i | D_i) = 
-\begin{cases}
-pr(T_i > E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) + pr(Y_i =1) & Z_i = 0 \\
-pr(T_i = E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) & Z_i = 1
-\end{cases}
-\]
+Censored observations:
+\begin{align*}
+pr(R_i, Z_i=0 | L_i) =& pr(T_i > R_i | T_i > L_i, Y_i =0) pr(Y_i =0)  + pr(Y_i =1)\\
+=&pr(T_i > R_i) pr(Y_i = 0)/pr(T_i > L_i) + pr(Y_i =1)\\
+=& \rho + (1-\rho) \left. \int_{R_i}^\infty f(x;\mu_i, \nu) dx \right/\int_{L_i}^\infty f(x;\mu_i, \nu) dx  \\
+= &  \rho + (1-\rho) \left. \exp\left[-\left(\frac{R_i}{\mu_i}\right)^\nu\right] \right/\exp\left[-\left(\frac{L_i}{\mu_i}\right)^\nu\right]\\
+= &  \rho + (1-\rho) \exp\left[\left(\frac{L_i}{\mu_i}\right)^\nu -\left(\frac{R_i}{\mu_i}\right)^\nu\right] \\
+\end{align*}
+
+Observed events:
+\begin{align*}
+pr(R_i, Z_i=1 | L_i) =& pr(T_i = R_i | T_i > L_i, Y_i =0)  pr(Y_i = 0) 
+\\
+=& 
+ pr(T_i = L_i) pr(Y_i = 0) /pr(T_i > L_i) \\
+=&
+ (1-\rho) f(L_i ; \mu_i, \nu) \left/ \int_{L_i}^\infty f(x;\mu_i, \nu) dx  \right.\\
+=&
+ (1-\rho) \frac{\nu}{\mu_i} \left(\frac{R_i}{\mu_i}\right)^{\nu-1} \exp\left[\left(\frac{L_i}{\mu_i}\right)^\nu-\left(\frac{R_i}{\mu_i}\right)^\nu\right]
+\end{align*}
 

### 2009-03-27 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@147 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@149 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2009-03-26 docs/survival/survival.tex
@@
 \usepackage{amsmath}
+\usepackage[margin=1in]{geometry}
+
 \begin{document}
 \section{The model}
-Person $i$ lives at location $s_i$, and is susceptible to cancer if $Y_i=1$.  If they are susceptible, they have a cancer incidence time $T_i$.
+Person $i, i=1\ldots N$ lives at location $s_i$, and is not susceptible to cancer (cured) if $Y_i=1$.  If they are susceptible, they have a cancer incidence time $T_i$.
+\begin{align*}
+Y_i & \sim \text{Bernoulli}(\rho)\\
+[T_i | Y_i = 0] &\sim \text{Weibull}(\gamma_i, \nu)\\
+[T_i | Y_i = 1] & = \infty \\
+\log(\gamma_i) &= X_i \beta + U(s_i)\\
+\text{cov}[U(s_0 + s), U(s_0)] &= \sigma^2 M(s; \theta, \kappa) 
+\end{align*}
+$\gamma_i$ and $\nu$ are the range and shape parameters of the Weibull respectively, and $\rho$ is the cure fraction.
+$U(s)$ is a Gaussian random field with a Matern correlation function with scale parameter $\theta$ and roughness $\kappa$.
+In practice $U(s)$ will be approximated by a Markov random field $\bar U$ on a grid $g_1 \ldots g_M$ with $U(g_k) \approx \bar U_k$.
+
+The density of a Weibull with range parameter $\gamma$ and shape parameter $\nu$ is 
+\[
+f(x;\gamma,\nu) =  \gamma \nu x^{\nu-1} \exp\left[-	\gamma x^\nu\right],
+\]
+and
+\[
+\int_0^x f(u;\gamma,\nu)du = 1-\exp(-\gamma y^\nu).
+\]
+If $\theta$ is the scale parameter in the parametrization of the Weibull in Wikipedia, then 
+$\log(\gamma) = -\alpha\log(\theta)$
+
+\section{Left truncated data}
+For each individual we observe an enrolment date $L_i$, an event (or death) date $R_i$, and an indicator variable $Z_i = 1$ if $R_i$ is a cancer incidence and 0 if it's a censoring time.
+Anyone with $T_i < L_i$ is excluded from the study, so we're conditioning on $T_i > L_i$.
+
+Right censored observations:
+\begin{align*}
+pr(R_i, Z_i=0 | L_i) =& pr(T_i > R_i | T_i > L_i, Y_i =0) pr(Y_i =0)  + pr(Y_i =1)\\
+=&pr(T_i > R_i) pr(Y_i = 0)/pr(T_i > L_i) + pr(Y_i =1)\\
+=& \rho + (1-\rho) \left. \int_{R_i}^\infty f(x;\gamma_i, \nu) dx \right/\int_{L_i}^\infty f(x;\gamma_i, \nu) dx  \\
+= &  \rho + (1-\rho) \left. \exp\left[-\gamma_i R_i^\nu\right] \right/\exp\left[-\gamma_i L_i^\nu\right]\\
+= &  \rho + (1-\rho) \exp\left[-\gamma_i\left(L_i^\nu -R_i^\nu\right)\right] \\
+\end{align*}
+
+Observed events:
+\begin{align*}
+pr(R_i, Z_i=1 | L_i) =& pr(T_i = R_i | T_i > L_i, Y_i =0)  pr(Y_i = 0) 
+\\
+=& 
+ pr(T_i = L_i) pr(Y_i = 0) /pr(T_i > L_i) \\
+=&
+ (1-\rho) f(R_i ; \mu_i, \nu) \left/ \int_{L_i}^\infty f(x;\mu_i, \nu) dx  \right.\\
+=&
+ (1-\rho) \gamma_i\nu R_i^{\nu-1} \exp\left[\gamma\left(L_i^\nu-R_i^\nu\right)\right]
+\end{align*}
+
+\section{Interval censored events}
+
+In this case the event occurs between $L_i$ and $R_i$, and since an event is known to have occurred, $Y_i$ must be 0.
 
 \begin{align*}
-Y_i & \sim \text{Bernoulli}(\rho)\\
-[T_i | Y_i = 1] &\sim \text{Weibull}(\mu_i, \nu)\\
-[T_i | Y_i = 0] & = \infty \\
-\log(\mu_i) &= X_i \beta + U(s_i)\\
-\text{cov}[U(s_0 + s), U(s_0)] &= \sigma^2 M(s; \theta, \kappa) 
+pr(L_i < T_i < R_i ) =& (1-\rho) \int_{L_i}^{R_i}f(u;\gamma_i, \nu) du 
+\\
+=& (1-\rho) \left[ \exp(-\gamma L_i^\nu) -  \exp(-\gamma R_i^\nu)\right]
+\\
 \end{align*}
 
-\section{The Data}
-For each individual we observe an enrolment date $D_i$, an event date $E_i$, and an indicator variable $Z_i = 1$ if $E_i$ is a cancer incidence and 0 otherwise.
+\section{Other observations}
+Untruncated right censored data or observed events can be seen as left truncated data with $L_i= 0$.  Left censored data can be regarded as interval censored data with $L_i = 0$.
 
-\[
-pr(E_i, Z_i | D_i) = 
-\begin{cases}
-pr(T_i > E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) + pr(Y_i =1) & Z_i = 0 \\
-pr(T_i = E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) & Z_i = 1
-\end{cases}
-\]
+Truncated and interval censored data are not exclusive.
+An event can have three times, $M_i$ is the left truncation time, $L_i$ is the left censoring time, and $R_i$ is the right censoring time.
+Sometimes $M_i$ or $L_i$ or both are zero, or $R_i = \infty$.
+
+Users could specify all 3 times, with $L_i =R_i$ for observed events.  R's Surv function specifies 2 times, and an option to indicate if it's interval censoring or  truncation.
 

### 2009-03-31 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@152 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@158 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \[
-\int_0^x f(u;\gamma,\nu)du = 1-\exp(-\gamma y^\nu).
+\int_x^\infty f(u;\gamma,\nu)du = \exp(-\gamma x^\nu).
 \]
@@
 = &  \rho + (1-\rho) \left. \exp\left[-\gamma_i R_i^\nu\right] \right/\exp\left[-\gamma_i L_i^\nu\right]\\
-= &  \rho + (1-\rho) \exp\left[-\gamma_i\left(L_i^\nu -R_i^\nu\right)\right] \\
+= &  \rho + (1-\rho) \exp\left[\gamma_i\left(L_i^\nu -R_i^\nu\right)\right] \\
 \end{align*}
@@
 \section{Other observations}
-Untruncated right censored data or observed events can be seen as left truncated data with $L_i= 0$.  Left censored data can be regarded as interval censored data with $L_i = 0$.
 
-Truncated and interval censored data are not exclusive.
-An event can have three times, $M_i$ is the left truncation time, $L_i$ is the left censoring time, and $R_i$ is the right censoring time.
-Sometimes $M_i$ or $L_i$ or both are zero, or $R_i = \infty$.
+Right censored data, is right-censored left truncated with $L_i= 0$
+\[
+pr(T_i> R_i) = \rho + (1-\rho) \exp\left[\gamma_i\left( -R_i^\nu\right)\right]
+\]
 
-Users could specify all 3 times, with $L_i =R_i$ for observed events.  R's Surv function specifies 2 times, and an option to indicate if it's interval censoring or  truncation.
+Left censored data is $pr(0<T_i < x)$ is interval censored data with $L_i=0$ and $R_i=x$.
 

### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@168 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[xcolor=dvipsnames]{beamer}

\usetheme{Singapore}
\usecolortheme[named=RawSienna]{structure}

\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]

\usepackage[nogin]{c:/brownpe/share/Sweave}
\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}

\graphicspath{{Figures/}}
\setkeys{Gin}{width=\textwidth}

\usepackage{amsmath}
\usepackage{c:/brownpe/share/Sweave}

\newcommand{\comment}[1]{}

\title{Generalized Linear Mixed Models}
\author{Patrick Brown}
\institute{Cancer Care Ontario, and \\ School of Public Health, University of Toronto}

\begin{document}
<<junk,fig=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
@

\frame{\maketitle}

\begin{frame}
\frametitle{Outline}
\begin{itemize}
	\item The model and the problem
	\item First generation of approximate methods and PQL
	\item Bayesian inference and Markov Chain Monte Carlo
	\item Integrated Nested Laplace Approximations --- the holy grail?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The model}
$Y_{ij}$ is the disease status of person $i$ in family $j$.
\begin{align*}
Y_{ij} \sim  & \text{Bernoulli}(p_{ij})\\\
logit(p_{ij}) = &x_{ij}\beta + V_{i}\\
V_{i} \sim &\text{N}(0, \sigma^2)
\end{align*}
$V_i$ is a family-level random effect.

\begin{itemize}
	\item What are the parameter estimates $\hat\sigma^2$ and $\hat\beta$?
	\item and their standard errors?
	\item And what are the predictors of family-level random effects $E(V_i | data)$, and their conditional distributions $[V_i | data]$.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Example: }
\begin{columns}
\column{0.5\textwidth}
\begin{itemize}
	\item Repeated measurements on the presence of a bacteria in children
	\item Placebo and two different treatments
\end{itemize}

<<thedata1,echo=true,fig=false>>=
library(MASS)
data(bacteria)
head(bacteria)
@

\column{0.5\textwidth}
\begin{itemize}
	\item $Y_{ij}$ is the bacteria presence in child $i$ at time $j$
	\item $X_{ij}$ is a vector of indicator variables for treatment
	\item $V_i$ is each child's relative risk
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Example: lung cancer in Ontario}
\begin{columns}
\column{0.65\textwidth}
\begin{itemize}
	\item $Y_i$ and $E_i$ are the observed and expected lung cancer cases in region $i$
	\item $X_i$ is income, a covariate
	\item $U_i$ and $V_i$ are spatial and independent random effects
\end{itemize}

\begin{align*}
Y_i \sim & \text{Posson}(\lambda_i E_i)\\
\log(\lambda_i) = & \mu + X_i \beta + U_i + V_i\\
U_i \sim &MRF(\sigma^2_u)\\
V_i \sim & N(0, \sigma^2_v)
\end{align*}

\column{0.35\textwidth}

Expected 
\includegraphics{ontarioE.jpg}
\comment{
library(diseasemapping)
library(sp)
data(popdata)
data(casedata)
model = getRates(casedata, popdata, ~age*sex)
ontario = getSMR(popdata,model,  casedata)
jpeg('Figures/ontarioE.jpg', height=600, width=800)
print(spplot(ontario, 'logExpected', lty=0))
dev.off()
}

Observed
\includegraphics{ontarioO.jpg}

\comment{
	jpeg('Figures/ontarioO.jpg', height=600, width=800)
 print(spplot(ontario, 'observed',lty=0))
dev.off()
}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Likelihood function?}
\begin{itemize}
	\item The $V_i$ are not parameters, and not observed, so they must be integrated out of the likelihood:
\end{itemize}
\[
L(\beta, \sigma) = \prod_i \int pr(Y_{ij}|V_i) pr(V_i) dV_i 
\]
\begin{itemize}
	\item The integral is intractable except in a few special cases
	\item such as Normally-distributed $Y_{ij}$ 
	\item or Gamma $Y_{ij}$ and $V_{i}$ with no covariates
	\item Empirical Bayes methods are based on conjugate priors
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{First generation of approximations}
\begin{block}{Can we use least squares?}
\begin{itemize}
	\item With no random effects, model fitting is done with Iterated Reweighted Least Squares, or Fisher Scoring
	\item For Poisson data, pretend $Y_i \sim N(\lambda_i, \lambda_i)$ and use least squares.  Though you need to know the variance to do this.
	\item Choose a starting value, then iterate between calculating the variances and estimating the parameters.
\end{itemize}
\end{block}

\begin{block}{Penalized Quazi-likelihood}
Start with initial values for $V_i$ and $\sigma$, then use normal-based likelihood and Fisher scores to:
\begin{itemize}
	\item estimate $\hat\beta$
	\item update predictions of  $V_i$, 
	\item Re-estimate $\sigma$
	\item repeat
\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\frametitle{Example: Bacteria}

\end{frame}

\begin{frame}
\frametitle{Markov Chain Monte Carlo}

\end{frame}

\begin{frame}
\frametitle{Example:}

\end{frame}

\begin{frame}
\frametitle{Example: Lung cancer in Ontario}

\end{frame}

\begin{frame}
\frametitle{Integrated Nested Laplace Approximations}

\end{frame}

\begin{frame}
\frametitle{How it works}

\end{frame}

\begin{frame}
\frametitle{Example: }

\end{frame}

\begin{frame}
\frametitle{Example: Lung cancer in Ontario}

\end{frame}

\end{document}

### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@169 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@170 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-<<thedata1,echo=true,fig=false>>=
+<<thedata1,fig=false>>=
 library(MASS)
 data(bacteria)
+bacterianew <- bacteria[,-(2:3)]
+bacterianew$y = as.integer(bacteria$y=="y")
+levels(bacterianew$trt) <- c("placebo", "drug", "drugplus")
+bacteria=bacterianew
+
 head(bacteria)
+#$
 @
@@
 \begin{itemize}
+\item the ID column denotes repeated measurements on each child.
 	\item $Y_{ij}$ is the bacteria presence in child $i$ at time $j$
@@
 \frametitle{First generation of approximations}
-\begin{block}{Can we use least squares?}
+
 \begin{itemize}
@@
 \end{itemize}
-\end{block}
 
@@
 
+\begin{frame}[fragile]
+\frametitle{Example: Bacteria}
+{\scriptsize
+<<bacteriaPQL,fig=false,echo=true>>=
+library(MASS)
+pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)
+summary(pql)
+@
+}
+
+\end{frame}
+
 \begin{frame}
-\frametitle{Example: Bacteria}
+\frametitle{Advantages/Disadvantages}
+
+\begin{block}{Advantages}
+\begin{itemize}
+	\item Quick and easy to use
+	\item It's better than ignoring dependence completely and using a glm.
+\end{itemize}
+\end{block}
+
+\begin{block}{Disadvantages}
+\begin{itemize}
+	\item It doesn't always produce good results
+	\item Especially if your data are very non-normal,  with small numbers in each group.
+	\item Theoretical justification isn't strong
+\end{itemize}
+\end{block}
 
@@
 \frametitle{Markov Chain Monte Carlo}
-
+\begin{itemize}
+	\item Put prior distributions on parameters $[\sigma]$, $[\beta]$.
+	\item Simulate from the posterior distributions
+	\[
+	[\sigma|Y],\ [\beta|Y],\ [V|Y]
+	\]
+	\item 
+\end{itemize}
 \end{frame}
@@
 \frametitle{Example: }
+<<bacteriaINlA>>=
+# do this pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)
+
+@
 
@@
 \frametitle{Example: Lung cancer in Ontario}
+<<cancerInla>>=
+library(diseasemapping)
+library(sp)
+data(popdata)
+data(casedata)
+model = getRates(casedata, popdata, ~age*sex)
+ontario = getSMR(popdata,model,  casedata)
+data(popDataAdjMat)
 
+# fit bym model 
+#forBugs = glmmBUGS(formula=Observed + logExpected ~ 1,
+#  effects="CSDUID",   family="poisson", spatial=popDataAdjMat,
+#  data=ontario)
+
+@
 \end{frame}

### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@171 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\]
-	\item 
+	\item Use $[V|Y] = [Y|V][V]/[Y]$, but $[Y]$ is impossible to evaluate.
+	\item Gibbs sampler: alternate between sampling from $[V|Y,\beta,\sigma]$ and $[\beta,\sigma|V,Y]$.
+	\item Random walk Metropolis: update $V_i^{(p)}$ to  $V_i^{(p+1)}~N(V_i^{(p)}, \nu^2)$, accept the move with probability
+	\[
+	\min[1, pr(Y|V_i^{(p+1)}) /  pr(Y|V_i^{(p)})) ]
+	\]
+	\item This is guaranteed to converge to a sample from $[V_i|Y]$\ldots
+	\item \ldots eventually
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Example:}
+\frametitle{Example:bacteria}
+<<bacmcmc,fig=false,echo=true>>=
+library(glmmBUGS)
+bac = glmmBUGS(y ~ trt, effects='ID', family='binomial',data=bacteria, file='modelx.bug')
+startingValues = bac$startingValues
+source("getInits.R")
+library(BRugs)
+library(R2WinBUGS)
+bacResult = bugs(bac$ragged, getInits, 
+	parameters.to.save = names(getInits()),
+  model.file="model.bug", n.chain=3, n.iter=1000, n.burnin=100, n.thin=10,
+                             program="openbugs") 
+bacParams = restoreParams(bacResult, bac$ragged)
 
+@
 \end{frame}
@@
 <<cancerInla>>=
+if(F) {
 library(diseasemapping)
@@
 
-# fit bym model 
+# fit bym model with inla, like
 #forBugs = glmmBUGS(formula=Observed + logExpected ~ 1,
@@
 #  data=ontario)
-
+}
 @

### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@172 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	parameters.to.save = names(getInits()),
-  model.file="model.bug", n.chain=3, n.iter=1000, n.burnin=100, n.thin=10,
+  model.file="model.bug", n.chain=3, n.iter=100, n.burnin=10, n.thin=2,
                              program="openbugs") 
@@
 \frametitle{Example: }
-<<bacteriaINlA>>=
+<<bacteriaINlA,fig=false>>=
+#library(MASS)
+#data(bacteria)
+#bacterianew <- bacteria[,-(2:3)]
+#bacterianew$y = as.integer(bacteria$y=="y")
+#levels(bacterianew$trt) <- c("placebo", "drug", "drugplus")
+#bacteria=bacterianew
 # do this pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)
@@
 \frametitle{Example: Lung cancer in Ontario}
-<<cancerInla>>=
+<<cancerInla,fig=false>>=
 if(F) {

### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@173 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\begin{frame}
+\begin{frame}[fragile]
 \frametitle{Example:bacteria}
+{\scriptsize
 <<bacmcmc,fig=false,echo=true>>=
 library(glmmBUGS)
-bac = glmmBUGS(y ~ trt, effects='ID', family='binomial',data=bacteria, file='modelx.bug')
+bac = glmmBUGS(y ~ trt, effects='ID', family='bernoulli',data=bacteria, modelFile='modelx.bug')
 startingValues = bac$startingValues
@@
 @
+}
 \end{frame}

### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@174 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2009-04-02 docs/inla/inla.rnw
@@
 	parameters.to.save = names(getInits()),
-  model.file="model.bug", n.chain=3, n.iter=100, n.burnin=10, n.thin=2,
-                             program="openbugs") 
+  n.chain=3, n.iter=100, n.burnin=10, n.thin=2,
+  program="openbugs", model.file='modelBac.bug') 
 bacParams = restoreParams(bacResult, bac$ragged)
+
+@
+}
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Posterior samples}
+<<bacSample,echo=true>>=
+hist(bacParams$SDID)
+@
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Parameter estimates}
+<<bacpar,fig=false,echo=true>>=
+bacSummary = summaryChain(bacParams)
+bacSummary$scalars[,c(1:3,5,8)]
+bacSummary$RID[1:4,c(1:3,5,8)]
 
@@
 
-\begin{frame}
+\begin{frame}[fragile]
+\frametitle{Did the chiain converge?}
+<<bacchain,echo=true>>=
+checkChain(bacParams)
+@
+
+\end{frame}
+
+\begin{frame}[fragile]
 \frametitle{Example: Lung cancer in Ontario}
+\begin{verbatim}
+popDataAdjMat = poly2nb(ontario,row.names=as.character(ontario[["CSDUID"]]))
 
+forBugs = glmmBUGS(formula=observed + logExpected ~ 1,
+  effects="CSDUID",   family="poisson", 
+  spatial=popDataAdjMat,
+  data=ontario@data)
+startingValues = forBugs$startingValues
+source("getInits.R")
+
+ontarioResult = bugs(forBugs$ragged, getInits, 
+  parameters.to.save = names(getInits()),
+    model.file="model.bug", n.chain=3, n.iter=50, 
+    n.burnin=10, n.thin=2,
+      program="winbugs", debug=T)
+
+ontarioParams = restoreParams(ontarioResult, forBugs$ragged)
+\end{verbatim}
+\end{frame}
+
+\comment(
+
+\comment{
+	jpeg('Figures/ontarioCancer.jpg', height=600, width=800)
+}
+
+\begin{frame}[fragile]
+\frametitle{Predicted cancer rates $E(U_i + V_i|Y)$}
+\begin{verbatim}
+ontarioSummary = summaryChain(ontarioParams)
+ontario = mergeBugsData(popdata, ontarioSummary)
+spplot(ontario, "RCSDUID.mean")
+\end{verbatim}
+\includegraphics[width=0.8\textwidth]{ontarioCancer.jpg}
 \end{frame}

### 2009-04-03 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@177 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\def\mm#1{\ensuremath{\boldsymbol{#1}}} % version: amsmath
 \newcommand{\comment}[1]{}
@@
 bacSummary$scalars[,c(1:3,5,8)]
+bacSummary$betas[,c(1:3,5,8)]
 bacSummary$RID[1:4,c(1:3,5,8)]
-
+#$
 @
@@
 
-\comment(
-
 \comment{
 	jpeg('Figures/ontarioCancer.jpg', height=600, width=800)
+
 }
@@
 \frametitle{Integrated Nested Laplace Approximations}
+\begin{itemize}
+\item Notice that
+\[
+\[
+\pi[\theta|Y] = \frac{\pi[U,\theta|Y]}{\pi[U|\theta,Y]}
+\]
+\item Laplace approximation
+\[
+\tilde\pi[\theta|Y] \propto \left. \frac{\pi[U,\theta|Y]}{\tilde\pi_G[U|\theta, Y]} \right| \theta=\theta^*
+\]
+where $\theta^*$ is the mode, and $\pi_G$ is a Gaussian approximation.
+	\item \[
+	\pi[U,\theta|Y] \propto \pi[\theta]\pi[U|\theta] \prod_{i} \pi[Y_i | U_i,\theta]
+	\]
 
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\begin{itemize}
+	\item Do the following integrals numerically
+\end{itemize}
+\frametitle{Marginal posteriors}
+           \begin{align}
+                \widetilde{\pi}(U_{i}\mid\mm{y}) &= \int
+                \widetilde{\pi}(\mm{\theta} \mid \mm{y}) \;
+                \widetilde{\pi}(U_{i} \mid \mm{\theta}, \mm{y})\; d\mm{\theta}\nonumber\\
+                \widetilde{\pi}(\theta_{j} \mid \mm{y}) &= \int
+                \widetilde{\pi}(\mm{\theta} \mid \mm{y}) \;
+                d\mm{\theta}_{-j} \nonumber
+            \end{align}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Remarks}
+    \begin{enumerate}
+    \item Expect $\widetilde{\pi}(\mm{\theta} | \mm{y})$ to be
+        accurate, since
+        \begin{itemize}
+        \item $\mm{x}|\mm{\theta}$ is \emph{a priori} Gaussian
+        \item Likelihood models are `well-behaved' so
+            \begin{displaymath}
+                \pi(\mm{x} \mid \mm{\theta}, \mm{y})
+            \end{displaymath}
+            is \emph{almost} Gaussian.
+        \end{itemize}
+    \item There are no distributional assumptions on $\mm{\theta} |
+        \mm{y}$
+    \item Similar remarks are valid to
+        \begin{displaymath}
+            \widetilde{\pi}(x_{i}\mid\mm{\theta}, \mm{y})
+        \end{displaymath}
+    \end{enumerate}
 \end{frame}
@@
 \frametitle{How it works}
+\begin{enumerate}
+	\item Explore the space to find the model of $[\theta|Y]$
+	\item Then compute the Laplace approximation for $[U|Y,\theta]$
+	\item Then numerically integrate out the $\theta_i$.
+	\item Use polynomials in place of the Gaussian to approximate the $[U_i|\theta,Y]$.
+\end{enumerate}
+\end{frame}
 
+\begin{frame}[fragile]
+\frametitle{Example: bacteria}
+{\small
+\begin{verbatim}
+bacteria$r<-bacteria$y=="y"
+bacteria$n<-1
+formula1<- r ~ trt + f(ID,model="iid",param=c(0.001,0.001))
+mod1 = inla(formula1,family="binomial",data=bacteria,Ntrials=n)
+summary(mod1) #This will give summaries of the posteriors
+
+Fixed effects:
+                 mean        sd 0.025quant 0.975quant    kld dist
+(Intercept)  2.2955447 0.4747077   1.498364  3.3689509 0.155881406
+trtdrug     -1.2147419 0.5831969  -2.442864 -0.1223520 0.027425825
+trtdrug+    -0.7292562 0.5917214  -1.953683  0.3977671 0.008737907
+
+Model hyperparameters:
+                mean   sd     0.025quant 0.975quant
+Precision for ID 0.9724 0.5651 0.2943     2.4230  
+Expected number of effective parameters(std dev) :18.03(7.335)
+Number of equivalent replicates : 12.20
+\end{verbatim}
+}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Example: }
-<<bacteriaINlA,fig=false>>=
-#library(MASS)
-#data(bacteria)
-#bacterianew <- bacteria[,-(2:3)]
-#bacterianew$y = as.integer(bacteria$y=="y")
-#levels(bacterianew$trt) <- c("placebo", "drug", "drugplus")
-#bacteria=bacterianew
-# do this pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)
+\frametitle{Posterior distribution of precision}
+\includegraphics[width=0.7\textwidth]{Prec_BACTERIA.pdf}
+\end{frame}
 
-@
+\begin{frame}[fragile]
+\frametitle{Example: Lung cancer in Ontario}
+{\small
+\begin{verbatim}
+nb2inla("ontario.graph",popDataAdjMat)
+data$observed[is.na(data$logExpected)]<-NA
+data$logExpected[is.na(data$logExpected)]<- -100 
+data$region<-0:(length(data$CSDUID)-1)
+data$region.spatial<-data$region
 
+formula2<-observed ~ f(region.spatial,
+    model="besag", graph.file="ontario.graph", 
+    param=c(20,15),initial=2) + 
+  f(region,model="iid", param=c(1,0.01))
+
+mod2 <- inla(formula2,family="poisson",data=data,
+  offset=logExpected,verbose=TRUE)
+
+data$V<-mod2$summary.random$region[,2]
+data$U<-mod2$summary.random$region.spatial[,2]
+
+writePolyShape(ontario, "ontario")
+
+\end{verbatim}
+}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Example: Lung cancer in Ontario}
-<<cancerInla,fig=false>>=
-if(F) {
-library(diseasemapping)
-library(sp)
-data(popdata)
-data(casedata)
-model = getRates(casedata, popdata, ~age*sex)
-ontario = getSMR(popdata,model,  casedata)
-data(popDataAdjMat)
+\frametitle{Smoothed cancer risk}
+\includegraphics{Map_LINEAR.pdf}
+\end{frame}
 
-# fit bym model with inla, like
-#forBugs = glmmBUGS(formula=Observed + logExpected ~ 1,
-#  effects="CSDUID",   family="poisson", spatial=popDataAdjMat,
-#  data=ontario)
+\begin{frame}[fragile]
+\frametitle{Results}
+{\small
+
+\begin{verbatim}
+summary(mod2)
+
+Fixed effects:
+                mean       sd 0.025quant 0.975quant     kld dist
+(Intercept) 0.7209629 3.654778  -6.456466   7.888327 7.134535e-05
+
+Random effects:
+Name      Model         Max KLD
+region.spatial   Besags ICAR model   0.00083
+region   IID model   0.01849
+
+Model hyperparameters:
+                            mean   sd     0.025quant 0.975quant
+Precision for region.spatial 2.2255 0.3251 1.6469     2.9213  
+Precision for region         1.9612 0.1960 1.6005     2.3700  
+Expected number of effective parameters(std dev) :357.87(2.378)
+Number of equivalent replicates : 1.115
+\end{verbatim}
+
 }
-@
+\end{frame}
+
+\begin{frame}
+\frametitle{Advantages/Disadvantanges}
+\begin{block}{Advantages}
+\begin{itemize}
+	\item It appears that this is a very good approximation, and is much less computationally intensive than MCMC
+	\item the code works in parallel, so is very fast on multi-core machines
+\end{itemize}
+\end{block}
+\begin{block}{Disadvantages}
+\begin{itemize}
+\item Only marginals, not joint posteriors, are produced.
+\item INLA is harder to modify than MCMC routines, though that might be only due to familiarity with MCMC
+\item There are some things which INLA can't do, though in theory MCMC can do anything.
+\item Starting values do have an effect, so INLA isn't fool-proof.
+\end{itemize}
+\end{block}
+\end{frame}
+
+\begin{frame}
+\frametitle{Conclusions:}
+\begin{itemize}
+	\item There's no excuse for avoiding GLMM's
+	\item or using PQL 
+	\item Good models and good tools are available for dependent non-Gaussian data.
+\end{itemize}
 \end{frame}

### 2009-04-06 docs/dionne/slides.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@178 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[xcolor=dvipsnames]{beamer}

\usetheme{Singapore}
\usecolortheme[named=RawSienna]{structure}

\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number] 
\usepackage[nogin]{c:/brownpe/share/Sweave}
\title{Modelling small-area disease incidence data in space and time }
\author{Patrick Brown}
\institute{Cancer Care Ontario, and \\ School of Public Health, University of Toronto}

\usepackage{amsmath}
\graphicspath{{Figures/}}
\setkeys{Gin}{width=\textwidth}
\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}

\begin{document}

\frame{\maketitle}

\begin{frame}
\frametitle{Large v Small Area Maps}

\begin{columns}
\column{0.5\textwidth}

\begin{block}{Coarse scale}

\vspace{10pt}

\includegraphics[width=0.9\textwidth]{lung.png}

\vspace{10pt}

Observed rate for Ontario, by Census sub-division

\end{block}

\column{0.5\textwidth}

\begin{block}{Fine scale}

\includegraphics[width=\textwidth]{torCTsmr.pdf}

SMR's for Toronto, by census tract

\end{block}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Issues with small area maps}

\begin{block}{Small counts}
	mostly 0 or 1
\begin{itemize}
\item The usual estimates of incidence rates perform poorly
	\item Methods based on the normal-distribution aren't appropriate
\end{itemize}

\end{block}
\begin{block}{Spatial dependence}
With regions close together being more similar than regions further apart
\begin{itemize}
	\item The usual regression models underestimate standard errors
  \item but this helps us deal with small counts
\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\frametitle{This talk}
Following this talk, those here will be able to:
\begin{itemize}
\item Understand a statistical model suitable for small-area disease incidence data with low counts;
\item Use the results of this model to answer questions in social and environmental epidemiology;
\item  Be aware of how to extend these models to problems in space and time.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Gonorrhoea in North Carolina}
describe data
map raw rates
map covariates, \% black, unemp
\end{frame}

\begin{frame}
\frametitle{The model (without equations)}

\begin{itemize}
	\item The true risk surface $\lambda(s)$ at location $s$ is unknown.  
	\item It depends on some measured risk factors, such as income and unemployment rate $X(s)$.
	\item And some residual spatial variation $W(s)$ which might be smoothly varying in space.

\begin{itemize}
	\item This might be caused by unmeasured risk factor such as air pollution
	\item Or be the result of direct transmission creating areas of high risk
\end{itemize}
\item The number of cases $Y_i$ in region $i$ is random, but we'd expect it to be related to the population in the region and the risk $\lambda(s)$ there.
\item In regions of small population we shouldn't be surprised to see zero cases, even if risk isn't zero.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Dealing with small counts}

\begin{columns}
\column{0.5\textwidth}
\begin{block}{The Model}
\[
Y_i \sim \text{Poisson}(P_i\lambda_i)
\]
\begin{itemize}
	\item $Y_i$ is the number of cases in region $i$
	\item $P_i$ is the population
	\item $\lambda_i$ is the risk
\end{itemize}
\end{block}
\column{0.5\textwidth}
\begin{block}{The Poisson Distriution}
<<poisson,fig=true,echo=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
xseq = 0:14
toplot = dpois(xseq, 0.5)
names(toplot) = as.character(xseq)
toplot2 = dpois(xseq, 1.5)
barplot(toplot, width=0.4, xlim = range(xseq), space=4, xlab='count', ylab='pr(count)',
	col='red',border=NA)
barplot(toplot, width=0.21, add=T, space=8.8, col='white', xlim = range(xseq),border=NA, axes=F, axisnames=F)

barplot(toplot2, width=0.2, add=T, space=9, col='blue', xlim = range(xseq),border=NA, axes=F)
@

mean = \textcolor{red}{0.5},  \textcolor{blue}{1.5} 
\end{block}

\end{columns}

\begin{block}{Interpretation}
\begin{itemize}
	\item We don't observe the true rate $\lambda_i$
	\item but rather a noisy version of it in the form of case counts
	\item The Poisson distribution can cope with small counts and lots of zeros
\end{itemize}
\end{block}

\end{frame}

\begin{frame}
\frametitle{Modelling the rates}
\begin{align*}
Y_i \sim & \text{Poisson}(P_i\lambda_i)\\
\log(\lambda_i) = & \mu + X_i \beta + U_i + V_i\\
V_i \sim &\text{N}(0, \tau^2)\\
U_i|U_{-i} \sim & \text{N}(\text{mean}\{U_j; j \sim i \}, \sigma^2/||\{j ; j \sim i\}||
\end{align*}
\begin{itemize}
  \item Random effects model: Risks are usually close to the average --- shrinkage
	\item Risk can depends on covariates $X_i$ (i.e.\ income)
	\item $U_i$ is a spatially smooth random effect, $V_i$ is a rough non-spatial effect
\item The spatial nature of $U_i$ results in regions neighbouring $i$ to influence risk estimation at $i$, somewhat mitigating the low counts problem.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Model fitting}
\begin{itemize}
	\item The data are not Normally-distributed, and there is an unobserved random effect.
	\item This makes the usual least squares or maximum likelihood methods problematic
	\item Bayesian inference using Markov Chain Monte Carlo simulation
	\item There are several implementations in the statistical package R.
	\item and the ArcGIS add-on the Rapid Inquiry Facility
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Example: North Carolina}

\begin{columns}
\column{0.5\textwidth}
\begin{block}{Where}
Predicted risk E$(\lambda_i | data)$

\end{block}
\column{0.5\textwidth}
\begin{block}{Why}
Predicted risk minus known risk factors \\
$E(\lambda_i - X_i \beta | data)= E(U_i + V_i)$

\end{block}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Parameter estimates}
<<parest,result=tex,figure=false>>=
load("data\\GONQResult.RData")
library(glmmBUGS)
  gonpar = restoreParams(GONQResult, GONQragged$ragged)
  save(gonpar, file="gonpar.RData")
@
load("gonpar.RData")

\end{frame}

\begin{frame}
\frametitle{Recap of spatial models:}
\begin{itemize}
	\item True risk is unobserved, use observed counts to predict risk
	\item Risk is random and may vary smoothly in space
\item Allow for known risk factors, as covariates
\item Map residual spatial variation or predicted risk
\item and look at parameter estimates
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Spatio-temporal models}
\begin{itemize}
	\item Risk changes in time as well as space
	\item $Y_{ij}$ is the incidence count in region $i$ at time $j$
	\begin{align*}
Y_{ij} \sim & \text{Poisson}(P_{ij}\lambda_{ij})\\
\log(\lambda_{ij}) = & \mu + X_{ij} \beta + U_i + V_i + W_{ij}\\
V_i \sim &\text{N}(0, \tau^2)\\
U_i|U_{-i} \sim & \text{N}(\text{mean}\{U_j; j \sim i \}, \sigma^2/||\{j ; j \sim i\}||\\
W_{ij} \sim &\text{N}(0, \nu^2)
	\end{align*}
\item $W_{ij}$ is the variation in time from the long-term average for each region.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Parameter Estimates}

\end{frame}

\begin{frame}
\frametitle{E($W_{ij}$), time 1}

\end{frame}

\begin{frame}
\frametitle{E($W_{ij}$), time 2}

\end{frame}

\begin{frame}
\frametitle{Conclusion}

\begin{itemize}
	\item Small area disease mapping in space and time is possible even when counts are low 
	\item Include covariates in the model to estimate the effect of potential risk factors
	\item There is software for implementing these models which is not terribly difficult to use.
\end{itemize}
\end{frame}

\end{document}

### 2009-04-06 docs/dionne/slides.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@179 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{frame}
-\frametitle{Gonorrhoea in North Carolina}
+\frametitle{ North Carolina}
 describe data
@@
 
-\begin{columns}
-\column{0.5\textwidth}
-\begin{block}{Where}
-Predicted risk E$(\lambda_i | data)$
+\begin{block}{Where:
+Predicted risk E$(\lambda_i | data)$}
+
+\includegraphics[width=0.8\textwidth]{rate.png}
+\end{block}
+
+\begin{block}{Why:
+Predicted risk minus known risk factors}
+$E(\lambda_i - X_i \beta | data)= E(U_i + V_i)$
+
+\includegraphics[width=0.8\textwidth]{relativeRate.png}
 
 \end{block}
-\column{0.5\textwidth}
-\begin{block}{Why}
-Predicted risk minus known risk factors \\
-$E(\lambda_i - X_i \beta | data)= E(U_i + V_i)$
-
-\end{block}
-
-\end{columns}
 
@@
 \begin{frame}
-\frametitle{Parameter estimates}
-<<parest,result=tex,figure=false>>=
+\vspace*{-15pt}
+Parameter estimates\\
+\vspace*{-5pt}
+{\small
+<<parest,results=tex,fig=false>>=
+if(F) {
 load("data\\GONQResult.RData")
@@
   save(gonpar, file="gonpar.RData")
+}
+load("gonpar.RData")
+library(glmmBUGS)
+gonSummary = summaryChain(gonpar)
+par = rbind(gonSummary[[1]], gonSummary[[2]])
+
+toprint =par[c(1,3,4,16:32),c(1,2,5,9)]
+rownames(toprint) = gsub("_", "-" , rownames(toprint))
+colnames(toprint) = gsub("%", "pct",colnames(toprint))
+latex(toprint, file='', dec=2, rowlabel='Parameter')
 @
-load("gonpar.RData")
-
+}
 \end{frame}
@@
 \frametitle{Parameter Estimates}
+<<parestst,results=tex,fig=false>>=
+
+toprint =par[2:12,c(1,2,5,9)]
+rownames(toprint) = gsub("_", "-" , rownames(toprint))
+colnames(toprint) = gsub("%", "pct",colnames(toprint))
+latex(toprint, file='', dec=2, rowlabel='Parameter')
+@
 
@@
 \frametitle{E($W_{ij}$), time 1}
-
+\includegraphics[width=1.2\textwidth]{R05Q2.png}
 \end{frame}
@@
 \frametitle{E($W_{ij}$), time 2}
+\includegraphics[width=1.2\textwidth]{R05Q3.png}
+\end{frame}
 
+\begin{frame}
+\frametitle{E($W_{ij}$), time 3}
+\includegraphics[width=1.2\textwidth]{R05Q4.png}
+\end{frame}
+
+\begin{frame}
+\frametitle{E($W_{ij}$), time 4}
+\includegraphics[width=1.2\textwidth]{R06Q1.png}
+\end{frame}
+
+\begin{frame}
+\frametitle{E($W_{ij}$), time 5}
+\includegraphics[width=1.2\textwidth]{R06Q2.png}
+\end{frame}
+\begin{frame}
+\frametitle{E($W_{ij}$), time 6}
+\includegraphics[width=1.2\textwidth]{R06Q3.png}
+\end{frame}
+\begin{frame}
+\frametitle{E($W_{ij}$), time 7}
+\includegraphics[width=1.2\textwidth]{R06Q4.png}
+\end{frame}
+\begin{frame}
+\frametitle{E($W_{ij}$), time 8}
+\includegraphics[width=1.2\textwidth]{R07Q1.png}
+\end{frame}
+\begin{frame}
+\frametitle{E($W_{ij}$), time 9}
+\includegraphics[width=1.2\textwidth]{R07Q2.png}
+\end{frame}
+\begin{frame}
+\frametitle{E($W_{ij}$), time 10}
+\includegraphics[width=1.2\textwidth]{R07Q3.png}
+\end{frame}
+\begin{frame}
+\frametitle{E($W_{ij}$), time 11}
+\includegraphics[width=1.2\textwidth]{R07Q4.png}
+\end{frame}
+\begin{frame}
+\frametitle{E($W_{ij}$), time 12}
+\includegraphics[width=1.2\textwidth]{R08Q1.png}
 \end{frame}

### 2009-04-07 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@188 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\section{Left truncated data}
-For each individual we observe an enrolment date $L_i$, an event (or death) date $R_i$, and an indicator variable $Z_i = 1$ if $R_i$ is a cancer incidence and 0 if it's a censoring time.
-Anyone with $T_i < L_i$ is excluded from the study, so we're conditioning on $T_i > L_i$.
+\subsection{Notation}
 
-Right censored observations:
+Each individual data for one or more of the following variables:
+\begin{itemize}
+\item Time of event $T_i$, often unobserved
+	\item left trunCation time $C_i$, with $C_i=0$ for untruncated observations
+\item Lower bound for the event time $L_i$
+\item Upper bound $U_i$
+\item Event type $E_i$ defined as
+\begin{itemize}
+	\item $E_i=1$: event observed at time $T_i$
+	\item $E_i=0$: event right censored at $L_i$, with $T_i$ unobserved but $T_i >L_i$.
+	\item $E_i=2$: left censored data at $U_i$, with $T_i < U_i$.
+	\item $E_i=3$: interval censored data with $L_i < T_i < U_i$. 
+\end{itemize}
+\end{itemize}
+The 'cure' indicator $Y_i$ is not part of the dataset.  Note that $E_i$ isnt strictly necessary since it can be worked out by which of $T_i$, $L_i$ and $U_i$ are specified.
+
+\section{Likelihood}
+
+\subsection{Right censored observations $E_i =0$}
+
 \begin{align*}
-pr(R_i, Z_i=0 | L_i) =& pr(T_i > R_i | T_i > L_i, Y_i =0) pr(Y_i =0)  + pr(Y_i =1)\\
-=&pr(T_i > R_i) pr(Y_i = 0)/pr(T_i > L_i) + pr(Y_i =1)\\
-=& \rho + (1-\rho) \left. \int_{R_i}^\infty f(x;\gamma_i, \nu) dx \right/\int_{L_i}^\infty f(x;\gamma_i, \nu) dx  \\
-= &  \rho + (1-\rho) \left. \exp\left[-\gamma_i R_i^\nu\right] \right/\exp\left[-\gamma_i L_i^\nu\right]\\
-= &  \rho + (1-\rho) \exp\left[\gamma_i\left(L_i^\nu -R_i^\nu\right)\right] \\
+pr(L_i, E_i=0 | C_i) =& pr(T_i > L_i | T_i > C_i, Y_i =0) pr(Y_i =0)  + pr(Y_i =1)\\
+=&pr(T_i > L_i) pr(Y_i = 0)/pr(T_i > C_i) + pr(Y_i =1)\\
+=& \rho + (1-\rho) \left. \int_{L_i}^\infty f(x;\gamma_i, \nu) dx \right/\int_{C_i}^\infty f(x;\gamma_i, \nu) dx  \\
+= &  \rho + (1-\rho) \left. \exp\left[-\gamma_i L_i^\nu\right] \right/\exp\left[-\gamma_i C_i^\nu\right]\\
+= &  \rho + (1-\rho) \exp\left[\gamma_i\left(C_i^\nu -L_i^\nu\right)\right] \\
+\end{align*}
+Note that $C_i=0$ implies the $C_i^\nu$ term can be omitted and the untruncated likelihood results.
+
+\subsection{Observed events $E_i =1$}
+\begin{align*}
+pr(T_i, E_i=1 | C_i) =& pr(T_i  | T_i > C_i, Y_i =0)  pr(Y_i = 0) 
+\\
+=& 
+ pr(T_i ) pr(Y_i = 0) /pr(T_i > C_i) \\
+=&
+ (1-\rho) f(T_i ; \mu_i, \nu) \left/ \int_{C_i}^\infty f(x;\mu_i, \nu) dx  \right.\\
+=&
+ (1-\rho) \gamma_i\nu T_i^{\nu-1} \exp\left[\gamma\left(C_i^\nu-T_i^\nu\right)\right]
 \end{align*}
 
-Observed events:
+\subsection{Left censored observations $E_i =2$}
+I'm a bit confused about this one...  For completeness we should define the likelihood for left censoring with left truncation.
 \begin{align*}
-pr(R_i, Z_i=1 | L_i) =& pr(T_i = R_i | T_i > L_i, Y_i =0)  pr(Y_i = 0) 
-\\
-=& 
- pr(T_i = L_i) pr(Y_i = 0) /pr(T_i > L_i) \\
-=&
- (1-\rho) f(R_i ; \mu_i, \nu) \left/ \int_{L_i}^\infty f(x;\mu_i, \nu) dx  \right.\\
-=&
- (1-\rho) \gamma_i\nu R_i^{\nu-1} \exp\left[\gamma\left(L_i^\nu-R_i^\nu\right)\right]
+pr(U_i, E_i=2 | C_i) =& pr(C_i < T_i < U_i | T_i > C_i, Y_i =0) pr(Y_i =0) \\
+=&pr(C_i < T_i <U_i) pr(Y_i = 0)/pr(T_i > C_i) \\
+=& (1-\rho) \left. \int_{C_i}^{U_i} f(x;\gamma_i, \nu) dx \right/\int_{C_i}^\infty f(x;\gamma_i, \nu) dx  \\
+= &    (1-\rho) \left. \left[\exp(-\gamma_i C_i^\nu) - \exp(-\gamma_i U_i^\nu) \right] \right/\exp\left[-\gamma_i C_i^\nu\right]\\
+= &  (1-\rho)\left\{1 -  \exp\left[\gamma_i\left(C_i^\nu -U_i^\nu\right)\right] \right\} \\
 \end{align*}
 
-\section{Interval censored events}
-
-In this case the event occurs between $L_i$ and $R_i$, and since an event is known to have occurred, $Y_i$ must be 0.
+\subsection{Interval censored events, $E_i =2$}
 
 \begin{align*}
-pr(L_i < T_i < R_i ) =& (1-\rho) \int_{L_i}^{R_i}f(u;\gamma_i, \nu) du 
+pr(L_i < T_i < U_i | T_i > C_i ) =& (1-\rho) \int_{L_i}^{U_i}f(u;\gamma_i, \nu) du 
 \\
-=& (1-\rho) \left[ \exp(-\gamma L_i^\nu) -  \exp(-\gamma R_i^\nu)\right]
+=& (1-\rho) \left[ \exp(-\gamma L_i^\nu) -  \exp(-\gamma U_i^\nu)\right] / \exp\left[-\gamma_i C_i^\nu\right]
 \\
 \end{align*}
-
-\section{Other observations}
-
-Right censored data, is right-censored left truncated with $L_i= 0$
-\[
-pr(T_i> R_i) = \rho + (1-\rho) \exp\left[\gamma_i\left( -R_i^\nu\right)\right]
-\]
-
-Left censored data is $pr(0<T_i < x)$ is interval censored data with $L_i=0$ and $R_i=x$.
+The denominator is 1 for untrancated $C_i =0$.
 

### 2009-04-14 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@214 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{Left censored observations $E_i =2$}
-I'm a bit confused about this one...  For completeness we should define the likelihood for left censoring with left truncation.
+For completeness wedefine the likelihood for left censoring with left truncation.
 \begin{align*}
-pr(U_i, E_i=2 | C_i) =& pr(C_i < T_i < U_i | T_i > C_i, Y_i =0) pr(Y_i =0) \\
+pr(U_i, E_i=2 | C_i) =& pr(T_i < U_i | T_i > C_i, Y_i =0) pr(Y_i =0) \\
 =&pr(C_i < T_i <U_i) pr(Y_i = 0)/pr(T_i > C_i) \\

### 2009-04-29 docs/hamilton/hamilton.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@253 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{subfigure,float}

\title{Lung Cancer, Smoking, and Industrial Pollution in Hamilton, Ontario}

\author{Patrick, Billy,  Susitha, Eric, ...}

\usepackage{natbib}

\begin{document}

\maketitle

\section*{Abstract}

do this last

\section{Introduction}

The link between environmental  pollution and cancer is a question of great public interest.   In the city of Hamilton, Ontario, lung cancer is clearly elevated near steel and iron foundries.  However, it might be expected that the individuals living further from the foundries have higher incomes and therefore lower smoking rates, and the correlation between cancer and proximity to foundries could be due to or enhanced by spatial variations in tobacco usage.  Cite some other papers which state smoking is a possible cause of pollution/cancer correlations?

This paper uses population-level data from the census  of Canada, the Ontario Cancer Registry (OCR), and the Canadian
Community Household Survey (CCHS) to assess the effect of pollution on lung cancer after adjusting for spatial variations in smoking rates.  As the estimated pollution levels in Hamilton vary greatly depending on distance from the foundries, cancer risk and smoking rates are modelled at the finest possible spatial resolution, the census dissemination areas (DA).  At this resolution the number of cancer cases and CCHS respondents in each area is small, making spatial models necessary for predicting the spatial risk surfaces for smoking and cancer.  The smoking surface is included as a confounder in the cancer model, with the cancer and smoking models integrated to allow for uncertainty in the smoking surface to be reflected in the cancer model in a manner similar to that used by \cite{BILY ADD REFERENCE}.  An additional explanatory variable in the model is environmental air pollution, and the effect of pollution on cancer risk is a parameter estimated.

map of pollution surface, with foundries, and DA boundries.

\section{Methods}
\subsection{Data}
describe cancer registry, pollution surface, cchs.

\subsection{Model}

Billy, can you start filling this bit in?

describe model

Bayesian inference

chain convergence assesed by examining trace plots, and we did thinning by ??, burnin?? chain length ??

\section{Results}

\begin{table}[H]

\subfigure[Female]{\hspace{.6in}a table\hspace{.6in}}
\subfigure[Male]{\hspace{.6in}a table\hspace{.6in}}

\caption{Model parameter estimates, with 95\% posterior credible intervals for lung cancer in Hamilton, Ontario}
\label{tab:parameters}
\end{table}

\begin{figure}[H]
\subfigure[Cancer risk]{\hspace{.6in} a plot\hspace{.6in}}
\subfigure[Smoking risk]{\hspace{.6in} a plot\hspace{.6in}}
\subfigure[Probability of residual 20\% excess risk]{\hspace{.6in}a plot\hspace{.6in}}
\caption{Predicted risk surfaces and exceedance probability plot for female lung cancer in Hamilton, Ontario.  Residual spatial variation is the predicted log cancer risk for each DA minus the predicted effects of smoking and pollution.}
\label{fig:female}
\end{figure}

\begin{figure}[H]
\subfigure[Cancer risk]{\hspace{.6in} a plot\hspace{.6in}}
\subfigure[Smoking risk]{\hspace{.6in} a plot\hspace{.6in}}
\subfigure[Probability of residual 20\% excess risk]{\hspace{.6in}a plot\hspace{.6in}}
\caption{Predicted risk surfaces and exceedance probability plot for male lung cancer in Hamilton, Ontario.  Residual spatial variation is the predicted log cancer risk for each DA minus the predicted effects of smoking and pollution.}
\label{fig:female}
\end{figure}

\section{Discussion}

\section*{Acknowledgements}

\bibliography{}

\end{document}

### 2009-05-11 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@289 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2008-12-15 pkg/diseasemapping/R/New Folder/SpatialFun/SpatialFun.tex
@@
 \documentclass{article}
-\usepackage{graphicx}
-\textwidth 6.0 true in
-\textheight 9 true in
-\topmargin -1.0 true in
-\headsep 0.375 true in
-\oddsidemargin 0.0 true in
-\evensidemargin 0.5 true in
+\usepackage{Sweave}
+\title{Priors for swine flu}
 
-\title{The Spatial Model Functions}
-\def\ffmc{\mbox{ffmc}}
-\def\wind{\mbox{wind}}
-\def\ROS{\mbox{ROS}}
-\def\ISI{\mbox{ISI}}
-\def\BUI{\mbox{BUI}}
 \begin{document}
+
 \maketitle
 
-\large
-\section{Introduction}
-
-\section{Method}
-\subsection{The Poisson Model}
-The followings are the theoretical functions:
-
-Let $Y_k$ denotes the number of cancer cases in group k.
-$P_k$ denotes the number of people (population size).
-
-We have
-     \[Y_k \sim Poisson (P_k\lambda_k),\]  we won't put time under our consideration at the beginning.  %or the function should be Y_k ~ Poisson(\lambda_i T)
-     \[\log(\lambda_k) = \mu +  x_k\beta + U_k + V_k,\]
-where      $\mu$ is the intercept, it could be \log(P_k);
-
-           $x_k$ is the covariates i.e income variables;
-
-           $\beta$ is the log relative risk;
-
-           $U_k$ is the spatial random variable and $U_k$ \sim $N(0, \sigma_u^2)$
-
-           $V_k$ is the non-spatial random variable and $V_k$ \sim $N(0, \sigma_v^2)$
-
-Therefore, we will fit a $glm$ (generalized linear model) as:
-    \[model = glm( cases \sim offset(logpop) + factor(age) factor(sex) )\]
-
-\subsection{The R Functions}
-Currently the diseasemapping package contains $6$ functions:
-
-\begin{enumerate}
-\item
-
-\begin{verbatim}
-''formatCases'' <- function(casedata)
-\end{verbatim}
-
-We get the data sets from different kinds of sources. Therefore, we need to clean and format the data set
-for the future manipulation. The function dealing with two scenarios: when there is  \texttt{group} column in the data set
-or when there is  \small \texttt{AGE\_SEX\_GROUP} column in data set. Essentially, we want to add in two
-columns : sex and age.
-\large
-
-Group is usually a column looks like:
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{|c|c|}
-\hline
- $M0\_5$ \\
- $M5\_9$ \\
- $M10\_14$ \\
-  \vdots  \\
- $M85PLUS$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-AGE\_SEX\_Group is usually a column of numbers composed with $3$ digits:
-Male are the numbers begin with $1$, the rest two digits are corresponding to the age.
-i.e:
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{|c|c|}
-\hline
-$100$  &  $M0\_4$ \\
-$101$  &  $M5\_9$ \\
-$102$  &  $M10\_14$ \\
-\vdots  &  \vdots  \\
-$117$  &  $M85PLUS$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-Female is the same, but starting at $200$.
-
-\item
-\begin{verbatim}
-''formatPopulation'' <- function(popdata)
-\end{verbatim}
-
-This function is similar to the formatCases function.
-In some of the population data set, there will be a group of the people who are
-age $85$ and plus, marked as: \small $M85PLUS$ or $F85PLUS$ \large for male and female respectively.
-The first part of the formatPopulation function is to change them to \small $M85\_89$ or $F85\_89$,
-\large so that they will have the same format as the other sex and age group, and that will
-simplifies our manipulation later.
-The second part of the function is to reshape the population data set using the function
-\texttt{reshape}. So that it changes the population data set from the wide format
-to the long format, and it will have the same format as the case data set. This step make
-the merge of case data set and population data set easier in the future calculations.
-
-\item 
-\begin{verbatim}
-''getRates'' <-
-function(casedata, popdata, formula, family=poisson,
-minimumAge=0, maximumAge=100, S=c("M", "F"))
-\end{verbatim}
-
-This is the function we constructed to fit a Generalized Linear Model (\texttt{glm} function) to
-the case data set and the population data set. Then we get out estimated coefficients for the future
-prediction.
-The argument \texttt{casedata} and \texttt{popdata} are the places where we put case data set
-and population data set correspondingly. \texttt{formula} is the place we fit a formula function
-in the \texttt{glm} model. i.e : $~age*sex $. \texttt{family} is the place we
-choose what distribution we want to fit in the \texttt{glm} model. By default we set it to be Poisson model, it
-can be binomial as well.
-\texttt{$minimumAge=0$} and \texttt{$maximumAge=100$} are the arguments to control the sample's age range. 
-By default it is $age \in ($0, 100$)$ , which includes all the cases. You can change the range to ($0$, $10$) 
-for chose all the kids' data set.  \texttt{S} is in charge of control what kind of sex group we want to choose. 
-By default it choose both the male and female.
-
-\item 
-\begin{verbatim}
-''getSMR'' <-
-function(model, population, cases=NULL, regionCode="CSDUID",
-  regionCodeCases="CSD2006")
-\end{verbatim}
-
-Function \texttt{getSMR} is used to do the prediction using the model we get from the
-\texttt{getRates} function. Eventually, it returns a shape file (if the population data set
-we insert the function is a shape file) contains everything of the original population
-data set and several columns : expected, logExpected, cases, SMR. expected are the expected
-case numbers we estimated using the model we get from \texttt{getRates} function. logExpected
-are the log values of the expected values. cases are the number of cases we got from the
-case data set (if we fit a poisson distribution it will be the aggregate number of cases by
-different Decent Areas. SMR is the rate of observed number of cases over expected number of cases.
-
-The argument \texttt{model} is the glm model estimated coefficients we got from the \texttt{getRates} function.
-\texttt{population} is the population data set, it can be a shape file. \texttt{cases} is the number of
-case data set. \texttt{regionCodeCases} and \texttt{regionCode} are used to indicate the region for case data and
-population data sets respectively.
-
-\item 
-\begin{verbatim}
-''area'' <- function(sp)
-\end{verbatim}
-Function \texttt{area} is used to calculate the area of each region, return a vector of areas. The argument \texttt{sp} is a single 
-spatial polygon object.
-\item 
-\begin{verbatim}
-''mergeBugsData'' <- function(x, bugsSummary, by.x = NULL, newcol="mean", ...) 
-\end{verbatim}
-Function \texttt{mergeBugsData} is used to merge the results from the \texttt{bugs} function to the other data set i.e population data set.
-
-The argument \texttt{x} is the data set we want the result from the \texttt{bugs} to merge to. \texttt{bugsSummary} is the simulation result
-get from the \texttt{bugs} function.
-\texttt{by.x} is the common column in both the {x} and {bugsSummary} data set.
-\texttt{newcol} is the summary statistic that to be merged back to the data frame.
-\end{enumerate}
-
-\subsection{The Examples}
-We will take the ontario long cancer data set as an example. Due to confidential needs, we simulated the case (cancer) data set.  
-For example, we have a case data set as:
-
-library
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{cccccc}
-Year& $CSD2006$ & $CD2006$ & $PR2006$ & AGE\_SEX\_GROUP & Cases \\
-$1999$&$3501005$&$3501$&$35$&$210$&$3$ \\
-$1999$&$3501005$&$3501$&$35$&$213$&$1$ \\
-$1999$&$3501005$&$3501$&$35$&$214$&$6$ \\
-$1999$&$3501005$&$3501$&$35$&$215$&$1$ \\
-$1999$&$3501005$&$3501$&$35$&$216$&$3$ \\
-\end{tabular}
-\end{center}
-\end{table}
-
-After applying the \texttt{formatCases} function, it will be:
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{cccccccc}
- Year& $CSD2006$ & $CD2006$ & $PR2006$ & AGE\_SEX\_GROUP &Cases &sex &age    \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $109$   &  $1$ &  M  &$45$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $110$   &  $1$ &  M  &$50$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $111$   &  $3$ &  M  &$55$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $112$   &  $1$ &  M  &$60$ \\
- $1999$& $3501005 $ & $3501 $  &  $35 $     &     $113$   &  $2$ &  M  &$65$ \\
- \end{tabular}
-\end{center}
-\end{table}
-
-So that we can see, there are two new columns: sex and age.
-
-Now let's look at the \texttt{formatPopulation} function:
-
-For example, we have a population data set as:
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{cccccc}
-
-  CSDUID   &     CSDNAME  & CSDTYPE & PRUID  & PRNAME & CDUID \\
- $3501005$ & South Glengarry  &    TP  & $35$ &Ontario & $3501$ \\
- $3501011$ & South Stormont   &   TP   & $35$ &Ontario & $3501$  \\
- $3501012$ &      Cornwall    &  CY    &$35$  &Ontario & $3501$  \\
- $3501020$ &  South Dundas    &  TP    &$35$  &Ontario & $3501$  \\
- $3501030$ &  North Dundas    &  TP    &$35$  &Ontario & $3501$  \\
- $3501042$ &  North Stormont  &    TP  &  $35$& Ontario&  $3501$ \\
-
-\end{tabular}
-\end{center}
-\end{table}
-
-After applying the \texttt{formatPopulation} function, it will be:
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{cccccccc}
-   CSDUID &    cases&\ldots & CDUID & \ldots &expected    &logExpected  &     SMR    \\
- $3501005$&    $28$& \ldots & $3501$& \ldots &$8.433922$  &  $2.132262$ &$3.319926 $ \\
- $3501007$&    $ 0$& \ldots & $3501$& \ldots &$      NA$  &  $      NA$ &$      NA $ \\
- $3501011$&    $24$& \ldots & $3501$& \ldots &$7.143624$  &  $1.966220$ &$3.359639 $ \\
- $3501012$&    $52$& \ldots & $3501$& \ldots &$28.673675$ &  $ 3.355979$&$ 1.813510$ \\
- $3501020$&    $22$& \ldots & $3501$& \ldots &$6.654331$  &  $1.895268$ &$3.306117 $ \\
- $3501030$&    $19$& \ldots & $3501$& \ldots &$5.855935$  &  $1.767456$ &$3.244572 $ \\
- \end{tabular}
- \end{center}
- \end{table}
-
-There are four new columns added in: cases, this is the number we got from the case data set;
-expected, this is the predicted number of cases we calculated using the glm function;
-logexpected are the \mbox{log} values of the expected numbers; SMR is the rate of \[\frac {oberved(cases)}{expected(expected)}.\]
-
-if we ran the following code:
-\begin{verbatim}
-> model <- getRates( casedata, popdata, ~age*sex, S=c("M","F"))
-\end{verbatim}
-
-The result of model will be the predicted coefficients of different age and sex groups:
-\begin{verbatim}
-> model
-
-Call:  glm(formula = formula, family = family, data = newdata)
-
-Coefficients:
-(Intercept)         age5        age10        age15        age20        age25
-  -13.41885     -0.07403      0.52666      0.87985      1.62710      2.35525
-      age30        age35        age40        age45        age50        age55
-    3.49147      4.32618      5.12292      5.49001      5.96051      5.79087
-      age60        age65        age70        age75        age80         sexM
-    7.03365      7.11180      7.22133      7.24461      7.10451      0.67377
-  age5:sexM   age10:sexM   age15:sexM   age20:sexM   age25:sexM   age30:sexM
-         NA           NA           NA     -1.08827     -0.20732     -1.65308
- age35:sexM   age40:sexM   age45:sexM   age50:sexM   age55:sexM   age60:sexM
-   -0.81485     -1.05938     -0.60803     -0.59874           NA     -0.73841
- age65:sexM   age70:sexM   age75:sexM   age80:sexM
-   -0.38252     -0.25474     -0.20245           NA
-
-Degrees of Freedom: 28 Total (i.e. Null);  0 Residual
-Null Deviance:      10680
-Residual Deviance: -5.529e-14   AIC: 229.2
-
-\end{verbatim}
-
-\section{Summary and Future Development}
-
-\section{Bibliography}
-
 \end{document}

### 2009-05-11 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@293 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-\documentclass{article}
-\usepackage{Sweave}
-\title{Priors for swine flu}
+\documentclass{beamer}
+\usepackage[nogin]{Sweave}
+\title{Bayesian statistics and  swine flu}
+\author{Patrick Brown, Dionne Gesink Law, Peter Neal} 
+\usepackage{amsmath,subfigure}
+
+\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}
 
 \begin{document}
-
+\begin{frame}
 \maketitle
 
+\end{frame}
+
+<<setup,fig=false>>=
+options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
+# source in all the R files
+Rfiles=system("CMD /K dir /B ..\\..\\pkg\\pandemic\\R\\*.R", intern=T)
+Rfiles = grep("\\.R$", Rfiles, value=T)
+for(D in Rfiles) {
+	source(paste("..\\..\\pkg\\pandemic\\R\\", D, sep=""))
+#$
+}
+@
+
+\begin{frame}
+\frametitle{The Problem}
+\begin{itemize}
+	\item Some people will get infected with swine flu
+	\item Some will get hospitalized, some will die
+	\item How do we fatality rate and the hospitalization rate?
+	\item Problem: censoring, some individuals in the dataset have not yet had their hospitalization and/or fatality
+	\item Crude rates underestimate the rates, and result in estimates increasing over time
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{The Model}
+
+\begin{block}{Mild}
+\begin{center}
+ Infection $\stackrel{f_I(t)}{\Longrightarrow}$ Onset $\stackrel{f_{OM}(t;)}{\Longrightarrow}$ \\ Medical  $\stackrel{f_{MR}(t)}{\Longrightarrow}$ Recovery
+\end{center}
+
+\end{block}
+
+\begin{block}{Serious} 
+\begin{center}
+Infection$\stackrel{f_I(t)}{\Longrightarrow}$ Onset $\stackrel{f_{OS}(t)}{\Longrightarrow}$ Medical $\stackrel{f_{SH}(t)}{\Longrightarrow}$\\ Hospitalization $\stackrel{f_{SR}(t;)}{\Longrightarrow}$ Recovery
+\end{center}\end{block}
+\begin{block}{Deadly} \begin{center}Infection $\stackrel{f_I(t)}{\Longrightarrow}$ Onset $\stackrel{f_{OD}(t)}{\Longrightarrow}$ Medical $\stackrel{f_{HD}(t)}{\Longrightarrow}$\\ Hospitalization $\stackrel{f_{D}(t)}{\Longrightarrow}$ Death
+\end{center}
+\end{block}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Distributions}
+\begin{itemize}
+	\item $f(t;\mu,\nu,\delta)$ is the distribution for time to an event.  
+	\item \ldots a rounded, zero-inflated Weibull
+\begin{align*}
+Z\sim&\text{Bern}(\delta)\\
+[Y|Z=1] =& 0 \text{w.p.} 1\\
+pr(T=t|Z=0) =&  \text{pr}(\max(t-0.5,0) < y < t+0.5)\\
+y \sim & \text{Weibull}(\mu/\Gamma(1 + 1/\nu), \nu)
+\end{align*}
+\item $\delta$ is the probability of transitioning immediately for an administrative (not biological) reason, 
+\begin{itemize}
+	\item i.e.\ arriving dead at the hospital.
+\end{itemize}
+\item Barring administrative zeros, $\mu = E(T|Z=0)$ is the average time to an event
+\begin{itemize}
+	\item $\mu_I$ is average time to infection
+\end{itemize}
+\item $\nu$ is a Weibull shape parameter
+\begin{itemize}
+	\item $\nu=1$ is constant hazards
+	\item $\nu > 1$ results in more regular event times.
+\end{itemize}
+
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Priors}
+\begin{itemize}
+	\item What does expert opinion tell us these parameters should be?
+	\item Based on previous experience, not the current data.
+	\item We need a distribution, not a single value
+\begin{itemize}
+	\item If we `knew' the correct value, there would be no need for this model
+\end{itemize}
+\item Gamma or Beta priors, parametrized by mean and standard deviation
+\item Average time to onset of symptoms is expected to be 3.5 days, but the average time could be a day longer or shorter.
+\begin{itemize}
+	\item $E(\mu_I = 3.5)$ and $sd(\mu_I = 0.5$
+	\item $\mu_I \sim \text{Gamma}(3.5^2/0.5^2 , 3.5/0.5^2 )$
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Proportion parameters}
+
+\begin{itemize}
+	\item $\rho_D$ is the fatality ratio, or probability of a deadly infection
+	\item $\rho_S$ is the proportion of non-fatalities which are hospitalized
+	\item so the probability of being hospitalized is $\rho_D + (1-\rho_D)\rho_S$
+	\item parametrize this way to ensure the probabilities of being Mild, Severe, or Deadly add to 1.
+	\item beta priors on $\rho_D$ and $\rho_S$, parametrized by mean and standard deviation
+	\item i.e.\ $\rho_D$ is expected to be $0.1$, with $sd(\rho_D = 0.05)$
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{What we need}
+\begin{verbatim}
+transition parameter mean sd distribution
+InfOns	mean	1	1	gamma	
+InfOns	shape	1	1	gamma	
+InfOns	zeros	0.1	0.1	beta	
+OnsMedM	mean	1	1	gamma	
+OnsMedM	shape	1	1	gamma	
+OnsMedM	zeros	0.1	0.1	beta	
+OnsMedS	mean	1	1	gamma	
+OnsMedS	shape	1	1	gamma	
+OnsMedS	zeros	0.1	0.1	beta	
+OnsMedD	mean	1	1	gamma	
+OnsMedD	shape	1	1	gamma	
+OnsMedD	zeros	0.1	0.1	beta	
+\end{verbatim}
+\end{frame}
+
+\begin{frame}
+\frametitle{Prior graphs}
+<<priorGraph, fig=false,results=tex>>=
+priors= pandemicPriors()
+plotPrior(priors$InfOns, file="InfOns", tex=TRUE)
+@
+\end{frame}
+
+\begin{frame}
+\frametitle{Inference}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Simulated data}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{To do}
+\begin{itemize}
+	\item model the infection process
+	\item simulate hospital loads
+	\item covariates
+	\item spatial infection process
+\end{itemize}
+\end{frame}
+
 \end{document}

### 2009-05-11 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@294 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 OnsMedD	zeros	0.1	0.1	beta	
+MedRec	mean	1	1	gamma	
+MedRec	shape	1	1	gamma	
+MedRec	zeros	0.1	0.1	beta	
+MedRec	lost	1	1	beta	
 \end{verbatim}
@@
 plotPrior(priors$InfOns, file="InfOns", tex=TRUE)
+#$
+@
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Inference}
+\begin{itemize}
+	\item We have data $Y$ and unknown parameters $\theta$.
+	\item The model specifies $[Y|\theta]$
+	\item Priors are $[\theta]$.
+	\item Posteriors: $[\theta|Y] \propto [Y|\theta][\theta]$
+	\item Posteriors show what we think the parameters could be, given the data and the prior distributions
+	\item There's no data on infection times, so the posterior for $\mu_{I}$ is the same as the prior.
+\end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Simulated data}
+<<simData,echo=true,fig=false>>=
+params = pandemicParams()
+ data = simEpidemic(params, 30)
 @
@@
 \begin{frame}
-\frametitle{Inference}
-
-\end{frame}
-
-\begin{frame}
 \frametitle{Simulated data}
-
+<<simDataTable,results=tex,fig=false>>=
+toPrint = data[1:12,]
+toPrint[is.na(toPrint)] = " "
+toPrint = as.matrix(toPrint)
+latex(toPrint, file="",center="none", title="", table.env=F)
+@
 \end{frame}

### 2009-08-12 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@318 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@354 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \title{Bayesian statistics and  swine flu}
-\author{Patrick Brown, Dionne Gesink Law, Peter Neal} 
+\author{Patrick Brown$^1$,  Peter Neal$^2$}
+\institute{1:Cancer Care Ontario and Dalla Lana School of Public Health, University of Toronto; 2: University of Manchester} 
 \usepackage{amsmath,subfigure}
@@
 # source in all the R files
-Rfiles=system("CMD /K dir /B ..\\..\\pkg\\pandemic\\R\\*.R", intern=T)
-Rfiles = grep("\\.R$", Rfiles, value=T)
-for(D in Rfiles) {
-	source(paste("..\\..\\pkg\\pandemic\\R\\", D, sep=""))
-#$
-}
+library(pandemic)
 @
+
+\begin{frame}
+\frametitle{Swine Flu in Ontario}
+\begin{itemize}
+	\item Many Canadians holiday in Mexico during the winter and  spring
+
+\begin{itemize}
+	\item Swine flu was probably established in Ontario before the outbreak was detected.
+\item Few control measures were adopted
+\end{itemize}
+\item Ontario's had recent public health crises
+\begin{itemize}
+	\item SARS, Walkerton water quality problems
+	\item Ontario Agency for Health Protection and Promotion created
+	\item Expert opinion abounds from mathematical modellers and infectious disease epidemiologists
+	\item Detailed models are used for the biological process, with largely ad-hoc inference and treatment of uncertainty. 
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{The Data}
+\begin{itemize}
+	\item During the first 6 weeks of the epidemic, we can assume most cases were identified.
+
+\begin{itemize}
+	\item Ill individuals were encouraged to seek medical help
+	\item Practitioners were encouraged to send samples to labs for testing
+\end{itemize}
+	\item After 6 weeks, testing was discouraged to ease the burden on public health labs.
+	\item Roughly 1000 cases in the first 6 weeks? 
+\end{itemize}
+\end{frame}
 
@@
 \begin{frame}
+\frametitle{The Goal}
+\begin{itemize}
+	\item Provide good estimates (and inference) about hospitalization and fatality rates in the early stages of an epidemic
+	\item To cope with sparse data, allow for expert opinion.
+	\item Allow for variations in rates by age
+	\item and account for individual's previous medical conditions?
+\item Predict the number of hospital beds required for a given number of infections.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \frametitle{The Model}
@@
 \begin{center}
- Infection $\stackrel{f_I(t)}{\Longrightarrow}$ Onset $\stackrel{f_{OM}(t;)}{\Longrightarrow}$ \\ Medical  $\stackrel{f_{MR}(t)}{\Longrightarrow}$ Recovery
+ Infection $\stackrel{f_I(t)}{\Longrightarrow}$ Onset $\stackrel{f_{OM}(t)}{\Longrightarrow}$ \\ Medical  $\stackrel{f_{MR}(t)}{\Longrightarrow}$ Recovery
 \end{center}
@@
 \begin{frame}
+\frametitle{The Data (again)}
+\begin{itemize}
+	\item Everyone has a date of a medical visit
+	\item Some have date of onset of symptoms
+	\item Most infections are mild and no additional data results
+	\item Some end up in hospital, resulting in either recovery or death
+	\item Longitudinal data on each case can be assembled\ldots in theory
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
 \frametitle{Distributions}
@@
 Z\sim&\text{Bern}(\delta)\\
-[Y|Z=1] =& 0 \text{w.p.} 1\\
+[Y|Z=1] =& 0 \text{ w.p.} 1\\
 pr(T=t|Z=0) =&  \text{pr}(\max(t-0.5,0) < y < t+0.5)\\
@@
 <<priorGraph, fig=false,results=tex>>=
-priors= pandemicPriors()
+priors= readPrior("bayesPandemicPriors.txt")
 plotPrior(priors$InfOns, file="InfOns", tex=TRUE)
@@
 \begin{itemize}
+	\item simulate hospital loads
 	\item model the infection process
-	\item simulate hospital loads
+	\item spatial infection process
 	\item covariates
-	\item spatial infection process
+
 \end{itemize}

### 2009-08-12 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@355 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Patrick Brown$^1$,  Peter Neal$^2$}
-\institute{1:Cancer Care Ontario and Dalla Lana School of Public Health, University of Toronto; 2: University of Manchester} 
+\institute{1:Cancer Care Ontario and Dalla Lana School of Public Health, University of Toronto; 2: School of Mathematics, University of Manchester} 
 \usepackage{amsmath,subfigure}
@@
 \end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Severity of infection}
+
 \end{frame}
@@
 
+\begin{frame}
+\frametitle{The Algorithm}
+data augmentation
+
+random walk metropolis on parameters
+\end{frame}
+
 \begin{frame}[fragile]
@@
 \begin{frame}
+\frametitle{Results}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Age?}
+
+piecewise constant 5yr?
+
+change points?
+
+note that after data augmentation it's a standard problem.
+
+\end{frame}
+
+\begin{frame}
 \frametitle{To do}

### 2009-08-12 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@356 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{frame}
-\frametitle{Severity of infection}
+\frametitle{Types of infection}
 
@@
 \frametitle{What we need}
-\begin{verbatim}
-transition parameter mean sd distribution
-InfOns	mean	1	1	gamma	
-InfOns	shape	1	1	gamma	
-InfOns	zeros	0.1	0.1	beta	
-OnsMedM	mean	1	1	gamma	
-OnsMedM	shape	1	1	gamma	
-OnsMedM	zeros	0.1	0.1	beta	
-OnsMedS	mean	1	1	gamma	
-OnsMedS	shape	1	1	gamma	
-OnsMedS	zeros	0.1	0.1	beta	
-OnsMedD	mean	1	1	gamma	
-OnsMedD	shape	1	1	gamma	
-OnsMedD	zeros	0.1	0.1	beta	
-MedRec	mean	1	1	gamma	
-MedRec	shape	1	1	gamma	
-MedRec	zeros	0.1	0.1	beta	
-MedRec	lost	1	1	beta	
-\end{verbatim}
+\input{bayesPandemicPriors.txt}
 \end{frame}
@@
 <<simData,echo=true,fig=false>>=
-params = pandemicParams()
- data = simEpidemic(params, 30)
+params = pandemicParams( 
+  MedHospD = c(mean = 1.5, shape = 2, zeros = 0.2)
+  )
+
+data = simEpidemic(params, delta=3, days=20)
+
+data[1:10,-1]
+@
+
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Results}
+<<mcmcResults,echo=true,fig=false>>=
+paramSample = mcmcPandemic(data,params,priors,sigma=0.1,runs=100)
+dim(paramSample)
+colnames(paramSample)
 @
@@
 \begin{frame}
-\frametitle{Simulated data}
-<<simDataTable,results=tex,fig=false>>=
-toPrint = data[1:12,]
-toPrint[is.na(toPrint)] = " "
-toPrint = as.matrix(toPrint)
-latex(toPrint, file="",center="none", title="", table.env=F)
+\frametitle{}
+\begin{block}{trace plots}
+<<tracePlot>>=
+plot(paramSample[,"probs.D"], type="l")
 @
-\end{frame}
-
-\begin{frame}
-\frametitle{Results}
+\end{block}
 

### 2009-08-12 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@357 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \institute{1:Cancer Care Ontario and Dalla Lana School of Public Health, University of Toronto; 2: School of Mathematics, University of Manchester} 
-\usepackage{amsmath,subfigure}
+\usepackage{amsmath,subfigure,alltt}
 
@@
 \frametitle{What we need}
+\begin{alltt}
 \input{bayesPandemicPriors.txt}
+\end{alltt}
 \end{frame}
@@
 \begin{frame}
-\frametitle{}
+\frametitle{Fatality probability}
+\begin{columns}
+\column{0.5\textwidth}
 \begin{block}{trace plots}
-<<tracePlot>>=
+<<tracePlot,fig=true>>=
 plot(paramSample[,"probs.D"], type="l")
+abline(h=params$probs["D"], col="red")
 @
 \end{block}
+\column{0.5\textwidth}
 
+\begin{block}{Histogram}
+<<histPost>>=
+hist(paramSample[,"probs.D"], breaks=12)
+abline(v=params$probs["D"], col="red")
+@
+\end{block}
+\end{columns}
 \end{frame}
+\end{document}
 

### 2009-08-12 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@358 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{amsmath,subfigure,alltt}
+\setkeys{Gin}{width=\columnwidth}
 
@@
 \end{frame}
-\end{document}
+
+\begin{frame}[fragile]
+\frametitle{Summary Statistics}
+<<summaryStats,fig=false>>=
+postSummary=t(apply(paramSample, 2, function(qq) 
+  c(mean=mean(qq), quantile(qq, prob=c(0.025, 0.5, 0.975)))))
+postSummary = cbind(true=unlist(params)[rownames(postSummary)], postSummary)
+
+postSummary[1:12,]
+@
+\end{frame}
+
+\begin{frame}
+\frametitle{Posterior and prior, fatality prob }
+<<priorPost, fig=false,results=tex>>=
+library(ellipse)
+plotPrior(priors, paramSample, transition="HospDeath",file="postFatality", tex=TRUE)
+@
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Hospital demand}
+<<hospLoad,echo=T>>=
+# generate hospital times
+hospSample = simHospitals(paramSample , Ndays=20, NinfectionsPerDay=10)
+
+# graph hospital times
+plotHospitals(hospSample)
+@
+\end{frame}
 

### 2009-08-13 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@359 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2009-08-12 docs/pandemic/bayesPandemic.rnw
@@
 <<setup,fig=false>>=
-options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
+options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)), width=58)
 # source in all the R files
@@
 	\item Some will get hospitalized, some will die
-	\item How do we fatality rate and the hospitalization rate?
+	\item How do we estimate the fatality and hospitalization rates?
 	\item Problem: censoring, some individuals in the dataset have not yet had their hospitalization and/or fatality
@@
 \frametitle{Types of infection}
+All infections are one of the following types:
+\begin{description}
+	\item[Mild:] The individual will recover without needing to be hospitalized
+	\item[Serious:] The infection is severe enough that hospitalization will be required before the individual recovers
+	\item[Deadly:] Following hospitalization, the patient dies of the infection.
+\end{description}
+When a case is first identified, it's not known what type it is.
 
@@
 \begin{itemize}
-	\item $f(t;\mu,\nu,\delta)$ is the distribution for time to an event.  
+	\item $ f(t;\mu,\nu,\delta) = pr(T=t)$ is the distribution for time to an event.  
 	\item \ldots a rounded, zero-inflated Weibull
@@
 Z\sim&\text{Bern}(\delta)\\
-[Y|Z=1] =& 0 \text{ w.p.} 1\\
+[T|Z=1] =& 0 \text{ w.p.} 1\\
 pr(T=t|Z=0) =&  \text{pr}(\max(t-0.5,0) < y < t+0.5)\\
@@
 \begin{itemize}
-	\item $E(\mu_I = 3.5)$ and $sd(\mu_I = 0.5$
+	\item $E(\mu_I = 3.5)$ and $sd(\mu_I)	 = 0.5$
 	\item $\mu_I \sim \text{Gamma}(3.5^2/0.5^2 , 3.5/0.5^2 )$
@@
 	\item beta priors on $\rho_D$ and $\rho_S$, parametrized by mean and standard deviation
-	\item i.e.\ $\rho_D$ is expected to be $0.1$, with $sd(\rho_D = 0.05)$
+	\item i.e.\ $\rho_D$ is expected to be $0.1$, with $sd(\rho_D) = 0.05$
 \end{itemize}
@@
 \frametitle{What we need}
+{\footnotesize
 \begin{alltt}
@@
 \end{alltt}
+}
 \end{frame}
@@
 \frametitle{The Algorithm}
-data augmentation
+\begin{block}{Data augmentation}
+\begin{itemize}
+	\item Fix the parameter values
+	\item Simulate the missing data: infection type, times of infection, onset, hospitalization, recovery, and/or death.
+	\item Can simulate directly, as conditioning is easy.
+\end{itemize}
+\end{block}
 
-random walk metropolis on parameters
+\begin{block}{Random walk metropolis}
+\begin{itemize}
+	\item With the full augmented data, the likelihood is simple as observations are independent.
+	\item Propose new parameters with a Gaussian random walk, accept with the appropriate probability.
+\end{itemize}
+\end{block}
+
 \end{frame}
@@
 params = pandemicParams( 
-  MedHospD = c(mean = 1.5, shape = 2, zeros = 0.2)
+	InfOns = c(mean=1.5, shape=2, zeros=.1),
+	OnsMedM= c(mean=1, shape=1, zeros=.1 ),
+	OnsMedS= c(mean=1, shape=1, zeros=.1 ),
+	OnsMedD= c(mean=1, shape=1, zeros=.1 ),
+	MedRec = c(mean=3, shape=2, zeros=.1, lost=0.95 ),
+  MedHospS = c(mean = 2, shape = 2, zeros = 0.1),
+  MedHospD = c(mean = 2, shape = 2, zeros = 0.1),
+	HospRec = c(mean=2, shape=1.5, zeros=.01 ),
+	HospDeath = c(mean=2, shape=0.1, zeros=.05 ),
+ 	probs = c(M=0.7, S=0.25, D=0.05)
   )
 
-data = simEpidemic(params, delta=3, days=20)
+data = simEpidemic(params, delta=4, days=20)
 
-data[1:10,-1]
+data[1:8,-1]
 @
@@
 <<mcmcResults,echo=true,fig=false>>=
-paramSample = mcmcPandemic(data,params,priors,sigma=0.1,runs=100)
+paramSample = mcmcPandemic(data,params,priors,sigma=0.05,runs=600)
+paramSample = paramSample[seq(100,dim(paramSample)[1], by=5),]
 dim(paramSample)
@@
 \end{frame}
-\end{document}
+
+\begin{frame}[fragile]
+\frametitle{Summary Statistics}
+<<summaryStats,fig=false>>=
+postSummary=t(apply(paramSample, 2, function(qq) 
+  c(mean=mean(qq), quantile(qq, prob=c(0.025, 0.5, 0.975)))))
+postSummary = cbind(true=unlist(params)[rownames(postSummary)], postSummary)
+
+postSummary[1:12,-4]
+@
+\end{frame}
+
+\begin{frame}
+\frametitle{Posterior and prior, Hosp $\Rightarrow$ Death }
+<<priorPost, fig=false,results=tex>>=
+library(ellipse)
+plotPrior(priors, paramSample, transition="MedHospS",file="MedHospS", tex=TRUE)
+@
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Hospital demand}
+<<hospLoad,echo=T>>=
+# generate hospital times
+hospSample = simHospitals(paramSample , Ndays=20, NinfectionsPerDay=20)
+
+# graph hospital times
+plotHospitals(hospSample)
+@
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{With rising and falling incidence}
+{\scrpitsize
+<<hospLoad2,echo=T,height=2,width=4>>=
+# generate hospital times
+hospSample2 = simHospitals(paramSample , Ndays=40, 
+	NinfectionsPerDay=c(seq(from=1,by=2, len=10), seq(from=20, len=10, by=2), rep(2, 20)))
+
+# graph hospital times
+plotHospitals(hospSample2)
+@
+}
+\end{frame}
 
@@
 \frametitle{Age?}
-
-piecewise constant 5yr?
-
-change points?
-
-note that after data augmentation it's a standard problem.
-
+\begin{itemize}
+	\item Assume the fatality and hospitalization probabilities vary with age
+	\item After data augmentation, we have a binary response with a smoothly-varying probability
+	\item Use a canned Bayesian semi-parametric routine to generate a new age curve at every iteration.
+	\item {\tt DPpackage}?
+	\item Informative prior on the average or rate at a baseline age
+	\item Strong prior that the curve is flat.
+\end{itemize}
 \end{frame}
@@
 \begin{itemize}
-	\item simulate hospital loads
+	\item more covariates	
 	\item model the infection process
 	\item spatial infection process
-	\item covariates
-
 \end{itemize}
@@
 
+\begin{frame}
+
+\frametitle{Conclusions}
+
+\begin{itemize}
+	\item Use the data even if there isn't much of it
+	\item If there are many cases and no fatalities, at least we can get an upper bound on the fatality rate
+	\item When data are plentiful and informative, there is a case to be made for non-parametric survival methods
+	\item But if rates change with age, sex, geography, or existing health conditions, even late stage epidemic data might not be sufficient for standard survival methods.
+\end{itemize}
+
+\end{frame}
+
 \end{document}
+
+D75A0BD6A6A1

### 2009-08-14 docs/pandemic/bayesPandemic.rnw
fixed chain mixing
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@360 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	HospRec = c(mean=2, shape=1.5, zeros=.01 ),
-	HospDeath = c(mean=2, shape=0.1, zeros=.05 ),
+	HospDeath = c(mean=2, shape=0.8, zeros=.05 ),
  	probs = c(M=0.7, S=0.25, D=0.05)
+# 	probs = c(M=0.979, S=0.02, D=0.001)
   )
@@
 <<mcmcResults,echo=true,fig=false>>=
-paramSample = mcmcPandemic(data,params,priors,sigma=0.05,runs=600)
+scale = mcmcScale(params, 0.075,minScale = 0.05 )
+paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=60)
 paramSample = paramSample[seq(100,dim(paramSample)[1], by=5),]

### 2009-08-21 docs/hamilton/graph.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@362 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

% C:\brownpe\research\diseasemapping\docs\hamilton\Figures\graph.jdr
% Created by Jpgfdraw version 0.5.4b
% Aug 18, 2009 9:28:53 AM
\begin{pgfpicture}{0bp}{0bp}{347.0bp}{247.5bp}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{93.75bp}{213.5bp}}
\pgfpathcurveto{\pgfpoint{93.75bp}{196.103031bp}}{\pgfpoint{76.624893bp}{182.0bp}}{\pgfpoint{55.5bp}{182.0bp}}
\pgfpathcurveto{\pgfpoint{34.375109bp}{182.0bp}}{\pgfpoint{17.25bp}{196.103031bp}}{\pgfpoint{17.25bp}{213.5bp}}
\pgfpathcurveto{\pgfpoint{17.25bp}{230.89697bp}}{\pgfpoint{34.375109bp}{245.0bp}}{\pgfpoint{55.5bp}{245.0bp}}
\pgfpathcurveto{\pgfpoint{76.624893bp}{245.0bp}}{\pgfpoint{93.75bp}{230.89697bp}}{\pgfpoint{93.75bp}{213.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{318.5bp}{214.5bp}}
\pgfpathcurveto{\pgfpoint{318.5bp}{197.103031bp}}{\pgfpoint{301.374893bp}{183.0bp}}{\pgfpoint{280.25bp}{183.0bp}}
\pgfpathcurveto{\pgfpoint{259.125109bp}{183.0bp}}{\pgfpoint{242.0bp}{197.103031bp}}{\pgfpoint{242.0bp}{214.5bp}}
\pgfpathcurveto{\pgfpoint{242.0bp}{231.89697bp}}{\pgfpoint{259.125109bp}{246.0bp}}{\pgfpoint{280.25bp}{246.0bp}}
\pgfpathcurveto{\pgfpoint{301.374893bp}{246.0bp}}{\pgfpoint{318.5bp}{231.89697bp}}{\pgfpoint{318.5bp}{214.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{181.25bp}{105.75bp}}
\pgfpathcurveto{\pgfpoint{181.25bp}{88.353031bp}}{\pgfpoint{164.124893bp}{74.25bp}}{\pgfpoint{143.0bp}{74.25bp}}
\pgfpathcurveto{\pgfpoint{121.875109bp}{74.25bp}}{\pgfpoint{104.75bp}{88.353031bp}}{\pgfpoint{104.75bp}{105.75bp}}
\pgfpathcurveto{\pgfpoint{104.75bp}{123.14697bp}}{\pgfpoint{121.875109bp}{137.25bp}}{\pgfpoint{143.0bp}{137.25bp}}
\pgfpathcurveto{\pgfpoint{164.124893bp}{137.25bp}}{\pgfpoint{181.25bp}{123.14697bp}}{\pgfpoint{181.25bp}{105.75bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{266.75bp}{107.25bp}}
\pgfpathcurveto{\pgfpoint{266.75bp}{89.853031bp}}{\pgfpoint{249.624893bp}{75.75bp}}{\pgfpoint{228.5bp}{75.75bp}}
\pgfpathcurveto{\pgfpoint{207.375109bp}{75.75bp}}{\pgfpoint{190.25bp}{89.853031bp}}{\pgfpoint{190.25bp}{107.25bp}}
\pgfpathcurveto{\pgfpoint{190.25bp}{124.64697bp}}{\pgfpoint{207.375109bp}{138.75bp}}{\pgfpoint{228.5bp}{138.75bp}}
\pgfpathcurveto{\pgfpoint{249.624893bp}{138.75bp}}{\pgfpoint{266.75bp}{124.64697bp}}{\pgfpoint{266.75bp}{107.25bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{166.5bp}{29.75bp}}
\pgfpathcurveto{\pgfpoint{166.5bp}{14.838318bp}}{\pgfpoint{129.563492bp}{2.75bp}}{\pgfpoint{84.0bp}{2.75bp}}
\pgfpathcurveto{\pgfpoint{38.436508bp}{2.75bp}}{\pgfpoint{1.5bp}{14.838318bp}}{\pgfpoint{1.5bp}{29.75bp}}
\pgfpathcurveto{\pgfpoint{1.5bp}{44.661682bp}}{\pgfpoint{38.436508bp}{56.75bp}}{\pgfpoint{84.0bp}{56.75bp}}
\pgfpathcurveto{\pgfpoint{129.563492bp}{56.75bp}}{\pgfpoint{166.5bp}{44.661682bp}}{\pgfpoint{166.5bp}{29.75bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{345.5bp}{28.5bp}}
\pgfpathcurveto{\pgfpoint{345.5bp}{13.588318bp}}{\pgfpoint{308.563492bp}{1.5bp}}{\pgfpoint{263.0bp}{1.5bp}}
\pgfpathcurveto{\pgfpoint{217.436508bp}{1.5bp}}{\pgfpoint{180.5bp}{13.588318bp}}{\pgfpoint{180.5bp}{28.5bp}}
\pgfpathcurveto{\pgfpoint{180.5bp}{43.411682bp}}{\pgfpoint{217.436508bp}{55.5bp}}{\pgfpoint{263.0bp}{55.5bp}}
\pgfpathcurveto{\pgfpoint{308.563492bp}{55.5bp}}{\pgfpoint{345.5bp}{43.411682bp}}{\pgfpoint{345.5bp}{28.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{77.25bp}{84.5bp}}
\pgfpathlineto{\pgfpoint{77.25bp}{125.75bp}}
\pgfpathlineto{\pgfpoint{19.5bp}{125.75bp}}
\pgfpathlineto{\pgfpoint{19.5bp}{84.5bp}}
\pgfpathlineto{\pgfpoint{77.25bp}{84.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{338.75bp}{89.25bp}}
\pgfpathlineto{\pgfpoint{338.75bp}{130.5bp}}
\pgfpathlineto{\pgfpoint{281.0bp}{130.5bp}}
\pgfpathlineto{\pgfpoint{281.0bp}{89.25bp}}
\pgfpathlineto{\pgfpoint{338.75bp}{89.25bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{297.0bp}{182.75bp}}
\pgfpathlineto{\pgfpoint{313.5bp}{131.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{42.0bp}{126.5bp}}
\pgfpathlineto{\pgfpoint{50.25bp}{179.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{81.0bp}{190.25bp}}
\pgfpathlineto{\pgfpoint{127.5bp}{136.25bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{265.5bp}{184.25bp}}
\pgfpathlineto{\pgfpoint{240.75bp}{137.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{164.25bp}{134.0bp}}
\pgfpathlineto{\pgfpoint{246.75bp}{194.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{87.75bp}{59.75bp}}
\pgfpathlineto{\pgfpoint{126.0bp}{74.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{233.25bp}{73.25bp}}
\pgfpathlineto{\pgfpoint{261.75bp}{56.0bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{251.25bp}{216.5bp}}
\pgftext[left,base]{Cancer  $Y_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{30.0bp}{211.25bp}}
\pgftext[left,base]{Smoking $Z_{ik}$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{124.5bp}{115.25bp}}
\pgftext[left,base]{Smoking}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{124.25bp}{103.0bp}}
\pgftext[left,base]{Risk}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{124.25bp}{91.0bp}}
\pgftext[left,base]{$\rho_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{216.75bp}{115.25bp}}
\pgftext[left,base]{Cancer}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{217.5bp}{103.75bp}}
\pgftext[left,base]{Risk}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{224.25bp}{93.25bp}}
\pgftext[left,base]{$U_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{32.25bp}{109.25bp}}
\pgftext[left,base]{Income}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{32.25bp}{97.0bp}}
\pgftext[left,base]{$X_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{294.75bp}{114.5bp}}
\pgftext[left,base]{Pollution}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{295.25bp}{102.0bp}}
\pgftext[left,base]{$P_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{28.5bp}{28.25bp}}
\pgftext[left,base]{Neighbouring smoking}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{28.25bp}{16.0bp}}
\pgftext[left,base]{$W_j | j \sim i$ }
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{209.25bp}{36.5bp}}
\pgftext[left,base]{Neighbouring Cancer}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{209.25bp}{24.0bp}}
\pgftext[left,base]{$U_j | j \sim i$}
\end{pgfscope}
\end{pgfpicture}

### 2009-08-21 docs/hamilton/slides.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@362 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[xcolor=dvipsnames]{beamer}

\usetheme{Singapore}
\usecolortheme[named=RawSienna]{structure}

\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number] 
\usepackage[nogin]{c:/brownpe/share/Sweave}
\title{Beyond disease mapping: models for disease incidence data with spatial risk factors}
\author{Patrick Brown\\ Billy Chang, Susitha Wanigaratne}
\institute{Cancer Care Ontario\\ Dalla Lana School of Public Health, University of Toronto}

\usepackage{amsmath,tikz}
\graphicspath{{Figures/}}
\setkeys{Gin}{width=\textwidth}
\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}

\begin{document}

\frame{\maketitle}

\begin{frame}
\frametitle{Hamilton, Canada}
\includegraphics[trim=1cm 12cm 1cm 6cm,clip=true]{hamiltonMap.pdf}

\includegraphics{YorkBlvdHamiltonA.JPG}
\end{frame}

\begin{frame}

\frametitle{Lung cancer in Hamilton}
% left bottom right top

\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{Female-CancerExpected.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{Female-CancerExpected.pdf}
\end{frame}

\begin{frame}
\frametitle{Smoking in Hamilton}
\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{Female-eversmoking.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{Female-eversmoking.pdf}
\end{frame}

\begin{frame}
\frametitle{The Problem}
\begin{itemize}
	\item Create a model for predicting cancer rates in Hamilton, allowing for low counts in small areas
	\item Allow explanatory variables (smoking)
	\item But how do we get smoking rates?

\begin{itemize}
	\item And how well can we predict them?
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Issues with small area maps}

\begin{block}{Small counts}
	mostly 0 or 1
\begin{itemize}
\item The usual estimates of incidence rates perform poorly
	\item Methods based on the normal-distribution aren't appropriate
\end{itemize}

\end{block}
\begin{block}{Spatial dependence}
With regions close together being more similar than regions further apart
\begin{itemize}
	\item The usual regression models underestimate standard errors
  \item but this helps us deal with small counts
\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\frametitle{This talk}
will accomplish the following:
\begin{itemize}
\item  present a model for mapping disease risk in small areas with low case counts

\item  show how to incorporate risk factors observed imprecisely, like smoking rates

\item  apply these methods to mapping cancer risk in Hamilton at the DA level, adjusting for smoking rates.\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Cancer Data}
For each census area $i$ we have
\begin{itemize}
	\item Expected cancer case count $E_i$

\begin{itemize}
\item $E_i = C_i \theta$
	\item $C_i$ is the (vector) of populations by 5-year age group
	\item $\theta$ are the (assumed known) rates by age group
\end{itemize}
\item Count of cancer cases $Y_i$, often 0 or 1
\item Estimated pollution levels $P_i$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Smoking data}
From the Canadian Community Household Survey (CCHS), we have
\begin{itemize}
	\item $N_i$ individuals surveyed in region $i$, often 0.
	\item The age $A_{ij}$ and smoking status $Z_{ij}$ of individual $j$ in region $i$
	\item The average household income $X_i$ in the region.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Smoking model}
\vspace*{-0.5cm}
\begin{align*}
pr(Z_{ij} =1) = & \pi_{ij}\\
 \text{logit}(\pi_{ij}) = & f(A_{ij};\psi) + \rho_i\\
\rho_i = & X_i \alpha + W_i\\
 W_i \sim & \text{BYM}(\sigma_1, \sigma_2)
\end{align*}
\vspace*{-0.5cm}
\begin{itemize}
\item $f(a;\psi)$ is a smoothly-varying function of age
\begin{itemize}
	\item $\psi$ are the coefficients on linear splines with fixed knots
\end{itemize}
	\item The smoking risk surface $\rho_i$  is unknown.  
	\item It depends on income $X_i$ and residual spatial variation $W_i$
	\item $W_i$ is a Besag, York and Mollie model, which is smoothly varying in space

\begin{itemize}
	\item $W_i$ depends on the values $W_k$ in neighbouring regions $k$
	\item Technically, it's the sum of a Markov random field and a spatially independent process
	\item $\sigma_1$ is the variance of the spatially smooth part of $W_i$, and $\sigma_2$ is the variance of the rough part.
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Dealing with small cancer counts}

\begin{columns}
\column{0.5\textwidth}
\begin{block}{The Model}
\[
Y_i \sim \text{Poisson}(E_i\lambda_i)
\]
\begin{itemize}
	\item $Y_i$ is the number of cases in region $i$
	\item $E_i$ is the expected count
	\item $\lambda_i$ is the relative risk
\end{itemize}
\end{block}
\column{0.5\textwidth}
\begin{block}{The Poisson Distriution}
<<poisson,fig=true,echo=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
xseq = 0:14
toplot = dpois(xseq, 0.5)
names(toplot) = as.character(xseq)
toplot2 = dpois(xseq, 1.5)
barplot(toplot, width=0.4, xlim = range(xseq), space=4, xlab='count', ylab='pr(count)',
	col='red',border=NA)
barplot(toplot, width=0.21, add=T, space=8.8, col='white', xlim = range(xseq),border=NA, axes=F, axisnames=F)

barplot(toplot2, width=0.2, add=T, space=9, col='blue', xlim = range(xseq),border=NA, axes=F)
@

mean = \textcolor{red}{0.5},  \textcolor{blue}{1.5} 
\end{block}

\end{columns}

\begin{block}{Interpretation}
\begin{itemize}
	\item We don't observe the true rate $\lambda_i$
	\item but rather a noisy version of it in the form of case counts
	\item The Poisson distribution can cope with small counts and lots of zeros
\end{itemize}
\end{block}

\end{frame}

\begin{frame}
\frametitle{Cancer model}
\begin{align*}
Y_i \sim & \text{Poisson}(\lambda_i E_i)\\
\log(\lambda_{i}) = &   P_i \beta + \rho_i \gamma + U_i\\
 U_i \sim & \text{BYM}(\tau_1, \tau_2)
\end{align*}
\begin{itemize}
	\item The cancer relative risk surface $\lambda_i$  is unknown.  
	\item It depends on pollution $P_i$, smoking $\rho_i$ and residual spatial variation $U_i$
	\item $U_i$ is a spatially-smooth Besag, York and Mollie process which depends on it's neighbours
\item Similar models used by Haining et al (2007) and Best et al.
\end{itemize}

\end{frame}

\begin{frame}
\hspace*{-1cm}\input{graph.tex}
\end{frame}

\begin{frame}
\frametitle{Model fitting}
\begin{itemize}
	\item The data are not Normally-distributed, and there are  unobserved random effects.
	\item Uncertainty in smoking should carry over into the cancer model
	\item This makes the usual least squares or maximum likelihood methods problematic
	\item Bayesian inference using Markov Chain Monte Carlo simulation
	\item WinBUGS, with help from the R package R2WinBUGS and glmmBUGS
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Bayesian inference in a nutshell}
\begin{itemize}
	\item The models give the distributions

\begin{itemize}
	\item Cancer: $[Y | P, \rho, U, \beta, \gamma]$
	\item Smoking: $[Z  | X, W, \alpha]$
	\item Spatial surfaces $[U|\tau]$, $[W|\sigma]$
	\item Priors $[\gamma]$,$[\beta]$, $[\alpha]$ $[\sigma]$
$[\tau]$
\end{itemize}
\item Given the data $Y$ and $Z$, what are the distributions of the surfaces
\begin{itemize}
	\item Smoking: $[W|Y,Z]$
	\item Cancer: $[U|Y,Z]$
\end{itemize}
\item MCMC draws a number of equally likely values from these distributions
\item The average of these values is our best guess at the risk surfaces
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Adjusted risk surfaces}
\begin{enumerate}
	\item Cancer relative risk surface:
	\[\lambda_{i} =   \exp(P_i \beta) \exp( \rho_i \gamma)\exp( U_i)\]
	a value of $\lambda_4=1.3$ means an individual living in region 4 has a 30\% higher risk than the Ontario average for their age group 
	\item Remove the effect of smoking:
	\[\lambda_{i}/ \exp( \rho_i \gamma) = \exp(P_i \beta) \exp( U_i)\]
	A 1.3 means that compared to a typical region in Ontario with the same smoking rate region 4 but average pollution, risk is 30\% higher.

	\item Remove the effect of smoking and pollution:
	\[\lambda_{i}/ \exp( \rho_i \gamma)\exp(P_i \beta) =  \exp( U_i)\]
	Compared to a typical region the same smoking rate and pollution as region 4, risk is 30\% higher.
\end{enumerate}
\end{frame}

\begin{frame}

\frametitle{Cancer risk}
% left bottom right top

\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{Female-CancerExpected.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{Female-CancerExpected.pdf}
\end{frame}

\begin{frame} 
\frametitle{Cancer risk, removing smoking effect}
\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{FemalesCancerRandomEffectPolllution.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{FemalesCancerRandomEffectPolllution.pdf}
\end{frame}

\begin{frame}
\frametitle{Cancer risk, removing smoking and pollution}
\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{FemalesCancerRandomEffect.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{FemalesCancerRandomEffect.pdf}
\end{frame}

\begin{frame}
\frametitle{Parameter estimates}

\end{frame}

\begin{frame}
\frametitle{Chain convergence}

\end{frame}

\begin{frame}
\frametitle{Limitations and Extensions}
\begin{description}
	\item[Continuous space]:  $\lambda(s)$ instead of $\lambda_i$
\begin{itemize}
	\item The real risk surface is continuous
	\item We observed spatially aggregated counts.
\end{itemize}
	\item[Changes in time]: $\lambda(s,t)$

\begin{itemize}
	\item To gain more power, use data from a long time period
	\item Allow for possible changes in risk over time
	\item Changing census boundaries are a problem!
\end{itemize}
\item[]
\end{description}
\end{frame}

\begin{frame}
\frametitle{Extensions continued}
\begin{description}
\item[Better smoking model]: make better use of survey data
\begin{itemize}
	\item Individuals are either current, former, or never smokers
	\item Rates at which starting and quitting happens vary with age
	\item We have cross-sectional, not longitudinal, information
	\item Multi-state model with time varying rates
	\item Spatially-varying start rate
\end{itemize}

\item[Uncertainty in pollution]:
\begin{itemize}
	\item Monitoring data are available
	\item pollution is another indirectly observed covariate?
\end{itemize}
\end{description}
\end{frame}

\begin{frame}
\frametitle{Conclusion}

\begin{itemize}
	\item Small area disease mapping in space and time is possible even when counts are low 
	\item Include covariates in the model to estimate the effect of potential risk factors
	\item It's possible to include covariates which are indirectly or imprecisely observed.
	\item So always adjust for smoking when mapping diseases?
\end{itemize}
\end{frame}

\end{document}

### 2009-09-11 docs/flowchart/flowchart.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@387 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{pdfsync}

\title{Flow chart for spatial analysis}
\author{Patrick, Todd, \ldots}

\begin{document}
\maketitle
\section{Introduction}

Question: What is the spatial distribution of cancer X in region Y?  This question could be interpreted in the following ways, some or all of which might be of interest:
\begin{itemize}
	\item It is known that income and the age-sex distribution of the population affects cancer rates.  Is there evidence further spatial variation in risk beyond what population and income would predict?  If so, can we predict and map it?
	\item What is the spatial distribution of cancer risk in the region?  How much of this is explained by income?
	\item What is the expected future incidence of cancer in each region, in absolute numbers and per person, incorporating (rather than adjusting for) the age distribution of each region?
\end{itemize}

This paper describes a step-by-step method for addressing the above questions.  A  statistical model for spatially varying disease incidence is presented, steps for fitting this model to disease data are explained, and results are derived to address each of the questions above.

\section{Methods}

\subsection{A flow chart}

The steps below will be followed:
\begin{enumerate}
	\item Estimating cancer rates in the reference population as a function of age, for each sex and income quintile
	\item Testing for spatial randomness: do age, sex and income explain the distribution of cancer or is there evidence of residual spatial variation? 
	\item If yes, a spatial model is fit to produce:
\begin{enumerate}
	\item A map of log relative risk by region, after subtracting the effect of income.  Answers the question of ``Does my region have cancer risk higher (or lower) than would be expected given it's income and the age-sex structure of the population?
	\item The absolute risk for a reference age-sex group, including the effect of income.
	\item The age-sex standardized incidence rate, based on the Canada 1991  population.
	\item The incidence rate per inhabitant, irrespective of age and sex.
	\item The expected number of future cases.
	\end{enumerate}
\end{enumerate}

This is accomplished by 
\begin{enumerate}
	\item Using the RIF
	\item Moran or Tango with parametric bootstrap
	\item Use the BYM results to compute
\begin{enumerate}
	\item $E(U_i+V_i + \alpha | Y)$
	\item $E(\exp(U_i+V_i + \alpha) \theta_{M,65}(X_i) | Y)$
	\item $E(\exp(U_i+V_i + \alpha) \sum_j P_j \theta_j(X_i) | Y)$
	\item $E(\exp(U_i+V_i + \alpha) \sum_j P_{ij} \theta_j(X_i)/P_i | Y)$
	\item $E(\exp(U_i+V_i + \alpha) \sum_j P_{ij} \theta_j(X_i)| Y)$
\end{enumerate}
\end{enumerate}

\section{Results}

\subsection{Rates}

show one cancer which depends on income, another which does not.

\subsection{Testing for spatial variation}

Show p-values from several cancers, pick one which shows significant variation

\subsection{Mapping risk surfaces}

show (some of) the five maps

\section{Discussion}

Say how the maps are useful

Note that we've only fit one model, and derived all the maps from it.  This is preferable than running the BYM model without using income to calculate expecteds as income is not spatially smooth and can also cause the data to be non-normal.  Running the model without income should be done if income is probably not a confounder but is correlated with an exposure.

\end{document}

### 2009-09-24 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@362 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@408 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-\documentclass{beamer}
+\documentclass[xcolor=dvipsnames]{beamer}
 \usepackage[nogin]{Sweave}
+
+\usetheme{Singapore}
+\usecolortheme[named=RawSienna]{structure}
+
+\usefonttheme{serif}
+\setbeamertemplate{navigation symbols}{}
+\setbeamertemplate{footline}[frame number]
+
 \title{Bayesian statistics and  swine flu}
@@
 \SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}
+\newcommand{\comment}[1]{}
 
@@
 
+\comment{
 <<setup,fig=false>>=
-options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)), width=58)
+options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)), width=54)
 # source in all the R files
@@
 @
+}
 
@@
 \frametitle{Swine Flu in Ontario}
+\begin{block}{The Flu}
 \begin{itemize}
@@
 
-\begin{itemize}
 	\item Swine flu was probably established in Ontario before the outbreak was detected.
 \item Few control measures were adopted
+
 \end{itemize}
-\item Ontario's had recent public health crises
+\end{block}
+
+\begin{block}{The Science}
 \begin{itemize}
-	\item SARS, Walkerton water quality problems
-	\item Ontario Agency for Health Protection and Promotion created
 	\item Expert opinion abounds from mathematical modellers and infectious disease epidemiologists
-	\item Detailed models are used for the biological process, with largely ad-hoc inference and treatment of uncertainty. 
+	\item For the most part, very elaborate models are combined with largely ad-hoc inference and treatment of uncertainty
+	\item Prior knowledge + limited data = Bayesian statistics?
 \end{itemize}
-\end{itemize}
+
+\end{block}
 
@@
 \frametitle{The Data}
+\begin{block}{Early stages}
 \begin{itemize}
@@
 
-\begin{itemize}
 	\item Ill individuals were encouraged to seek medical help
 	\item Practitioners were encouraged to send samples to labs for testing
-\end{itemize}
-	\item After 6 weeks, testing was discouraged to ease the burden on public health labs.
 	\item Roughly 1000 cases in the first 6 weeks? 
 \end{itemize}
+\end{block}
+\begin{block}{Later stages}
+\begin{itemize}
+	\item After 6 weeks, testing was discouraged to ease the burden on public health labs.
+\item Data is more extensive but less complete
+\end{itemize}
+\end{block}
 \end{frame}
@@
 	\item Crude rates underestimate the rates, and result in estimates increasing over time
+\item In early stages, data are sparse
+\item If data are divided into age groups, the problem is worse.
 \end{itemize}
@@
 	\item[Serious:] The infection is severe enough that hospitalization will be required before the individual recovers
-	\item[Deadly:] Following hospitalization, the patient dies of the infection.
+	\item[Deadly:] Following hospitalization, the patient will die of the infection.
 \end{description}
-When a case is first identified, it's not known what type it is.
+Until a case recovers or dies, it's not known what type it is.
 
@@
 \begin{frame}
-\frametitle{Distributions}
+\frametitle{Event time distributions}
 \begin{itemize}
@@
 pr(T=t|Z=0) =&  \text{pr}(\max(t-0.5,0) < y < t+0.5)\\
-y \sim & \text{Weibull}(\mu/\Gamma(1 + 1/\nu), \nu)
+y \sim & \text{Weibull}[\mu/\Gamma(1 + 1/\nu), \nu]
 \end{align*}
@@
 \begin{frame}
-\frametitle{Priors}
+\frametitle{Infection type probabilities}
+
+\begin{itemize}
+	\item For individual $i$,
+	\item $W_i \in \{M,S,D\}$ is infection type
+	\item $X_i$ is their age
+\end{itemize}
+\begin{block}{Fatality rate}
+\[
+Pr(W_i = D) = g_D(X_i)
+\]
+\end{block}
+
+\begin{block}{(non-fatal) Hospitalization rate}
+\[
+Pr(W_i = S| W_i \neq D) = g_S(X_i)
+\]
+\end{block}
+\begin{itemize}
+\item Parameterizing the probabilities this way ensures they sum to 1.
+	\item $g_D$ and $g_S$ are functions which vary smoothly with age
+	\item More specifically, log-linear cubic splines with fixed knots and random coefficients 
+\item As implemented in the DPpackage in R
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Parameters (for simulating data)}
+<<echo=false,fig=false>>=
+options(width=45)
+ageSeq = 0:100
+
+params = pandemicParams( 
+	InfOns = c(mean=1.5, shape=2, zeros=.1),
+	OnsMedM= c(mean=1, shape=1, zeros=.1 ),
+	OnsMedS= c(mean=1, shape=1, zeros=.1 ),
+	OnsMedD= c(mean=1, shape=1, zeros=.1 ),
+	MedRec = c(mean=3, shape=2, zeros=.1, lost=0.95 ),
+  MedHospS = c(mean = 2, shape = .5, zeros = 0.1),
+  MedHospD = c(mean = 2, shape = 2, zeros = 0.1),
+	HospRec = c(mean=3, shape=1.5, zeros=.01 ),
+	HospDeath = c(mean=5, shape=0.8, zeros=.05 ),
+ageProbs=list(
+	S=addAgeProbs(ageSeq, 0.02+0.8*dnorm(ageSeq, mean=20, sd=10)),
+	D=addAgeProbs(ageSeq, 0.01+0.5*dnorm(ageSeq, mean=30, sd=5))  ),
+	probs=NULL
+	)
+tempList = (params[c("InfOns","OnsMedM","OnsMedS", "OnsMedD", "MedRec","MedHospS","MedHospD","HospRec","HospDeath")])
+temp = unlist(tempList)
+temp = temp [-grep("lost",names(temp))]
+tempMat = matrix(temp, ncol=4, byrow=T)
+rownames(tempMat) = names(tempList)
+colnames(tempMat) = names(tempList[[1]])
+paramMat = tempMat[,c("mean","shape","zeros")]
+library(Hmisc)
+@
+
+<<parameters,echo=false,fig=false,results=tex>>=
+latex(paramMat, file="")
+@
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Hospitalization and Death probabilities}
+
+\begin{columns}
+
+\column{0.5\textwidth}
+\begin{block}{Hospitalization}
+<<probS>>=
+plot(params$age$S, type="l", ylim=c(0,0.06))
+@
+\end{block}
+
+\column{0.5\textwidth}
+\begin{block}{Death}
+<<probD>>=
+plot(params$age$D, type="l", ylim=c(0,0.06))
+@
+
+\end{block}
+
+\end{columns}
+
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{Simulated data}
+<<echo=false,fig=false>>=
+options(width=60)
+@
+<<simData,echo=true,fig=false>>=
+
+data = simEpidemic(params, delta=20, days=28)
+
+table(data$observedType)
+@
+<<fig=false>>=
+toshow = NULL
+for(Dshow in levels(data$observedType))
+	toshow = c(toshow, which(data$observedType==Dshow)[1:2])
+toshow = na.omit(toshow)	
+@
+<<simData2,echo=true,fig=false>>=
+data[toshow,-c(1,7)]
+@
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Prior Distributions}
 \begin{itemize}
@@
 \end{itemize}
-\item Gamma or Beta priors, parametrized by mean and standard deviation
+\item Gamma priors for Weibull  mean and shape parameters
+\item Beta priors for zero-inflation parameters
+\item parametrized by mean and standard deviation
 \item Average time to onset of symptoms is expected to be 3.5 days, but the average time could be a day longer or shorter.
@@
 \begin{frame}
-\frametitle{Proportion parameters}
-
+\frametitle{Priors for infection types}
 \begin{itemize}
-	\item $\rho_D$ is the fatality ratio, or probability of a deadly infection
-	\item $\rho_S$ is the proportion of non-fatalities which are hospitalized
-	\item so the probability of being hospitalized is $\rho_D + (1-\rho_D)\rho_S$
-	\item parametrize this way to ensure the probabilities of being Mild, Severe, or Deadly add to 1.
-	\item beta priors on $\rho_D$ and $\rho_S$, parametrized by mean and standard deviation
-	\item i.e.\ $\rho_D$ is expected to be $0.1$, with $sd(\rho_D) = 0.05$
+	\item Specify prior mean for the probability an individual in some reference group is
+\begin{itemize}
+	\item Serious
+	\item Deadly
+\end{itemize}
+	and give upper 95\% prior quantile for each probability
+\begin{itemize}
+	\item i.e.\ for hospitalization rate, 0.001 and 0.005
+\end{itemize}
+	\item and Gamma shape and scale parameters for spline coefficients
+	\item i.e.\ shape=2, scale=10 would give fairly flat curves
 \end{itemize}
@@
 @
-
+\vspace*{-0.8cm}
+Density and hazard curves are shown for combinations of the 2.5\% and 97.5\% quantiles of each parameter's prior distribution.
 \end{frame}
@@
 \begin{frame}
-\frametitle{Inference}
+\frametitle{Bayesian Inference}
 \begin{itemize}
@@
 	\item Posteriors show what we think the parameters could be, given the data and the prior distributions
-	\item There's no data on infection times, so the posterior for $\mu_{I}$ is the same as the prior.
+	\item If there's no data (i.e.\ infection times), the posterior is the same as the prior.
+	\item More data will pull the posterior away from the prior to be ``more like'' the data.
 \end{itemize}
@@
 \begin{frame}
-\frametitle{The Algorithm}
+\frametitle{Updating infection type probabilities}
+
+\begin{itemize}
+\item If we knew what would happen in the future, the problem would be easy
+\begin{itemize}
+	\item Fatality ratio = fatalities/cases
+	\item GAM or logistic regression with fatality v age
+\end{itemize}
+\item So, use the model to simulate a future.
+\end{itemize}
+
 \begin{block}{Data augmentation}
 \begin{itemize}
-	\item Fix the parameter values
-	\item Simulate the missing data: infection type, times of infection, onset, hospitalization, recovery, and/or death.
-	\item Can simulate directly, as conditioning is easy.
+\item Choose some starting values for the probabilities
+\item Use these probabilities to simulate infection types
+\begin{itemize}
+	\item Suppose person $i$ was hospitalized 3 days ago
+	\item compute $p_i = pr(W_i =D | T_i > 3;X_i)$ 
+	\item simulate $(W_i = D) \sim \text{Bern}(p_i)$
+\end{itemize}
+\item Use these simulated data to get better parameter values (which is easy because the data are complete)
+\begin{itemize}
+	\item $[g_D(x) | W_1 \ldots W_N]$
+	\item using the PSgam function in the DPpackage
+\end{itemize}
+\item Repeat
 \end{itemize}
 \end{block}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Event time parameters}
 
@@
 	\item With the full augmented data, the likelihood is simple as observations are independent.
+	\item We have a event times, and some censored event times
+	\item and because of the augmented data, we know which distribution each of these times come from
 	\item Propose new parameters with a Gaussian random walk, accept with the appropriate probability.
@@
 
-\begin{frame}[fragile]
-\frametitle{Simulated data}
-<<simData,echo=true,fig=false>>=
-params = pandemicParams( 
-	InfOns = c(mean=1.5, shape=2, zeros=.1),
-	OnsMedM= c(mean=1, shape=1, zeros=.1 ),
-	OnsMedS= c(mean=1, shape=1, zeros=.1 ),
-	OnsMedD= c(mean=1, shape=1, zeros=.1 ),
-	MedRec = c(mean=3, shape=2, zeros=.1, lost=0.95 ),
-  MedHospS = c(mean = 2, shape = 2, zeros = 0.1),
-  MedHospD = c(mean = 2, shape = 2, zeros = 0.1),
-	HospRec = c(mean=2, shape=1.5, zeros=.01 ),
-	HospDeath = c(mean=2, shape=0.8, zeros=.05 ),
- 	probs = c(M=0.7, S=0.25, D=0.05)
-# 	probs = c(M=0.979, S=0.02, D=0.001)
-  )
-
-data = simEpidemic(params, delta=4, days=20)
-
-data[1:8,-1]
-@
-
-\end{frame}
+\end{document}
 
@@
 
-postSummary[1:12,-4]
+postSummary[1:12,-c(1,4)]
 @

### 2009-09-24 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@409 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\begin{frame}[fragile]
+\frametitle{Infection type priors}
+
+\begin{verbatim}
+PS Priors
+transition	distribution	taub1	taub2	priorMean	upper95
+probs	fatality	psPrior	   2	   0.05	  0.01	   0.02
+probs	hosp    	psPrior	   2	   0.05	  0.01	   0.02
+\end{verbatim}
+
+\end{frame}
+
 \begin{frame}
@@
 <<priorGraph, fig=false,results=tex>>=
-priors= readPrior("bayesPandemicPriors.txt")
+priors = readPrior("bayesPandemicPriors.txt")
 plotPrior(priors$InfOns, file="InfOns", tex=TRUE)

### 2009-09-24 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@411 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 scale = mcmcScale(params, 0.075,minScale = 0.05 )
-paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=60)
-paramSample = paramSample[seq(100,dim(paramSample)[1], by=5),]
+paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=60, thin = 5)
+
 dim(paramSample)

### 2009-09-25 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@412 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<setup,fig=false>>=
-options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)), width=54)
+options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)), width=54, digits=3)
 # source in all the R files
@@
 	HospRec = c(mean=3, shape=1.5, zeros=.01 ),
-	HospDeath = c(mean=5, shape=0.8, zeros=.05 ),
+	HospDeath = c(mean=5, shape=0.3, zeros=.05 ),
 ageProbs=list(
-	S=addAgeProbs(ageSeq, 0.02+0.8*dnorm(ageSeq, mean=20, sd=10)),
-	D=addAgeProbs(ageSeq, 0.01+0.5*dnorm(ageSeq, mean=30, sd=5))  ),
+	S=addAgeProbs(ageSeq, 0.02+1.5*dnorm(ageSeq, mean=20, sd=10)),
+	D=addAgeProbs(ageSeq, 0.01+dnorm(ageSeq, mean=30, sd=8))  ),
 	probs=NULL
@@
 
-data = simEpidemic(params, delta=20, days=28)
+data = simEpidemic(params, delta=50, days=42)
 
@@
 toshow = na.omit(toshow)	
+#$
 @
@@
 
-\end{document}
-
 \begin{frame}[fragile]
-\frametitle{Results}
+\frametitle{Fitting the model}
+(to the simulated data)
 <<mcmcResults,echo=true,fig=false>>=
 scale = mcmcScale(params, 0.075,minScale = 0.05 )
-paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=60, thin = 5)
+paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=100, thin = 5)
 
@@
 \begin{frame}
-\frametitle{Fatality probability}
+\frametitle{Fatality probability, age 20}
 \begin{columns}
@@
 <<tracePlot,fig=true>>=
-plot(paramSample[,"probs.D"], type="l")
-abline(h=params$probs["D"], col="red")
+toplot = "ageProbs.D.prob20"
+plot(paramSample[,toplot], type="l")
+abline(h=params$ageProbs$D[20,"prob"], col="red")
 @
@@
 <<histPost>>=
-hist(paramSample[,"probs.D"], breaks=12)
-abline(v=params$probs["D"], col="red")
+hist(paramSample[,toplot], breaks=12)
+abline(v=params$ageProbs$D[20,"prob"], col="red")
 @
@@
 
-postSummary[1:12,-c(1,4)]
+postSummary[grep("^(OnsMed|MedHosp)(M|S|D).(mean|shape)$",#$
+
+rownames(postSummary)),-c(1,4)]
 @
@@
 
+\begin{frame}
+\frametitle{Fatality rate}
+<<fatalityAgeGraph>>=
+ageCols = grep("^ageProbs.D.prob.[[:digit:]]+$", colnames(paramSample), value=T)
+Sage= as.numeric(gsub("^ageProbs.D.prob", "", ageCols))
+
+matplot(Sage, t(paramSample[,ageCols]), type="l", col=2:8, ylim=quantile(paramSample[,ageCols], probs=c(0.01, 0.99))) 
+lines(Sage, apply(paramSample[,ageCols], 2, mean), lwd=3, col="black")
+#$
+@
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Hospitalization rate}
+<<fatalityAgeGraph>>=
+ageColsS = grep("^ageProbs.S.prob.[[:digit:]]+$", colnames(paramSample), value=T)
+
+theRate = paramSample[,ageCols] + (1-paramSample[,ageCols])*paramSample[,ageColsS]
+
+matplot(Sage, t(theRate), type="l", col=2:8, ylim=quantile(theRate, probs=c(0.01, 0.99))) 
+lines(Sage, apply(theRate, 2, mean), lwd=3, col="black")
+#$
+@
+
+\end{frame}
+
+\end{document}
+
 \begin{frame}[fragile]
@@
 # generate hospital times
-hospSample = simHospitals(paramSample , Ndays=20, NinfectionsPerDay=20)
+hospSample = simHospitals(paramSample , Ndays=60, 
+	NinfectionsPerDay=c(seq(10, 100, len=30),seq(100, 10, len=30) ) )
 

### 2009-09-25 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@413 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 # generate hospital times
-hospSample = simHospitals(paramSample , Ndays=60, 
-	NinfectionsPerDay=c(seq(10, 100, len=30),seq(100, 10, len=30) ) )
+hospSample = simHospitals(paramSample , Ndays=80, 
+	NinfectionsPerDay=c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) ) )
 

### 2009-09-25 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@414 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2009-09-25 docs/pandemic/bayesPandemic.rnw
@@
 <<probS>>=
-plot(params$age$S, type="l", ylim=c(0,0.06))
+plot(params$age$S, type="l", ylim=c(0,0.1))
 @
@@
 <<probD>>=
-plot(params$age$D, type="l", ylim=c(0,0.06))
+plot(params$age$D, type="l", ylim=c(0,0.1))
 @
@@
 \begin{frame}
-\frametitle{Posterior and prior, Hosp $\Rightarrow$ Death }
+\frametitle{Posterior and prior, Med $\Rightarrow$ Hosp }
 <<priorPost, fig=false,results=tex>>=
@@
 
-\end{document}
-
 \begin{frame}[fragile]
@@
 # generate hospital times
-hospSample = simHospitals(paramSample , Ndays=60, 
-	NinfectionsPerDay=c(seq(10, 100, len=30),seq(100, 10, len=30) ) )
+hospSample = simHospitals(paramSample , Ndays=80, 
+	NinfectionsPerDay=20 )
 
@@
 # generate hospital times
-hospSample2 = simHospitals(paramSample , Ndays=40, 
-	NinfectionsPerDay=c(seq(from=1,by=2, len=10), seq(from=20, len=10, by=2), rep(2, 20)))
+hospSample2 = simHospitals(paramSample , Ndays=80, 
+	NinfectionsPerDay=c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) ) )
 

### 2009-09-25 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@415 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 plotHospitals(hospSample)
+abline(h=20)
 @
@@
 \frametitle{With rising and falling incidence}
-{\scrpitsize
+{\scriptsize
 <<hospLoad2,echo=T,height=2,width=4>>=
 # generate hospital times
-hospSample2 = simHospitals(paramSample , Ndays=80, 
-	NinfectionsPerDay=c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) ) )
+Ninfect=	c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) ) )
+
+hospSample2 = simHospitals(paramSample , Ndays=80, NinfectionsPerDay=Ninfect)
 
@@
 plotHospitals(hospSample2)
+lines(Ninfect)
 @
 }
-\end{frame}
-
-\begin{frame}
-\frametitle{Age?}
-\begin{itemize}
-	\item Assume the fatality and hospitalization probabilities vary with age
-	\item After data augmentation, we have a binary response with a smoothly-varying probability
-	\item Use a canned Bayesian semi-parametric routine to generate a new age curve at every iteration.
-	\item {\tt DPpackage}?
-	\item Informative prior on the average or rate at a baseline age
-	\item Strong prior that the curve is flat.
-\end{itemize}
 \end{frame}

### 2009-09-25 docs/pandemic/bayesPandemic.rnw
pandemicresults
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@416 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 scale = mcmcScale(params, 0.075,minScale = 0.05 )
-paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=100, thin = 5)
+paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=500, thin = 10)
 
@@
 # generate hospital times
-Ninfect=	c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) ) )
+Ninfect=c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) )
 

### 2009-09-25 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@417 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 library(pandemic)
+doSims=F
 @
@@
 options(width=60)
+if(doSims){
+data = simEpidemic(params, delta=50, days=42)
+} else{
+load("results.RData")
+}
 @
+\verb!data = simEpidemic(params, delta=50, days=42)!
 <<simData,echo=true,fig=false>>=
-
-data = simEpidemic(params, delta=50, days=42)
-
 table(data$observedType)
@@
 scale = mcmcScale(params, 0.075,minScale = 0.05 )
+if(doSims){
 	paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=500, thin = 10)
-
+}
 dim(paramSample)
@@
 \frametitle{Hospitalization rate}
-<<fatalityAgeGraph>>=
+<<hospitalAgeGraph>>=
 ageColsS = grep("^ageProbs.S.prob.[[:digit:]]+$", colnames(paramSample), value=T)
@@
 # generate hospital times
+if(doSims) {
 hospSample = simHospitals(paramSample , Ndays=80, 
 	NinfectionsPerDay=20 )
+}
 
@@
 plotHospitals(hospSample)
-abline(h=20)
+abline(h=20/5)
 @
@@
 plotHospitals(hospSample2)
-lines(Ninfect)
+lines(Ninfect/5)
 @

### 2009-10-13 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@423 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 dim(paramSample)
-colnames(paramSample)
+colnames(paramSample)[c(1, 100, 185, 197)]
 @
@@
 \end{columns}
+True value is in red
 \end{frame}
@@
 
+\begin{frame}
+\frametitle{Predicting hospital loads}
+
+\begin{itemize}
+	\item For a given proposed outbreak
+
+\begin{itemize}
+	\item Either a hypothetical scenario
+	\item or projections from an infectious disease model
+\end{itemize}
+\item What will the hospital load be?
+\item \ldots depends on the parameters
+\item So, use each of the 1000 posterior samples to simulate 1000 hospital loads
+\item and graph the mean and quantiles for hospital demand
+\end{itemize}
+
+\end{frame}
+
 \begin{frame}[fragile]
 \frametitle{Hospital demand}
-<<hospLoad,echo=T>>=
-# generate hospital times
+<<fig=false>>=
 if(doSims) {
@@
 	NinfectionsPerDay=20 )
+Ninfect=c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) ) 
+hospSample2 = simHospitals(paramSample , Ndays=80, NinfectionsPerDay=Ninfect)
 }
+@
+\verb!# generate hospital times!
+\verb!hospSample = simHospitals(paramSample , Ndays=80, !\\
+\verb!	NinfectionsPerDay=20 )!
+
+\begin{minipage}{0.8\textwidth}
+<<hospLoad,echo=T>>=
 
@@
 @
+
+\end{minipage}
 \end{frame}
@@
 \frametitle{With rising and falling incidence}
-{\scriptsize
+\verb!hospSample2 = simHospitals(paramSample , Ndays=80,! \\
+\verb!  NinfectionsPerDay=Ninfect)!
+
+\begin{minipage}{0.8\textwidth}
 <<hospLoad2,echo=T,height=2,width=4>>=
-# generate hospital times
-Ninfect=c(seq(1, 100, len=20),seq(100, 1, len=20), rep(2, 40) )
+Ninfect=c(seq(1, 100, len=20),
+  seq(100, 1, len=20), rep(2, 40) )
 
-hospSample2 = simHospitals(paramSample , Ndays=80, NinfectionsPerDay=Ninfect)
-
-# graph hospital times
 plotHospitals(hospSample2)
@@
 @
-}
+\end{minipage}
 \end{frame}
@@
 \begin{itemize}
-	\item more covariates	
+\item fine-tune the MCMC
+\begin{itemize}
+	\item check convergence
+	\item optimize scaling
+	\item look for more conjugacy?
+\end{itemize}
+	\item more covariates	--- prior health conditions?
 	\item model the infection process
 	\item spatial infection process
+\item use the real data?	
 \end{itemize}
@@
 \begin{frame}
+\frametitle{Conclusions 1 (specific)}
+\begin{itemize}
+	\item This model can
 
-\frametitle{Conclusions}
+\begin{enumerate}
+	\item Estimate fatality and hospitalization rates, allowing for censoring
+	\item Take infection trajectories from an infectious disease model and translate them into hospital loads
+\end{enumerate}
+\item If this is what you want, this model is much lighter than a full-blown infectious disease model
+\item and will require much less prior information.
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+
+\frametitle{Conclusions (general)}
 
 \begin{itemize}
-	\item Use the data even if there isn't much of it
+	\item Use the data for estimating parameters even if there isn't much of it
 	\item If there are many cases and no fatalities, at least we can get an upper bound on the fatality rate

### 2009-10-23 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@431 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \documentclass[xcolor=dvipsnames]{beamer}
+
 \usepackage[nogin]{Sweave}
+\usepackage{tikz,pgflibraryshapes}
 
@@
 
-\title{Bayesian statistics and  swine flu}
+\title{Applications of Time-Inhomogeneous Multi-State Models (including Bayesian statistics and swine flu)}
 \author{Patrick Brown$^1$,  Peter Neal$^2$}
@@
 
+\begin{frame}
+\frametitle{Multi-state models}
+\begin{itemize}
+	\item The Swin flu application was \emph{ordinal}
+	\item Once an individual makes a transtion they can't go back
+	\item but what if they can go back?
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Endemic diseases in Tanzanian cows}
+\begin{columns}
+\column{0.5\textwidth}
+\includegraphics{Figures/G-cowTS}
+\column{0.5\textwidth}
+\begin{itemize}
+	\item Sampled up to 11 different occasions
+	\item Approximately 6 weeks apart
+	\item 60 farms, 100 animals
+\end{itemize}
+\end{columns}
+\end{frame}
+
+\begin{frame}
+\frametitle{Questions}
+\begin{itemize}
+	\item How long does an episode last?
+	\item How does age affect infection probabilities?
+	\item Do cows located close together have similar infection probabilities?
+
+\begin{itemize}
+	\item Transmission between animals?
+ \item A spatially varying risk factor (density of ticks)?
+\end{itemize}
+\item Note: We don't observe infection and recovery times.
+\item We only know infection status on certain days.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{A Susceptible-Infected-Susceptible model}
+\begin{tikzpicture}[fill=blue!20,line width=2pt,shorten >=5pt]
+\tikzstyle{every node}=[draw,fill,ellipse,minimum height=2cm,minimum width=3cm] 
+\path (0,0) node(S) {Susceptible}
+	 		(5,0) node(I)  {Infected};
+\draw[blue,->](S) .. controls +(2,2) and +(-2,2)  ..(I);
+\draw[blue,->](I) .. controls +(-2,-2) and +(2,-2)  ..  (S);
+\end{tikzpicture}
+\begin{itemize}
+	\item Every animal alternates between two states
+	\item Infections happen at rate $\delta$
+	\item Recoveries have rate $\rho$
+\end{itemize}
+\end{frame}
+\begin{frame}
+
+\frametitle{Infection probability changes with age}
+\begin{tikzpicture}[fill=white,font=\sffamily]
+
+\begin{scope}[line width=2pt,->]
+\draw(0,0) -- (0,5) node[pos=0.5,sloped,above=7pt]{$\delta$(age)};
+\draw(0,0) -- (8,0) node[pos=0.5,sloped,below=7pt]{age};
+\end{scope}
+
+\draw (0.5,4) -- (2.5,1) --  (8,1);
+
+\begin{scope}[dotted,line width=0.5pt]
+\draw (2.5,0) -- (2.5, 1) node[below,pos=0]{$\tau_0$};
+\end{scope}
+
+\end{tikzpicture}
+\begin{itemize}
+	\item Recovery rate $\rho$ is constant
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{The Model}
+Animal $i$ located at $s_i$ has infection rate
+\[
+\delta_{i}(t) =\begin{cases} \exp(\delta_0 + X_{i}' \beta + w_i \gamma + A(s_i)) & w_i(t) < \tau_0\\
+\exp(\delta_0 + X_{i}' \beta + w_i \tau_0 +A(s_i)) & w_i(t) \geq \tau_0 \end{cases} 
+\]
+\begin{itemize}
+	\item $X_{i}$ is indicator variables denoting if the animal grazes, is sprayed, or eats grass
+	\item $w_i(t)$ is the age of animal $i$ at time $t$.
+	\item $A(\cdot)$ is a Gaussian random field with covariance function 
+	\[
+	\text{cov}(A(s_0), A(s_0+u)) = \sigma^2 \exp(-||u||/\phi)
+	\]
+	\item Recovery rate is $\rho$
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Inference with constant rates}
+\begin{itemize}
+	\item Alternating states model, see Grimmett and Stirzaker.
+	\item Transition model.
+	\item To construct the likelihood, we need the transition probability matrix:
+	\[
+	P_u = \left(\begin{array}{ll} Pr(Y_{t+u} = 0 | Y_t = 0) &  Pr(Y_{t+u} = 1 | Y_t = 0)\\
+ Pr(Y_{t+u} = 0 | Y_t = 1) &  Pr(Y_{t+u} = 1 | Y_t = 1) \end{array}\right) 
+	\]
+	\item Cow $i$ has infection rate $\delta_i$ and recovery rate $\rho$ 
+	\begin{align*}
+	G_i = &\left(\begin{array}{cc} -\delta_i & \delta_i \\ \rho & -\rho  \end{array}\right) \Leftarrow \text{Generating Matrix}
+\\
+P_u = &\exp(u G_i)
+\end{align*}
+\item This is a matrix exponential $\exp(X) = \sum_{n=1}^\infty X^n/n!$
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Probabilities}
+\begin{itemize}
+	\item Unconditional distribution
+	\[
+	 Pr(Y_{it} = 1) = \frac{\delta_i}{\delta_i + \rho} 
+	\]
+	\item Susceptible $\rightarrow$ infected
+	\[
+	  Pr(Y_{i,t+u} = 1 | Y_{it} = 0) = \frac{\delta_i}{\delta_i + \rho} \left( 1- \exp[-u(\delta_i + \rho)]\right)
+	\]
+	\item Infected $\rightarrow$ infected
+	\[
+	  Pr(Y_{i,t+u} = 1 | Y_{it} = 1) = \frac{\delta_i}{\delta_i + \rho} + 
+	  	\frac{\rho}{\delta_{i} + \rho} \exp[-u(\delta_{i} + \rho)]
+	  \]
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Notes}
+\begin{itemize}
+	\item The observations are Bernoulli random variables, with a peculiar formula for calculating the probabilities
+	\[
+	p_{it} = pr(Y_{it} = 1) = f(X_i, Y_{i,t-1})
+	\]
+	\item The likelihood function is a product of the probabilities (or 1 minus them).
+	\[
+	L(\delta_0, \rho, \beta ) =  \prod_{it : Y_{it}=1}p_{it} \prod_{it : Y_{it}=0}(1-p_{it} )
+	\]
+	\item  The Markov assumption is why this problem is tractable
+	\item Markov property: the future depends only on the present, not the past.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{More  Notes}
+\begin{itemize}
+\item Number of infections in an interval of length $u$ $\sim \text{Poisson}(u \delta)$
+\item Duration of a disease episode $\sim \text{Exponential}(1/\rho)$
+\item The Exponential-Poisson combination is a result of assuming the Markov property.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+
+\frametitle{Changing rates?}
+\begin{itemize}
+\item A surprisingly difficult problem
+	\item Peter Neal (University of Manchester) found the solution:
+\end{itemize}
+\begin{align*}
+G_i(t)= &\begin{pmatrix} -\exp(\delta_0 + \beta t) & \exp(\delta_0 + \beta t) \\ \rho &-\rho  \end{pmatrix}\\
+P(t, t+u) = & \exp\left( \int_0^{u}  G(t+x) dx\right)\\
+\end{align*}
+\pause
+Later, Peter says: \emph{The above expression is unfortunately wrong.  The problem being that matrix multiplication is non-commutative.  In other words $e^Ae^B \neq e^{A+B}$ unless $AB=BA$.  }
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Solutions}
+\begin{block}{Brute force}
+\begin{itemize}
+	\item Reversible jump MCMC, data augmentation
+	\item number of transitions is a latent variable, augmented
+	\item locations of those transitions is also augmented
+\end{itemize}
+\end{block}
+\begin{block}{Approximations}
+\begin{itemize}
+	\item Peter's found upper and lower bounds to the probabilities
+	\item When updating, check if proposal is above upper or below lower bound
+	\item If not, recalculate boundaries with greater precision
+\end{itemize}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Local Likelihood}
+with Gerald Lebovic
+
+\begin{itemize}
+\item Write $Y_{ij}$ as the state of individual $i$ at time $T_{ij}$.
+	\item To estimate $\delta(t)$, use $\{Y_{ij} |t-h< T_{ij} < t+h \}$
+	\item or some weighted kernel kernel
+	\item suppose data are unbalanced?
+	\item There may be no observations in the interval
+	\item but observations just outside could be informative
+	\item with a kernel, observations might be in places of low weight, but be informative about the (missing) data with high weight.
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Local EM}
+\begin{itemize}
+	\item Put the data on a fine grid, $\tilde Y_{ik}, \tilde T_{ik}$, $k=1 \ldots K$
+	\item Most of the $\tilde Y_{ik}$ are missing
+\item Start with an estimate $\hat\delta^{(0)}(t)$:
+\end{itemize}
+\begin{description}
+\item[E step]: compute $\text{E}[\tilde Y_{ik}|Y;\hat\delta^{(p)}(t)]$
+\item [M step]: use local likelihood to estimate $\hat\delta^{(p+1)}(t)$
+\end{description}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{What is $\tilde Y_{ik}$?}
+\begin{itemize}
+	\item If three states, $\tilde Y_{ik} = (p_{ik1}, p_{ik2}, p_{ik1})$, probability of being in each of the three states
+	\item No\ldots, not easy to calculate the likelihood from these data
+	\item $\tilde Y_{ik}$ is a 3x3 matrix:
+\[
+\tilde Y_{ik}(a,b) = pr(Y_{ik-1} = a, Y_{ik}=b)
+\] 
+\item If $Y_{ik}$ and $Y_{ik-1}$ are known, one ``1'' and zeros
+\item If one of $Y_{ik}$ and $Y_{ik-1}$ are known, one non-zero row or column
+\item then if $P_k$ is the transition matrix 
+\[
+\log[L(\tilde Y)] = \sum_{i,k} \tilde Y_{ik}' P \tilde Y_{ik}
+\]
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Another example: smoking}
+\includegraphics[width=2in]{Figures/G-ukDataCurrent.pdf}\includegraphics[width=2in]{Figures/G-ukDataFormer.pdf}
+
+Current smokers \hspace{1.5in} Never smoked
+\end{frame}
+
+\begin{frame}
+\frametitle{A simple model}
+\begin{tikzpicture}[fill=white,line width=2pt,shorten >=5pt,font=\sffamily\small]
+
+\begin{scope}
+\tikzstyle{every node}=[draw,fill,ellipse,minimum height=1.4cm,minimum width=1.5cm]
+
+% the nodes
+\node (N) at (0,0) {Never Smoked};
+\node (C)	at (4,0)  {Current Smoker};
+\node	(F) at (8,0)  {Former Smoker};
+
+\end{scope}
+
+% paths between the nodes
+\draw[black,->] (N) .. controls +(1.5,1.5) and +(-1.5,1.5)  .. node[below]{start} (C);
+\draw[black,->](C) .. controls +(1.5,1.5) and +(-1.5,1.5)  .. node[below]{quit}(F);
+
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+\frametitle{A more complicated model}
+
+\hspace*{-0.6in}\begin{tikzpicture}[fill=white,line width=2pt,shorten >=5pt,font=\sffamily]
+
+\begin{scope}
+\tikzstyle{every node}=[draw,fill,ellipse,minimum height=1.4cm,minimum width=1.5cm]
+
+% the nodes
+\node (N) at (0,0) {Never Smoked};
+\node (C)	at (5,0)  {Current Smoker};
+\node	(F) at (10,0)  {Former Smoker};
+\node	(D) at (5,3)  {Death};
+
+\end{scope}
+
+% paths between the nodes
+\draw[black,->] (N) .. controls +(1.5,1.5) and +(-1.5,1.5)  .. node[below]{start} (C);
+\draw[black,->](C) .. controls +(1.5,1.5) and +(-1.5,1.5)  .. node[below]{quit}(F);
+\draw[black,->](F) .. controls +(-1.5,-1.5) and +(1.5,-1.5)  .. node[above]{lapse}(C);
+\draw[black,->](F) .. controls +(-0.5,-2.5) and +(0.5,-2.5)  .. node[above]{reclassify}(N);
+\draw[black,->](N)..controls + (0, 1.5) .. (D);
+\draw[black,->](C)--(D);
+\draw[black,->](F).. controls + (0, 1.5) ..(D);
+
+\end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Conclusions}
+\begin{itemize}
+\item Transition models can model complex correlation structures with a small number of parameters
+\item Extensions to time-varying rates are useful but non-trivial
+\item Since these models make scientific sense, they're easier to build and interpret
+\item Possible extensions
+\begin{itemize}
+\item Weibull distribution for disease episode length?
+\item Non-linear rate functions?
+\item Likely to be impossibly complicated.
+\end{itemize}
+\pause
+\item Don't dismiss stochastic processes as unnecessarily complex and of little use!
+\end{itemize}
+
+\end{frame}
 \end{document}
-
-D75A0BD6A6A1

### 2012-10-05 docs/pandemic/pandemicPaper.tex
aded MRFraster
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@495 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}
\textwidth 6.5in \textheight 9.0in \topmargin -.5in \oddsidemargin
0.0in \evensidemargin 0.0in

\usepackage{graphicx}
\usepackage[nogin]{Sweave}
\usepackage{tikz,pgflibraryshapes}
\usepackage{cite}
\usepackage{subfigure}
\usepackage[authoryear]{natbib}
\bibliographystyle{newapa}

\title{\textbf{Time-Inhomogeneous Multi-State Model for a Swine Flu Epidemic}}
\date{\today}

       \usepackage{amsmath,subfigure,alltt}
\setkeys{Gin}{width=0.5\columnwidth}

\newcommand{\comment}[1]{}

\comment{
}

\begin{document}

\maketitle

\section{Introduction}
\label{sec:introduction}

At the beginning of epidemics, such as H1N1 (``swine flu'') or severe acute respiratory syndrome (SARS), the major concern is the severity of the illness,
and the large number of individuals that have the potential to be hospitalized after being infected.  
Accordingly, during the early stages of an epidemic
we can assume that most cases are identified as ill individuals are encouraged to seek medical help and practitioners
to send samples to the labs for testing.  After the first few
weeks of an epidemic, testing of sick individuals is often discouraged to ease the burned on public health labs.

In infectious disease modelling, we are often interested in estimating the case fatality ratio.  
That is, we are interesting in the proportion of cases
that eventually die from the disease.  However, crude rates calculated
throughout the epidemic often underestimate this quantity due to the large proportion of censoring present in the data.  The reason for this being that many 
individuals have yet to reach the terminal state of their illness, either recovery or death.
Additionally, these crude rates are often misleading due to shifts in case ascertainment which occur as the epidemic progresses because 
efforts from medical professionals become focussed on the more severe cases.  If these biases are not accounted for, they result in underestimation 
of the case fatality ratio (\cite{garske-bmj-2009}).  In fact, \cite{ghani-AmericanJournalEpi-2005}, even omit performing analyses during the first two months 
of the SARS epidemic due to a lack of data.

We propose a Bayesian approach to this problem using time-inhomogeneous multi-state models.  This approach takes into account expert optinion which allows us to 
get reasonable estimates at the beginning an epidemic when the data are sparse.  The use of survival analysis is
a logical approach because this is the standard way of dealing with the large amounts of censoring present in the data.  We illustrate our approach through a simulation study
with the aim of appropriately estimating the fatality rate, predicting the hospital load and predicting the number of unobserved cases over the course
of the epidemic.  We are also interested in forecasting the number of
cases beyond the completion our simulated epidemic.  The strength of this method is appropriately estimating these quantities at the 
beginning of the epidemic while the data are sparse and censored, rather than at the completion.

We also compare our methods to similar analyses...

This remainder of this paper proceeds as follows.  Section \ref{sec:multiStateModel} goes into detail describing our multi-state model and Section \ref{sec:modelInference}
discusses model inference.  The results of our simulation study are shown in Section \ref{sec:simStudyResults}.  Finally, our conclusions
are given in Section \ref{sec:conclusions}.

\section{Time-Inhomogeneous Multi-State Models}
\label{sec:multiStateModel}

\subsection{Infection Types}
\label{subsec:InfectionTypes}

We assumed that all individuals had one of three infection types:
mild, serious or deadly.  Individuals were classified as mild if
they recovered without needing to be hospitalized.  If an infection
was severe enough that hospitalization was required before recovery,
they were classified as having a serious infection type. Individuals
with a deadly infection, as the name suggests, were those that died
following hospitalization. For mild, serious and deadly cases,
respectively, we modelled the disease progression as follows:

\begin{equation}
\mbox{Infection} \stackrel{f_I(t)}{\Longrightarrow} \mbox{Onset} \stackrel{f_{OM}(t)}{\Longrightarrow} \mbox{Medical}  \stackrel{f_{MR}(t)}{\Longrightarrow} \mbox{Recovery} \label{eq:mildProgress}
\end{equation}

\begin{equation}
\mbox{Infection} \stackrel{f_I(t)}{\Longrightarrow} \mbox{Onset} \stackrel{f_{OS}(t)}{\Longrightarrow} \mbox{Medical} \stackrel{f_{SH}(t)}{\Longrightarrow} \mbox{Hospitalization} \stackrel{f_{SR}(t)}{\Longrightarrow} \mbox{Recovery} \label{eq:seriousProgress}
\end{equation}

\begin{equation}
\mbox{Infection} \stackrel{f_I(t)}{\Longrightarrow} \mbox{Onset} \stackrel{f_{OD}(t)}{\Longrightarrow} \mbox{Medical} \stackrel{f_{HD}(t)}{\Longrightarrow} \mbox{Hospitalization} \stackrel{f_{D}(t)}{\Longrightarrow} \mbox{Death}
\label{eq:deadlyProgress}
\end{equation}

Regardless of an individual's infection type, progression from one
stage to the next occurred according to a Weibull distribution with
parameters that corresponded to the respective transition and
infection type. Note that until an individual died or recovered
their infection type was unknown.  We also assumed that a small
proportion of mild cases were lost to follow up. That is, they
received medical consultation, and were assumed to recover since no
further information was available.

In this analysis, we assumed that the infection type probabilities varied with respect to age which we generated using log-linear cubic splines.  
This will be discussed further in Section \ref{subsec:infectionTypeProbs}.

\subsection{The Model}
\label{subsec:theModel}

We used a zero-inflated rounded Weibull for the time-to-event
distribution, denoted $f(t)$.  We used a rounded distribution
because the data only included the date the event took place. For
example, if there was an event recorded as being on the first day of
the epidemic, the event would actually have occurred at some point
between day 0.5 and day 1.5. To account for the large number of
administrative zeros recorded in the data, we used a zero-inflated
distribution.  That is, we let $\delta$ be the probability of
transitioning immediately to the death stage for an administrative
reason, such as arriving dead at the hospital. Furthermore, if we
let $Z$ denote the random variable for administrative zeros, we can
assume $Z$ follows a Bernoulli distribution with probability
$\delta$ and hence,

\begin{align*}
\text{Pr}(T=t|Z=1) =&\text{ }0 \text{ with probability } 1
\end{align*}

\noindent For the case where the individual was not an
administrative zero

\begin{align*}
\text{Pr}(T=t|Z=0) =&  \text{Pr}[\max(t-0.5,0) < y < t+0.5]\\
y \sim & \text{Weibull}\left [\frac{\mu}{\Gamma(1 + 1/\nu)}, \nu
\right]
\end{align*}

\noindent where $\mu = \mbox{E}[T = t|Z = 0]$ is the average time to
an event and $\nu$ is the Weibull shape parameter. We used beta
priors for the zero-inflation parameters and gamma priors for the
Weibull mean and shape parameters.

\subsection{Infection Type Probabilties}
\label{subsec:infectionTypeProbs}

As mentioned above in Section \ref{subsec:InfectionTypes}, we assumed the infection type probabilities varied with respect to age 
and generated them using log-linear cubic splines.  For individual $i$ we let $W_i$ $\in$ \{$M$, $S$, $D$\}
denote the infection type, where $M$, $S$ and $D$ represent mild,
serious and deadly infections, respectively and $X_i$ represents the
individual's age. The fatality rate is the marginal probability
$\mbox{Pr}(W_i = D) = g_D(X_i)$ and the non-fatal hospitalization
rate is the conditional probability $\mbox{Pr}(W_i = S|W_i \neq D)=
g_S(X_i)$  where $g_D$ and $g_S$ are log-linear cubic splines with
fixed knots and random coefficients that vary smoothly with age.  Note that this was done using the R package, DPpackage.

 \begin{figure}[!h]
 \begin{center}
\includegraphics{Figures/G-ageVaryingDeadly}
 \end{center}
 \caption{Death probabilities.}
 \label{fig:ageVaryingDeadly}
 \end{figure}

\subsection{Discussion of Prior Distributions}

\section{Model Inference}
\label{sec:modelInference}

Inference for this model was done through data augmentation and
random walk Metropolis simulations for calculating event time
parameters and predicting the number of unobserved cases.  We also forecasted the number of cases beyond the end of the simulated epidemic.

\subsection{Data Augmentation}
\label{subsec:dataAugmentation}

In order to deal with the censoring, we performed data augmentation
to impute the lifetimes as follows:

\begin{enumerate}

\item choose starting values for the probabilities

\item simulate $[W_i = D ] \sim \mbox{Bernoulli}(p_i)$, where $p_i = \mbox{Pr}(W_i = D | T_i > k_i; X_i)$,
 $k_i$ is the number of days previous that individual $i$ was hospitalized and $X_i$ is the age of individual $i$

\item simulate $[g_D|W_1, \ldots, W_n]$ to update
the parameters

\end{enumerate}

\subsection{Event Time Parameters}
\label{subsec:eventTimeParams}

After the data augmentation step, the likelihood was straightforward
to calculate since the observations were assumed to be independent.
Moreover, we could easily use random walk Metropolis to calculate
the event time parameters.
 If we denote $\theta$ as the current step and $\phi$ as
the proposed value, then $q(\theta, \phi)$ is the proposal
distribution for moving from the current position $\theta$ to the
proposed position $\phi$. For this simulation the proposal
distribution was absolute normal. The acceptance probability for the
random walk was:

\begin{eqnarray*}
\alpha(\theta, \phi) =  \mbox{min}\left\{1,
\frac{\pi(\phi)}{\pi(\theta)}\right\}
\end{eqnarray*}

\noindent where $\pi(^.) \propto \ell(^.) p(^.)$.  Using this
notation $p(^.)$ is the prior distribution, $\ell(^.)$ is the
likelihood and
 $\pi(^.)$ is the posterior distribution.  We accepted the move if $u \leq \alpha$ where $u \sim \mbox{Unif}(0, 1)$ and rejected otherwise.

\subsection{Number of Unobserved Cases}
\label{subsec:unobservedCases}

Again, we used random walk Metropolis to calculate the number of
unobserved infections.  We let $y_j$ denote the number of observed
cases on day $j$ of a $J$ day epidemic and we assumed that for each day of
the simulation:

\begin{eqnarray*}
y_j \sim \mbox{Binomial} (N_j, p_j), j = 1, \ldots, J
\end{eqnarray*}

\noindent where $N_j$ is the total number of cases (including both
observed and unobserved) on the $j^{th}$ day of the epidemic and
$p_j$ is the probability of observing a case on day $j$. A Poisson
prior with mean $\lambda_{j}$ is used for $N_j$, where
$\lambda_{j}$ is $[N_{j_{\mbox{inf}}} \theta + \omega]\gamma\mbox{ , }j = 1, \ldots, J$.  Using this notation, $N_{j_{\mbox{inf}}}$ is the number of infective individuals on the $j^{th}$ day of the epidemic,
$\theta$ is the rate parameter, $\omega$ is the immigration parameter and $\gamma$ is the probability of the case being mild, serious
or deadly.

For this simulation, we accepted the proposed value with
probability:

\begin{eqnarray*}
\alpha(\theta, \phi) = \mbox{min} \left\{1, \frac{\pi(\phi)q(\phi,
\theta)}{\pi(\theta)q(\theta, \phi)} \right\}
\end{eqnarray*}

\noindent where $\pi(^.)$ is the posterior distribution and
$q(\theta, \phi)$ is the Poisson proposal distribution for day $j$ with mean $N_{j-1}$ + \pi_0$, where $\pi_0$ 
denotes the proposal offset or the mean increase in the number of new cases each day of the epidemic.  Note that for this simulation $q(\theta, \phi)$ is the 
conditional distribution, given $N_j$ is greater than $y_j - 0.5$.  Hence, in order for $q(\phi, \theta)$ to be a
valid probability density function we divided the conditional distribution by $P(Y_j > N_{j-1} + \pi_0)$,  where $Y_j$ are the 
individuals that have been infected, but have not had their medical consultation by day $j$. 
Similary, for $q(\theta, \phi)$, we divided by $P(Y_j > N_j + \pi_0)$.  
We accepted the proposed value if $u
\leq \alpha$ where $u \sim \mbox{Unif}(0,1)$.

Once this was complete, we performed data augmentation to estimate the progression of the cases from stage to stage by infection type as described in the 
multi-state models, \ref{eq:mildProgress} - \ref{eq:deadlyProgress} (Section \ref{subsec:InfectionTypes}).

\subsection{Forecasting the Number of Cases}
\label{subsec:forecasting}

The final goal of this paper was to predict the cases for a number of days beyond the completion of a simulated epidemic.  For a $J$ day
epidemic, if we denote $N_k$, $k$ = $J+1, \ldots, K$, as the number of cases on the $k^{th}$ day of the epidemic.  Then,

\begin{eqnarray*}
N_{k} \sim \mbox{Poisson}([N_{k_{\mbox{inf}}} \theta + \omega]\gamma), k = {J+1}, \ldots, K
\end{eqnarray*}

\noindent with $N_{k_{\mbox{inf}}}$, $\theta$, $\omega$ and $\gamma$ as defined in Section \ref{subsec:unobservedCases}.  Note that these forecasts were done separately
for each infection type.  Once this was complete, we performed data augmentation, as discussed in Section \ref{subsec:dataAugmentation}.

\section{Results of the Swine Flu Simulation Study}
\label{sec:simStudyResults}

This section summarizes the results of a simulated 20 day epidemic and then forecasted the number of cases for the next 10 days beyond the end of the
epidemic.

\subsection{Fatality Rate}
\label{subsec:fatalityRate}

\end{document}

### 2013-02-19 pkg/geostatsp/R/krige.R.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@510 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

krige = function(obj.model, geodata,  locations, covariates, locations.mean=locations,
		factor.info=NULL, exp.pred=FALSE,rasterMethod = c("ngb", "bilinear"),
		nugget.in.prediction=TRUE) {

NsimBoxCox=40

data = geodata@data	
theCovs = attributes(terms(obj.model$formula))$term.labels

varTypes = unlist(lapply(data, class))[theCovs]
thefactors = names(varTypes[varTypes=="factor"])

# if necessary, turn covariates into a raster stack
if(class(covariates) == "list") {

	themethod = rep(rasterMethod[1], length(covariates))
	names(themethod) = names(covariates)
	themethod[thefactors] = "ngb"

	locations.mean = stackRasterList(covariates,locations.mean, themethod)	
} else { # covariates must be a raster stack, use as-is
	if(length(names(covariates))==1 & length(theCovs)==1) {
		names(covariates) = theCovs
	}
	locations.mean = covariates
}

if(!all(names(locations.mean)%in% theCovs))
	warning("some covariates in the model formula weren't supplied")

# data frame of random field prediction locations 
locationsDF = as.data.frame(locations, xy=TRUE)

# convert data to factors, using same levels for model fit data and prediction data

# construct the fixed effects component
meanRaster = raster(locations.mean)
meanRaster[] = obj.model$beta["(Intercept)"]

for(D in theCovs){

	if(varTypes[D] == "factor") {
		if (D %in% names(factor.info)) {
			tofac = factor.info[[D]]
			if(all(c("levels","labels") %in% names(tofac)))
				tofac=data.frame(tofac[["levels"]],tofac[["labels"]])
			rownames(tofac) = tofac[,2]
			thelevels = tofac[levels(data[[D]]),1]
		}  else {
			thelevels = levels(data[[D]])	
		}
		theBetaNames = paste(D, levels(data[[D]]), sep="")
		betaHasPar = theBetaNames %in% names(obj.model$beta)
		betasHere = rep(0, length(theBetaNames))
		names(betasHere) = theBetaNames
		betasHere[betaHasPar] = obj.model$beta[theBetaNames[betaHasPar]]
		names(betasHere) = thelevels

		toAdd = setValues(locations.mean[[D]],
				betasHere[as.character(
								locations.mean[[D]]@data@values
						)]
		)

		meanRaster = meanRaster + toAdd

	} else {

		meanRaster = meanRaster + obj.model$beta[D]*locations.mean[[D]]
	}
}
names(meanRaster) = "fixed"
# do the kriging

data.col = strsplit(as.character(obj.model$formula), "~")[[2]]
geodataForKrige = as.geodata(geodata, 
		data.col=data.col, 
		covar.col=theCovs)
if(obj.model$lambda != 1) {
	if(obj.model$lambda == 0) {
		geodataForKrige$data =log(geodataForKrige$data)
	} else{
		geodataForKrige$data =(geodataForKrige$data^obj.model$lambda -1)/obj.model$lambda
	} 
}
dummyDF = locationsDF
for(D in theCovs){
	if(varTypes[D] == "factor") {
		dummyDF[,D] = factor(levels(data[,D])[1], levels=levels(data[,D]))
	}
	else
		dummyDF[,D] = 0	
}
trend.d= geostatsp:::trend.spatial(obj.model$trend, data)	
trend.l = geostatsp:::trend.spatial(obj.model$trend, dummyDF)

obj.modelNoLambda = obj.model
obj.modelNoLambda$lambda =1
thecontrol = krige.control(obj.model=obj.modelNoLambda,
		trend.d=trend.d, 
		trend.l=trend.l)

krigeResult = krige.conv(
		geodataForKrige, locations=	locationsDF[,1:2],
		krige=thecontrol
)

rastKrige = setValues(locations[[1]], krigeResult$predict - obj.model$beta["(Intercept)"])
rastKrige = addLayer(rastKrige, 
		setValues(locations[[1]], krigeResult$krige.var)
)
names(rastKrige) = c("predict","krige.var")

if(as(meanRaster, "BasicRaster")==as(rastKrige, "BasicRaster")) {
	result = addLayer(meanRaster,
			rastKrige
	)
}	 else {
	result = addLayer(meanRaster,
			raster::resample(rastKrige, meanRaster, method="ngb")
	)
}

names(result)[names(result)=="predict"] = "random"
result = addLayer(result, 
		predict=result[["fixed"]] + result[["random"]]
)
names(result)[names(result)=="layer"] = "predict"

if(exp.pred | obj.model$lambda==0){
	names(result)[names(result)=="predict"] = "predict.log"
	result = addLayer(result, 
			exp(result[["predict.log"]]+ 0.5*result[["krige.var"]] +
							0.5* nugget.in.prediction * obj.model$nugget)
	)
	names(result)[names(result)=="layer"] = "predict"

}

# box-cox
if(!any(obj.model$lambda==c(0,1))){

	names(result)[names(result)=="predict"] = "predict.boxcox"

	bcpred= 0
	Ndata = length(result[["predict.boxcox"]][])

	themean = result[["predict.boxcox"]][]
	if(nugget.in.prediction){
		thesd = sqrt(result[["predict.boxcox"]][] + obj.model$nugget)
 	} else {
		thesd = sqrt(result[["predict.boxcox"]][])
	}
	thesd[is.na(thesd)] = 0 
	themean[is.na(themean)] = 0

	invlambda = 1/obj.model$lambda

	for(D in 1:NsimBoxCox) {
		bcpred = bcpred + (obj.model$lambda *rnorm(Ndata, themean, thesd)+1)^invlambda
	}
	bcpred = bcpred / NsimBoxCox
	bcpred = raster(matrix(bcpred, ncol=result@ncols,byrow=T), 
			result@extent@xmin,result@extent@xmax,
			result@extent@ymin,result@extent@ymax,crs=result@crs)

	result = addLayer(result, 
			bcpred)

	names(result)[names(result)=="layer"] = "predict"

}

if(as(result, 'BasicRaster')!=as(rastKrige, "BasicRaster")) {
	result = list(random = rastKrige, prediction=result)			
}

result
}

### 2013-11-13 pkg/mapmisc/inst/doc/mapmisc.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@625 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Overview of mapping with mapmisc}

\usepackage[margin=1in,landscape]{geometry}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.png}

\usepackage{subcaption,float}
\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}

\title{The mapmisc package}
\author{Patrick Brown}

\begin{document}
\maketitle

<<setup,echo=FALSE,cache=FALSE,results='hide',message=FALSE>>=
library('knitr')
opts_chunk$set(echo=TRUE,message=FALSE,
		dev='png',
		fig.align='center',fig.pos='H',width=80,
		fig.height=6,fig.width=4,out.width="0.3\\textwidth"
)
par(bg="white")
library(mapmisc)
library(rgdal)
@

\section{Introduction}

This package provides a few utilities for making nice maps in R, with an
emphasis on enabling maps in short tidy code chunks which are suitable for
Sweave and knitr documents.  The package duplicates the capabilities of
packages such as classInt, geonames, and OpenStreetMap, and the price of having
tidier code is much of the flexibility from these other packages has been lost
here.

The meuse data from the sp package
<<meusedata,echo=TRUE>>=
library('mapmisc')
library(sp)
data(meuse)
coordinates(meuse) <- ~x+y
proj4string(meuse) <- CRS("+init=epsg:28992")
class(meuse)
@

Get elevation data with
<<elevation,eval=FALSE>>=
meuseLL = spTransform(meuse, CRS("+init=epsg:4326"))
getData("SRTM", lon=xmin(extent(meuseLL)), 
	lat=ymin(extent(meuseLL)),path=tempdir())
nldElev = raster(paste(tempdir(), "/", "srtm_38_02.tif", sep=""))
nldElev = crop(nldElev, extent(meuseLL))
nldElev = projectRaster(nldElev, crs=proj4string(meuse))
@
Or simply do
<<dataneth>>=
data(netherlands)
@

The elevation data is a Raster.
<<classElev>>=
class(nldElev)
nldElev = crop(nldElev, extend(extent(meuse), 1000))
@

\section{Downloading background maps and city locations}

Get a background map covering the extent of the meuse data
<<openmap,eval=FALSE>>=
nldTiles = openmap(meuse)
@
{\tt nldTiles} is a Raster with the sampe projection as {\tt meuse}
<<openmapInfo>>=
class(nldTiles)
proj4string(nldTiles)
proj4string(meuse)
@

{\tt GNcities} is a wrapper for the function of the same name in the geonames
package.

<<cities,eval=FALSE>>=
nldCities = GNcities(meuse, maxRows=6)
@
A SpatialPointsDataFrame, with same map projection.
<<citiesInfo>>=
class(nldCities)
names(nldCities)
proj4string(nldCities)
@

\section{Making maps}

The {\tt map.new} function sets up a map in the current plot window with the
correct limits and aspect ratio for the object supplied, and without margins or
white space. {\tt scaleBar} adds a scale and north arrow.  It uses the map
projection of the argument supplied to calculate distances and find north.

<<simpleMap,fig.cap="simple map", fig.subcap=c("data locations","cities","elevation")>>=

# plot the data locations
map.new(meuse)
plot(nldTiles, add=TRUE)
points(meuse,col="red", cex=0.3)
scaleBar(proj4string(meuse),pos="topleft", bg="white")

# plot city names
map.new(meuse)
plot(nldTiles, add=TRUE)
points(nldCities)
text(nldCities, labels=nldCities$name, pos=3)
scaleBar(proj4string(meuse),pos="topleft", bg="white")

# plot elevation
map.new(meuse, legendRight=TRUE)
plot(nldTiles, add=TRUE)
plot(nldElev,add=TRUE,col=terrain.colors(8),alpha=0.6,legend.mar=2, legend.line=0,)
scaleBar(meuse,pos="topleft",bg="white")
@

Reproject the data so it's roughly aligned with the x-axis.

<<reproj,cache=TRUE>>=
rotatedCRS = CRS("+proj=omerc +lat_0=51 +lonc=5.75 +alpha=-50.0000000 +gamma=0.0 +k=1.000000 +x_0=500000.000 +y_0=0.000 +ellps=WGS84	+units=m")
meuseRot = spTransform(meuse, rotatedCRS)
tilesRot = projectRaster(nldTiles, crs=rotatedCRS, method='ngb')
tilesRot@legend@colortable = nldTiles@legend@colortable
elevRot = projectRaster(nldElev, crs=rotatedCRS)
@

And create new plots

<<rotatedMap,fig.cap="Rotated map", fig.subcap=c("elevation","data"), fig.height=3,fig.width=4,out.width="0.48\\textwidth">>=
# first elevation
map.new(meuseRot)
plot(tilesRot, add=TRUE)
plot(elevRot,add=TRUE,alpha=0.5,col=terrain.colors(8), legend=FALSE)
points(nldCities)
text(nldCities, labels=nldCities$name, pos=3)
scaleBar(meuseRot,pos="topleft", bg="white")

# then data locations
map.new(meuseRot)
plot(tilesRot, add=TRUE)
points(meuseRot,col="red", cex=0.3)
scaleBar(proj4string(meuseRot), bg="white")
@

\section{Inset maps}

<<world,eval=FALSE>>=
world = openmap(extent(-10,30,40,60),path="osm-no-labels")
@

<<inset,fig.cap="Inset map", fig.subcap=c("elevation","rotated"), fig.height=3,fig.width=4,out.width="0.48\\textwidth">>=

# not rotated
map.new(meuse)
plot(nldTiles, add=TRUE)
points(meuse)
scaleBar(proj4string(meuse),pos="left", bg="white")
insetMap(meuse, "topleft",map=world)

# rotated
map.new(meuseRot)
plot(tilesRot, add=TRUE)
points(meuseRot,col="red", cex=0.3)
scaleBar(proj4string(meuseRot), bg="white")
insetMap(meuseRot, "bottomleft",map=world)
@

\section{Legends}

Create a colour scale for plotting copper  concentrations
<<meuseScale>>=
cuScale = colourScale(meuse$copper, breaks=5, style='equal', 
		opacity=0.8, dec=-1, firstBreak=0)
@
and elevation, with transparency decreasing as elevation increases.
<<elevScale>>=
elevScale = colourScale(values(nldElev), breaks=6, col=terrain.colors,
		firstBreak=0, dec=-1,opacity=c(0.2, 0.9))
@
Soil type is a categorical variable, create a factor and create a colour scale
of unique values
<<soilScale>>=
meuse$soilFac = factor(meuse$soil, levels=c(1,2,3), 
		labels=c("Calcareous","Non-Calc's","Red Brick"))
soilScale = colourScale(meuse$soilFac, style="unique",
		col="Set2")
@

<<meuseLegends,fig.subcap=c("Copper","soil", "elevation"), fig.cap="Meuse data again">>=
map.new(meuse)
plot(nldTiles, add=TRUE)
plot(meuse, col=cuScale$col,add=TRUE,pch=16)
scaleBar(meuse, pos="topleft",bg="white")
legendBreaks("bottomright",  breaks=cuScale,
		title="gals/firkin",bg="white",adj=c(0,-2))

map.new(meuse)
plot(nldTiles, add=TRUE)
plot(meuse, col=soilScale$col,add=TRUE,pch=16)
scaleBar(meuse, pos="topleft",bg="white")
legend("bottomright", fill=soilScale$legendCol, legend=soilScale$breaks,
		title="soil type", cex=0.7,bg="white")

map.new(meuse)
plot(nldTiles, add=TRUE)
image(nldElev, breaks=elevScale$breaks, col=elevScale$withTrans, legend=FALSE,add=TRUE)
scaleBar(meuse, pos="topleft",bg="white")
legendBreaks("left", breaks=elevScale, title='Metres',bg="white")
@

\end{document}

### 2013-11-14 pkg/mapmisc/inst/doc/mapmisc.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@626 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[margin=1in,landscape]{geometry}
-\usepackage{graphicx}
+\usepackage{graphicx,hyperref}
 \DeclareGraphicsExtensions{.png}
@@
 		fig.align='center',fig.pos='H',width=80,
-		fig.height=6,fig.width=4,out.width="0.3\\textwidth"
+		fig.height=5,fig.width=3,out.width="0.3\\textwidth"
 )
@@
 @
+Maps which can be downloaded are shown at
+\url{http://diseasemapping.r-forge.r-project.org/openmap/}
 
@@
 # not rotated
-map.new(meuse)
+map.new(meuse,legendRight=TRUE)
 plot(nldTiles, add=TRUE)
 points(meuse)
-scaleBar(proj4string(meuse),pos="left", bg="white")
-insetMap(meuse, "topleft",map=world)
+scaleBar(proj4string(meuse),pos="bottomright", bg="white")
+insetMap(meuse, "topright",map=world)
 

### 2013-11-18 pkg/mapmisc/inst/doc/mapmisc.rnw
changed results of colourScale to plot, col, colOpacity, also made methods for factors, rasters
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@629 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<elevScale>>=
-elevScale = colourScale(values(nldElev), breaks=6, col=terrain.colors,
+elevScale = colourScale(nldElev, style='equal',
+		breaks=6, col=terrain.colors,
 		firstBreak=0, dec=-1,opacity=c(0.2, 0.9))
@@
 plot(nldTiles, add=TRUE)
-plot(meuse, col=cuScale$col,add=TRUE,pch=16)
+plot(meuse, col=cuScale$plot,add=TRUE,pch=16)
 scaleBar(meuse, pos="topleft",bg="white")
@@
 plot(nldTiles, add=TRUE)
-plot(meuse, col=soilScale$col,add=TRUE,pch=16)
+plot(meuse, col=soilScale$plot,add=TRUE,pch=16)
 scaleBar(meuse, pos="topleft",bg="white")
-legend("bottomright", fill=soilScale$legendCol, legend=soilScale$breaks,
+legendBreaks("bottomright", breaks=soilScale,
 		title="soil type", cex=0.7,bg="white")
@@
 plot(nldTiles, add=TRUE)
-image(nldElev, breaks=elevScale$breaks, col=elevScale$withTrans, legend=FALSE,add=TRUE)
+image(nldElev, breaks=elevScale$breaks, col=elevScale$colOpacity, 
+		legend=FALSE,add=TRUE)
 scaleBar(meuse, pos="topleft",bg="white")

### 2015-03-27 pkg/gmrf/vignettesStuff/sim.rnw
waiting for cholmod_solve2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1094 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{caption,subcaption,float,amsmath,filecontents}
\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}

<<setup,include=FALSE>>=
library('knitr')

chapter=''
opts_chunk$set(echo=TRUE, 
		dev='pdf', marginsp=1,  
		fig.align='',fig.pos='H',
		fig.width=5,fig.height=4,out.width='\\textwidth',
		tidy=TRUE,tidy.opts=list(reindent.spaces=2,
				width.cutoff=45)
)

knit_hooks$set(
		marginsp = function(before, options, envir){	
			if(!before) return()
# use small margins				
			par(mar=c(1.5+0.9*options$marginsp,
							1.5+0.9*options$marginsp,0.2,0.2), 
					mgp=c(1.45, 0.45, 0),cex=1.25)
		}		
)

options(width=67)

@

\begin{document}

Hello $1^5=1^6$ therefore $5=6$.

Simulate some data
<<simData,cache=TRUE>>=
set.seed(0)
library(geostatsp)

myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=6*xres(myraster),shape=2,variance=900)

theU = RFsimulate(myraster,model=themodel)

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
		rep(4,ncell(thecov)/2))
beta.x=5
theY = theU + beta.x*thecov

theData = brick(theY,thecov)
names(theData) = c("y","x")

fracNugget = 1/2
nuggetSd = sqrt(themodel['variance']*fracNugget)
thedf = as.data.frame(theData)
thedf$intercept = 1
Xmat = as.matrix(thedf[,c("intercept","x")])
thedf$yNoise = rnorm(nrow(thedf),
		thedf$y,nuggetSd)

@

nearest neighbours
<<nnmat,cache=TRUE>>=
theNN = NNmat(theData)
@

fit the model
<<fit,cache=TRUE>>=

Sar = exp(seq(log(0.01),log(0.1),len=24))

source("../R/loglikGmrf.R")
source("../R/lgmrfm.R")
#source("../R/conditionalGmrf.R")

Sar2 =  seq((0.05),(0.4),len=24)
Snugget =  seq((0.1), (1),len=25)
maternShape = themodel['shape']

resVanilla = loglikGmrf(
		oneminusar=Sar2 ,
		propNugget = Snugget ,
		Yvec=thedf$yNoise,
		Xmat=Xmat,
		shape=maternShape,
		NN=theNN,mc.cores=4
)

@

make some plots
<<plots,out.width='0.45\\textwidth',fig.cap='the results',fig.subcap=c('L','second'),echo=FALSE,marginsp=TRUE>>=

dseq = rev(c(0,0.5, 1,2,4,8,20,100))
thecol = mapmisc::colourScale(resVanilla['logL.ml',,],
		breaks=max(resVanilla['logL.ml',,]) - dseq, 
		col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resVanilla['propNugget',,1]), 
		range(resVanilla['oneminusar',1,]),type='n',
		xlab='tausq/xisq',ylab='1-phi')
.filled.contour(resVanilla['propNugget',,1], resVanilla['oneminusar',1,],resVanilla['logL.ml',,],
		col=thecol$col,levels=thecol$breaks
)

mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq,
		inset=0.025)

plot(1:10)

@
\end{document}

### 2015-03-27 pkg/gmrf/vignettesStuff/so2.rnw
waiting for cholmod_solve2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1094 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{caption,subcaption,float,amsmath,filecontents}
\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}

<<setup,include=FALSE>>=
library('knitr')

chapter=''
opts_chunk$set(echo=TRUE, 
		dev='pdf', marginsp=1,  
		fig.align='',fig.pos='H',
		fig.width=4,fig.height=3,out.width='\\textwidth',
		fig.path=file.path('Figures', chapter, ""),
		cache.path=file.path('cache', chapter, ""),
		tidy=TRUE,tidy.opts=list(reindent.spaces=2,
				code.width=45)
)

heavyCacheDate=NULL
lightCacheDate=NULL

knit_hooks$set(
		marginsp = function(before, options, envir){	
			if(!before) return()
# use small margins				
			par(mar=c(1.5+0.9*options$marginsp,
							1.5+0.9*options$marginsp,0.2,0.2), 
					mgp=c(1.45, 0.45, 0),cex=1.25)
		}		
)

library('RandomFields')
library('raster')
library('rgdal')
library('geostatsp')
library('mapmisc')
library('Hmisc')

options(width=67)

dataDir='/store/patrick/spatialData'

@

\title{SO2 Analysis}
\author{Patrick Brown}
\begin{document}
\maketitle

<<>>=
source("../R/loglikGmrf.R")
source("../R/loglikGmrfNugget.R")
@

<<world,cache=TRUE,echo=FALSE,cache.extra=1>>=
worldShapeFile =  system(paste("find ", dataDir, 
				" -name *countries*.shp", sep=""),TRUE)
worldShapeFile = gsub("\\.shp$","",worldShapeFile)
world = readOGR(dataDir,gsub("[[:print::]+/","",worldShapeFile),
		stringsAsFactors=FALSE,verbose=FALSE)
@

\section{SO2}
<<>>=
load("chinaSo2.RData")
@
<<>>=
china2
@

<<so2Plot,fig.cap='so2 data',out.width='0.5\\textwidth',fig.subcap=c('so2','population'),fig.height=4,fig.width=5,echo=FALSE,dev='png',cache=TRUE>>=
library(mapmisc)
map.new(china2)
plot(china2[['SO2.median']],add=TRUE)
plot(world,add=TRUE)
map.new(china2)
plot(china2[['pop']],add=TRUE)
plot(world,add=TRUE)

@

\subsection{Simplest model}

\begin{align*}
Y_i  =  & X_i\beta + U_i  \\
[U_1 \ldots U_N] \sim &\text{GMRF}
\end{align*}

<<chinaForGmrf,cache=TRUE,cache.extra=1>>=
chinaForGmrf = 
		list(Xmat = 
			model.matrix(
  ~ popLow + popHigh + popVeryHigh + latLow,
			chinaDF),
			Yvec=chinaDF$SO2.median,
				maternShape=2
		)
	chinaForGmrf$NN = 
				NNmat(china2,
						nearest=chinaForGmrf$maternShape+1)

@

<<so2OneL,echo=TRUE>>=
so2oneres = loglikGmrf(ar=0.8, Yvec=chinaForGmrf$Yvec, 
		Xmat=chinaForGmrf$Xmat, 
		NN=chinaForGmrf$NN, 
		maternShape=chinaForGmrf$maternShape)
as.vector(so2oneres)
@

<<so2CarbonFitGmrf,cache=TRUE,dependson='chinaForGmrf'>>=
arseq = seq(0.5,1,len=200)
arseq = arseq[-length(arseq)]
library(parallel)
so2Res = mcmapply(loglikGmrf, arseq,
		MoreArgs=chinaForGmrf,
		mc.cores=4,SIMPLIFY=TRUE)
@

<<so2summary>>=
so2summary = summaryGmrfFit(so2Res,chinaForGmrf) 
so2summary$summary
@

<<so2profLplot,fig.width=6,fig.height=4,out.width='0.6\\textwidth',fig.cap='SO2 profile l',echo=FALSE>>=
plot(so2summary$profL[,'rangeInCells'], 
		so2summary$profL[,'logL'], type='l',
		xlab='range (in cells)',ylab='log L')
abline(v=so2summary$summary['rangeInCells',
				c('mle','q0.025','q0.975')])

@

\subsection{Model with nugget effect}
\begin{align*}
Y_i|U_i  \sim  & \text{N}(\mu_i, \tau^2)\\
\mu_i = & X_i\beta + U_i  \\
[U_1 \ldots U_N] \sim &\text{GMRF}
\end{align*}

<<so2OneLnugget,echo=TRUE,cache=TRUE>>=
so2oneNugget = loglikGmrfNugget(ar=0.95,
		propNugget = seq(0, len=5,by=0.001),
		Yvec=chinaForGmrf$Yvec, 
		Xmat=chinaForGmrf$Xmat, 
		NN=chinaForGmrf$NN, 
		maternShape=chinaForGmrf$maternShape)
@

<<so2oneNuggetRes>>=
so2oneNugget[c('logL','propNugget','sigmasq','tausq'),]
@

<<so2NuggetFit,cache=TRUE >>=
arseqN = seq(0.94,0.99,len=41)
arseqN=arseqN[arseqN < 1]
chinaForGmrf$propNugget = seq(0,0.003,len=21)
library(parallel)
so2ResN = mcmapply(loglikGmrfNugget, 
		ar=arseqN,
		MoreArgs=chinaForGmrf,
		mc.cores=4,SIMPLIFY='array')
@

<<so2NuggetProf,out.width='0.5\\textwidth',fig.cap='SO2 Prof L',fig.height=5,fig.width=7,fig.subcap=c('gmrf params','matern params'),echo=FALSE,warnings=FALSE>>=
Sprob = c(1, 0.9999, 0.999, 0.99, 0.95, 0.8, 0.5,   0) 
Schisq = qchisq(Sprob,df=2)
thebreaks = max(so2ResN['logL',,]) - Schisq

thebreaks[thebreaks==-Inf] = min(so2ResN['logL',,])-1

thecol = rev(RColorBrewer::brewer.pal( length(Sprob)-1,"Spectral"))

image(
		so2ResN['propNugget',,1],
		so2ResN['ar',1,],
		so2ResN['logL',,],
		breaks = thebreaks,
		col=thecol
)

image(
		so2ResN['propNugget',,1],
		so2ResN['rangeInCells',1,],
		so2ResN['logL',,],
		breaks = thebreaks,
		col=thecol,
		log='y'
)
mapmisc::legendBreaks('right',breaks=Sprob,col=thecol)
@

<<summarySo2N,cache=TRUE>>=
summarySo2N = summaryGmrfFitNugget(so2ResN,chinaForGmrf)
@

<<so2NuggetProfUniv,out.width='0.3\\textwidth',fig.cap='univariate prof l', fig.subcap=names(summarySo2N$profL),fig.width=4,fig.height=5,echo=FALSE>>=
for(D in names(summarySo2N$profL)) {

	plot(summarySo2N$profL[[D]] ,xlab=D,ylab='log L',type='l')
	abline(v=summarySo2N$summary[D,c('mle','q0.025','q0.975')],
			col=c('red','orange','orange'))

}

@

<<condnlGmrf,cache=TRUE>>=
so2cond=conditionalGmrf(summarySo2N$summary[,"mle"], 
		chinaForGmrf,mc.cores=4)
@
<<so2map>>=
so2map = brick(china2,nl=ncol(so2cond))
values(so2map) = as.vector(so2cond)
names(so2map) = colnames(so2cond)
@

\section{Black carbon}
<<>>=
load("blackCarbon.RData")
@

<< blackCarbonLm,cache=TRUE>>=
		bcLm = lm(y ~ popHigh + popLow + nightLow + nightHigh,
				data=bcAsiaDF)
@

<< bcLmSummary>>=
		summary(bcLm)$coef
@

<< bcForGmrf,cache=TRUE>>=
		bcForGmrf = list(Xmat = model.matrix(bcLm),
				Yvec=bcLm$model$y,
				maternShape=2
		)
		bcForGmrf$NN = 
				NNmat(bcAsiaBrick,
						nearest=bcForGmrf$maternShape+1)

@

<<bcOneL>>=
loglikGmrf(ar=0.8, Yvec=bcForGmrf$Yvec, Xmat=bcForGmrf$Xmat, 
		NN=bcForGmrf$NN, maternShape=bcForGmrf$maternShape)
@

<< blackCarbonFitGmrf,cache=TRUE>>=
arseq = seq(0.9,1,len=100)
library(parallel)
bcRes = mcmapply(loglikGmrf, arseq,
		MoreArgs=bcForGmrf,
		mc.cores=4,SIMPLIFY=TRUE)
@

<<>>=
bcSummary = summaryGmrfFit(bcRes,bcForGmrf) 
bcSummary$summary
@

\end{document}

### 2015-03-27 pkg/gmrf/vignettesStuff/so2Data.rnw
waiting for cholmod_solve2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1094 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2015-03-27 pkg/gmrf/vignettesStuff/so2.rnw
@@
 
+\title{SO2 Data}
+
 <<setup,include=FALSE>>=
@@
 chapter=''
-opts_chunk$set(echo=TRUE, 
+opts_chunk$set(echo=FALSE, 
 		dev='pdf', marginsp=1,  
@@
 		cache.path=file.path('cache', chapter, ""),
-		tidy=TRUE,tidy.opts=list(reindent.spaces=2,
-				code.width=45)
+		tidy=TRUE,tidy.opts=list(reindent.spaces=2,code.width=45)
 )
@@
 
-options(width=67)
-
-dataDir='/store/patrick/spatialData'
+evalP = Sys.info()['user'] =='patrick'
 
@@
 
-\title{SO2 Analysis}
-\author{Patrick Brown}
 \begin{document}
-\maketitle
 
-<<>>=
-source("../R/loglikGmrf.R")
-source("../R/loglikGmrfNugget.R")
+<<child="/home/patrick/workspace/ssr/data/ozone.Rnw">>=
 @
 
-<<world,cache=TRUE,echo=FALSE,cache.extra=1>>=
-worldShapeFile =  system(paste("find ", dataDir, 
-				" -name *countries*.shp", sep=""),TRUE)
-worldShapeFile = gsub("\\.shp$","",worldShapeFile)
-world = readOGR(dataDir,gsub("[[:print::]+/","",worldShapeFile),
-		stringsAsFactors=FALSE,verbose=FALSE)
-@
-
-\section{SO2}
-<<>>=
-load("chinaSo2.RData")
-@
-<<>>=
-china2
-@
-
-<<so2Plot,fig.cap='so2 data',out.width='0.5\\textwidth',fig.subcap=c('so2','population'),fig.height=4,fig.width=5,echo=FALSE,dev='png',cache=TRUE>>=
-library(mapmisc)
-map.new(china2)
-plot(china2[['SO2.median']],add=TRUE)
-plot(world,add=TRUE)
-map.new(china2)
-plot(china2[['pop']],add=TRUE)
-plot(world,add=TRUE)
-
-@
-
-\subsection{Simplest model}
-
-\begin{align*}
-Y_i  =  & X_i\beta + U_i  \\
-[U_1 \ldots U_N] \sim &\text{GMRF}
-\end{align*}
-
-<<chinaForGmrf,cache=TRUE,cache.extra=1>>=
-chinaForGmrf = 
-		list(Xmat = 
-			model.matrix(
-  ~ popLow + popHigh + popVeryHigh + latLow,
-			chinaDF),
-			Yvec=chinaDF$SO2.median,
-				maternShape=2
-		)
-	chinaForGmrf$NN = 
-				NNmat(china2,
-						nearest=chinaForGmrf$maternShape+1)
-
-@
-
-<<so2OneL,echo=TRUE>>=
-so2oneres = loglikGmrf(ar=0.8, Yvec=chinaForGmrf$Yvec, 
-		Xmat=chinaForGmrf$Xmat, 
-		NN=chinaForGmrf$NN, 
-		maternShape=chinaForGmrf$maternShape)
-as.vector(so2oneres)
-@
-
-<<so2CarbonFitGmrf,cache=TRUE,dependson='chinaForGmrf'>>=
-arseq = seq(0.5,1,len=200)
-arseq = arseq[-length(arseq)]
-library(parallel)
-so2Res = mcmapply(loglikGmrf, arseq,
-		MoreArgs=chinaForGmrf,
-		mc.cores=4,SIMPLIFY=TRUE)
-@
-
-<<so2summary>>=
-so2summary = summaryGmrfFit(so2Res,chinaForGmrf) 
-so2summary$summary
-@
-
-<<so2profLplot,fig.width=6,fig.height=4,out.width='0.6\\textwidth',fig.cap='SO2 profile l',echo=FALSE>>=
-plot(so2summary$profL[,'rangeInCells'], 
-		so2summary$profL[,'logL'], type='l',
-		xlab='range (in cells)',ylab='log L')
-abline(v=so2summary$summary['rangeInCells',
-				c('mle','q0.025','q0.975')])
-
-@
-
-\subsection{Model with nugget effect}
-\begin{align*}
-Y_i|U_i  \sim  & \text{N}(\mu_i, \tau^2)\\
-\mu_i = & X_i\beta + U_i  \\
-[U_1 \ldots U_N] \sim &\text{GMRF}
-\end{align*}
-
-<<so2OneLnugget,echo=TRUE,cache=TRUE>>=
-so2oneNugget = loglikGmrfNugget(ar=0.95,
-		propNugget = seq(0, len=5,by=0.001),
-		Yvec=chinaForGmrf$Yvec, 
-		Xmat=chinaForGmrf$Xmat, 
-		NN=chinaForGmrf$NN, 
-		maternShape=chinaForGmrf$maternShape)
-@
-
-<<so2oneNuggetRes>>=
-so2oneNugget[c('logL','propNugget','sigmasq','tausq'),]
-@
-
-<<so2NuggetFit,cache=TRUE >>=
-arseqN = seq(0.94,0.99,len=41)
-arseqN=arseqN[arseqN < 1]
-chinaForGmrf$propNugget = seq(0,0.003,len=21)
-library(parallel)
-so2ResN = mcmapply(loglikGmrfNugget, 
-		ar=arseqN,
-		MoreArgs=chinaForGmrf,
-		mc.cores=4,SIMPLIFY='array')
-@
-
-<<so2NuggetProf,out.width='0.5\\textwidth',fig.cap='SO2 Prof L',fig.height=5,fig.width=7,fig.subcap=c('gmrf params','matern params'),echo=FALSE,warnings=FALSE>>=
-Sprob = c(1, 0.9999, 0.999, 0.99, 0.95, 0.8, 0.5,   0) 
-Schisq = qchisq(Sprob,df=2)
-thebreaks = max(so2ResN['logL',,]) - Schisq
-
-thebreaks[thebreaks==-Inf] = min(so2ResN['logL',,])-1
-
-thecol = rev(RColorBrewer::brewer.pal( length(Sprob)-1,"Spectral"))
-
-image(
-		so2ResN['propNugget',,1],
-		so2ResN['ar',1,],
-		so2ResN['logL',,],
-		breaks = thebreaks,
-		col=thecol
-)
-
-image(
-		so2ResN['propNugget',,1],
-		so2ResN['rangeInCells',1,],
-		so2ResN['logL',,],
-		breaks = thebreaks,
-		col=thecol,
-		log='y'
-)
-mapmisc::legendBreaks('right',breaks=Sprob,col=thecol)
-@
-
-<<summarySo2N,cache=TRUE>>=
-summarySo2N = summaryGmrfFitNugget(so2ResN,chinaForGmrf)
-@
-
-<<so2NuggetProfUniv,out.width='0.3\\textwidth',fig.cap='univariate prof l', fig.subcap=names(summarySo2N$profL),fig.width=4,fig.height=5,echo=FALSE>>=
-for(D in names(summarySo2N$profL)) {
-
-	plot(summarySo2N$profL[[D]] ,xlab=D,ylab='log L',type='l')
-	abline(v=summarySo2N$summary[D,c('mle','q0.025','q0.975')],
-			col=c('red','orange','orange'))
-
-}
-
-@
-
-<<condnlGmrf,cache=TRUE>>=
-so2cond=conditionalGmrf(summarySo2N$summary[,"mle"], 
-		chinaForGmrf,mc.cores=4)
-@
-<<so2map>>=
-so2map = brick(china2,nl=ncol(so2cond))
-values(so2map) = as.vector(so2cond)
-names(so2map) = colnames(so2cond)
-@
-
-\section{Black carbon}
-<<>>=
-load("blackCarbon.RData")
-@
-
-<< blackCarbonLm,cache=TRUE>>=
-		bcLm = lm(y ~ popHigh + popLow + nightLow + nightHigh,
-				data=bcAsiaDF)
-@
-
-<< bcLmSummary>>=
-		summary(bcLm)$coef
-@
-
-<< bcForGmrf,cache=TRUE>>=
-		bcForGmrf = list(Xmat = model.matrix(bcLm),
-				Yvec=bcLm$model$y,
-				maternShape=2
-		)
-		bcForGmrf$NN = 
-				NNmat(bcAsiaBrick,
-						nearest=bcForGmrf$maternShape+1)
-
-@
-
-<<bcOneL>>=
-loglikGmrf(ar=0.8, Yvec=bcForGmrf$Yvec, Xmat=bcForGmrf$Xmat, 
-		NN=bcForGmrf$NN, maternShape=bcForGmrf$maternShape)
-@
-
-<< blackCarbonFitGmrf,cache=TRUE>>=
-arseq = seq(0.9,1,len=100)
-library(parallel)
-bcRes = mcmapply(loglikGmrf, arseq,
-		MoreArgs=bcForGmrf,
-		mc.cores=4,SIMPLIFY=TRUE)
+<<chinaDF>>=
+china2 = crop(china, extent(80,100,20,35))
+chinaDF = as.data.frame(china2, xy=TRUE)
+chinaDF[is.na(chinaDF$pop),'pop'] = 0
+chinaDF$logpop = log(chinaDF$pop+1)
+chinaDF$popLow = pmin(0,chinaDF$logpop-1)
+chinaDF$popHigh = pmax(0,chinaDF$logpop-1)
+chinaDF$popVeryHigh = pmax(0,chinaDF$logpop-4)
+chinaDF$latLow = pmin(0, chinaDF$y-45)
+chinaDF$latHigh = pmax(0, chinaDF$y-45)
 @
@@
 <<>>=
-bcSummary = summaryGmrfFit(bcRes,bcForGmrf) 
-bcSummary$summary
+save(chinaDF,china2,file="chinaSo2.RData")
 @
 
+<<>>=
+#plot(gam(y ~  s(lognight) + s(logpop,k=7,fx=T),data=bcAsiaDF))
+@
 \end{document}

### 2015-03-27 pkg/gmrf/vignettesStuff/v1.0.rnw
waiting for cholmod_solve2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1094 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{caption,float,amsmath,filecontents,subfig}

<<setup,include=FALSE>>=
library('knitr')

chapter=''
opts_chunk$set(echo=TRUE,
  	dev='pdf', marginsp=1,
		fig.align='',fig.pos='H',
		fig.width=5,fig.height=4,out.width='\\textwidth',
		tidy=TRUE,tidy.opts=list(reindent.spaces=2,
				width.cutoff=45)
)

knit_hooks$set(
		marginsp = function(before, options, envir){	
			if(!before) return()
# use small margins				
			par(mar=c(1.5+0.9*options$marginsp,
							1.5+0.9*options$marginsp,0.2,0.2),
					mgp=c(1.45, 0.45, 0),cex=1.25)
		}		
)

options(width=55)

@

\begin{document}
Make 10 raster datas, fit lgmrfm function
\begin{enumerate}

\item The base raster
<<Model1, cache=TRUE>>=
themodel = c(range=4*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

@
<<Sim_data1, cache=TRUE,echo=FALSE>>=
set.seed(1234)
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=4*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.05),log(0.3),len=24))
Snugget =  seq((0.01), (0.1),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)

@

Plot the result with and without Edge corrections 
<<plot1_mat,out.width='0.45\\textwidth',fig.cap='Model1 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot1_gmrf,out.width='0.45\\textwidth',fig.cap='Model1 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@

\item A bigger range
<<Model2, cache=TRUE>>=
themodel = c(range=8*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

@

 <<Sim_data2, cache=TRUE, echo=FALSE>>=
set.seed(1234)
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=8*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.01),log(0.07),len=24))
Snugget =  seq((0.1), (0.4),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)
 @

Plot the result with and without Edge corrections 
<<plot2_mat,out.width='0.45\\textwidth',fig.cap='Model2 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot2_gmrf,out.width='0.45\\textwidth',fig.cap='Model2 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@

\item A smaller variance
<<Model3, cache=TRUE>>=
themodel = c(range=4*xres(myraster),shape=1,variance=9,cellSize = xres(myraster))

@

 <<Sim_data3, cache=TRUE, echo=FALSE, warning=FALSE>>=
set.seed(1234)
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=4*xres(myraster),shape=1,variance=9,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.05),log(0.25),len=24))
Snugget =  seq((0.025), (0.1),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)
 @

Plot the result with and without Edge corrections 
<<plot3_mat,out.width='0.45\\textwidth',fig.cap='Model3 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot3_gmrf,out.width='0.45\\textwidth',fig.cap='Model3 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@
\item A different cellSize
<<Sim_data4, cache=TRUE, echo=FALSE, warning=FALSE>>=
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=20,nrows=15))

themodel = c(range=4*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.05),log(0.25),len=24))
Snugget =  seq((0.025), (0.1),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)
 @

Plot the result with and without Edge corrections 
<<plot4_mat,out.width='0.45\\textwidth',fig.cap='Model4 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot4_gmrf,out.width='0.45\\textwidth',fig.cap='Model4 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@
\end{enumerate}
\end{document}

### 2015-08-25 pkg/gmrf/vignettesStuff/so2.rnw
should work with covariates having elements which aren't in the model
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1220 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 worldShapeFile =  system(paste("find ", dataDir, 
-				" -name *countries*.shp", sep=""),TRUE)
+				" -name *countries*.shp", sep=""),TRUE)[1]
 worldShapeFile = gsub("\\.shp$","",worldShapeFile)

### 2017-02-23 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1648 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

```{r setup, include=FALSE}

if(any(commandArgs()=='mdToTex')) {
# if you'll be converting md to tex, can use the following
  knitr::knit_hooks$set(plot=knitr::hook_plot_tex)  
  knitr::opts_chunk$set(
   out.width='0.45\\textwidth', 
			dev='pdf', fig.align='center')
} else {
  knitr::knit_hooks$set(plot=Pmisc::hook_plot_mdsubfig)
  knitr::opts_chunk$set(dev='png', out.width='30%')
}

knitr::knit_hooks$set(
	margins = Pmisc::hook_plot_margins
)

knitr::opts_chunk$set(
  echo=FALSE, prompt=TRUE, comment=NA,
		margins=1, fig.cap=' ',
		fig.width=2.75, fig.height=2.75,  dpi=250, 
		tidy=TRUE, tidy.opts=list(indent=2, width.cutoff=55)
)

options(width=80)

bibFile = 'myDocBib.bib'
cat('
@article{lindgren2011explicit,
  title={An explicit link between {G}aussian fields and {G}aussian {M}arkov random fields the stochastic partial differential equation approach},
  author={Lindgren, Finn and Rue, Haavard and Lindstrom, Johan},
  journal={Journal of the Royal Statistical Society B},
  volume={73},
  number={4},
  pages={423-498},
  year={2011},
  publisher={Wiley}
}
', file=bibFile, encoding='UTF8')

```

```{r theHeader, echo=FALSE}
Pmisc::markdownHeader(
	title="Example of an html document with subfigures", 		
	author='Patrick Brown',
	slides=FALSE,
	bibliography = bibFile,
	geometry='margin=1in',
#	csl = system.file('src','apa-no-doi-no-issue.csl', package='Pmisc'),
	css = system.file('src','article.css', package='Pmisc')
)
```

# Introduction

@lindgren2011explicit is a very good paper.

$$
f(x) = x + \int_0^\infty \exp(-\theta u^2) du
$$

```{r aplot, fig.ncol=2, fig.cap='Uninteresting plots', fig.subcap=c('stuff','qq','eeee','www')}
plot(1:10)

plot(1:10+2)

plot(1:10-4)

plot(1:10+4)
```

# References

### 2017-02-23 pkg/Pmisc/inst/docs/slides.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1648 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

```{r setup, include=FALSE}
fName = 'slides.Rmd'
Makefile = system.file('src','knitrMakefile', package='Pmisc')
toCompile = paste("make -f ", Makefile, ' ', gsub("Rmd$","pdf", fName), sep='')
toCompile

# packages
library('knitr')
library('Pmisc')

# hook for multicolumn plots
knitr::knit_hooks$set(plot=hook_plot_beamer)

# hook for smaller plot margins
knitr::knit_hooks$set(
	marginsp = function(before, options, envir){	
		if(!before) return()
		par(mar=c(1.5+0.9*options$marginsp,
				1.5+0.9*options$marginsp,0.2,0.2),
				mgp=c(1.45, 0.45, 0),cex=1.25)
	}		
)

# hook for half width
knitr::opts_hooks$set(half = function(options) {
  options$tidy.opts = list(width.cutoff = options$tidy.opts$width.cutoff/2)
  options
})

# default chunk options
knitr::opts_chunk$set(echo=FALSE, prompt=TRUE, comment=NA,
		dev='pdf', marginsp=1, fig.start='#',
		fig.width=8, fig.height=4.5, half=NULL,
		tidy=TRUE,tidy.opts=list(indent=2, width.cutoff=55)
)

options(width=80)

# write a bibliography
knitr::write_bib(
	c('geostatsp', 'knitr'),
	file='myBib.bib'
)
```

```{r theHeader, echo=FALSE}
Pmisc::markdownHeader(
	title="Pmisc slides example", 		
	author='Patrick Brown',
 	bibliography='myBib.bib',
	theme='Singapore',
	beamer=TRUE,
	classoption=c(
  		aspectratio=169,
		xcolor='dvipsnames'),
	biblatexoptions = c(
		maxbibnames=20,
		maxcitenames=2,doi='false'),
	headerincludes =c(
  		'\\usecolortheme[named=MidnightBlue]{structure}',
  		'\\DeclareMathOperator{\\var}{var}',
		'\\DeclareMathOperator{\\E}{E}')
)
```

```{r theHeaderAgain, eval=FALSE, include=FALSE}
cat(
		Pmisc::markdownHeader(
				title="Pmisc slides example", 		
				author='Patrick Brown',
				bibliography='myBib.bib',
				theme='Singapore',
				beamer=TRUE,
				classoption=c(
						aspectratio=169,
						xcolor='dvipsnames'),
				biblatexoptions = c(
						maxbibnames=20,
						maxcitenames=2,doi='false'),
				headerincludes =c(
						'\\usecolortheme[named=MidnightBlue]{structure}',
						'\\DeclareMathOperator{\\var}{var}',
						'\\DeclareMathOperator{\\E}{E}')
		)
)
```

# Some references

- `@R-knitr`: @R-knitr
- `[@R-knitr]`: @R-knitr
- `[see @R-geostatsp for example]`: [see @R-geostatsp for example]
- `\fullcite{R-geostatsp}`: \fullcite{R-geostatsp}

# Markdown citations

```{r aplot, fig.height=3, fig.width=8, fig.ncol=2, fig.cap='A few plots', fig.subcap=c('a','b','c','d'), echo=FALSE, fig.start='#'}

plot(1:10)

plot(1:10+2)

plot(1:10-4)

plot(1:10+4)

```

# A few more plots

\bcolb{Code}

```{r twoplots, echo=TRUE, eval=FALSE, half=TRUE}
plot(rnorm(10), xlab='index', ylab='value', pch=1:10)

plot(runif(10), xlab='index', ylab='value', pch=1:10)
```

\newcolb{Output}

```{r twoplots, fig.height=3, fig.width=8, fig.subcap=c('normal','uniform'), echo=FALSE, fig.start=NULL}
```

\ecolb

### 2017-02-23 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1649 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 			dev='pdf', fig.align='center')
+  mdToTex = TRUE
 } else {
   knitr::knit_hooks$set(plot=Pmisc::hook_plot_mdsubfig)
-  knitr::opts_chunk$set(dev='png', out.width='30%')
+  knitr::opts_chunk$set(dev='png', out.width='40%')
+  mdToTex = FALSE
 }
@@
 
-@lindgren2011explicit is a very good paper.
+@lindgren2011explicit is a very good paper, but Figure [@fig:aplot] is dull.
 
 $$
-f(x) = x + \int_0^\infty \exp(-\theta u^2) du
+\begin{aligned}
+f(x) = &x + \int_0^\infty \exp(-\theta u^2) du\\
+\rho(x) = & \sum_i x^i
+\end{aligned}
 $$
@@
 
+```{r anotherplot, fig.ncol=3, fig.cap='Other uninteresting plots', fig.subcap=month.name[1:6], out.width=c('30%', '0.3\\textwidth')[1+mdToTex]}
+
+for(D in knitr::opts_current$get()$fig.subcap) {
+  hist(runif(100))
+}
+
+```
+
 # References

### 2017-02-23 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1650 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 }
+```
+
+Here is a table
+
+```{r theTable}
+
+mytable = as.data.frame(matrix(runif(6),,3))
+names(mytable) = month.name[1:ncol(mytable)]
+
+tableFun = function(...) {
+  if(mdToTex) {
+    Hmisc::latex(...)
+  } else {
+    Hmisc::html(...)
+  }
+}
+tableFun(mytable, digits=3, file='',
+  caption = 'the table', caption.loc='bottom',
+  label = 'tbl:theTable'
+    )
 

### 2017-02-24 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1651 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	geometry='margin=1in',
-#	csl = system.file('src','apa-no-doi-no-issue.csl', package='Pmisc'),
+	csl = system.file('src','apa-no-doi-no-issue.csl', package='Pmisc'),
 	css = system.file('src','article.css', package='Pmisc')
@@
 
-Here is a table
+Below one will find Table [@tbl:theTable]
 
@@
 
-tableFun = function(...) {
-  if(mdToTex) {
-    Hmisc::latex(...)
-  } else {
-    Hmisc::html(...)
-  }
-}
-tableFun(mytable, digits=3, file='',
-  caption = 'the table', caption.loc='bottom',
-  label = 'tbl:theTable'
-    )
+mdTable(x=mytable, digits=3,
+  caption = 'the table', caption.loc='bottom')
 

### 2017-02-27 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1655 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   echo=FALSE, prompt=TRUE, comment=NA,
-		margins=1, fig.cap=' ',
+		margins=1, fig.cap=' ', fig.pos='ht',
 		fig.width=2.75, fig.height=2.75,  dpi=250, 
@@
 
-mdTable(x=mytable, digits=3,
+Pmisc::mdTable(x=mytable, digits=3,
   caption = 'the table', caption.loc='bottom')

### 2017-02-27 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1658 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2017-02-24 pkg/Pmisc/inst/docs/document.Rmd
@@
   echo=FALSE, prompt=TRUE, comment=NA,
-		margins=1, fig.cap=' ',
-		fig.width=2.75, fig.height=2.75,  dpi=250, 
+		margins=1, fig.cap=' ', fig.pos='ht',
+		fig.width=2.75, fig.height=2,  dpi=250, 
 		tidy=TRUE, tidy.opts=list(indent=2, width.cutoff=55)
@@
 Pmisc::markdownHeader(
-	title="Example of an html document with subfigures", 		
+	title="Example of a simple R markdown document with subfigures", 		
 	author='Patrick Brown',
-	slides=FALSE,
 	bibliography = bibFile,
@@
 for(D in knitr::opts_current$get()$fig.subcap) {
-  hist(runif(100))
+  hist(runif(100), main='')
 }
@@
 
-mdTable(x=mytable, digits=3,
-  caption = 'the table', caption.loc='bottom')
+Pmisc::mdTable(x=mytable, digits=3,
+  caption = 'the table', caption.loc='bottom', 
+  where = 'htb')
 

### 2017-05-10 docs/survival/weibullcure.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1717 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[a4paper,11pt]{article}
\usepackage[scale={0.8,0.9},centering,includeheadfoot]{geometry}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{verbatim}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\begin{document}
\section*{Weibull With Cure Fraction}

\subsection*{Background}

\begin{align*}
Z_i & \sim \text{Bernoulli}(\rho)\\
[Y_i | Z_i = 0] &\sim \text{Weibull}(\lambda_i, \alpha)\\
[Y_i | Z_i = 1] & = \infty
\end{align*}

\subsection*{Parametrisation}

The Weibull is parametrized as \texttt{variant=0} of the \texttt{weibull} family.
\begin{displaymath}
    f(y) = 
    (1-\rho) \alpha y^{\alpha-1}
    \lambda\exp( - \lambda  y^{\alpha}), 
    \qquad  0 \leq y < \infty,
    \qquad \alpha>0, \qquad \lambda>0
\end{displaymath}

\begin{description}
\item[$\alpha$:] shape parameter.
\item[$\rho$:] the cure fraction parameter
\end{description}

\subsection*{Link-function}

The parameter $\lambda$ is linked to the linear predictor as:
\[
    \lambda = \exp(\eta)
\]

\subsection*{Hyperparameters}

The $\alpha$ parameter is represented as
\[
    \theta_1 = \log\alpha
\]
and $\rho$ is transformed to
\[
\theta_2 = \log[\rho / (1-\rho)].
\]
The priors are defined on $\theta$.

\subsection*{Specification}

Response variable $y$ must be given using $\texttt{inla.surv()}$

\subsubsection*{Hyperparameter spesification and default values}

\begin{quote}
    \input{../hyper/likelihood/weibullcure.tex}
%    \input{weibullcureHyper.tex}
\end{quote}

\subsection*{Example}

In the following example we estimate the parameters in a simulated
case \verbatiminput{example-weibullcure.R}

\subsection*{Notes}

\begin{itemize}
\item Weibull model can be used for right censored, left censored,
    interval censored data. If the observed times $y$ are large/huge,
    then this can cause numerical overflow in the likelihood routine.
    If you encounter this problem, try to scale the observatios,
    \verb|time = time / max(time)| or similar.
\end{itemize}

\end{document}

% LocalWords:  np Hyperparameters Ntrials

%%% Local Variables: 
%%% TeX-master: t
%%% End:

### 2017-05-18 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1746 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1748 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1749 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   bibliography = bibFile,
-	geometry='margin=1in',
-	csl = system.file('src','apa-no-doi-no-issue.csl', package='Pmisc'),
-	css = system.file('src','article.css', package='Pmisc')
+  geometry='margin=1in'
 )
@@
 
-# Introduction
+# Introduction {#sec:intro}
 
 @lindgren2011explicit is a very good paper, but Figure [@fig:aplot] is dull.
+There are some formulas in this document, from ([@eq:myeqn]) we can see
+nothing useful.  If you are unsure of which Section we are in, be aware that
+this is Section [@sec:intro].
 
@@
 f(x) = &x + \int_0^\infty \exp(-\theta u^2) du\\
-\rho(x) = & \sum_i x^i
+\rho(x_1 \ldots x_N) = & \sum_i x^i
 \end{aligned}
-$$
+$$ {#eq:myeqn}
+
+Now some plots.
 

### 2017-05-18 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1751 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 )
+
+# consider adding
+automatic-styles = '<style:style style:name="OLE" style:family="graphic"><style:graphic-properties text:anchor-type="paragraph" style:wrap="none"/></style:style>'
+# to make equations better in odt document
 ```

### 2017-07-07 pkg/Pmisc/inst/docs/slides.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1770 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 # hook for multicolumn plots
-knitr::knit_hooks$set(plot=hook_plot_beamer)
+knitr::knit_hooks$set(Pmisc::plot=hook_plot_beamer)
 
@@
 knitr::knit_hooks$set(
-	marginsp = function(before, options, envir){	
-		if(!before) return()
-		par(mar=c(1.5+0.9*options$marginsp,
-				1.5+0.9*options$marginsp,0.2,0.2),
-				mgp=c(1.45, 0.45, 0),cex=1.25)
-	}		
+	margins = Pmisc::hook_plot_margins
 )
@@
 knitr::opts_chunk$set(echo=FALSE, prompt=TRUE, comment=NA,
-		dev='pdf', marginsp=1, fig.start='#',
+		dev='pdf', margins=1, fig.start='#',
 		fig.width=8, fig.height=4.5, half=NULL,

### 2017-09-19 pkg/mapmisc/inst/extdoc/world.Rmd
out.width function created
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1856 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

```{r theHeader, echo=FALSE}
knitr::knit_hooks$set(plot=knitr::hook_plot_tex) 
Pmisc::markdownHeader(
  title="Maps of the world", 		
  author='Patrick Brown',
  geometry='margin=1in'
)

```

```{r worldCrs, tidy=TRUE, warning=FALSE}
require('mapmisc', quietly=TRUE) 
options(mapmiscCachePath = system.file('extdata', package='mapmisc'))

worldLL = openmap(extent(-180,180,-80,80), zoom=1, crs=mapmisc::crsLL)

Scentre = c('Grenwich'=0, Toronto=-79, 'Hong kong' = 114)

mollList = mapply(
    mapmisc::moll,     x= Scentre     )

worldMoll = mapply(     mapmisc::openmap,     x=mollList,     MoreArgs = list(zoom=1,
fact=2)     )

```

```{r worldMaps, fig.width=8, fig.height=3, fig.cap='World maps in different projections', fig.subcap = c('Long-lat', paste('Mollweide', names(mollList))), dev='png', out.width='0.8\\textwidth', warning=FALSE, fig.align='default'}
map.new(worldLL)
plot(worldLL, add=TRUE)

for(D in names(worldMoll)) {
  map.new(mollList[[D]])
  plot(worldMoll[[D]], add=TRUE)
}

```

### 2018-02-05 pkg/mapmisc/inst/extdoc/world.Rmd
hopefully all the caching stuff works now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1903 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 )
-
 ```
@@
 require('mapmisc', quietly=TRUE)
-options(mapmiscCachePath = system.file('extdata', package='mapmisc'))
 

### 2018-03-29 pkg/mapmisc/inst/extdoc/mapLayers.Rmd
mapLayers added back
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1963 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Examples of map layers}
-->

# Examples of map layers

This document will be incomplete if `rgdal` is unavailable or there is on
internet connection when this document is compiled.  The full document is at
[diseasemapping.r-forge.r-project.org](http://diseasemapping.r-forge.r-project.org).

See [mc.bbbike.org/mc](http://mc.bbbike.org/mc/) for examples more map tiles 
(not all of which are compatible with `mapmisc`).

```{r setup, include=FALSE}
library('knitr')
library('mapmisc')

fact = 0.5
allTiles = FALSE

# uncomment for nicer images
# fact = 1 
# reRun = TRUE
# or run in sandbox	
# R --slave -e "knitr::knit('../pkg/mapmisc/vignettes/mapLayers.Rmd')" fact=1 allTiles=TRUE
# pandoc --to=html5 --output=mapLayers.html mapLayers.md
# cp -r figure ../www
# cp -r mapLayers.html ../www
theArgs = grep("^(fact|allTiles)=", commandArgs(), value=TRUE)
for(D in theArgs)
  eval(parse(text=D))

opts_chunk$set(dpi=50*fact, 
  dev.args=list(bg='white'))

if(allTiles) {
  theNames = names(osmTiles())
  cachePath = file.path(tempdir(), 'mapLayersVignetteCache')
} else {
  # smaller selection of tiles, which are distributed with the package
  theNames = names(grep(
      "cyclemap|thunderforest|waze|wikimedia|korona|opentopomap|maptoolkit|nrcan", 
    osmTiles(), invert=TRUE, value=TRUE))
  cachePath = system.file('extdata', package='mapmisc')
}

```

```{r theExtent, include=FALSE}
africaExtent = c(25, 25)
trafalgarSquare=c(-0.128056,51.508056)
stjohns=c(-52.683320, 47.566285)
```

```{r downloadTiles,include=FALSE}

  africaTiles = openmap(africaExtent, path=theNames,zoom=2,
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath)

  londonTiles = openmap(trafalgarSquare,zoom=12,path=theNames,
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath)

  nfTiles = openmap(stjohns, zoom=14, path=theNames,
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath)

```

The following are some examples of maps produced with `mymap=openmap(x, path="something")`

```{r someTiles, echo=FALSE, results='asis', fig.height=5, fig.width=5}

mytiles = list(africaTiles, londonTiles, nfTiles)

for(D in theNames){
  Dsub = gsub("-",".", D)

  threeTiles = length(
    grep(
      paste("^", Dsub, "([rR]ed|[gG]reen|[bB]lue)$", sep=""), 
      names(mytiles[[1]])
    )
  ) == 3

  cat("\n\n# ", D, "\n\n")
  cat("`plot", "RGB"[threeTiles], 
    "(openmap(x,path=\"", D, "\"))`\n\n", 
    sep='')
  cat(openmapAttribution(D, 'html'), '\n\n')

  cat("\n|     |     |     |\n|-----|-----|-----|\n")
  for(Dplot in 1:length(mytiles)) {
    cat("|")

    oneTileName = grep(paste("^", Dsub, "$", sep=""), 
      names(mytiles[[Dplot]]), value=TRUE)
    threeTilesName = grep(paste("^", Dsub, 
        "([rR]ed|[gG]reen|[bB]lue)$", sep=""), 
      names(mytiles[[Dplot]]),value=TRUE)

    if(length(threeTilesName)==3) { # plotRGB
      map.new(mytiles[[Dplot]])

      plotRGB(mytiles[[Dplot]][[
          paste(Dsub, c("Red","Green","Blue"),sep="")
        ]],add=TRUE)

    } else if (length(oneTileName)==1){

      map.new(mytiles[[Dplot]])
      plot(mytiles[[Dplot]][[oneTileName]],add=TRUE)

    } else {

      plot(0.5, 0.8)
      text(0.5, 0.8, paste(D, "is\n not currently\navailable"))

    }

  }
  cat("|")

}

```

# Additional map layers

See [mc.bbbike.org/mc](http://mc.bbbike.org/mc/)

Michelin

```{r additionalMichelin, echo=TRUE, purl=FALSE, eval=unlist(mget("allTiles", ifnotfound=FALSE))}

mPath = osmTiles(paste(
  'http://map3.viamichelin.com/map/mapdirect?map=',
  c('light','viamichelin'), '&', sep=''),
  xyz='zxy=', suffix = 
    '&format=png&version=201503191157&layer=background')

africaMichelin = openmap(
  africaExtent, path=mPath,
  zoom=2,
  verbose=TRUE, 
  crs = crsMerc)

for(D in names(africaMichelin)) {
  map.new(africaMichelin)
  plot(africaMichelin[[D]], add=TRUE)
  mtext(D, side=3, line=-2, outer=FALSE)
}

```
Stamen retina

```{r stamenRetina, echo=TRUE, purl=FALSE, eval=unlist(mget("allTiles", ifnotfound=FALSE))}

sPath = osmTiles('stamen-toner', 
  suffix = '@2x.png')

africaS = openmap(
  africaTiles, 
  buffer=-1,
  zoom=attributes(africaTiles)$tiles$zoom+1,
  path=sPath, verbose=TRUE)
africaS2 = tonerToTrans(africaS)

map.new(africaTiles)
plot(africaTiles[['opentopomap']], add=TRUE)
plot(africaS2, add=TRUE)

mtext(sPath, side=3, line=-2, outer=FALSE)
```

Komoot, Yandex, ESRI, F4map, and Sigma

```{r aditionalEsri, echo=TRUE, purl=FALSE, eval=unlist(mget("allTiles", ifnotfound=FALSE))}

otherPaths = c(
  'https://a.tile.hosted.thunderforest.com/komoot-2/',
  'https://sat01.maps.yandex.net/tiles?l=sat&v=1.35.0&',
  'https://vec02.maps.yandex.net/tiles?l=map&v=4.40&',
  'https://server.arcgisonline.com/ArcGIS/rest/services/World_Physical_Map/MapServer/tile/',
  'https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/',
  'https://services.arcgisonline.com/ArcGIS/rest/services/Ocean/World_Ocean_Base/MapServer/tile/',
  'https://tile2.f4map.com/tiles/f4_2d/',
  'http://tiles1.sigma-dc-control.com/layer8/')

africaExtra = openmap(
  africaExtent, path=otherPaths,
  zoom=2,
  verbose=TRUE, 
  crs = crsMerc)

Spaths = unique(
  gsub("(red|green|blue|trans)$", "",
    names(africaExtra), ignore.case=TRUE)
)

for(D in Spaths) {

  oneTileName = grep(paste("^", D, "$", sep=""), 
    names(africaExtra), value=TRUE)

  threeTilesName = grep(paste("^", D, 
      "([rR]ed|[gG]reen|[bB]lue)$", sep=""), 
    names(africaExtra),value=TRUE)

  map.new(africaExtra)
  if(length(threeTilesName)==3) {
    plotRGB(africaExtra[[threeTilesName]], add=TRUE)
  } else {
    plot(africaExtra[[oneTileName]], add=TRUE)
  }
  mtext(D, side=3, line=-2, outer=FALSE)
}

```

### 2018-03-29 pkg/mapmisc/inst/extdoc/mapLayers.Rmd
fixing the caching, should solve cran errors
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1964 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-<!--
-%\VignetteEngine{knitr::knitr}
-%\VignetteIndexEntry{Examples of map layers}
--->
+```{r theHeader, echo=FALSE, result='asis'}
+if(requireNamespace("Pmisc", quietly=TRUE)) {
+  Pmisc::markdownHeader(
+    title="Examples of map layers in the mapmisc package",    
+    author='Patrick Brown'
+    )
+} else {
+cat("# Examples of map layers in the mapmisc package")  
+}
+```
 
-# Examples of map layers
-
-This document will be incomplete if `rgdal` is unavailable or there is on
-internet connection when this document is compiled.  The full document is at
+The full document is at
 [diseasemapping.r-forge.r-project.org](http://diseasemapping.r-forge.r-project.org).
@@
 
-fact = 0.5
+fact = 2
 allTiles = FALSE
 
-# uncomment for nicer images
-# fact = 1 
-# reRun = TRUE
-# or run in sandbox	
-# R --slave -e "knitr::knit('../pkg/mapmisc/vignettes/mapLayers.Rmd')" fact=1 allTiles=TRUE
-# pandoc --to=html5 --output=mapLayers.html mapLayers.md
-# cp -r figure ../www
-# cp -r mapLayers.html ../www
-theArgs = grep("^(fact|allTiles)=", commandArgs(), value=TRUE)
-for(D in theArgs)
-  eval(parse(text=D))
+knitr::opts_chunk$set(dpi=150, dev.args=list(bg='white'),
+    fig.path = 'mapLayersFigures/', out.width='30%')
 
-opts_chunk$set(dpi=50*fact, 
-  dev.args=list(bg='white'))
-
-if(allTiles) {
 theNames = names(osmTiles())
-  cachePath = file.path(tempdir(), 'mapLayersVignetteCache')
-} else {
-  # smaller selection of tiles, which are distributed with the package
-  theNames = names(grep(
-      "cyclemap|thunderforest|waze|wikimedia|korona|opentopomap|maptoolkit|nrcan", 
-    osmTiles(), invert=TRUE, value=TRUE))
-  cachePath = system.file('extdata', package='mapmisc')
-}
+cachePath = file.path(getwd(), 'mapLayersCache')
 
@@
     cachePath = cachePath)
-
 ```
@@
 
+Maps shown because it is not clear how they should be attributed.
+
 Michelin
 
-```{r additionalMichelin, echo=TRUE, purl=FALSE, eval=unlist(mget("allTiles", ifnotfound=FALSE))}
+```{r additionalMichelin, echo=TRUE, purl=FALSE, eval=FALSE}
 
@@
 
-```{r stamenRetina, echo=TRUE, purl=FALSE, eval=unlist(mget("allTiles", ifnotfound=FALSE))}
+```{r stamenRetina, echo=TRUE, purl=FALSE, eval=FALSE}
 
@@
 
-```{r aditionalEsri, echo=TRUE, purl=FALSE, eval=unlist(mget("allTiles", ifnotfound=FALSE))}
+```{r aditionalEsri, echo=TRUE, purl=FALSE, eval=FALSE}
 

### 2018-04-09 pkg/mapmisc/inst/extdoc/perspective.Rmd
moved perspective to extdoc
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1978 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<!--
%\VignetteEngine{knitr::docco_linear}
%\VignetteIndexEntry{Space Station Map Projections}
-->

# Space Station Map Projections

```{r setup}
library('mapmisc', quietly=TRUE)
haveRgdal = requireNamespace('rgdal', quietly=TRUE)
haveGeosphere = requireNamespace('geosphere', quietly=TRUE)

if(!exists('mapmiscCachePath'))
  mapmiscCachePath = system.file('extdata', package='mapmisc')

if(!exists('mapmiscCacheReadOnly'))
  mapmiscCacheReadOnly = TRUE

if(!exists('testing'))
  testing=FALSE

if(!exists('fact'))
  fact = 0.6
# uncomment for nicer images
# fact = 1

options(
  mapmiscCachePath = mapmiscCachePath,
  mapmiscCacheReadOnly = mapmiscCacheReadOnly
  )
```

```{r mapmiscCache, echo=FALSE, purl=FALSE}
  getOption('mapmiscCachePath')
  getOption('mapmiscCacheReadOnly')
if(testing) {
  options(mapmiscVerbose = TRUE)
}
```

```{r data}
okinawa = mapmisc::geocode("Okinawa")
hk = y = mapmisc::geocode("Hong Kong")
guam = mapmisc::geocode("Guam")
x = SpatialPoints(cbind(130,15), proj4string=crsLL)
```

```{r projection}
if(haveGeosphere & haveRgdal) {
  myProj = tpers(x=bind(x,y), hKm = 2*1000,
      tilt =-15, axis='seu')
  xProj = spTransform(guam, myProj)
  yProj = spTransform(hk, myProj)
  okinawaProj = spTransform(okinawa, myProj)
}
```

```{r background}
if(haveGeosphere & haveRgdal)
  myMap = openmap(myProj, path='landscape', zoom=2,fact=3)
```

```{r backgroundOverDateline, include=FALSE, eval=FALSE, purl=FALSE}
if(length(attributes(myProj)$ellipseSafeLL)>2 & haveRgdal) {

  myMap1 = openmap(
      x=attributes(myProj)$regionLL[1,],
      crs=crsMerc, zoom=3,
      path='https://sat01.maps.yandex.net/tiles?l=sat&v=1.35.0&')

  myMap2 = openmap(
      x=attributes(myProj)$regionLL[2,],
      zoom=attributes(myMap1)$tiles$zoom,
      crs=crsMerc, zoom=3,
      path='https://sat01.maps.yandex.net/tiles?l=sat&v=1.35.0&')

  mapRaster = raster(
      extent(attributes(myProj)$ellipse), 
      res = mean(res(myMap1)/1.5),
      crs=myProj
  )

  myMap1a = projectRaster(
      from=myMap1, to=mapRaster, method='ngb')
  myMap2a = projectRaster(
      from=myMap2, to=mapRaster, method='ngb')

  myMap = mosaic(myMap1a, myMap2a, fun=mean)
}
```

```{r map}
if(haveGeosphere & haveRgdal) {
  map.new(myProj, bg='black', buffer=2*100*1000)
  plot(myMap, add=TRUE)

  if(requireNamespace('maptools')) {
    data("wrld_simpl", package='maptools')
    wrld4 = wrapPoly(x=wrld_simpl, crs=myProj)
    plot(wrld4, add=TRUE)
  }

  gridlinesWrap(myProj, col='yellow', 
      north=seq(-40,40,by=10), easts=seq(-150,180,by=10),
      lty=2)

  points(xProj, col='red', pch=16, cex=2)
  text(xProj@coords, 
      as.character(xProj$originalPlace),
      pos=2, col='red')

  points(okinawaProj, col='red', pch=16, cex=2)
  text(okinawaProj@coords, 
      as.character(okinawaProj$originalPlace),
      pos=2, col='red')

  points(yProj, col='red', pch=16, cex=2)
  text(yProj@coords, 
      as.character(yProj$originalPlace),
      pos=2, col='red')

}
```

### 2019-02-12 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1752 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
subfigures fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2046 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   knitr::knit_hooks$set(plot=Pmisc::hook_plot_mdsubfig)
-  knitr::opts_chunk$set(dev='png', out.width='40%')
+  knitr::opts_chunk$set(dev='png')
   mdToTex = FALSE
@@
   margins=1, fig.cap=' ', fig.pos='ht',
-  fig.width=2.75, fig.height=2,  dpi=250, 
+  fig.width=5.5, fig.height=4, 
   tidy=TRUE, tidy.opts=list(indent=2, width.cutoff=55)
@@
 
-# consider adding
-automatic-styles = '<style:style style:name="OLE" style:family="graphic"><style:graphic-properties text:anchor-type="paragraph" style:wrap="none"/></style:style>'
-# to make equations better in odt document
 ```
@@
 
-```{r anotherplot, fig.ncol=3, fig.cap='Other uninteresting plots', fig.subcap=month.name[1:6], out.width=c('30%', '0.3\\textwidth')[1+mdToTex]}
+```{r anotherplot, fig.ncol=3, fig.cap='Other uninteresting plots', fig.subcap=month.name[1:6], out.width=Pmisc::out.width(0.32)}
 

### 2021-07-21 pkg/gpuRandom/R/CreateStreamsGpu.R
r documentation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2759 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -28,7 +28,7 @@ CreateStreamsGpu = function(seedR=c(12345, 12345, 12345, 12345, 12345, 12345),
   
   
     myseed <- gpuR::vclVector(as.integer(seedR), type="integer")  
-    streamsR<-vclMatrix(0L, nrow=Nstreams, ncol=18, type="integer")
+    streamsR<-gpuR::vclMatrix(0L, nrow=Nstreams, ncol=18, type="integer")
     
     gpuRandom:::CreateStreamsGpuBackend(myseed, streamsR, keepInitial)
   

### 2021-07-21 pkg/gpuRandom/R/runif.R
r documentation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2759 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,4 +1,12 @@
 #' @title runif
+#' 
+#' @param n a number or a vector specifying the size of output matrix
+#' @param streams streams object
+#' @param Nglobal the size of the index space for use
+#' @param type "double" or "float" format of generated random numbers
+#' 
+#' @return a vclVector or vclMatrix of uniform random numbers
+#' 
 #' @useDynLib gpuRandom
 #' @export
 

### 2021-08-18 pkg/gpuRandom/src/maternBatch.cpp
matern
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2761 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -51,11 +51,7 @@ std::string maternBatchKernelString(
   
   std::string typeString = openclTypeString<T>();
   std::string result = "";
-  
-  if(NmatrixMax * NlocalParams > NlocalParamsCache) {
-    Rcpp::Rcout << "warning creating matern kernel: need to store " << NmatrixMax << 
-      " sets of " << NlocalParams<< " parameters, cache of " << NlocalParamsCache <<" is insufficient\n";
-    }
+
     
   if(typeString == "double") {
     result += "\n#pragma OPENCL EXTENSION cl_khr_fp64 : enable\n\n"
@@ -253,13 +249,29 @@ std::string maternBatchKernelString(
   result += "event_t wait;\n\n";
   
   result += 
-    // dimension 0 is cell, dimension 1 is matrix
+    "// dimension 0 is cell, dimension 1 is matrix\n"
     "const int isFirstLocal = (get_local_id(0)==0 & get_local_id(1)==0);\n"
-    "const int isFirstLocal1 = (get_local_id(1)==0);\n";
+    "const int isFirstLocal1 = (get_local_id(1)==0);\n\n";
     
-    // copy parameters to local storage
-    result += "wait = (event_t) 0;\n";
-  result += //Dcell is really DmatrixBlock
+    result +=
+      "// copy parameters to local storage\n";
+  result += 
+    "if(get_local_id(0)==0){\n"
+    " for(DmatrixLocal = get_local_id(1),Dmatrix = get_global_id(1);\n"
+    "     Dmatrix < Nmatrix;\n"
+    "     DmatrixLocal+= get_local_size(1), Dmatrix += get_global_size(1)){\n"
+    "   for(Dcell=0;Dcell < NlocalParams;Dcell++){\n"
+    "     localParams[DmatrixLocal * NlocalParams + Dcell]=\n"
+    "        params[(startrow + Dmatrix) * NpadParams + Dcell];\n"  
+    "  }//Dcell\n"
+    " }//Dmatrix\n"
+    "}//if\n";
+result +=  "barrier(CLK_LOCAL_MEM_FENCE);\n\n";
+  
+// old stuff
+#ifdef UNDEF  
+  result += 
+    "//Dcell is really DmatrixLocal\n"
     "for(Dcell = 0,Dmatrix = get_group_id(1)*get_local_size(1);\n"
[…]

### 2021-08-19 pkg/geostatsp/R/param.R
fill param bug fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2762 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -65,7 +65,7 @@ model
 
 fillParam = function(param) {
 
-	if(is.numeric(param))
+	if(is.vector(param))
 		param = matrix(param, ncol=length(param), nrow=1,
 				dimnames=list("1", names(param)))
 
@@ -137,11 +137,12 @@ fillParam = function(param) {
 	otherParams = setdiff(colnames(param), theOrder)
 	param = param[,c(theOrder, otherParams)]
 
+	param[is.na(param[, 'range']), 'range'] = 1
+	param[is.na(param[, 'anisoRatio']), 'anisoRatio'] = 1
+	param[is.na(param[, 'anisoAngleRadians']), 'anisoAngleRadians'] = 1
+	param[is.na(param[, 'anisoAngleDegrees']), 'anisoAngleDegrees'] = 1
+	
 	param = drop(param)
-	if(is.na(param['range'])) param['range'] = 1
-	if(is.na(param['anisoRatio'])) param['anisoRatio'] = 1
-
-	if(is.na(param['anisoAngleDegrees'])) 
-		param['anisoAngleRadians'] = param['anisoAngleDegrees'] =0
+	
 	param
 }

### 2021-08-26 pkg/gpuRandom/R/crossprodBatch.R
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2768 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -9,14 +9,13 @@
 
  
 crossprodBatch <- function(C,   # must be a batch of square matrices 
-                      A,
-                      D,
-                      invertD,
-                      Nglobal, Nlocal, 
-                      NlocalCache,
-                      Cstartend,
-                      Astartend,
-                      Dstartend) {
+                           A,
+                           D,
+                           invertD,
+                           Nglobal, Nlocal, NlocalCache,
+                           Cstartend,
+                           Astartend,
+                           Dstartend) {
   
   Nbatches = nrow(C)/ncol(C)
   

### 2021-08-26 pkg/gpuRandom/src/likfitGpu_P.cpp

@@ -487,11 +487,6 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
   
   
   
-  
-  
-  
-  
-  
   viennacl::ocl::program & my_prog_matern = viennacl::ocl::current_context().add_program(maternClString, "mykernelmatern");
   viennacl::ocl::program & my_prog_chol = viennacl::ocl::current_context().add_program(cholClString, "mykernelchol");
   viennacl::ocl::program & my_prog_backsolve = viennacl::ocl::current_context().add_program(backsolveString, "mykernelbacksolve");
@@ -689,8 +684,11 @@ void likfitGpuP_Templated(
     Rcpp::IntegerVector localSize,
     Rcpp::IntegerVector NlocalCache,
     Rcpp::IntegerVector verbose,
-    Rcpp::S4 ssqYX,    Rcpp::S4 ssqYXcopy,
-    Rcpp::S4 LinvYX, Rcpp::S4 QinvSsqYx, Rcpp::S4 cholXVXdiag,
+    Rcpp::S4 ssqYX,    
+    Rcpp::S4 ssqYXcopy,
+    Rcpp::S4 LinvYX, 
+    Rcpp::S4 QinvSsqYx, 
+    Rcpp::S4 cholXVXdiag,
     Rcpp::S4 varMat,
     Rcpp::S4 cholDiagMat){  //Rcpp::S4 aTDinvb_beta,Rcpp::S4 aTDinvb_beta_diag
   
@@ -719,9 +717,6 @@ void likfitGpuP_Templated(
   //  std::shared_ptr<viennacl::matrix<T> > aTDinvb_beta_diaggpu = getVCLptr<T>(aTDinvb_beta_diag.slot("address"), BisVCL, ctx_id);
   
   
-  
-  
-  
   addBoxcoxToData<T>(
     *yxGpu,
     *boxcoxGpu,
@@ -731,6 +726,9 @@ void likfitGpuP_Templated(
     NlocalCache,
     ctx_id,
     verbose);
+  
+  
+  
   
   likfitGpuP<T>(
     *yxGpu, 
@@ -743,8 +741,17 @@ void likfitGpuP_Templated(
     NparamPerIter,
     workgroupSize, 
     localSize, 
[…]

### 2021-08-26 pkg/brnd.sh
copy
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2769 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,5 +1,5 @@
 R -e "Rcpp::compileAttributes('gpuRandom')"
-R -e "devtools::document('gpuRandom')"
+#R -e "devtools::document('gpuRandom')"
 R CMD build --no-build-vignettes gpuRandom
 R CMD INSTALL gpuRandom_0.1.tar.gz
 

### 2021-08-26 pkg/gpuRandom/src/RcppExports.cpp

@@ -65,6 +65,17 @@ BEGIN_RCPP
     return R_NilValue;
 END_RCPP
 }
+// copyToCpu
+void copyToCpu(Rcpp::NumericMatrix output, Rcpp::S4 input);
+RcppExport SEXP _gpuRandom_copyToCpu(SEXP outputSEXP, SEXP inputSEXP) {
+BEGIN_RCPP
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< Rcpp::NumericMatrix >::type output(outputSEXP);
+    Rcpp::traits::input_parameter< Rcpp::S4 >::type input(inputSEXP);
+    copyToCpu(output, input);
+    return R_NilValue;
+END_RCPP
+}
 // crossprodBatchBackend
 SEXP crossprodBatchBackend(Rcpp::S4 C, Rcpp::S4 A, Rcpp::S4 D, const int invertD, Rcpp::IntegerVector Cstartend, Rcpp::IntegerVector Astartend, Rcpp::IntegerVector Dstartend, Rcpp::IntegerVector Nglobal, Rcpp::IntegerVector Nlocal, const int NlocalCache);
 RcppExport SEXP _gpuRandom_crossprodBatchBackend(SEXP CSEXP, SEXP ASEXP, SEXP DSEXP, SEXP invertDSEXP, SEXP CstartendSEXP, SEXP AstartendSEXP, SEXP DstartendSEXP, SEXP NglobalSEXP, SEXP NlocalSEXP, SEXP NlocalCacheSEXP) {
@@ -406,6 +417,7 @@ static const R_CallMethodDef CallEntries[] = {
     {"_gpuRandom_backsolveBatchBackend", (DL_FUNC) &_gpuRandom_backsolveBatchBackend, 11},
     {"_gpuRandom_backsolveBatchBackend2", (DL_FUNC) &_gpuRandom_backsolveBatchBackend2, 7},
     {"_gpuRandom_cholBatchBackend", (DL_FUNC) &_gpuRandom_cholBatchBackend, 8},
+    {"_gpuRandom_copyToCpu", (DL_FUNC) &_gpuRandom_copyToCpu, 2},
     {"_gpuRandom_crossprodBatchBackend", (DL_FUNC) &_gpuRandom_crossprodBatchBackend, 10},
[…]

### 2021-08-26 pkg/gpuRandom/src/copy.cpp
copy
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2769 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,52 @@
+#include "gpuRandom.hpp"
+
+
+//#define DEBUGKERNEL
+
+// TO DO: use async_copy to copy by column and vector_iterator
+template<typename T> 
+void copyToCpuTemplated(
+    Rcpp::NumericMatrix output,
+    Rcpp::S4 input) {
+  
+  const bool BisVCL=1;
+  const int ctx_id = INTEGER(input.slot(".context_index"))[0]-1;
+  std::shared_ptr<viennacl::matrix_base<T> > xMat = getVCLptr<T>(input.slot("address"), BisVCL, ctx_id);
+
+  viennacl::vector_base<T> x((*xMat).handle(), (*xMat).internal_size(), 0, 1);
+  
+  
+  int Drow=0;
+  const int Nrow = (*xMat).size1(), Ncol = (*xMat).size2(), NcolInt = (*xMat).internal_size2();
+  for(Drow = 0; Drow < Nrow; Drow++) {
+    viennacl::fast_copy(
+      x.begin() + Drow*NcolInt,// + Drow*NcolInt, 
+      x.begin() + Drow*NcolInt + Ncol,// + Drow*NcolInt+Ncol, 
+      output.begin() +Drow*Ncol
+    );
+  }
+  
+}
+
+// copy output = t(input)
+// I don't think it will work with float.
+//[[Rcpp::export]]
+void copyToCpu(
+    Rcpp::NumericMatrix output,
+    Rcpp::S4 input
+  ) {
+  
+  
+  Rcpp::traits::input_parameter< std::string >::type classVarR(RCPP_GET_CLASS(input));
+  std::string precision_type = (std::string) classVarR;
+  
+  if(precision_type == "fvclMatrix") {
+    copyToCpuTemplated<float>(
+      output, input);
+  } else if (precision_type == "dvclMatrix") {
+    copyToCpuTemplated<double>(
+      output, input);
+  } else {
+    Rcpp::warning("class of input must be fvclMatrix or dvclMatrix");
+  }
+}

### 2021-08-27 pkg/gpuRandom/src/copy.cpp
fixed with clFinish
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2770 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,52 +1,86 @@
 #include "gpuRandom.hpp"
 
 
-//#define DEBUGKERNEL
+#define DEBUG
 
-// TO DO: use async_copy to copy by column and vector_iterator
-template<typename T> 
-void copyToCpuTemplated(
-    Rcpp::NumericMatrix output,
-    Rcpp::S4 input) {
-  
-  const bool BisVCL=1;
-  const int ctx_id = INTEGER(input.slot(".context_index"))[0]-1;
-  std::shared_ptr<viennacl::matrix_base<T> > xMat = getVCLptr<T>(input.slot("address"), BisVCL, ctx_id);
 
-  viennacl::vector_base<T> x((*xMat).handle(), (*xMat).internal_size(), 0, 1);
-  
-  
-  int Drow=0;
-  const int Nrow = (*xMat).size1(), Ncol = (*xMat).size2(), NcolInt = (*xMat).internal_size2();
-  for(Drow = 0; Drow < Nrow; Drow++) {
-    viennacl::fast_copy(
-      x.begin() + Drow*NcolInt,// + Drow*NcolInt, 
-      x.begin() + Drow*NcolInt + Ncol,// + Drow*NcolInt+Ncol, 
-      output.begin() +Drow*Ncol
-    );
-  }
-  
-}
 
 // copy output = t(input)
 // I don't think it will work with float.
 //[[Rcpp::export]]
-void copyToCpu(
-    Rcpp::NumericMatrix output,
+Rcpp::NumericMatrix copyToCpu(
     Rcpp::S4 input
   ) {
-  
   
   Rcpp::traits::input_parameter< std::string >::type classVarR(RCPP_GET_CLASS(input));
   std::string precision_type = (std::string) classVarR;
   
-  if(precision_type == "fvclMatrix") {
-    copyToCpuTemplated<float>(
-      output, input);
-  } else if (precision_type == "dvclMatrix") {
-    copyToCpuTemplated<double>(
-      output, input);
-  } else {
-    Rcpp::warning("class of input must be fvclMatrix or dvclMatrix");
+  if(precision_type != "dvclMatrix") {
+        Rcpp::warning("class of input must be dvclMatrix");
   }
+  
+  // TO DO: use async_copy to copy by column and vector_iterator
+
+  const bool BisVCL=1;
+  const int ctx_id = INTEGER(input.slot(".context_index"))[0]-1;
[…]

### 2021-08-27 pkg/gpuRandom/src/backsolve.cpp
fixed with clFinish for more functions
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2771 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -529,8 +529,9 @@ void backsolveBatch(
   backsolveKernel.local_work_size(0, Nlocal[0]);
   backsolveKernel.local_work_size(1, Nlocal[1]);
   
-  
+  viennacl::ocl::command_queue theQueue = backsolveKernel.context().get_queue();
   viennacl::ocl::enqueue(backsolveKernel(C, A, B, Nmatrix));
+  clFinish(theQueue.handle().get());
   
 }
 
@@ -564,7 +565,7 @@ SEXP backsolveBatchTyped(
   std::shared_ptr<viennacl::matrix<T> > AG = getVCLptr<T>(AR.slot("address"), BisVCL, ctx_id);
   std::shared_ptr<viennacl::matrix<T> > BG = getVCLptr<T>(BR.slot("address"), BisVCL, ctx_id);
   std::shared_ptr<viennacl::matrix<T> > CG = getVCLptr<T>(CR.slot("address"), BisVCL, ctx_id);
-  
+
   backsolveBatch<T>(*CG, *AG, *BG, 
                     Cstartend, Astartend, Bstartend,
                     numbatchB,diagIsOne, 

### 2021-08-27 pkg/gpuRandom/src/multiplyLowerTri.cpp

@@ -397,9 +397,10 @@ void multiplyLowerDiagonalBatch(
   multiplyKernel.local_work_size(2, 1L);//Nlocal[2]);
 
     // diagonals and diagTimesRowOfA
+  viennacl::ocl::command_queue theQueue = multiplyKernel.context().get_queue();
   viennacl::ocl::enqueue(multiplyKernel(
-      C, A, D, B));
-
+      C, A, D, B), theQueue);
+  clFinish(theQueue.handle().get());
 }
 
 template <typename T> 

### 2021-08-31 pkg/gpuRandom/src/crossprodBatch.cpp
crossprodbatch.cpp
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2779 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -199,9 +199,8 @@ std::string crossprodBatchString(
   }
   
   result += 
-    "  barrier(CLK_GLOBAL_MEM_FENCE);\n"
-    "      Cout=0.0;\n\n";
-  
+    "      Cout=0.0;\n\n"
+  "  barrier(CLK_LOCAL_MEM_FENCE);\n";
   result += 
     
     "      // cached parts\n";

### 2021-09-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
updated document folder extdata folder
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2782 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article,nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\usepackage{float}
\usepackage{framed}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{listings}
\renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }

%\addbibresource{paper2.bib}
\def\cov{{\text{COV}}}
\def\E{{\text{E}}}
\def\T{{\footnotesize{^{_{\sf T}}}}}
\setkeys{Gin}{width=0.45\textwidth}
\setcitestyle{square}

\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

\author{Ruoyong Xu\\University of Toronto
   \And Patrick Brown\\University of Toronto}
\Plainauthor{Ruoyong Xu, Patrick Brown}

\title{Random number simulation and applications on GPU’s in R?}
\Plaintitle{}
\Shorttitle{}

\Abstract{
Parallel processing by graphics processing unit (GPU) can be used to speed up computationally intensive tasks, which when combined with \proglang{R}, it can largely improve \proglang{R}’s downsides in terms of slow speed, memory usage and computation mode.  There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling. We introduce the \proglang{R} package \pkg{gpuRandom} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by utilizing the \pkg{clRNG} (OpenCL) library, the package enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.  \pkg{gpuRandom} is portable and flexible, developers can use its random number generation kernel for various other purposes and applications. 
}

\Keywords{GPU, \pkg{gpuRandom} package, parallel computing, \pkg{clRNG} library}
\Plainkeywords{GPU, gpuRandom package, parallel computing, clRNG library}

\Address{
  Ruoyong Xu\\
  Department of Statistics\\
  University of Toronto\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}

\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  University of Toronto\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
%\SweaveOpts{concordance=FALSE}

\section[Introduction]{Introduction}

\begin{leftbar}
% Why?
% 
% gpu massively parallel and relatively cheap, useful for stats
% there are gpu packages in R, gpuR
% no random numbers yet.  complicated to do on gpu, need multiple seeds becuase of parallel generation
% reproducability: need to save current state and restore

In recent years, parallel computing with R \citep{r2021} has become a very important topic and attracted lots of interest from researchers \citep[see][for a review]{eddelbuettel2021parallel}. Although \proglang{R} is one of the most popular statistical software with many advantages, it has drawbacks in memory usage and computation mode aspects \citep{zhao_2016}. To be more specific, (1) \proglang{R} requires all data to be loaded into the main memory (RAM) and thus can handle a very limited size of data; (2) \proglang{R} is a single-threaded program, it can not effectively use all the computing cores of multi-core processors. Parallel computing is the solution to these drawbacks of \proglang{R}, for an overview of current parallel computing approaches with \proglang{R}, see CRAN Task View by \citet{cran2021} at \url{https://cran.r-project.org/web/views/HighPerformanceComputing.html}. One major way of parallel computing with \proglang{R} is using Graphics Processing Units (GPUs). GPUs can perform thousands of tests simultaneously, which makes them powerful at doing massive parallel computing and they are relatively cheap compared to multicore CPU's. Although there has been a few \proglang{R} packages developed that provide some GPU capability in the past years, they come with some restrictions. Packages such as \pkg{gputools}, \pkg{gpumatrix}, \pkg{cudaBayesreg}, \pkg{rpud} (available on github), are now no longer maintained, the popular \pkg{tensorflow} \citep{tensorflow1} package uses GPU via \proglang{Python}, which makes it hard to be incorporated in a new \proglang{R} package that uses GPU. All of these mentioned packages are restricted to \proglang{R} users with NVIDIA GPUs. \pkg{gpuR} \citep{gpur1} is the only \proglang{R} package with a convenient interface between \proglang{R} and GPUs and it is created for any \proglang{R} user with a GPU device. By utilizing the \pkg{ViennaCL} \citep*{rupp2016viennacl} library, it provides a bridge between \proglang{R} and non-proprietary GPUs through an \proglang{OpenCL} (Open Computing Language) backend, which when combined with \pkg{Rcpp} \citep{rcpp1} gives a building block for other \proglang{R} packages.

Random number generator (RNG) is critical in simulation-based statistical inference, there is not yet an \proglang{R} package that is able to generate random numbers on GPU. This is complicate to do on GPU because multiple streams will be needed for parallel generation of random numbers, and the stream states need to be able to be restored for reproducibility of simulation results. We introduce the \pkg{gpuRandom} \proglang{R} package that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.

The remaining sections are organized as follows. 
In Section 2, we introduce streams and the use of streams in work items on a GPU device, and describe and check the functions for generating uniform and Normal random numbers.
In Section 3, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we parallelise this Monte Carlo simulation task and implement it on GPU, then we provide two real data examples to demonstrate the function usage and test its R performance.
Section 4 simulates simultaneously multiple Gaussian random surfaces by applying GPU-generated Normal random numbers and GPU-computed batches of Mat\'ern covariance matrices, then we plot the generated random surfaces in the end. 
Finally, the paper concludes with a short summary and a discussion in Section 5.

\end{leftbar}

\section{Generating random numbers} \label{}
Random numbers are the basis in statistical simulations methods, and are critical in many other fields. \cite{l2012random} summarized the usual two steps to generate a random variable %from a specified distribution 
in computational statistics: (1) generating independent and identically distributed (i.i.d.) uniform random variables on the interval $(0, 1)$, (2) applying transformations to these i.i.d. $U(0, 1)$ random variables to get samples from the desired distribution. \cite{l2012random} and \cite{robert2004random} present several general transformation methods for generating non-uniform random variables, for example, the most frequently used inverse transform method for generating a random variable $X$ from some arbitrary distribution function $F$; The Box-Muller algorithm \citep{box1958note} for Gaussian random variable generation, and so on, which are all built on uniform random variables. Therefore, uniform random number generators (RNGs) are essential in simulating random numbers from all types of probability distributions. In the rest paper, ``RNG'' refers to uniform pseudo random number generator.

\proglang{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators, these RNGs are actually stream objects of different forms. \pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG} library, making it able to generate random numbers on GPUs. In what follows, we will illustrate how to create stream objects and how to use streams to generate (standard) Uniform and (standard) Gaussian random numbers.

\subsection{Creating streams}
In \proglang{clRNG} library, a stream object is a structure that contains three states (see the \proglang{R} output below): the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream. The \proglang{clRNG} function \fct{clrngMrg31k3pCreateStreams} creates streams on the host computer using the MRG31k3p RNG. If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory. In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device. \fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams. \fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on the GPU device. From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used. Below, we show the usage of these two functions by generating 4 streams on the host and on the device respectively.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on host
streams_on_CPU<-gpuRandom::CreateStreamsCpu(Nstreams=4)
t(streams_on_CPU)
##                 [,1]       [,2]       [,3]       [,4]
## current.g1.1   12345  336690377  502033783  739421137
## current.g1.2   12345  597094797 1322587635 1475938232
## current.g1.3   12345 1245771585 1964121530  730262207
## current.g2.1   12345   85196284 1949818481 1630192198
## current.g2.2   12345  523477687 1607232546  324551134
## current.g2.3   12345 2094976052 1462898381  795289868
## initial.g1.1   12345  336690377  502033783  739421137
## initial.g1.2   12345  597094797 1322587635 1475938232
## initial.g1.3   12345 1245771585 1964121530  730262207
## initial.g2.1   12345   85196284 1949818481 1630192198
## initial.g2.2   12345  523477687 1607232546  324551134
## initial.g2.3   12345 2094976052 1462898381  795289868
## substream.g1.1 12345  336690377  502033783  739421137
## substream.g1.2 12345  597094797 1322587635 1475938232
## substream.g1.3 12345 1245771585 1964121530  730262207
## substream.g2.1 12345   85196284 1949818481 1630192198
## substream.g2.2 12345  523477687 1607232546  324551134
## substream.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create. For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) has a default value ``12345''. By setting a seed identical to the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as the streams created on host as follows,
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on GPU
seed <- c(12345, 12345, 12345, 12345, 12345, 12345)
streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
streams_from_GPU <- as.matrix(streams1)
streams2 = deepcopy(streams1)
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
t(streams_from_GPU)   
##                 [,1]       [,2]       [,3]       [,4]
## current.g1.1   12345  336690377  502033783  739421137
## current.g1.2   12345  597094797 1322587635 1475938232
## current.g1.3   12345 1245771585 1964121530  730262207
## current.g2.1   12345   85196284 1949818481 1630192198
## current.g2.2   12345  523477687 1607232546  324551134
## current.g2.3   12345 2094976052 1462898381  795289868
## initial.g1.1   12345  336690377  502033783  739421137
## initial.g1.2   12345  597094797 1322587635 1475938232
## initial.g1.3   12345 1245771585 1964121530  730262207
## initial.g2.1   12345   85196284 1949818481 1630192198
## initial.g2.2   12345  523477687 1607232546  324551134
## initial.g2.3   12345 2094976052 1462898381  795289868
## substream.g1.1 12345  336690377  502033783  739421137
## substream.g1.2 12345  597094797 1322587635 1475938232
## substream.g1.3 12345 1245771585 1964121530  730262207
## substream.g2.1 12345   85196284 1949818481 1630192198
## substream.g2.2 12345  523477687 1607232546  324551134
## substream.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
\begin{itemize}
  \item \code{seedR}: an R vector of length 6, which is the seed of streams. Default is set to a vector of six ``12345''.
  \item \code{Nstreams}: number of streams to create.
  \item \code{keepInitial}: logical specifying whether to keep the seed in generated streams. Default is set to TRUE.
\end{itemize}
Streams in \pkg{gpuRandom} are like \fct{.Random.seed} in \proglang{R}. By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \pkg{gpuRandom}, with the \code{seedR} argument, we are able to produce exactly the same sequence of streams as many times as we want.

\subsection{Using streams to generate  i.i.d. $U (0, 1)$ random numbers on GPU}
The streams created on GPU are then used in work items that executes in parallel on GPU device, in \pkg{gpuRandom} each work item takes one distinct stream to generate random numbers, so number of streams should equal to the number of total work items in use. A part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown below. Kernels are written in \proglang{OpenCL C} language, \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}. Pointers to the \code{streams} and to the matrix \code{out} on the global memory are passed to the kernel as arguments. \code{Nrow} and \code{Ncol} represent number of rows and columns of the matrix \code{out} respectively. \code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is internal number of columns of the \code{streams} matrix (these variables are defined in macros not shown here). Each stream is taken to the private memory of each work item by the function \fct{streamsToPrivate}. Here we use a 2-dimensional index space with work-group size $1 \times 2$.  \code{index} represents the position of the work item when mapping it to a one-dimensional vector. The function \fct{clrngMrg31k3pNextState} gives a uniform random integer named \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here. At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
\begin{framed}
\begin{lstlisting}[language=C,basicstyle=\small,label=uniform]
 __kernel void mrg31k3pMatrix(
    __global int* streams,
    __global double* out){

const int 
index = get_global_id(0)*get_global_size(1) + get_global_id(1),
DrowInc = 2*get_global_size(0), 
DrowStartInc = DrowInc * NpadCol,
startvalue=index * NpadStreams;
int Drow, Dcol, DrowStart, Dentry;
uint temp;
uint g1[3], g2[3];

streamsToPrivate(streams,g1,g2,startvalue);

//filling in random numbers in the output matrix
for(Drow=2*get_global_id(0), 
    DrowStart = (Drow + get_local_id(1))* NpadCol;
    Drow < Nrow; Drow += DrowInc, DrowStart += DrowStartInc) {

   for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
       Dcol < Ncol;
       Dcol += get_num_groups(1), Dentry += get_num_groups(1)) {

       temp = clrngMrg31k3pNextState(g1, g2);
       out[Dentry] = mrg31k3p_NORM_cl * temp;

    }//Dcol
 }//Drow

streamsFromPrivate(streams,g1,g2,startvalue);
}
\end{lstlisting}
\end{framed}

The kernel code may not be straightforward enough in showing how streams are used in work items to generate random numbers, so here we give a small example to illustrate the matrix filling process. As shown in Figure \ref{fig1}, suppose we use a $2 \times 4$ NDRange (index space size) ((a) in Figure \ref{fig1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig1}) of uniform random numbers using 8 streams. In each grid of (a) that represents a work item, there are two rows of numbers, the number in the above row is the stream index (from 0 to 7), and a pair of numbers in the second row is the work item index (each work item has a unique 2-dimensional global ID). Each stream object is passed to each work item. The number in a cell of matrix in (b) is the stream index, that indicates which stream does the random number generation for this matrix cell. For example, stream 0 which is allocated in work item $(0,0)$ generates a random number in matrix $(0,0)$ (row and column index of the matrix) in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it goes to matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.

\begin{figure}[H]%
    \centering
    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f1} }}%
    \qquad
    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f2} }}%
    \caption{name?\label{fig1}}%
\end{figure}

In the following code, we use the 4 streams created in Section 2.1 and 4 work items to generate a vector \code{sim_1} of 6 i.i.d. $U (0,1)$ random numbers by the function \fct{gpuRandom::runif}. By default \code{Nglobal} is set as (64,8), and a matrix of \code{prod(Nglobal)} streams created on GPU is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host computer.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_1 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_1)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a number specifying the length if to create a vector, or a vector of length 2 specifying the row and column number if to create a matrix.
  \item \code{streams}: streams used for random number generating.
  \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}
Next we show that using the same streams (\code{streams1} and \code{streams2}),  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_2 <- gpuRandom::runif(n = 6, streams = streams2, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_2)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}
Note that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used, reusing the \code{streams1} will produce a vector  different from \code{sim_1} (result of \code{sim_3}), because each time a random number is generated, the current state of the stream is changed (advances by one position). See below the first six rows (current states) of \code{streams1} after it being used.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_3 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_3)
## [1] 0.6487742 0.1112075 0.3661944 0.5018562 0.1088229 0.3114331
t(as.matrix(streams1))[1:6, ]
##            [,1]       [,2]       [,3]       [,4]
## [1,] 1167281028  640923766  502033783  739421137
## [2,] 1918428443 1912157188 1322587635 1475938232
## [3,] 1858462085  286315187 1964121530  730262207
## [4,]  933585541 2119609883 1949818481 1630192198
## [5,] 1132031887  834429287 1607232546  324551134
## [6,]  465230163   47498872 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
In order to reproduce the same sequence of random numbers, we can save the streams in a file on CPU, because GPU device does not keep memory for streams objects after each execution. Below is the \proglang{R} code in which we use the saved streams \code{streams_from_GPU.rds} to produce a vector \code{sim_saved}, which is also same as \code{sim_1}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
saveRDS(streams_from_GPU, "streams_from_GPU.rds")
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Load the streams object as streams_saved
streams_saved <- gpuR::vclMatrix(readRDS("streams_from_GPU.rds"))
sim_saved = gpuRandom::runif(n=6,streams=streams_saved, 
                             Nglobal=c(1,4), type="double")
as.vector(sim_saved)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}
We evaluate the distribution of 50000 GPU-generated uniform random numbers by making a density histogram  (the left panel of Figure \ref{fig0}), plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel of Figure \ref{fig0}). The two lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} should come from a standard uniform distribution.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =256, keepInitial=1)
random_vector<-gpuRandom::runif(n=50000, streams=streams, 
                                Nglobal=c(64,4), type="double")
random_vector<-as.vector(as.matrix(random_vector))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-1} 
\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-2} 
\end{knitrout}
\caption{\label{fig0} histogram and empirical cumulative distribution of the GPU-generated random numbers.}
\end{figure}

\subsection{Using streams to generate i.i.d. $N (0, 1)$ random numbers on GPU}
We apply the Box-Muller transformation to $U (0,1)$ random numbers to generate Gaussian random numbers. Box-Muller algorithm (see algorithm \ref{algorithm1}) takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables. The derivation basically uses transformation from Cartesian coordinates to polar coordinates to represent two independent, Normally distributed random variables. The Box-Muller algorithm turns out to be the best choice for Gaussian transform on GPU compared to other transform methods \citep{howes2007efficient}, because this algorithm has no branching or looping, which are the things GPU is not good at, but GPU is good at the high computational load of sine and cosine functions in the algorithm.

\begin{algorithm*}[H] \label{algorithm1}
\SetAlgoLined
 1, Generate $U_1$, $U_2$ i.i.d. from $U (0,1)$ \;
 2, Define \begin{align*} 
& R = \sqrt{-2*\log U_1},\\
&  \Theta = 2\pi*U_2,\\
&  X=R*\cos(\Theta),\\
& Y=R*\sin(\Theta);\;\end{align*}
 3, Take $X$ and $Y$ as two independent draws from $N(0,1)$\;
 \caption{Box-Muller algorithm.}
\end{algorithm*}
%
Below is a fragment of the kernel function that generates standard Gaussian random numbers. 
%
\begin{framed}
\begin{lstlisting}[language=C,basicstyle=\small,label=normal]
for(Drow=2*get_global_id(0), 
    DrowStart=(Drow + get_local_id(1))* NpadCol;
    Drow < Nrow; 
    Drow += DrowInc, DrowStart += DrowStartInc){

for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
      Dcol < Ncol;
      Dcol += get_num_groups(1), Dentry += get_num_groups(1)){

      temp = clrngMrg31k3pNextState(g1, g2);

     if(get_local_id(1)) {
       part[1] = TWOPI_mrg31k3p_NORM_cl * temp;
       cosPart1 = cos(part[1]);
       sinPart1 = sin(part[1]);
     } else {
       part[0] = sqrt(-2.0*log(mrg31k3p_NORM_cl * temp));
     }

     barrier(CLK_LOCAL_MEM_FENCE);

     if(get_local_id(1)) {
       out[Dentry] = part[0]*sinPart1;
     } else {
       out[Dentry] = part[0]*cosPart1;
     }
     barrier(CLK_LOCAL_MEM_FENCE);
   }//Dcol
   }//Drow
 }
\end{lstlisting}
\end{framed}
The kernel is adjusted from the kernel for generating uniform random numbers in \ref{uniform}, we only add the part that does the Box-Muller transformation. If developers want to use other Gaussian transform methods, this kernel code is easy to be modified and used in other \proglang{R} packages.  In the code, \code{part0} and \code{part1} correspond to $R$ and $\Theta$ in the formulas respectively. \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that $(X_1,Y_1), (X_2,Y_2), \dots, (X_n,Y_n)$ are generated correctly in pairs, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ will not occur. We illustrate the matrix filling process with the previous example: to create a $4 \times 6$ matrix ((b) in Figure \ref{fig3}) of Gaussian random numbers using a $2 \times 4$ NDRange ((a) in Figure \ref{fig3}) and 8 streams.

\begin{figure}[H]%
    \centering
    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f3} }}%
    \qquad
    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f4} }}%
    \caption{name???}%
    \label{fig3}
\end{figure}
As Gaussian numbers are generated in pairs using the Box-Muller method, work items are organized to be $1 \times 2$ work groups in the index space (a), the shading in the grids of (a) distinguishes work groups. Each work group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where the work item of `\code{get_local_id(1)}$=0$' generates the \code{X} in a pair, and work item of `\code{get_local_id(1)}$=1$' generates the \code{Y} in a pair. The numbers in the third row in each gird of (a) represents the local index of each work item. Cells of the matrix are filled from left to right, that is, in the first iteration of execution, the 8 work items generate random numbers in the leftmost two columns (8 cells) of matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration  fills the rest two columns of the matrix simultaneously. Therefore, each work item generates one Gaussian random number in each iteration, the filling process is the same as the filling process of Figure \ref{fig1}.

We show the usage of \fct{gpuRandom::rnorm} by generating a $512 \times 1024$ matrix named \code{normal_matrix} of double presicion Gaussian random numbers, and assess Normality by making a histogram (left panel of Figure \ref{fig2}) and a Q-Q plot (right panel of Figure \ref{fig2}) of the generated random numbers. The points in the Q-Q plot in Figure \ref{fig2} fall along the diagonal line, we have good evidence the data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part (adapted from \fct{stat::qnorm}) on GPU, which speeds up the Q-Q plot greatly.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="double")
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
avector<-as.vector(as.matrix(normal_matrix))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
\end{verbatim}
\end{kframe}
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-1} 
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-2} 
\end{knitrout}
\caption{\label{fig2} Histogram and Q-Q plot of the GPU-generated random numbers.}
\end{figure}
Let's compare the time of generating a large-size matrix of more than 26.8 million Gaussian random numbers of double precision using the \fct{stats::rnorm} and using the \fct{gpuRandom::rnorm}. Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $128 \times 128$ work-items is about 10 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
system.time(gpuRandom::rnorm(c(16384,16384), streams=streams, 
                             Nglobal=c(512,128), type="double"))
##    user  system elapsed 
##   0.112   0.001   1.279
system.time(matrix(stats::rnorm(16384*16384),16384,16384))
##    user  system elapsed 
##  11.962   0.898  12.838
\end{verbatim}
\end{kframe}
\end{knitrout}

\section{Monte Carlo simulation for Fisher's exact test}
One application of GPU-generated random numbers in \pkg{gpuRandom} is Monte Carlo simulation for Fisher's exact test. Fisher’s exact test is usually applied for analyzing $2 \times 2$ contingency tables when one of the expected values in table is less than 5. Different from methods which rely on approximation, Fisher's exact test computes directly the probability of obtaining each possible combination of the data for the same row and column totals (marginal totals) as the observed table, and get the exact p-value by adding together all the probabilities of tables as extreme or more extreme than the one observed. However, when the observed table gets large in terms of sample size and table dimensions, the number of combinations of cell frequencies with the same marginal totals gets very large, \cite[][p. 23]{mehta2011ibm} shows a $5 \times 6$ observed table that has 1.6 billion possible tables. Calculating the exact P-values may lead to very long running time and can sometimes exceed the memory limits of your computer. Hence, the option \code{simulate.p.value = TRUE} in \fct{stats::fisher.test} is provided, which enables computing p-values by Monte Carlo simulation for tables larger than 2 $\times$ 2. Given an observed table and a number of replicates $B$, the Monte Carlo simulation does the following steps: 
\begin{enumerate}
  \item Calculate the test statistic for the observed table.
	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic (minus log-factorial of table) from the random table.
	\item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
	\item Estimate p-value using $\frac{1 + \emph{Counts}}{B + 1}$.
\end{enumerate}
The Monte Carlo simulation solves the computational problem of Fisher's exact test but can still take long running time when $B$ reaches millions in scale, thus we make the function \fct{gpuRandom::fisher.sim} that does the steps 1, 2 and 3 parallelly on GPU, and step 2 of saving the test statistics from the random tables as an option in \fct{gpuRandom::fisher.sim}, which largely saves processing time.

Next we compare the run time between using \fct{gpuRandom::fisher.sim} and \fct{stats::fisher.test} using two real data examples with small and large p-values respectively. These 2-way contingency tables (\ref{tab:month} and \ref{tab:week}) consists of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}. The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday. The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.

\subsection{Comparing running time: Month data example}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]

\caption{\label{tab:monthdata}Month\label{tab:month}}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
\hline
Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
\hline
Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
\hline
Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
\hline
Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
\hline
May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
\hline
Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
\hline
Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
\hline
Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
\hline
Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
\hline
Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
\hline
Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
\hline
Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
\hline
\end{tabular}
\end{table}

\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using CPU
stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
## [1] 0.4035866
system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
##    user  system elapsed 
##   9.753   0.004   9.740
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for one million simulations, has a p-value of 0.4035.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
almost.1 <- 1 + 64 * .Machine$double.eps
observed_m= -gpuRandom::logfactSum(month, c(16,16))/almost.1
month_GPU<-vclMatrix(month,type="integer")
system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = Nglobal))
##    user  system elapsed 
##   0.680   0.008   2.291
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::fisher.sim} takes about 2.3 seconds, compared to the elapsed time of \fct{gpuRandom::fisher.sim},  \fct{gpuRandom::fisher.sim} takes less than a quater of the time. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = Nglobal)
result_monthgpu$simnum
## [1] 999424
result_monthgpu$counts
## [1] 403274
result_monthgpu$p.value
## [1] 0.403507
\end{verbatim}
\end{kframe}
\end{knitrout}
We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.4035, which is close enough to the p-value from \fct{stats::fisher.test}: 0.4035.

\subsection{Comparing running time: Week data example}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]

\caption{\label{tab:weekdata}Week\label{tab:week}}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
\hline
Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
\hline
Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
\hline
Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
\hline
Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
\hline
Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
\hline
Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
\hline
Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
\hline
\end{tabular}
\end{table}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using CPU
fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
## [1] 0.0001189
system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
##    user  system elapsed 
##  60.412   0.088  60.395
\end{verbatim}
\end{kframe}
\end{knitrout}
The ``week'' table has a much smaller p-value: around 0.000119, which should require a larger number of simulations to get a more accurate p-value.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
week_GPU<-gpuR::vclMatrix(week,type="integer")
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
##    user  system elapsed 
##   0.613   0.010  10.839
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes 60.4 seconds, while \fct{fisher.sim} takes about 10.8 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = Nglobal)
result_weekgpu$simnum
## [1] 9994240
result_weekgpu$counts
## [1] 1272
result_weekgpu$p.value
## [1] 0.0001273734
\end{verbatim}
\end{kframe}
\end{knitrout}
With about ten million simulations, we get 1272 cases and a p-value around 0.0001273. We summarized the results in Table \ref{tab:summary}.

\begin{table}[H]
\begin{tabular}{ |c|c|c|c|c|c| } 
 \hline
 test example & B & p-value stats & p-value gpuRandom & time stats & time gpuRandom \\ \hline
 Month & 1e6 & 0.4029 & 0.4035 & 9.8 & 2.3\\ \hline
 Week & 1e7 & 0.000119 & 0.000127 & 60.4 & 10.8\\ \hline
\end{tabular}
\caption{\label{tab:summary} Comparison summary of \fct{gpuRandom::fisher.sim} and \fct{stats::fisher.test}.}
\end{table}

Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example. The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot. 
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
\includegraphics[width=0.45\textwidth]{figure/fighistMonth-1} 
\includegraphics[width=0.45\textwidth]{figure/fighistMonth-2} 
\end{knitrout}
\caption{\label{fig4} Frequency distribution for test statistics from month and week data tables.}
\end{figure}

\section{Generating Gaussian random fields} 
Given a parameter space $X$, a random field $U$ on $X$ is a collection of random variables $\{U_x :x\in X\}$. A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})$ is multivariate Gaussian. One reason that Gaussian random fields play an important role is that the specification of their finite-dimensional distributions is simple \citep{abrahamsen1997review}, the expectation $\mu(x)$ and covariance function $\Sigma$ completely determine the distribution of $U$, where 
\begin{gather*} 
\mu(x) = \E (U(x)),\\
\Sigma_{ij} = \cov (U(x_i),U(x_j)).
\end{gather*}
There exist many choices for the covariance function $\Sigma$, one common choice is the Mat\'ern covariance function \citep{matern1960spatial}. The Mat\'ern covariance between $U(x)$ and $U(x')$ takes the form 
\begin{equation*}
\Sigma(d;\sigma^2,\phi,\kappa)=\sigma^2*\frac{2^{\kappa-1}}{\Gamma(\kappa)} (\sqrt{8\kappa} \frac{d}{\phi})^\kappa  K_\kappa(\sqrt{8\kappa}  \frac{d}{\phi}),  \quad \text{where} \quad \phi 	\geq 0,  \kappa 	\geq 0,
\end{equation*} 
$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
$\sigma^2$ the variance of the random field $U$.
$\Gamma(\cdot)$ is the standard gamma function,  $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as $d$ increases. $K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$, specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$. (For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.)  There are several alternative parameterisations of Mat\'ern covariance functions appeared in literatures  \citep[see][]{haskard2007anisotropic}. In \pkg{gpuRandom}, we use the above form for computing Mat\'ern covariance.

Due to the importance of Gaussian random field, there are a large amount of studies devoted to Gaussian random field generation techniques. \cite{LiuandLi2019} gives a comprehensive review on 7 popular methods for Gaussian random field generation, which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}. All of these methods except matrix decomposition method, are approximations and has specific requirements on the type of grid or covariance functions. The matrix decomposition method is exact, it works for all covariance functions and can generate random field on all types of grids. Other \proglang{R} packages that offer simulation of Gaussian random fields like \pkg{geoR} \citep{geoR2001}, does not work for large number of locations; the \pkg{RandomFields} \citep{RandomFields2015,RandomFields2020} package can use different methods for simulation of Gaussian fields, among which the (modified) circulant embedding method \citep{Dietrich1997FastAE} for covariance matrix decomposition is also an exact method (need to confirm???), however, it works only for isotropic Gaussian fields and on rectangular grids. \pkg{gpuRandom} does exact simulation of exact Gaussian fields on the GPU as it relies on the matrix decomposition method. The implementation of this method is as follows.

Suppose $U=(U_1, U_2, \dots, U_n)$ is a Gaussian random field with mean $\mu$ and covariance matrix $\Sigma$, without loss of generality, we assume mean value zero $(\mu = 0)$. Since covariance matrix $\Sigma$ is symmetric and positive-definite, we can take Cholesky decomposition of $\Sigma$, then we can simulate random samples from $U$ using $U=L*D^{1/2}*Z$, where 
\begin{itemize}
\item $L$ is the lower unit triangular matrix in Cholesky decomposition of $\Sigma = L*D*L^\top$,
\item $D$ is a diagonal matrix.
\item $Z=(Z_1, Z_2, \dots, Z_n) \sim \text{MVN}(0,I_n)$, where $I_n$ is a $n \times n$ identity matrix.
\end{itemize}
The matrix decomposition method is straightforward to implement, however, when a large number ($n$) of locations is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.  \pkg{gpuRandom} package takes advantage of GPU's parallel computing power to not only do the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU. The following shows an example, in which we simulate 10 Gaussian random fields of mat\'ern covariance with 5 sets of parameters at one time, and introduces several GPU functions used in the simulation process.

\subsection{Simulating Gaussian random fields with Mat\'ern covariances}

Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
Ngrid = c(60, 80)
NlocalCache = 1000
Nglobal = c(128, 64, 2)
Nlocal = c(4, 2, 2)
theType = "float"
myRaster = raster(
  extent(0,Ngrid[1]/Ngrid[2],5,6),
  Ngrid[2], Ngrid[1]
)
coordsSp = sp::SpatialPoints(xyFromCell(myRaster, 1:ncell(myRaster)))
coordsGpu = vclMatrix(coordsSp@coords, 
                      nrow(coordsSp@coords), 
                      ncol(coordsSp@coords), type=theType)
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets. Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuRandom::maternGpuParam}.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
myParamsBatch
##      shape range variance nugget anisoRatio anisoAngleRadians
## [1,]  1.25  0.50      1.5      0          1         0.0000000
## [2,]  2.15  0.25      2.0      0          4         0.4487990
## [3,]  0.55  1.50      2.0      0          4         0.4487990
## [4,]  2.15  0.50      2.0      0          4        -0.4487990
## [5,]  2.15  0.50      2.0      0          2         0.7853982
paramsGpu = gpuRandom::maternGpuParam(myParamsBatch, type=theType)
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 3, compute Mat\'ern covariance matrices using \fct{gpuRandom::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices 
are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
                         nrow(coordsGpu),type=theType)
dim(maternCov)
## [1] 24000  4800
gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
                       Nglobal=c(64,64,16), Nlocal=c(4,4,4))
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 4, the first argument \code{maternCov} in \fct{gpuRandom::cholBatch} specifies the matrix object to take Cholesky decomposition. Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
                     Nglobal,Nlocal,NlocalCache)
\end{verbatim}
\end{kframe}
\end{knitrout}
% \begin{gather}
% \begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
% \end{bmatrix}
%  \rightarrow
%  \begin{bmatrix}
%   L_{1} \\ L_{2} \\L_{3} \\ \vdots
%   \end{bmatrix} \text{and}
%   \begin{bmatrix}
%   D_{1} \\ D_{2} \\D_{3} \\ \vdots
%   \end{bmatrix}
% \end{gather}
Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{gpuRandom::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streamsGpu <- gpuRandom::CreateStreamsGpu(Nstreams=128*64, keepInitial=TRUE)
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
zmatGpu = gpuRandom::rnorm(c(nrow(maternCov),2), 
                           streams=streamsGpu, Nglobal=c(128,64),
                           type = theType)
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 6, use \fct{gpuRandom::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration. \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
                   type = theType)

gpuRandom::multiplyLowerDiagonalBatch(
  simMat, maternCov, diagMat, zmatGpu,
  diagIsOne = 1L, # diagonal of L is one
  transformD = "sqrt", # take the square root of each element of D
  Nglobal,
  Nlocal,
  NlocalCache)
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{gather}
 \begin{bmatrix}  L_{1} \\ L_{2} \\L_{3} \\ \vdots
 \end{bmatrix} 
 *
  \begin{bmatrix}
   D_{1} \\ D_{2} \\D_{3} \\ \vdots
   \end{bmatrix} 
   *
   \begin{bmatrix}
   Z_{11} & Z_{12}
   \end{bmatrix}
  =
 \begin{bmatrix}
   L_1D_1Z_{11} & L_1D_1Z_{12} \\
   L_2D_2Z_{11} & L_2D_2Z_{12} \\
   L_3D_3Z_{11} & L_3D_3Z_{12} 
  \end{bmatrix}
\end{gather}
Step 7, Finally, plot the 10 simulated Gaussian random surfaces.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
library(raster)
simRaster = brick(myRaster, nl = ncol(simMat)*nrow(paramsGpu))
values(simRaster) = as.vector(as.matrix(simMat))
names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
                                     'sim', 1:ncol(simMat)), 1, paste, collapse='')
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
par(mar=rep(0.1, 4))
for(D in names(simRaster)) {
  plot(extent(simRaster))
  plot(simRaster[[D]], legend=FALSE, add=TRUE)
}
\end{verbatim}
\end{kframe}
\includegraphics[width=0.45\textwidth]{figure/maternplot-1} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-2} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-3} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-4} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-5} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-6} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-7} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-8} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-9} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-10} 
\end{knitrout}
\caption{\label{fig5} }
\end{figure}

\section{Discussion}
The package \pkg{gpuRandom} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU. We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the Normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU. Most of these functions uses local memory on GPU, all of them uses parallel programming and avoids unnecessary data transfers between host and device. Comparison of performance between using \pkg{gpuRandom} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.

By leveraging the \pkg{gpuR} package, \pkg{gpuRandom} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.  \pkg{gpuRandom} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for further development.

One of the main difficulties in developing the package is bug fixes and uncertain software behaviors. Error messages are often vague and it is time-consuming to locate the error within the backend \proglang{C} program. The usual steps that we take to debug is commenting out some parts of codes and then recompile many times until we find the problem, or printing out flag messages in several places in the code to find out where the program stops at due to error. \pkg{gpuRandom} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.

Future work on \pkg{gpuRandom} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs. (2) Now that the package can do Cholesky decomposition and matrix multiplication in batches on GPU, which is a motivation for us to work on parallel likelihood evaluations on GPU for Gaussian spatial models in the next step. (3) Create a ``sparse'' matrix class on GPU and use it for simulating Gaussian random fields with sparse correlation structures such as the Gaussian Markov random fields.

\bibliography{paper1}

\newpage

\begin{appendix}

\end{appendix}

\end{document}

### 2021-09-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
updated document folder extdata folder
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2782 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-09-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
-\documentclass[article,nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
-% maxwidth is the original width if it is less than linewidth
-% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-\let\hlipl\hlkwb
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
+\documentclass[article,nojss]{jss}
 
@@
 
+<<echo=FALSE,results="hide">>=
+options(width=65)
+@
+
 \newcommand{\class}[1]{`\code{#1}'}
 \newcommand{\fct}[1]{\code{#1()}}
+
+<<preliminaries, echo=FALSE, results="hide">>=
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+knitr::opts_chunk$set(highlight=FALSE, background='#FFFFFF', fig.height=4, fig.width=6, out.width='0.45\\textwidth')
+library("MASS")
+@
 
@@
   E-mail: \email{ruoyong.xu@mail.utoronto.ca}}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+
 \begin{document}
@@
 \end{leftbar}
+<<echo=FALSE, results="hide", message=FALSE>>=
+library("gpuR")
+library("gpuRandom")
+setContext(   grep('gpu', listContexts()$device_type) [1]    )
+setwd("/home/ruoyong/paper1_2021_5")
+@
 
@@
 Below, we show the usage of these two functions by generating 4 streams on the host and on the device respectively.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<streams on CPU>>=
 # Create streams on host
@@
 t(streams_on_CPU)
-##                 [,1]       [,2]       [,3]       [,4]
-## current.g1.1   12345  336690377  502033783  739421137
-## current.g1.2   12345  597094797 1322587635 1475938232
-## current.g1.3   12345 1245771585 1964121530  730262207
-## current.g2.1   12345   85196284 1949818481 1630192198
-## current.g2.2   12345  523477687 1607232546  324551134
-## current.g2.3   12345 2094976052 1462898381  795289868
-## initial.g1.1   12345  336690377  502033783  739421137
-## initial.g1.2   12345  597094797 1322587635 1475938232
-## initial.g1.3   12345 1245771585 1964121530  730262207
-## initial.g2.1   12345   85196284 1949818481 1630192198
-## initial.g2.2   12345  523477687 1607232546  324551134
-## initial.g2.3   12345 2094976052 1462898381  795289868
-## substream.g1.1 12345  336690377  502033783  739421137
-## substream.g1.2 12345  597094797 1322587635 1475938232
-## substream.g1.3 12345 1245771585 1964121530  730262207
-## substream.g2.1 12345   85196284 1949818481 1630192198
-## substream.g2.2 12345  523477687 1607232546  324551134
-## substream.g2.3 12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create.
@@
 By setting a seed identical to the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as the streams created on host as follows,
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<streams on GPU1, eval=TRUE>>=
 # Create streams on GPU
@@
 streams2 = deepcopy(streams1)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+@ 
+<<streams on GPU2, eval=TRUE, echo=FALSE>>=
+gpuR::colnames(streams_from_GPU) = c("current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
+"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3",
+"substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")   
+@ 
+<<streams on GPU3,eval=TRUE>>=
 t(streams_from_GPU)   
-##                 [,1]       [,2]       [,3]       [,4]
-## current.g1.1   12345  336690377  502033783  739421137
-## current.g1.2   12345  597094797 1322587635 1475938232
-## current.g1.3   12345 1245771585 1964121530  730262207
-## current.g2.1   12345   85196284 1949818481 1630192198
-## current.g2.2   12345  523477687 1607232546  324551134
-## current.g2.3   12345 2094976052 1462898381  795289868
-## initial.g1.1   12345  336690377  502033783  739421137
-## initial.g1.2   12345  597094797 1322587635 1475938232
-## initial.g1.3   12345 1245771585 1964121530  730262207
-## initial.g2.1   12345   85196284 1949818481 1630192198
-## initial.g2.2   12345  523477687 1607232546  324551134
-## initial.g2.3   12345 2094976052 1462898381  795289868
-## substream.g1.1 12345  336690377  502033783  739421137
-## substream.g1.2 12345  597094797 1322587635 1475938232
-## substream.g1.3 12345 1245771585 1964121530  730262207
-## substream.g2.1 12345   85196284 1949818481 1630192198
-## substream.g2.2 12345  523477687 1607232546  324551134
-## substream.g2.3 12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
@@
 The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host computer.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-sim_1 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
-    type = "double")
+<<eval=TRUE, tidy=TRUE>>=
+sim_1 = gpuRandom::runif(n=6,streams=streams1, Nglobal=c(1,4), type="double")
 as.vector(sim_1)
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 The arguments of the \fct{gpuRandom::runif} are described as follows:
@@
 Next we show that using the same streams (\code{streams1} and \code{streams2}),  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-sim_2 <- gpuRandom::runif(n = 6, streams = streams2, Nglobal = c(1, 4),
-    type = "double")
+<<eval=TRUE,tidy=TRUE>>=
+sim_2 <-gpuRandom::runif(n=6, streams=streams2, Nglobal=c(1,4), type = "double")
 as.vector(sim_2)
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 Note that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used, reusing the \code{streams1} will produce a vector  different from \code{sim_1} (result of \code{sim_3}), because each time a random number is generated, the current state of the stream is changed (advances by one position).
 See below the first six rows (current states) of \code{streams1} after it being used.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-sim_3 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
-    type = "double")
+<<eval=TRUE,tidy=TRUE>>=
+sim_3 = gpuRandom::runif(n=6,streams=streams1, Nglobal=c(1,4), type="double")
 as.vector(sim_3)
-## [1] 0.6487742 0.1112075 0.3661944 0.5018562 0.1088229 0.3114331
-t(as.matrix(streams1))[1:6, ]
-##            [,1]       [,2]       [,3]       [,4]
-## [1,] 1167281028  640923766  502033783  739421137
-## [2,] 1918428443 1912157188 1322587635 1475938232
-## [3,] 1858462085  286315187 1964121530  730262207
-## [4,]  933585541 2119609883 1949818481 1630192198
-## [5,] 1132031887  834429287 1607232546  324551134
-## [6,]  465230163   47498872 1462898381  795289868
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+t(as.matrix(streams1))[1:6,]
+@
 In order to reproduce the same sequence of random numbers, we can save the streams in a file on CPU, because GPU device does not keep memory for streams objects after each execution.
 Below is the \proglang{R} code in which we use the saved streams \code{streams_from_GPU.rds} to produce a vector \code{sim_saved}, which is also same as \code{sim_1}.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<echo=TRUE,eval=FALSE>>=
 saveRDS(streams_from_GPU, "streams_from_GPU.rds")
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<save streams,eval=TRUE>>=
 # Load the streams object as streams_saved
@@
 as.vector(sim_saved)
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 We evaluate the distribution of 50000 GPU-generated uniform random numbers by making a density histogram  (the left panel of Figure \ref{fig0}), plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel of Figure \ref{fig0}).
 The two lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} should come from a standard uniform distribution.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<uniforms_random_vector,eval=TRUE>>=
 streams <- CreateStreamsGpu(Nstreams =256, keepInitial=1)
@@
 random_vector<-as.vector(as.matrix(random_vector))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \begin{figure}[H]
 \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
-\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-1} 
-\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-2} 
-\end{knitrout}
+<<largenumberuniforms, eval=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold'>>=
+hist(random_vector, prob = TRUE, main="Histogram")
+plot(ecdf(random_vector), ylab="F(x)",main="Empirical cumulative distribution")
+curve(punif(x, 0, 1), lty=2, add=TRUE, col="red")
+@
 \caption{\label{fig0} histogram and empirical cumulative distribution of the GPU-generated random numbers.}
@@
 We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part (adapted from \fct{stat::qnorm}) on GPU, which speeds up the Q-Q plot greatly.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<histAndQQ_prepare,eval=TRUE>>=
 streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
@@
                                 Nglobal=c(512,128), type="double")
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+@
+<<histAndQQ_prepare_for_cache, eval=TRUE, cache=TRUE>>=
 avector<-as.vector(as.matrix(normal_matrix))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \begin{figure}[H]
 \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+ <<histAndQQ, echo=TRUE, eval=TRUE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold'>>=
 hist(avector,breaks=40, main = "histogram")
 gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=0.45\textwidth]{figure/histAndQQ-1} 
-\includegraphics[width=0.45\textwidth]{figure/histAndQQ-2} 
-\end{knitrout}
+@
 \caption{\label{fig2} Histogram and Q-Q plot of the GPU-generated random numbers.}
@@
 The difference in elapsed time becomes larger when matrix size goes larger.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<random normal time compare, eval=TRUE, cache=FALSE>>=
 system.time(gpuRandom::rnorm(c(16384,16384), streams=streams, 
                              Nglobal=c(512,128), type="double"))
-##    user  system elapsed 
-##   0.112   0.001   1.279
 system.time(matrix(stats::rnorm(16384*16384),16384,16384))
-##    user  system elapsed 
-##  11.962   0.898  12.838
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 
@@
 \subsection{Comparing running time: Month data example}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
-
-\caption{\label{tab:monthdata}Month\label{tab:month}}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
-\hline
-Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
-\hline
-Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
-\hline
-Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
-\hline
-Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
-\hline
-May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
-\hline
-Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
-\hline
-Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
-\hline
-Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
-\hline
-Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
-\hline
-Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
-\hline
-Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
-\hline
-Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
-\hline
-\end{tabular}
-\end{table}
-
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<Timecomparemonth,echo=FALSE,eval=TRUE, results="hide">>=
+data2<-read.csv("/home/ruoyong/paper2/month.csv") 
+data2<-data2[,-1]
+month<-as.matrix(data2)
+rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
+colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
+Nglobal = c(128,128)
+streams <- CreateStreamsGpu(Nstreams =128*128, keepInitial=1)
+@
+<<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
+library(knitr)
+library(kableExtra)
+knitr::kable(head(month, n=12),format="latex",  caption = "Month\\label{tab:month}") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F, position = "center",latex_options = "HOLD_position")
+@
+<<timecomparemonth2Cpu, cache=TRUE, eval=TRUE>>=
 #using CPU
 stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
-## [1] 0.4035866
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
-##    user  system elapsed 
-##   9.753   0.004   9.740
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for one million simulations, has a p-value of 0.4035.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
 #using GPU
@@
             type="double", returnStatistics=FALSE, Nglobal = Nglobal))
-##    user  system elapsed 
-##   0.680   0.008   2.291
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \fct{gpuRandom::fisher.sim} takes about 2.3 seconds, compared to the elapsed time of \fct{gpuRandom::fisher.sim},  \fct{gpuRandom::fisher.sim} takes less than a quater of the time. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<timecomparemonth3Gpu, eval=TRUE,cache=TRUE>>=
 result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
@@
 result_monthgpu$simnum
-## [1] 999424
 result_monthgpu$counts
-## [1] 403274
 result_monthgpu$p.value
-## [1] 0.403507
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.4035, which is close enough to the p-value from \fct{stats::fisher.test}: 0.4035.
+<<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
+result_monthgpu$cpu <- as.vector(result_monthgpu$sim)
+@
 
 \subsection{Comparing running time: Week data example}
+<<Timecompareweek, echo=FALSE, eval=TRUE, results="hide">>=
+library(latex2exp)
+data<-read.csv("/home/ruoyong/paper2/weekday.csv") 
+#exampletwo<-as.matrix(as.data.frame(lapply(data, as.numeric)))
+data<-data[,-1]
+week<-as.matrix(data)
+rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
+colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
+@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
+knitr::kable(head(week, n=7),format="latex",caption = "Week\\label{tab:week}") %>%
+  kable_styling(bootstrap_options = "striped", full_width = F, position = "center",latex_options = "HOLD_position")
+@
 
-\caption{\label{tab:weekdata}Week\label{tab:week}}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
-\hline
-Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
-\hline
-Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
-\hline
-Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
-\hline
-Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
-\hline
-Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
-\hline
-Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
-\hline
-Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
-\hline
-\end{tabular}
-\end{table}
-
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<Timecompareweek2Cpu, cache=TRUE, eval=TRUE>>=
 #using CPU
 fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
-## [1] 0.0001189
 system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
-##    user  system elapsed 
-##  60.412   0.088  60.395
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 The ``week'' table has a much smaller p-value: around 0.000119, which should require a larger number of simulations to get a more accurate p-value.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<Timecompareweek2Gpu, eval=TRUE, cache=FALSE>>=
 #using GPU
@@
             type="double", returnStatistics=FALSE,Nglobal = Nglobal))
-##    user  system elapsed 
-##   0.613   0.010  10.839
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \fct{stats::fisher.test} takes 60.4 seconds, while \fct{fisher.sim} takes about 10.8 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<Timecompareweek2GpuTest, eval=TRUE, cache=TRUE>>=
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
@@
 result_weekgpu$simnum
-## [1] 9994240
 result_weekgpu$counts
-## [1] 1272
 result_weekgpu$p.value
-## [1] 0.0001273734
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 With about ten million simulations, we get 1272 cases and a p-value around 0.0001273. We summarized the results in Table \ref{tab:summary}.
-
+<<Timecompareweek3GpuTest, echo=FALSE, eval=TRUE, cache=TRUE>>=
+result_weekgpu$cpu = as.vector(result_weekgpu$sim)
+@
 \begin{table}[H]
@@
 \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
-\includegraphics[width=0.45\textwidth]{figure/fighistMonth-1} 
-\includegraphics[width=0.45\textwidth]{figure/fighistMonth-2} 
-\end{knitrout}
+<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE>>=
+hist(result_monthgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}))$'), breaks=40, ylab="Frequency", main="Month data")
+abline(v = observed_m, col = "blue", lwd = 1.5)
+
+hist(result_weekgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}))$'), ylab="Frequency", main="Week data")
+abline(v = observed_w, col = "blue", lwd = 1.5)
+@
 \caption{\label{fig4} Frequency distribution for test statistics from month and week data tables.}
@@
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
-
+<<setupData, eval=TRUE,echo=FALSE, message=FALSE>>=
+library("gpuR")
+library("gpuRandom")
+library('geostatsp')
+@
 Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<setupcoords, eval=TRUE>>=
 Ngrid = c(60, 80)
@@
                       ncol(coordsSp@coords), type=theType)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets.
 Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuRandom::maternGpuParam}.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<setup paramsBatch, eval=TRUE,echo=FALSE>>=
+myParamsBatch = 
+rbind(c(shape=1.25, range=0.5, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
+c(shape=2.15, range=0.25, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
+c(shape=0.55, range=1.5, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
+c(shape=2.15, range=0.5, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = -pi/7),
+c(shape=2.15, range=0.5, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/4))
+@
+<<show parameterBatch,eval=TRUE>>=
 myParamsBatch
-##      shape range variance nugget anisoRatio anisoAngleRadians
-## [1,]  1.25  0.50      1.5      0          1         0.0000000
-## [2,]  2.15  0.25      2.0      0          4         0.4487990
-## [3,]  0.55  1.50      2.0      0          4         0.4487990
-## [4,]  2.15  0.50      2.0      0          4        -0.4487990
-## [5,]  2.15  0.50      2.0      0          2         0.7853982
 paramsGpu = gpuRandom::maternGpuParam(myParamsBatch, type=theType)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 Step 3, compute Mat\'ern covariance matrices using \fct{gpuRandom::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
 are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<r outputBatchCreate, eval=TRUE>>=
 maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
@@
 dim(maternCov)
-## [1] 24000  4800
 gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
                        Nglobal=c(64,64,16), Nlocal=c(4,4,4))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 Step 4, the first argument \code{maternCov} in \fct{gpuRandom::cholBatch} specifies the matrix object to take Cholesky decomposition.
 Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<r cholBach, eval=TRUE>>=
 diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
@@
                      Nglobal,Nlocal,NlocalCache)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 % \begin{gather}
@@
 Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{gpuRandom::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
 streamsGpu <- gpuRandom::CreateStreamsGpu(Nstreams=128*64, keepInitial=TRUE)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+@
+<<r randomNormalsSim2, eval=TRUE, cache=FALSE>>=
 zmatGpu = gpuRandom::rnorm(c(nrow(maternCov),2), 
@@
                            type = theType)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 Step 6, use \fct{gpuRandom::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
 \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<r multLowerdiag_LDZ, eval=TRUE, cache=FALSE, results='hide'>>=
 simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
@@
   NlocalCache)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \begin{gather}
@@
 Step 7, Finally, plot the 10 simulated Gaussian random surfaces.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<r plotSimSetup, echo=TRUE, eval=TRUE, cache=TRUE>>=
 library(raster)
@@
                                      'sim', 1:ncol(simMat)), 1, paste, collapse='')
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+@
 \begin{figure}[H]
 \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+<<maternplot, echo=TRUE, fig.show='hold', eval=TRUE, dev='pdf',cache=TRUE>>=
 par(mar=rep(0.1, 4))
@@
 }
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=0.45\textwidth]{figure/maternplot-1} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-2} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-3} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-4} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-5} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-6} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-7} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-8} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-9} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-10} 
-\end{knitrout}
+@
 \caption{\label{fig5} }

### 2021-09-06 pkg/gpuRandom/src/gpuFisher_test.cpp
logfactori and streamsGpu
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2790 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -313,7 +313,7 @@ int gpuFisher_test(
  
   std::string Fisher_kernel_string = FisherSimkernelString<T>(nr, nc, streams.internal_size2());
   if(resultSize >= B) {
-   kernel_string = "\n#define returnResults\n" + kernel_string;
+    Fisher_kernel_string = "\n#define returnResults\n" + Fisher_kernel_string;
   }
    #ifdef DEBUGKERNEL
    Rcpp::Rcout << Fisher_kernel_string << "\n\n";
@@ -323,16 +323,16 @@ int gpuFisher_test(
 
   std::string lfactorialKernelString = logfactString<double>();
   #ifdef DEBUGKERNEL
-   Rcpp::Rcout << logKernelString << "\n\n";
+   Rcpp::Rcout << lfactorialKernelString << "\n\n";
   #endif  
  
  // the context
   viennacl::ocl::context ctx(viennacl::ocl::get_context(ctx_id));  
-  viennacl::ocl::program &my_prog = ctx.add_program(Fisher_kernel_string, "my_kernel");
-  viennacl::ocl::program & my_prog = viennacl::ocl::current_context().add_program(lfactorialKernelString, "my_kernel");
+  viennacl::ocl::program &my_prog_Fisher = ctx.add_program(Fisher_kernel_string, "Fisher_kernel");
+  viennacl::ocl::program & my_prog_lf = viennacl::ocl::current_context().add_program(lfactorialKernelString, "lfactorialKernel");
 
-  viennacl::ocl::kernel &fisher_sim = my_prog.get_kernel("fisher_sim_gpu"); 
-  viennacl::ocl::kernel &lfactorialKernel = my_prog.get_kernel("logfactorial");
+  viennacl::ocl::kernel &fisher_sim = my_prog_Fisher.get_kernel("fisher_sim_gpu"); 
+  viennacl::ocl::kernel &lfactorialKernel = my_prog_lf.get_kernel("logfactorial");
  
  
  
@@ -367,7 +367,7 @@ int gpuFisher_test(
    
    countss = viennacl::linalg::sum(count);
    
-   #ifdef DEBUG
+   #ifdef DEBUGKERNEL
    Rcpp::Rcout << "threshold " << threshold << " countss " << countss << " count0 " << count(0) << " size " << B <<  "\n";
    #endif  
 

### 2021-09-08 pkg/gpuRandom/src/lgmlikFit.hpp
loop for ssqbetahat fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2800 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -188,8 +188,8 @@ void maternBatchVcl(
             Rcpp::IntegerVector Dstartend,  
             Rcpp::IntegerVector Nglobal,
             Rcpp::IntegerVector Nlocal,
-            const int NlocalCache, 
-            const int ctx_id);      
+            const int NlocalCache, //int NrowStartC,
+            const int ctx_id);     
       
         
         template <typename T> 

### 2021-09-08 pkg/gpuRandom/src/likfitGpu_P.cpp

@@ -265,6 +265,10 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
   int DiterIndex, NthisIteration;
   int verboseMatern = verbose[0]>1;
   
+  Rcpp::IntegerVector chol_localSize=localSize;
+  chol_localSize[1]=workgroupSize[1];
+  
+  
   viennacl::ocl::switch_context(ctx_id);
   viennacl::ocl::context ctx(viennacl::ocl::get_context(ctx_id));
   
@@ -287,9 +291,9 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
   int Ncell = Nobs * (Nobs - 1)/2, maxIter = 1500;
   
   std::string maternClString = maternBatchKernelString<T>(
-                 maxIter,
-                 Nobs, Ncell, 
-                 NparamPerIter[0],//NmatrixMax
+    maxIter,
+    Nobs, Ncell, 
+    NparamPerIter[0],//NmatrixMax
                  Vbatch.internal_size2(), //NpadVbatch
                  Vbatch.internal_size2()*Nobs, //NpadBetweenMatrices,
                  coords.internal_size2(), 
@@ -311,7 +315,7 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
     0,//NstartA
     0,// NstartD
     NlocalCache, //Ncache
-    localSize, //Nlocal
+    chol_localSize, //Nlocal
     1,//allowOverflow, // allowoverflow
     1 // do log determinant
   );
@@ -326,7 +330,7 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
     Ndatasets * ssqYX.internal_size2() + Ndatasets,//NstartA start at entry ssqYX[Ndatasets, Ndatasets]
     0,// NstartD
     NlocalCache, //Ncache
-    localSize, //Nlocal
+    chol_localSize, //Nlocal
     1, // allowoverflow
     1 // do log determinant
   );
@@ -417,7 +421,7 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
     0,//const int NstartA,  
     0,//const int NstartD,  
     ssqBetahat.internal_size2(), //const int NpadBetweenMatricesC
-    QinvSsqYx.internal_size2()*Ndatasets, //const int NpadBetweenMatricesA,
[…]

### 2021-09-08 pkg/gpuRandom/src/multiplyLowerTri.cpp
backsolve bugs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2808 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,5 +1,5 @@
 #include "gpuRandom.hpp"
-//#define DEBUG
+#define DEBUG
 
 // C = A B, A diagonal
 
@@ -156,7 +156,7 @@ result +=  "	__global "+ typeString+ " *B) {\n\n" +
   "int AHere, BHere=0, CHere, BcacheHere=0;\n"
   "local "+ typeString+ " Acache[NlocalCacheA];\n" 
   "local "+ typeString+ " Bcache[NlocalCache];\n"
-  "int Dmatrix, Drow, Dcol, Dinner, DrowNpadC, DrowNpadA, DcolInCache, DinnerCache;\n"
+  "int Dmatrix, Drow, DrowBlock, Dcol, DcolBlock, Dinner, DrowNpadC, DrowNpadA, DcolInCache, DinnerCache;\n"
   "const int incInCacheLocal = get_local_size(0)*NcolInCache;\n"
   "const int incInCacheGlobal = get_global_size(0)*NcolInCache;\n"
   "const int doCacheA = (get_local_id(1) == 0);\n";
@@ -183,16 +183,18 @@ if(NpadD) {
   
   result += "\n  // cache first rows of B for Dmatrix\n";
   result +=  
-    "  for(Drow = get_local_id(0),BcacheHere = get_local_id(0)*NcolInCache;\n"
-    "      Drow < DinnerStop; Drow += get_local_size(0),BcacheHere += incInCacheLocal){\n"
-    
+    "  for(DrowBlock = 0,BcacheHere = get_local_id(0)*NcolInCache;\n"
+    "      DrowBlock < DinnerStop; DrowBlock += get_local_size(0),BcacheHere += incInCacheLocal){\n"
+    "    Drow = DrowBlock + get_local_id(0);\n"
     "    DrowNpadA = BHere+Drow*NpadB;\n"
     "    barrier(CLK_LOCAL_MEM_FENCE);\n";
   
   if(NpadD) {
-    result +=  "    if(doCacheA) {\n"
+    result +=  "    if(doCacheA & Drow < DinnerStop) {\n"
     "      Dcache[get_local_id(0)] = " +
       transformD + "(D[DHere + Drow]);\n"
+    "    } else {\n"
+    "      Dcache[get_local_id(0)] =0;\n"
     "    }\n"
     "    barrier(CLK_LOCAL_MEM_FENCE);\n";
   }
@@ -216,12 +218,15 @@ result +=
 result += 
   
 "\n// looped through rows which are all cached\n"
[…]

### 2021-09-10 pkg/gpuRandom/R/RcppExports.R
add matrix element divide vector cppfile
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2819 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -75,8 +75,8 @@ likfitGpu_Backend <- function(coordsGpuR, bigparamsBatchR, yXR, betasR, bigvaria
     invisible(.Call('_gpuRandom_likfitGpu_Backend', PACKAGE = 'gpuRandom', coordsGpuR, bigparamsBatchR, yXR, betasR, bigvariancesR, jacobianR, finalssqBetaR, finalssqXR, finalssqYR, finallogDR, finallogPR, finalbetahatR, finalLogLikR, n, p, groupsize, colbatch, form, workgroupSize, localSize, NlocalCache, verbose))
 }
 
-likfitGpu_BackendP <- function(yx, coords, params, boxcox, betas, ssqY, ssqBetahat, detVar, detReml, jacobian, NparamPerIter, workgroupSize, localSize, NlocalCache, verbose, ssqYX, XVYXVX, ssqYXcopy, LinvYX, QinvSsqYx, cholXVXdiag, varMat, cholDiagMat) {
-    invisible(.Call('_gpuRandom_likfitGpu_BackendP', PACKAGE = 'gpuRandom', yx, coords, params, boxcox, betas, ssqY, ssqBetahat, detVar, detReml, jacobian, NparamPerIter, workgroupSize, localSize, NlocalCache, verbose, ssqYX, XVYXVX, ssqYXcopy, LinvYX, QinvSsqYx, cholXVXdiag, varMat, cholDiagMat))
+likfitGpu_BackendP <- function(yx, coords, params, boxcox, betas, ssqY, XVYXVX, ssqBetahat, detVar, detReml, jacobian, NparamPerIter, workgroupSize, localSize, NlocalCache, verbose, ssqYX, ssqYXcopy, LinvYX, QinvSsqYx, cholXVXdiag, varMat, cholDiagMat) {
+    invisible(.Call('_gpuRandom_likfitGpu_BackendP', PACKAGE = 'gpuRandom', yx, coords, params, boxcox, betas, ssqY, XVYXVX, ssqBetahat, detVar, detReml, jacobian, NparamPerIter, workgroupSize, localSize, NlocalCache, verbose, ssqYX, ssqYXcopy, LinvYX, QinvSsqYx, cholXVXdiag, varMat, cholDiagMat))
 }
 
 logfactsumBackend <- function(xR, numWorkItems) {
@@ -85,6 +85,10 @@ logfactsumBackend <- function(xR, numWorkItems) {
 
[…]

### 2021-09-13 pkg/gpuRandom/src/cholBatch.cpp
fixing chol
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2832 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -50,9 +50,9 @@ template <typename T> std::string cholBatchKernelString(
       ",\n	__global " + typeString + " *logDet,\n"
       "                int logDetIndex\n";
     }
+
     
   result += "\n){\n"
-  " barrier(CLK_LOCAL_MEM_FENCE);\n"
   " const int localIndex = get_local_id(0)*get_local_size(1) + get_local_id(1);\n"
   " const int localIndexIsZero = (localIndex==0);\n"
   " const int NlocalTotal = get_local_size(0)*get_local_size(1);\n"
@@ -64,7 +64,7 @@ template <typename T> std::string cholBatchKernelString(
   " int Drow, DrowBlock, Dk, Dmatrix, DmatrixBlock;\n";
   
   if(allowOverflow) {
-    result += "	int minDcolNcache;\n";
+    result += "int minDcolNcache;\n";
   }
   
   result += " " +  typeString + " DL;\n" 
@@ -90,6 +90,8 @@ result +=
     "  DcolNpad = Dcol*Npad;\n"
   "  AHereDcol = AHere+DcolNpad;\n"
   //"  Dcolm1 = Dcol - 1;\n"
+  "  DL = 0.0;\n"
+  "  diagLocal[localIndex]=0.0;\n"
   "  toAddLocal[localIndex]=0.0;\n";
   
   

### 2021-09-14 pkg/gpuRandom/src/cholBatch.cpp
matern and chol done
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2835 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,5 +1,5 @@
 #include "gpuRandom.hpp"
-#define DEBUG
+//#define DEBUG
 
 // dimension 0 is row, local dimension 1 inner loop over columns, group 0 is matrix
 // local and global work sizes should be identical for dimension 1 (second dimension), only 1 group for dimension 1

### 2021-09-14 pkg/gpuRandom/src/multiplyLowerTri.cpp
matern and chol done
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2835 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -160,7 +160,8 @@ result +=  "	__global "+ typeString+ " *B) {\n\n" +
   "int AHere, BHere=0, CHere, BcacheHere=0;\n"
   "local "+ typeString+ " Acache[NlocalCacheA];\n" 
   "local "+ typeString+ " Bcache[NlocalCache];\n"
-  "int Dmatrix, DmatrixBlock, Drow, DrowBlock, DrowBlockEnd, Dcol, DcolBlock, Dinner, DrowNpadC, DrowNpadA;\n"
+  "int Dmatrix, DmatrixBlock, Drow, DrowBlock, DrowBlockEnd,\n"
+  "    Dcol, DcolBlock, Dinner, DrowNpadC, DrowNpadA;\n"
   "int DcolInCache, DinnerCache;\n"
   "const int incInCacheLocal = get_local_size(0)*NcolInCache;\n"
   "const int incInCacheGlobal = get_global_size(0)*NcolInCache;\n"
@@ -190,13 +191,15 @@ if(NpadD) {
   result += "\n  // cache first rows of B for Dmatrix\n";
   result +=  
     "  for(DrowBlock = 0,BcacheHere = get_local_id(0)*NcolInCache;\n"
-    "      DrowBlock < DinnerStop; DrowBlock += get_local_size(0),BcacheHere += incInCacheLocal){\n"
+    "      DrowBlock < DinnerStop;\n"
+    "      DrowBlock += get_local_size(0),BcacheHere += incInCacheLocal){\n"
     "    Drow = DrowBlock + get_local_id(0);\n"
-   "    DrowNpadA = BHere+Drow*NpadB;\n"
+    "    DrowNpadA = BHere+Drow*NpadB;\n"
     "    barrier(CLK_LOCAL_MEM_FENCE);\n";
   
   if(NpadD) {
-    result +=  "    if(doCacheA & (Drow < DinnerStop) & (Dmatrix < Nmatrix) ) {\n"
+    result +=  
+    "    if(doCacheA & (Drow < DinnerStop) & (Dmatrix < Nmatrix) ) {\n"
     "      Dcache[get_local_id(0)] = " +
       transformD + "(D[DHere + Drow]);\n"
     "    } else {\n"
@@ -205,30 +208,34 @@ if(NpadD) {
     "    barrier(CLK_LOCAL_MEM_FENCE);\n";
   }
     result +=     
-      "    for(DcolBlock = 0, DcolInCache=get_local_id(1);\n"
[…]

### 2021-09-14 pkg/gpuRandom/src/cholBatch.cpp
multiploy
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2844 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -81,7 +81,7 @@ result +=
   "     Dmatrix+= get_num_groups(0),DmatrixBlock+= get_num_groups(0)){\n"
  
   " diagHere = Dmatrix*NpadDiag + NstartD;\n"
-  " AHere = Dmatrix*NpadBetweenMatrices + NstartA;\n";
+  " AHere = (Dmatrix < Nmatrix) * Dmatrix*NpadBetweenMatrices + NstartA;\n";
   if(logDet){
     result += " logDetHere = 0.0;\n";
   }

### 2021-09-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
add paper1 to inst
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2851 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\title{Random number simulation and applications on GPU’s in R?}
+\title{Random number generation and applications on GPU’s in R?}
 \Plaintitle{}
@@
 ##    user  system elapsed 
-##   0.112   0.001   1.279
+##   0.102   0.008   1.278
 system.time(matrix(stats::rnorm(16384*16384),16384,16384))
 ##    user  system elapsed 
-##  11.962   0.898  12.838
+##  12.028   0.857  12.862
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.680   0.008   2.291
+##   0.672   0.012   2.287
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.613   0.010  10.839
+##   0.598   0.017  10.831
 \end{verbatim}

### 2021-09-16 pkg/gpuRandom/src/multiplyLowerTri.cpp
multiply tri fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2854 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -122,7 +122,8 @@ std::string multiplyLowerBatchString(
   ) {
 
   std::string typeString = openclTypeString<T>();
-  std::string result = "", cacheBstring="", DhereString="", DhereStringZero="";
+  std::string result = "", cacheBstring="", DhereString="", DhereStringZero="",
+    checkString = "";
   const int rowsToCache = std::floor(NlocalCacheB/Ncol);
 
   if(typeString == "double") {
@@ -165,9 +166,8 @@ result +=  "	__global "+ typeString+ " *B) {\n\n" +
   "int DcolInCache, DinnerCache, DmatrixInBounds, DrowInBounds, DcolInBounds;\n"
   "const int incInCacheLocal = get_local_size(0)*NcolInCache;\n"
   "const int incInCacheGlobal = get_global_size(0)*NcolInCache;\n"
-  "const int localSize0m1  = get_local_size(0) - 1;\n"
-  "const int localSize1m1  = get_local_size(1) - 1;\n"
-  "int doCacheAhere;\n"
+//  "const int localSize0m1  = get_local_size(0) - 1;\n"
+//  "const int localSize1m1  = get_local_size(1) - 1;\n"
   "const int doCacheA = (get_local_id(1) == 0);\n";
 
 if(NpadD) {
@@ -273,7 +273,8 @@ result +=
 result += 
   "    Drow = DrowBlock + get_global_id(0);\n";
 
-result +=  "     DrowBlockEnd = DrowBlock + get_local_size(0)-1;\n";
+result +=  
+  "     DrowBlockEnd = DrowBlock + get_global_size(0)-1;\n";
 
 result += 
   "    if(DmatrixInBounds & (Drow < DinnerStop) ){\n"
@@ -308,26 +309,30 @@ if(diagIsOne) {
     "      } else {\n"
     "        Dout = 0.0;\n"
     "      }\n";
+  checkString = "<";
 } else { 
   result +=   "      Dout = 0.0;\n";
+  checkString = "<=";
 }
 
 result +=
-  "      for(Dinner = 0,DinnerCache=0; Dinner < DrowBlockEnd; Dinner++,DinnerCache += NcolInCache){\n";
-
[…]

### 2021-09-17 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
add paper1 to inst
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2851 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
updates paper1 in inst folder
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2855 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{figure}
-Let's compare the time of generating a large-size matrix of more than 26.8 million Gaussian random numbers of double precision using the \fct{stats::rnorm} and using the \fct{gpuRandom::rnorm}.
-Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $128 \times 128$ work-items is about 10 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
+Let's compare the time of generating a large-size matrix of 100 million Gaussian random numbers of double precision using the \fct{stats::rnorm} and using the \fct{gpuRandom::rnorm}.
+Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is 117.6 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size goes larger.
-<<random normal time compare, eval=TRUE, cache=FALSE>>=
-system.time(gpuRandom::rnorm(c(16384,16384), streams=streams, 
+<<random normal time compare, eval=TRUE, cache=TRUE>>=
+system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
                              Nglobal=c(512,128), type="double"))
-system.time(matrix(stats::rnorm(16384*16384),16384,16384))
+system.time(matrix(stats::rnorm(10000^2),10000,10000))
 @
@@
 
-Next we compare the run time between using \fct{gpuRandom::fisher.sim} and \fct{stats::fisher.test} using two real data examples with small and large p-values respectively.
+We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and a normal GPU, the other is equipped with an excellent GPU and a normal CPU.
+The device information is included in the table \ref{tab:device}.
+The \proglang{R} outputs for testing on computer 1 is shown in the following sections.
+\begin{table}[H]
+\begin{tabular}{ |c|c|c|c|c|c| } 
+\hline
+           & CPU   & GPU \\ \hline
+ Computer1 & pthread-Intel(R) Xeon(R) W-2145 CPU @ 3.70GHz & 2 Radeon VII's : ROCm AMDGPU  \\ \hline
+ Computer2 & dontknow & GRID V100D-8C \\ \hline
+\end{tabular}
+\caption{\label{tab:device} Device information for run-time compare.}
+\end{table}
+
 These 2-way contingency tables (\ref{tab:month} and \ref{tab:week}) consists of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
@@
 <<Timecomparemonth,echo=FALSE,eval=TRUE, results="hide">>=
-data2<-read.csv("/home/ruoyong/paper2/month.csv") 
+data2<-read.csv("/home/ruoyong/paper1_2021_5/month.csv") 
 data2<-data2[,-1]
@@
 system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = Nglobal))
+            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
 @
 \fct{gpuRandom::fisher.sim} takes about 2.3 seconds, compared to the elapsed time of \fct{gpuRandom::fisher.sim},  \fct{gpuRandom::fisher.sim} takes less than a quater of the time. 
-<<timecomparemonth3Gpu, eval=TRUE,cache=TRUE>>=
+<<timecomparemonth3Gpu, eval=TRUE,cache=FALSE>>=
 result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = Nglobal)
+                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
 result_monthgpu$simnum
@@
 @
-We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.4035, which is close enough to the p-value from \fct{stats::fisher.test}: 0.4035.
+We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.4035, which is close enough to the p-value from \fct{stats::fisher.test}: 0.4034.
 <<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
@@
 library(latex2exp)
-data<-read.csv("/home/ruoyong/paper2/weekday.csv") 
+data<-read.csv("/home/ruoyong/paper1_2021_5/weekday.csv") 
 #exampletwo<-as.matrix(as.data.frame(lapply(data, as.numeric)))
@@
 @
-
-<<Timecompareweek2Cpu, cache=TRUE, eval=TRUE>>=
+We perform 10 million simulations for the ``week'' table.
+<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
 #using CPU
@@
 The ``week'' table has a much smaller p-value: around 0.000119, which should require a larger number of simulations to get a more accurate p-value.
-<<Timecompareweek2Gpu, eval=TRUE, cache=FALSE>>=
+<<TimecompareweekGpu, eval=TRUE, cache=FALSE>>=
 #using GPU
@@
 \fct{stats::fisher.test} takes 60.4 seconds, while \fct{fisher.sim} takes about 10.8 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
-<<Timecompareweek2GpuTest, eval=TRUE, cache=TRUE>>=
+<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
@@
 @
-With about ten million simulations, we get 1272 cases and a p-value around 0.0001273. We summarized the results in Table \ref{tab:summary}.
-<<Timecompareweek3GpuTest, echo=FALSE, eval=TRUE, cache=TRUE>>=
+With about ten million simulations, we get 1272 cases and a p-value around 0.0001274. We summarized the results in Table \ref{tab:summary}.
+<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
 result_weekgpu$cpu = as.vector(result_weekgpu$sim)
@@
 \begin{table}[H]
+\centering
 \begin{tabular}{ |c|c|c|c|c|c| } 
  \hline
- test example & B & p-value stats & p-value gpuRandom & time stats & time gpuRandom \\ \hline
- Month & 1e6 & 0.4029 & 0.4035 & 9.8 & 2.3\\ \hline
- Week & 1e7 & 0.000119 & 0.000127 & 60.4 & 10.8\\ \hline
+ data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
+ Month & 1e6 & 0.4038 & 0.403507  & 0.403507     \\ \hline
+ Week &  1e7  & 0.0001208 & 0.0001274 & 0.0001274\\ \hline
 \end{tabular}
-\caption{\label{tab:summary} Comparison summary of \fct{gpuRandom::fisher.sim} and \fct{stats::fisher.test}.}
+\caption{\label{tab:summary} Summary of p-values obtained on different devices.}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{ |c|c|c|c|c|c| } 
+ \hline
+ data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
+ Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
+ Week &  1e7  & 60.19 & 90.19 & 10.82 & 1.90\\ \hline
+\end{tabular}
+\caption{\label{tab:summary} Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.}
 \end{table}
@@
 Nlocal = c(4, 2, 2)
-theType = "float"
-myRaster = raster(
-  extent(0,Ngrid[1]/Ngrid[2],5,6),
-  Ngrid[2], Ngrid[1]
-)
-coordsSp = sp::SpatialPoints(xyFromCell(myRaster, 1:ncell(myRaster)))
+theType = "double"
+
+myRaster = raster::raster( raster::extent(0,Ngrid[1]/Ngrid[2],5,6),Ngrid[1], Ngrid[2])
+coordsSp = sp::SpatialPoints(raster::xyFromCell(myRaster, 1:raster::ncell(myRaster)))
 coordsGpu = vclMatrix(coordsSp@coords, 
@@
 gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
-                       Nglobal=c(64,64,16), Nlocal=c(4,4,4))
+                       Nglobal=c(128,64), Nlocal=c(16,4))
 @
@@
 gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
-                     Nglobal,Nlocal,NlocalCache)
+                     Nglobal= c(512, 8),
+                     Nlocal= c(32, 8),
+                     NlocalCache=2000)
 @
@@
 Step 7, Finally, plot the 10 simulated Gaussian random surfaces.
-<<r plotSimSetup, echo=TRUE, eval=TRUE, cache=TRUE>>=
+<<r plotSimSetup, echo=TRUE, eval=TRUE, cache=FALSE>>=
 library(raster)
@@
 \centering
-<<maternplot, echo=TRUE, fig.show='hold', eval=TRUE, dev='pdf',cache=TRUE>>=
+<<maternplot, echo=TRUE, fig.show='hold', eval=TRUE, dev='pdf',cache=FALSE>>=
 par(mar=rep(0.1, 4))

### 2021-09-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
add paper1 draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2865 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{listings}
+\usepackage[shortcuts]{extdash}
 \renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }
@@
 \subsection{Creating streams}
-In \proglang{clRNG} library, a stream object is a structure that contains three states (see the \proglang{R} output below): the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream.
-The \proglang{clRNG} function \fct{clrngMrg31k3pCreateStreams} creates streams on the host computer using the MRG31k3p RNG.
+In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
+The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
 If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory.
@@
 \fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams.
-\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on the GPU device.
+\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device.
 From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used.
-Below, we show the usage of these two functions by generating 4 streams on the host and on the device respectively.
+In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
 \begin{knitrout}
@@
 \fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create.
-For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) has a default value ``12345''.
-By setting a seed identical to the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as the streams created on host as follows,
+For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
+By setting a seed same as the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
 \begin{knitrout}
@@
 \end{itemize}
-Streams in \pkg{gpuRandom} are like \fct{.Random.seed} in \proglang{R}.
+\fct{CreateStreamsGpu} in \pkg{gpuRandom} are similar to \code{.Random.seed} in \proglang{R}.
 By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
@@
 \subsection{Using streams to generate  i.i.d. $U (0, 1)$ random numbers on GPU}
-The streams created on GPU are then used in work items that executes in parallel on GPU device, in \pkg{gpuRandom} each work item takes one distinct stream to generate random numbers, so number of streams should equal to the number of total work items in use.
-A part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown below.
-Kernels are written in \proglang{OpenCL C} language, \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
-Pointers to the \code{streams} and to the matrix \code{out} on the global memory are passed to the kernel as arguments.
+The streams created on GPU are then used in work items that executes in parallel on GPU device.
+In \pkg{gpuRandom} each work item takes one distinct stream to generate random numbers, it is important to notice that the number of streams should always be equal to the number of total work items in use.
+The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
+Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
+Here the pointers to \code{streams} and to the matrix \code{out} on the global memory are passed to the kernel as arguments.
 \code{Nrow} and \code{Ncol} represent number of rows and columns of the matrix \code{out} respectively.
 \code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is internal number of columns of the \code{streams} matrix (these variables are defined in macros not shown here).
-Each stream is taken to the private memory of each work item by the function \fct{streamsToPrivate}.
-Here we use a 2-dimensional index space with work-group size $1 \times 2$.
-\code{index} represents the position of the work item when mapping it to a one-dimensional vector.
-The function \fct{clrngMrg31k3pNextState} gives a uniform random integer named \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
+Each stream's current state is copied to the private memory of each work item by the function \fct{streamsToPrivate}.
+\code{index} represents the position of the work item when mapping it to a one-dimensional data structure.
+The function \fct{clrngMrg31k3pNextState} generates a uniform random integer named \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
 At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label=uniform]
+\begin{lstlisting}[caption={captiontext},language=C,basicstyle=\small,label={lst:uniformkernel}]
  __kernel void mrg31k3pMatrix(
@@
 uint g1[3], g2[3];
-
 streamsToPrivate(streams,g1,g2,startvalue);
@@
  }//Drow
-
 streamsFromPrivate(streams,g1,g2,startvalue);
@@
 \end{framed}
-
-The kernel code may not be straightforward enough in showing how streams are used in work items to generate random numbers, so here we give a small example to illustrate the matrix filling process.
-As shown in Figure \ref{fig1}, suppose we use a $2 \times 4$ NDRange (index space size) ((a) in Figure \ref{fig1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig1}) of uniform random numbers using 8 streams.
-In each grid of (a) that represents a work item, there are two rows of numbers, the number in the above row is the stream index (from 0 to 7), and a pair of numbers in the second row is the work item index (each work item has a unique 2-dimensional global ID).
+This kernel executes over a 2-dimensional index space.
+Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example.
+Suppose we have a $2 \times 4$ NDRange (index space size) ((a) in Figure \ref{fig:1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig:1}) of uniform random numbers using 8 streams.
+In each grid of (a) that represents a work item, there are two rows of numbers, the number in the above row is the stream index (from 0 to 7), and a pair of numbers in the second row is the 2-dimensional global ID of the work item.
 Each stream object is passed to each work item.
-The number in a cell of matrix in (b) is the stream index, that indicates which stream does the random number generation for this matrix cell.
-For example, stream 0 which is allocated in work item $(0,0)$ generates a random number in matrix $(0,0)$ (row and column index of the matrix) in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it goes to matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
-
+The number in a cell of the matrix in (b) corresponds to the stream, it indicates which stream does the random number generation for this matrix cell.
+For example, stream 0 which is allocated in work item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
 \begin{figure}[H]%
@@
     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f2} }}%
-    \caption{name?\label{fig1}}%
+    \caption{name?\label{fig:1}}%
 \end{figure}
 
-In the following code, we use the 4 streams created in Section 2.1 and 4 work items to generate a vector \code{sim_1} of 6 i.i.d.
-$U (0,1)$ random numbers by the function \fct{gpuRandom::runif}.
-By default \code{Nglobal} is set as (64,8), and a matrix of \code{prod(Nglobal)} streams created on GPU is supplied.
-The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host computer.
+Now we use the 4 streams created in Section 2.1 and an NDRange of 1 by 4 work items to generate a vector \code{sim_1} of 6 i.i.d.
+$U (0,1)$ random numbers using the function \fct{gpuRandom::runif}.
+If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied.
+The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
 \begin{knitrout}
@@
 \begin{itemize}
-  \item \code{n}: a number specifying the length if to create a vector, or a vector of length 2 specifying the row and column number if to create a matrix.
-  \item \code{streams}: streams used for random number generating.
+  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or specifying the length if to create a vector.
+  \item \code{streams}: streams used for random number generation.
   \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
@@
 \end{itemize}
-Next we show that using the same streams (\code{streams1} and \code{streams2}),  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
+Next we show that by using a same stream \code{streams2} as \code{streams1},  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
 \begin{knitrout}
@@
 \end{knitrout}
-Note that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used, reusing the \code{streams1} will produce a vector  different from \code{sim_1} (result of \code{sim_3}), because each time a random number is generated, the current state of the stream is changed (advances by one position).
-See below the first six rows (current states) of \code{streams1} after it being used.
+Notice that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used.
+Reusing \code{streams1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
+Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{streams1} after it being used.
 \begin{knitrout}
@@
 \end{knitrout}
-In order to reproduce the same sequence of random numbers, we can save the streams in a file on CPU, because GPU device does not keep memory for streams objects after each execution.
-Below is the \proglang{R} code in which we use the saved streams \code{streams_from_GPU.rds} to produce a vector \code{sim_saved}, which is also same as \code{sim_1}.
+Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart.
+Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
+However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations.
+Below is a trivial example, in which we demonstrate saving \code{streams1} to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU to produce the vector called \code{sim_saved}, which is also the same as \code{sim_1}.
 \begin{knitrout}
@@
 \end{knitrout}
-
 \begin{knitrout}
@@
 \end{knitrout}
-We evaluate the distribution of 50000 GPU-generated uniform random numbers by making a density histogram  (the left panel of Figure \ref{fig0}), plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel of Figure \ref{fig0}).
-The two lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} should come from a standard uniform distribution.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streams <- CreateStreamsGpu(Nstreams =256, keepInitial=1)
+
+We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
+The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} does come from a standard uniform distribution.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+streams <- CreateStreamsGpu(Nstreams =512, keepInitial=1)
 random_vector<-gpuRandom::runif(n=50000, streams=streams, 
-                                Nglobal=c(64,4), type="double")
-random_vector<-as.vector(as.matrix(random_vector))
+                                Nglobal=c(64,8), type="double")
+random_vector<-as.vector(random_vector)
 \end{verbatim}
@@
 \end{knitrout}
-\caption{\label{fig0} histogram and empirical cumulative distribution of the GPU-generated random numbers.}
+\caption{histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
 \end{figure}
 
-\subsection{Using streams to generate i.i.d. $N (0, 1)$ random numbers on GPU}
+\subsection{Using streams to generate i.i.d. N (0, 1) random numbers on GPU}
 We apply the Box-Muller transformation to $U (0,1)$ random numbers to generate Gaussian random numbers.
-Box-Muller algorithm (see algorithm \ref{algorithm1}) takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables.
+As shown in Algorithm \ref{algorithm1}, Box-Muller algorithm takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables.
 The derivation basically uses transformation from Cartesian coordinates to polar coordinates to represent two independent, Normally distributed random variables.
@@
 \end{algorithm*}
-%
-Below is a fragment of the kernel function that generates standard Gaussian random numbers. 
+
+Listing \ref{lst:normal} shows a fragment of the kernel that generates standard Gaussian random numbers.
+Aside from using local memory and the part that does the Box-Muller transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}.
+If developers want to use other Gaussian transform methods, this kernel is easy to be modified and incorporated in other \proglang{R} packages.
+The kernel executes over a 2-dimensional index space of $1 \times 2$ work-groups.
+\code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
+As the work-items in each work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), (X_2,Y_2), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
 %
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label=normal]
+\begin{lstlisting}[language=C,basicstyle=\small,caption={name},label={lst:normal}]
+const int 
+index = get_global_id(0)*get_global_size(1) + get_global_id(1);
+const int 
+DrowInc = 2*get_global_size(0), DrowStartInc = DrowInc * NpadCol;
+int Drow, Dcol, DrowStart, Dentry;
+uint temp;
+uint g1[3], g2[3];
+const int startvalue=index * NpadStreams;
+
+local double part[2], cosPart1, sinPart1;
+
+streamsToPrivate(streams,g1,g2,startvalue);
+
+// Iterate for the rows
 for(Drow=2*get_global_id(0), 
-    DrowStart=(Drow + get_local_id(1))* NpadCol;
+    DrowStart = (Drow + get_local_id(1))* NpadCol;
     Drow < Nrow; 
-    Drow += DrowInc, DrowStart += DrowStartInc){
-
+    Drow += DrowInc, DrowStart += DrowStartInc) {
+  // Iterate for the columns
   for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
       Dcol < Ncol;
-      Dcol += get_num_groups(1), Dentry += get_num_groups(1)){
-
+      Dcol += get_num_groups(1), Dentry += get_num_groups(1)) {
+
+    // Generates an uniform random integer
     temp = clrngMrg31k3pNextState(g1, g2);
@@
     if(get_local_id(1)) {
+      // Cache data to local memory
       part[1] = TWOPI_mrg31k3p_NORM_cl * temp;
@@
     } else {
-       part[0] = sqrt(-2.0*log(mrg31k3p_NORM_cl * temp));
+      part[0] = sqrt(-2.0*log(temp * mrg31k3p_NORM_cl));
     }
 
+    // Wait until all work items have read the data and 
+    // it becomes visible
     barrier(CLK_LOCAL_MEM_FENCE);
 
+    // Perform the operation and output the data
     if(get_local_id(1)) {
@@
     }
+
     barrier(CLK_LOCAL_MEM_FENCE);
+
   }//Dcol
 }//Drow
- }
+
+streamsFromPrivate(streams,g1,g2,startvalue);
 \end{lstlisting}
 \end{framed}
-The kernel is adjusted from the kernel for generating uniform random numbers in \ref{uniform}, we only add the part that does the Box-Muller transformation.
-If developers want to use other Gaussian transform methods, this kernel code is easy to be modified and used in other \proglang{R} packages.
-In the code, \code{part0} and \code{part1} correspond to $R$ and $\Theta$ in the formulas respectively.
-\code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that $(X_1,Y_1), (X_2,Y_2), \dots, (X_n,Y_n)$ are generated correctly in pairs, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ will not occur.
-We illustrate the matrix filling process with the previous example: to create a $4 \times 6$ matrix ((b) in Figure \ref{fig3}) of Gaussian random numbers using a $2 \times 4$ NDRange ((a) in Figure \ref{fig3}) and 8 streams.
+We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange and 8 streams.
+Work items are organized to be 1 by 2 work groups as Gaussian numbers are generated in pairs using the Box-Muller method.
+The shading on the grids in the NDRange distinguishes work groups.
+Each work group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where work item of `\code{get_local_id(1)}$=0$' generates the \code{X} of a pair, and work item of `\code{get_local_id(1)}$=1$' generates the \code{Y} of a pair.
+The numbers in the third row in each gird of NDRange represents the local index of each work item.
+Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work items generate random numbers in the leftmost two columns (8 cells) of matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
+Therefore, each work item generates one Gaussian random number in each iteration, the matrix filling process is the same as that shown in Figure \ref{fig:1}.
 
@@
 \end{figure}
-As Gaussian numbers are generated in pairs using the Box-Muller method, work items are organized to be $1 \times 2$ work groups in the index space (a), the shading in the grids of (a) distinguishes work groups.
-Each work group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where the work item of `\code{get_local_id(1)}$=0$' generates the \code{X} in a pair, and work item of `\code{get_local_id(1)}$=1$' generates the \code{Y} in a pair.
-The numbers in the third row in each gird of (a) represents the local index of each work item.
-Cells of the matrix are filled from left to right, that is, in the first iteration of execution, the 8 work items generate random numbers in the leftmost two columns (8 cells) of matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration  fills the rest two columns of the matrix simultaneously.
-Therefore, each work item generates one Gaussian random number in each iteration, the filling process is the same as the filling process of Figure \ref{fig1}.
-
-We show the usage of \fct{gpuRandom::rnorm} by generating a $512 \times 1024$ matrix named \code{normal_matrix} of double presicion Gaussian random numbers, and assess Normality by making a histogram (left panel of Figure \ref{fig2}) and a Q-Q plot (right panel of Figure \ref{fig2}) of the generated random numbers.
-The points in the Q-Q plot in Figure \ref{fig2} fall along the diagonal line, we have good evidence the data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part (adapted from \fct{stat::qnorm}) on GPU, which speeds up the Q-Q plot greatly.
+
+In the following code, We create a $1024 \times 512$ matrix named \code{normal_matrix} of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we adapted the \proglang{R} function \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on GPU, which can speed up the Q-Q plot greatly.
 \begin{knitrout}
@@
 normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
-                                Nglobal=c(512,128), type="double")
+                                Nglobal=c(512,128), type="float")
 \end{verbatim}
@@
 \end{knitrout}
-\caption{\label{fig2} Histogram and Q-Q plot of the GPU-generated random numbers.}
+\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
 \end{figure}
-Let's compare the time of generating a large-size matrix of more than 26.8 million Gaussian random numbers of double precision using the \fct{stats::rnorm} and using the \fct{gpuRandom::rnorm}.
-Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $128 \times 128$ work-items is about 10 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
+
+We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the runtime between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
+Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size goes larger.
@@
 \begin{verbatim}
-system.time(gpuRandom::rnorm(c(16384,16384), streams=streams, 
+system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
                              Nglobal=c(512,128), type="double"))
 ##    user  system elapsed 
-##   0.102   0.008   1.278
-system.time(matrix(stats::rnorm(16384*16384),16384,16384))
+##   0.056   0.008   0.064
+system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##  12.028   0.857  12.862
+##   6.932   0.942   7.870
 \end{verbatim}
@@
 
-Next we compare the run time between using \fct{gpuRandom::fisher.sim} and \fct{stats::fisher.test} using two real data examples with small and large p-values respectively.
-These 2-way contingency tables (\ref{tab:month} and \ref{tab:week}) consists of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
+We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the runtime with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
+The device information is included in table \ref{tab:device}.
+The \proglang{R} outputs for testing on computer 2 is shown in the following.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+
+\caption{\label{tab:deviceinfo}Device information for run-time comparison.\label{tab:device}}
+\centering
+\begin{tabular}[t]{c|c|c}
+\hline
+ & Host & Device\\
+\hline
+Computer1 & Intel Xenon W-2145 3.7Ghz & AMD Radeon VII\\
+\hline
+Computer2 & VCPU Intel Xenon Skylake 2.5Ghz & VGPU Nvidia Tesla V100\\
+\hline
+\end{tabular}
+\end{table}
+
+\end{knitrout}
+% \begin{table}[H]
+% \begin{tabular}{ |c|c|c|c|c|c| } 
+% \hline
+%            & host   &                     device \\ \hline
+%  Computer1 & INtel Xenon W-2145 3.7 Ghz & Amd Radeon VII  \\ \hline
+%  Computer2 & VCPU Intel Xenon Skylake 2.5ghz & VGPU Nvidia Tesla V100 \\ \hline
+% \end{tabular}
+% \caption{ Device information for run-time compare.\label{tab:device}}
+% \end{table}
+
+The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
 The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
@@
 \end{knitrout}
+
 \begin{knitrout}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
-## [1] 0.4035866
+## [1] 0.4038036
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
 ##    user  system elapsed 
-##   9.753   0.004   9.740
+##  14.591   0.006  14.590
 \end{verbatim}
@@
 \begin{verbatim}
-#using GPU
 almost.1 <- 1 + 64 * .Machine$double.eps
 observed_m= -gpuRandom::logfactSum(month, c(16,16))/almost.1
+
 month_GPU<-vclMatrix(month,type="integer")
+result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
+                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
+result_monthgpu$simnum
+## [1] 1015808
+result_monthgpu$counts
+## [1] 410098
+result_monthgpu$p.value
+## [1] 0.4037166
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+#using GPU
 system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = Nglobal))
+            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
 ##    user  system elapsed 
-##   0.672   0.012   2.287
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{gpuRandom::fisher.sim} takes about 2.3 seconds, compared to the elapsed time of \fct{gpuRandom::fisher.sim},  \fct{gpuRandom::fisher.sim} takes less than a quater of the time. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = Nglobal)
-result_monthgpu$simnum
-## [1] 999424
-result_monthgpu$counts
-## [1] 403274
-result_monthgpu$p.value
-## [1] 0.403507
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.4035, which is close enough to the p-value from \fct{stats::fisher.test}: 0.4035.
+##   0.355   0.006   0.361
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.
 
@@
 \end{knitrout}
-
+We perform 10 million simulations for the ``week'' table.
 \begin{knitrout}
@@
 fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
-## [1] 0.0001189
+## [1] 0.0001208
 system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
 ##    user  system elapsed 
-##  60.412   0.088  60.395
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-The ``week'' table has a much smaller p-value: around 0.000119, which should require a larger number of simulations to get a more accurate p-value.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using GPU
+##  90.024   0.080  90.069
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
 observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
+
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
-##    user  system elapsed 
-##   0.598   0.017  10.831
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{stats::fisher.test} takes 60.4 seconds, while \fct{fisher.sim} takes about 10.8 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
@@
 result_weekgpu$counts
-## [1] 1272
+## [1] 1223
 result_weekgpu$p.value
-## [1] 0.0001273734
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-With about ten million simulations, we get 1272 cases and a p-value around 0.0001273. We summarized the results in Table \ref{tab:summary}.
-
-\begin{table}[H]
-\begin{tabular}{ |c|c|c|c|c|c| } 
- \hline
- test example & B & p-value stats & p-value gpuRandom & time stats & time gpuRandom \\ \hline
- Month & 1e6 & 0.4029 & 0.4035 & 9.8 & 2.3\\ \hline
- Week & 1e7 & 0.000119 & 0.000127 & 60.4 & 10.8\\ \hline
+## [1] 0.0001224705
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+#using GPU
+system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
+            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
+##    user  system elapsed 
+##   1.899   0.000   1.898
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.
+
+\subsection{A summary of the results}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+
+\caption{\label{tab:summarycompare}Summary of comparions of Fisher's test simulation on different devices.\label{tab:summary}}
+\centering
+\begin{tabular}[t]{l|l|l|l|l|l}
+\hline
+\multicolumn{1}{c|}{ } & \multicolumn{2}{c|}{Computer 1} & \multicolumn{2}{c|}{Computer 2} & \multicolumn{1}{c}{ } \\
+\cline{2-3} \cline{4-5}
+B & Intel 2.5ghz & AMD Radeon & Intel 3.7ghz & NVIDIA V100 & Data\\
+\hline
+\multicolumn{6}{l}{\textbf{P-value}}\\
+\hline
+\hspace{1em}1m & 0.403804 & 0.403507 & 0.4035606 & 0.403507 & month\\
+\hline
+\hspace{1em}10m & 0.0001251 & 0.0001274 & 0.0001202 & 0.0001274 & week\\
+\hline
+\multicolumn{6}{l}{\textbf{Run-time}}\\
+\hline
+\hspace{1em}1m & 9.74 & 2.28 & 14.58 & 0.32 & month\\
+\hline
+\hspace{1em}10m & 60.19 & 10.82 & 90.38 & 1.89 & week\\
+\hline
 \end{tabular}
-\caption{\label{tab:summary} Comparison summary of \fct{gpuRandom::fisher.sim} and \fct{stats::fisher.test}.}
 \end{table}
+
+\end{knitrout}
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ |c|c|c|c|c|c| } 
+%  \hline
+%  data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
+%  Month & 1e6 & 0.403804 & 0.403507  & 0.403507     \\ \hline
+%  Week &  1e7  & 0.0001251 & 0.0001274 & 0.0001274\\ \hline
+% \end{tabular}
+% \caption{ Summary of p-values obtained on different devices.\label{tab:summary}}
+% \end{table}
+% 
+% 
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ |c|c|c|c|c|c| } 
+%  \hline
+%  data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
+%  Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
+%  Week &  1e7  & 60.19 & 90.38 & 10.82 & 1.90\\ \hline
+% \end{tabular}
+% \caption{ Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.\label{tab:compare time}}
+% \end{table}
 
@@
 \end{knitrout}
-\caption{\label{fig4} Frequency distribution for test statistics from month and week data tables.}
+\caption{ Frequency distribution for test statistics from month and week data tables.\label{fig4}}
 \end{figure}
@@
 gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
-                       Nglobal=c(64,64,16), Nlocal=c(4,4,4))
+                       Nglobal=c(128,64), Nlocal=c(16,4))
 \end{verbatim}
@@
 gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
-                     Nglobal,Nlocal,NlocalCache)
+                     Nglobal= c(512, 8),
+                     Nlocal= c(32, 8),
+                     NlocalCache=2000)
 \end{verbatim}
@@
 \end{knitrout}
-\caption{\label{fig5} }
+\caption{\label{fig:5} }
 \end{figure}

### 2021-09-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
add paper1 draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2865 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams.
-\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on the GPU device.
+\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device.
 From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used.
-Below, we show the usage of these two functions by generating 4 streams on the host and on the device respectively.
+In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
 <<streams on CPU>>=
@@
 \fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create.
-For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) has a default value ``12345''.
-By setting a seed identical to the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as the streams created on host as follows,
+For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
+By setting a seed same as the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
 <<streams on GPU1, eval=TRUE>>=
@@
     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f2} }}%
-    \caption{name?\label{fig1}}%
+    \caption{name?\label{fig:1}}%
 \end{figure}
 
-In the following code, we use the 4 streams created in Section 2.1 and 4 work items to generate a vector \code{sim_1} of 6 i.i.d.
-$U (0,1)$ random numbers by the function \fct{gpuRandom::runif}.
-By default \code{Nglobal} is set as (64,8), and a matrix of \code{prod(Nglobal)} streams created on GPU is supplied.
-The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host computer.
+Now we use the 4 streams created in Section 2.1 and an NDRange of 1 by 4 work items to generate a vector \code{sim_1} of 6 i.i.d.
+$U (0,1)$ random numbers using the function \fct{gpuRandom::runif}.
+If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied.
+The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
 <<eval=TRUE, tidy=TRUE>>=
@@
 \end{itemize}
-Next we show that using the same streams (\code{streams1} and \code{streams2}),  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
+Next we show that by using a same stream \code{streams2} as \code{streams1},  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
 <<eval=TRUE,tidy=TRUE>>=
@@
 @
-Note that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used, reusing the \code{streams1} will produce a vector  different from \code{sim_1} (result of \code{sim_3}), because each time a random number is generated, the current state of the stream is changed (advances by one position).
-See below the first six rows (current states) of \code{streams1} after it being used.
+Notice that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used.
+Reusing \code{streams1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
+Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{streams1} after it being used.
 <<eval=TRUE,tidy=TRUE>>=
@@
 @
-In order to reproduce the same sequence of random numbers, we can save the streams in a file on CPU, because GPU device does not keep memory for streams objects after each execution.
-Below is the \proglang{R} code in which we use the saved streams \code{streams_from_GPU.rds} to produce a vector \code{sim_saved}, which is also same as \code{sim_1}.
+Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart.
+Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
+However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations.
+Below is a trivial example, in which we demonstrate saving \code{streams1} to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU to produce the vector called \code{sim_saved}, which is also the same as \code{sim_1}.
 <<echo=TRUE,eval=FALSE>>=
@@
 @
-
 <<save streams,eval=TRUE>>=
@@
 @
-We evaluate the distribution of 50000 GPU-generated uniform random numbers by making a density histogram  (the left panel of Figure \ref{fig0}), plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel of Figure \ref{fig0}).
-The two lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} should come from a standard uniform distribution.
+
+We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
+The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} does come from a standard uniform distribution.
 <<uniforms_random_vector,eval=TRUE>>=
-streams <- CreateStreamsGpu(Nstreams =256, keepInitial=1)
+streams <- CreateStreamsGpu(Nstreams =512, keepInitial=1)
 random_vector<-gpuRandom::runif(n=50000, streams=streams, 
-                                Nglobal=c(64,4), type="double")
-random_vector<-as.vector(as.matrix(random_vector))
+                                Nglobal=c(64,8), type="double")
+random_vector<-as.vector(random_vector)
 @
@@
 @
-\caption{\label{fig0} histogram and empirical cumulative distribution of the GPU-generated random numbers.}
+\caption{histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
 \end{figure}
 
-\subsection{Using streams to generate i.i.d. $N (0, 1)$ random numbers on GPU}
+\subsection{Using streams to generate i.i.d. N (0, 1) random numbers on GPU}
 We apply the Box-Muller transformation to $U (0,1)$ random numbers to generate Gaussian random numbers.
-Box-Muller algorithm (see algorithm \ref{algorithm1}) takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables.
+As shown in Algorithm \ref{algorithm1}, Box-Muller algorithm takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables.
 The derivation basically uses transformation from Cartesian coordinates to polar coordinates to represent two independent, Normally distributed random variables.
@@
 \end{algorithm*}
-%
-Below is a fragment of the kernel function that generates standard Gaussian random numbers. 
+
+Listing \ref{lst:normal} shows a fragment of the kernel that generates standard Gaussian random numbers.
+Aside from using local memory and the part that does the Box-Muller transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}.
+If developers want to use other Gaussian transform methods, this kernel is easy to be modified and incorporated in other \proglang{R} packages.
+The kernel executes over a 2-dimensional index space of $1 \times 2$ work-groups.
+\code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
+As the work-items in each work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), (X_2,Y_2), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
 %
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label=normal]
+\begin{lstlisting}[language=C,basicstyle=\small,caption={name},label={lst:normal}]
+const int 
+index = get_global_id(0)*get_global_size(1) + get_global_id(1);
+const int 
+DrowInc = 2*get_global_size(0), DrowStartInc = DrowInc * NpadCol;
+int Drow, Dcol, DrowStart, Dentry;
+uint temp;
+uint g1[3], g2[3];
+const int startvalue=index * NpadStreams;
+
+local double part[2], cosPart1, sinPart1;
+
+streamsToPrivate(streams,g1,g2,startvalue);
+
+// Iterate for the rows
 for(Drow=2*get_global_id(0), 
-    DrowStart=(Drow + get_local_id(1))* NpadCol;
+    DrowStart = (Drow + get_local_id(1))* NpadCol;
     Drow < Nrow; 
-    Drow += DrowInc, DrowStart += DrowStartInc){
-
+    Drow += DrowInc, DrowStart += DrowStartInc) {
+  // Iterate for the columns
   for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
       Dcol < Ncol;
-      Dcol += get_num_groups(1), Dentry += get_num_groups(1)){
+      Dcol += get_num_groups(1), Dentry += get_num_groups(1)) {
 
+    // Generates an uniform random integer
     temp = clrngMrg31k3pNextState(g1, g2);
@@
 \end{figure}
-As Gaussian numbers are generated in pairs using the Box-Muller method, work items are organized to be $1 \times 2$ work groups in the index space (a), the shading in the grids of (a) distinguishes work groups.
-Each work group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where the work item of `\code{get_local_id(1)}$=0$' generates the \code{X} in a pair, and work item of `\code{get_local_id(1)}$=1$' generates the \code{Y} in a pair.
-The numbers in the third row in each gird of (a) represents the local index of each work item.
-Cells of the matrix are filled from left to right, that is, in the first iteration of execution, the 8 work items generate random numbers in the leftmost two columns (8 cells) of matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration  fills the rest two columns of the matrix simultaneously.
-Therefore, each work item generates one Gaussian random number in each iteration, the filling process is the same as the filling process of Figure \ref{fig1}.
 
-We show the usage of \fct{gpuRandom::rnorm} by generating a $512 \times 1024$ matrix named \code{normal_matrix} of double presicion Gaussian random numbers, and assess Normality by making a histogram (left panel of Figure \ref{fig2}) and a Q-Q plot (right panel of Figure \ref{fig2}) of the generated random numbers.
-The points in the Q-Q plot in Figure \ref{fig2} fall along the diagonal line, we have good evidence the data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part (adapted from \fct{stat::qnorm}) on GPU, which speeds up the Q-Q plot greatly.
+In the following code, We create a $1024 \times 512$ matrix named \code{normal_matrix} of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we adapted the \proglang{R} function \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on GPU, which can speed up the Q-Q plot greatly.
 <<histAndQQ_prepare,eval=TRUE>>=
@@
 normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
-                                Nglobal=c(512,128), type="double")
+                                Nglobal=c(512,128), type="float")
 @
@@
 @
-\caption{\label{fig2} Histogram and Q-Q plot of the GPU-generated random numbers.}
+\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
 \end{figure}
-Let's compare the time of generating a large-size matrix of 100 million Gaussian random numbers of double precision using the \fct{stats::rnorm} and using the \fct{gpuRandom::rnorm}.
-Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is 117.6 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
+
+We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the runtime between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
+Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size goes larger.
@@
 
-We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and a normal GPU, the other is equipped with an excellent GPU and a normal CPU.
-The device information is included in the table \ref{tab:device}.
-The \proglang{R} outputs for testing on computer 1 is shown in the following sections.
-\begin{table}[H]
-\begin{tabular}{ |c|c|c|c|c|c| } 
-\hline
-           & CPU   & GPU \\ \hline
- Computer1 & pthread-Intel(R) Xeon(R) W-2145 CPU @ 3.70GHz & 2 Radeon VII's : ROCm AMDGPU  \\ \hline
- Computer2 & dontknow & GRID V100D-8C \\ \hline
-\end{tabular}
-\caption{\label{tab:device} Device information for run-time compare.}
-\end{table}
+We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the runtime with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
+The device information is included in table \ref{tab:device}.
+The \proglang{R} outputs for testing on computer 2 is shown in the following.
+<<deviceinfo, echo=FALSE,cache=TRUE>>=
+library(kableExtra)
+a <- data.frame(
+  var1 = c(" ", "Computer1", "Computer2"),
+  var2 = c('Host', 'Intel Xenon W-2145 3.7Ghz', 	'VCPU Intel Xenon Skylake 2.5Ghz'),
+  var3 = c('Device', 'AMD Radeon VII', 'VGPU Nvidia Tesla V100')
+)
 
-These 2-way contingency tables (\ref{tab:month} and \ref{tab:week}) consists of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
+knitr::kable(a, col.names = NULL, align='c', caption = "Device information for run-time comparison.\\label{tab:device}") %>%
+  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position")
+@
+% \begin{table}[H]
+% \begin{tabular}{ |c|c|c|c|c|c| } 
+% \hline
+%            & host   &                     device \\ \hline
+%  Computer1 & INtel Xenon W-2145 3.7 Ghz & Amd Radeon VII  \\ \hline
+%  Computer2 & VCPU Intel Xenon Skylake 2.5ghz & VGPU Nvidia Tesla V100 \\ \hline
+% \end{tabular}
+% \caption{ Device information for run-time compare.\label{tab:device}}
+% \end{table}
+
+The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
 The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
@@
 \subsection{Comparing running time: Month data example}
-<<Timecomparemonth,echo=FALSE,eval=TRUE, results="hide">>=
+<<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
 data2<-read.csv("/home/ruoyong/paper1_2021_5/month.csv") 
@@
 @
-<<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
+<<monthdata,eval=TRUE,echo=FALSE, cache=FALSE>>=
 library(knitr)
 library(kableExtra)
-knitr::kable(head(month, n=12),format="latex",  caption = "Month\\label{tab:month}") %>%
-  kable_styling(bootstrap_options = "striped", full_width = F, position = "center",latex_options = "HOLD_position")
+knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
+  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
 @
+
 <<timecomparemonth2Cpu, cache=TRUE, eval=TRUE>>=
@@
 \fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for one million simulations, has a p-value of 0.4035.
-<<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
-#using GPU
+<<timecomparemonth3Gpu, eval=TRUE,cache=FALSE>>=
 almost.1 <- 1 + 64 * .Machine$double.eps
 observed_m= -gpuRandom::logfactSum(month, c(16,16))/almost.1
+
 month_GPU<-vclMatrix(month,type="integer")
-system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
-@
-\fct{gpuRandom::fisher.sim} takes about 2.3 seconds, compared to the elapsed time of \fct{gpuRandom::fisher.sim},  \fct{gpuRandom::fisher.sim} takes less than a quater of the time. 
-<<timecomparemonth3Gpu, eval=TRUE,cache=FALSE>>=
 result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
@@
 @
-We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.4035, which is close enough to the p-value from \fct{stats::fisher.test}: 0.4034.
+We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
+<<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
+#using GPU
+system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
+            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
+@
+\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU. 
 <<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
@@
 \subsection{Comparing running time: Week data example}
-<<Timecompareweek, echo=FALSE, eval=TRUE, results="hide">>=
+<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
 library(latex2exp)
@@
 <<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
-knitr::kable(head(week, n=7),format="latex",caption = "Week\\label{tab:week}") %>%
-  kable_styling(bootstrap_options = "striped", full_width = F, position = "center",latex_options = "HOLD_position")
+knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
+  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
 @
@@
 @
-The ``week'' table has a much smaller p-value: around 0.000119, which should require a larger number of simulations to get a more accurate p-value.
-<<TimecompareweekGpu, eval=TRUE, cache=FALSE>>=
-#using GPU
+The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
+<<TimecompareweekGpu2, eval=TRUE, cache=TRUE>>=
 observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
+
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
-@
-\fct{stats::fisher.test} takes 60.4 seconds, while \fct{fisher.sim} takes about 10.8 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
-<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
@@
 @
-With about ten million simulations, we get 1272 cases and a p-value around 0.0001274. We summarized the results in Table \ref{tab:summary}.
+With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
+<<TimecompareweekGpu, eval=TRUE, cache=TRUE>>=
+#using GPU
+system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
+            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
+@
+\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
 <<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
@@
 @
-\begin{table}[H]
-\centering
-\begin{tabular}{ |c|c|c|c|c|c| } 
- \hline
- data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
- Month & 1e6 & 0.4038 & 0.403507  & 0.403507     \\ \hline
- Week &  1e7  & 0.0001208 & 0.0001274 & 0.0001274\\ \hline
-\end{tabular}
-\caption{\label{tab:summary} Summary of p-values obtained on different devices.}
-\end{table}
 
-\begin{table}[H]
-\centering
-\begin{tabular}{ |c|c|c|c|c|c| } 
- \hline
- data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
- Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
- Week &  1e7  & 60.19 & 90.19 & 10.82 & 1.90\\ \hline
-\end{tabular}
-\caption{\label{tab:summary} Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.}
-\end{table}
+\subsection{A summary of the results}
+<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE>>=
+library(kableExtra)
+dt <- data.frame(
+  var1 = c("B", "1m", "10m", "1m", "10m"),
+  var2 = c('Intel 2.5ghz', 0.403804, 	0.0001251, 9.74, 60.19),
+  var3 = c('AMD Radeon', 0.403507, 0.0001274, 2.28, 10.82),
+  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202,14.58,90.38),
+  var5 = c('NVIDIA V100', 0.403507,0.0001274,0.32, 1.89),
+  var6 = c('Data','month','week','month','week')
+)
+
+knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices.\\label{tab:summary}") %>%
+  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position") %>%
+  kableExtra::group_rows(index = c("P-value" = 2, "Run-time" = 2)) %>%
+  add_header_above(c(" " = 1, "Computer 1" = 2, "Computer 2" = 2, " " = 1))
+@
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ |c|c|c|c|c|c| } 
+%  \hline
+%  data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
+%  Month & 1e6 & 0.403804 & 0.403507  & 0.403507     \\ \hline
+%  Week &  1e7  & 0.0001251 & 0.0001274 & 0.0001274\\ \hline
+% \end{tabular}
+% \caption{ Summary of p-values obtained on different devices.\label{tab:summary}}
+% \end{table}
+% 
+% 
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ |c|c|c|c|c|c| } 
+%  \hline
+%  data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
+%  Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
+%  Week &  1e7  & 60.19 & 90.38 & 10.82 & 1.90\\ \hline
+% \end{tabular}
+% \caption{ Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.\label{tab:compare time}}
+% \end{table}
 
@@
 @
-\caption{\label{fig4} Frequency distribution for test statistics from month and week data tables.}
+\caption{ Frequency distribution for test statistics from month and week data tables.\label{fig4}}
 \end{figure}
@@
                      Nglobal= c(512, 8),
-                     Nlocal= c(32, 8),
+                     Nlocal= c(128, 8),
                      NlocalCache=2000)
@@
 @
-\caption{\label{fig5} }
+\caption{\label{fig:5} }
 \end{figure}

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
better simulation number in fisher and solved a concern
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2869 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 gpuR::colnames(streams_from_GPU) = c("current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
-"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3",
-"substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")   
+"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3")   
 @ 
@@
 We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the runtime with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
-The device information is included in table \ref{tab:device}.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
-<<deviceinfo, echo=FALSE,cache=TRUE>>=
+<<deviceinfo, include=FALSE,cache=FALSE>>=
 library(kableExtra)
@@
 
-knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices.\\label{tab:summary}") %>%
+knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices.
+Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII.
+Computer 2 is equipped with	VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
   kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position") %>%

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
better simulation number in fisher and solved a concern
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2869 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ## initial.g2.3 12345 2094976052 1462898381  795289868
-## substream.g1.1 12345  336690377  502033783  739421137
-## substream.g1.2 12345  597094797 1322587635 1475938232
-## substream.g1.3 12345 1245771585 1964121530  730262207
-## substream.g2.1 12345   85196284 1949818481 1630192198
-## substream.g2.2 12345  523477687 1607232546  324551134
-## substream.g2.3 12345 2094976052 1462898381  795289868
 \end{verbatim}
@@
 We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the runtime with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
-The device information is included in table \ref{tab:device}.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
-
-\caption{\label{tab:deviceinfo}Device information for run-time comparison.\label{tab:device}}
-\centering
-\begin{tabular}[t]{c|c|c}
-\hline
- & Host & Device\\
-\hline
-Computer1 & Intel Xenon W-2145 3.7Ghz & AMD Radeon VII\\
-\hline
-Computer2 & VCPU Intel Xenon Skylake 2.5Ghz & VGPU Nvidia Tesla V100\\
-\hline
-\end{tabular}
-\end{table}
-
-\end{knitrout}
+
 % \begin{table}[H]
@@
 ##    user  system elapsed 
-##   0.355   0.006   0.361
+##   0.699   0.005   0.703
 \end{verbatim}
@@
 ##    user  system elapsed 
-##  90.024   0.080  90.069
+##  90.691   0.126  90.750
 \end{verbatim}
@@
 
-\caption{\label{tab:summarycompare}Summary of comparions of Fisher's test simulation on different devices.\label{tab:summary}}
+\caption{\label{tab:summarycompare}Summary of comparions of Fisher's test simulation on different devices.
+Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII.
+Computer 2 is equipped with	VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\label{tab:summary}}
 \centering
@@
                      Nglobal= c(512, 8),
-                     Nlocal= c(32, 8),
+                     Nlocal= c(128, 8),
                      NlocalCache=2000)
+## 
+## #pragma OPENCL EXTENSION cl_khr_fp64 : enable
+## 
+## 
+## #define N 4800  //dimension of matrix
+## #define colStart 0  //column to start at
+## #define colEnd 4800
+## //internal number of columns
+## #define Npad 4864
+## //internal columns for matrix holding diagonals
+## #define NpadDiag 4864
+## #define NstartA 0
+## #define NstartD 0
+## //elements in internal cache
+## #define Ncache 2000
+## //extra rows between stacked matrices
+## #define NpadBetweenMatrices 23347200
+## #define maxLocalItems 1024
+## 
+## __kernel void cholBatch(
+## 	__global double *A,
+## 	__global double *diag,
+##               int Nmatrix
+## ){
+##  const int localIndex = get_local_id(0)*get_local_size(1) + get_local_id(1);
+##  const int localIndexIsZero = (localIndex==0);
+##  const int NlocalTotal = get_local_size(0)*get_local_size(1);
+##  local double diagLocal[Ncache];//local cache of diagonals
+##  local double toAddLocal[maxLocalItems];
+##  int Dcol, DcolNpad;
+##  int Drow, DrowBlock, Dk, Dmatrix, DmatrixBlock;
+## int minDcolNcache;
+##  double DL;
+##  local double diagDcol;
+##  int AHere, AHereDcol, AHereDrow, diagHere;
+##  barrier(CLK_LOCAL_MEM_FENCE);
+## 
+## for(Dmatrix = get_group_id(0);
+##      Dmatrix < Nmatrix;
+##      Dmatrix+= get_num_groups(0)){
+## 
+##  diagHere = Dmatrix*NpadDiag + NstartD;
+##  AHere =  Dmatrix*NpadBetweenMatrices + NstartA;
+## 
+##  for(Dcol = colStart; Dcol < colEnd; Dcol++) {
+##   DcolNpad = Dcol*Npad;
+##   AHereDcol = AHere+DcolNpad;
+##   DL = 0.0;
+##   diagLocal[localIndex]=0.0;
+##   toAddLocal[localIndex]=0.0;
+##   minDcolNcache = min(Dcol, Ncache);
+##   for(Dk = localIndex; Dk < minDcolNcache; Dk += NlocalTotal) {
+##     DL = A[AHereDcol+Dk];
+##     diagLocal[Dk] = diag[diagHere+Dk] * DL;
+##     toAddLocal[localIndex] += diagLocal[Dk] * DL;
+##   }// Dk
+##   for(Dk=minDcolNcache+localIndex; Dk < Dcol; Dk += NlocalTotal) {
+##     DL = A[AHereDcol+Dk];
+##     toAddLocal[localIndex] += diag[diagHere+Dk] * DL * DL;
+##   }// Dk
+## 
+## // reduction on dimension 1
+##   barrier(CLK_LOCAL_MEM_FENCE);
+##   if( (get_local_id(1) == 0)){
+##    for(Dk = 1; Dk < get_local_size(1); Dk++) {
+##     toAddLocal[localIndex] +=  toAddLocal[localIndex + Dk];
+##    }//for Dk
+##   }//get_local_id(1) == 0
+##   barrier(CLK_LOCAL_MEM_FENCE);
+## 
+## // final reduction on dimension 0
+##   if(localIndexIsZero ){
+##    for(Dk = get_local_size(1); Dk < NlocalTotal; Dk+= get_local_size(1)) {
+##     toAddLocal[localIndex] +=  toAddLocal[Dk];
+##    } //Dk
+##    diagDcol = A[AHereDcol+Dcol] - toAddLocal[localIndex];
+##    diag[diagHere+Dcol] = diagDcol;
+## #ifdef diagToOne
+##    A[AHereDcol+Dcol] = 1.0;
+## #endif
+##   }  //localIndex==0 and Dmatrix < Nmatrix
+##   barrier(CLK_LOCAL_MEM_FENCE);
+## 
+## // off diagonals
+##  for(DrowBlock = Dcol+1; DrowBlock < N; DrowBlock += get_local_size(0)) {
+##   Drow = DrowBlock + get_local_id(0);
+##   AHereDrow = AHere+Drow*Npad;
+##   DL = 0.0;
+##   if(Drow < N){
+##   for(Dk = get_local_id(1); Dk < minDcolNcache; Dk+=get_local_size(1)) {
+##    DL += A[AHereDrow+Dk] * diagLocal[Dk];
+##   } // Dk
+## 	 for(minDcolNcache + get_local_id(1); Dk < Dcol; Dk+=get_local_size(1)) {
+##     DL += A[AHereDrow+Dk] * diag[diagHere+Dk] * A[AHereDcol+Dk];
+##   }// Dk
+##    }//Drow < N
+##   toAddLocal[localIndex] = DL;
+## 
+##   // local reduction
+##   barrier(CLK_LOCAL_MEM_FENCE);
+##   if( (get_local_id(1) == 0) & (Drow < N)){
+##    DL = toAddLocal[localIndex];
+##    for(Dk = 1; Dk < get_local_size(1); Dk++) {
+##     DL +=  toAddLocal[localIndex + Dk];
+##    }//Dk
+##    A[AHereDrow+Dcol] = (A[AHereDrow+Dcol] - DL)/diagDcol;
+##   }//get_local_id(1) == 0
+##   barrier(CLK_GLOBAL_MEM_FENCE);
+## 
+##  }//DrowBlock
+## barrier(CLK_GLOBAL_MEM_FENCE);
+## } // Dcol loop
+## 
+## barrier(CLK_GLOBAL_MEM_FENCE);
+## } // Dmatrix loop
+## 
+## barrier(CLK_LOCAL_MEM_FENCE);
+## }// kernel
 \end{verbatim}

### 2021-09-28 pkg/gpuRandom/R/CreateStreamsCpu.R

@@ -16,7 +16,7 @@ CreateStreamsCpu = function(Nstreams) {
 
   colnames(result) = c("current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
                        "initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3")
-                #       "substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")
+                     # "substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")
 
 
   cpp_mrg31k3pCreateStreams(result)

### 2021-09-28 pkg/gpuRandom/R/RcppExports.R

@@ -49,8 +49,8 @@ gemmBatch2backend <- function(A, B, C, transposeABC, submatrixA, submatrixB, sub
     .Call('_gpuRandom_gemmBatch2backend', PACKAGE = 'gpuRandom', A, B, C, transposeABC, submatrixA, submatrixB, submatrixC, batches, workgroupSize, NlocalCache, verbose)
 }
 
-cpp_gpuFisher_test <- function(xR, resultsR, B, streamsR, max_global_size, max_local_size) {
-    .Call('_gpuRandom_cpp_gpuFisher_test', PACKAGE = 'gpuRandom', xR, resultsR, B, streamsR, max_global_size, max_local_size)
+cpp_gpuFisher_test <- function(xR, resultsR, NstartofResult, B, streamsR, max_global_size, max_local_size) {
+    .Call('_gpuRandom_cpp_gpuFisher_test', PACKAGE = 'gpuRandom', xR, resultsR, NstartofResult, B, streamsR, max_global_size, max_local_size)
 }
 
 #' Random number generation

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
changes in paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2870 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
                    type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
-result_monthgpu$simnum
+result_monthgpu$simNum
 result_monthgpu$counts
@@
                 type="double",returnStatistics=TRUE,Nglobal = Nglobal)
-result_weekgpu$simnum
+result_weekgpu$simNum
 result_weekgpu$counts

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
changes in paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2871 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
                    type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
-result_monthgpu$simnum
-## [1] 1015808
+result_monthgpu$simNum
+## NULL
 result_monthgpu$counts
-## [1] 410098
+## [1] 403788
 result_monthgpu$p.value
-## [1] 0.4037166
+## [1] 0.4037628
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.699   0.005   0.703
+##   0.975   0.002   1.043
 \end{verbatim}
@@
                 type="double",returnStatistics=TRUE,Nglobal = Nglobal)
-result_weekgpu$simnum
-## [1] 9994240
+result_weekgpu$simNum
+## NULL
 result_weekgpu$counts
-## [1] 1223
+## [1] 1224
 result_weekgpu$p.value
-## [1] 0.0001224705
+## [1] 0.0001225
 \end{verbatim}

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
back to previous version of fisher
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2872 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-09-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
                    type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
-result_monthgpu$simnum
-## [1] 1015808
+result_monthgpu$simNum
+## [1] 1000064
 result_monthgpu$counts
-## [1] 410098
+## [1] 403788
 result_monthgpu$p.value
-## [1] 0.4037166
+## [1] 0.4037628
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.355   0.006   0.361
+##   0.793   0.000   0.792
 \end{verbatim}

### 2021-09-28 pkg/gpuRandom/R/fisher_sim.R

@@ -2,7 +2,7 @@
 #'
 #'
 #' @param x a vclMatrix for Fisher's test
-#' @param B requested number of simulations.
+#' @param N requested number of simulations.
 #' @param streams streams objects. 
 #' @param type "double" or "float" of test statistics.
 #' @param Nglobal the size of the index space for use.
@@ -18,7 +18,7 @@
 
 fisher.sim=function(
   x, # a vclMatrix
-  B, # requested number of simualtion,
+  N, # requested number of simualtion,
   streams, 
   type = c('float','double')[1+gpuR::gpuInfo()$double_support],
   returnStatistics = FALSE,
@@ -47,7 +47,9 @@ fisher.sim=function(
   #       storage.mode(x) <- "integer"
   #   }
   # 
-  #  }else{stop("'x'  must be matrix")}  
+  #  }else{stop("'x'  must be matrix")}
+  
+  
   # lfactorial(x)=lgamma(x+1)
   # STATISTIC <- -sum(lfactorial(x))          ##STATISTIC is negative
   # almost.1 <- 1 + 64 * .Machine$double.eps
@@ -59,13 +61,13 @@ fisher.sim=function(
       Nglobal = c(64,16)
       seedR = as.integer(as.integer(2^31-1)*(2*stats::runif(6) - 1) ) 
       seed <- gpuR::vclVector(seedR, type="integer")  
-      streams<-vclMatrix(0L, nrow=1024, ncol=12, type="integer")
+      streams<-vclMatrix(0L, nrow=1024, ncol=18, type="integer")
       CreateStreamsGpuBackend(seed, streams, keepInitial=1)
       
     }else{
       seedR = as.integer(as.integer(2^31-1)*(2*stats::runif(6) - 1) ) 
       seed <- gpuR::vclVector(seedR, type="integer")  
-      streams<-vclMatrix(0L, nrow=prod(Nglobal), ncol=12, type="integer")
+      streams<-vclMatrix(0L, nrow=prod(Nglobal), ncol=18, type="integer")
       CreateStreamsGpuBackend(seed, streams, keepInitial=1)
     }
   }else {
@@ -97,50 +99,29 @@ fisher.sim=function(
   
   #  print(class(xVcl))
   
-  TotalWorkItems <- prod(Nglobal)
-  
-  
-  simPerItem1<-B %/% TotalWorkItems  
[…]

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
back to previous version of fisher
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2873 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
-Nglobal = c(128,128)
-streams <- CreateStreamsGpu(Nstreams =128*128, keepInitial=1)
+Nglobal = c(512,128)
+streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
 @
@@
 result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
+                   type="double", returnStatistics=TRUE,  Nglobal = c(512,128))
 result_monthgpu$simNum
@@
 system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
+            type="double", returnStatistics=FALSE, Nglobal = c(512,128)))
 @
@@
 The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
-<<TimecompareweekGpu2, eval=TRUE, cache=TRUE>>=
+<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
 observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
@@
 With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
-<<TimecompareweekGpu, eval=TRUE, cache=TRUE>>=
+<<TimecompareweekGpu, eval=TRUE, cache=FALSE>>=
 #using GPU

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
back to previous version of fisher
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2873 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
+                   type="double", returnStatistics=TRUE,  Nglobal = c(512,128))
 result_monthgpu$simNum
-## [1] 1000064
+## [1] 1048576
 result_monthgpu$counts
-## [1] 403788
+## [1] 422440
 result_monthgpu$p.value
-## [1] 0.4037628
+## [1] 0.4028707
 \end{verbatim}
@@
 system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
+            type="double", returnStatistics=FALSE, Nglobal = c(512,128)))
 ##    user  system elapsed 
-##   0.793   0.000   0.792
+##   0.719   0.011   0.730
 \end{verbatim}
@@
 result_weekgpu$simNum
-## NULL
+## [1] 10027008
 result_weekgpu$counts
-## [1] 1224
+## [1] 1226
 result_weekgpu$p.value
-## [1] 0.0001225
+## [1] 0.0001223695
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.899   0.000   1.898
+##   4.649   0.004   4.710
 \end{verbatim}

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
add likfit for Betas matrix given
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2876 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
 ## initial.g2.3   12345 2094976052 1462898381  795289868
+## substream.g1.1 12345  336690377  502033783  739421137
+## substream.g1.2 12345  597094797 1322587635 1475938232
+## substream.g1.3 12345 1245771585 1964121530  730262207
+## substream.g2.1 12345   85196284 1949818481 1630192198
+## substream.g2.2 12345  523477687 1607232546  324551134
+## substream.g2.3 12345 2094976052 1462898381  795289868
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.056   0.008   0.064
+##   0.140   0.004   0.143
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   6.932   0.942   7.870
+##   6.610   0.862   7.466
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
-## [1] 0.4038036
+## [1] 0.4042196
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
 ##    user  system elapsed 
-##  14.591   0.006  14.590
+##  14.578   0.002  14.569
 \end{verbatim}
@@
 result_monthgpu$simNum
-## [1] 1000064
+## [1] 1015808
 result_monthgpu$counts
-## [1] 403788
+## [1] 410098
 result_monthgpu$p.value
-## [1] 0.4037628
+## [1] 0.4037166
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.793   0.000   0.792
+##   0.702   0.005   0.706
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
-## [1] 0.0001208
+## [1] 0.0001284
 system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
 ##    user  system elapsed 
-##  90.691   0.126  90.750
+##  90.259   0.150  90.439
 \end{verbatim}
@@
 result_weekgpu$simNum
-## NULL
+## [1] 10010624
 result_weekgpu$counts
-## [1] 1224
+## [1] 1260
 result_weekgpu$p.value
-## [1] 0.0001225
+## [1] 0.0001259662
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.899   0.000   1.898
+##   4.477   0.009   4.483
 \end{verbatim}

### 2021-09-28 pkg/gpuRandom/R/fisher_sim.R

@@ -61,13 +61,13 @@ fisher.sim=function(
       Nglobal = c(64,16)
       seedR = as.integer(as.integer(2^31-1)*(2*stats::runif(6) - 1) ) 
       seed <- gpuR::vclVector(seedR, type="integer")  
-      streams<-vclMatrix(0L, nrow=1024, ncol=12, type="integer")
+      streams<-vclMatrix(0L, nrow=1024, ncol=18, type="integer")
       CreateStreamsGpuBackend(seed, streams, keepInitial=1)
       
     }else{
       seedR = as.integer(as.integer(2^31-1)*(2*stats::runif(6) - 1) ) 
       seed <- gpuR::vclVector(seedR, type="integer")  
-      streams<-vclMatrix(0L, nrow=prod(Nglobal), ncol=12, type="integer")
+      streams<-vclMatrix(0L, nrow=prod(Nglobal), ncol=18, type="integer")
       CreateStreamsGpuBackend(seed, streams, keepInitial=1)
     }
   }else {

### 2021-09-28 pkg/gpuRandom/R/rnorm.R

@@ -37,12 +37,12 @@ rnorm = function(
       Nglobal = c(64,8)
       seedR = as.integer(as.integer(2^31-1)*(2*stats::runif(6) - 1) ) 
       seed <- gpuR::vclVector(seedR, type="integer")  
-      streams<-vclMatrix(0L, nrow=512, ncol=12, type="integer")
+      streams<-vclMatrix(0L, nrow=512, ncol=18, type="integer")
       CreateStreamsGpuBackend(seed, streams, keepInitial=1)
     }else{
       seedR = as.integer(as.integer(2^31-1)*(2*stats::runif(6) - 1) ) 
       seed <- gpuR::vclVector(seedR, type="integer")  
-      streams<-vclMatrix(0L, nrow=prod(Nglobal), ncol=12, type="integer")
+      streams<-vclMatrix(0L, nrow=prod(Nglobal), ncol=18, type="integer")
       CreateStreamsGpuBackend(seed, streams, keepInitial=1)
     }
   }else if(missing(Nglobal)){

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
add likfit for Betas matrix given
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2876 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
@@
 gpuR::colnames(streams_from_GPU) = c("current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
-"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3")   
+"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3",
+"substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")   
 @ 
@@
 Nglobal = c(128,128)
-streams <- CreateStreamsGpu(Nstreams =128*128, keepInitial=1)
+streams <- CreateStreamsGpu(Nstreams =128^2, keepInitial=1)
 @

### 2021-09-28 pkg/gpuRandom/src/streamsCpu.cpp

@@ -244,13 +244,13 @@ void convertclRngMat(clrngMrg31k3pStream* streams, Rcpp::IntegerMatrix result) {
       result(Ditem,Dcis*6 + Dg*3 + Delement) = streams[Ditem].initial.g1[Delement];//0,6; 0,7
       Dg=1;
       result(Ditem,Dcis*6 + Dg*3 + Delement) = streams[Ditem].initial.g2[Delement];//0,9; 0,10
-   /*    
+      
       Dcis=2;
       Dg=0;
       result(Ditem,Dcis*6 + Dg*3 + Delement) = streams[Ditem].substream.g1[Delement];//0,12
       Dg=1;
       result(Ditem,Dcis*6 + Dg*3 + Delement) = streams[Ditem].substream.g2[Delement];//0,15
-   */
+  
     }
   }
   

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
streams length back to 12
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2877 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.140   0.004   0.143
+##   0.141   0.004   0.145
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   6.610   0.862   7.466
+##   6.600   0.743   7.338
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
-## [1] 0.4042196
+## [1] 0.4037316
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
 ##    user  system elapsed 
-##  14.578   0.002  14.569
+##  14.505   0.008  14.530
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.702   0.005   0.706
+##   0.734   0.005   0.737
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
-## [1] 0.0001284
+## [1] 0.0001274
 system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
 ##    user  system elapsed 
-##  90.259   0.150  90.439
+##  90.278   0.157  90.506
 \end{verbatim}
@@
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = Nglobal)
+                type="double",returnStatistics=TRUE,Nglobal = c(128,128))
 result_weekgpu$simNum
@@
 system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
+            type="double", returnStatistics=FALSE,Nglobal = c(128,128)))
 ##    user  system elapsed 
-##   4.477   0.009   4.483
+##   4.148   0.008   4.153
 \end{verbatim}

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
streams length back to 12
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2877 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 gpuR::colnames(streams_from_GPU) = c("current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
-"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3",
-"substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")   
+"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3")
+#"substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")   
 @ 
@@
 system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
+            type="double", returnStatistics=FALSE,Nglobal = c(128,128)))
 @

### 2021-10-01 pkg/gpuRandom/R/RcppExports.R
working on paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2886 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -134,8 +134,14 @@ multiplyLowerBatchBackend <- function(C, A, B, diagIsOne, Nglobal, Nlocal, Nloca
     .Call('_gpuRandom_multiplyLowerBatchBackend', PACKAGE = 'gpuRandom', C, A, B, diagIsOne, Nglobal, Nlocal, NlocalCache)
 }
 
-cpp_mrg31k3pCreateStreams <- function(result) {
-    .Call('_gpuRandom_cpp_mrg31k3pCreateStreams', PACKAGE = 'gpuRandom', result)
+#' @name createStreamsCpu
+#' @title create streams stored on the CPU
+#' @description streams for random numbers
+#' @param n number of streams to create
+#' @param seed random seed, length 6, recycled if shorter
+#' @export
+createStreamsCpu <- function(n, seed) {
+    .Call('_gpuRandom_createStreamsCpu', PACKAGE = 'gpuRandom', n, seed)
 }
 
 CreateStreamsGpuBackend <- function(creatorInitialGlobalR, streamsR, keepInitial) {

### 2021-10-01 pkg/gpuRandom/src/RcppExports.cpp

@@ -408,14 +408,15 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
-// cpp_mrg31k3pCreateStreams
-Rcpp::IntegerMatrix cpp_mrg31k3pCreateStreams(Rcpp::IntegerMatrix result);
-RcppExport SEXP _gpuRandom_cpp_mrg31k3pCreateStreams(SEXP resultSEXP) {
+// createStreamsCpu
+Rcpp::IntegerMatrix createStreamsCpu(Rcpp::IntegerVector n, Rcpp::IntegerVector seed);
+RcppExport SEXP _gpuRandom_createStreamsCpu(SEXP nSEXP, SEXP seedSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
-    Rcpp::traits::input_parameter< Rcpp::IntegerMatrix >::type result(resultSEXP);
-    rcpp_result_gen = Rcpp::wrap(cpp_mrg31k3pCreateStreams(result));
+    Rcpp::traits::input_parameter< Rcpp::IntegerVector >::type n(nSEXP);
+    Rcpp::traits::input_parameter< Rcpp::IntegerVector >::type seed(seedSEXP);
+    rcpp_result_gen = Rcpp::wrap(createStreamsCpu(n, seed));
     return rcpp_result_gen;
 END_RCPP
 }
@@ -456,7 +457,7 @@ static const R_CallMethodDef CallEntries[] = {
     {"_gpuRandom_multiplyLowerDiagonalBatchBackend", (DL_FUNC) &_gpuRandom_multiplyLowerDiagonalBatchBackend, 9},
     {"_gpuRandom_multiplyDiagonalBatchBackend", (DL_FUNC) &_gpuRandom_multiplyDiagonalBatchBackend, 6},
     {"_gpuRandom_multiplyLowerBatchBackend", (DL_FUNC) &_gpuRandom_multiplyLowerBatchBackend, 7},
-    {"_gpuRandom_cpp_mrg31k3pCreateStreams", (DL_FUNC) &_gpuRandom_cpp_mrg31k3pCreateStreams, 1},
+    {"_gpuRandom_createStreamsCpu", (DL_FUNC) &_gpuRandom_createStreamsCpu, 2},
[…]

### 2021-10-01 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
working on paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2886 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{Creating streams}
+
+{\tt changed createStreams functions to be compatible with each other}
+
+create streams on cpu
+<<createStreams, size="tiny">>=
+library("gpuRandom")
+myStreamsCpu <- createStreamsCpu(n=4, seed = 12345)
+myStreamsCpu
+@
+
+streams on gpu
+<<createStreamsGpu>>=
+myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu2 = createStreams(n=4, seed=12345)
+range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
+@
+
+<<simStuff, eval=TRUE,tidy=TRUE>>=
+sim_2 <-gpuRandom::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double")
+as.vector(sim_2)
+@
+
+check seed
+<<simSeed, size='tiny'>>=
+matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
+@
+
+check reproducibility
+<<simStuff, eval=TRUE,tidy=TRUE>>=
+c(as.vector(sim_2), as.vector(gpuRandom::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double")))
+as.vector(gpuRandom::runif(n=12, streams=vclMatrix(myStreamsCpu), Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+@
+
 In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
@@
 # Create streams on host
-streams_on_CPU<-gpuRandom::CreateStreamsCpu(Nstreams=4)
+streams_on_CPU<-gpuRandom::CreateStreamsCpu(seed = rep(12345, 6), Nstreams=4)
 t(streams_on_CPU)

### 2021-10-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
streams
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2887 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<simStuff, eval=TRUE,tidy=TRUE>>=
-c(as.vector(sim_2), as.vector(gpuRandom::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double")))
-as.vector(gpuRandom::runif(n=12, streams=vclMatrix(myStreamsCpu), Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
 @

### 2021-10-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
no big changes,just to save
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2888 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 setContext(   grep('gpu', listContexts()$device_type) [1]    )
-setwd("/home/ruoyong/paper1_2021_5")
+#setwd("/home/ruoyong/paper1_2021_5")
 @
@@
 <<createStreams, size="tiny">>=
-library("gpuRandom")
 myStreamsCpu <- createStreamsCpu(n=4, seed = 12345)
-myStreamsCpu
+t(myStreamsCpu)
 @
@@
 check reproducibility
-<<simStuff, eval=TRUE,tidy=TRUE>>=
+<<simStuff2, eval=TRUE,tidy=TRUE>>=
 as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))

### 2021-10-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
no big changes,just to save
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2888 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{Creating streams}
-In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
-The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
-If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory.
-In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device.
-\fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams.
-\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device.
-From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used.
-In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-# Create streams on host
-streams_on_CPU<-gpuRandom::CreateStreamsCpu(Nstreams=4)
-t(streams_on_CPU)
+
+{\tt changed createStreams functions to be compatible with each other}
+
+create streams on cpu
+\begin{knitrout}\tiny
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+myStreamsCpu <- createStreamsCpu(n=4, seed = 12345)
+t(myStreamsCpu)
 ##               [,1]       [,2]       [,3]       [,4]
@@
 ## initial.g2.3 12345 2094976052 1462898381  795289868
-## substream.g1.1 12345  336690377  502033783  739421137
-## substream.g1.2 12345  597094797 1322587635 1475938232
-## substream.g1.3 12345 1245771585 1964121530  730262207
-## substream.g2.1 12345   85196284 1949818481 1630192198
-## substream.g2.2 12345  523477687 1607232546  324551134
-## substream.g2.3 12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+streams on gpu
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu2 = createStreams(n=4, seed=12345)
+range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
+## [1] 0 0
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+sim_2 <- gpuRandom::runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1,
+    nrow(myStreamsGpu1)), type = "double")
+as.vector(sim_2)
+## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+check seed
+\begin{knitrout}\tiny
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
+##      current.g1.1 current.g1.2 current.g1.3 current.g2.1 current.g2.2
+## [1,]    878672095    240667857    240667857    642281259   1069151070
+## [2,]   2113333390    559530223   1309565828   1335994581     61444481
+## [3,]    502033783   1322587635   1964121530   1949818481   1607232546
+## [4,]    739421137   1475938232    730262207   1630192198    324551134
+##      current.g2.3 initial.g1.1 initial.g1.2 initial.g1.3 initial.g2.1
+## [1,]    809054265        12345        12345        12345        12345
+## [2,]    197003928    336690377    597094797   1245771585     85196284
+## [3,]   1462898381    502033783   1322587635   1964121530   1949818481
+## [4,]    795289868    739421137   1475938232    730262207   1630192198
+##      initial.g2.2 initial.g2.3
+## [1,]        12345        12345
+## [2,]    523477687   2094976052
+## [3,]   1607232546   1462898381
+## [4,]    324551134    795289868
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+check reproducibility
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+as.vector(gpuRandom::runif(n = 4, streams = myStreamsGpu1, Nglobal = c(1,
+    nrow(myStreamsGpu1)), type = "double"))
+## [1] 0.6487742 0.1112075 0.3661944 0.5018562
+as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
+    nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
+## [1] 0.6487742 0.1112075 0.3661944 0.5018562
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
+The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
+If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory.
+In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device.
+\fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams.
+\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device.
+From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used.
+In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+# Create streams on host
+streams_on_CPU<-gpuRandom::CreateStreamsCpu(seed = rep(12345, 6), Nstreams=4)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'CreateStreamsCpu' is not an exported object from 'namespace:gpuRandom'}}\begin{verbatim}
+t(streams_on_CPU)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 't': object 'streams\_on\_CPU' not found}}\end{kframe}
 \end{knitrout}
@@
 streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'CreateStreamsGpu' is not an exported object from 'namespace:gpuRandom'}}\begin{verbatim}
 streams_from_GPU <- as.matrix(streams1)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(streams1): object 'streams1' not found}}\begin{verbatim}
 streams2 = deepcopy(streams1)
 \end{verbatim}
-\end{kframe}
-\end{knitrout}
-
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'object' in selecting a method for function 'deepcopy': object 'streams1' not found}}\end{kframe}
+\end{knitrout}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuR::colnames(streams\_from\_GPU) = c("{}current.g1.1"{}, "{}current.g1.2"{}, : object 'streams\_from\_GPU' not found}}\end{kframe}
+\end{knitrout}
 \begin{knitrout}
@@
 t(streams_from_GPU)   
-##                 [,1]       [,2]       [,3]       [,4]
-## current.g1.1   12345  336690377  502033783  739421137
-## current.g1.2   12345  597094797 1322587635 1475938232
-## current.g1.3   12345 1245771585 1964121530  730262207
-## current.g2.1   12345   85196284 1949818481 1630192198
-## current.g2.2   12345  523477687 1607232546  324551134
-## current.g2.3   12345 2094976052 1462898381  795289868
-## initial.g1.1   12345  336690377  502033783  739421137
-## initial.g1.2   12345  597094797 1322587635 1475938232
-## initial.g1.3   12345 1245771585 1964121530  730262207
-## initial.g2.1   12345   85196284 1949818481 1630192198
-## initial.g2.2   12345  523477687 1607232546  324551134
-## initial.g2.3   12345 2094976052 1462898381  795289868
-## substream.g1.1 12345  336690377  502033783  739421137
-## substream.g1.2 12345  597094797 1322587635 1475938232
-## substream.g1.3 12345 1245771585 1964121530  730262207
-## substream.g2.1 12345   85196284 1949818481 1630192198
-## substream.g2.2 12345  523477687 1607232546  324551134
-## substream.g2.3 12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 't': object 'streams\_from\_GPU' not found}}\end{kframe}
 \end{knitrout}
@@
     type = "double")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams1' not found}}\begin{verbatim}
 as.vector(sim_1)
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.vector(sim\_1): object 'sim\_1' not found}}\end{kframe}
 \end{knitrout}
@@
     type = "double")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams2' not found}}\begin{verbatim}
 as.vector(sim_2)
@@
     type = "double")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams1' not found}}\begin{verbatim}
 as.vector(sim_3)
-## [1] 0.6487742 0.1112075 0.3661944 0.5018562 0.1088229 0.3114331
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.vector(sim\_3): object 'sim\_3' not found}}\begin{verbatim}
 t(as.matrix(streams1))[1:6, ]
-##            [,1]       [,2]       [,3]       [,4]
-## [1,] 1167281028  640923766  502033783  739421137
-## [2,] 1918428443 1912157188 1322587635 1475938232
-## [3,] 1858462085  286315187 1964121530  730262207
-## [4,]  933585541 2119609883 1949818481 1630192198
-## [5,] 1132031887  834429287 1607232546  324551134
-## [6,]  465230163   47498872 1462898381  795289868
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 't': object 'streams1' not found}}\end{kframe}
 \end{knitrout}
@@
 streams <- CreateStreamsGpu(Nstreams =512, keepInitial=1)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in CreateStreamsGpu(Nstreams = 512, keepInitial = 1): could not find function "{}CreateStreamsGpu"{}}}\begin{verbatim}
 random_vector<-gpuRandom::runif(n=50000, streams=streams, 
                                 Nglobal=c(64,8), type="double")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams' not found}}\begin{verbatim}
 random_vector<-as.vector(random_vector)
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.vector(random\_vector): object 'random\_vector' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
-\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-1} 
-\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-2} 
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hist(random\_vector, prob = TRUE, main = "{}Histogram"{}): object 'random\_vector' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in sort(x): object 'random\_vector' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot.xy(xy.coords(x, y), type = type, ...): plot.new has not been called yet}}\end{kframe}
 \end{knitrout}
@@
 streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in CreateStreamsGpu(Nstreams = 512 * 128, keepInitial = 1): could not find function "{}CreateStreamsGpu"{}}}\begin{verbatim}
 normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams' not found}}\end{kframe}
 \end{knitrout}
@@
 \subsection{Comparing running time: Month data example}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in CreateStreamsGpu(Nstreams = 128\textasciicircum{}2, keepInitial = 1): could not find function "{}CreateStreamsGpu"{}}}\end{kframe}
+\end{knitrout}
 \begin{knitrout}
@@
                    type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuRandom::fisher.sim(month\_GPU, 1e+06, streams = streams, type = "{}double"{}, : object 'streams' not found}}\begin{verbatim}
 result_monthgpu$simNum
-## [1] 1015808
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'result\_monthgpu' not found}}\begin{verbatim}
 result_monthgpu$counts
-## [1] 410098
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'result\_monthgpu' not found}}\begin{verbatim}
 result_monthgpu$p.value
-## [1] 0.4037166
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'result\_monthgpu' not found}}\end{kframe}
 \end{knitrout}
@@
             type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
-##    user  system elapsed 
-##   0.734   0.005   0.737
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuRandom::fisher.sim(month\_GPU, 1e+06, streams = streams, type = "{}double"{}, : object 'streams' not found}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Timing stopped at: 0.001 0 0}}\end{kframe}
 \end{knitrout}
@@
                 type="double",returnStatistics=TRUE,Nglobal = c(128,128))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuRandom::fisher.sim(week\_GPU, 1e+07, streams = streams, type = "{}double"{}, : object 'streams' not found}}\begin{verbatim}
 result_weekgpu$simNum
-## [1] 10010624
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'result\_weekgpu' not found}}\begin{verbatim}
 result_weekgpu$counts
-## [1] 1260
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'result\_weekgpu' not found}}\begin{verbatim}
 result_weekgpu$p.value
-## [1] 0.0001259662
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'result\_weekgpu' not found}}\end{kframe}
 \end{knitrout}
@@
             type="double", returnStatistics=FALSE,Nglobal = c(128,128)))
-##    user  system elapsed 
-##   4.148   0.008   4.153
-\end{verbatim}
-\end{kframe}
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuRandom::fisher.sim(week\_GPU, 1e+07, streams = streams, type = "{}double"{}, : object 'streams' not found}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Timing stopped at: 0 0 0.001}}\end{kframe}
 \end{knitrout}
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'CreateStreamsGpu' is not an exported object from 'namespace:gpuRandom'}}\end{kframe}
 \end{knitrout}
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streamsGpu' not found}}\end{kframe}
 \end{knitrout}
@@
                    type = theType)
-
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'zmatGpu' not found}}\begin{verbatim}
 gpuRandom::multiplyLowerDiagonalBatch(
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuRandom:::multiplyLowerDiagonalBatchBackend(output, L, D, B, : object 'simMat' not found}}\end{kframe}
 \end{knitrout}
@@
 simRaster = brick(myRaster, nl = ncol(simMat)*nrow(paramsGpu))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\begin{verbatim}
 values(simRaster) = as.vector(as.matrix(simMat))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'as.matrix': object 'simMat' not found}}\begin{verbatim}
 names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\end{kframe}
 \end{knitrout}
@@
 \end{verbatim}
-\end{kframe}
-\includegraphics[width=0.45\textwidth]{figure/maternplot-1} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-2} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-3} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-4} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-5} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-6} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-7} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-8} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-9} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-10} 
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'simRaster' not found}}\end{kframe}
 \end{knitrout}

### 2021-10-04 pkg/gpuRandom/src/streamsCpu.cpp

@@ -340,6 +340,8 @@ clrngMrg31k3pStreamCreator buildStreamCreator(Rcpp::IntegerVector seed){
   return streamCreatorHere;
 }
 
+
+
 //' @name createStreamsCpu
 //' @title create streams stored on the CPU
 //' @description streams for random numbers
@@ -357,7 +359,6 @@ Rcpp::IntegerMatrix  createStreamsCpu(
   colnames(result) = CharacterVector::create(
     "current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
     "initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3");
-  //,
   //  "substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3");
 
   size_t streamBufferSize;
@@ -368,8 +369,8 @@ Rcpp::IntegerMatrix  createStreamsCpu(
   
   clrngMrg31k3pStreamCreator streamCreatorHere = buildStreamCreator(seed2);
   
-  Rcpp::Rcout << "a" << streamCreatorHere.initialState.g1[0]<< " " << streamCreatorHere.initialState.g1[1]<< " " << streamCreatorHere.initialState.g1[2]<< "\n";
-  Rcpp::Rcout << "b" << streamCreatorHere.initialState.g2[0]<< " " << streamCreatorHere.initialState.g2[1]<< " " << streamCreatorHere.initialState.g2[2]<< "\n";
+  // Rcpp::Rcout << "a" << streamCreatorHere.initialState.g1[0]<< " " << streamCreatorHere.initialState.g1[1]<< " " << streamCreatorHere.initialState.g1[2]<< "\n";
+  // Rcpp::Rcout << "b" << streamCreatorHere.initialState.g2[0]<< " " << streamCreatorHere.initialState.g2[1]<< " " << streamCreatorHere.initialState.g2[2]<< "\n";
 
   clrngMrg31k3pStream* streams = clrngMrg31k3pCreateStreams(&streamCreatorHere, n[0], 
                                                             &streamBufferSize, &err);//line 299 in mrg31k3p.c

### 2021-10-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
no important changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2889 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
 The remaining sections are organized as follows. 
-In Section 2, we introduce streams and the use of streams in work items on a GPU device, and describe and check the functions for generating uniform and Normal random numbers.
+In Section 2, we introduce streams and the use of streams in work-items on GPU device, and describe and check the functions for generating uniform and Normal random numbers.
 In Section 3, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we parallelise this Monte Carlo simulation task and implement it on GPU, then we provide two real data examples to demonstrate the function usage and test its R performance.
@@
 
-\subsection{Creating streams}
-In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
+\subsection{Creating streams}\label{create_streams}
+In \proglang{clRNG} library, a stream object is a structure that contains three states: the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
 The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
-If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory.
-In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device.
-\fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams.
-\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device.
-From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used.
+If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item picks streams from the global memory and copies their current states only to its private memory.
+\pkg{gpuRandom} is able to create streams both on the host computer and on the GPU device.
+\fct{createStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on CPU as \proglang{R} matrices (see the \proglang{R} output below), in which the substreams are not created because they are not actually used in random number generation.
+\fct{createStreamsGpu} adapted the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on GPU device, so that it can avoid unnecessary data transfer between host and device.
+From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
 In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
@@
 \begin{verbatim}
-# Create streams on host
-streams_on_CPU<-gpuRandom::CreateStreamsCpu(Nstreams=4)
-t(streams_on_CPU)
+# creating streams on cpu
+myStreamsCpu <- createStreamsCpu(n=4)
+t(myStreamsCpu)
 ##               [,1]       [,2]       [,3]       [,4]
@@
 ## initial.g2.3 12345 2094976052 1462898381  795289868
-## substream.g1.1 12345  336690377  502033783  739421137
-## substream.g1.2 12345  597094797 1322587635 1475938232
-## substream.g1.3 12345 1245771585 1964121530  730262207
-## substream.g2.1 12345   85196284 1949818481 1630192198
-## substream.g2.2 12345  523477687 1607232546  324551134
-## substream.g2.3 12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create.
-For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
-By setting a seed same as the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-# Create streams on GPU
-seed <- c(12345, 12345, 12345, 12345, 12345, 12345)
-streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
-streams_from_GPU <- as.matrix(streams1)
-streams2 = deepcopy(streams1)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-t(streams_from_GPU)   
-##                 [,1]       [,2]       [,3]       [,4]
-## current.g1.1   12345  336690377  502033783  739421137
-## current.g1.2   12345  597094797 1322587635 1475938232
-## current.g1.3   12345 1245771585 1964121530  730262207
-## current.g2.1   12345   85196284 1949818481 1630192198
-## current.g2.2   12345  523477687 1607232546  324551134
-## current.g2.3   12345 2094976052 1462898381  795289868
-## initial.g1.1   12345  336690377  502033783  739421137
-## initial.g1.2   12345  597094797 1322587635 1475938232
-## initial.g1.3   12345 1245771585 1964121530  730262207
-## initial.g2.1   12345   85196284 1949818481 1630192198
-## initial.g2.2   12345  523477687 1607232546  324551134
-## initial.g2.3   12345 2094976052 1462898381  795289868
-## substream.g1.1 12345  336690377  502033783  739421137
-## substream.g1.2 12345  597094797 1322587635 1475938232
-## substream.g1.3 12345 1245771585 1964121530  730262207
-## substream.g2.1 12345   85196284 1949818481 1630192198
-## substream.g2.2 12345  523477687 1607232546  324551134
-## substream.g2.3 12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\fct{createStreamsCpu} has two arguments:
 \begin{itemize}
-  \item \code{seedR}: an R vector of length 6, which is the seed of streams. Default is set to a vector of six ``12345''.
-  \item \code{Nstreams}: number of streams to create.
-  \item \code{keepInitial}: logical specifying whether to keep the seed in generated streams. Default is set to TRUE.
+  \item \code{n}: number of streams to create.
+  \item \code{seed}: random seed of streams of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
 \end{itemize}
-\fct{CreateStreamsGpu} in \pkg{gpuRandom} are similar to \code{.Random.seed} in \proglang{R}.
-By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \pkg{gpuRandom}, with the \code{seedR} argument, we are able to produce exactly the same sequence of streams as many times as we want.
+For the MRG31k3p RNG, each state of the stream is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
+By setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+# creating streams on gpu
+#myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu1 = createStreamsGpu(n=4, seed=12345)
+range(as.matrix(myStreamsGpu1) - myStreamsCpu)
+## [1] 0 0
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\fct{createStreamsGpu} also has the two arguments \code{n}, which is the number of streams to create, and \code{seed}, which is the seed of streams, and is also by default a vector of six ``12345''.
+\fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
+By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
+Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we are able to produce exactly the same sequence of streams as many times as we want.
 
 \subsection{Using streams to generate  i.i.d. $U (0, 1)$ random numbers on GPU}
-The streams created on GPU are then used in work items that executes in parallel on GPU device.
-In \pkg{gpuRandom} each work item takes one distinct stream to generate random numbers, it is important to notice that the number of streams should always be equal to the number of total work items in use.
+The streams created on GPU are then used in work-items that executes in parallel on GPU device.
+In \pkg{gpuRandom} each work-item takes one distinct stream to generate random numbers, it is important to notice that the number of streams should always be equal to the number of total work-items in use.
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
@@
 \code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is internal number of columns of the \code{streams} matrix (these variables are defined in macros not shown here).
-Each stream's current state is copied to the private memory of each work item by the function \fct{streamsToPrivate}.
-\code{index} represents the position of the work item when mapping it to a one-dimensional data structure.
+Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}.
+\code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
 The function \fct{clrngMrg31k3pNextState} generates a uniform random integer named \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
@@
 Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example.
-Suppose we have a $2 \times 4$ NDRange (index space size) ((a) in Figure \ref{fig:1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig:1}) of uniform random numbers using 8 streams.
-In each grid of (a) that represents a work item, there are two rows of numbers, the number in the above row is the stream index (from 0 to 7), and a pair of numbers in the second row is the 2-dimensional global ID of the work item.
-Each stream object is passed to each work item.
-The number in a cell of the matrix in (b) corresponds to the stream, it indicates which stream does the random number generation for this matrix cell.
-For example, stream 0 which is allocated in work item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
+Suppose we have a $2 \times 4$ NDRange ((a) in Figure \ref{fig:1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig:1}) of uniform random numbers.
+In each grid of (a) that represents a work-item, there are two rows of numbers, the number in the above row is the stream index (from 0 to 7), and a pair of numbers in the second row is the 2-dimensional global ID of the work-item.
+Each stream object is passed to each work-item.
+The number in a cell of the matrix corresponds to the stream index, it indicates which stream does the random number generation for this matrix cell.
+For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
 \begin{figure}[H]%
@@
 
-Now we use the 4 streams created in Section 2.1 and an NDRange of 1 by 4 work items to generate a vector \code{sim_1} of 6 i.i.d.
+Now we use the 4 streams created in section \ref{create_streams} and an NDRange of 1 by 4 work-items to generate a vector \code{sim_1} of 6 i.i.d.
 $U (0,1)$ random numbers using the function \fct{gpuRandom::runif}.
@@
 \begin{verbatim}
-sim_1 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
-    type = "double")
+myStreamsGpu2 = deepcopy(myStreamsGpu1)
+sim_1 = runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, 4), type = "double")
 as.vector(sim_1)
@@
 \begin{itemize}
-  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or specifying the length if to create a vector.
+  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
   \item \code{streams}: streams used for random number generation.
-  \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
+  \item \code{Nglobal}: global index space (NDRange). Default is set as $(64,8)$.
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.
 \end{itemize}
-Next we show that by using a same stream \code{streams2} as \code{streams1},  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-sim_2 <- gpuRandom::runif(n = 6, streams = streams2, Nglobal = c(1, 4),
-    type = "double")
+Next we show that by using a stream \code{myStreamsGpu2} same as \code{myStreamsGpu1},  \fct{runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+sim_2 <- runif(n = 6, streams = myStreamsGpu2, Nglobal = c(1, 4), type = "double")
 as.vector(sim_2)
@@
 \end{knitrout}
-Notice that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used.
-Reusing \code{streams1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
-Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{streams1} after it being used.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-sim_3 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
-    type = "double")
+Notice that \code{myStreamsGpu2} is a \code{deepcopy} of \code{myStreamsGpu1} before \code{myStreamsGpu1} is used.
+Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
+Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{myStreamsGpu1} after it being used.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+sim_3 = runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, 4), type = "double")
 as.vector(sim_3)
 ## [1] 0.6487742 0.1112075 0.3661944 0.5018562 0.1088229 0.3114331
-t(as.matrix(streams1))[1:6, ]
+t(as.matrix(myStreamsGpu1))[1:6, ]
 ##            [,1]       [,2]       [,3]       [,4]
@@
 \begin{verbatim}
-saveRDS(streams_from_GPU, "streams_from_GPU.rds")
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
 # Load the streams object as streams_saved
-streams_saved <- gpuR::vclMatrix(readRDS("streams_from_GPU.rds"))
-sim_saved = gpuRandom::runif(n=6,streams=streams_saved, 
-                             Nglobal=c(1,4), type="double")
+streams_saved <- vclMatrix(readRDS("myStreams.rds"))
+sim_saved = runif(n=6,streams=streams_saved, Nglobal=c(1,4), type="double")
 as.vector(sim_saved)
@@
 We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
-The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} does come from a standard uniform distribution.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streams <- CreateStreamsGpu(Nstreams =512, keepInitial=1)
-random_vector<-gpuRandom::runif(n=50000, streams=streams, 
-                                Nglobal=c(64,8), type="double")
-random_vector<-as.vector(random_vector)
+The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+streams <- createStreamsGpu(n = 512)
+random_u<-runif(n=50000, streams=streams, Nglobal=c(64,8), type="double")
+random_u<-as.vector(random_u)
 \end{verbatim}
@@
 \end{knitrout}
-\caption{histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
+\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
 \end{figure}
@@
 \code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
-As the work-items in each work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), (X_2,Y_2), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
+As the work-items in a work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
 %
@@
 const int startvalue=index * NpadStreams;
-
 local double part[2], cosPart1, sinPart1;
@@
     }
-
     barrier(CLK_LOCAL_MEM_FENCE);
@@
 \end{framed}
-We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange and 8 streams.
-Work items are organized to be 1 by 2 work groups as Gaussian numbers are generated in pairs using the Box-Muller method.
-The shading on the grids in the NDRange distinguishes work groups.
-Each work group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where work item of `\code{get_local_id(1)}$=0$' generates the \code{X} of a pair, and work item of `\code{get_local_id(1)}$=1$' generates the \code{Y} of a pair.
-The numbers in the third row in each gird of NDRange represents the local index of each work item.
-Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work items generate random numbers in the leftmost two columns (8 cells) of matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
-Therefore, each work item generates one Gaussian random number in each iteration, the matrix filling process is the same as that shown in Figure \ref{fig:1}.
+We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange.
+Each stream is copied to the private memory of each work-item.
+Work-items are organized to be 1 by 2 work-groups as the random numbers are generated in pairs using the Box-Muller method.
+The shading on the grids in the NDRange distinguishes work-groups.
+The numbers in the third row in each gird of NDRange represents the local index of each work-item.
+Each work-group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where work-item of `\code{get_local_id(1)}$=0$' generates the \code{X} of a pair, and work-item of `\code{get_local_id(1)}$=1$' generates the \code{Y} of a pair.
+Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
+Therefore, each work-item generates one Gaussian random number in each iteration, the matrix filling process is the same as that shown in Figure \ref{fig:1}.
 
@@
 \begin{verbatim}
-streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
+streams <- createStreamsGpu(n =512*128)
 normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                 Nglobal=c(512,128), type="float")
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
 avector<-as.vector(as.matrix(normal_matrix))
@@
 ##    user  system elapsed 
-##   0.140   0.004   0.143
+##   0.143   0.004   0.147
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   6.610   0.862   7.466
+##   6.678   0.850   7.522
 \end{verbatim}
@@
 \begin{verbatim}
+almost.1 <- 1 + 64 * .Machine$double.eps
+observed_m= -logfactSum(month, c(16,16))/almost.1
+
+month_gpu<-vclMatrix(month,type="integer")
+result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
+                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
+result_month$simNum
+## [1] 1015808
+result_month$counts
+## [1] 410644
+result_month$p.value
+## [1] 0.4042541
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+#using GPU
+system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
+            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
+##    user  system elapsed 
+##   0.769   0.001   0.770
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
 #using CPU
-stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
-## [1] 0.4042196
-system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
+stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
+## [1] 0.4037993
+system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  14.578   0.002  14.569
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for one million simulations, has a p-value of 0.4035.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-almost.1 <- 1 + 64 * .Machine$double.eps
-observed_m= -gpuRandom::logfactSum(month, c(16,16))/almost.1
-
-month_GPU<-vclMatrix(month,type="integer")
-result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
-result_monthgpu$simNum
-## [1] 1015808
-result_monthgpu$counts
-## [1] 410098
-result_monthgpu$p.value
-## [1] 0.4037166
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using GPU
-system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
-##    user  system elapsed 
-##   0.702   0.005   0.706
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.
+##  14.962   0.017  14.970
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for 1015808 simulations, has a p-value of 0.4035.
 
@@
 \end{knitrout}
-We perform 10 million simulations for the ``week'' table.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using CPU
-fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
-## [1] 0.0001284
-system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
-##    user  system elapsed 
-##  90.259   0.150  90.439
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
+
 \begin{knitrout}
@@
 observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
-
 week_GPU<-gpuR::vclMatrix(week,type="integer")
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = Nglobal)
+                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
 result_weekgpu$simNum
-## [1] 10010624
+## [1] 10027008
 result_weekgpu$counts
-## [1] 1260
+## [1] 1242
 result_weekgpu$p.value
-## [1] 0.0001259662
+## [1] 0.0001239652
 \end{verbatim}
@@
 system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
+            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
 ##    user  system elapsed 
-##   4.477   0.009   4.483
+##   4.640   0.012   4.649
 \end{verbatim}
@@
 \fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+#using CPU
+fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
+## [1] 0.0001263657
+system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
+##    user  system elapsed 
+##  90.996   0.170  91.109
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
 
@@
 \begin{verbatim}
-streamsGpu <- gpuRandom::CreateStreamsGpu(Nstreams=128*64, keepInitial=TRUE)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-zmatGpu = gpuRandom::rnorm(c(nrow(maternCov),2), 
-                           streams=streamsGpu, Nglobal=c(128,64),
+streamsGpu <- createStreamsGpu(n=128*64)
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+zmatGpu = rnorm(c(nrow(maternCov),2), streams=streamsGpu, Nglobal=c(128,64),
                  type = theType)

### 2021-10-05 pkg/gpuRandom/R/CreateStreamsCpu.R

@@ -1,1 +1,0 @@
-# now in the cpp file streamsCpu.cpp

### 2021-10-05 pkg/gpuRandom/R/CreateStreamsGpu.R

@@ -12,14 +12,13 @@
 
 
 createStreamsGpu = function(n, 
-                        seed=12345){
+                            seed=12345){
 
     seed = rep_len(seed, 6)
     seedVec <- gpuR::vclVector(as.integer(seed), type="integer")  
     streamsR<-gpuR::vclMatrix(0L, nrow=as.integer(n), ncol=12, type="integer")
 
-    keepInitial=1
-    gpuRandom:::CreateStreamsGpuBackend(seedVec, streamsR, keepInitial)
+    gpuRandom:::CreateStreamsGpuBackend(seedVec, streamsR, keepInitial=1)
   
     streamsR
   

### 2021-10-05 pkg/gpuRandom/R/fisher_sim.R

@@ -99,8 +99,8 @@ fisher.sim=function(
   
   #  print(class(xVcl))
   
-  simPerWork<-ceiling(N/prod(Nglobal))
-  TotalSim<-simPerWork*prod(Nglobal)
+  simPerItem<-ceiling(N/prod(Nglobal))
+  TotalSim<-simPerItem*prod(Nglobal)
   #remainder<-C%%prod(Nglobal)
   
   if(returnStatistics) {
@@ -112,7 +112,7 @@ fisher.sim=function(
   
   PVAL <- NULL
   
-  counts<-cpp_gpuFisher_test(x, results, simPerWork, streams, Nglobal,Nlocal)
+  counts<-cpp_gpuFisher_test(x, results, simPerItem, streams, Nglobal,Nlocal)
   
   #theTime<-system.time(cpp_gpuFisher_test(x, results, as.integer(B), streams, Nglobal,Nlocal))
   
@@ -137,7 +137,7 @@ fisher.sim=function(
     
   }else {
     
-    theResult = list(p.value=PVAL, simNum=TotalSim, counts=counts, streams=streams)
+    theResult = list(p.value = PVAL, simNum=TotalSim, counts=counts, streams=streams)
   }
   
   theResult

### 2021-10-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
no important changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2889 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 setContext(   grep('gpu', listContexts()$device_type) [1]    )
-#setwd("/home/ruoyong/paper1_2021_5")
+#setwd("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5")
 @
@@
 
-\subsection{Creating streams}
-
-{\tt changed createStreams functions to be compatible with each other}
-
-create streams on cpu
-<<createStreams, size="tiny">>=
-myStreamsCpu <- createStreamsCpu(n=4, seed = 12345)
+\subsection{Creating streams}\label{create_streams}
+In \proglang{clRNG} library, a stream object is a structure that contains three states: the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
+The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
+If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item picks streams from the global memory and copies their current states only to its private memory.
+\pkg{gpuRandom} is able to create streams both on the host computer and on the GPU device.
+\fct{createStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on CPU as \proglang{R} matrices (see the \proglang{R} output below), in which the substreams are not created because they are not actually used in random number generation.
+\fct{createStreamsGpu} adapted the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on GPU device, so that it can avoid unnecessary data transfer between host and device.
+From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
+In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
+<<createStreams_CPU>>=
+# creating streams on cpu
+myStreamsCpu <- createStreamsCpu(n=4)
 t(myStreamsCpu)
 @
-
-streams on gpu
-<<createStreamsGpu>>=
-myStreamsGpu1 = vclMatrix(myStreamsCpu)
-myStreamsGpu2 = createStreams(n=4, seed=12345)
-range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
+\fct{createStreamsCpu} has two arguments:
+\begin{itemize}
+  \item \code{n}: number of streams to create.
+  \item \code{seed}: random seed of streams of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
+\end{itemize}
+For the MRG31k3p RNG, each state of the stream is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
+By setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
+<<createStreams_Gpu>>=
+# creating streams on gpu
+#myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu1 = createStreamsGpu(n=4, seed=12345)
+range(as.matrix(myStreamsGpu1) - myStreamsCpu)
+matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
 @
-
-<<simStuff, eval=TRUE,tidy=TRUE>>=
-sim_2 <-gpuRandom::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double")
-as.vector(sim_2)
-@
+\fct{createStreamsGpu} also has the two arguments \code{n}, which is the number of streams to create, and \code{seed}, which is the seed of streams, and is also by default a vector of six ``12345''.
+\fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
+By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
+Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we are able to produce exactly the same sequence of streams as many times as we want.
 
@@
 
-In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
-The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
-If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory.
-In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device.
-\fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams.
-\fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device.
-From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used.
-In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
-<<streams on CPU>>=
-# Create streams on host
-streams_on_CPU<-gpuRandom::CreateStreamsCpu(seed = rep(12345, 6), Nstreams=4)
-t(streams_on_CPU)
-@
-\fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create.
-For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
-By setting a seed same as the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
-<<streams on GPU1, eval=TRUE>>=
-# Create streams on GPU
-seed <- c(12345, 12345, 12345, 12345, 12345, 12345)
-streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
-streams_from_GPU <- as.matrix(streams1)
-streams2 = deepcopy(streams1)
-@ 
-<<streams on GPU2, eval=TRUE, echo=FALSE>>=
-gpuR::colnames(streams_from_GPU) = c("current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
-"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3")
-#"substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")   
-@ 
-<<streams on GPU3,eval=TRUE>>=
-t(streams_from_GPU)   
-@
-The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
-\begin{itemize}
-  \item \code{seedR}: an R vector of length 6, which is the seed of streams. Default is set to a vector of six ``12345''.
-  \item \code{Nstreams}: number of streams to create.
-  \item \code{keepInitial}: logical specifying whether to keep the seed in generated streams. Default is set to TRUE.
-\end{itemize}
-\fct{CreateStreamsGpu} in \pkg{gpuRandom} are similar to \code{.Random.seed} in \proglang{R}.
-By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \pkg{gpuRandom}, with the \code{seedR} argument, we are able to produce exactly the same sequence of streams as many times as we want.
-
 \subsection{Using streams to generate  i.i.d. $U (0, 1)$ random numbers on GPU}
-The streams created on GPU are then used in work items that executes in parallel on GPU device.
-In \pkg{gpuRandom} each work item takes one distinct stream to generate random numbers, it is important to notice that the number of streams should always be equal to the number of total work items in use.
+The streams created on GPU are then used in work-items that executes in parallel on GPU device.
+In \pkg{gpuRandom} each work-item takes one distinct stream to generate random numbers, it is important to notice that the number of streams should always be equal to the number of total work-items in use.
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
@@
 <<eval=TRUE, tidy=TRUE>>=
-sim_1 = gpuRandom::runif(n=6,streams=streams1, Nglobal=c(1,4), type="double")
+myStreamsGpu2 = deepcopy(myStreamsGpu1)
+sim_1 = runif(n=6,streams=myStreamsGpu1, Nglobal=c(1,4), type="double")
 as.vector(sim_1)
@@
 \begin{itemize}
-  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or specifying the length if to create a vector.
+  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
   \item \code{streams}: streams used for random number generation.
-  \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
+  \item \code{Nglobal}: global index space (NDRange). Default is set as $(64,8)$.
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.
 \end{itemize}
-Next we show that by using a same stream \code{streams2} as \code{streams1},  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
+Next we show that by using a stream \code{myStreamsGpu2} same as \code{myStreamsGpu1},  \fct{runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
 <<eval=TRUE,tidy=TRUE>>=
-sim_2 <-gpuRandom::runif(n=6, streams=streams2, Nglobal=c(1,4), type = "double")
+sim_2 <-runif(n=6, streams=myStreamsGpu2, Nglobal=c(1,4), type = "double")
 as.vector(sim_2)
 @
-Notice that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used.
-Reusing \code{streams1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
-Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{streams1} after it being used.
+Notice that \code{myStreamsGpu2} is a \code{deepcopy} of \code{myStreamsGpu1} before \code{myStreamsGpu1} is used.
+Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
+Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{myStreamsGpu1} after it being used.
 <<eval=TRUE,tidy=TRUE>>=
-sim_3 = gpuRandom::runif(n=6,streams=streams1, Nglobal=c(1,4), type="double")
+sim_3 = runif(n=6,streams=myStreamsGpu1, Nglobal=c(1,4), type="double")
 as.vector(sim_3)
-t(as.matrix(streams1))[1:6,]
+t(as.matrix(myStreamsGpu1))[1:6,]
 @
@@
 Below is a trivial example, in which we demonstrate saving \code{streams1} to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU to produce the vector called \code{sim_saved}, which is also the same as \code{sim_1}.
-<<echo=TRUE,eval=FALSE>>=
-saveRDS(streams_from_GPU, "streams_from_GPU.rds")
-@
-<<save streams,eval=TRUE>>=
+<<save streams,echo=TRUE,eval=TRUE>>=
+saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
 # Load the streams object as streams_saved
-streams_saved <- gpuR::vclMatrix(readRDS("streams_from_GPU.rds"))
-sim_saved = gpuRandom::runif(n=6,streams=streams_saved, 
-                             Nglobal=c(1,4), type="double")
+streams_saved <- vclMatrix(readRDS("myStreams.rds"))
+sim_saved = runif(n=6,streams=streams_saved, Nglobal=c(1,4), type="double")
 as.vector(sim_saved)
@@
 We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
-The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} does come from a standard uniform distribution.
+The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
 <<uniforms_random_vector,eval=TRUE>>=
-streams <- CreateStreamsGpu(Nstreams =512, keepInitial=1)
-random_vector<-gpuRandom::runif(n=50000, streams=streams, 
-                                Nglobal=c(64,8), type="double")
-random_vector<-as.vector(random_vector)
+streams <- createStreamsGpu(n = 512)
+random_u<-runif(n=50000, streams=streams, Nglobal=c(64,8), type="double")
+random_u<-as.vector(random_u)
 @
@@
 <<largenumberuniforms, eval=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold'>>=
-hist(random_vector, prob = TRUE, main="Histogram")
-plot(ecdf(random_vector), ylab="F(x)",main="Empirical cumulative distribution")
+hist(random_u, prob = TRUE, main="Histogram")
+plot(ecdf(random_u), ylab="F(x)",main="Empirical cumulative distribution")
 curve(punif(x, 0, 1), lty=2, add=TRUE, col="red")
 @
-\caption{histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
+\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
 \end{figure}
@@
 We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we adapted the \proglang{R} function \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on GPU, which can speed up the Q-Q plot greatly.
-<<histAndQQ_prepare,eval=TRUE>>=
-streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
+<<histAndQQ_prepare,eval=TRUE,cache=TRUE>>=
+streams <- createStreamsGpu(n =512*128)
 normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                 Nglobal=c(512,128), type="float")
-@
-<<histAndQQ_prepare_for_cache, eval=TRUE, cache=TRUE>>=
 avector<-as.vector(as.matrix(normal_matrix))
@@
 colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
-Nglobal = c(128,128)
-streams <- CreateStreamsGpu(Nstreams =128^2, keepInitial=1)
+Nglobal = c(256,128)
+streams <- createStreamsGpu(n =256*128)
 @
@@
 
-<<timecomparemonth2Cpu, cache=TRUE, eval=TRUE>>=
-#using CPU
-stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
-system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
-@
-\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for one million simulations, has a p-value of 0.4035.
 <<timecomparemonth3Gpu, eval=TRUE,cache=FALSE>>=
 almost.1 <- 1 + 64 * .Machine$double.eps
-observed_m= -gpuRandom::logfactSum(month, c(16,16))/almost.1
+observed_m= -logfactSum(month, c(16,16))/almost.1
 
-month_GPU<-vclMatrix(month,type="integer")
-result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
-result_monthgpu$simNum
-result_monthgpu$counts
-result_monthgpu$p.value
+month_gpu<-vclMatrix(month,type="integer")
+result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
+                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
+result_month$simNum
+result_month$counts
+result_month$p.value
 @
-We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
+We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
 <<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
 #using GPU
-system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
+system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
+            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
 @
@@
 <<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
-result_monthgpu$cpu <- as.vector(result_monthgpu$sim)
+month_stats <- as.vector(result_month$sim)
 @
+
+<<timecomparemonth2Cpu, cache=TRUE, eval=TRUE>>=
+#using CPU
+stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
+system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
+@
+\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for 1015808 simulations, has a p-value of 0.4035.
 
@@
 @
-
 <<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
@@
 @
-We perform 10 million simulations for the ``week'' table.
-<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
-#using CPU
-fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
-system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
-@
-The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
+
 <<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
 observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
-
 week_GPU<-gpuR::vclMatrix(week,type="integer")
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = c(128,128))
+                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
 result_weekgpu$simNum
@@
 system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = c(128,128)))
+            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
 @
@@
 @
+
+<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
+#using CPU
+fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
+system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
+@
+The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
 
@@
 <<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE>>=
-hist(result_monthgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}))$'), breaks=40, ylab="Frequency", main="Month data")
+hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}))$'), breaks=40, ylab="Frequency", main="Month data")
 abline(v = observed_m, col = "blue", lwd = 1.5)
@@
 <<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
-streamsGpu <- gpuRandom::CreateStreamsGpu(Nstreams=128*64, keepInitial=TRUE)
+streamsGpu <- createStreamsGpu(n=128*64)
 @
 <<r randomNormalsSim2, eval=TRUE, cache=FALSE>>=
-zmatGpu = gpuRandom::rnorm(c(nrow(maternCov),2), 
-                           streams=streamsGpu, Nglobal=c(128,64),
+zmatGpu = rnorm(c(nrow(maternCov),2), streams=streamsGpu, Nglobal=c(128,64),
                  type = theType)

### 2021-10-05 pkg/gpuRandom/R/RcppExports.R

@@ -140,7 +140,7 @@ multiplyLowerBatchBackend <- function(C, A, B, diagIsOne, Nglobal, Nlocal, Nloca
 #' @param n number of streams to create
 #' @param seed random seed, length 6, recycled if shorter
 #' @export
-createStreamsCpu <- function(n, seed) {
+createStreamsCpu <- function(n, seed = as.integer( c(12345,12345,12345,12345,12345,12345))) {
     .Call('_gpuRandom_createStreamsCpu', PACKAGE = 'gpuRandom', n, seed)
 }
 

### 2021-10-05 pkg/gpuRandom/R/runif.R

@@ -1,9 +1,9 @@
 #' @title runif
 #' 
 #' @param n a number or a vector specifying the size of output matrix
-#' @param streams streams object
-#' @param Nglobal the size of the index space for use
-#' @param type "double" or "float" format of generated random numbers
+#' @param streams streams matrix of random number seeds
+#' @param Nglobal vector of length 2, number of work items
+#' @param type random number type, one of "uniform" or "normal"
 #' 
 #' @return a vclVector or vclMatrix of uniform random numbers
 #' 

### 2021-10-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
adds swissrain in vignettes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2890 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 # creating streams on gpu
-#myStreamsGpu1 = vclMatrix(myStreamsCpu)
-myStreamsGpu1 = createStreamsGpu(n=4, seed=12345)
-range(as.matrix(myStreamsGpu1) - myStreamsCpu)
-matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
+myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu2 = createStreamsGpu(n=4, seed=12345)
+matrix(as.matrix(myStreamsGpu2), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
+range(as.matrix(myStreamsGpu2) - myStreamsCpu)
 @
@@
 Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we are able to produce exactly the same sequence of streams as many times as we want.
-
-  check seed
-<<simSeed, size='tiny'>>=
-  matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
-@
-
-  check reproducibility
-<<simStuff2, eval=TRUE,tidy=TRUE>>=
-  as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
-as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
-@
 
@@
 The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
+<<simStuff2, eval=TRUE,tidy=TRUE>>=
+as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
+@
+
 <<eval=TRUE, tidy=TRUE>>=
-myStreamsGpu2 = deepcopy(myStreamsGpu1)
-sim_1 = runif(n=6,streams=myStreamsGpu1, Nglobal=c(1,4), type="double")
+myStreamsGpu3 = deepcopy(myStreamsGpu2)
+sim_1 = runif(n=6,streams=myStreamsGpu2, Nglobal=c(1,4), type="double")
 as.vector(sim_1)
@@
 \end{itemize}
-Next we show that by using a stream \code{myStreamsGpu2} same as \code{myStreamsGpu1},  \fct{runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
+Next we show that by using a stream \code{myStreamsGpu3} same as \code{myStreamsGpu2},  \fct{runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
 <<eval=TRUE,tidy=TRUE>>=
-sim_2 <-runif(n=6, streams=myStreamsGpu2, Nglobal=c(1,4), type = "double")
+sim_2 <-runif(n=6, streams=myStreamsGpu3, Nglobal=c(1,4), type = "double")
 as.vector(sim_2)
 @
-Notice that \code{myStreamsGpu2} is a \code{deepcopy} of \code{myStreamsGpu1} before \code{myStreamsGpu1} is used.
-Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
-Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{myStreamsGpu1} after it being used.
+Notice that \code{myStreamsGpu2} is a \code{deepcopy} of \code{myStreamsGpu2} before \code{myStreamsGpu2} is used.
+Reusing \code{myStreamsGpu2} will produce a vector different from \code{sim_1} (see \code{sim_3}).
+Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{myStreamsGpu2} after it being used.
 <<eval=TRUE,tidy=TRUE>>=
-sim_3 = runif(n=6,streams=myStreamsGpu1, Nglobal=c(1,4), type="double")
+sim_3 = runif(n=6,streams=myStreamsGpu2, Nglobal=c(1,4), type="double")
 as.vector(sim_3)
-t(as.matrix(myStreamsGpu1))[1:6,]
+t(as.matrix(myStreamsGpu2))[1:6,]
 @

### 2021-10-06 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2891 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{Creating streams}\label{create_streams}
-In \proglang{clRNG} library, a stream object is a structure that contains three states: the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream.
+In \proglang{clRNG} library, a stream object is a structure that contains three states: the current state of the stream, the initial state (or seed) of the stream, and the initial state of the current substream.
 The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
@@
 For the MRG31k3p RNG, each state of the stream is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
-By setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
+We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams directly on GPU using \fct{createStreamsGpu}.
+Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
 \begin{knitrout}
@@
 # creating streams on gpu
-#myStreamsGpu1 = vclMatrix(myStreamsCpu)
-myStreamsGpu1 = createStreamsGpu(n=4, seed=12345)
-range(as.matrix(myStreamsGpu1) - myStreamsCpu)
+myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu2 = createStreamsGpu(n=4)
+range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
 ## [1] 0 0
@@
 \end{knitrout}
-\fct{createStreamsGpu} also has the two arguments \code{n}, which is the number of streams to create, and \code{seed}, which is the seed of streams, and is also by default a vector of six ``12345''.
+\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''.
 \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
@@
 
-Now we use the 4 streams created in section \ref{create_streams} and an NDRange of 1 by 4 work-items to generate a vector \code{sim_1} of 6 i.i.d.
-$U (0,1)$ random numbers using the function \fct{gpuRandom::runif}.
+Now we use the streams created in section \ref{create_streams} to generate a vector \code{sim_1} of 6 i.i.d.
+$U (0,1)$ random numbers using the function \fct{runif}.
 If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied.
@@
 \begin{verbatim}
-myStreamsGpu2 = deepcopy(myStreamsGpu1)
-sim_1 = runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, 4), type = "double")
+sim_1 <- runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, nrow(myStreamsGpu1)),
+    type = "double")
 as.vector(sim_1)
@@
   \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: streams used for random number generation.
-  \item \code{Nglobal}: global index space (NDRange). Default is set as $(64,8)$.
+  \item \code{streams}: streams for random number generation.
+  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.
 \end{itemize}
-Next we show that by using a stream \code{myStreamsGpu2} same as \code{myStreamsGpu1},  \fct{runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
+Next we show that by using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers as \code{sim_1}. 
 \begin{knitrout}
@@
 \begin{verbatim}
-sim_2 <- runif(n = 6, streams = myStreamsGpu2, Nglobal = c(1, 4), type = "double")
+sim_2 = runif(n = 7, streams = myStreamsGpu2, Nglobal = c(1, nrow(myStreamsGpu1)),
+    type = "double")
 as.vector(sim_2)
 ## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
+## [7] 0.6487742
 \end{verbatim}
@@
 \end{knitrout}
-Notice that \code{myStreamsGpu2} is a \code{deepcopy} of \code{myStreamsGpu1} before \code{myStreamsGpu1} is used.
-Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1} (see \code{sim_3}).
-Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{myStreamsGpu1} after it being used.
+Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
+Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
 \begin{knitrout}
@@
 \begin{verbatim}
-sim_3 = runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, 4), type = "double")
-as.vector(sim_3)
-## [1] 0.6487742 0.1112075 0.3661944 0.5018562 0.1088229 0.3114331
-t(as.matrix(myStreamsGpu1))[1:6, ]
+t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu),
+    dimnames = dimnames(myStreamsCpu)))
 ##                    [,1]       [,2]       [,3]       [,4]
-## [1,] 1167281028  640923766  502033783  739421137
-## [2,] 1918428443 1912157188 1322587635 1475938232
-## [3,] 1858462085  286315187 1964121530  730262207
-## [4,]  933585541 2119609883 1949818481 1630192198
-## [5,] 1132031887  834429287 1607232546  324551134
-## [6,]  465230163   47498872 1462898381  795289868
+## current.g1.1  878672095 2113333390  502033783  739421137
+## current.g1.2  240667857  559530223 1322587635 1475938232
+## current.g1.3  240667857 1309565828 1964121530  730262207
+## current.g2.1  642281259 1335994581 1949818481 1630192198
+## current.g2.2 1069151070   61444481 1607232546  324551134
+## current.g2.3  809054265  197003928 1462898381  795289868
+## initial.g1.1      12345  336690377  502033783  739421137
+## initial.g1.2      12345  597094797 1322587635 1475938232
+## initial.g1.3      12345 1245771585 1964121530  730262207
+## initial.g2.1      12345   85196284 1949818481 1630192198
+## initial.g2.2      12345  523477687 1607232546  324551134
+## initial.g2.3      12345 2094976052 1462898381  795289868
 \end{verbatim}
@@
 However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations.
-Below is a trivial example, in which we demonstrate saving \code{streams1} to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU to produce the vector called \code{sim_saved}, which is also the same as \code{sim_1}.
+Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
 \begin{knitrout}
@@
 streams_saved <- vclMatrix(readRDS("myStreams.rds"))
-sim_saved = runif(n=6,streams=streams_saved, Nglobal=c(1,4), type="double")
-as.vector(sim_saved)
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.769   0.001   0.770
+##   0.844   0.004   0.847
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.640   0.012   4.649
+##   4.555   0.012   4.564
 \end{verbatim}

### 2021-10-06 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2891 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 For the MRG31k3p RNG, each state of the stream is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
-By setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
+We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams directly on GPU using \fct{createStreamsGpu}.
+Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
 <<createStreams_Gpu>>=
@@
 myStreamsGpu1 = vclMatrix(myStreamsCpu)
-myStreamsGpu2 = createStreamsGpu(n=4, seed=12345)
-matrix(as.matrix(myStreamsGpu2), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
-range(as.matrix(myStreamsGpu2) - myStreamsCpu)
+myStreamsGpu2 = createStreamsGpu(n=4)
+range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
 @
-\fct{createStreamsGpu} also has the two arguments \code{n}, which is the number of streams to create, and \code{seed}, which is the seed of streams, and is also by default a vector of six ``12345''.
+\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''.
 \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
@@
 <<simStuff2, eval=TRUE,tidy=TRUE>>=
-as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
-as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
-@
-
-<<eval=TRUE, tidy=TRUE>>=
-myStreamsGpu3 = deepcopy(myStreamsGpu2)
-sim_1 = runif(n=6,streams=myStreamsGpu2, Nglobal=c(1,4), type="double")
+sim_1 <-runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type="double")
 as.vector(sim_1)
@@
   \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: streams used for random number generation.
-  \item \code{Nglobal}: global index space (NDRange). Default is set as $(64,8)$.
+  \item \code{streams}: streams for random number generation.
+  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.
 \end{itemize}
-Next we show that by using a stream \code{myStreamsGpu3} same as \code{myStreamsGpu2},  \fct{runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
-<<eval=TRUE,tidy=TRUE>>=
-sim_2 <-runif(n=6, streams=myStreamsGpu3, Nglobal=c(1,4), type = "double")
+Next we show that by using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers as \code{sim_1}. 
+<<eval=TRUE, tidy=TRUE>>=
+sim_2 = runif(n=7,streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type="double")
 as.vector(sim_2)
 @
-Notice that \code{myStreamsGpu2} is a \code{deepcopy} of \code{myStreamsGpu2} before \code{myStreamsGpu2} is used.
-Reusing \code{myStreamsGpu2} will produce a vector different from \code{sim_1} (see \code{sim_3}).
-Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{myStreamsGpu2} after it being used.
+Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
+Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
 <<eval=TRUE,tidy=TRUE>>=
-sim_3 = runif(n=6,streams=myStreamsGpu2, Nglobal=c(1,4), type="double")
-as.vector(sim_3)
-t(as.matrix(myStreamsGpu2))[1:6,]
+t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
 @
@@
 However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations.
-Below is a trivial example, in which we demonstrate saving \code{streams1} to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU to produce the vector called \code{sim_saved}, which is also the same as \code{sim_1}.
+Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
 <<save streams,echo=TRUE,eval=TRUE>>=
@@
 streams_saved <- vclMatrix(readRDS("myStreams.rds"))
-sim_saved = runif(n=6,streams=streams_saved, Nglobal=c(1,4), type="double")
-as.vector(sim_saved)
 @

### 2021-10-06 pkg/gpuRandom/src/likfitGpu_0.cpp
matern fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2892 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,4 +1,3 @@
-
 #include "lgmlikFit.hpp"
 //#include "gpuRandom.hpp"
 //#include<Rmath.h>
@@ -628,7 +627,6 @@ void likfitGpu_Backend(
 }
 
 
-//#endif
 
 
 

### 2021-10-08 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1 update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2896 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 # creating streams on cpu
-myStreamsCpu <- createStreamsCpu(n=4)
+myStreamsCpu <- createStreamsCpu(n=4, seed=12345)
 t(myStreamsCpu)
@@
 myStreamsGpu1 = vclMatrix(myStreamsCpu)
-myStreamsGpu2 = createStreamsGpu(n=4)
+myStreamsGpu2 = createStreamsGpu(n=4, seed=12345)
 range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
@@
 
-Now we use the streams created in section \ref{create_streams} to generate a vector \code{sim_1} of 6 i.i.d.
+Now we use the streams created in section \ref{create_streams} to generate two vectors of 6 i.i.d.
 $U (0,1)$ random numbers using the function \fct{runif}.
@@
 The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
+By using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers.
 \begin{knitrout}
@@
 \begin{verbatim}
-sim_1 <- runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, nrow(myStreamsGpu1)),
-    type = "double")
-as.vector(sim_1)
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
+as.vector(gpuRandom::runif(n = 4, streams = myStreamsGpu1, Nglobal = c(1,
+    nrow(myStreamsGpu1)), type = "double"))
+## [1] 0.7353245 0.5180770 0.6142074 0.2319392
+as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
+    nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
+## [1] 0.6487742 0.1112075 0.3661944 0.5018562
 \end{verbatim}
@@
 \end{itemize}
-Next we show that by using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers as \code{sim_1}. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-sim_2 = runif(n = 7, streams = myStreamsGpu2, Nglobal = c(1, nrow(myStreamsGpu1)),
-    type = "double")
-as.vector(sim_2)
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
-## [7] 0.6487742
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+
 Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
@@
 ##                    [,1]       [,2]       [,3]       [,4]
-## current.g1.1  878672095 2113333390  502033783  739421137
-## current.g1.2  240667857  559530223 1322587635 1475938232
-## current.g1.3  240667857 1309565828 1964121530  730262207
-## current.g2.1  642281259 1335994581 1949818481 1630192198
-## current.g2.2 1069151070   61444481 1607232546  324551134
-## current.g2.3  809054265  197003928 1462898381  795289868
+## current.g1.1  240667857  559530223  502033783  739421137
+## current.g1.2  240667857 1309565828 1322587635 1475938232
+## current.g1.3      12345  336690377 1964121530  730262207
+## current.g2.1 1069151070   61444481 1949818481 1630192198
+## current.g2.2  809054265  197003928 1607232546  324551134
+## current.g2.3      12345   85196284 1462898381  795289868
 ## initial.g1.1      12345  336690377  502033783  739421137
@@
 ##    user  system elapsed 
-##   0.844   0.004   0.847
+##   0.764   0.000   0.764
 \end{verbatim}
@@
 ##    user  system elapsed 
-##  14.962   0.017  14.970
+##  14.747   0.004  14.743
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.555   0.012   4.564
+##   4.343   0.012   4.352
 \end{verbatim}
@@
 ##    user  system elapsed 
-##  90.996   0.170  91.109
+##  90.524   0.155  90.629
 \end{verbatim}
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in vectorToVCL(data, 4L, context\_index - 1): ViennaCL: FATAL ERROR: CL\_OUT\_OF\_RESOURCES \\\#\# \ ViennaCL tried to launch a compute kernel, but the device does not provide enough resources.
+Try changing the global and local work item sizes.\\\#\# If you think that this is a bug in ViennaCL, please report it at viennacl-support@lists.sourceforge.net and supply at least the following information:\\\#\# \ * Operating System\\\#\# \ * Which OpenCL implementation (AMD, NVIDIA, etc.)\\\#\# \ * ViennaCL version\\\#\# Many thanks in advance!}}\end{kframe}
 \end{knitrout}
@@
                    type = theType)
+\end{verbatim}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'zmatGpu' not found}}\begin{verbatim}
 gpuRandom::multiplyLowerDiagonalBatch(

### 2021-10-08 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
matern fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2892 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
paper 1 update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2896 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
+By using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers.
 <<simStuff2, eval=TRUE,tidy=TRUE>>=
-sim_1 <-runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type="double")
-as.vector(sim_1)
+as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
 @
@@
 \end{itemize}
-Next we show that by using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers as \code{sim_1}. 
-<<eval=TRUE, tidy=TRUE>>=
+<<include=FALSE, tidy=TRUE>>=
 sim_2 = runif(n=7,streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type="double")

### 2021-10-10 pkg/gpuRandom/R/crossprodBatch.R
new examples in vignnettes, weird problems
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2902 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,8 +1,13 @@
 #' @title crossprodBatch function on GPU
 #'
-
+#' @param A rectangular matrices
+#' @param D rectangular matrix, columns are diagonals
+#' @param invertD set to 1 for C = t(A) D^(-1) A
+#' @param Nglobal vector of number of global work items
+#' @param Nlocal vector of number of local work items
+#' @param NlocalCache elements in local cache
 #' 
-#' @return  C = A^T A or A^T D A or A^T D^(-1) A 
+#' @return  C = A^T A or A^T D A or A^T D^(-1) A, output matrices, stacked row-wise 
 #' @useDynLib gpuRandom
 #' @export
 

### 2021-10-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
debugging
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2903 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<simStuff2, eval=TRUE,tidy=TRUE>>=
-as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
-as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
+as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
 @
@@
 \end{itemize}
-<<include=FALSE, tidy=TRUE>>=
-sim_2 = runif(n=7,streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type="double")
-as.vector(sim_2)
-@
 Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.

### 2021-10-12 pkg/gpuRandom/src/crossprodBatch.cpp

@@ -1,5 +1,5 @@
 #include "gpuRandom.hpp"
-//#define DEBUG
+#define DEBUG
 //#define NOKERNELS
 
 // C = A^T A or A^T D A or A^T D^(-1) A 
@@ -391,7 +391,7 @@ void crossprodBatch(
     Rcpp::IntegerVector Dstartend,  
     Rcpp::IntegerVector Nglobal,
     Rcpp::IntegerVector Nlocal,
-    const int NlocalCache, //int NrowStartC,
+    const int NlocalCache, 
     const int ctx_id) {
   
   const int Ncol = Astartend[3];
@@ -469,7 +469,7 @@ SEXP crossprodBatchTyped(
     Rcpp::IntegerVector Dstartend,  
     Rcpp::IntegerVector Nglobal,
     Rcpp::IntegerVector Nlocal, 
-    const int NlocalCache    //, int NrowStartC
+    const int NlocalCache    
 ) {
   /*
    std::vector<int> Nglobal = Rcpp::as<std::vector<int> >(NglobalR);
@@ -508,7 +508,7 @@ SEXP crossprodBatchBackend(
     Rcpp::IntegerVector Dstartend, 
     Rcpp::IntegerVector Nglobal,
     Rcpp::IntegerVector Nlocal, 
-    const int NlocalCache  //, int NrowStartC
+    const int NlocalCache  
 ) {
   
   SEXP result;

### 2021-10-12 pkg/gpuRandom/src/likfitGpu_P.cpp

@@ -490,7 +490,7 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
     0,//const int NstartD,  
     ssqYX.internal_size2()*yx.size2(), //const int NpadBetweenMatricesC,
     LinvYX.internal_size2()*Nobs, //const int NpadBetweenMatricesA,
-    NlocalCache[0]/2, // NlocalCacheA, 
+    NlocalCache[0], // NlocalCacheA, 
     localSize// Nlocal// cache a Nlocal[0] by Nlocal[1] submatrix of C
   );
   
@@ -721,7 +721,7 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
   ///////////////////////////Loop starts !!!//////////////////////////////////////////////////////////////////////////
   
   viennacl::ocl::command_queue theQueue = maternKernel.context().get_queue();
-  
+#ifdef UNDEF 
   if(verbose[1]==1){  // given betas
   for (Diter=0,DiterIndex=0; 
        Diter< Niter; 
@@ -736,24 +736,24 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
       Diter << " endThisIteration " <<  endThisIteration <<
         " NthisIteration " << NthisIteration << " localMemorySize " << localMemory.size() << " DiterIndex " << DiterIndex <<"\n";
       }
-
+    // matern
     viennacl::ocl::enqueue(maternKernel(Vbatch, coords, params, 
                                         localMemory,
                                         DiterIndex, NthisIteration),
                            theQueue);
 
-    //#Vbatch=LDL^T, cholesky decomposition
+    // cholesky Vbatch=LDL^T, cholesky decomposition
     viennacl::ocl::enqueue(cholKernel(Vbatch, cholDiagMat, 
                                       localMemory, 
                                       NthisIteration, detVar, DiterIndex),
                            theQueue);
 
-    // LinvYX = L^(-1) YX,   Nobs by (Ndatasets + Ncovariates)
+    // backsolve  LinvYX = L^(-1) YX,   Nobs by (Ndatasets + Ncovariates)
     viennacl::ocl::enqueue(backsolveKernel(LinvYX, Vbatch, yx, NthisIteration),
                            theQueue);
     
 
[…]

### 2021-10-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
debugging
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2903 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.143   0.004   0.147
+##   0.087   0.009   0.096
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   6.678   0.850   7.522
+##   6.695   0.819   7.511
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.764   0.000   0.764
+##   0.819   0.000   0.819
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4037993
+## [1] 0.4042226
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  14.747   0.004  14.743
+##  14.841   0.027  14.860
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.343   0.012   4.352
+##   4.638   0.001   4.637
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001263657
+## [1] 0.0001221702
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  90.524   0.155  90.629
+##  90.222   0.098  90.305
 \end{verbatim}

### 2021-10-14 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2911 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{A summary of the results}
+(Results obtained in first week of October)
 <<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE>>=

### 2021-10-14 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2911 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{verbatim}
-as.vector(gpuRandom::runif(n = 4, streams = myStreamsGpu1, Nglobal = c(1,
+as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu1, Nglobal = c(1,
     nrow(myStreamsGpu1)), type = "double"))
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392
+##  [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
+##  [7] 0.6487742 0.1112075 0.3661944 0.5018562
 as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
-    nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
-## [1] 0.6487742 0.1112075 0.3661944 0.5018562
+    nrow(myStreamsGpu1)), type = "double"))
+##  [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
+##  [7] 0.6487742 0.1112075 0.3661944 0.5018562
 \end{verbatim}
@@
 \end{itemize}
-
 Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
@@
 ##                    [,1]       [,2]       [,3]       [,4]
-## current.g1.1  240667857  559530223  502033783  739421137
-## current.g1.2  240667857 1309565828 1322587635 1475938232
-## current.g1.3      12345  336690377 1964121530  730262207
-## current.g2.1 1069151070   61444481 1949818481 1630192198
-## current.g2.2  809054265  197003928 1607232546  324551134
-## current.g2.3      12345   85196284 1462898381  795289868
+## current.g1.1 1918428443 1912157188  502033783  739421137
+## current.g1.2 1858462085  286315187 1322587635 1475938232
+## current.g1.3  878672095 2113333390 1964121530  730262207
+## current.g2.1 1132031887  834429287 1949818481 1630192198
+## current.g2.2  465230163   47498872 1607232546  324551134
+## current.g2.3  642281259 1335994581 1462898381  795289868
 ## initial.g1.1      12345  336690377  502033783  739421137
@@
 ##    user  system elapsed 
-##   0.087   0.009   0.096
+##   0.122   0.009   0.130
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   6.695   0.819   7.511
+##   6.592   0.867   7.529
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.819   0.000   0.819
+##   0.746   0.005   0.750
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4042226
+## [1] 0.4037127
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  14.841   0.027  14.860
+##  14.913   0.030  14.980
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.638   0.001   4.637
+##   4.570   0.013   4.579
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001221702
+## [1] 0.0001141787
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  90.222   0.098  90.305
+##  91.084   0.186  91.258
 \end{verbatim}
@@
 \subsection{A summary of the results}
+(Results obtained in first week of October)
 \begin{knitrout}

### 2021-10-17 pkg/gpuRandom/src/likfitGpu_P.cpp
crossprod local mem
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2915 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -867,7 +867,6 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
                                              NthisIteration),
                              theQueue);
       
-#ifdef UNDEF      
       // ssqYX = YX^Y L^(-1)T D^(-1) L^(-1) YX  square matrix, (Ndatasets + Ncovariates)
       viennacl::ocl::enqueue(crossprodKernel(ssqYX, LinvYX, cholDiagMat, 0, NthisIteration),
                              theQueue);
@@ -887,6 +886,7 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
       viennacl::ocl::enqueue(extractBlockKernel(XVYXVX, ssqYX, DiterIndex, NthisIteration),
                              theQueue);  
       
+#ifdef UNDEF      
       
       
       // cholesky X^T V^(-1) X = QPQ^T, save determinant as detReml, changes Ncovariates by Ncovariates part

### 2021-10-17 pkg/gpuRandom/src/crossprodBatch.cpp
crossprod local mem
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2915 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -49,7 +49,7 @@ std::string crossprodBatchString(
     "#define NstartC " + std::to_string(NstartC) + "\n"   
     "#define NstartA " + std::to_string(NstartA) + "\n" 
     "#define NstartD " + std::to_string(NstartD) + "\n" 
-    "#define NpadLocal " + std::to_string(Nlocal[1]) + "\n"    
+//    "#define NpadLocal " + std::to_string(Nlocal[1]) + "\n"    
     "#define NpadBetweenMatricesC " + std::to_string(NpadBetweenMatricesC) + "\n"    
     "#define NpadBetweenMatricesA " + std::to_string(NpadBetweenMatricesA) + "\n"    
     "#define NrowStop " + std::to_string(NrowStop) + "\n"    
@@ -63,19 +63,20 @@ std::string crossprodBatchString(
   if (NpadD >0 ) {
     result +=  "const global " + typeString + " *D,\n";
   }
-  
-  result += "int NrowStartC,\n"
+  result +=
+    " __local " + typeString + " *localCache,\n"
+ "int NrowStartC,\n"
   "int Nmatrix) {\n\n"
   
-  "local " + typeString + " Acache[" + 
+  "local " + typeString + " *Acache=localCache;\n" 
+  "local " + typeString + " *Dcache = &localCache[" + 
     std::to_string(NlocalCacheA) + "];\n" 
-  "local " + typeString + " Dcache[" + 
-    std::to_string(NlocalCacheA) + "];\n" 
-  "local " + typeString + " Ccache[" + 
-    std::to_string(Nlocal[0] * Nlocal[1]) + "];\n" +
+  "local " + typeString + " *Ccache = &localCache[" + 
+    std::to_string(2*NlocalCacheA) + "];\n" +
     
     typeString + " Cout, Ctemp;\n"
   "event_t ev;\n"
+  "const int NpadLocal = get_local_size(1);"
   "int AHere, CHere;\n"
   "int Dmatrix, Drow, Dcol, DrowNpadC, Dinner, DinnerA, DinnerAcol, DrowBlock, DinnerBlock, DcolBlock, DmatrixBlock, DcolInBounds, DmatrixInBounds;\n"
   "int A0Dcol, A0Drow;// location of elements A[0,Dcol] and A[0,Drow]\n";
@@ -445,8 +446,8 @@ void crossprodBatch(
     Rcpp::IntegerVector Astartend,
[…]

### 2021-10-18 pkg/gpuRandom/src/backsolve.cpp
no important changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2918 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -8,18 +8,18 @@
  
  
  
- - Dgroup[0] is matrix Dlocal[0] is row of A and C
- - Dgroup[1] is column of B, Dlocal[1] is Dinner
- - each Dgroup[1] does NcolsGroup columns of B
- - NcolsPerGroup = ceil( Ncol/Ngroup[1])
+ - Dglobal[0] is matrix, Dlocal[0] is row of A and C
+ - Dglobal[1] is column of B, Dlocal[1] is Dinner
+ - each Dglobal[1] does NcolsGroup columns of B
+ - NcolsPerGroup = ceil( Ncol/Nglobal[1])
  - local cache Nlocal[1] submatrices of size Nlocal[0] by NcolsPerGroup of C
  - don't cache A ?
  
- - loop Dmatrix = Dgroup[0]
+ - loop Dmatrix = Dglobal[0]
  - loop Drow = Dlocal[0]
  - loop Dinner = Dlocal[1]
  - cache A ?
- - loop DcolB = Dgroup[1]
+ - loop DcolB = Dglobal[1]
  - compute part of Linv b
  - end Dinner, DcolB, 
  - sum Linv b over Dlocal[1] in cache
@@ -275,7 +275,7 @@ result +=
   
   result +=
     "        // create C in cache and copy to C\n"
-    "        if((Dcol < Ncol) & DrowInBounds & localIsFirstCol){\n"
+    "        if( DrowInBounds & localIsFirstCol){\n"
     "        if((get_local_id(0) == Dinner)){\n"   
 
     "          cacheSum[NpadBetweenMatricesSum*DcolCache + DinnerC] = (B[BHereRow + Dcol] -\n"
@@ -294,16 +294,16 @@ result +=
 
   result +=  
     "        } //if(get_local_id(0) == Dinner)\n"
-    "        }// if(Dcol < Ncol) & DrowInBounds & localIsFirstCol\n"
+    "        }// if DrowInBounds & localIsFirstCol\n"
     "        barrier(CLK_LOCAL_MEM_FENCE);\n";  
   result +=
     "        // update A[Drow, ] * B[, Dcol] \n"
-    "        if((Dcol < Ncol) & DrowInBounds & localIsFirstCol){\n"    
+    "        if( DrowInBounds & localIsFirstCol){\n"    
     "        if((get_local_id(0) > Dinner)){\n"
     "          cacheSum[NpadBetweenMatricesSum*DcolCache + DlocalCache] +=\n"
[…]

### 2021-10-18 pkg/gpuRandom/R/RcppExports.R
no important changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2918 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -16,15 +16,6 @@ cholBatchBackend <- function(A, D, Astartend, Dstartend, numbatchD, Nglobal, Nlo
 #' Multiply crossproduct matrices
 #' 
 #' Computes C = t(A) D A
-#'
-#' @param C output matrices, stacked row-wise
-#' @param A rectangular matrices
-#' @param D rectangular matrix, columns are diagonals
-#' @param invertD set to 1 for C = t(A) D^(-1) A
-#' @param Nglobal vector of number of global work items
-#' @param Nlocal vector of number of local work items
-#' @param NlocalCache elements in local cache
-#' @export
 crossprodBatchBackend <- function(C, A, D, invertD, Cstartend, Astartend, Dstartend, Nglobal, Nlocal, NlocalCache) {
     .Call('_gpuRandom_crossprodBatchBackend', PACKAGE = 'gpuRandom', C, A, D, invertD, Cstartend, Astartend, Dstartend, Nglobal, Nlocal, NlocalCache)
 }

### 2021-10-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2920 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\title{Random number generation and applications on GPU’s in R?}
+\title{Random number generation and applications on GPU’s in R(?)}
 \Plaintitle{}
@@
 
-In recent years, parallel computing with R \citep{r2021} has become a very important topic and attracted lots of interest from researchers \citep[see][for a review]{eddelbuettel2021parallel}.
+In recent years, parallel computing with \proglang{R} \citep{r2021} has become a very important topic and attracted lots of interest from researchers \citep[see][for a review]{eddelbuettel2021parallel}.
 Although \proglang{R} is one of the most popular statistical software with many advantages, it has drawbacks in memory usage and computation mode aspects \citep{zhao_2016}.
@@
 Parallel computing is the solution to these drawbacks of \proglang{R}, for an overview of current parallel computing approaches with \proglang{R}, see CRAN Task View by \citet{cran2021} at \url{https://cran.r-project.org/web/views/HighPerformanceComputing.html}.
-One major way of parallel computing with \proglang{R} is using Graphics Processing Units (GPUs).
+
+An important way of parallel computing with \proglang{R} is using Graphics Processing Units (GPUs).
 GPUs can perform thousands of tests simultaneously, which makes them powerful at doing massive parallel computing and they are relatively cheap compared to multicore CPU's.
@@
 \pkg{gpuR} \citep{gpur1} is the only \proglang{R} package with a convenient interface between \proglang{R} and GPUs and it is created for any \proglang{R} user with a GPU device.
-By utilizing the \pkg{ViennaCL} \citep*{rupp2016viennacl} library, it provides a bridge between \proglang{R} and non-proprietary GPUs through an \proglang{OpenCL} (Open Computing Language) backend, which when combined with \pkg{Rcpp} \citep{rcpp1} gives a building block for other \proglang{R} packages.
+By utilizing the \pkg{ViennaCL} \citep*{rupp2016viennacl} library, it provides a bridge between \proglang{R} and non-proprietary GPUs through the \proglang{OpenCL} (Open Computing Language) backend, which when combined with \pkg{Rcpp} \citep{rcpp1} gives a building block for other \proglang{R} packages.
 
-Random number generator (RNG) is critical in simulation-based statistical inference, there is not yet an \proglang{R} package that is able to generate random numbers on GPU.
-This is complicate to do on GPU because multiple streams will be needed for parallel generation of random numbers, and the stream states need to be able to be restored for reproducibility of simulation results.
-We introduce the \pkg{gpuRandom} \proglang{R} package that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
+Random number generation is critical in simulation-based statistical inference, machine learning and many other scientific fields.
+While most random number generators are sequential, the \proglang{R} packages \pkg{parallel} (not on cran, how to cite?) and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
+More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG), \pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented (?).
 
-The remaining sections are organized as follows. 
-In Section 2, we introduce streams and the use of streams in work-items on GPU device, and describe and check the functions for generating uniform and Normal random numbers.
-In Section 3, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we parallelise this Monte Carlo simulation task and implement it on GPU, then we provide two real data examples to demonstrate the function usage and test its R performance.
+There is not yet an \proglang{R} package that is able to generate random numbers on GPU, this is complicated to do because multiple streams will be needed for parallel generation of random numbers, and the stream states need to be able to be restored for reproducibility of results, which is essential in many simulation applications.
+We introduce the \proglang{R} package \pkg{gpuRandom} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
+
+The remaining sections are organized as follows:
+In Section 2, we introduce streams and the use of streams in work-items on GPU device, and describe and check the functions for generating uniform and normal random numbers.
+In Section 3, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we parallelise this Monte Carlo simulation and implement it on GPU, then we provide two real data examples to demonstrate the function usage and test its R performance.
 Section 4 simulates simultaneously multiple Gaussian random surfaces by applying GPU-generated Normal random numbers and GPU-computed batches of Mat\'ern covariance matrices, then we plot the generated random surfaces in the end.
@@
 \section{Generating random numbers} \label{}
-Random numbers are the basis in statistical simulations methods, and are critical in many other fields.
-\cite{l2012random} summarized the usual two steps to generate a random variable %from a specified distribution
-in computational statistics: (1) generating independent and identically distributed (i.i.d.) uniform random variables on the interval $(0, 1)$, (2) applying transformations to these i.i.d.
+Uniform random number generators (RNGs) are essential in simulating random numbers from all types of probability distributions.
+\cite{l2012random} summarized the usual two steps to generate a random variable in computational statistics: (1) generating independent and identically distributed (i.i.d.) uniform random variables on the interval $(0, 1)$, (2) applying transformations to these i.i.d.
 $U(0, 1)$ random variables to get samples from the desired distribution.
-\cite{l2012random} and \cite{robert2004random} present several general transformation methods for generating non-uniform random variables, for example, the most frequently used inverse transform method for generating a random variable $X$ from some arbitrary distribution function $F$; The Box-Muller algorithm \citep{box1958note} for Gaussian random variable generation, and so on, which are all built on uniform random variables.
-Therefore, uniform random number generators (RNGs) are essential in simulating random numbers from all types of probability distributions.
-In the rest paper, ``RNG'' refers to uniform pseudo random number generator.
+\cite{l2012random} and \cite{robert2004random} present many general transformation methods for generating non-uniform random variables, for example, the most frequently used inverse transform method, the Box-Muller algorithm \citep{box1958note} for Gaussian random variable generation, and so on, which are all built on uniform random variables.
+%In the rest paper, ``RNG'' refers to uniform pseudo random number generator.
 
-\proglang{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators, these RNGs are actually stream objects of different forms.
-\pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG} library, making it able to generate random numbers on GPUs.
-In what follows, we will illustrate how to create stream objects and how to use streams to generate (standard) Uniform and (standard) Gaussian random numbers.
+\pkg{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators.
+\pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG}, making it able to generate random numbers on GPUs.
+In what follows, we will illustrate how to create streams and how to use streams to generate (standard) uniform and (standard) Gaussian random numbers.
 
 \subsection{Creating streams}\label{create_streams}
-In \proglang{clRNG} library, a stream object is a structure that contains three states: the current state of the stream, the initial state (or seed) of the stream, and the initial state of the current substream.
-The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
-If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item picks streams from the global memory and copies their current states only to its private memory.
-\pkg{gpuRandom} is able to create streams both on the host computer and on the GPU device.
-\fct{createStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on CPU as \proglang{R} matrices (see the \proglang{R} output below), in which the substreams are not created because they are not actually used in random number generation.
-\fct{createStreamsGpu} adapted the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on GPU device, so that it can avoid unnecessary data transfer between host and device.
-From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
+The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
+Creating a new stream amounts to computing its starting point (seed).
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object}, which are occasionally useful.
+
+In general, a stream object contains three states: the current state of the stream， the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+Whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
+
+In \proglang{clRNG} library, the initial state of the first stream has a default value (12345).
+For the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+Each stream is further partitioned into substreams of length $W = 2^{72}$.
+The state (and seed) of each stream is a vector of six 31-bit integers.
+This size of state is appropriate for having streams running in work items on GPU cards, for example, while providing a sufficient period length for most applications.
+
+The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG.
+If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
+\pkg{gpuRandom} is able to create streams both on the host and on the GPU device.
+\fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices (see the \proglang{R} output below), creating streams on host makes it easy for us to view and check the streams behavior.
+\fct{createStreamsGpu} is adapted from the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on a GPU device.
+Substreams are not created because currently we use one stream per work-item.
+%(From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
+delete?)
 In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively. 
 <<createStreams_CPU>>=
-# creating streams on cpu
+# creating streams on CPU
+library("gpuR")
+library("gpuRandom")
 myStreamsCpu <- createStreamsCpu(n=4, seed=12345)
@@
   \item \code{n}: number of streams to create.
-  \item \code{seed}: random seed of streams of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
+  \item \code{seed}: vector of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
 \end{itemize}
-For the MRG31k3p RNG, each state of the stream is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''.
 We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams directly on GPU using \fct{createStreamsGpu}.
@@
 <<createStreams_Gpu>>=
-# creating streams on gpu
+# creating streams on GPU
 myStreamsGpu1 = vclMatrix(myStreamsCpu)
@@
 By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we are able to produce exactly the same sequence of streams as many times as we want.
+Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.
 
 \subsection{Using streams to generate  i.i.d. $U (0, 1)$ random numbers on GPU}
-The streams created on GPU are then used in work-items that executes in parallel on GPU device.
-In \pkg{gpuRandom} each work-item takes one distinct stream to generate random numbers, it is important to notice that the number of streams should always be equal to the number of total work-items in use.
+The streams created sequentially on a GPU are then used in work-items that executes in parallel on a GPU device.
+In \pkg{gpuRandom} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be equal to the number of total work-items in use.
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
 Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
-Here the pointers to \code{streams} and to the matrix \code{out} on the global memory are passed to the kernel as arguments.
-\code{Nrow} and \code{Ncol} represent number of rows and columns of the matrix \code{out} respectively.
-\code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is internal number of columns of the \code{streams} matrix (these variables are defined in macros not shown here).
+Here the pointers to \code{streams} and to the output matrix \code{out} on the global memory are passed to the kernel as arguments.
+\code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively.
+\code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of the \code{streams} (these variables are defined in macros not shown here).
 Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}.
 \code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates a uniform random integer named \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
 At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
@@
 \end{framed}
-This kernel executes over a 2-dimensional index space.
+This kernel executes over a 2-dimensional NDRange.
 Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example.
@@
 Each stream object is passed to each work-item.
-The number in a cell of the matrix corresponds to the stream index, it indicates which stream does the random number generation for this matrix cell.
+The number in a cell of the matrix corresponds to the stream index, it indicates which stream generates a random number for this matrix cell.
 For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
@@
     \centering
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f1} }}%
+    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f1} }}%
     \qquad
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f2} }}%
+    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f2} }}%
     \caption{name?\label{fig:1}}%
@@
 
-Now we use the streams created in section \ref{create_streams} to generate two vectors of 6 i.i.d.
-$U (0,1)$ random numbers using the function \fct{runif}.
-If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied.
-The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
-By using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers.
+Now we use the streams created in section \ref{create_streams} to generate two vectors of double-precison i.i.d.
+$U(0,1)$ random numbers with the function \fct{runif}.
+To view the generated random numbers, we need to convert them to R vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
 <<simStuff2, eval=TRUE,tidy=TRUE>>=
-as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
-as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+as.vector(gpuRandom::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+
+as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
 @
@@
   \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
-  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
+  \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
 \end{itemize}
+\fct{runif} returns the same sequence of random numbers if it uses the same streams in the same current state (here \code{myStreamsGpu1} and \code{myStreamsGpu2}), not streams starting from the same states.
 Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
-Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
-<<eval=TRUE,tidy=TRUE>>=
+Because each time a random number is generated, the current state of the stream advances by one position.
+Below shows the \code{myStreamsGpu1} after it being used.
+<<eval=TRUE,tidy=TRUE, include=TRUE>>=
 t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
@@
 Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
-However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations.
+However, we can save the streams on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, in this way we can reproduce the exact same sequence of random numbers in simulations.
 Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
@@
 streams <- createStreamsGpu(n = 512)
-random_u<-runif(n=50000, streams=streams, Nglobal=c(64,8), type="double")
-random_u<-as.vector(random_u)
+
+random_u<-as.vector(runif(n=50000, streams=streams, Nglobal=c(64,8), type="double"))
 @
@@
 
-\subsection{Using streams to generate i.i.d. N (0, 1) random numbers on GPU}
-We apply the Box-Muller transformation to $U (0,1)$ random numbers to generate Gaussian random numbers.
+\subsection{Using streams to generate i.i.d. N(0, 1) random numbers on GPU}
+We apply the Box-Muller transformation to $U(0,1)$ random numbers to generate standard Gaussian random numbers.
 As shown in Algorithm \ref{algorithm1}, Box-Muller algorithm takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables.
-The derivation basically uses transformation from Cartesian coordinates to polar coordinates to represent two independent, Normally distributed random variables.
-The Box-Muller algorithm turns out to be the best choice for Gaussian transform on GPU compared to other transform methods \citep{howes2007efficient}, because this algorithm has no branching or looping, which are the things GPU is not good at, but GPU is good at the high computational load of sine and cosine functions in the algorithm.
+%The derivation basically uses transformation from Cartesian coordinates to polar coordinates to represent two independent, Normally distributed random variables.
+The Box-Muller algorithm turns out to be the best choice for Gaussian transform on GPU compared to other transform methods \citep{howes2007efficient}, because this algorithm has no branching or looping, which are the things GPU is not good at.
+%but GPU is good at the high computational load of sine and cosine functions in the algorithm.
 
@@
 Listing \ref{lst:normal} shows a fragment of the kernel that generates standard Gaussian random numbers.
-Aside from using local memory and the part that does the Box-Muller transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}.
-If developers want to use other Gaussian transform methods, this kernel is easy to be modified and incorporated in other \proglang{R} packages.
-The kernel executes over a 2-dimensional index space of $1 \times 2$ work-groups.
+Aside from using local memory and the part that does the transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}.
+If developers want to use other Gaussian transform methods, this kernel is easy to be adapted and incorporated in other \proglang{R} packages.
+The kernel executes over a 2-dimensional NDRange with $1 \times 2$ work-groups.
 \code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
@@
 We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange.
-Each stream is copied to the private memory of each work-item.
-Work-items are organized to be 1 by 2 work-groups as the random numbers are generated in pairs using the Box-Muller method.
+%Each stream is copied to the private memory of each work-item.
+Work-items are organized to be in 1 by 2 work-groups as the random numbers are generated in pairs using the Box-Muller method.
 The shading on the grids in the NDRange distinguishes work-groups.
@@
 Each work-group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where work-item of `\code{get_local_id(1)}$=0$' generates the \code{X} of a pair, and work-item of `\code{get_local_id(1)}$=1$' generates the \code{Y} of a pair.
-Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
-Therefore, each work-item generates one Gaussian random number in each iteration, the matrix filling process is the same as that shown in Figure \ref{fig:1}.
+Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of the matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
+%Therefore, each work-item generates one Gaussian random number in each iteration,
+The matrix filling process is the same as that shown in Figure \ref{fig:1}.
 
@@
     \centering
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f3} }}%
+    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f3} }}%
     \qquad
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/paper1_2021_5/f4} }}%
+    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
     \caption{name???}%
@@
 
-In the following code, We create a $1024 \times 512$ matrix named \code{normal_matrix} of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+In the following code, We create a $1024 \times 512$ matrix %named \code{normal_matrix} 
+of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we adapted the \proglang{R} function \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on GPU, which can speed up the Q-Q plot greatly.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
 <<histAndQQ_prepare,eval=TRUE,cache=TRUE>>=
@@
 
-We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the runtime between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
+We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
 Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
@@
 One application of GPU-generated random numbers in \pkg{gpuRandom} is Monte Carlo simulation for Fisher's exact test.
-Fisher’s exact test is usually applied for analyzing $2 \times 2$ contingency tables when one of the expected values in table is less than 5.
+Fisher’s exact test is applied for analyzing usually $2 \times 2$ contingency tables when one of the expected values in table is less than 5.
 Different from methods which rely on approximation, Fisher's exact test computes directly the probability of obtaining each possible combination of the data for the same row and column totals (marginal totals) as the observed table, and get the exact p-value by adding together all the probabilities of tables as extreme or more extreme than the one observed.
@@
 23]{mehta2011ibm} shows a $5 \times 6$ observed table that has 1.6 billion possible tables.
-Calculating the exact P-values may lead to very long running time and can sometimes exceed the memory limits of your computer.
+Calculating the exact P-values may lead to very long run-time and can sometimes exceed the memory limits of your computer.
 Hence, the option \code{simulate.p.value = TRUE} in \fct{stats::fisher.test} is provided, which enables computing p-values by Monte Carlo simulation for tables larger than 2 $\times$ 2.
-Given an observed table and a number of replicates $B$, the Monte Carlo simulation does the following steps:
+Given an observed table and a number of replicates $B$, the Monte Carlo simulation for Fisher's exact test does the following steps:
 \begin{enumerate}
   \item Calculate the test statistic for the observed table.
-	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic (minus log-factorial of table) from the random table.
-	\item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
+	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic from the random table.
+	\item Count the number of iterations that have test statistics less or equal to the one from the observed table (\emph{Counts}).
 	\item Estimate p-value using $\frac{1 + \emph{Counts}}{B + 1}$.
 \end{enumerate}
-The Monte Carlo simulation solves the computational problem of Fisher's exact test but can still take long running time when $B$ reaches millions in scale, thus we make the function \fct{gpuRandom::fisher.sim} that does the steps 1, 2 and 3 parallelly on GPU, and step 2 of saving the test statistics from the random tables as an option in \fct{gpuRandom::fisher.sim}, which largely saves processing time.
+The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table.
+This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}.
 
-We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the runtime with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
+\fct{stats::fisher.test} can take a long running time when $B$ reaches millions as the $B$ random tables are constructed sequentially, one in each iteration.
+\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration), and the steps 2 and 3
+are also done in parallel by work-items at each iteration on GPU. Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
+
+We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
@@
 <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
-data2<-read.csv("/home/ruoyong/paper1_2021_5/month.csv") 
+data2<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv") 
 data2<-data2[,-1]
@@
 library(latex2exp)
-data<-read.csv("/home/ruoyong/paper1_2021_5/weekday.csv") 
+data<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
 #exampletwo<-as.matrix(as.data.frame(lapply(data, as.numeric)))
@@
 <<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE>>=
-hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}))$'), breaks=40, ylab="Frequency", main="Month data")
+hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
 abline(v = observed_m, col = "blue", lwd = 1.5)
 
-hist(result_weekgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}))$'), ylab="Frequency", main="Week data")
+hist(result_weekgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
 abline(v = observed_w, col = "blue", lwd = 1.5)
@@
 A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})$ is multivariate Gaussian.
-One reason that Gaussian random fields play an important role is that the specification of their finite-dimensional distributions is simple \citep{abrahamsen1997review}, the expectation $\mu(x)$ and covariance function $\Sigma$ completely determine the distribution of $U$, where
+%One reason that Gaussian random fields play an important role is that the specification of their finite-dimensional distributions is simple \citep{abrahamsen1997review},
+The expectation $\mu(x)$ and covariance function $\Sigma$ completely determine the distribution of $U$, where 
 \begin{gather*} 
@@
 \end{gather*}
-There exist many choices for the covariance function $\Sigma$, one common choice is the Mat\'ern covariance function \citep{matern1960spatial}.
-The Mat\'ern covariance between $U(x)$ and $U(x')$ takes the form
+The common choice for the covariance function $\Sigma$ is the Mat\'ern covariance function \citep{matern1960spatial}. The Mat\'ern covariance between $U(x)$ and $U(x')$ takes the form 
 \begin{equation*}
@@
 $K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$, specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.
-(For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.)  There are several alternative parameterisations of Mat\'ern covariance functions appeared in literatures  \citep[see][]{haskard2007anisotropic}.
+%(For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.)
+There are several alternative parameterisations of Mat\'ern covariance functions in literatures  \citep[see][]{haskard2007anisotropic}.
 In \pkg{gpuRandom}, we use the above form for computing Mat\'ern covariance.
 
-Due to the importance of Gaussian random field, there are a large amount of studies devoted to Gaussian random field generation techniques.
+%Due to the importance of Gaussian random field, there are a large amount of studies devoted to Gaussian random field generation techniques. 
 \cite{LiuandLi2019} gives a comprehensive review on 7 popular methods for Gaussian random field generation, which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
@@
 \end{itemize}
-The matrix decomposition method is straightforward to implement, however, when a large number ($n$) of locations is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
-\pkg{gpuRandom} package takes advantage of GPU's parallel computing power to not only do the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
-The following shows an example, in which we simulate 10 Gaussian random fields of mat\'ern covariance with 5 sets of parameters at one time, and introduces several GPU functions used in the simulation process.
+The matrix decomposition method is straightforward to implement, however, when a large number of locations ($n$) is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$ \citep{LiuandLi2019},  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
+\pkg{gpuRandom} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
+The following shows an example, in which we simulate 10 Gaussian random fields of mat\'ern covariance with 5 sets of parameters at one time.
+%and introduces several GPU functions used in the simulation process.
 
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
-<<setupData, eval=TRUE,echo=FALSE, message=FALSE>>=
+<<setupData, eval=TRUE,echo=TRUE, message=FALSE>>=
 library("gpuR")
@@
 gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
-                     Nglobal= c(512, 8),
-                     Nlocal= c(128, 8),
-                     NlocalCache=2000)
+                     Nglobal= c(128, 8),
+                     Nlocal= c(32, 8),
+                     NlocalCache=1000)
 @
@@
 We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the Normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
-Most of these functions uses local memory on GPU, all of them uses parallel programming and avoids unnecessary data transfers between host and device.
+Most of these functions uses local memory on GPU %all of them uses parallel programming and avoids unnecessary data transfers between host and device.
 Comparison of performance between using \pkg{gpuRandom} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
@@
 The usual steps that we take to debug is commenting out some parts of codes and then recompile many times until we find the problem, or printing out flag messages in several places in the code to find out where the program stops at due to error.
+
+talk about using substreams in future?
+
 \pkg{gpuRandom} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.

### 2021-10-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2920 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-10-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
 \pkg{gpuR} \citep{gpur1} is the only \proglang{R} package with a convenient interface between \proglang{R} and GPUs and it is created for any \proglang{R} user with a GPU device.
-By utilizing the \pkg{ViennaCL} \citep*{rupp2016viennacl} library, it provides a bridge between \proglang{R} and non-proprietary GPUs through an \proglang{OpenCL} (Open Computing Language) backend, which when combined with \pkg{Rcpp} \citep{rcpp1} gives a building block for other \proglang{R} packages.
-
-Random number generator (RNG) is critical in simulation-based statistical inference, there is not yet an \proglang{R} package that is able to generate random numbers on GPU.
-This is complicate to do on GPU because multiple streams will be needed for parallel generation of random numbers, and the stream states need to be able to be restored for reproducibility of simulation results.
-We introduce the \pkg{gpuRandom} \proglang{R} package that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
-
-The remaining sections are organized as follows. 
-In Section 2, we introduce streams and the use of streams in work-items on GPU device, and describe and check the functions for generating uniform and Normal random numbers.
-In Section 3, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we parallelise this Monte Carlo simulation task and implement it on GPU, then we provide two real data examples to demonstrate the function usage and test its R performance.
+By utilizing the \pkg{ViennaCL} \citep*{rupp2016viennacl} library, it provides a bridge between \proglang{R} and non-proprietary GPUs through the \proglang{OpenCL} (Open Computing Language) backend, which when combined with \pkg{Rcpp} \citep{rcpp1} gives a building block for other \proglang{R} packages.
+
+Random number generation is critical in simulation-based statistical inference, machine learning and many other scientific fields.
+While most random number generators are sequential, the \proglang{R} packages \pkg{parallel} (not on cran, how to cite?) and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
+More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG), \pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented (?).
+
+There is not yet an \proglang{R} package that is able to generate random numbers on GPU, this is complicated to do because multiple streams will be needed for parallel generation of random numbers, and the stream states need to be able to be restored for reproducibility of results, which is essential in many simulation applications.
+We introduce the \proglang{R} package \pkg{gpuRandom} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
+
+The remaining sections are organized as follows:
+In Section 2, we introduce streams and the use of streams in work-items on GPU device, and describe and check the functions for generating uniform and normal random numbers.
+In Section 3, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we parallelise this Monte Carlo simulation and implement it on GPU, then we provide two real data examples to demonstrate the function usage and test its R performance.
 Section 4 simulates simultaneously multiple Gaussian random surfaces by applying GPU-generated Normal random numbers and GPU-computed batches of Mat\'ern covariance matrices, then we plot the generated random surfaces in the end.
@@
 \section{Generating random numbers} \label{}
-Random numbers are the basis in statistical simulations methods, and are critical in many other fields.
-\cite{l2012random} summarized the usual two steps to generate a random variable %from a specified distribution
-in computational statistics: (1) generating independent and identically distributed (i.i.d.) uniform random variables on the interval $(0, 1)$, (2) applying transformations to these i.i.d.
+Uniform random number generators (RNGs) are essential in simulating random numbers from all types of probability distributions.
+\cite{l2012random} summarized the usual two steps to generate a random variable in computational statistics: (1) generating independent and identically distributed (i.i.d.) uniform random variables on the interval $(0, 1)$, (2) applying transformations to these i.i.d.
 $U(0, 1)$ random variables to get samples from the desired distribution.
-\cite{l2012random} and \cite{robert2004random} present several general transformation methods for generating non-uniform random variables, for example, the most frequently used inverse transform method for generating a random variable $X$ from some arbitrary distribution function $F$; The Box-Muller algorithm \citep{box1958note} for Gaussian random variable generation, and so on, which are all built on uniform random variables.
-Therefore, uniform random number generators (RNGs) are essential in simulating random numbers from all types of probability distributions.
-In the rest paper, ``RNG'' refers to uniform pseudo random number generator.
-
-\proglang{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators, these RNGs are actually stream objects of different forms.
-\pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG} library, making it able to generate random numbers on GPUs.
-In what follows, we will illustrate how to create stream objects and how to use streams to generate (standard) Uniform and (standard) Gaussian random numbers.
+\cite{l2012random} and \cite{robert2004random} present many general transformation methods for generating non-uniform random variables, for example, the most frequently used inverse transform method, the Box-Muller algorithm \citep{box1958note} for Gaussian random variable generation, and so on, which are all built on uniform random variables.
+%In the rest paper, ``RNG'' refers to uniform pseudo random number generator.
+
+\pkg{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators, these RNGs are actually stream objects of different forms.
+\pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG}, making it able to generate random numbers on GPUs.
+In what follows, we will illustrate how to create streams and how to use streams to generate (standard) uniform and (standard) Gaussian random numbers by \pkg{gpuRandom}.
 
 \subsection{Creating streams}\label{create_streams}
-In \proglang{clRNG} library, a stream object is a structure that contains three states: the current state of the stream, the initial state (or seed) of the stream, and the initial state of the current substream.
-The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG.
-If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item picks streams from the global memory and copies their current states only to its private memory.
-\pkg{gpuRandom} is able to create streams both on the host computer and on the GPU device.
-\fct{createStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on CPU as \proglang{R} matrices (see the \proglang{R} output below), in which the substreams are not created because they are not actually used in random number generation.
-\fct{createStreamsGpu} adapted the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on GPU device, so that it can avoid unnecessary data transfer between host and device.
-From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
+The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
+Creating a new stream amounts to computing its starting point (seed).
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object}, which are occasionally useful.
+
+In general, a stream object contains three states: the initial state of the stream (or seed), the initial state of the current substream (by default it is equal to the seed), and the current state of the stream.
+Whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
+
+In \proglang{clRNG} library, the initial state of the first stream has a default value (12345).
+For the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+Each stream is further partitioned into substreams of length $W = 2^{72}$.
+The state (and seed) of each stream is a vector of six 31-bit integers.
+This size of state is appropriate for having streams running in work items on GPU cards, for example, while providing a sufficient period length for most applications.
+
+The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG.
+If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
+\pkg{gpuRandom} is able to create streams both on the host and on the GPU device.
+\fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices (see the \proglang{R} output below), creating streams on host makes it easy for us to view and check the streams behavior.
+\fct{createStreamsGpu} is adapted from the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on a GPU device.
+Substreams are not created because currently we use one stream per work-item.
+%(From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
+delete?)
 In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively. 
@@
 \begin{verbatim}
-# creating streams on cpu
+# creating streams on CPU
+library("gpuR")
+library("gpuRandom")
 myStreamsCpu <- createStreamsCpu(n=4, seed=12345)
@@
 \begin{verbatim}
-# creating streams on gpu
+# creating streams on GPU
 myStreamsGpu1 = vclMatrix(myStreamsCpu)
@@
 \subsection{Using streams to generate  i.i.d. $U (0, 1)$ random numbers on GPU}
-The streams created on GPU are then used in work-items that executes in parallel on GPU device.
-In \pkg{gpuRandom} each work-item takes one distinct stream to generate random numbers, it is important to notice that the number of streams should always be equal to the number of total work-items in use.
+The streams created sequentially on a GPU are then used in work-items that executes in parallel on a GPU device.
+In \pkg{gpuRandom} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be equal to the number of total work-items in use.
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
 Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
-Here the pointers to \code{streams} and to the matrix \code{out} on the global memory are passed to the kernel as arguments.
-\code{Nrow} and \code{Ncol} represent number of rows and columns of the matrix \code{out} respectively.
-\code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is internal number of columns of the \code{streams} matrix (these variables are defined in macros not shown here).
+Here the pointers to \code{streams} and to the output matrix \code{out} on the global memory are passed to the kernel as arguments.
+\code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively.
+\code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of the \code{streams} (these variables are defined in macros not shown here).
 Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}.
 \code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates a uniform random integer named \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
 At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
@@
 
-Now we use the streams created in section \ref{create_streams} to generate two vectors of 6 i.i.d.
-$U (0,1)$ random numbers using the function \fct{runif}.
-If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied.
-The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
-By using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-as.vector(gpuRandom::runif(n = 4, streams = myStreamsGpu1, Nglobal = c(1,
+Now we use the streams created in section \ref{create_streams} to generate two vectors of double-precison i.i.d.
+$U(0,1)$ random numbers with the function \fct{runif}.
+To view the generated random numbers, we need to convert them to R vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+as.vector(gpuRandom::runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1,
     nrow(myStreamsGpu1)), type = "double"))
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392
+## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
 as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
@@
   \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
-  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
+  \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
 \end{itemize}
-
+\fct{runif} returns the same sequence of random numbers if it uses the same streams in the same current state (here \code{myStreamsGpu1} and \code{myStreamsGpu2}), not streams starting from the same states.
 Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
-Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
+Because each time a random number is generated, the current state of the stream advances by one position.
+Below shows the \code{myStreamsGpu1} after it being used.
 \begin{knitrout}
@@
 ##                    [,1]       [,2]       [,3]       [,4]
-## current.g1.1  240667857  559530223  502033783  739421137
-## current.g1.2  240667857 1309565828 1322587635 1475938232
-## current.g1.3      12345  336690377 1964121530  730262207
-## current.g2.1 1069151070   61444481 1949818481 1630192198
-## current.g2.2  809054265  197003928 1607232546  324551134
-## current.g2.3      12345   85196284 1462898381  795289868
+## current.g1.1  878672095 2113333390  502033783  739421137
+## current.g1.2  240667857  559530223 1322587635 1475938232
+## current.g1.3  240667857 1309565828 1964121530  730262207
+## current.g2.1  642281259 1335994581 1949818481 1630192198
+## current.g2.2 1069151070   61444481 1607232546  324551134
+## current.g2.3  809054265  197003928 1462898381  795289868
 ## initial.g1.1      12345  336690377  502033783  739421137
@@
 streams <- createStreamsGpu(n = 512)
-random_u<-runif(n=50000, streams=streams, Nglobal=c(64,8), type="double")
-random_u<-as.vector(random_u)
+
+random_u<-as.vector(runif(n=50000, streams=streams, Nglobal=c(64,8), type="double"))
 \end{verbatim}
@@
 
-In the following code, We create a $1024 \times 512$ matrix named \code{normal_matrix} of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+In the following code, We create a $1024 \times 512$ matrix %named \code{normal_matrix} 
+of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we adapted the \proglang{R} function \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on GPU, which can speed up the Q-Q plot greatly.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
 \begin{knitrout}
@@
 ##    user  system elapsed 
-##   0.087   0.009   0.096
+##   0.100   0.004   0.459
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   6.695   0.819   7.511
+##   4.640   0.304   4.939
 \end{verbatim}
@@
 23]{mehta2011ibm} shows a $5 \times 6$ observed table that has 1.6 billion possible tables.
-Calculating the exact P-values may lead to very long running time and can sometimes exceed the memory limits of your computer.
+Calculating the exact P-values may lead to very long run-time and can sometimes exceed the memory limits of your computer.
 Hence, the option \code{simulate.p.value = TRUE} in \fct{stats::fisher.test} is provided, which enables computing p-values by Monte Carlo simulation for tables larger than 2 $\times$ 2.
-Given an observed table and a number of replicates $B$, the Monte Carlo simulation does the following steps:
+Given an observed table and a number of replicates $B$, the Monte Carlo simulation for Fisher's exact test does the following steps:
 \begin{enumerate}
   \item Calculate the test statistic for the observed table.
-	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic (minus log-factorial of table) from the random table.
-	\item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
+	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic from the random table.
+	\item Count the number of iterations that have test statistics less or equal to the one from the observed table (\emph{Counts}).
 	\item Estimate p-value using $\frac{1 + \emph{Counts}}{B + 1}$.
 \end{enumerate}
-The Monte Carlo simulation solves the computational problem of Fisher's exact test but can still take long running time when $B$ reaches millions in scale, thus we make the function \fct{gpuRandom::fisher.sim} that does the steps 1, 2 and 3 parallelly on GPU, and step 2 of saving the test statistics from the random tables as an option in \fct{gpuRandom::fisher.sim}, which largely saves processing time.
-
-We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the runtime with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
+The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table.
+This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}.
+
+\fct{stats::fisher.test} can take a long running time when $B$ reaches millions as the $B$ random tables are constructed sequentially, one in each iteration.
+\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration), and the steps 2 and 3
+are also done in parallel by work-items at each iteration on GPU. Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
+
+We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
@@
 ##    user  system elapsed 
-##   0.819   0.000   0.819
+##   0.693   0.008   2.317
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4042226
+## [1] 0.4039421
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  14.841   0.027  14.860
+##  10.248   0.000  10.242
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.638   0.001   4.637
+##   0.589   0.025  10.759
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001221702
+## [1] 0.0001263657
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  90.222   0.098  90.305
+##  62.792   0.036  62.770
 \end{verbatim}
@@
 \end{itemize}
-The matrix decomposition method is straightforward to implement, however, when a large number ($n$) of locations is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
-\pkg{gpuRandom} package takes advantage of GPU's parallel computing power to not only do the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
-The following shows an example, in which we simulate 10 Gaussian random fields of mat\'ern covariance with 5 sets of parameters at one time, and introduces several GPU functions used in the simulation process.
+The matrix decomposition method is straightforward to implement, however, when a large number of locations ($n$) is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$ \citep{LiuandLi2019},  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
+\pkg{gpuRandom} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
+The following shows an example, in which we simulate 10 Gaussian random fields of mat\'ern covariance with 5 sets of parameters at one time.
+%and introduces several GPU functions used in the simulation process.
 
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+library("gpuR")
+library("gpuRandom")
+library('geostatsp')
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
@@
 gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
-                     Nglobal= c(512, 8),
-                     Nlocal= c(128, 8),
-                     NlocalCache=2000)
+                     Nglobal= c(128, 8),
+                     Nlocal= c(32, 8),
+                     NlocalCache=1000)
 \end{verbatim}
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in vectorToVCL(data, 4L, context\_index - 1): ViennaCL: FATAL ERROR: CL\_OUT\_OF\_RESOURCES \\\#\# \ ViennaCL tried to launch a compute kernel, but the device does not provide enough resources.
-Try changing the global and local work item sizes.\\\#\# If you think that this is a bug in ViennaCL, please report it at viennacl-support@lists.sourceforge.net and supply at least the following information:\\\#\# \ * Operating System\\\#\# \ * Which OpenCL implementation (AMD, NVIDIA, etc.)\\\#\# \ * ViennaCL version\\\#\# Many thanks in advance!}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streamsGpu' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
                    type = theType)
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'zmatGpu' not found}}\begin{verbatim}
+
 gpuRandom::multiplyLowerDiagonalBatch(
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuRandom:::multiplyLowerDiagonalBatchBackend(output, L, D, B, : object 'simMat' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 simRaster = brick(myRaster, nl = ncol(simMat)*nrow(paramsGpu))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\begin{verbatim}
 values(simRaster) = as.vector(as.matrix(simMat))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'as.matrix': object 'simMat' not found}}\begin{verbatim}
 names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'simRaster' not found}}\end{kframe}
+\end{kframe}
+\includegraphics[width=0.45\textwidth]{figure/maternplot-1} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-2} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-3} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-4} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-5} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-6} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-7} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-8} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-9} 
+\includegraphics[width=0.45\textwidth]{figure/maternplot-10} 
 \end{knitrout}

### 2021-10-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
global same with local
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2921 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-In general, a stream object contains three states: the current state of the stream， the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.

### 2021-10-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
global same with local
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2921 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\pkg{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators, these RNGs are actually stream objects of different forms.
+\pkg{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators.
 \pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG}, making it able to generate random numbers on GPUs.
-In what follows, we will illustrate how to create streams and how to use streams to generate (standard) uniform and (standard) Gaussian random numbers by \pkg{gpuRandom}.
+In what follows, we will illustrate how to create streams and how to use streams to generate (standard) uniform and (standard) Gaussian random numbers.
 
@@
 
-In general, a stream object contains three states: the initial state of the stream (or seed), the initial state of the current substream (by default it is equal to the seed), and the current state of the stream.
+In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
@@
 By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we are able to produce exactly the same sequence of streams as many times as we want.
+Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.
 
@@
 ##    user  system elapsed 
-##   0.693   0.008   2.317
+##   0.709   0.004   2.326
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.589   0.025  10.759
+##   0.607   0.012  10.764
 \end{verbatim}

### 2021-10-26 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
chol fixed ithink
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2930 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.100   0.004   0.459
+##   0.144   0.004   0.189
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   4.640   0.304   4.939
+##   6.762   0.868   7.624
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.709   0.004   2.326
+##   0.720   0.001   0.721
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4039421
+## [1] 0.4035335
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  10.248   0.000  10.242
+##  14.706   0.004  14.730
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.607   0.012  10.764
+##   4.585   0.001   4.582
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001263657
+## [1] 0.0001292627
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  62.792   0.036  62.770
+##  90.434   0.196  90.561
 \end{verbatim}
@@
                      NlocalCache=1000)
+## 
+## #pragma OPENCL EXTENSION cl_khr_fp64 : enable
+## 
+## 
+## #define N 4800  //dimension of matrix
+## #define colStart 0  //column to start at
+## #define colEnd 4800
+## //internal number of columns
+## #define Npad 4864
+## //internal columns for matrix holding diagonals
+## #define NpadDiag 4864
+## #define NstartA 0
+## #define NstartD 0
+## //elements in internal cache
+## #define Ncache 1000
+## //extra rows between stacked matrices
+## #define NpadBetweenMatrices 23347200
+## #define maxLocalItems 256
+## 
+## __kernel void cholBatch(
+## 	__global double *A,
+## 	__global double *diag,
+##  __local double*diagLocal,
+##               int Nmatrix
+## ){
+##  const int localIndex = get_local_id(0)*get_local_size(1) + get_local_id(1);
+##  const int localIndexIsZero = (localIndex==0);
+##  const int NlocalTotal = get_local_size(0)*get_local_size(1);
+##  local double toAddLocal[maxLocalItems];
+##  int Dcol, DcolNpad;
+##  int Drow, DrowBlock, Dk, Dmatrix, DmatrixBlock;
+## int minDcolNcache;
+##  double DL;
+##  local double diagDcol;
+##  int AHere, AHereDcol, AHereDrow, diagHere;
+##  barrier(CLK_LOCAL_MEM_FENCE);
+## 
+## for(Dmatrix = get_group_id(0);
+##      Dmatrix < Nmatrix;
+##      Dmatrix+= get_num_groups(0)){
+## 
+##  diagHere = Dmatrix*NpadDiag + NstartD;
+##  AHere =  Dmatrix*NpadBetweenMatrices + NstartA;
+## 
+##  for(Dcol = colStart; Dcol < colEnd; Dcol++) {
+##   DcolNpad = Dcol*Npad;
+##   AHereDcol = AHere+DcolNpad;
+##   DL = 0.0;
+##   diagLocal[localIndex]=0.0;
+##   toAddLocal[localIndex]=0.0;
+##   minDcolNcache = min(Dcol, Ncache);
+##   for(Dk = localIndex; Dk < minDcolNcache; Dk += NlocalTotal) {
+##     DL = A[AHereDcol+Dk];
+##     diagLocal[Dk] = diag[diagHere+Dk] * DL;
+##     toAddLocal[localIndex] += diagLocal[Dk] * DL;
+##   }// Dk
+##   for(Dk=minDcolNcache+localIndex; Dk < Dcol; Dk += NlocalTotal) {
+##     DL = A[AHereDcol+Dk];
+##     toAddLocal[localIndex] += diag[diagHere+Dk] * DL * DL;
+##   }// Dk
+## 
+## // reduction on dimension 1
+##   barrier(CLK_LOCAL_MEM_FENCE);
+##   if( (get_local_id(1) == 0)){
+##    for(Dk = 1; Dk < get_local_size(1); Dk++) {
+##     toAddLocal[localIndex] +=  toAddLocal[localIndex + Dk];
+##    }//for Dk
+##   }//get_local_id(1) == 0
+##   barrier(CLK_LOCAL_MEM_FENCE);
+## 
+## // final reduction on dimension 0
+##   if(localIndexIsZero ){
+##    for(Dk = get_local_size(1); Dk < NlocalTotal; Dk+= get_local_size(1)) {
+##     toAddLocal[localIndex] +=  toAddLocal[Dk];
+##    } //Dk
+##    diagDcol = A[AHereDcol+Dcol] - toAddLocal[localIndex];
+##    diag[diagHere+Dcol] = diagDcol;
+## #ifdef diagToOne
+##    A[AHereDcol+Dcol] = 1.0;
+## #endif
+##   }  //localIndex==0 and Dmatrix < Nmatrix
+##   barrier(CLK_LOCAL_MEM_FENCE);
+## 
+## // off diagonals
+##  for(DrowBlock = Dcol+1; DrowBlock < N; DrowBlock += get_local_size(0)) {
+##   Drow = DrowBlock + get_local_id(0);
+##   AHereDrow = AHere+Drow*Npad;
+##   DL = 0.0;
+##   if(Drow < N){
+##   for(Dk = get_local_id(1); Dk < minDcolNcache; Dk+=get_local_size(1)) {
+##    DL += A[AHereDrow+Dk] * diagLocal[Dk];
+##   } // Dk
+## 	 for(minDcolNcache + get_local_id(1); Dk < Dcol; Dk+=get_local_size(1)) {
+##     DL += A[AHereDrow+Dk] * diag[diagHere+Dk] * A[AHereDcol+Dk];
+##   }// Dk
+##    }//Drow < N
+##   toAddLocal[localIndex] = DL;
+## 
+##   // local reduction
+##   barrier(CLK_LOCAL_MEM_FENCE);
+##   if( (get_local_id(1) == 0) & (Drow < N)){
+##    DL = toAddLocal[localIndex];
+##    for(Dk = 1; Dk < get_local_size(1); Dk++) {
+##     DL +=  toAddLocal[localIndex + Dk];
+##    }//Dk
+##    A[AHereDrow+Dcol] = (A[AHereDrow+Dcol] - DL)/diagDcol;
+##   }//get_local_id(1) == 0
+##   barrier(CLK_GLOBAL_MEM_FENCE);
+## 
+##  }//DrowBlock
+## barrier(CLK_GLOBAL_MEM_FENCE);
+## } // Dcol loop
+## 
+## barrier(CLK_GLOBAL_MEM_FENCE);
+## } // Dmatrix loop
+## 
+## barrier(CLK_LOCAL_MEM_FENCE);
+## }// kernel
 \end{verbatim}

### 2021-10-26 pkg/gpuRandom/src/maternBatch.cpp
matern fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2932 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -485,13 +485,17 @@ void maternBatchVcl(
     NpadBetweenMatrices = Npad*N; // change to Npad*(Nmat+k) to insert extra rows between matrices
   
     const int Ncell = N * (N - 1)/2, maxIter = 1500;
+    int NgroupsOfParameters = numWorkItems[1]/numLocalItems[1];
+    int NparametersPerGroup = std::min(
+      numLocalItems[1], 
+      ( (int) ceil( ((T) Nmatrix) / ((T) NgroupsOfParameters) ) )
+    );
   
   fill22params(param);
       // memory for local copies of parameters
   
   viennacl::ocl::local_mem localParameters(sizeof(param(0,0)) *
-                           NlocalParams * numLocalItems[1] * 
-                           (1+ Nmatrix * numLocalItems[1] / numWorkItems[1]));
+                           NlocalParams * NparametersPerGroup);
   
   //  cl_device_type type_check = ctx.current_device().type();
 
@@ -545,7 +549,9 @@ if(verbose > 1) {
   }
   
   if(verbose) {
-    Rcpp::Rcout << startrow << " " << Nmatrix << " " << param.size1() << "\n";
+    Rcpp::Rcout << startrow << " " << Nmatrix << " " << param.size1() << " " <<
+      NgroupsOfParameters << " " << NparametersPerGroup << " " <<
+        localParameters.size() / sizeof(param(0,0)) << " " << sizeof(param(0,0)) <<  "\n";
   }
   
   
@@ -639,8 +645,8 @@ void maternBatchBackend(
     Rcpp::S4 var,
     Rcpp::S4 coords,
     Rcpp::S4 param, //22 columns 
-    Rcpp::IntegerVector Nglobal,
-    Rcpp::IntegerVector Nlocal,
+    Rcpp::IntegerVector Nglobal, // dimension 0 cells, dimension 1 parameter sets
+    Rcpp::IntegerVector Nlocal, // groups in 0 share coordinates, groups 1 share parameters
     int startrow,   // new added
     int numberofrows,
     int verbose=0) {

### 2021-10-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
integer runif completed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2934 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 0 and 4294967295,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/4294967295$ in macro not shown here.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 0 and 2147483648,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ in macro not shown here.
 At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.

### 2021-10-27 pkg/gpuRandom/R/runif.R

@@ -16,7 +16,7 @@ runif = function(
   n, 
   streams, 
   Nglobal,
-  type=c("double","float")) {
+  type=c("double","float","integer")) {
   
   
   if(length(n)>=3){
@@ -50,14 +50,15 @@ runif = function(
   
 
 
-  xVcl<-gpuR::vclMatrix(0, nrow=n[1], ncol=n[2], type=type[1])    
+  xVcl<-gpuR::vclMatrix(0L, nrow=n[1], ncol=n[2], type=type[1])    
   
   
   gpuRnBackend(xVcl,streams,Nglobal,"uniform") 
   
   invisible(streams)
   
-  if(ncol(xVcl)==1) xVcl <- as.vclVector(xVcl)
+  if(ncol(xVcl)==1) capture.output(xVcl <- as.vclVector(xVcl), file='NUL')      #xVcl <- as.vclVector(xVcl)
+  
   xVcl
   
 }

### 2021-10-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
integer runif completed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2936 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-10-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
 ##    user  system elapsed 
-##   0.100   0.004   0.459
+##   0.221   0.008   0.241
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   4.640   0.304   4.939
+##   6.549   0.997   7.544
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.709   0.004   2.326
+##   0.691   0.006   0.697
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4039421
+## [1] 0.403125
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  10.248   0.000  10.242
+##  14.814   0.007  14.814
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.607   0.012  10.764
+##   4.591   0.004   4.595
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001263657
+## [1] 0.0001293626
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  62.792   0.036  62.770
+##  90.558   0.102  90.615
 \end{verbatim}

### 2021-10-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
integer runif
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2937 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG}, making it able to generate random numbers on GPUs.
-In what follows, we will illustrate how to create streams and how to use streams to generate (standard) uniform and (standard) Gaussian random numbers.
+In what follows, we will illustrate how to create streams and how to use streams to generate uniform and (standard) Gaussian random numbers.
 
@@
 
-\subsection{Using streams to generate  i.i.d. $U (0, 1)$ random numbers on GPU}
+\subsection{Using streams to generate uniform random numbers on GPU}
 The streams created sequentially on a GPU are then used in work-items that executes in parallel on a GPU device.
@@
 \code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 0 and 2147483648,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ in macro not shown here.
-At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ if to generate double-precision random numbers.
+(ADD THINGS ABOUT RANDOM INTEGERS) At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
 \begin{framed}
@@
 
-Now we use the streams created in section \ref{create_streams} to generate two vectors of double-precison i.i.d.
+Now we use the streams created in section \ref{create_streams} to generate two vectors of double-precision i.i.d.
 $U(0,1)$ random numbers with the function \fct{runif}.
@@
 ##    user  system elapsed 
-##   0.691   0.006   0.697
+##   0.676   0.007   0.684
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.591   0.004   4.595
+##   4.173   0.009   4.182
 \end{verbatim}

### 2021-11-01 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
removed unnecessary cpp files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2943 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Ruoyong Xu\\University of Toronto
-   \And Patrick Brown\\University of Toronto}
-\Plainauthor{Ruoyong Xu, Patrick Brown}
+   \And Patrick Brown\\University of Toronto
+   \And Pierre L’Ecuyer\\University of Montreal}
+\Plainauthor{Ruoyong Xu, Patrick Brown, Pierre L’Ecuyer}
 
-\title{Random number generation and applications on GPU’s in R(?)}
+\title{A tool set for random number generation on GPUs in R(?)}
 \Plaintitle{}
@@
 \Abstract{
+We introduce the \proglang{R} package \pkg{gpuRandom} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by utilizing the \pkg{clRNG} (OpenCL) library.
 Parallel processing by graphics processing unit (GPU) can be used to speed up computationally intensive tasks, which when combined with \proglang{R}, it can largely improve \proglang{R}’s downsides in terms of slow speed, memory usage and computation mode.
 There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling.
-We introduce the \proglang{R} package \pkg{gpuRandom} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by utilizing the \pkg{clRNG} (OpenCL) library, the package enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
+The package enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
 \pkg{gpuRandom} is portable and flexible, developers can use its random number generation kernel for various other purposes and applications.
@@
 While most random number generators are sequential, the \proglang{R} packages \pkg{parallel} (not on cran, how to cite?) and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
-More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG), \pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented (?).
+More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG).
+\pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented (?).
 
@@
 \pkg{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators.
+These four RNGs use different types of constructions.
+But we can create multiple streams with any of those four types.
 \pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG}, making it able to generate random numbers on GPUs.
@@
 Creating a new stream amounts to computing its starting point (seed).
-Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object}, which are occasionally useful.
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object,rLEC15a}, which are occasionally useful.
 
@@
 
-In \proglang{clRNG} library, the initial state of the first stream has a default value (12345).
-For the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
 Each stream is further partitioned into substreams of length $W = 2^{72}$.
@@
 This size of state is appropriate for having streams running in work items on GPU cards, for example, while providing a sufficient period length for most applications.
+The initial state of the first stream for the MRG31k3p is by default a vector of six integers ``12345''.
 
@@
 Because each time a random number is generated, the current state of the stream advances by one position.
-Below shows the \code{myStreamsGpu1} after it being used.
-<<eval=TRUE,tidy=TRUE, include=TRUE>>=
+%Below shows the \code{myStreamsGpu1} after it being used.
+<<eval=TRUE,tidy=TRUE, include=FALSE>>=
 t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
@@
 \centering
-<<largenumberuniforms, eval=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold'>>=
+<<largenumberuniforms, eval=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold', dev='png'>>=
 hist(random_u, prob = TRUE, main="Histogram")
@@
 \centering
- <<histAndQQ, echo=TRUE, eval=TRUE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold'>>=
+ <<histAndQQ, echo=TRUE, eval=TRUE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
 hist(avector,breaks=40, main = "histogram")
@@
 Hence, the option \code{simulate.p.value = TRUE} in \fct{stats::fisher.test} is provided, which enables computing p-values by Monte Carlo simulation for tables larger than 2 $\times$ 2.
+
+The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table.
+This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}.
 Given an observed table and a number of replicates $B$, the Monte Carlo simulation for Fisher's exact test does the following steps:
@@
 \end{enumerate}
-The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table.
-This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}.
 
-\fct{stats::fisher.test} can take a long running time when $B$ reaches millions as the $B$ random tables are constructed sequentially, one in each iteration.
-\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration), and the steps 2 and 3
-are also done in parallel by work-items at each iteration on GPU. Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
+In \fct{stats::fisher.test}, the $B$ random tables are constructed sequentially, one in each iteration.
+\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration).
+Each random table requires multiple calls to \fct{clrngMrg31k3pNextState} in \fct{rcont2} to generate $U(0,1)$ random numbers, the number of calls depends on the row and column number of the observed table.
+For more about the algorithm used the function \fct{rcont2} for generating random two-way table with given marginal totals, see \citep{patefield1981efficient}.
+Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially, the test statistics (if returned) and the count?
+are saved on a GPU, the last step for calculating the p-value is done in R.
+Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
+p-value is calculated using the simple random sampling method.
 
@@
 \centering
-<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE>>=
+<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE, dev='png'>>=
 hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
@@
 %Due to the importance of Gaussian random field, there are a large amount of studies devoted to Gaussian random field generation techniques. 
-\cite{LiuandLi2019} gives a comprehensive review on 7 popular methods for Gaussian random field generation, which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
-All of these methods except matrix decomposition method, are approximations and has specific requirements on the type of grid or covariance functions.
+\cite{LiuandLi2019} gives a comprehensive review on 7 popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
+all of these methods except matrix decomposition method, are approximations and has specific requirements on the type of grid or covariance functions.
 The matrix decomposition method is exact, it works for all covariance functions and can generate random field on all types of grids.
@@
 \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
-<<r multLowerdiag_LDZ, eval=TRUE, cache=FALSE, results='hide'>>=
+<<r multLowerdiag_LDZ, eval=TRUE, cache=FALSE, results='hide',dev='png'>>=
 simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
@@
 \centering
-<<maternplot, echo=TRUE, fig.show='hold', eval=TRUE, dev='pdf',cache=FALSE>>=
+<<maternplot, echo=TRUE, fig.show='hold', eval=TRUE, dev='pdf',cache=FALSE,dev='png'>>=
 par(mar=rep(0.1, 4))
@@
 
-One of the main difficulties in developing the package is bug fixes and uncertain software behaviors.
+%One of the main difficulties in developing the package is bug fixes and uncertain software behaviors.
 Error messages are often vague and it is time-consuming to locate the error within the backend \proglang{C} program.

### 2021-11-01 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
removed unnecessary cpp files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2943 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Ruoyong Xu\\University of Toronto
-   \And Patrick Brown\\University of Toronto}
-\Plainauthor{Ruoyong Xu, Patrick Brown}
-
-\title{Random number generation and applications on GPU’s in R(?)}
+   \And Patrick Brown\\University of Toronto
+   \And Pierre L’Ecuyer\\University of Montreal}
+\Plainauthor{Ruoyong Xu, Patrick Brown, Pierre L’Ecuyer}
+
+\title{A tool set for random number generation on GPUs in R(?)}
 \Plaintitle{}
@@
 Because each time a random number is generated, the current state of the stream advances by one position.
-Below shows the \code{myStreamsGpu1} after it being used.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu),
-    dimnames = dimnames(myStreamsCpu)))
-##                    [,1]       [,2]       [,3]       [,4]
-## current.g1.1  878672095 2113333390  502033783  739421137
-## current.g1.2  240667857  559530223 1322587635 1475938232
-## current.g1.3  240667857 1309565828 1964121530  730262207
-## current.g2.1  642281259 1335994581 1949818481 1630192198
-## current.g2.2 1069151070   61444481 1607232546  324551134
-## current.g2.3  809054265  197003928 1462898381  795289868
-## initial.g1.1      12345  336690377  502033783  739421137
-## initial.g1.2      12345  597094797 1322587635 1475938232
-## initial.g1.3      12345 1245771585 1964121530  730262207
-## initial.g2.1      12345   85196284 1949818481 1630192198
-## initial.g2.2      12345  523477687 1607232546  324551134
-## initial.g2.3      12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+%Below shows the \code{myStreamsGpu1} after it being used.
+
 Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart.
@@
                              Nglobal=c(512,128), type="double"))
-##    user  system elapsed 
-##   0.221   0.008   0.241
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in cpp\_vcl\_nrow(x@address, 4L): external pointer is not valid}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Timing stopped at: 0.001 0 0.002}}\begin{verbatim}
 system.time(matrix(stats::rnorm(10000^2),10000,10000))
 ##    user  system elapsed 
-##   6.549   0.997   7.544
+##   6.695   0.749   7.438
 \end{verbatim}
@@
 \end{enumerate}
-The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table.
-This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}.
-
-\fct{stats::fisher.test} can take a long running time when $B$ reaches millions as the $B$ random tables are constructed sequentially, one in each iteration.
-\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration), and the steps 2 and 3
-are also done in parallel by work-items at each iteration on GPU. Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
+
+In \fct{stats::fisher.test}, the $B$ random tables are constructed sequentially, one in each iteration.
+\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration).
+Each random table requires multiple calls to \fct{clrngMrg31k3pNextState} in \fct{rcont2} to generate $U(0,1)$ random numbers, the number of calls depends on the row and column number of the observed table.
+For more about the algorithm used the function \fct{rcont2} for generating random two-way table with given marginal totals, see \citep{patefield1981efficient}.
+Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially, the test statistics (if returned) and the count?
+are saved on a GPU, the last step for calculating the p-value is done in R.
+Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
+p-value is calculated using the simple random sampling method.
 
@@
 ##    user  system elapsed 
-##   0.676   0.007   0.684
+##   0.705   0.011   0.714
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.403125
+## [1] 0.403691
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  14.814   0.007  14.814
+##  14.717   0.006  14.713
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.173   0.009   4.182
+##   4.371   0.012   4.382
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001293626
+## [1] 0.0001341575
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  90.558   0.102  90.615
+##  90.251   0.089  90.287
 \end{verbatim}

### 2021-11-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update pkg
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2944 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling.
-The package enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
-\pkg{gpuRandom} is portable and flexible, developers can use its random number generation kernel for various other purposes and applications.
+\pkg{gpuRandom} enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
+This package is portable and flexible, developers can use its random number generation kernel for various other purposes and applications.
 }
@@
 To be more specific, (1) \proglang{R} requires all data to be loaded into the main memory (RAM) and thus can handle a very limited size of data; (2) \proglang{R} is a single-threaded program, it can not effectively use all the computing cores of multi-core processors.
-Parallel computing is the solution to these drawbacks of \proglang{R}, for an overview of current parallel computing approaches with \proglang{R}, see CRAN Task View by \citet{cran2021} at \url{https://cran.r-project.org/web/views/HighPerformanceComputing.html}.
+Parallel computing is the solution to these drawbacks, for an overview of current parallel computing approaches with \proglang{R}, see CRAN Task View by \citet{cran2021} at \url{https://cran.r-project.org/web/views/HighPerformanceComputing.html}.
 
@@
 \pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented (?).
+For up-to-date review papers on the generation of random numbers on parallel devices, and GPUs in particular, see:  \cite{rLEC15a,rLEC17p,rLEC21a}.
 
@@
 
-\section{Generating random numbers} \label{}
+\section{Uniform random number generation} \label{}
 Uniform random number generators (RNGs) are essential in simulating random numbers from all types of probability distributions.
@@
 These four RNGs use different types of constructions.
-But we can create multiple streams with any of those four types.
-\pkg{gpuRandom} package uses the MRG31k3p RNG from the \proglang{clRNG}, making it able to generate random numbers on GPUs.
-In what follows, we will illustrate how to create streams and how to use streams to generate uniform and (standard) Gaussian random numbers.
+\pkg{gpuRandom} package uses the MRG31k3p RNG, making it able to generate random numbers on GPUs.
+We choose MRG31k3p as the base generator for the following reasons:  the original \pkg{RNGStreams} package \citep{l2002object} was built with MRG32k3a, which was designed to be implemented in ``double'' in floating point(?), and not with 32-bit integers.
+The MRG31k3p generator was designed later, specifically for 32-bit integer arithmetic, so it runs faster on the 32-bit GPUs.
+It is also faster than Philox-4×32-10 \citep{rLEC00b}.
+MRG31k3p was also statistically tested extensively and successfully in \citep{rLEC07b}.
+
+In what follows, we will illustrate how to create streams and how to use streams to generate uniform random numbers.
 
@@
 Creating a new stream amounts to computing its starting point (seed).
-Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object,rLEC15a}, which are occasionally useful.
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, which are occasionally useful.
 
@@
 The state (and seed) of each stream is a vector of six 31-bit integers.
-This size of state is appropriate for having streams running in work items on GPU cards, for example, while providing a sufficient period length for most applications.
+This size of state is appropriate for having streams running in work-items on GPU cards, for example, while providing a sufficient period length for most applications.
 The initial state of the first stream for the MRG31k3p is by default a vector of six integers ``12345''.
@@
 \pkg{gpuRandom} is able to create streams both on the host and on the GPU device.
-\fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices (see the \proglang{R} output below), creating streams on host makes it easy for us to view and check the streams behavior.
-\fct{createStreamsGpu} is adapted from the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on a GPU device.
-Substreams are not created because currently we use one stream per work-item.
-%(From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
-delete?)
-In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively. 
+\fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
+Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item.
 <<createStreams_CPU>>=
@@
 \begin{itemize}
+\itemsep0em 
   \item \code{n}: number of streams to create.
@@
 \end{itemize}
-We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams directly on GPU using \fct{createStreamsGpu}.
+\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
+We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}.
 Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
@@
 
-\subsection{Using streams to generate uniform random numbers on GPU}
+\subsection{Using streams to generate uniform random numbers}
 The streams created sequentially on a GPU are then used in work-items that executes in parallel on a GPU device.
@@
 \code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ if to generate double-precision random numbers.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$, if to generate double-precision random numbers.
 (ADD THINGS ABOUT RANDOM INTEGERS) At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
 \begin{framed}
-\begin{lstlisting}[caption={captiontext},language=C,basicstyle=\small,label={lst:uniformkernel}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel}]
  __kernel void mrg31k3pMatrix(
@@
 \begin{itemize}
+\itemsep0em 
   \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
@@
 \end{itemize}
-\fct{runif} returns the same sequence of random numbers if it uses the same streams in the same current state (here \code{myStreamsGpu1} and \code{myStreamsGpu2}), not streams starting from the same states.
+\fct{runif} returns the same sequence of random numbers if it uses the same streams (here \code{myStreamsGpu1} and \code{myStreamsGpu2}), not streams starting from the same states.
 Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
@@
 However, we can save the streams on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, in this way we can reproduce the exact same sequence of random numbers in simulations.
-Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
+Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
 <<save streams,echo=TRUE,eval=TRUE>>=
@@
 
+\subsection{Plot the uniform random numbers}
 We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
@@
 streams <- createStreamsGpu(n = 512)
-
 random_u<-as.vector(runif(n=50000, streams=streams, Nglobal=c(64,8), type="double"))
@@
 \centering
-<<largenumberuniforms, eval=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold', dev='png'>>=
+<<largenumberuniforms, eval=FALSE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold', dev='png'>>=
 hist(random_u, prob = TRUE, main="Histogram")
@@
 
-\subsection{Using streams to generate i.i.d. N(0, 1) random numbers on GPU}
+\section{Normal random number generation}
 We apply the Box-Muller transformation to $U(0,1)$ random numbers to generate standard Gaussian random numbers.
@@
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,caption={name},label={lst:normal}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal}]
 const int 
@@
 
-In the following code, We create a $1024 \times 512$ matrix %named \code{normal_matrix} 
-of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
+Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
+The difference in elapsed time becomes larger when matrix size goes larger.
+<<random normal time compare, eval=TRUE, cache=FALSE>>=
+streams <- createStreamsGpu(n = 512 * 128)
+system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
+                             Nglobal=c(512,128), type="double"))
+@
+<<random normal time compare2, eval=TRUE, cache=TRUE>>=
+system.time(matrix(stats::rnorm(10000^2),10000,10000))
+@
+
+TO BE PUT IN APPENDIX OR DELETE? \\
+In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
 We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-<<histAndQQ_prepare,eval=TRUE,cache=TRUE>>=
+<<histAndQQ_prepare,eval=FALSE,cache=TRUE>>=
 streams <- createStreamsGpu(n =512*128)
@@
 \centering
- <<histAndQQ, echo=TRUE, eval=TRUE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
+ <<histAndQQ, echo=TRUE, eval=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
 hist(avector,breaks=40, main = "histogram")
@@
 \end{figure}
-
-We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
-Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
-The difference in elapsed time becomes larger when matrix size goes larger.
-<<random normal time compare, eval=TRUE, cache=TRUE>>=
-system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
-                             Nglobal=c(512,128), type="double"))
-system.time(matrix(stats::rnorm(10000^2),10000,10000))
-@
 
@@
 \begin{enumerate}
+\setlength\itemsep{0em}
   \item Calculate the test statistic for the observed table.
 	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic from the random table. 
-	\item Count the number of iterations that have test statistics less or equal to the one from the observed table (\emph{Counts}).
+	\item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
 	\item Estimate p-value using $\frac{1 + \emph{Counts}}{B + 1}$.
@@
 
-In \fct{stats::fisher.test}, the $B$ random tables are constructed sequentially, one in each iteration.
-\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration).
-Each random table requires multiple calls to \fct{clrngMrg31k3pNextState} in \fct{rcont2} to generate $U(0,1)$ random numbers, the number of calls depends on the row and column number of the observed table.
-For more about the algorithm used the function \fct{rcont2} for generating random two-way table with given marginal totals, see \citep{patefield1981efficient}.
-Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially, the test statistics (if returned) and the count?
-are saved on a GPU, the last step for calculating the p-value is done in R.
+Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially.
+For step 2, \fct{gpuRandom::fisher.sim} adapted the function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random two-way tables with given marginal totals.
+The algorithm \citep[see][]{patefield1981efficient} samples the entries row by row, one at a time, conditional on the values of the entries already sampled.
+The conditional probabilities for the possible values of the next entry are updated dynamically each time a new entry is sampled.
+Then this entry is sampled by standard inversion of the cumulative distribution function, using one $U(0,1)$ random number.
+For an $I \times J$ table, this requires $(I-1)(J-1)$ random numbers (the last column and last row do not need to be sampled).
+Finally, one can compute the test statistic for this newly sampled table.
+On a GPU, this step can be replicated say $n$ times in parallel by creating $n$ distinct random streams and launching $n$ separate work-items.
+Each work-item takes a random stream as input, performs all of Step 2 and 3, and returns the value of the test statistic and count on the GPU.
+Computing the p-value on the CPU (Step 4) is then a trivial operation.
 Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
-p-value is calculated using the simple random sampling method.
+%There are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo.
+The use of streams and GPU would be quite different in that case.
+See for example
 
@@
 The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
-The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
+%The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
 The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
 
-\subsection{Comparing running time: Month data example}
+\subsection{Comparing run-time: Month data example}
 <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
@@
 @
-<<monthdata,eval=TRUE,echo=FALSE, cache=FALSE>>=
+<<monthdata,eval=TRUE,include=FALSE, cache=TRUE>>=
 library(knitr)
@@
 
-\subsection{Comparing running time: Week data example}
+\subsection{Comparing run-time: Week data example}
 <<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
@@
 @
-<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
+<<weekdata, eval=TRUE, include=FALSE, cache=TRUE>>=
 knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
@@
 
-\section{Generating Gaussian random fields} 
+\section{Gaussian surface simulation} 
 Given a parameter space $X$, a random field $U$ on $X$ is a collection of random variables $\{U_x :x\in X\}$.
 A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})$ is multivariate Gaussian.
-%One reason that Gaussian random fields play an important role is that the specification of their finite-dimensional distributions is simple \citep{abrahamsen1997review},
 The expectation $\mu(x)$ and covariance function $\Sigma$ completely determine the distribution of $U$, where
@@
 \begin{itemize}
+\itemsep0em 
 \item $L$ is the lower unit triangular matrix in Cholesky decomposition of $\Sigma = L*D*L^\top$,
@@
 \end{itemize}
-The matrix decomposition method is straightforward to implement, however, when a large number of locations ($n$) is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$ \citep{LiuandLi2019},  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
+The matrix decomposition method is straightforward to implement, however, when a large number of locations ($n$) is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
 \pkg{gpuRandom} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
@@
 \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
-<<r multLowerdiag_LDZ, eval=TRUE, cache=FALSE, results='hide',dev='png'>>=
+<<r multLowerdiag_LDZ, eval=TRUE, cache=FALSE, results='hide'>>=
 simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
@@
 The package \pkg{gpuRandom} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
-We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the Normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
-Most of these functions uses local memory on GPU %all of them uses parallel programming and avoids unnecessary data transfers between host and device.
+We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
+Most of these functions uses local memory on GPU.
 Comparison of performance between using \pkg{gpuRandom} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
@@
 By leveraging the \pkg{gpuR} package, \pkg{gpuRandom} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.
-\pkg{gpuRandom} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for further development.
+\pkg{gpuRandom} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for other applications.
 

### 2021-11-02 pkg/gpuRandom/src/lgmlikFit.hpp

@@ -12,7 +12,7 @@
 
 
 template <typename T> std::string cholBatchKernelString(
-    int colStart,
+int colStart,
 int colEnd,
 int N,
 int Npad,
@@ -150,12 +150,12 @@ void maternBatchVcl(
         const int ctx_id);
     
     
-    template <typename T>
-    void rowsum(
-        viennacl::matrix<T> &x,
-        viennacl::vector_base<T> &Sum,
-        std::string type,
-        int log);
+    // template <typename T>
+    // void rowsum(
+    //     viennacl::matrix<T> &x,
+    //     viennacl::vector_base<T> &Sum,
+    //     std::string type,
+    //     int log);
         
         
         template <typename T> 
@@ -207,23 +207,23 @@ void maternBatchVcl(
         
         
         
-template<typename T> 
-void matrix_vector_sum(
-     viennacl::matrix<T> &matrix,// viennacl::vector_base<int>  rowSum, viennacl::vector_base<int>  colSum,  
-     viennacl::vector_base<T> &vector,
-     viennacl::matrix<T> &sum,
-     const int byrow,
-     Rcpp::IntegerVector numWorkItems,
-     const int ctx_id);            
-                
-                
-                template<typename T> 
-                void matrix_scalar_sum(
-                    viennacl::matrix<T> &matrix,// viennacl::vector_base<int>  rowSum, viennacl::vector_base<int>  colSum,  
-                    T  value,         //viennacl::scalar<T> value,     
-                    viennacl::matrix<T> &sum,
-                    Rcpp::IntegerVector numWorkItems,
-                    const int ctx_id);                
+// template<typename T> 
+// void matrix_vector_sum(
+//      viennacl::matrix<T> &matrix,// viennacl::vector_base<int>  rowSum, viennacl::vector_base<int>  colSum,  
+//      viennacl::vector_base<T> &vector,
+//      viennacl::matrix<T> &sum,
+//      const int byrow,
+//      Rcpp::IntegerVector numWorkItems,
+//      const int ctx_id);            
+//                 
+//                 
+//                 template<typename T> 
+//                 void matrix_scalar_sum(
+//                     viennacl::matrix<T> &matrix,// viennacl::vector_base<int>  rowSum, viennacl::vector_base<int>  colSum,  
+//                     T  value,         //viennacl::scalar<T> value,     
+//                     viennacl::matrix<T> &sum,
+//                     Rcpp::IntegerVector numWorkItems,
+//                     const int ctx_id);                
                 
                 
                 

### 2021-11-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update pkg
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2944 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \pkg{gpuRandom} is able to create streams both on the host and on the GPU device.
-\fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices (see the \proglang{R} output below), creating streams on host makes it easy for us to view and check the streams behavior.
-\fct{createStreamsGpu} is adapted from the \fct{clrngMrg31k3pCreateStreams} and creates streams directly on a GPU device.
-Substreams are not created because currently we use one stream per work-item.
-%(From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{createStreamsGpu} is more efficient and more frequently used.
-delete?)
-In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively. 
+\fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
+Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item.
 \begin{knitrout}
@@
 However, we can save the streams on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, in this way we can reproduce the exact same sequence of random numbers in simulations.
-Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
+Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
 \begin{knitrout}
@@
 \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
-\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-1} 
-\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-2} 
-\end{knitrout}
+
 \caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
@@
 
-In the following code, We create a $1024 \times 512$ matrix %named \code{normal_matrix} 
-of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
+Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
+The difference in elapsed time becomes larger when matrix size goes larger.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+streams <- createStreamsGpu(n = 512 * 128)
+system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
+                             Nglobal=c(512,128), type="double"))
+##    user  system elapsed 
+##   0.138   0.000   0.137
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+system.time(matrix(stats::rnorm(10000^2),10000,10000))
+##    user  system elapsed 
+##   6.608   0.796   7.399
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+TO BE PUT IN APPENDIX OR DELETE? \\
+In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
@@
 \end{kframe}
-\includegraphics[width=0.45\textwidth]{figure/histAndQQ-1} 
-\includegraphics[width=0.45\textwidth]{figure/histAndQQ-2} 
 \end{knitrout}
@@
 \end{figure}
-
-We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
-Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
-The difference in elapsed time becomes larger when matrix size goes larger.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
-                             Nglobal=c(512,128), type="double"))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in cpp\_vcl\_nrow(x@address, 4L): external pointer is not valid}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Timing stopped at: 0.001 0 0.002}}\begin{verbatim}
-system.time(matrix(stats::rnorm(10000^2),10000,10000))
-##    user  system elapsed 
-##   6.695   0.749   7.438
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 
@@
 \begin{enumerate}
+\setlength\itemsep{-0.1em}
   \item Calculate the test statistic for the observed table.
 	\item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic from the random table. 
-	\item Count the number of iterations that have test statistics less or equal to the one from the observed table (\emph{Counts}).
+	\item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
 	\item Estimate p-value using $\frac{1 + \emph{Counts}}{B + 1}$.
@@
 
-In \fct{stats::fisher.test}, the $B$ random tables are constructed sequentially, one in each iteration.
-\fct{gpuRandom::fisher.sim} adapted the backend function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random tables, so that it generates \code{prod(Nglobal)} random tables in parallel in each iteration (i.e., one random table each work-item per iteration).
-Each random table requires multiple calls to \fct{clrngMrg31k3pNextState} in \fct{rcont2} to generate $U(0,1)$ random numbers, the number of calls depends on the row and column number of the observed table.
-For more about the algorithm used the function \fct{rcont2} for generating random two-way table with given marginal totals, see \citep{patefield1981efficient}.
-Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially, the test statistics (if returned) and the count?
-are saved on a GPU, the last step for calculating the p-value is done in R.
+Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially.
+For step 2, \fct{gpuRandom::fisher.sim} adapted the function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random two-way tables with given marginal totals.
+The algorithm \citep[see][]{patefield1981efficient} samples the entries row by row, one at a time, conditional on the values of the entries already sampled.
+The conditional probabilities for the possible values of the next entry are updated dynamically each time a new entry is sampled.
+Then this entry is sampled by standard inversion of the cumulative distribution function, using one $U(0,1)$ random number.
+For an $I \times J$ table, this requires $(I-1)(J-1)$ random numbers (the last column and last row do not need to be sampled).
+Finally, one can compute the test statistic for this newly sampled table.
+On a GPU, this step can be replicated say $n$ times in parallel by creating $n$ distinct random streams and launching $n$ separate work-items.
+Each work-item takes a random stream as input, performs all of Step 2, and returns the value of the test statistic and count on the GPU.
+Computing the p-value on the CPU (Step 4) is then a trivial operation.
 Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
-p-value is calculated using the simple random sampling method.
+%There are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo.
+The use of streams and GPU would be quite different in that case.
+See for example
 
@@
 The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
-The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
+%The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
 The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
 
-\subsection{Comparing running time: Month data example}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
-
-\caption{\label{tab:monthdata}Month\label{tab:month}}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
-\hline
-Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
-\hline
-Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
-\hline
-Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
-\hline
-Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
-\hline
-May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
-\hline
-Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
-\hline
-Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
-\hline
-Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
-\hline
-Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
-\hline
-Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
-\hline
-Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
-\hline
-Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
-\hline
-\end{tabular}
-\end{table}
-
-\end{knitrout}
+\subsection{Comparing run-time: Month data example}
 
@@
 ##    user  system elapsed 
-##   0.705   0.011   0.714
+##   0.708   0.011   0.718
 \end{verbatim}
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.403691
+## [1] 0.4029291
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  14.717   0.006  14.713
+##  14.762   0.003  14.757
 \end{verbatim}
@@
 
-\subsection{Comparing running time: Week data example}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
-
-\caption{\label{tab:weekdata}Week\label{tab:week}}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
-\hline
-Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
-\hline
-Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
-\hline
-Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
-\hline
-Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
-\hline
-Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
-\hline
-Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
-\hline
-Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
-\hline
-\end{tabular}
-\end{table}
-
-\end{knitrout}
+\subsection{Comparing run-time: Week data example}
 
@@
 ##    user  system elapsed 
-##   4.371   0.012   4.382
+##   4.350   0.017   4.364
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001341575
+## [1] 0.0001209715
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  90.251   0.089  90.287
+##  90.139   0.098  90.181
 \end{verbatim}

### 2021-11-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
random seeds in streams GPU
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2946 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 library("gpuRandom")
-myStreamsCpu <- createStreamsCpu(n=4, seed=12345)
+myStreamsCpu <- createStreamsCpu(n=4, initial=12345)
 t(myStreamsCpu)
@@
 myStreamsGpu1 = vclMatrix(myStreamsCpu)
-myStreamsGpu2 = createStreamsGpu(n=4, seed=12345)
+myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
 range(as.matrix(myStreamsGpu1 - myStreamsGpu2))

### 2021-11-04 pkg/gpuRandom/R/CreateStreamsGpu.R

@@ -11,19 +11,25 @@
 
 
 
+
 createStreamsGpu = function(n, 
-                            seed=12345){
-
-    seed = rep_len(seed, 6)
-    seedVec <- gpuR::vclVector(as.integer(seed), type="integer")  
-    streamsR<-gpuR::vclMatrix(0L, nrow=as.integer(n), ncol=12, type="integer")
-
-    gpuRandom:::CreateStreamsGpuBackend(seedVec, streamsR, keepInitial=1)
+                            initial){
   
-    streamsR
+  if(missing(initial)) {
+    seed = as.integer(as.integer(2^31-1)*(2*stats::runif(6) - 1) ) 
+    seedVec <- gpuR::vclVector(seed, type="integer") 
+  }else{
+    seed = rep_len(initial, 6)
+    seedVec <- gpuR::vclVector(as.integer(seed), type="integer")  
+  }
+  
+  streamsR<-gpuR::vclMatrix(0L, nrow=as.integer(n), ncol=12, type="integer")
+  
+  gpuRandom:::CreateStreamsGpuBackend(seedVec, streamsR, keepInitial=1)
+  
+  streamsR
   
 }
-
 
 
 

### 2021-11-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
random exponential finished
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2947 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.138   0.000   0.137
+##   0.133   0.000   0.133
 \end{verbatim}
@@
 \end{knitrout}
-
-TO BE PUT IN APPENDIX OR DELETE? \\
-In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
-The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streams <- createStreamsGpu(n =512*128)
-normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
-                                Nglobal=c(512,128), type="float")
-avector<-as.vector(as.matrix(normal_matrix))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{figure}[H]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-hist(avector,breaks=40, main = "histogram")
-gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
-\end{figure}
 
@@
 \begin{enumerate}
-\setlength\itemsep{-0.1em}
+\setlength\itemsep{0em}
   \item Calculate the test statistic for the observed table.
@@
 On a GPU, this step can be replicated say $n$ times in parallel by creating $n$ distinct random streams and launching $n$ separate work-items.
-Each work-item takes a random stream as input, performs all of Step 2, and returns the value of the test statistic and count on the GPU.
+Each work-item takes a random stream as input, performs all of Step 2 and 3, and returns the value of the test statistic and count on the GPU.
 Computing the p-value on the CPU (Step 4) is then a trivial operation.
@@
 
-We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
+The arguments of the \fct{gpuRandom::fisher.sim} are described as follows:
+\begin{itemize}
+\itemsep0em 
+  \item \code{x}: a vclMatrix.
+  \item \code{N}: number of simulation runs.
+  \item \code{streams}: streams.
+  \item \code{type}: ``double'' or ``float'' or returned test statistics.
+  \item \code{returnStatistics}: if TRUE, return all test statistics. 
+  \item \code{Nglobal}: global index space. Default is set as $(64,16)$.
+\end{itemize}
+We show the advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
 
-% \begin{table}[H]
-% \begin{tabular}{ |c|c|c|c|c|c| } 
-% \hline
-%            & host   &                     device \\ \hline
-%  Computer1 & INtel Xenon W-2145 3.7 Ghz & Amd Radeon VII  \\ \hline
-%  Computer2 & VCPU Intel Xenon Skylake 2.5ghz & VGPU Nvidia Tesla V100 \\ \hline
-% \end{tabular}
-% \caption{ Device information for run-time compare.\label{tab:device}}
-% \end{table}
-
-The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
-%The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
-The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
+The 2-way contingency tables %\ref{tab:month} and table \ref{tab:week} 
+consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
 
@@
 result_month$counts
-## [1] 410644
+## [1] 410533
 result_month$p.value
-## [1] 0.4042541
+## [1] 0.4041449
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.708   0.011   0.718
+##   0.702   0.005   0.707
 \end{verbatim}
@@
 result_weekgpu$counts
-## [1] 1242
+## [1] 1259
 result_weekgpu$p.value
-## [1] 0.0001239652
+## [1] 0.0001256606
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.350   0.017   4.364
+##   4.324   0.011   4.332
 \end{verbatim}
@@
 \end{knitrout}
-% \begin{table}[H]
-% \centering
-% \begin{tabular}{ |c|c|c|c|c|c| } 
-%  \hline
-%  data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
-%  Month & 1e6 & 0.403804 & 0.403507  & 0.403507     \\ \hline
-%  Week &  1e7  & 0.0001251 & 0.0001274 & 0.0001274\\ \hline
-% \end{tabular}
-% \caption{ Summary of p-values obtained on different devices.\label{tab:summary}}
-% \end{table}
-% 
-% 
-% \begin{table}[H]
-% \centering
-% \begin{tabular}{ |c|c|c|c|c|c| } 
-%  \hline
-%  data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
-%  Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
-%  Week &  1e7  & 60.19 & 90.38 & 10.82 & 1.90\\ \hline
-% \end{tabular}
-% \caption{ Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.\label{tab:compare time}}
-% \end{table}
 
-Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example.
-The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
 \begin{figure}[H]
@@
 \end{knitrout}
-\caption{ Frequency distribution for test statistics from month and week data tables.\label{fig4}}
+\caption{Frequency distribution for test statistics from month and week data tables.
+Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example.
+The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
+\label{fig4}}
 \end{figure}
@@
 \begin{appendix}
+TO BE PUT IN APPENDIX OR DELETE? \\
+In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+streams <- createStreamsGpu(n =512*128)
+normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
+                                Nglobal=c(512,128), type="float")
+avector<-as.vector(as.matrix(normal_matrix))
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\begin{figure}[H]
+\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+hist(avector,breaks=40, main = "histogram")
+gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
+\end{figure}
 
+The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
+The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+
+\caption{\label{tab:monthdata}Month\label{tab:month}}
+\centering
+\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
+\hline
+  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
+\hline
+Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
+\hline
+Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
+\hline
+Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
+\hline
+Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
+\hline
+May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
+\hline
+Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
+\hline
+Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
+\hline
+Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
+\hline
+Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
+\hline
+Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
+\hline
+Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
+\hline
+Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
+\hline
+\end{tabular}
+\end{table}
+
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+
+\caption{\label{tab:weekdata}Week\label{tab:week}}
+\centering
+\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
+\hline
+  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
+\hline
+Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
+\hline
+Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
+\hline
+Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
+\hline
+Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
+\hline
+Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
+\hline
+Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
+\hline
+Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
+\hline
+\end{tabular}
+\end{table}
+
+\end{knitrout}
 \end{appendix}

### 2021-11-05 pkg/gpuRandom/R/runif.R

@@ -3,7 +3,7 @@
 #' @param n a number or a vector specifying the size of output matrix
 #' @param streams streams matrix of random number seeds
 #' @param Nglobal vector of length 2, number of work items
-#' @param type random number type, one of "uniform" or "normal"
+#' @param type precision type of random numbers, "double" or "float" or "integer"
 #' 
 #' @return a vclVector or vclMatrix of uniform random numbers
 #' 

### 2021-11-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
random exponential finished
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2947 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-TO BE PUT IN APPENDIX OR DELETE? \\
-In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
-The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-<<histAndQQ_prepare,eval=FALSE,cache=TRUE>>=
-streams <- createStreamsGpu(n =512*128)
-normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
-                                Nglobal=c(512,128), type="float")
-avector<-as.vector(as.matrix(normal_matrix))
-@
-\begin{figure}[H]
-\centering
- <<histAndQQ, echo=TRUE, eval=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
-hist(avector,breaks=40, main = "histogram")
-gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
-@
-\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
-\end{figure}
-
 \section{Monte Carlo simulation for Fisher's exact test}
@@
 
-We show the usage and advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
+The arguments of the \fct{gpuRandom::fisher.sim} are described as follows:
+\begin{itemize}
+\itemsep0em 
+  \item \code{x}: a vclMatrix.
+  \item \code{N}: number of simulation runs.
+  \item \code{streams}: streams.
+  \item \code{type}: ``double'' or ``float'' or returned test statistics.
+  \item \code{returnStatistics}: if TRUE, return all test statistics. 
+  \item \code{Nglobal}: global index space. Default is set as $(64,16)$.
+\end{itemize}
+We show the advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
@@
 @
-% \begin{table}[H]
-% \begin{tabular}{ |c|c|c|c|c|c| } 
-% \hline
-%            & host   &                     device \\ \hline
-%  Computer1 & INtel Xenon W-2145 3.7 Ghz & Amd Radeon VII  \\ \hline
-%  Computer2 & VCPU Intel Xenon Skylake 2.5ghz & VGPU Nvidia Tesla V100 \\ \hline
-% \end{tabular}
-% \caption{ Device information for run-time compare.\label{tab:device}}
-% \end{table}
-
-The 2-way contingency table \ref{tab:month} and table \ref{tab:week} consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
-%The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
-The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
+The 2-way contingency tables %\ref{tab:month} and table \ref{tab:week} 
+consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
 
@@
 streams <- createStreamsGpu(n =256*128)
-@
-<<monthdata,eval=TRUE,include=FALSE, cache=TRUE>>=
-library(knitr)
-library(kableExtra)
-knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
-  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
 @
@@
 colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
-@
-<<weekdata, eval=TRUE, include=FALSE, cache=TRUE>>=
-knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
-  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
 @
@@
 @
-% \begin{table}[H]
-% \centering
-% \begin{tabular}{ |c|c|c|c|c|c| } 
-%  \hline
-%  data &   B  & Intel CPU &  AMD GPU & NVIDIA GPU \\ \hline
-%  Month & 1e6 & 0.403804 & 0.403507  & 0.403507     \\ \hline
-%  Week &  1e7  & 0.0001251 & 0.0001274 & 0.0001274\\ \hline
-% \end{tabular}
-% \caption{ Summary of p-values obtained on different devices.\label{tab:summary}}
-% \end{table}
-% 
-% 
-% \begin{table}[H]
-% \centering
-% \begin{tabular}{ |c|c|c|c|c|c| } 
-%  \hline
-%  data &   B  & Intel CPU & sencha CPU & AMD GPU & NVIDIA GPU \\ \hline
-%  Month & 1e6 & 9.74 & 14.58 & 2.28 & 0.31\\ \hline
-%  Week &  1e7  & 60.19 & 90.38 & 10.82 & 1.90\\ \hline
-% \end{tabular}
-% \caption{ Run-time compare of Fisher's test simulation on different devices, all numbers are in seconds.\label{tab:compare time}}
-% \end{table}
 
-Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example.
-The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
 \begin{figure}[H]
@@
 @
-\caption{ Frequency distribution for test statistics from month and week data tables.\label{fig4}}
+\caption{Frequency distribution for test statistics from month and week data tables.
+Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example.
+The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
+\label{fig4}}
 \end{figure}
@@
 \begin{appendix}
+TO BE PUT IN APPENDIX OR DELETE? \\
+In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
+<<histAndQQ_prepare,eval=FALSE,cache=TRUE>>=
+streams <- createStreamsGpu(n =512*128)
+normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
+                                Nglobal=c(512,128), type="float")
+avector<-as.vector(as.matrix(normal_matrix))
+@
+\begin{figure}[H]
+\centering
+ <<histAndQQ, echo=TRUE, eval=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
+hist(avector,breaks=40, main = "histogram")
+gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
+@
+\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
+\end{figure}
 
+The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
+The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
+<<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
+library(knitr)
+library(kableExtra)
+knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
+  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
+@
+
+<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
+knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
+  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
+@
 \end{appendix}

### 2021-11-06 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
modify rexp kernel
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2948 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
    \And Patrick Brown\\University of Toronto
-   \And Pierre L’Ecuyer\\University of Montreal}
+   \And Pierre L’Ecuyer\\University of Montr\'eal}
 \Plainauthor{Ruoyong Xu, Patrick Brown, Pierre L’Ecuyer}
 
-\title{A tool set for random number generation on GPUs in R(?)}
+\title{A tool set for random number generation on GPUs in R}
 \Plaintitle{}
@@
 \Address{
-  Ruoyong Xu\\
+  Ruoyong Xu, Patrick Brown\\
   Department of Statistics\\
   University of Toronto\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}
+  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
+  E-mail: \email{ruoyong.xu@mail.utoronto.ca}, \email{patrick.brown@utoronto.ca}}
 
 \Address{
-  Achim Zeileis\\
-  Journal of Statistical Software\\
-  \emph{and}\\
-  Department of Statistics\\
-  University of Toronto\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}
+  Pierre L’Ecuyer\\
+  Department of Computer Science and Operations Research\\
+  University of Montr\'eal\\
+  Pavillon André-Aisenstadt 2920, chemin de la Tour Montréal QCH3T 1J4\\
+  E-mail: \email{lecuyer@iro.umontreal.ca}}
 
@@
 Random number generation is critical in simulation-based statistical inference, machine learning and many other scientific fields.
-While most random number generators are sequential, the \proglang{R} packages \pkg{parallel} (not on cran, how to cite?) and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
+While most random number generators are sequential, the \proglang{R} packages \pkg{parallel} (not available on cran), \pkg{rlecuyer} \citep{sevcikova2015package} and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
 More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG).
-\pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented (?).
+\pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented.
 For up-to-date review papers on the generation of random numbers on parallel devices, and GPUs in particular, see:  \cite{rLEC15a,rLEC17p,rLEC21a}.
@@
 The remaining sections are organized as follows:
-In Section 2, we introduce streams and the use of streams in work-items on GPU device, and describe and check the functions for generating uniform and normal random numbers.
-In Section 3, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we parallelise this Monte Carlo simulation and implement it on GPU, then we provide two real data examples to demonstrate the function usage and test its R performance.
-Section 4 simulates simultaneously multiple Gaussian random surfaces by applying GPU-generated Normal random numbers and GPU-computed batches of Mat\'ern covariance matrices, then we plot the generated random surfaces in the end.
-Finally, the paper concludes with a short summary and a discussion in Section 5.
+In Section 2, we introduce streams and the use of streams in work-items on a GPU device for generating uniform random numbers, and the usage of \fct{gpuRandom::runif}.
+In Section 3, we introduce two non-uniform RNGs in \pkg{gpuRandom}, as examples for users to develop other RNGs of interest on GPUs in R.
+In Section 4, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we introduce how the random numbers are used and how the algorithm is parallelized and implemented on GPU.
+Then we provide two real data examples to demonstrate the function usage and its R performance.
+Section 5, we show an useful application of normal random numbers on GPU, using them to simulate batches of Gaussian random surfaces with Mat\'ern covariance matrices simultaneously, the simulation also uses GPU-enabled functions from our another package \pkg{gpuBatchMatrix}.
+Finally, the paper concludes with a short summary and a discussion in Section 6.
 
@@
 It is also faster than Philox-4×32-10 \citep{rLEC00b}.
-MRG31k3p was also statistically tested extensively and successfully in \citep{rLEC07b}.
+MRG31k3p was also statistically tested extensively and successfully, \cite[see][]{rLEC07b}.
 
@@
 
-\subsection{Creating streams}\label{create_streams}
+\subsection{Creating streams}\label{createstreams}
 The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
@@
   \item \code{n}: number of streams to create.
-  \item \code{seed}: vector of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
+  \item \code{initial}: vector of length 6, recycled if shorter.
 \end{itemize}
+We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}.
 \fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
-We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}.
-Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
+Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we produce on device exactly the same streams as those created on host as follows.
+Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
 <<createStreams_Gpu>>=
@@
 @
-\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''.
+\fct{createStreamsGpu} also has the two arguments \code{n} and \code{initial}.
 \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
 By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.
+Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{initial} argument, we make the sequence of streams repeatable.
 
@@
 The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$, if to generate double-precision random numbers.
-(ADD THINGS ABOUT RANDOM INTEGERS) At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
+At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel},caption={Using streams to produce random U(0,1) numbers}]
  __kernel void mrg31k3pMatrix(
@@
 For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
-\begin{figure}[H]%
+\begin{figure}[H]
     \centering
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f1} }}%
+    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f1} }}
     \qquad
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f2} }}%
-    \caption{name?\label{fig:1}}%
+    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f2} }}
+    \caption{Using streams to produce a $4 \times 6$ matrix of random U(0,1) numbers. \label{fig:1}}
 \end{figure}
 
-Now we use the streams created in section \ref{create_streams} to generate two vectors of double-precision i.i.d.
+Now we use the streams created in section \ref{createstreams} to generate two vectors of double-precision i.i.d.
 $U(0,1)$ random numbers with the function \fct{runif}.
+\fct{runif} can also be used to produce random integers in $[1, 2147483647]$.
 To view the generated random numbers, we need to convert them to R vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
@@
 \end{itemize}
-\fct{runif} returns the same sequence of random numbers if it uses the same streams (here \code{myStreamsGpu1} and \code{myStreamsGpu2}), not streams starting from the same states.
+\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu1} and \code{myStreamsGpu2} here.
 Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
@@
 
-\subsection{Plot the uniform random numbers}
-We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
-The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
-<<uniforms_random_vector,eval=TRUE>>=
-streams <- createStreamsGpu(n = 512)
-random_u<-as.vector(runif(n=50000, streams=streams, Nglobal=c(64,8), type="double"))
-@
-\begin{figure}[H]
-\centering
-<<largenumberuniforms, eval=FALSE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold', dev='png'>>=
-hist(random_u, prob = TRUE, main="Histogram")
-plot(ecdf(random_u), ylab="F(x)",main="Empirical cumulative distribution")
-curve(punif(x, 0, 1), lty=2, add=TRUE, col="red")
-@
-\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
-\end{figure}
+\subsection{Plot the uniform random numbers (now deleted?)}
 
-\section{Normal random number generation}
-We apply the Box-Muller transformation to $U(0,1)$ random numbers to generate standard Gaussian random numbers.
+\section{Some non-uniform random number generation}
+\subsection{Normal random number generation}
+We apply the Box-Muller transformation to $U(0,1)$ random numbers to generate standard normal random numbers.
 As shown in Algorithm \ref{algorithm1}, Box-Muller algorithm takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables.
@@
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},caption={Using streams to produce standard normal random numbers}]
 const int 
@@
     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
-    \caption{name???}%
+    \caption{Using streams to produce a $4 \times 6$ matrix of normal random numbers.}%
     \label{fig3}
@@
 We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
-Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $512 \times 128$ work-items is around 120 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}.
+The best performance we have seen using \fct{gpuRandom::rnorm} is around 120 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size goes larger.
@@
 
-\section{Monte Carlo simulation for Fisher's exact test}
+\subsection{Exponential random number generation}
+The Exponential random variates are produced by applying the inverse transform method on i.i.d.
+$U(0,1)$ random numbers.
+The random variable $X \sim \text{Exponential}(\lambda)$ has cumulative distribution function $F_{X}(x)=1-e^{-\lambda x}$ for $x \geq 0$ and $\lambda > 0$.
+The inverse of $F^{-1}_{X}$ is $F^{-1}_{x}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
+The kernel is not shown because it is mostly same as the kernel for uniform and normal random numbers, basically just by applying the inverse transform formula to the \code{clrngMrg31k3pNextState(g1, g2)} in the kernel.
+Developers can make use of our kernel for generating other types of random variables on GPU in \proglang{R}.
+
+Below is the code that we produce a 2 by 2 matrix of exponential random number with expectation 1.
+<<random exponential, eval=TRUE, cache=FALSE>>=
+r_matrix <- gpuRandom::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
+as.matrix(r_matrix)
+@
+The arguments of the \fct{gpuRandom::rexp} are described as follows:
+\begin{itemize}
+\itemsep0em 
+  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
+  \item \code{streams}: streams. If missing, a stream with random initial state is supplied.
+  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
+  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
+\end{itemize}
+
+\section{An application of uniform RNG: Fisher's simulation}
+The GPU-generated random numbers can be applied in suitable statistical simulations to accelerate the performance.
 One application of GPU-generated random numbers in \pkg{gpuRandom} is Monte Carlo simulation for Fisher's exact test.
@@
 \itemsep0em 
-  \item \code{x}: a vclMatrix.
+  \item \code{x}: a ``vclMatrix''.
   \item \code{N}: number of simulation runs.
   \item \code{streams}: streams.
-  \item \code{type}: ``double'' or ``float'' or returned test statistics.
+  \item \code{type}: ``double'' or ``float'' of returned test statistics.
   \item \code{returnStatistics}: if TRUE, return all test statistics. 
@@
 
-\subsection{Comparing run-time: Month data example}
+\subsection{Comparing run-time: Month data example}\label{fisher_month}
  <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
@@
  colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
-Nglobal = c(256,128)
-streams <- createStreamsGpu(n =256*128)
+ streams <- createStreamsGpu(n =256*64, initial=666)
  @
 
-<<timecomparemonth3Gpu, eval=TRUE,cache=FALSE>>=
+ <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE>>=
  almost.1 <- 1 + 64 * .Machine$double.eps
@@
  result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
+                    type="double", returnStatistics=TRUE,  Nglobal = c(256,64))
  result_month$simNum
@@
  @
-We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
+
  <<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
@@
  system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
+             type="double", returnStatistics=FALSE, Nglobal = c(256,64)))
  @
-\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU. 
-<<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
+We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
+\fct{gpuRandom::fisher.sim} takes about 0.69 seconds, \pkg{gpuRandom} accounts for 4.64\% of the elapsed time on CPU.
+ <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
  month_stats <- as.vector(result_month$sim)
@@
 
-<<timecomparemonth2Cpu, cache=TRUE, eval=TRUE>>=
+ <<timecomparemonth4Cpu, cache=TRUE, eval=TRUE>>=
  #using CPU
@@
  @
-\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for 1015808 simulations, has a p-value of 0.4035.
 
-\subsection{Comparing run-time: Week data example}
+\subsection{Comparing run-time: Week data example}\label{fisher_week}
 <<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
@@
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
+                type="double",returnStatistics=TRUE,Nglobal = c(256,64))
 result_weekgpu$simNum
@@
 @
-With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
+
 <<TimecompareweekGpu, eval=TRUE, cache=FALSE>>=
@@
 system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
+            type="double", returnStatistics=FALSE,Nglobal = c(256,64)))
 @
-\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
+The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value.
+With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}.
+\fct{stats::fisher.test} takes 91.6 seconds, while \fct{fisher.sim} takes about 4.282 seconds, the elapsed time is decreased to about 4.67\% of the time taken on CPU.
 <<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
@@
 @
-The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
 
 \subsection{A summary of the results}
-(Results obtained in first week of October)
+We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
 <<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE>>=
@@
   var1 = c("B", "1m", "10m", "1m", "10m"),
-  var2 = c('Intel 2.5ghz', 0.403804, 	0.0001251, 9.74, 60.19),
+  var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 10.74,  63.24),
   var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.28,    10.82),
-  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202,14.58,90.38),
-  var5 = c('NVIDIA V100', 0.403507,0.0001274,0.32, 1.89),
+  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 15.00,  91.58),
+  var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.72,   4.18),
   var6 = c('Data','month','week','month','week')
@@
 @
-\caption{Frequency distribution for test statistics from month and week data tables.
-Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example.
-The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
+\caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''.
+The test statistic values of the observed tables are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
 \label{fig4}}
@@
 
-\section{Gaussian surface simulation} 
+\section{An application of normal RNG: Gaussian surface simulation} 
 Given a parameter space $X$, a random field $U$ on $X$ is a collection of random variables $\{U_x :x\in X\}$.
@@
 \centering
-<<maternplot, echo=TRUE, fig.show='hold', eval=TRUE, dev='pdf',cache=FALSE,dev='png'>>=
+<<maternplot, echo=TRUE,eval=TRUE, fig.show='hold', dev='pdf',cache=FALSE,dev='png'>>=
 par(mar=rep(0.1, 4))
@@
 @
-\caption{\label{fig:5} }
+\caption{Simulated Gaussian surface\label{fig:5} }
 \end{figure}
@@
 
-talk about using substreams in future?
-
 \pkg{gpuRandom} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
@@
 \begin{appendix}
-TO BE PUT IN APPENDIX OR DELETE? \\
-In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
+The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
+<<uniforms_random_vector,include=FALSE>>=
+streams <- createStreamsGpu(n = 512)
+random_u<-as.vector(runif(n=50000, streams=streams, Nglobal=c(64,8), type="double"))
+@
+%\begin{figure}[H]
+%\centering
+<<largenumberuniforms, include=FALSE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold', dev='png'>>=
+hist(random_u, prob = TRUE, main="Histogram")
+plot(ecdf(random_u), ylab="F(x)",main="Empirical cumulative distribution")
+curve(punif(x, 0, 1), lty=2, add=TRUE, col="red")
+@
+%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
+%\end{figure}
+
+%In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
 We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-<<histAndQQ_prepare,eval=FALSE,cache=TRUE>>=
+<<histAndQQ_prepare,include=FALSE,cache=TRUE>>=
 streams <- createStreamsGpu(n =512*128)
@@
 @
-\begin{figure}[H]
-\centering
- <<histAndQQ, echo=TRUE, eval=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
+% \begin{figure}[H]
+% \centering
+ <<histAndQQ, echo=TRUE, include=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
 hist(avector,breaks=40, main = "histogram")
@@
 @
-\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
-\end{figure}
+% \caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
+% \end{figure}
 
-The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
+The ``month'' and ``week'' tables for testing \fct{fisher.sim} in section \ref{fisher_month} and \ref{fisher_week} are shown below, where 
+The row variable of the ``month'' table represents the twelve months: from January to December, and for the ``week'' table it represents the seven weekdays: Monday to Sunday.
 The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.

### 2021-11-06 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
modify rexp kernel
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2948 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \Address{
-  Ruoyong Xu\\
+  Ruoyong Xu, Patrick Brown\\
   Department of Statistics\\
   University of Toronto\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}
+  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
+  E-mail: \email{ruoyong.xu@mail.utoronto.ca}, \email{patrick.brown@utoronto.ca}}
 
 \Address{
-  Achim Zeileis\\
-  Journal of Statistical Software\\
-  \emph{and}\\
-  Department of Statistics\\
-  University of Toronto\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}
+  Pierre L’Ecuyer\\
+  Department of Computer Science and Operations Research\\
+  University of Montr\'eal\\
+  Pavillon André-Aisenstadt 2920, chemin de la Tour Montréal QCH3T 1J4\\
+  E-mail: \email{lecuyer@iro.umontreal.ca}}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
   \item \code{n}: number of streams to create.
-  \item \code{seed}: vector of length 6, recycled if shorter. Default is set to a vector of six ``12345''.
+  \item \code{initial}: vector of length 6, recycled if shorter.
 \end{itemize}
+We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}.
 \fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
-We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}.
-Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
+Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we produce on device exactly the same streams as those created on host as follows.
+Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
 \begin{knitrout}
@@
 \end{knitrout}
-\fct{createStreamsGpu} also has the two arguments \code{n} and \code{seed}, \code{seed} is by default a vector of six ``12345''.
+\fct{createStreamsGpu} also has the two arguments \code{n} and \code{initial}.
 \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
 By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{seed} argument, we make the sequence of streams repeatable.
+Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{initial} argument, we make the sequence of streams repeatable.
 
@@
 
-\subsection{Plot the uniform random numbers}
-We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
+\subsection{Plot the uniform random numbers (now deleted?)}
+%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
 The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streams <- createStreamsGpu(n = 512)
-random_u<-as.vector(runif(n=50000, streams=streams, Nglobal=c(64,8), type="double"))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{figure}[H]
-\centering
-
-\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
-\end{figure}
-
-\section{Normal random number generation}
-We apply the Box-Muller transformation to $U(0,1)$ random numbers to generate standard Gaussian random numbers.
+
+%\begin{figure}[H]
+%\centering
+
+%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
+%\end{figure}
+
+\section{Some non-uniform random number generation}
+\subsection{Normal random number generation}
+We apply the Box-Muller transformation to $U(0,1)$ random numbers to generate standard normal random numbers.
 As shown in Algorithm \ref{algorithm1}, Box-Muller algorithm takes two independent, standard uniform random variables $U_1$ and $U_2$ and produces two independent, standard Gaussian random variables $X$ and $Y$, where $R$ and $\Theta$ are polar coordinate random variables.
@@
 ##    user  system elapsed 
-##   0.133   0.000   0.133
+##   0.142   0.003   0.145
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   6.608   0.796   7.399
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\section{Monte Carlo simulation for Fisher's exact test}
+##   6.689   0.830   7.513
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\subsection{Exponential random number generation}
+The Exponential random variates are produced by applying the inverse transform method on i.i.d.
+$U(0,1)$ random numbers.
+The random variable $X \sim \text{Exponential}(\lambda)$ has cumulative distribution function $F_{X}(x)=1-e^{-\lambda x}$ for $x \geq 0$ and $\lambda > 0$.
+The inverse of $F^{-1}_{X}$ is $F^{-1}_{x}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
+The kernel is not shown because it is mostly same as the kernel for uniform and normal random numbers, basically just by applying the inverse transform formula to the \code{clrngMrg31k3pNextState(g1, g2)} in the kernel.
+Developers can make use of our kernel for generating other types of random variables on GPU in \proglang{R}.
+
+Below is the code that we produce a 2 by 2 matrix of exponential random number with expectation 1.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+r_matrix <- gpuRandom::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
+as.matrix(r_matrix)
+##           [,1]      [,2]
+## [1,] 0.9260462 0.1431169
+## [2,] 0.6722247 0.8337073
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+The arguments of the \fct{gpuRandom::rexp} are described as follows:
+\begin{itemize}
+\itemsep0em 
+  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
+  \item \code{streams}: streams. If missing, a stream with random initial state is supplied.
+  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
+  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
+\end{itemize}
+
+\section{An application of uniform RNG: Fisher's simulation}
+The GPU-generated random numbers can be applied in suitable statistical simulations to accelerate the performance.
 One application of GPU-generated random numbers in \pkg{gpuRandom} is Monte Carlo simulation for Fisher's exact test.
@@
 
-\subsection{Comparing run-time: Month data example}
+\subsection{Comparing run-time: Month data example}\label{fisher_month}
 
@@
 result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
+                   type="double", returnStatistics=TRUE,  Nglobal = c(256,64))
 result_month$simNum
@@
 result_month$counts
-## [1] 410533
+## [1] 410825
 result_month$p.value
-## [1] 0.4041449
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.
+## [1] 0.4044323
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
 \begin{knitrout}
@@
 system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
+            type="double", returnStatistics=FALSE, Nglobal = c(256,64)))
 ##    user  system elapsed 
-##   0.702   0.005   0.707
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.
+##   0.739   0.004   0.742
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
+\fct{gpuRandom::fisher.sim} takes about 0.69 seconds, \pkg{gpuRandom} accounts for 4.64\% of the elapsed time on CPU.
 
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4029291
+## [1] 0.4044816
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  14.762   0.003  14.757
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for 1015808 simulations, has a p-value of 0.4035.
-
-\subsection{Comparing run-time: Week data example}
+##  15.006   0.009  15.002
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\subsection{Comparing run-time: Week data example}\label{fisher_week}
 
@@
 result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
+                type="double",returnStatistics=TRUE,Nglobal = c(256,64))
 result_weekgpu$simNum
-## [1] 10027008
+## [1] 10010624
 result_weekgpu$counts
-## [1] 1259
+## [1] 1205
 result_weekgpu$p.value
-## [1] 0.0001256606
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
+## [1] 0.000120472
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
 \begin{knitrout}
@@
 system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
+            type="double", returnStatistics=FALSE,Nglobal = c(256,64)))
 ##    user  system elapsed 
-##   4.324   0.011   4.332
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.
+##   4.524   0.009   4.530
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value.
+With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}.
+\fct{stats::fisher.test} takes 91.6 seconds, while \fct{fisher.sim} takes about 4.282 seconds, the elapsed time is decreased to about 4.67\% of the time taken on CPU.
 
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001209715
+## [1] 0.0001249672
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  90.139   0.098  90.181
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
+##  91.498   0.128  91.580
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 \subsection{A summary of the results}
-(Results obtained in first week of October)
+We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
 \begin{knitrout}
@@
 \hline
-\hspace{1em}1m & 9.74 & 2.28 & 14.58 & 0.32 & month\\
-\hline
-\hspace{1em}10m & 60.19 & 10.82 & 90.38 & 1.89 & week\\
+\hspace{1em}1m & 10.74 & 2.28 & 15 & 0.72 & month\\
+\hline
+\hspace{1em}10m & 63.24 & 10.82 & 91.58 & 4.18 & week\\
 \hline
@@
 \end{knitrout}
-\caption{Frequency distribution for test statistics from month and week data tables.
-Figure \ref{fig4} visualizes the approximate sampling distributions of the test statistics from the two examples "Month" and "Week", by plotting a histogram for each of the example.
-The values of observed statistics are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
+\caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''.
+The test statistic values of the observed tables are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
 \label{fig4}}
@@
 \end{knitrout}
-\caption{\label{fig:5} }
+\caption{Simulated Gaussian surface\label{fig:5} }
 \end{figure}
@@
 \begin{appendix}
-TO BE PUT IN APPENDIX OR DELETE? \\
-In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+%In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
 We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streams <- createStreamsGpu(n =512*128)
-normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
-                                Nglobal=c(512,128), type="float")
-avector<-as.vector(as.matrix(normal_matrix))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{figure}[H]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-hist(avector,breaks=40, main = "histogram")
-gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
-\end{figure}
-
-The row variable of the first table represents the twelve months: from January to December, and for the second table it represents the seven weekdays: Monday to Sunday.
+
+% \begin{figure}[H]
+% \centering
+
+% \caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
+% \end{figure}
+
+The ``month'' and ``week'' tables for testing \fct{fisher.sim} in section \ref{fisher_month} and \ref{fisher_week} are shown below, where 
+The row variable of the ``month'' table represents the twelve months: from January to December, and for the ``week'' table it represents the seven weekdays: Monday to Sunday.
 The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.

### 2021-11-06 pkg/gpuRandom/R/rexp.R

@@ -16,7 +16,7 @@
 
 rexp = function(
   n, 
-  rate,
+  rate=1,
   streams, 
   Nglobal,
   type=c("double","float")) {
@@ -32,6 +32,14 @@ rexp = function(
     n<-c(n,1)
   }
 
+  if(Nglobal[2]<2){
+    stop("Nglobal[2] should be larger than 1")
+  }
+
+  if(rate <= 0 || !is.finite(rate)){
+    stop("invalid rate value")
+  }
+  
   
   if(missing(streams)) {
     if(missing(Nglobal)) {
@@ -58,13 +66,13 @@ rexp = function(
   xVcl<-gpuR::vclMatrix(0, nrow=n[1], ncol=n[2], type=type[1])     
   
   
-  gpuRnBackend(xVcl,streams,rate, Nglobal,"exponential") 
+  gpuRnBackend(xVcl,streams, Nglobal,"exponential") 
   
   invisible(streams)
   
   if(ncol(xVcl)==1) xVcl = xVcl[,1]
   
-  xVcl
+  rate * xVcl
   
 }
 

### 2021-11-06 pkg/gpuRandom/R/rnorm.R

@@ -30,7 +30,9 @@ rnorm = function(
     n<-c(n,1)
   }
   
-  #x<-matrix(0,nrow=n[1],ncol=n[2])
+  if(Nglobal[2]<2){
+    stop("Nglobal[2] should be larger than 1")
+  }
   
   if(missing(streams)) {
     if(missing(Nglobal)) {
@@ -57,7 +59,7 @@ rnorm = function(
   xVcl<-gpuR::vclMatrix(0, nrow=n[1], ncol=n[2], type=type[1])     
   
   
-  gpuRnBackend(xVcl,streams,Nglobal,"normal") 
+  gpuRnBackend(xVcl, streams, Nglobal,"normal") 
   
   invisible(streams)
   

### 2021-11-06 pkg/gpuRandom/src/gpuRnMatrix.cpp

@@ -1,7 +1,7 @@
 #include "gpuRandom.hpp"
 
 //#define DEBUG
-//#define DEBUGKERNEL
+#define DEBUGKERNEL
 
 
 
@@ -18,8 +18,7 @@ std::string mrg31k3pMatrixString(
     const int Ncol,
     const int NpadCol,
     const int NpadStreams,
-    const std::string random_type,
-    T lambda) { 
+    const std::string random_type) { 
   
   std::string typeString = openclTypeString<T>();
   std::string result = "";
@@ -55,10 +54,32 @@ std::string mrg31k3pMatrixString(
     "\n#define Nrow " + std::to_string(Nrow) + "\n"    
     "#define Ncol " + std::to_string(Ncol) + "\n"
     "#define NpadStreams " + std::to_string(NpadStreams) + "\n"
-    "#define NpadCol " + std::to_string(NpadCol) + "\n"
-    "#define lambda " + std::to_string(lambda) + "\n";    
+    "#define NpadCol " + std::to_string(NpadCol) + "\n";    
   
     result += mrg31k3pString();
+  
+  if(random_type == "exponential"){
+    result += 
+      "__global double q[]= {\n"
+      " 0.6931471805599453,\n"
+      " 0.9333736875190459,\n"
+      " 0.9888777961838675,\n"
+      " 0.9984959252914960,\n"
+      " 0.9998292811061389,\n"
+      " 0.9999833164100727,\n"
+      " 0.9999985691438767,\n"
+      " 0.9999998906925558,\n"
+      " 0.9999999924734159,\n"
+      " 0.9999999995283275,\n"
+      " 0.9999999999728814,\n"
+      " 0.9999999999985598,\n"
+      " 0.9999999999999289,\n"
+      " 0.9999999999999968,\n"
+      " 0.9999999999999999,\n"
+      " 1.0000000000000000"
+      "};\n";
+  }
+  
   
   result += 
     "\n\n__kernel void mrg31k3pMatrix(\n"
@@ -83,6 +104,7 @@ std::string mrg31k3pMatrixString(
       "local " + typeString + "  part[2], cosPart1, sinPart1;\n";// local size must be 1,2
   }
   
+
   
   /*result +=  " for(Drow = 0, DrowStart = startvalue, Dcol = DrowStart + 3;\n"
[…]

### 2021-11-06 pkg/gpuRandom/src/RcppExports.cpp
modify rexp kernel
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2948 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -126,17 +126,16 @@ BEGIN_RCPP
 END_RCPP
 }
 // gpuRnBackend
-SEXP gpuRnBackend(Rcpp::S4 x, Rcpp::S4 streams, SEXP lambdaR, IntegerVector max_global_size, std::string random_type);
-RcppExport SEXP _gpuRandom_gpuRnBackend(SEXP xSEXP, SEXP streamsSEXP, SEXP lambdaRSEXP, SEXP max_global_sizeSEXP, SEXP random_typeSEXP) {
+SEXP gpuRnBackend(Rcpp::S4 x, Rcpp::S4 streams, IntegerVector max_global_size, std::string random_type);
+RcppExport SEXP _gpuRandom_gpuRnBackend(SEXP xSEXP, SEXP streamsSEXP, SEXP max_global_sizeSEXP, SEXP random_typeSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< Rcpp::S4 >::type x(xSEXP);
     Rcpp::traits::input_parameter< Rcpp::S4 >::type streams(streamsSEXP);
-    Rcpp::traits::input_parameter< SEXP >::type lambdaR(lambdaRSEXP);
     Rcpp::traits::input_parameter< IntegerVector >::type max_global_size(max_global_sizeSEXP);
     Rcpp::traits::input_parameter< std::string >::type random_type(random_typeSEXP);
-    rcpp_result_gen = Rcpp::wrap(gpuRnBackend(x, streams, lambdaR, max_global_size, random_type));
+    rcpp_result_gen = Rcpp::wrap(gpuRnBackend(x, streams, max_global_size, random_type));
     return rcpp_result_gen;
 END_RCPP
 }
@@ -341,7 +340,7 @@ static const R_CallMethodDef CallEntries[] = {
     {"_gpuRandom_gemmBatchBackend", (DL_FUNC) &_gpuRandom_gemmBatchBackend, 9},
     {"_gpuRandom_gemmBatch2backend", (DL_FUNC) &_gpuRandom_gemmBatch2backend, 11},
     {"_gpuRandom_cpp_gpuFisher_test", (DL_FUNC) &_gpuRandom_cpp_gpuFisher_test, 6},
[…]

### 2021-11-08 pkg/gpuRandom/src/likfitGpu_P.cpp
fixed crossprod, there is still a concern
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2951 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -424,7 +424,7 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
   int NlocalCacheC = NcolsPerGroup * NrowsToCache;
   
   if(NrowsToCache < 0) {
-    Rcpp::warning("NlocaalCache too small for this number of work groups");
+    Rcpp::warning("NlocalCache too small for this number of work groups");
   }
   
   std::string backsolveString = backsolveBatchString<T>(
@@ -480,6 +480,10 @@ void likfitGpuP(viennacl::matrix_base<T> &yx,
   
   
   const int NlocalCacheAD = std::max( 0, (NlocalCache[0] - localSize[0]*localSize[1])/2 );
+  
+  if((NlocalCache[0] - localSize[0]*localSize[1]) < 0) {
+    Rcpp::warning("NlocalCache too small for this number of work groups");
+  }
   
   std::string crossprodKernelString = crossprodBatchString<T>(
     Nobs,//const int Nrow, 

### 2021-11-09 pkg/gpuRandom/src/backsolve.cpp
fixing random plot problem
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2952 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -611,7 +611,7 @@ void backsolveBatch(
     NcolsPerGroup,  // 15
     NlocalCacheC,   // 16
     NlocalCacheSum, // 17
-    Nlocal[0] * Nlocal[1] //NpadBetweenMatricesSum   18
+    Nlocal[0] * Nlocal[1] //  18
   );  //NcolsInCache
   
   if(verbose > 2) {

### 2021-11-09 pkg/gpuRandom/src/cholBatch.cpp

@@ -44,7 +44,7 @@ template <typename T> std::string cholBatchKernelString(
   result += "\n__kernel void cholBatch(\n"
   "	__global " + typeString + " *A,\n" 
   "	__global " + typeString + " *diag,\n"
-  " __local " + typeString + "*diagLocal,\n"
+  " __local " + typeString + " *diagLocal,\n"
   "              int Nmatrix";
   
   if(logDet){

### 2021-11-10 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2954 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \Abstract{
-We introduce the \proglang{R} package \pkg{gpuRandom} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by utilizing the \pkg{clRNG} (OpenCL) library.
+We introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by utilizing the \pkg{clRNG} (OpenCL) library.
 Parallel processing by graphics processing unit (GPU) can be used to speed up computationally intensive tasks, which when combined with \proglang{R}, it can largely improve \proglang{R}’s downsides in terms of slow speed, memory usage and computation mode.
 There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling.
-\pkg{gpuRandom} enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
+\pkg{clrng} enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
 This package is portable and flexible, developers can use its random number generation kernel for various other purposes and applications.
@@
 
-\Keywords{GPU, \pkg{gpuRandom} package, parallel computing, \pkg{clRNG} library}
-\Plainkeywords{GPU, gpuRandom package, parallel computing, clRNG library}
+\Keywords{GPU, \pkg{clrng} package, parallel computing, \pkg{clRNG} library}
+\Plainkeywords{GPU, clrng package, parallel computing, clRNG library}
 
@@
 There is not yet an \proglang{R} package that is able to generate random numbers on GPU, this is complicated to do because multiple streams will be needed for parallel generation of random numbers, and the stream states need to be able to be restored for reproducibility of results, which is essential in many simulation applications.
-We introduce the \proglang{R} package \pkg{gpuRandom} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
+We introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
 
 The remaining sections are organized as follows:
-In Section 2, we introduce streams and the use of streams in work-items on a GPU device for generating uniform random numbers, and the usage of \fct{gpuRandom::runif}.
-In Section 3, we introduce two non-uniform RNGs in \pkg{gpuRandom}, as examples for users to develop other RNGs of interest on GPUs in R.
+In Section 2, we introduce streams and the use of streams in work-items on a GPU device for generating uniform random numbers, and the usage of \fct{clrng::runif}.
+In Section 3, we introduce two non-uniform RNGs in \pkg{clrng}, as examples for users to develop other RNGs of interest on GPUs in R.
 In Section 4, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we introduce how the random numbers are used and how the algorithm is parallelized and implemented on GPU.
@@
 library("gpuR")
-library("gpuRandom")
+library("clrng")
 setContext(   grep('gpu', listContexts()$device_type) [1]    )
-#setwd("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5")
+#setwd("/home/ruoyong/diseasemapping/pkg/clrng/inst/documents/paper1_2021_5")
 @
@@
 These four RNGs use different types of constructions.
-\pkg{gpuRandom} package uses the MRG31k3p RNG, making it able to generate random numbers on GPUs.
+\pkg{clrng} package uses the MRG31k3p RNG, making it able to generate random numbers on GPUs.
 We choose MRG31k3p as the base generator for the following reasons:  the original \pkg{RNGStreams} package \citep{l2002object} was built with MRG32k3a, which was designed to be implemented in ``double'' in floating point(?), and not with 32-bit integers.
@@
 If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
-\pkg{gpuRandom} is able to create streams both on the host and on the GPU device.
+\pkg{clrng} is able to create streams both on the host and on the GPU device.
 \fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
-Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item.
+Substreams are not created in \pkg{clrng} because currently we use one stream per work-item.
 <<createStreams_CPU>>=
@@
 library("gpuR")
-library("gpuRandom")
+library("clrng")
 myStreamsCpu <- createStreamsCpu(n=4, initial=12345)
@@
 The streams created sequentially on a GPU are then used in work-items that executes in parallel on a GPU device.
-In \pkg{gpuRandom} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be equal to the number of total work-items in use.
+In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be equal to the number of total work-items in use.
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
@@
 <<simStuff2, eval=TRUE,tidy=TRUE>>=
-as.vector(gpuRandom::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
+as.vector(clrng::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
 
-as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
+as.vector(clrng::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
 @
-The arguments of the \fct{gpuRandom::runif} are described as follows:
+The arguments of the \fct{clrng::runif} are described as follows:
 \begin{itemize}
@@
 
-We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{gpuRandom::rnorm}.
-The best performance we have seen using \fct{gpuRandom::rnorm} is around 120 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
+We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{clrng::rnorm}.
+The best performance we have seen using \fct{clrng::rnorm} is around 120 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size goes larger.
@@
 streams <- createStreamsGpu(n = 512 * 128)
-system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
+system.time(clrng::rnorm(c(10000,10000), streams=streams, 
                              Nglobal=c(512,128), type="double"))
@@
 <<random exponential, eval=TRUE, cache=FALSE>>=
-r_matrix <- gpuRandom::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
+r_matrix <- clrng::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
 as.matrix(r_matrix)
 @
-The arguments of the \fct{gpuRandom::rexp} are described as follows:
+The arguments of the \fct{clrng::rexp} are described as follows:
 \begin{itemize}
@@
 The GPU-generated random numbers can be applied in suitable statistical simulations to accelerate the performance.
-One application of GPU-generated random numbers in \pkg{gpuRandom} is Monte Carlo simulation for Fisher's exact test.
+One application of GPU-generated random numbers in \pkg{clrng} is Monte Carlo simulation for Fisher's exact test.
 Fisher’s exact test is applied for analyzing usually $2 \times 2$ contingency tables when one of the expected values in table is less than 5.
@@
 The test statistic calculated for each random table is $-\sum_{i,j}\log(n_{ij}!), i=(1,\dots, I), j=(1,\dots,J)$, (i.e., minus log-factorial of table), where $I$ and $J$ are the row and column number of the observed table.
-This test statistic can also be independently calculated for a table by \fct{gpuRandom::logfactSum}.
+This test statistic can also be independently calculated for a table by \fct{clrng::logfactSum}.
 Given an observed table and a number of replicates $B$, the Monte Carlo simulation for Fisher's exact test does the following steps:
@@
 Step 1 and step 2-3 are done on a GPU with two kernels enqueued sequentially.
-For step 2, \fct{gpuRandom::fisher.sim} adapted the function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random two-way tables with given marginal totals.
+For step 2, \fct{clrng::fisher.sim} adapted the function \fct{rcont2} used by \fct{stats::fisher.test} for constructing random two-way tables with given marginal totals.
 The algorithm \citep[see][]{patefield1981efficient} samples the entries row by row, one at a time, conditional on the values of the entries already sampled.
@@
 On a GPU, this step can be replicated say $n$ times in parallel by creating $n$ distinct random streams and launching $n$ separate work-items.
-Each work-item takes a random stream as input, performs all of Step 2 and 3, and returns the value of the test statistic and count on the GPU.
+Each work-item takes a random stream as input, performs all of Step 2 and 3, and returns the value of the test statistic on the GPU.
 Computing the p-value on the CPU (Step 4) is then a trivial operation.
 Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
-%There are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo.
-The use of streams and GPU would be quite different in that case.
-See for example
+By the way, there are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo, the use of streams and GPU would be quite different in that case.
+See for examples \citep{10.1214/13-AOS1131, KAYIBI2018298, dyer1997sampling}.
+Doing the \fct{fisher.sim} function on GPU opens up many possibilities for future work, the specific implementation we've done for the tables isn't necessarily the optimal one.
 
-The arguments of the \fct{gpuRandom::fisher.sim} are described as follows:
+The arguments of the \fct{clrng::fisher.sim} are described as follows:
 \begin{itemize}
@@
 \end{itemize}
-We show the advantage of \fct{gpuRandom::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
+We show the advantage of \fct{clrng::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
@@
  @
-We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
-\fct{gpuRandom::fisher.sim} takes about 0.69 seconds, \pkg{gpuRandom} accounts for 4.64\% of the elapsed time on CPU.
+We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
+\fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
  <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
@@
 <<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
-observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
+observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
+result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                 type="double",returnStatistics=TRUE,Nglobal = c(256,64))
@@
 #using GPU
-system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
+system.time(clrng::fisher.sim(week_GPU, 1e7, streams=streams,
             type="double", returnStatistics=FALSE,Nglobal = c(256,64)))
@@
 \caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''.
-The test statistic values of the observed tables are calculated by \fct{gpuRandom::logfactSum} and are marked with a blue line on each plot.
+The test statistic values of the observed tables are calculated by \fct{clrng::logfactSum} and are marked with a blue line on each plot.
 \label{fig4}}
@@
 There are several alternative parameterisations of Mat\'ern covariance functions in literatures  \citep[see][]{haskard2007anisotropic}.
-In \pkg{gpuRandom}, we use the above form for computing Mat\'ern covariance.
+In \pkg{clrng}, we use the above form for computing Mat\'ern covariance.
 
@@
 Other \proglang{R} packages that offer simulation of Gaussian random fields like \pkg{geoR} \citep{geoR2001}, does not work for large number of locations; the \pkg{RandomFields} \citep{RandomFields2015,RandomFields2020} package can use different methods for simulation of Gaussian fields, among which the (modified) circulant embedding method \citep{Dietrich1997FastAE} for covariance matrix decomposition is also an exact method (need to confirm???), however, it works only for isotropic Gaussian fields and on rectangular grids.
-\pkg{gpuRandom} does exact simulation of exact Gaussian fields on the GPU as it relies on the matrix decomposition method.
+\pkg{clrng} does exact simulation of exact Gaussian fields on the GPU as it relies on the matrix decomposition method.
 The implementation of this method is as follows.
@@
 The matrix decomposition method is straightforward to implement, however, when a large number of locations ($n$) is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
-\pkg{gpuRandom} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
+\pkg{clrng} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
 The following shows an example, in which we simulate 10 Gaussian random fields of mat\'ern covariance with 5 sets of parameters at one time.
@@
 library("gpuR")
-library("gpuRandom")
+library("clrng")
+library("gpuBatchMatrix")
 library('geostatsp')
@@
 Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets.
-Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuRandom::maternGpuParam}.
+Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuBatchMatrix::maternGpuParam}.
 <<setup paramsBatch, eval=TRUE,echo=FALSE>>=
@@
 myParamsBatch
-paramsGpu = gpuRandom::maternGpuParam(myParamsBatch, type=theType)
+paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
 @
-Step 3, compute Mat\'ern covariance matrices using \fct{gpuRandom::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
+Step 3, compute Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
 are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
@@
 dim(maternCov)
-gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
+gpuBatchMatrix::maternBatch(maternCov, coordsGpu, paramsGpu,
                        Nglobal=c(128,64), Nlocal=c(16,4))
 @
-Step 4, the first argument \code{maternCov} in \fct{gpuRandom::cholBatch} specifies the matrix object to take Cholesky decomposition.
+Step 4, the first argument \code{maternCov} in \fct{gpuBatchMatrix::cholBatch} specifies the matrix object to take Cholesky decomposition.
 Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
@@
 diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
+gpuBatchMatrix::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
                      Nglobal= c(128, 8),
@@
 % \end{gather}
-Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{gpuRandom::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 <<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
@@
 @
-Step 6, use \fct{gpuRandom::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
+Step 6, use \fct{gpuBatchMatrix::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
 \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
@@
 
-gpuRandom::multiplyLowerDiagonalBatch(
+gpuBatchMatrix::multiplyLowerDiagonalBatch(
   simMat, maternCov, diagMat, zmatGpu,
@@
 \section{Discussion}
-The package \pkg{gpuRandom} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
+The package \pkg{clrng} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
 We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
 Most of these functions uses local memory on GPU.
-Comparison of performance between using \pkg{gpuRandom} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
+Comparison of performance between using \pkg{clrng} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
 
-By leveraging the \pkg{gpuR} package, \pkg{gpuRandom} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.
-\pkg{gpuRandom} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for other applications.
+By leveraging the \pkg{gpuR} package, \pkg{clrng} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.
+\pkg{clrng} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for other applications.
 
@@
 
-\pkg{gpuRandom} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
+\pkg{clrng} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
 
-Future work on \pkg{gpuRandom} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs.
+Future work on \pkg{clrng} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs.
 (2) Now that the package can do Cholesky decomposition and matrix multiplication in batches on GPU, which is a motivation for us to work on parallel likelihood evaluations on GPU for Gaussian spatial models in the next step.
@@
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{clrng::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
 <<histAndQQ_prepare,include=FALSE,cache=TRUE>>=
 streams <- createStreamsGpu(n =512*128)
-normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
+normal_matrix<-clrng::rnorm(c(1024,512), streams=streams, 
                                 Nglobal=c(512,128), type="float")
@@
 hist(avector,breaks=40, main = "histogram")
-gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
+clrng::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
 @

### 2021-11-10 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2954 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
-\pkg{gpuRandom} is able to create streams both on the host and on the GPU device.
+\pkg{clrng} is able to create streams both on the host and on the GPU device.
 \fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
-Substreams are not created in \pkg{gpuRandom} because currently we use one stream per work-item.
+Substreams are not created in \pkg{clrng} because currently we use one stream per work-item.
 \begin{knitrout}
@@
 \begin{verbatim}
-as.vector(gpuRandom::runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1,
-    nrow(myStreamsGpu1)), type = "double"))
+as.vector(clrng::runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, nrow(myStreamsGpu1)),
+    type = "double"))
 ## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
-as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
+as.vector(clrng::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
     nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
@@
 \end{knitrout}
-The arguments of the \fct{gpuRandom::runif} are described as follows:
+The arguments of the \fct{clrng::runif} are described as follows:
 \begin{itemize}
@@
 \subsection{Plot the uniform random numbers (now deleted?)}
-%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
-The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
-
-%\begin{figure}[H]
-%\centering
-
-%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
-%\end{figure}
 
@@
 streams <- createStreamsGpu(n = 512 * 128)
-system.time(gpuRandom::rnorm(c(10000,10000), streams=streams, 
+system.time(clrng::rnorm(c(10000,10000), streams=streams, 
                              Nglobal=c(512,128), type="double"))
 ##    user  system elapsed 
-##   0.142   0.003   0.145
+##   0.098   0.004   0.390
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   6.689   0.830   7.513
+##   4.629   0.348   4.971
 \end{verbatim}
@@
 \begin{verbatim}
-r_matrix <- gpuRandom::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
+r_matrix <- clrng::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
 as.matrix(r_matrix)
 ##           [,1]      [,2]
-## [1,] 0.9260462 0.1431169
-## [2,] 0.6722247 0.8337073
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-The arguments of the \fct{gpuRandom::rexp} are described as follows:
+## [1,] 1.2243377 0.9593675
+## [2,] 0.7679958 0.5158689
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+The arguments of the \fct{clrng::rexp} are described as follows:
 \begin{itemize}
@@
 On a GPU, this step can be replicated say $n$ times in parallel by creating $n$ distinct random streams and launching $n$ separate work-items.
-Each work-item takes a random stream as input, performs all of Step 2 and 3, and returns the value of the test statistic and count on the GPU.
+Each work-item takes a random stream as input, performs all of Step 2 and 3, and returns the value of the test statistic on the GPU.
 Computing the p-value on the CPU (Step 4) is then a trivial operation.
 Step 2 of saving test statistics from the random tables is made optional, which can reduce the run-time.
-%There are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo.
-The use of streams and GPU would be quite different in that case.
-See for example
-
-The arguments of the \fct{gpuRandom::fisher.sim} are described as follows:
+By the way, there are a lot of other more recent methods for sampling (larger) contingency tables, many of them use Markov Chain Monte Carlo, the use of streams and GPU would be quite different in that case.
+See for examples \citep{10.1214/13-AOS1131, KAYIBI2018298, dyer1997sampling}.
+Doing the \fct{fisher.sim} function on GPU opens up many possibilities for future work, the specific implementation we've done for the tables isn't necessarily the optimal one.
+
+The arguments of the \fct{clrng::fisher.sim} are described as follows:
 \begin{itemize}
@@
 ##    user  system elapsed 
-##   0.739   0.004   0.742
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
-\fct{gpuRandom::fisher.sim} takes about 0.69 seconds, \pkg{gpuRandom} accounts for 4.64\% of the elapsed time on CPU.
+##   0.633   0.013   2.178
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
+\fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
 
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4044816
+## [1] 0.4039588
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  15.006   0.009  15.002
+##  10.184   0.004  10.178
 \end{verbatim}
@@
 \begin{verbatim}
-observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
+observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
+result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                 type="double",returnStatistics=TRUE,Nglobal = c(256,64))
@@
 #using GPU
-system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
+system.time(clrng::fisher.sim(week_GPU, 1e7, streams=streams,
             type="double", returnStatistics=FALSE,Nglobal = c(256,64)))
 ##    user  system elapsed 
-##   4.524   0.009   4.530
+##   0.503   0.017  10.226
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001249672
+## [1] 0.0001252669
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  91.498   0.128  91.580
+##  62.503   0.052  62.503
 \end{verbatim}
@@
 Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets.
-Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuRandom::maternGpuParam}.
+Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuBatchMatrix::maternGpuParam}.
 
@@
 ## [5,]  2.15  0.50      2.0      0          2         0.7853982
-paramsGpu = gpuRandom::maternGpuParam(myParamsBatch, type=theType)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-Step 3, compute Mat\'ern covariance matrices using \fct{gpuRandom::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
+paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+Step 3, compute Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
 are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
@@
 ## [1] 24000  4800
-gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
+gpuBatchMatrix::maternBatch(maternCov, coordsGpu, paramsGpu,
                        Nglobal=c(128,64), Nlocal=c(16,4))
@@
 \end{knitrout}
-Step 4, the first argument \code{maternCov} in \fct{gpuRandom::cholBatch} specifies the matrix object to take Cholesky decomposition.
+Step 4, the first argument \code{maternCov} in \fct{gpuBatchMatrix::cholBatch} specifies the matrix object to take Cholesky decomposition.
 Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
@@
 % \end{gather}
-Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{gpuRandom::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 \begin{knitrout}
@@
 \end{knitrout}
-Step 6, use \fct{gpuRandom::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
+Step 6, use \fct{gpuBatchMatrix::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
 \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
@@
 \section{Discussion}
-The package \pkg{gpuRandom} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
+The package \pkg{clrng} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
 We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
 Most of these functions uses local memory on GPU.
-Comparison of performance between using \pkg{gpuRandom} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
-
-By leveraging the \pkg{gpuR} package, \pkg{gpuRandom} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.
-\pkg{gpuRandom} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for other applications.
+Comparison of performance between using \pkg{clrng} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
+
+By leveraging the \pkg{gpuR} package, \pkg{clrng} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.
+\pkg{clrng} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for other applications.
 
@@
 
-\pkg{gpuRandom} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
-
-Future work on \pkg{gpuRandom} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs.
+\pkg{clrng} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
+
+Future work on \pkg{clrng} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs.
 (2) Now that the package can do Cholesky decomposition and matrix multiplication in batches on GPU, which is a motivation for us to work on parallel likelihood evaluations on GPU for Gaussian spatial models in the next step.
@@
 \begin{appendix}
+%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
+The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
+
+%\begin{figure}[H]
+%\centering
+
+%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
+%\end{figure}
+
 %In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
 The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{clrng::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
 

### 2021-11-10 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2955 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-An important way of parallel computing with \proglang{R} is using Graphics Processing Units (GPUs).
-GPUs can perform thousands of tests simultaneously, which makes them powerful at doing massive parallel computing and they are relatively cheap compared to multicore CPU's.
-Although there has been a few \proglang{R} packages developed that provide some GPU capability in the past years, they come with some restrictions.
-Packages such as \pkg{gputools}, \pkg{gpumatrix}, \pkg{cudaBayesreg}, \pkg{rpud} (available on github), are now no longer maintained, the popular \pkg{tensorflow} \citep{tensorflow1} package uses GPU via \proglang{Python}, which makes it hard to be incorporated in a new \proglang{R} package that uses GPU.
+Graphics Processing Units (GPUs) have the potential to make important contribution to parallel computing with \proglang{R}.
+GPUs can perform thousands of tests simultaneously, which makes them powerful for doing massively parallel computing, and they are relatively cheap compared to multicore CPU's.
+Although there have been a number of \proglang{R} packages developed which provide some GPU capability, they inevitably come with some limitations.
+Packages such as \pkg{gputools}, \pkg{gpumatrix}, \pkg{cudaBayesreg}, \pkg{rpud} (available on github), are now no longer maintained, the popular \pkg{tensorflow} \citep{tensorflow1} package uses GPU via \proglang{Python}, which makes it difficult to include as a dependency for new \proglang{R} packages.
 All of these mentioned packages are restricted to \proglang{R} users with NVIDIA GPUs.
-\pkg{gpuR} \citep{gpur1} is the only \proglang{R} package with a convenient interface between \proglang{R} and GPUs and it is created for any \proglang{R} user with a GPU device.
+\pkg{gpuR} \citep{gpur1} is the only \proglang{R} package with a convenient and flexible interface between \proglang{R} and GPUs, and it is compatible with many GPU devices.
 By utilizing the \pkg{ViennaCL} \citep*{rupp2016viennacl} library, it provides a bridge between \proglang{R} and non-proprietary GPUs through the \proglang{OpenCL} (Open Computing Language) backend, which when combined with \pkg{Rcpp} \citep{rcpp1} gives a building block for other \proglang{R} packages.
@@
 Random number generation is critical in simulation-based statistical inference, machine learning and many other scientific fields.
-While most random number generators are sequential, the \proglang{R} packages \pkg{parallel} (not available on cran), \pkg{rlecuyer} \citep{sevcikova2015package} and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
+While most random number generators are sequential, the \proglang{R} packages \pkg{parallel}, \pkg{rlecuyer} \citep{sevcikova2015package} and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
 More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG).
-\pkg{future} also uses the combined MRG algorithm for generating random numbers, however, it is not very clear how the streams are implemented.
+\pkg{future} also uses the combined MRG algorithm for generating random numbers.
 For up-to-date review papers on the generation of random numbers on parallel devices, and GPUs in particular, see:  \cite{rLEC15a,rLEC17p,rLEC21a}.
 
-There is not yet an \proglang{R} package that is able to generate random numbers on GPU, this is complicated to do because multiple streams will be needed for parallel generation of random numbers, and the stream states need to be able to be restored for reproducibility of results, which is essential in many simulation applications.
-We introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
+The \pkg{clrng} package described here is currently the only \proglang{R} package that is able to generate random numbers on GPUs.
+Accomplishing this is complicated to do because each process must produce an independent sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of numbers at any point.
+Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
 
@@
 \section{Uniform random number generation} \label{}
-Uniform random number generators (RNGs) are essential in simulating random numbers from all types of probability distributions.
+Uniform random number generators (RNGs) are the foundation for simulating random numbers from all types of probability distributions.
 \cite{l2012random} summarized the usual two steps to generate a random variable in computational statistics: (1) generating independent and identically distributed (i.i.d.) uniform random variables on the interval $(0, 1)$, (2) applying transformations to these i.i.d.
@@
 \pkg{clrng} package uses the MRG31k3p RNG, making it able to generate random numbers on GPUs.
-We choose MRG31k3p as the base generator for the following reasons:  the original \pkg{RNGStreams} package \citep{l2002object} was built with MRG32k3a, which was designed to be implemented in ``double'' in floating point(?), and not with 32-bit integers.
+We choose MRG31k3p as the base generator for the following reasons:  the original \pkg{RNGStreams} package \citep{l2002object} was built with MRG32k3a, which was designed to be implemented in double precision, and not with 32-bit integers.
 The MRG31k3p generator was designed later, specifically for 32-bit integer arithmetic, so it runs faster on the 32-bit GPUs.
@@
 \subsection{Creating streams}\label{createstreams}
+
+THE PROBLEM:  each process needs to be independent, there's a risk of overlap.
+solution: streams.
+mrg31kp3 a stream is a vector of three positive integers.
+initial state is first stream, compute new streams sequentially.
+
 The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
@@
 
+In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implement in \pkg{clrng}.
+The state (and seed) of each stream is a vector of six 31-bit integers.
+This size of state is appropriate for having streams running in work-items on GPU cards, for example, while providing a sufficient period length for most applications.
+The initial state of the first stream for the MRG31k3p is by default $(12345, 12345, 12345)$.
+
 In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
-
-In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
-Each stream is further partitioned into substreams of length $W = 2^{72}$.
-The state (and seed) of each stream is a vector of six 31-bit integers.
-This size of state is appropriate for having streams running in work-items on GPU cards, for example, while providing a sufficient period length for most applications.
-The initial state of the first stream for the MRG31k3p is by default a vector of six integers ``12345''.
 

### 2021-11-11 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2956 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\subsection{Creating streams}\label{createstreams}
-
-THE PROBLEM:  each process needs to be independent, there's a risk of overlap.
+%THE PROBLEM:  each process needs to be independent, there's a risk of overlap.
 solution: streams.
@@
 initial state is first stream, compute new streams sequentially.
-
+\subsection{Creating streams}\label{createstreams}
+Random number generation is more complicated when the RNG algorithm is run in parallel processes in \proglang{R}, there is a risk that the generated sequences of random numbers have correlation.
+\pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
 The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
 Creating a new stream amounts to computing its starting point (seed).
+In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, which are occasionally useful.
 
-In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
-Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implement in \pkg{clrng}.
+in \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implemented in \pkg{clrng}.
 The state (and seed) of each stream is a vector of six 31-bit integers.
-This size of state is appropriate for having streams running in work-items on GPU cards, for example, while providing a sufficient period length for most applications.
-The initial state of the first stream for the MRG31k3p is by default $(12345, 12345, 12345)$.
-
-In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
-Whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
+This size of state is appropriate for having streams running in work-items on GPU cards, while providing a sufficient period length for most applications.
+The initial state of the first stream (also called ``initial seed for the package'' in \proglang{clRNG} and ``initial'' in \pkg{clrng}) for the MRG31k3p is by default $(12345, 12345, 12345, 12345, 12345, 12345)$.
+Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
 

### 2021-11-11 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2956 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+%THE PROBLEM:  each process needs to be independent, there's a risk of overlap.
+solution: streams.
+mrg31kp3 a stream is a vector of three positive integers.
+initial state is first stream, compute new streams sequentially.
 \subsection{Creating streams}\label{createstreams}
+Random number generation is more complicated when the RNG algorithm is run in parallel processes in \proglang{R}, there is a risk that the generated sequences of random numbers have correlation.
+\pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
 The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
 Creating a new stream amounts to computing its starting point (seed).
+In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, which are occasionally useful.
 
-In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
-Whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
-
-In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
-Each stream is further partitioned into substreams of length $W = 2^{72}$.
+in \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implemented in \pkg{clrng}.
 The state (and seed) of each stream is a vector of six 31-bit integers.
-This size of state is appropriate for having streams running in work-items on GPU cards, for example, while providing a sufficient period length for most applications.
-The initial state of the first stream for the MRG31k3p is by default a vector of six integers ``12345''.
+This size of state is appropriate for having streams running in work-items on GPU cards, while providing a sufficient period length for most applications.
+The initial state of the first stream (also called ``initial seed for the package'' in \proglang{clRNG} and ``initial'' in \pkg{clrng}) for the MRG31k3p is by default $(12345, 12345, 12345, 12345, 12345, 12345)$.
+Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
 
@@
 ##    user  system elapsed 
-##   0.098   0.004   0.390
+##   0.093   0.008   0.391
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.633   0.013   2.178
+##   0.619   0.009   2.158
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.503   0.017  10.226
+##   0.514   0.010  10.215
 \end{verbatim}

### 2021-11-11 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2957 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-in \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
 Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implemented in \pkg{clrng}.

### 2021-11-11 pkg/gpuRandom/src/crossprodBatch.cpp
crossprod
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2958 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -77,7 +77,7 @@ std::string crossprodBatchString(
   "event_t ev;\n"
   "const int NpadLocal = get_local_size(1);\n"
   "int AHere, CHere;\n"
-  "int Dmatrix, Drow, Dcol, DrowNpadC, Dinner, DinnerA, DinnerAcol, DrowBlock, DinnerBlock, DcolBlock, DmatrixBlock, DcolInBounds, DmatrixInBounds;\n"
+  "int Dmatrix, Drow, Dcol, DrowNpadC, Dinner, DinnerA, DinnerAcol, DrowBlock, DinnerBlock, DcolBlock, DcolInBounds, DrowInBounds;\n"
   "int A0Dcol, A0Drow;// location of elements A[0,Dcol] and A[0,Drow]\n"
   "const int AHereInc = get_num_groups(1)*NpadBetweenMatricesA;\n";
   
@@ -141,29 +141,23 @@ std::string crossprodBatchString(
   //   "  ){\n";
   
   result +=  "\n\n"
-  "   for(DmatrixBlock = 0;\n"
-  "    DmatrixBlock < Nmatrix;\n"
-  "    DmatrixBlock += get_num_groups(1)\n"
-  "    ){\n"
-  "    Dmatrix = DmatrixBlock + get_group_id(1);\n"
-  "    if(Dmatrix < Nmatrix) {\n"
-  "     DmatrixInBounds = 1;\n"
-  "     } else {\n"
-  "     DmatrixInBounds = 0;\n"
-  "     Dmatrix = 0;\n"
-  "     }\n"
-  
+  "   for(Dmatrix = get_group_id(1);\n"
+  "       Dmatrix < Nmatrix;\n"
+  "       Dmatrix += get_num_groups(1)\n"
+  "    ){\n";
+
+  result += 
   "     AHere = Dmatrix * NpadBetweenMatricesA + NstartA;\n";
   if(NpadD) {
     result +=
-      "    DHere = Dmatrix * NpadD + NstartD,\n"; 
+      "     DHere = Dmatrix * NpadD + NstartD,\n"; 
   }
   if(onlyDiagC) {
     result +=
-      "    CHere = (Dmatrix + NrowStartC) * NpadC + NstartC;\n"; 
+      "     CHere = (Dmatrix + NrowStartC) * NpadC + NstartC;\n"; 
   }else{
     result +=  
[…]

### 2021-11-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2959 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
 Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, which are occasionally useful.
@@
 The initial state of the first stream (also called ``initial seed for the package'' in \proglang{clRNG} and ``initial'' in \pkg{clrng}) for the MRG31k3p is by default $(12345, 12345, 12345, 12345, 12345, 12345)$.
-Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
 
@@
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
+Users can set \code{verbose>1} in \fct{runif} to print out the kernels.
 Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
@@
 \code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively.
-\code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of the \code{streams} (these variables are defined in macros not shown here).
-Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}.
+\code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of \code{streams} (these variables are defined in macros not shown here).
+Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, where \code{g1} and \code{g2} point to the first three and second three elements of stream states.
 \code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$, if to generate double-precision random numbers.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ for generating double-precision random numbers.
+If to generate random integers, \code{temp} is not scaled.
 At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
@@
   __global int* streams,
-    __global double* out){
+  __global float* out){
 
-const int 
-index = get_global_id(0)*get_global_size(1) + get_global_id(1),
-DrowInc = 2*get_global_size(0), 
-DrowStartInc = DrowInc * NpadCol,
-startvalue=index * NpadStreams;
-int Drow, Dcol, DrowStart, Dentry;
-uint temp;
+const int index = get_global_id(0)*get_global_size(1) + get_global_id(1);
+int Drow, Dcol, DrowStart, Dentry, DrowBlock, DcolBlock, DrowInBounds;
+const int DrowStartInc = get_global_size(0) * NpadCol;
 uint g1[3], g2[3];
+const int startvalue=index * NpadStreams;
+float temp, fact = mrg31k3p_NORM_cl;
 streamsToPrivate(streams,g1,g2,startvalue);
 
-//filling in random numbers in the output matrix
-for(Drow=2*get_global_id(0), 
-    DrowStart = (Drow + get_local_id(1))* NpadCol;
-    Drow < Nrow; Drow += DrowInc, DrowStart += DrowStartInc) {
+// filling in random numbers in the output matrix
+for(DrowBlock = 0, Drow=get_global_id(0), DrowStart = Drow * NpadCol;
+    DrowBlock < Nrow;
+    Drow += get_global_size(0), 
+    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc) {
 
-   for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
-       Dcol < Ncol;
-       Dcol += get_num_groups(1), Dentry += get_num_groups(1)) {
+    DrowInBounds = Drow < Nrow;
 
-       temp = clrngMrg31k3pNextState(g1, g2);
-       out[Dentry] = mrg31k3p_NORM_cl * temp;
+    for(DcolBlock = 0, Dcol=get_global_id(1), Dentry = DrowStart + Dcol;
+        DcolBlock < Ncol;
+        DcolBlock += get_global_size(1), Dentry += get_global_size(1) ) {
+        temp = fact * clrngMrg31k3pNextState(g1, g2);
+        if(DrowInBounds) out[Dentry] = temp;
+  }//Dcol
 
-    }//Dcol
 }//Drow
+
 streamsFromPrivate(streams,g1,g2,startvalue);
-}
+
+}//kernel
 \end{lstlisting}
@@
 The number in a cell of the matrix corresponds to the stream index, it indicates which stream generates a random number for this matrix cell.
-For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
+%For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
 \begin{figure}[H]
@@
 $U(0,1)$ random numbers with the function \fct{runif}.
-\fct{runif} can also be used to produce random integers in $[1, 2147483647]$.
 To view the generated random numbers, we need to convert them to R vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
@@
   \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
+  \item \code{verbose}: verbose print extra information if >1.
 \end{itemize}
@@
 
-\subsection{Plot the uniform random numbers (now deleted?)}
+%\subsection{Plot the uniform random numbers (now deleted?)}
 
@@
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},caption={Using streams to produce standard normal random numbers}]
-const int 
-index = get_global_id(0)*get_global_size(1) + get_global_id(1);
-const int 
-DrowInc = 2*get_global_size(0), DrowStartInc = DrowInc * NpadCol;
-int Drow, Dcol, DrowStart, Dentry;
-uint temp;
-uint g1[3], g2[3];
-const int startvalue=index * NpadStreams;
-local double part[2], cosPart1, sinPart1;
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},caption={Using streams to produce standard normal random numbers},escapeinside={(*@}{@*)}]
+__kernel void mrg31k3pMatrix(
+  __global int* streams,
+  __global float* out){
 
+(*@{\hspace*{1cm}\vdots}@*)
+local float part[2];
+float sinOrCosPart1, addForSine = get_local_id(1) * PI;
+if(get_local_id(1)){
+  fact = TWOPI_mrg31k3p_NORM_cl;
+}
 streamsToPrivate(streams,g1,g2,startvalue);
+for(DrowBlock = 0, Drow=get_global_id(0), DrowStart = Drow * NpadCol;
+    DrowBlock < Nrow;
+    Drow += get_global_size(0), 
+    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc){
+    DrowInBounds = Drow < Nrow;
+    for(DcolBlock = 0, Dcol=get_global_id(1), Dentry = DrowStart + Dcol;
+        DcolBlock < Ncol;
+        DcolBlock += get_global_size(1), Dentry += get_global_size(1)){
 
-// Iterate for the rows
-for(Drow=2*get_global_id(0), 
-    DrowStart = (Drow + get_local_id(1))* NpadCol;
-    Drow < Nrow; 
-    Drow += DrowInc, DrowStart += DrowStartInc) {
-  // Iterate for the columns
-  for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
-      Dcol < Ncol;
-      Dcol += get_num_groups(1), Dentry += get_num_groups(1)) {
+      temp = fact * clrngMrg31k3pNextState(g1, g2);
+      part[get_local_id(1)] = temp;
 
-    // Generates an uniform random integer
-    temp = clrngMrg31k3pNextState(g1, g2);
-
-    if(get_local_id(1)) {
-      // Cache data to local memory
-      part[1] = TWOPI_mrg31k3p_NORM_cl * temp;
-      cosPart1 = cos(part[1]);
-      sinPart1 = sin(part[1]);
-    } else {
-      part[0] = sqrt(-2.0*log(temp * mrg31k3p_NORM_cl));
+      if(!get_local_id(1)) {
+        part[0] = sqrt(-2.0*log(part[0]));
       }
-
-    // Wait until all work items have read the data and 
-    // it becomes visible
       barrier(CLK_LOCAL_MEM_FENCE);
 
-    // Perform the operation and output the data
-    if(get_local_id(1)) {
-      out[Dentry] = part[0]*sinPart1;
-    } else {
-      out[Dentry] = part[0]*cosPart1;
-    }
+      // cos for local0, sine for local1
+      sinOrCosPart1 = cos(part[1] + addForSine);
+      temp = part[0]*sinOrCosPart1;
+        if(DrowInBounds) out[Dentry] = temp;
       barrier(CLK_LOCAL_MEM_FENCE);
@@
 }//Drow
-
 streamsFromPrivate(streams,g1,g2,startvalue);
+}//kernel
 \end{lstlisting}
@@
 Each work-group generates a pair of Gaussian random numbers \code{(X,Y)} in each iteration, where work-item of `\code{get_local_id(1)}$=0$' generates the \code{X} of a pair, and work-item of `\code{get_local_id(1)}$=1$' generates the \code{Y} of a pair.
-Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of the matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
+%Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of the matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
 %Therefore, each work-item generates one Gaussian random number in each iteration,
@@
 The random variable $X \sim \text{Exponential}(\lambda)$ has cumulative distribution function $F_{X}(x)=1-e^{-\lambda x}$ for $x \geq 0$ and $\lambda > 0$.
-The inverse of $F^{-1}_{X}$ is $F^{-1}_{x}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
-The kernel is not shown because it is mostly same as the kernel for uniform and normal random numbers, basically just by applying the inverse transform formula to the \code{clrngMrg31k3pNextState(g1, g2)} in the kernel.
-Developers can make use of our kernel for generating other types of random variables on GPU in \proglang{R}.
+The inverse of $F_{X}$ is $F^{-1}_{x}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
+The kernel is not shown because it is mostly same as the kernel for uniform and normal random numbers, except for the part that does the inverse transform.
+%Developers can make use of our kernel for generating other types of random variables on GPU in \proglang{R}.
 
-Below is the code that we produce a 2 by 2 matrix of exponential random number with expectation 1.
+Below is an example that we produce a 2 by 2 matrix of exponential random number with expectation 1.
 <<random exponential, eval=TRUE, cache=FALSE>>=
@@
   \item \code{streams}: streams. If missing, a stream with random initial state is supplied.
-  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
+  \item \code{Nglobal}: global index space. Default is $(64,8)$.
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.

### 2021-11-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2959 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
 Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, which are occasionally useful.
 
-in \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
 Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implemented in \pkg{clrng}.
@@
   __global int* streams,
-    __global double* out){
-
-const int 
-index = get_global_id(0)*get_global_size(1) + get_global_id(1),
-DrowInc = 2*get_global_size(0), 
-DrowStartInc = DrowInc * NpadCol,
-startvalue=index * NpadStreams;
-int Drow, Dcol, DrowStart, Dentry;
-uint temp;
+  __global float* out){
+
+const int index = get_global_id(0)*get_global_size(1) + get_global_id(1);
+int Drow, Dcol, DrowStart, Dentry, DrowBlock, DcolBlock, DrowInBounds;
+const int DrowStartInc = get_global_size(0) * NpadCol;
 uint g1[3], g2[3];
+const int startvalue=index * NpadStreams;
+float temp, fact = mrg31k3p_NORM_cl;
 streamsToPrivate(streams,g1,g2,startvalue);
 
-//filling in random numbers in the output matrix
-for(Drow=2*get_global_id(0), 
-    DrowStart = (Drow + get_local_id(1))* NpadCol;
-    Drow < Nrow; Drow += DrowInc, DrowStart += DrowStartInc) {
-
-   for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
-       Dcol < Ncol;
-       Dcol += get_num_groups(1), Dentry += get_num_groups(1)) {
-
-       temp = clrngMrg31k3pNextState(g1, g2);
-       out[Dentry] = mrg31k3p_NORM_cl * temp;
-
+// filling in random numbers in the output matrix
+for(DrowBlock = 0, Drow=get_global_id(0), DrowStart = Drow * NpadCol;
+    DrowBlock < Nrow;
+    Drow += get_global_size(0), 
+    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc) {
+
+    DrowInBounds = Drow < Nrow;
+
+    for(DcolBlock = 0, Dcol=get_global_id(1), Dentry = DrowStart + Dcol;
+        DcolBlock < Ncol;
+        DcolBlock += get_global_size(1), Dentry += get_global_size(1) ) {
+        temp = fact * clrngMrg31k3pNextState(g1, g2);
+        if(DrowInBounds) out[Dentry] = temp;
   }//Dcol
+
 }//Drow
+
 streamsFromPrivate(streams,g1,g2,startvalue);
-}
+
+}//kernel
 \end{lstlisting}
@@
     type = "double"))
-## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
+## [1] 0.7353245 0.6142074 0.1100781 0.6487742 0.3661944 0.1088229
 as.vector(clrng::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
     nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
-## [1] 0.6487742 0.1112075 0.3661944 0.5018562
+## [1] 0.5330548 0.9783798 0.9151238 0.8509746
 \end{verbatim}
@@
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},caption={Using streams to produce standard normal random numbers}]
-const int 
-index = get_global_id(0)*get_global_size(1) + get_global_id(1);
-const int 
-DrowInc = 2*get_global_size(0), DrowStartInc = DrowInc * NpadCol;
-int Drow, Dcol, DrowStart, Dentry;
-uint temp;
-uint g1[3], g2[3];
-const int startvalue=index * NpadStreams;
-local double part[2], cosPart1, sinPart1;
-
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},caption={Using streams to produce standard normal random numbers},escapeinside={(*@}{@*)}]
+__kernel void mrg31k3pMatrix(
+  __global int* streams,
+  __global float* out){
+
+(*@{\hspace*{3cm}\vdots}@*)
+local float part[2];
+float sinOrCosPart1, addForSine = get_local_id(1) * PI;
+if(get_local_id(1)){
+  fact = TWOPI_mrg31k3p_NORM_cl;
+}
 streamsToPrivate(streams,g1,g2,startvalue);
-
-// Iterate for the rows
-for(Drow=2*get_global_id(0), 
-    DrowStart = (Drow + get_local_id(1))* NpadCol;
-    Drow < Nrow; 
-    Drow += DrowInc, DrowStart += DrowStartInc) {
-  // Iterate for the columns
-  for(Dcol=get_group_id(1), Dentry = DrowStart + Dcol;
-      Dcol < Ncol;
-      Dcol += get_num_groups(1), Dentry += get_num_groups(1)) {
-
-    // Generates an uniform random integer
-    temp = clrngMrg31k3pNextState(g1, g2);
-
-    if(get_local_id(1)) {
-      // Cache data to local memory
-      part[1] = TWOPI_mrg31k3p_NORM_cl * temp;
-      cosPart1 = cos(part[1]);
-      sinPart1 = sin(part[1]);
-    } else {
-      part[0] = sqrt(-2.0*log(temp * mrg31k3p_NORM_cl));
-    }
-
-    // Wait until all work items have read the data and 
-    // it becomes visible
-    barrier(CLK_LOCAL_MEM_FENCE);
-
-    // Perform the operation and output the data
-    if(get_local_id(1)) {
-      out[Dentry] = part[0]*sinPart1;
-    } else {
-      out[Dentry] = part[0]*cosPart1;
+for(DrowBlock = 0, Drow=get_global_id(0), DrowStart = Drow * NpadCol;
+    DrowBlock < Nrow;
+    Drow += get_global_size(0), 
+    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc){
+    DrowInBounds = Drow < Nrow;
+    for(DcolBlock = 0, Dcol=get_global_id(1), Dentry = DrowStart + Dcol;
+        DcolBlock < Ncol;
+        DcolBlock += get_global_size(1), Dentry += get_global_size(1)){
+
+      temp = fact * clrngMrg31k3pNextState(g1, g2);
+      part[get_local_id(1)] = temp;
+
+      if(!get_local_id(1)) {
+        part[0] = sqrt(-2.0*log(part[0]));
       }
@@
 
+      // cos for local0, sine for local1
+      sinOrCosPart1 = cos(part[1] + addForSine);
+      temp = part[0]*sinOrCosPart1;
+        if(DrowInBounds) out[Dentry] = temp;
+      barrier(CLK_LOCAL_MEM_FENCE);
+
   }//Dcol
 }//Drow
-
 streamsFromPrivate(streams,g1,g2,startvalue);
+}//kernel
 \end{lstlisting}
@@
 ##    user  system elapsed 
-##   0.093   0.008   0.391
+##   0.092   0.004   0.393
 \end{verbatim}
@@
 The random variable $X \sim \text{Exponential}(\lambda)$ has cumulative distribution function $F_{X}(x)=1-e^{-\lambda x}$ for $x \geq 0$ and $\lambda > 0$.
-The inverse of $F^{-1}_{X}$ is $F^{-1}_{x}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
-The kernel is not shown because it is mostly same as the kernel for uniform and normal random numbers, basically just by applying the inverse transform formula to the \code{clrngMrg31k3pNextState(g1, g2)} in the kernel.
-Developers can make use of our kernel for generating other types of random variables on GPU in \proglang{R}.
-
-Below is the code that we produce a 2 by 2 matrix of exponential random number with expectation 1.
+The inverse of $F_{X}$ is $F^{-1}_{x}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
+The kernel is not shown because it is mostly same as the kernel for uniform and normal random numbers, except for the part that does the inverse transform.
+%Developers can make use of our kernel for generating other types of random variables on GPU in \proglang{R}.
+
+Below is an example that we produce a 2 by 2 matrix of exponential random number with expectation 1.
 \begin{knitrout}
@@
 ##            [,1]      [,2]
-## [1,] 1.2243377 0.9593675
-## [2,] 0.7679958 0.5158689
+## [1,] 1.22433768 0.7679958
+## [2,] 0.06687963 1.9744549
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.619   0.009   2.158
+##   0.617   0.008   2.156
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.514   0.010  10.215
+##   0.491   0.023  10.210
 \end{verbatim}

### 2021-11-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2960 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
-The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
-Creating a new stream amounts to computing its starting point (seed).
-In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+A popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long disjoint pieces of equal length $Z$, and use each piece as a separate stream.
+Creating a new stream amounts to computing its starting point.
+In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
-Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, which are occasionally useful.
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}.
 
-In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+In \proglang{clRNG} library, the MRG31k3p RNG's entire period of length approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
 Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implemented in \pkg{clrng}.
@@
 
-The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG.
+%The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host using the MRG31k3p RNG.
 If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
-\pkg{clrng} is able to create streams both on the host and on the GPU device.
-\fct{createStreamsCpu} is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
+\pkg{clrng} is able to create streams both on the host and on the GPU device. The \fct{createStreamsCpu} function does the former. The \proglang{R} output below creates 4 streams on the host.
+%is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
 Substreams are not created in \pkg{clrng} because currently we use one stream per work-item.
@@
 \end{itemize}
-We can have streams on GPU by either converting the \code{myStreamsCpu} to a `\code{vclMatrix}' or producing streams using \fct{createStreamsGpu}.
-\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
+We move streams to the GPU by converting them to a `\code{vclMatrix}'.
+%\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
 Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we produce on device exactly the same streams as those created on host as follows.
@@
 # creating streams on GPU
-myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu = vclMatrix(myStreamsCpu)
 myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
-range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
 @
-\fct{createStreamsGpu} also has the two arguments \code{n} and \code{initial}.
+\fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
+However, any data is lost when the \proglang{R} session is terminated.
+%also has the two arguments \code{n} and \code{initial}.
 \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
@@
 \subsection{Using streams to generate uniform random numbers}
-The streams created sequentially on a GPU are then used in work-items that executes in parallel on a GPU device.
-In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be equal to the number of total work-items in use.
+The streams created sequentially are then used by the GPU's work-items to generate random numbers.
+In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be greater than the number of total work-items in use.
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
@@
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel},caption={Using streams to produce random U(0,1) numbers}]
+\begin{lstlisting}[language=C,basicstyle=\small,breaklines=true,label={lst:uniformkernel}]
 __kernel void mrg31k3pMatrix(
@@
     Drow += get_global_size(0), 
-    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc) {
+    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc){
 
@@
         DcolBlock < Ncol;
-        DcolBlock += get_global_size(1), Dentry += get_global_size(1) ) {
+        DcolBlock += get_global_size(1), Dentry += get_global_size(1)){
         temp = fact * clrngMrg31k3pNextState(g1, g2);
@@
 <<simStuff2, eval=TRUE,tidy=TRUE>>=
-as.vector(clrng::runif(n=6, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
-
-as.vector(clrng::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
+as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(1,nrow(myStreamsGpu)), type = "double"))
 @
@@
 \end{itemize}
-\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu1} and \code{myStreamsGpu2} here.
-Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
+\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here.
+Reusing \code{myStreamsGpu} will produce a vector different from \code{sim_1}.
 Because each time a random number is generated, the current state of the stream advances by one position.
-%Below shows the \code{myStreamsGpu1} after it being used.
+%Below shows the \code{myStreamsGpu} after it being used.
 <<eval=TRUE,tidy=TRUE, include=FALSE>>=
-t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
+t(matrix(as.matrix(myStreamsGpu), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
 @
@@
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},caption={Using streams to produce standard normal random numbers},escapeinside={(*@}{@*)}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},breaklines=true, escapeinside={(*@}{@*)}]
 __kernel void mrg31k3pMatrix(
@@
 
-(*@{\hspace*{1cm}\vdots}@*)
+(*@{\hspace*{0cm}\vdots}@*)
 local float part[2];
@@
   \item Calculate the test statistic for the observed table.
-    \item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table, compute and sometimes save the test statistic from the random table. 
+    \item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table using the \fct{rcont2} algorithm, compute and sometimes save the test statistic from the random table.
     \item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.
@@
 Suppose $U=(U_1, U_2, \dots, U_n)$ is a Gaussian random field with mean $\mu$ and covariance matrix $\Sigma$, without loss of generality, we assume mean value zero $(\mu = 0)$.
-Since covariance matrix $\Sigma$ is symmetric and positive-definite, we can take Cholesky decomposition of $\Sigma$, then we can simulate random samples from $U$ using $U=L*D^{1/2}*Z$, where
-\begin{itemize}
-\itemsep0em 
-\item $L$ is the lower unit triangular matrix in Cholesky decomposition of $\Sigma = L*D*L^\top$,
-\item $D$ is a diagonal matrix.
-\item $Z=(Z_1, Z_2, \dots, Z_n) \sim \text{MVN}(0,I_n)$, where $I_n$ is a $n \times n$ identity matrix.
-\end{itemize}
+Since covariance matrix $\Sigma$ is symmetric and positive-definite, we can take Cholesky decomposition of $\Sigma$, then we can simulate random samples from $U$ using Algorithm \ref{algorithm2},
+
+\begin{algorithm*}[H] \label{algorithm2}
+\SetAlgoLined
+ 1, Calculate the covariance matrix $\Sigma$ between locations\;
+ 2, Compute the Cholesky decomposition of $\Sigma = L \cdot D \cdot L^\top$, $L$ is the lower unit triangular matrix, $D$ is a diagonal matrix\;
+ 3, Generate on GPU a random matrix $Z=(Z_1, Z_2, \dots, Z_n) \sim \text{MVN}(0,I_n)$, where $I_n$ is a $n \times n$ identity matrix\;
+ 4, Compute the random samples from $U$ in batches using $U = L \cdot D^{\frac{1}{2}} \cdot Z$ \;
+ \caption{Gaussian random fields simulation using covariance matrix decomposition method.}
+\end{algorithm*}
+
+% $U=L*D^{1/2}*Z$, where 
+% \begin{itemize}
+% \itemsep0em 
+% \item $L$ is the lower unit triangular matrix in Cholesky decomposition of $\Sigma = L*D*L^\top$,
+% \item $D$ is a diagonal matrix.
+% \item $Z=(Z_1, Z_2, \dots, Z_n) \sim \text{MVN}(0,I_n)$, where $I_n$ is a $n \times n$ identity matrix.
+% \end{itemize}
 The matrix decomposition method is straightforward to implement, however, when a large number of locations ($n$) is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
@@
 \end{gather}
-Step 7, Finally, plot the 10 simulated Gaussian random surfaces.
+Step 7, Finally, plot the 10 realizations.
 <<r plotSimSetup, echo=TRUE, eval=TRUE, cache=FALSE>>=

### 2021-11-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2960 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
-The popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long pieces, these pieces are adjacent and of equal length $Z$, so each piece can be used as a separate stream.
-Creating a new stream amounts to computing its starting point (seed).
-In general, a stream object contains three states: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+A popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long disjoint pieces of equal length $Z$, and use each piece as a separate stream.
+Creating a new stream amounts to computing its starting point.
+In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
-Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, which are occasionally useful.
-
-In \proglang{clRNG} library, for the MRG31k3p RNG, the entire period length of approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}.
+
+In \proglang{clRNG} library, the MRG31k3p RNG's entire period of length approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
 Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implemented in \pkg{clrng}.
@@
 # creating streams on GPU
-myStreamsGpu1 = vclMatrix(myStreamsCpu)
+myStreamsGpu = vclMatrix(myStreamsCpu)
 myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
-range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
-## [1] 0 0
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\fct{createStreamsGpu} also has the two arguments \code{n} and \code{initial}.
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
+However, any data is lost when the \proglang{R} session is terminated.
+%also has the two arguments \code{n} and \code{initial}.
 \fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
@@
 \begin{verbatim}
-as.vector(clrng::runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, nrow(myStreamsGpu1)),
+as.vector(clrng::runif(n = 6, streams = myStreamsGpu, Nglobal = c(1, nrow(myStreamsGpu)),
     type = "double"))
 ## [1] 0.7353245 0.6142074 0.1100781 0.6487742 0.3661944 0.1088229
-as.vector(clrng::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
-    nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
-## [1] 0.5330548 0.9783798 0.9151238 0.8509746
 \end{verbatim}
@@
 \end{itemize}
-\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu1} and \code{myStreamsGpu2} here.
-Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}.
+\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here.
+Reusing \code{myStreamsGpu} will produce a vector different from \code{sim_1}.
 Because each time a random number is generated, the current state of the stream advances by one position.
-%Below shows the \code{myStreamsGpu1} after it being used.
+%Below shows the \code{myStreamsGpu} after it being used.
 
@@
 
-(*@{\hspace*{3cm}\vdots}@*)
+(*@{\hspace*{0cm}\vdots}@*)
 local float part[2];
@@
 ##    user  system elapsed 
-##   0.092   0.004   0.393
+##   0.160   0.004   0.461
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.617   0.008   2.156
+##   1.138   0.019   2.691
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.491   0.023  10.210
+##   0.895   0.001  10.601
 \end{verbatim}
@@
 \end{gather}
-Step 7, Finally, plot the 10 simulated Gaussian random surfaces.
+Step 7, Finally, plot the 10 realizations.
 \begin{knitrout}

### 2021-11-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2961 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 In \proglang{clRNG} library, the MRG31k3p RNG's entire period of length approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
-Each stream can be further partitioned into substreams of length $W = 2^{72}$, although this is not currently implemented in \pkg{clrng}.
+Each stream can be further partitioned into substreams of length $2^{72}$, although this is not currently implemented in \pkg{clrng}.
 The state (and seed) of each stream is a vector of six 31-bit integers.
@@
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
-Users can set \code{verbose>1} in \fct{runif} to print out the kernels.
+Users can set \code{verbose=2} in \fct{runif} to print out the kernels.
 Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
@@
 \code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ for generating double-precision random numbers.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type.
 If to generate random integers, \code{temp} is not scaled.
@@
 \begin{framed}
-\begin{lstlisting}[language=C,basicstyle=\small,breaklines=true,label={lst:uniformkernel}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel}]
 __kernel void mrg31k3pMatrix(
@@
 \end{framed}
-This kernel executes over a 2-dimensional NDRange.
-Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example.
+%Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example.
 Suppose we have a $2 \times 4$ NDRange ((a) in Figure \ref{fig:1}) and want to create a $4 \times 6$ matrix ((b) in Figure \ref{fig:1}) of uniform random numbers.
@@
 %For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
-\begin{figure}[H]
-    \centering
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f1} }}
-    \qquad
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f2} }}
-    \caption{Using streams to produce a $4 \times 6$ matrix of random U(0,1) numbers. \label{fig:1}}
-\end{figure}
+% \begin{figure}[H]
+%     \centering
+%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f1} }}
+%     \qquad
+%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f2} }}
+%     \caption{Using streams to produce a $4 \times 6$ matrix of random U(0,1) numbers. \label{fig:1}}
+% \end{figure}
 
@@
 \end{itemize}
-\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here.
+%\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here. 
 Reusing \code{myStreamsGpu} will produce a vector different from \code{sim_1}.
@@
 %The derivation basically uses transformation from Cartesian coordinates to polar coordinates to represent two independent, Normally distributed random variables.
-The Box-Muller algorithm turns out to be the best choice for Gaussian transform on GPU compared to other transform methods \citep{howes2007efficient}, because this algorithm has no branching or looping, which are the things GPU is not good at.
+The Box-Muller algorithm is a very good choice for Gaussian transform on GPU compared to other transform methods \citep{howes2007efficient}, because this algorithm has no branching or looping, which are the things GPU is not good at.
 %but GPU is good at the high computational load of sine and cosine functions in the algorithm.
@@
 If developers want to use other Gaussian transform methods, this kernel is easy to be adapted and incorporated in other \proglang{R} packages.
-The kernel executes over a 2-dimensional NDRange with $1 \times 2$ work-groups.
+The kernel executes over a 2-dimensional index space with $1 \times 2$ work-groups.
 \code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
+\code{PI_2} equals to $\pi/2$.
 As the work-items in a work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
@@
   __global int* streams,
-  __global float* out){
-
-(*@{\hspace*{0cm}\vdots}@*)
-local float part[2];
-float sinOrCosPart1, addForSine = get_local_id(1) * PI;
+  __global double* out){
+
+(*@\centerline{\raisebox{-1pt}[0pt][0pt]{$\vdots$}}@*)
+
+double temp, fact = mrg31k3p_NORM_cl;
+local double part[2];
+double sinOrCosPart1, addForSine = - get_local_id(1) * PI_2;
 if(get_local_id(1)){
@@
 streamsToPrivate(streams,g1,g2,startvalue);
+
 for(DrowBlock = 0, Drow=get_global_id(0), DrowStart = Drow * NpadCol;
     DrowBlock < Nrow;
-    Drow += get_global_size(0), 
-    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc){
+    Drow += get_global_size(0), DrowBlock +=get_global_size(0), 
+    DrowStart += DrowStartInc) {
+
     DrowInBounds = Drow < Nrow;
@@
         DcolBlock < Ncol;
-        DcolBlock += get_global_size(1), Dentry += get_global_size(1)){
+        DcolBlock += get_global_size(1), Dentry += get_global_size(1) ){
 
@@
 
-      // cos for local0, sine for local1
+      // sine for local[1], cos for local[0]
       sinOrCosPart1 = cos(part[1] + addForSine);
@@
 \end{framed}
-We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange.
+%We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange.
 %Each stream is copied to the private memory of each work-item.
-Work-items are organized to be in 1 by 2 work-groups as the random numbers are generated in pairs using the Box-Muller method.
+%Work-items are organized to be in 1 by 2 work-groups as the random numbers are generated in pairs using the Box-Muller method.
 The shading on the grids in the NDRange distinguishes work-groups.
@@
 %Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of the matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
-%Therefore, each work-item generates one Gaussian random number in each iteration,
-The matrix filling process is the same as that shown in Figure \ref{fig:1}.
-
-\begin{figure}[H]%
-    \centering
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f3} }}%
-    \qquad
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
-    \caption{Using streams to produce a $4 \times 6$ matrix of normal random numbers.}%
-    \label{fig3}
-\end{figure}
+%Therefore, each work-item generates one Gaussian random number in each iteration, The matrix filling process is the same as that shown in Figure \ref{fig:1}.
+
+% \begin{figure}[H]%
+%     \centering
+%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f3} }}%
+%     \qquad
+%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
+%     \caption{Using streams to produce a $4 \times 6$ matrix of normal random numbers.}%
+%     \label{fig3}
+% \end{figure}
 
@@
 ##    user  system elapsed 
-##   0.160   0.004   0.461
+##   0.086   0.011   0.530
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.138   0.019   2.691
+##   0.620   0.010   2.162
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.895   0.001  10.601
+##   0.504   0.015  10.211
 \end{verbatim}

### 2021-11-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2961 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   __global int* streams,
-  __global float* out){
+  __global double* out){
 
-(*@{\hspace*{0cm}\vdots}@*)
-local float part[2];
-float sinOrCosPart1, addForSine = get_local_id(1) * PI;
+(*@\centerline{\raisebox{-1pt}[0pt][0pt]{$\vdots$}}@*)
+
+double temp, fact = mrg31k3p_NORM_cl;
+local double part[2];
+double sinOrCosPart1, addForSine = - get_local_id(1) * PI_2;
 if(get_local_id(1)){
@@
 %Here with the loop iterating over columns, the cells of matrix are filled from left to right, that is, in the first iteration, the 8 work-items generate random numbers in the leftmost two columns (8 cells) of the matrix simultaneously, the second iteration fills random numbers in the middle two columns of the matrix, and the third iteration fills the rest two columns of the matrix simultaneously.
-%Therefore, each work-item generates one Gaussian random number in each iteration,
-The matrix filling process is the same as that shown in Figure \ref{fig:1}.
+%Therefore, each work-item generates one Gaussian random number in each iteration, The matrix filling process is the same as that shown in Figure \ref{fig:1}.
 
-\begin{figure}[H]%
-    \centering
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f3} }}%
-    \qquad
-    \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
-    \caption{Using streams to produce a $4 \times 6$ matrix of normal random numbers.}%
-    \label{fig3}
-\end{figure}
+% \begin{figure}[H]%
+%     \centering
+%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f3} }}%
+%     \qquad
+%     \subfloat[]{{\includegraphics[width=8cm]{/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/f4} }}%
+%     \caption{Using streams to produce a $4 \times 6$ matrix of normal random numbers.}%
+%     \label{fig3}
+% \end{figure}
 

### 2021-11-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2962 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Random number generation is critical in simulation-based statistical inference, machine learning and many other scientific fields.
-While most random number generators are sequential, the \proglang{R} packages \pkg{parallel}, \pkg{rlecuyer} \citep{sevcikova2015package} and \pkg{future} \citep*{future1.19.1} are able to generate random numbers in parallel on multicore CPUs.
-More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG).
-\pkg{future} also uses the combined MRG algorithm for generating random numbers.
+While most random number generators are sequential, the \proglang{R} packages \pkg{parallel}, pkg{future} \citep*{future1.19.1} and \pkg{rlecuyer} \citep{sevcikova2015package} are able to generate random numbers in parallel on multicore CPUs.
+More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG) MRG32k3a.
+\pkg{future} and \pkg{rlecuyer} also uses the combined MRG algorithm for generating random numbers.
 For up-to-date review papers on the generation of random numbers on parallel devices, and GPUs in particular, see:  \cite{rLEC15a,rLEC17p,rLEC21a}.

### 2021-11-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2963 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.086   0.011   0.530
+##   0.093   0.000   0.524
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.620   0.010   2.162
+##   0.603   0.016   2.151
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.504   0.015  10.211
+##   0.492   0.032  10.217
 \end{verbatim}

### 2021-11-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2963 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Random number generation is critical in simulation-based statistical inference, machine learning and many other scientific fields.
-While most random number generators are sequential, the \proglang{R} packages \pkg{parallel}, pkg{future} \citep*{future1.19.1} and \pkg{rlecuyer} \citep{sevcikova2015package} are able to generate random numbers in parallel on multicore CPUs.
+While most random number generators are sequential, the \proglang{R} packages \pkg{parallel}, \pkg{future} \citep*{future1.19.1} and \pkg{rlecuyer} \citep{sevcikova2015package} are able to generate random numbers in parallel on multicore CPUs.
 More specifically, \pkg{parallel} writes an interface for the \pkg{RngStreams}, a \proglang{\texttt{C++}} library by \cite{l2002object} which is based on a combined multiple-recursive generator (MRG) MRG32k3a.

### 2021-11-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2964 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.093   0.000   0.524
+##   0.096   0.005   0.530
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.603   0.016   2.151
+##   0.618   0.011   2.160
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.492   0.032  10.217
+##   0.496   0.021  10.209
 \end{verbatim}

### 2021-11-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2965 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.096   0.005   0.530
+##   0.083   0.013   0.526
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.618   0.011   2.160
+##   0.610   0.009   2.150
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.496   0.021  10.209
+##   0.485   0.033  10.210
 \end{verbatim}
@@
 \begin{equation*}
-\Sigma(d;\sigma^2,\phi,\kappa)=\sigma^2*\frac{2^{\kappa-1}}{\Gamma(\kappa)} (\sqrt{8\kappa} \frac{d}{\phi})^\kappa  K_\kappa(\sqrt{8\kappa}  \frac{d}{\phi}),  \quad \text{where} \quad \phi    \geq 0,  \kappa     \geq 0,
+\Sigma(d;\sigma^2,\phi,\kappa)=\sigma^2*\frac{2^{\kappa-1}}{\Gamma(\kappa)} (\sqrt{8\kappa} \frac{d}{\phi})^\kappa  K_\kappa(\sqrt{8\kappa}  \frac{d}{\phi}),  \quad \text{where} \quad \phi> 0,  \kappa > 0,
 \end{equation*} 

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2965 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2967 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[shortcuts]{extdash}
+\usepackage{booktabs}
 \renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }
@@
 @
-The 2-way contingency tables %\ref{tab:month} and table \ref{tab:week} 
-consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
 
 \subsection{Comparing run-time: Month data example}\label{fisher_month}
+The 2-way contingency tables consist of selected data from the 2018 Natality public use data \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics (NCHS), the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
+Table \ref{tab:month} is a $12 \times 12$ table that shows frequencies for congenital anomalies of the newborn by birth month in 2018 within the United States.
+The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
+The first 5 rows of the table are displayed below.
  <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
- data2<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv") 
- data2<-data2[,-1]
- month<-as.matrix(data2)
+ datamonth<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv") 
+ month<-as.matrix(datamonth[,-1])
  rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
  colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
- streams <- createStreamsGpu(n =256*64, initial=666)
+ @
+
+ <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
+library(knitr)
+knitr::kable(head(month),format="latex", align = c("rrrrrrrrrrrr"), caption = "Month\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
+  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
 @
@@
  <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE>>=
+ streams <- createStreamsGpu(n =256*64, initial=666)
  almost.1 <- 1 + 64 * .Machine$double.eps
@@
  month_gpu<-vclMatrix(month,type="integer")
- result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
-                    type="double", returnStatistics=TRUE,  Nglobal = c(256,64))
+ system.time(result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
+                    type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
  result_month$simNum
@@
  result_month$p.value
- @
-
- <<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
- #using GPU
- system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
-             type="double", returnStatistics=FALSE, Nglobal = c(256,64)))
  @
@@
 \subsection{Comparing run-time: Week data example}\label{fisher_week}
+Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. The first 5 rows are displayed below.
 <<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
 library(latex2exp)
-data<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
-#exampletwo<-as.matrix(as.data.frame(lapply(data, as.numeric)))
-data<-data[,-1]
-week<-as.matrix(data)
+dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
+week<-as.matrix(dataweek[,-1])
 rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
 colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
+@
+
+<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
+knitr::kable(head(week),format="latex",align = c("rrrrrrrrrrrr"), caption = "Week\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
+  kable_styling(full_width = F, position = "center", latex_options = "HOLD_position")
 @
@@
 \begin{appendix}
+
 %We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
@@
 
-The ``month'' and ``week'' tables for testing \fct{fisher.sim} in section \ref{fisher_month} and \ref{fisher_week} are shown below, where 
-The row variable of the ``month'' table represents the twelve months: from January to December, and for the ``week'' table it represents the seven weekdays: Monday to Sunday.
-The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
-<<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
-library(knitr)
-library(kableExtra)
-knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
-  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
-@
-
-<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
-knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
-  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
-@
 \end{appendix}

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2967 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.083   0.013   0.526
+##   0.140   0.016   0.586
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   4.629   0.348   4.971
+##   7.388   0.432   7.814
 \end{verbatim}
@@
 ##           [,1]      [,2]
-## [1,] 1.22433768 0.7679958
-## [2,] 0.06687963 1.9744549
+## [1,] 0.9340107 5.0051762
+## [2,] 3.4506078 0.8855801
 \end{verbatim}
@@
 
-The 2-way contingency tables %\ref{tab:month} and table \ref{tab:week} 
-consist of selected data from the 2018 Natality public use file \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics, the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
-
 \subsection{Comparing run-time: Month data example}\label{fisher_month}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+The 2-way contingency tables consist of selected data from the 2018 Natality public use data \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics (NCHS), the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
+Table \ref{tab:month} is a $12 \times 12$ table that shows frequencies for congenital anomalies of the newborn by birth month in 2018 within the United States.
+The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
+The first 5 rows of the table are displayed below.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+
+\caption{\label{tab:monthdata}Month\label{tab:month}}
+\centering
+\begin{tabular}[t]{lrrrrrrrrrrrr}
+\toprule
+  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
+\midrule
+Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
+Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
+Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
+Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
+May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
+Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+streams <- createStreamsGpu(n =256*64, initial=666)
 almost.1 <- 1 + 64 * .Machine$double.eps
@@
 month_gpu<-vclMatrix(month,type="integer")
-result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(256,64))
+system.time(result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
+                   type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
+##    user  system elapsed 
+##   1.036   0.017   2.587
 result_month$simNum
@@
 \end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using GPU
-system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
-            type="double", returnStatistics=FALSE, Nglobal = c(256,64)))
-##    user  system elapsed 
-##   0.610   0.009   2.150
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
@@
 stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
-## [1] 0.4039588
+## [1] 0.4036733
 system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
 ##    user  system elapsed 
-##  10.184   0.004  10.178
+##  18.468   0.016  18.468
 \end{verbatim}
@@
 \subsection{Comparing run-time: Week data example}\label{fisher_week}
+Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. The first 5 rows are displayed below.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+
+\caption{\label{tab:weekdata}Week\label{tab:week}}
+\centering
+\begin{tabular}[t]{lrrrrrrrrrrrr}
+\toprule
+  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
+\midrule
+Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
+Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
+Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
+Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
+Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
+Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\end{knitrout}
 
@@
 result_weekgpu$counts
-## [1] 1205
+## [1] 1193
 result_weekgpu$p.value
-## [1] 0.000120472
+## [1] 0.0001192733
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.485   0.033  10.210
+##   0.851   0.020  10.580
 \end{verbatim}
@@
 fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-## [1] 0.0001252669
+## [1] 0.0001249672
 system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
 ##    user  system elapsed 
-##  62.503   0.052  62.503
+##  64.151   0.080  64.172
 \end{verbatim}
@@
 
-The ``month'' and ``week'' tables for testing \fct{fisher.sim} in section \ref{fisher_month} and \ref{fisher_week} are shown below, where 
-The row variable of the ``month'' table represents the twelve months: from January to December, and for the ``week'' table it represents the seven weekdays: Monday to Sunday.
-The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
-
-\caption{\label{tab:monthdata}Month\label{tab:month}}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
-\hline
-Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
-\hline
-Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
-\hline
-Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
-\hline
-Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
-\hline
-May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
-\hline
-Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
-\hline
-Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
-\hline
-Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
-\hline
-Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
-\hline
-Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
-\hline
-Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
-\hline
-Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
-\hline
-\end{tabular}
-\end{table}
-
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
-
-\caption{\label{tab:weekdata}Week\label{tab:week}}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
-\hline
-Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
-\hline
-Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
-\hline
-Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
-\hline
-Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
-\hline
-Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
-\hline
-Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
-\hline
-Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
-\hline
-\end{tabular}
-\end{table}
-
-\end{knitrout}
 \end{appendix}

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2968 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-const int index = get_global_id(0)*get_global_size(1) + get_global_id(1);
+const int index = get_global_id(0)*get_global_size(1) + 
+                  get_global_id(1);
 int Drow, Dcol, DrowStart, Dentry, DrowBlock, DcolBlock, DrowInBounds;
@@
 
-Now we use the streams created in section \ref{createstreams} to generate two vectors of double-precision i.i.d.
+Now we use the streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d.
 $U(0,1)$ random numbers with the function \fct{runif}.
@@
   \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
-  \item \code{verbose}: verbose print extra information if >1.
+  \item \code{verbose}: print extra information if $\text{verbose} > 1$.
 \end{itemize}
@@
 \code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
-\code{PI_2} equals to $\pi/2$.
+\code{PI_2} is defined to be $\pi/2$.
 As the work-items in a work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
@@
 \end{itemize}
+Users request $N$ number of replicates, while the actual number of replicates to be executed on GPU is \code{ceiling(N/prod(Nglobal))*prod(Nglobal)}.
 We show the advantage of \fct{clrng::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
@@
  @
-We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
+We got 409524 cases whose test statistics are below the observed threshold when there are an actual number of 1015808 simulations executed, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
 \fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
@@
 \subsection{Comparing run-time: Week data example}\label{fisher_week}
-Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. The first 5 rows are displayed below.
+Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
 <<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
@@
 <<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
-knitr::kable(head(week),format="latex",align = c("rrrrrrrrrrrr"), caption = "Week\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
+knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"), caption = "Week\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
   kable_styling(full_width = F, position = "center", latex_options = "HOLD_position")
@@
 \centering
-<<maternplot, echo=TRUE,eval=TRUE, fig.show='hold', dev='pdf',cache=FALSE,dev='png'>>=
+<<maternplot, echo=FALSE,eval=TRUE, fig.show='hold', dev='pdf',cache=FALSE,dev='png'>>=
 par(mar=rep(0.1, 4))

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2968 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.140   0.016   0.586
+##   0.149   0.008   0.587
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.036   0.017   2.587
+##   1.064   0.008   2.607
 result_month$simNum
@@
 \end{knitrout}
-We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 409524 cases whose test statitsic is below the observed threshold, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40448.
+We got 409524 cases whose test statistics are below the observed threshold when there are an actual number of 1015808 simulations executed, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
 \fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
@@
 \subsection{Comparing run-time: Week data example}\label{fisher_week}
-Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. The first 5 rows are displayed below.
+Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States.
 
@@
 Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
+Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
 \bottomrule
@@
 ##    user  system elapsed 
-##   0.851   0.020  10.580
+##   0.535   0.016  10.259
 \end{verbatim}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-par(mar=rep(0.1, 4))
-for(D in names(simRaster)) {
-  plot(extent(simRaster))
-  plot(simRaster[[D]], legend=FALSE, add=TRUE)
-}
-\end{verbatim}
-\end{kframe}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
 \includegraphics[width=0.45\textwidth]{figure/maternplot-1} 

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2969 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
-  \item \code{verbose}: print extra information if $\text{verbose} > 1$.
+  \item \code{verbose}: print extra information if $\text{verbose} > 1$. Default is FALSE.
 \end{itemize}
@@
 ##    user  system elapsed 
-##   0.149   0.008   0.587
+##   0.148   0.008   0.585
 \end{verbatim}
@@
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.
+  \item \code{verbose}: print extra information if $\text{verbose} > 1$. Default is FALSE.
 \end{itemize}
@@
 ##    user  system elapsed 
-##   1.064   0.008   2.607
+##   1.030   0.024   2.590
 result_month$simNum
@@
 ##    user  system elapsed 
-##   0.535   0.016  10.259
+##   0.874   0.030  10.610
 \end{verbatim}

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2970 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.148   0.008   0.585
+##   0.154   0.003   0.588
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.030   0.024   2.590
+##   1.047   0.016   2.597
 result_month$simNum
@@
 ##    user  system elapsed 
-##   0.874   0.030  10.610
+##   0.887   0.020  10.614
 \end{verbatim}
@@
 ## [1] 24000  4800
-gpuBatchMatrix::maternBatch(maternCov, coordsGpu, paramsGpu,
+gpuBatchMatrix::maternBatch(var=maternCov, coords=coordsGpu, 
+                            param=paramsGpu,
                             Nglobal=c(128,64), Nlocal=c(16,4))
@@
 \end{knitrout}
-Step 4, the first argument \code{maternCov} in \fct{gpuBatchMatrix::cholBatch} specifies the matrix object to take Cholesky decomposition.
-Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
+Step 4, the first argument \code{A} in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
+Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}.
+Each row of \code{diagMat} stores the diagonal values of each $D_i$, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
 \begin{knitrout}
@@
 diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-gpuBatchMatrix::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
-                     Nglobal= c(128, 8),
-                     Nlocal= c(32, 8),
+gpuBatchMatrix::cholBatch(A=maternCov, 
+                          D=diagMat, 
+                          numbatchD=nrow(myParamsBatch),
+                          Nglobal= c(128, 8), Nlocal= c(32, 8),
                           NlocalCache=1000)
@@
 \end{knitrout}
-% \begin{gather}
-% \begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
-% \end{bmatrix}
-%  \rightarrow
-%  \begin{bmatrix}
-%   L_{1} \\ L_{2} \\L_{3} \\ \vdots
-%   \end{bmatrix} \text{and}
-%   \begin{bmatrix}
-%   D_{1} \\ D_{2} \\D_{3} \\ \vdots
-%   \end{bmatrix}
-% \end{gather}
+\begin{gather}
+\begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
+\end{bmatrix}
+ \rightarrow
+ \begin{bmatrix}
+  L_{1} \\ L_{2} \\L_{3} \\ \vdots
+  \end{bmatrix} \text{and}
+  \begin{bmatrix}
+  D_{1} \\ D_{2} \\D_{3} \\ \vdots
+  \end{bmatrix}
+\end{gather}
 Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
@@
 gpuBatchMatrix::multiplyLowerDiagonalBatch(
-  simMat, maternCov, diagMat, zmatGpu,
+  output=simMat, L=maternCov, 
+  D=diagMat, B=zmatGpu,
   diagIsOne = 1L, # diagonal of L is one
@@
    L_2D_2Z_{11} & L_2D_2Z_{12} \\
-   L_3D_3Z_{11} & L_3D_3Z_{12} 
+   L_3D_3Z_{11} & L_3D_3Z_{12} \\
+   \vdots  &   \vdots
   \end{bmatrix}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+par(mar=rep(0.1, 4))
+for(D in names(simRaster)) {
+  plot(extent(simRaster))
+  plot(simRaster[[D]], legend=FALSE, add=TRUE)
+}
+\end{verbatim}
+\end{kframe}
 \includegraphics[width=0.45\textwidth]{figure/maternplot-1} 

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2969 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2970 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 dim(maternCov)
-gpuBatchMatrix::maternBatch(maternCov, coordsGpu, paramsGpu,
+gpuBatchMatrix::maternBatch(var=maternCov, coords=coordsGpu, 
+                            param=paramsGpu,
                             Nglobal=c(128,64), Nlocal=c(16,4))
 @
-Step 4, the first argument \code{maternCov} in \fct{gpuBatchMatrix::cholBatch} specifies the matrix object to take Cholesky decomposition.
-Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
+Step 4, the first argument \code{A} in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
+Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}.
+Each row of \code{diagMat} stores the diagonal values of each $D_i$, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
 <<r cholBach, eval=TRUE>>=
 diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-gpuBatchMatrix::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
-                     Nglobal= c(128, 8),
-                     Nlocal= c(32, 8),
+gpuBatchMatrix::cholBatch(A=maternCov, 
+                          D=diagMat, 
+                          numbatchD=nrow(myParamsBatch),
+                          Nglobal= c(128, 8), Nlocal= c(32, 8),
                           NlocalCache=1000)
 @
-% \begin{gather}
-% \begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
-% \end{bmatrix}
-%  \rightarrow
-%  \begin{bmatrix}
-%   L_{1} \\ L_{2} \\L_{3} \\ \vdots
-%   \end{bmatrix} \text{and}
-%   \begin{bmatrix}
-%   D_{1} \\ D_{2} \\D_{3} \\ \vdots
-%   \end{bmatrix}
-% \end{gather}
+\begin{gather}
+\begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
+\end{bmatrix}
+ \rightarrow
+ \begin{bmatrix}
+  L_{1} \\ L_{2} \\L_{3} \\ \vdots
+  \end{bmatrix} \text{and}
+  \begin{bmatrix}
+  D_{1} \\ D_{2} \\D_{3} \\ \vdots
+  \end{bmatrix}
+\end{gather}
 Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
@@
 \centering
-<<maternplot, echo=FALSE,eval=TRUE, fig.show='hold', dev='pdf',cache=FALSE,dev='png'>>=
+<<maternplot, echo=TRUE,eval=TRUE, fig.show='hold', dev='pdf',cache=FALSE,dev='png'>>=
 par(mar=rep(0.1, 4))

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2971 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = c(256,64))
+system.time(result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
+                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
 result_weekgpu$simNum
@@
 result_weekgpu$p.value
-@
-
-<<TimecompareweekGpu, eval=TRUE, cache=FALSE>>=
-#using GPU
-system.time(clrng::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = c(256,64)))
 @

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2971 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.154   0.003   0.588
+##   0.180   0.008   0.617
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.047   0.016   2.597
+##   1.075   0.012   2.621
 result_month$simNum
@@
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = c(256,64))
+system.time(result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
+                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
+##    user  system elapsed 
+##   0.849   0.017  10.583
 result_weekgpu$simNum
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using GPU
-system.time(clrng::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = c(256,64)))
-##    user  system elapsed 
-##   0.887   0.020  10.614
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value.

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2972 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
 ##    user  system elapsed 
-##   0.154   0.003   0.588
+##   0.155   0.000   0.585
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.047   0.016   2.597
+##   1.057   0.004   2.596
 result_month$simNum
@@
 \end{knitrout}
-We got 409524 cases whose test statistics are below the observed threshold when there are an actual number of 1015808 simulations executed, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
+We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations executed, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
 \fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
@@
 #using CPU
-stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
+system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
+##    user  system elapsed 
+##  18.606   0.027  18.626
+result_monthcpu$p.value
 ## [1] 0.4036733
-system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
-##    user  system elapsed 
-##  18.468   0.016  18.468
 \end{verbatim}
@@
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = c(256,64))
+system.time(result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
+                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
+##    user  system elapsed 
+##   0.925   0.012  10.653
 result_weekgpu$simNum
@@
 ## [1] 0.0001192733
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using GPU
-system.time(clrng::fisher.sim(week_GPU, 1e7, streams=streams,
-            type="double", returnStatistics=FALSE,Nglobal = c(256,64)))
-##    user  system elapsed 
-##   0.887   0.020  10.614
 \end{verbatim}
@@
 #using CPU
-fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
+system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
+##    user  system elapsed 
+## 117.213   0.152 117.289
+result_weekcpu$p.value
 ## [1] 0.0001249672
-system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
-##    user  system elapsed 
-##  64.151   0.080  64.172
 \end{verbatim}

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2972 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
- colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
+ colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
  @
@@
  @
-We got 409524 cases whose test statistics are below the observed threshold when there are an actual number of 1015808 simulations executed, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
+We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
 \fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
@@
  #using CPU
- stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
- system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
+ system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
+ result_monthcpu$p.value
  @
@@
 rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
-colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
+colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
 @
@@
 #using CPU
-fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
-system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
+system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
+result_weekcpu$p.value
 @

### 2021-11-16 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2973 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.155   0.000   0.585
+##   0.156   0.000   0.586
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   7.388   0.432   7.814
+##   6.951   0.444   7.390
 \end{verbatim}
@@
 ##           [,1]        [,2]
-## [1,] 0.9340107 5.0051762
-## [2,] 3.4506078 0.8855801
+## [1,] 0.2627643 0.005952261
+## [2,] 0.4535731 0.186351621
 \end{verbatim}
@@
 \toprule
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
+  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Dow & Chr & Hyp\\
 \midrule
@@
 ##    user  system elapsed 
-##   1.057   0.004   2.596
+##   1.023   0.028   2.587
 result_month$simNum
@@
 \end{knitrout}
-We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations executed, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
+We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
 \fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
@@
 ##    user  system elapsed 
-##  18.606   0.027  18.626
+##  18.341   0.008  18.336
 result_monthcpu$p.value
-## [1] 0.4036733
+## [1] 0.4039637
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.925   0.012  10.653
+##   0.869   0.013  10.601
 result_weekgpu$simNum
@@
 ##    user  system elapsed 
-## 117.213   0.152 117.289
+## 115.767   0.111 115.784
 result_weekcpu$p.value
-## [1] 0.0001249672
+## [1] 0.0001297621
 \end{verbatim}

### 2021-11-17 pkg/gpuRandom/R/likfit_givenBetas.R
mamapply had prob
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2975 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -6,12 +6,12 @@
 
 likfit_givenBeta <- function(Betas, #a p x m matrix  given by the user 
                              #DatasetIndex=NA,     #an integer, which dataset are the betas for?,
-                             variances,
+                             variances=NULL,
                              Nobs,  # number of observations.
-                             detVar,
-                             ssqY,
-                             XVYXVX,
-                             jacobian,
+                             detVar, # vclVector
+                             ssqY,   # vclMatrix
+                             XVYXVX,   # vclMatrix
+                             jacobian, # vclVector
                              form = c("loglik", "profileforBeta"),
                              minustwotimes=TRUE){
 
@@ -34,52 +34,52 @@ likfit_givenBeta <- function(Betas, #a p x m matrix  given by the user
 
   XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
   ssqForBetas <- matrix(0, nrow=Nparam, ncol=m)
+  likForBeta<- matrix(0, nrow=1, ncol=m)
   #maximized over lambda
   minusTwoLogLikOverLambda <- matrix(0, nrow=Nparam, ncol=m)    #row is param, col is Betas
   index <- matrix(0, nrow=m, ncol=2) #row is Betas, col is row & col index 
   
-  
-  for(beta in 1:m){  
-    
   param_vec  <- rep(seq_len(Nparam), each = Ndata)
   lambda_vec <- rep(seq_len(Ndata), times = Nparam)
-
-  midItem <- parallel::mcmapply(param_vec, lambda_vec,
-                      FUN = function(i, j) {
-                        # This is where expensive operations should go
-                        t(XVYXVX[((i-1)*Ncov+1) : (i*Ncov), j] ) %*% Betas[ ,beta]
-                      }
-                      )
-
-  # computed was a vector, but we need to put it in the correct-size matrix
-  midItem <- matrix(midItem, ncol = Ndata, byrow=TRUE)
+  param_vec  <- seq_len(Nparam)
+  midItem <- matrix(-66, nrow=Nparam, ncol=Ndata)
+  
[…]

### 2021-11-22 pkg/gpuRandom/src/crossprodBatch.cpp
fixed crossprod
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2985 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -48,7 +48,7 @@ std::string crossprodBatchString(
     "#define NstartC " + std::to_string(NstartC) + "\n"   
     "#define NstartA " + std::to_string(NstartA) + "\n" 
     "#define NstartD " + std::to_string(NstartD) + "\n" 
-//    "#define NpadLocal " + std::to_string(Nlocal[1]) + "\n"    
+    //    "#define NpadLocal " + std::to_string(Nlocal[1]) + "\n"    
     "#define NpadBetweenMatricesC " + std::to_string(NpadBetweenMatricesC) + "\n"    
     "#define NpadBetweenMatricesA " + std::to_string(NpadBetweenMatricesA) + "\n"    
     "#define NrowStop " + std::to_string(NrowStop) + "\n"    
@@ -64,12 +64,12 @@ std::string crossprodBatchString(
   }
   result +=
     " __local " + typeString + " *localCache,\n"
- " int NrowStartC,\n"
-  " int Nmatrix) {\n\n"
-  
-  "local " + typeString + " *Acache=localCache;\n" 
-  "local " + typeString + " *Dcache = &localCache[" + 
-    std::to_string(NlocalCacheD) + "];\n" 
+    " int NrowStartC,\n"
+    " int Nmatrix) {\n\n"
+    
+    "local " + typeString + " *Acache=localCache;\n" 
+    "local " + typeString + " *Dcache = &localCache[" + 
+      std::to_string(NlocalCacheD) + "];\n" 
   "local " + typeString + " *Ccache = &localCache[" + 
     std::to_string(2*NlocalCacheD) + "];\n" +
     
@@ -82,13 +82,13 @@ std::string crossprodBatchString(
   "const int AHereInc = get_num_groups(1)*NpadBetweenMatricesA;\n";
   
   
-  // if(onlyDiagC) {
-  //   result +=
-  //     "const int CHereInc = get_num_groups(1)*NpadC;\n"; // made changes here
-  // }else{
-  //   result +=
-  //     "const int CHereInc = get_num_groups(1)*NpadBetweenMatricesC;\n"; // made changes here
-  // }
+  if(onlyDiagC) {
+    result +=
+      "const int CHereInc = get_num_groups(1)*NpadC;\n"; // made changes here
+  }else{
+    result +=
[…]

### 2021-11-22 pkg/gpuRandom/R/likLgm_betaProfile.R
fixed crossprod
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2987 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,4 +1,5 @@
 #' @title Estimate Log-likelihood for Gaussian random fields when Betas are given
+#' @import data.table
 #' @useDynLib gpuRandom
 #' @export
 
@@ -66,7 +67,22 @@ likLgm_betaProfile <- function(Betas, #a m x p R matrix  given by the user
     
   }
   
+  resultbeta1 = data.table::as.data.table(cbind(-0.5*likForBetas, Betas[,"beta1"]))
+  colnames(resultbeta1) <- c("lik", "beta1")
+  profileLogLik_beta1 <- resultbeta1[, .(profile=max(.SD)), by=beta1]
+  
+  
+  resultintercept = data.table::as.data.table(cbind(-0.5*likForBetas, Betas[,1]))
+  colnames(resultintercept) <- c("lik", "intercept")
+  profileLogLik_intercept <- resultintercept[, .(profile=max(.SD)), by=intercept]
+  
+  
+  #plot(profileLogLik_beta1$beta1,profileLogLik_beta1$profile)
+  #plot(profileLogLik_intercept$intercept,profileLogLik_intercept$profile)
+  
   Theoutput <- list(likForBetas=likForBetas,
+                    profileLogLik_beta1 =profileLogLik_beta1,
+                    resultintercept=resultintercept,
                     ssqForBetas = ssqForBetas)
   
   
@@ -77,6 +93,25 @@ likLgm_betaProfile <- function(Betas, #a m x p R matrix  given by the user
   
   
   
-  
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
   
   

### 2021-11-23 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
gpu paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2991 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   __global int* streams,
-  __global float* out){
+  __global float* out,
+  int Nrow, int Ncol, int Npad){
 
-const int index = get_global_id(0)*get_global_size(1) + 
-                  get_global_id(1);
-int Drow, Dcol, DrowStart, Dentry, DrowBlock, DcolBlock, DrowInBounds;
-const int DrowStartInc = get_global_size(0) * NpadCol;
+int Drow, Dcol;
 uint g1[3], g2[3];
-const int startvalue=index * NpadStreams;
 float temp, fact = mrg31k3p_NORM_cl;
+
 streamsToPrivate(streams,g1,g2,startvalue);
 
-// filling in random numbers in the output matrix
-for(DrowBlock = 0, Drow=get_global_id(0), DrowStart = Drow * NpadCol;
-    DrowBlock < Nrow;
-    Drow += get_global_size(0), 
-    DrowBlock +=get_global_size(0), DrowStart += DrowStartInc){
+for(Drow=get_global_id(0); Drow < Nrow;
+    Drow += get_global_size(0)){
 
-    DrowInBounds = Drow < Nrow;
+    for(Dcol=get_global_id(1); DcolBlock < Ncol; 
+        Dcol += get_global_size(1)){
 
-    for(DcolBlock = 0, Dcol=get_global_id(1), Dentry = DrowStart + Dcol;
-        DcolBlock < Ncol;
-        DcolBlock += get_global_size(1), Dentry += get_global_size(1)){
         temp = fact * clrngMrg31k3pNextState(g1, g2);
-        if(DrowInBounds) out[Dentry] = temp;
+        out[Drow * Npad + Dcol] = temp;
+
   }//Dcol
-
 }//Drow
@@
 To view the generated random numbers, we need to convert them to R vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
-<<simStuff2, eval=TRUE,tidy=TRUE>>=
-as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(1,nrow(myStreamsGpu)), type = "double"))
+<<simStuff2, eval=TRUE,tidy=FALSE>>=
+as.vector(
+  clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
 @
@@
   __global int* streams,
-  __global double* out){
+  __global double* out,
+  int Nrow, int Ncol, int Npad, int NpadStreams){
 
-(*@\centerline{\raisebox{-1pt}[0pt][0pt]{$\vdots$}}@*)
+int Drow, Dcol;
+uint g1[3], g2[3];
+local double part[2];
+int startvalue = (
+    get_global_id (0)∗ get_global_size (1) + get_global_id (1)
+  ) * NpadStreams;
 
+double sinOrCosPart1, addForSine = - get_local_id(1) * PI_2;
 double temp, fact = mrg31k3p_NORM_cl;
-local double part[2];
-double sinOrCosPart1, addForSine = - get_local_id(1) * PI_2;
 if(get_local_id(1)){
@@
 }
-streamsToPrivate(streams,g1,g2,startvalue);
 
-for(DrowBlock = 0, Drow=get_global_id(0), DrowStart = Drow * NpadCol;
-    DrowBlock < Nrow;
-    Drow += get_global_size(0), DrowBlock +=get_global_size(0), 
-    DrowStart += DrowStartInc) {
+streamsToPrivate(streams,g1,g2, startvalue);
 
-    DrowInBounds = Drow < Nrow;
-    for(DcolBlock = 0, Dcol=get_global_id(1), Dentry = DrowStart + Dcol;
-        DcolBlock < Ncol;
-        DcolBlock += get_global_size(1), Dentry += get_global_size(1) ){
+for(Drow=get_global_id(0); Drow < Nrow;
+    Drow += get_global_size(0)){
+
+    for(Dcol=get_global_id(1); DcolBlock < Ncol; 
+        Dcol += get_global_size(1)){
 
@@
       sinOrCosPart1 = cos(part[1] + addForSine);
-      temp = part[0]*sinOrCosPart1;
-      if(DrowInBounds) out[Dentry] = temp;
+      out[Drow * Npad + Dcol] = part[0]*sinOrCosPart1;
+
       barrier(CLK_LOCAL_MEM_FENCE);
-
   }//Dcol
@@
 
-Below is an example that we produce a 2 by 2 matrix of exponential random number with expectation 1.
+Below is an example that we produce a matrix of exponential random number with expectation 1.
 <<random exponential, eval=TRUE, cache=FALSE>>=
-r_matrix <- clrng::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
+r_matrix <- clrng::rexp(c(2,4), rate=1, Nglobal=c(2,2), type="double")
 as.matrix(r_matrix)
@@
 library(knitr)
-knitr::kable(head(month),format="latex", align = c("rrrrrrrrrrrr"), caption = "Month\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
+knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), 
+  caption = "Monthly birth anomaly data\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
   kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@@
  streams <- createStreamsGpu(n =256*64, initial=666)
- almost.1 <- 1 + 64 * .Machine$double.eps
- observed_m= -logfactSum(month, c(16,16))/almost.1
+# almost.1 <- 1 + 64 * .Machine$double.eps
+# observed_m= -logfactSum(month, c(16,16))/almost.1
 
  month_gpu<-vclMatrix(month,type="integer")
- system.time(result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
+ system.time(result_month <- clrng::fisher.sim(month_gpu, 1e6, streams=streams,
                     type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
@@
  #using CPU
- system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
+ system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e6))
  result_monthcpu$p.value
@@
 <<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
-knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"), caption = "Week\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
+knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"), caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
   kable_styling(full_width = F, position = "center", latex_options = "HOLD_position")
@@
 <<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
-observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
+#observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
 week_GPU<-gpuR::vclMatrix(week,type="integer")
@@
 dt <- data.frame(
-  var1 = c("B", "1m", "10m", "1m", "10m"),
+  var1 = c("B", "1M", "10M", "1M", "10M"),
   var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 10.74,  63.24),
@@
 \begin{gather*} 
-\mu(x) = \E (U(x)),\\
-\Sigma_{ij} = \cov (U(x_i),U(x_j)).
+[U(s_1) \ldots U(s_N)] \sim \text{MVN}(0, \Sigma) \\
+\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|s_i = s_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|s_i = s_j|}{\phi}\right)
 \end{gather*}
+
 The common choice for the covariance function $\Sigma$ is the Mat\'ern covariance function \citep{matern1960spatial}. The Mat\'ern covariance between $U(x)$ and $U(x')$ takes the form 
-\begin{equation*}
-\Sigma(d;\sigma^2,\phi,\kappa)=\sigma^2*\frac{2^{\kappa-1}}{\Gamma(\kappa)} (\sqrt{8\kappa} \frac{d}{\phi})^\kappa  K_\kappa(\sqrt{8\kappa}  \frac{d}{\phi}),  \quad \text{where} \quad \phi> 0,  \kappa > 0,
-\end{equation*} 
 $d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
@@
 <<setup paramsBatch, eval=TRUE,echo=FALSE>>=
-myParamsBatch = 
+params = 
 rbind(c(shape=1.25, range=0.5, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
@@
 <<show parameterBatch,eval=TRUE>>=
-myParamsBatch
-paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
+params
+#paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
 @
@@
 <<r outputBatchCreate, eval=TRUE>>=
-maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
-                         nrow(coordsGpu),type=theType)
+#maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
+#                         nrow(coordsGpu),type=theType)
+maternCov = gpuBatchMatrix::maternBatch(
+  params, coordsSp,           
+  Nglobal=c(128,64), Nlocal=c(16,4))
 dim(maternCov)
-gpuBatchMatrix::maternBatch(var=maternCov, coords=coordsGpu, 
-                            param=paramsGpu,
-                            Nglobal=c(128,64), Nlocal=c(16,4))
 @
@@
 <<r cholBach, eval=TRUE>>=
-diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-gpuBatchMatrix::cholBatch(A=maternCov, 
-                          D=diagMat, 
-                          numbatchD=nrow(myParamsBatch),
-                          Nglobal= c(128, 8), Nlocal= c(32, 8),
-                          NlocalCache=1000)
+# diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
+diagMat = gpuBatchMatrix::cholBatch(A=maternCov,
+#                          D=diagMat, 
+                          numbatchD=nrow(params),
+                          Nglobal= c(128, 8), Nlocal= c(32, 8)
+                        )
 @

### 2021-11-23 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
gpu paper again
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2992 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<preliminaries, echo=FALSE, results="hide">>=
+  knitr::knit_hooks$set(plot=knitr::hook_plot_tex)  
   options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
@@
 <<setupcoords, eval=TRUE>>=
-Ngrid = c(60, 80)
+Ngrid = 80
 NlocalCache = 1000
@@
 theType = "double"
+@
 
-myRaster = raster::raster( raster::extent(0,Ngrid[1]/Ngrid[2],5,6),Ngrid[1], Ngrid[2])
-coordsSp = sp::SpatialPoints(raster::xyFromCell(myRaster, 1:raster::ncell(myRaster)))
-coordsGpu = vclMatrix(coordsSp@coords, 
-                      nrow(coordsSp@coords), 
-                      ncol(coordsSp@coords), type=theType)
+<<swissRainBoundary, echo=TRUE>>=
+data("swissRain", package="geostatsp")
+myRaster = geostatsp::squareRaster(swissBorder, Ngrid)
+myRaster
 @
@@
 maternCov = gpuBatchMatrix::maternBatch(
-  params, coordsSp,           
+  params, myRaster,          
   Nglobal=c(128,64), Nlocal=c(16,4))
@@
 # diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-diagMat = gpuBatchMatrix::cholBatch(A=maternCov,
+diagMat = gpuBatchMatrix::cholBatch(
+  maternCov,
 #                          D=diagMat, 
-                          numbatchD=nrow(params),
+#                          numbatchD=nrow(params),
                           Nglobal= c(128, 8), Nlocal= c(32, 8)
@@
 <<r plotSimSetup, echo=TRUE, eval=TRUE, cache=FALSE>>=
-library(raster)
-simRaster = brick(myRaster, nl = ncol(simMat)*nrow(paramsGpu))
+simRaster = raster::brick(myRaster, nl = ncol(simMat)*nrow(params))
 values(simRaster) = as.vector(as.matrix(simMat))
@@
 @
-\begin{figure}[H]
-\centering
-<<maternplot, echo=TRUE,eval=TRUE, fig.show='hold', dev='pdf',cache=FALSE,dev='png'>>=
-par(mar=rep(0.1, 4))
+
+<<maternplot, echo=FALSE, eval=TRUE, cache=FALSE,dev='png', out.width = '0.45\\textwidth', fig.height=3, fig.width=6, fig.cap="Simulated Gaussian random fields", fig.subcap = names(simRaster)>>=
+myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=0)
 for(D in names(simRaster)) {
-  plot(extent(simRaster))
-  plot(simRaster[[D]], legend=FALSE, add=TRUE)
+  mapmisc::map.new(simRaster)
+  plot(simRaster[[D]], legend=FALSE, add=TRUE, col=myCol$col, breaks=myCol$breaks)
+  plot(swissBorder, add=TRUE)
 }
+mapmisc::legendBreaks("right", myCol)
 @
-\caption{Simulated Gaussian surface\label{fig:5} }
-\end{figure}
+
+<<maternplot, echo=TRUE, eval=FALSE>>=
+@
 

### 2021-11-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2993 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
                     type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
+ result_month$threshold
  result_month$simNum
@@
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-system.time(result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
+system.time(result_week<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                 type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
-result_weekgpu$simNum
-result_weekgpu$counts
-result_weekgpu$p.value
+result_week$threshold
+result_week$simNum
+result_week$counts
+result_week$p.value
 @
@@
 <<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
-result_weekgpu$cpu = as.vector(result_weekgpu$sim)
+result_week$cpu = as.vector(result_week$sim)
 @
@@
 hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
-abline(v = observed_m, col = "blue", lwd = 1.5)
+abline(v = result_month$threshold, col = "blue", lwd = 1.5)
 
-hist(result_weekgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
-abline(v = observed_w, col = "blue", lwd = 1.5)
+hist(result_week$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
+abline(v = result_week$threshold, col = "blue", lwd = 1.5)
 @

### 2021-11-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2993 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   E-mail: \email{lecuyer@iro.umontreal.ca}}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+
 \begin{document}
@@
 \begin{verbatim}
-as.vector(clrng::runif(n = 6, streams = myStreamsGpu, Nglobal = c(1, nrow(myStreamsGpu)),
-    type = "double"))
-## [1] 0.7353245 0.6142074 0.1100781 0.6487742 0.3661944 0.1088229
+as.vector(
+  clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
+## [1] 0.7353245 0.8423426 0.6142074 0.2159195 0.1100781 0.8698300
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.156   0.000   0.586
+##   0.094   0.004   0.528
 \end{verbatim}
@@
 
-Below is an example that we produce a 2 by 2 matrix of exponential random number with expectation 1.
+Below is an example that we produce a matrix of exponential random number with expectation 1.
 \begin{knitrout}
@@
 \begin{verbatim}
-r_matrix <- clrng::rexp(c(2,2), rate=1, Nglobal=c(1,4), type="double")
+r_matrix <- clrng::rexp(c(2,4), rate=1, Nglobal=c(2,2), type="double")
 as.matrix(r_matrix)
-##           [,1]        [,2]
-## [1,] 0.2627643 0.005952261
-## [2,] 0.4535731 0.186351621
+##           [,1]        [,2]      [,3]      [,4]
+## [1,] 0.2627643 0.005952261 0.4535731 0.1863516
+## [2,] 0.5783051 0.241706604 1.2820227 0.4889530
 \end{verbatim}
@@
 
-\caption{\label{tab:monthdata}Month\label{tab:month}}
+\caption{\label{tab:monthdata}Monthly birth anomaly data\label{tab:month}}
 \centering
@@
 Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
+Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
+Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
+Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
+Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
+Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
+Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
 \bottomrule
@@
 streams <- createStreamsGpu(n =256*64, initial=666)
-almost.1 <- 1 + 64 * .Machine$double.eps
-observed_m= -logfactSum(month, c(16,16))/almost.1
+# almost.1 <- 1 + 64 * .Machine$double.eps
+# observed_m= -logfactSum(month, c(16,16))/almost.1
 
 month_gpu<-vclMatrix(month,type="integer")
-system.time(result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
+system.time(result_month <- clrng::fisher.sim(month_gpu, 1e6, streams=streams,
                    type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
 ##    user  system elapsed 
-##   1.023   0.028   2.587
+##   0.593   0.020   2.148
+result_month$threshold
+## [1] -47954.8
 result_month$simNum
@@
 result_month$counts
-## [1] 410825
+## [1] 4.34485e-309
 result_month$p.value
-## [1] 0.4044323
+## [1] 9.84437e-07
 \end{verbatim}
@@
 #using CPU
-system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
+system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e6))
 ##    user  system elapsed 
-##  18.341   0.008  18.336
+##  10.081   0.008  10.084
 result_monthcpu$p.value
-## [1] 0.4039637
+## [1] 0.4039026
 \end{verbatim}
@@
 
-\caption{\label{tab:weekdata}Week\label{tab:week}}
+\caption{\label{tab:weekdata}Day-of-week birth anomaly data\label{tab:week}}
 \centering
@@
 \begin{verbatim}
-observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
+#observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-system.time(result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
+system.time(result_week<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                 type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
 ##    user  system elapsed 
-##   0.869   0.013  10.601
-result_weekgpu$simNum
+##   0.512   0.009  10.241
+result_week$threshold
+## [1] -54989.56
+result_week$simNum
 ## [1] 10010624
-result_weekgpu$counts
-## [1] 1193
-result_weekgpu$p.value
-## [1] 0.0001192733
+result_week$counts
+## [1] 1.266831e-311
+result_week$p.value
+## [1] 9.989386e-08
 \end{verbatim}
@@
 \hline
-\hspace{1em}1m & 0.403804 & 0.403507 & 0.4035606 & 0.403507 & month\\
+\hspace{1em}1M & 0.403804 & 0.403507 & 0.4035606 & 0.403507 & month\\
 \hline
-\hspace{1em}10m & 0.0001251 & 0.0001274 & 0.0001202 & 0.0001274 & week\\
+\hspace{1em}10M & 0.0001251 & 0.0001274 & 0.0001202 & 0.0001274 & week\\
 \hline
@@
 \hline
-\hspace{1em}1m & 10.74 & 2.28 & 15 & 0.72 & month\\
+\hspace{1em}1M & 10.74 & 2.28 & 15 & 0.72 & month\\
 \hline
-\hspace{1em}10m & 63.24 & 10.82 & 91.58 & 4.18 & week\\
+\hspace{1em}10M & 63.24 & 10.82 & 91.58 & 4.18 & week\\
 \hline
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
-\includegraphics[width=0.45\textwidth]{figure/fighistMonth-1} 
-\includegraphics[width=0.45\textwidth]{figure/fighistMonth-2} 
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+
+\includegraphics[width=0.45\textwidth]{figure/fighistMonth-1} \includegraphics[width=0.45\textwidth]{figure/fighistMonth-2} \end{kframe}
 \end{knitrout}
@@
 \begin{verbatim}
-Ngrid = c(60, 80)
+Ngrid = 80
 NlocalCache = 1000
@@
 theType = "double"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
-myRaster = raster::raster( raster::extent(0,Ngrid[1]/Ngrid[2],5,6),Ngrid[1], Ngrid[2])
-coordsSp = sp::SpatialPoints(raster::xyFromCell(myRaster, 1:raster::ncell(myRaster)))
-coordsGpu = vclMatrix(coordsSp@coords, 
-                      nrow(coordsSp@coords), 
-                      ncol(coordsSp@coords), type=theType)
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+data("swissRain", package="geostatsp")
+myRaster = geostatsp::squareRaster(swissBorder, Ngrid)
+myRaster
+## class      : RasterLayer 
+## dimensions : 51, 80, 4080  (nrow, ncol, ncell)
+## resolution : 4358.065, 4358.065  (x, y)
+## extent     : 2485351, 2833996, 1075013, 1297275  (xmin, xmax, ymin, ymax)
+## crs        : +proj=somerc +lat_0=46.9524055555556 +lon_0=7.43958333333333 +k_0=1 +x_0=2600000 +y_0=1200000 +ellps=GRS80 +units=m +no_defs
 \end{verbatim}
@@
 \begin{verbatim}
-myParamsBatch
+params
 ##      shape range variance nugget anisoRatio anisoAngleRadians
@@
 ## [5,]  2.15  0.50      2.0      0          2         0.7853982
-paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
+#paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
 \end{verbatim}
@@
 \begin{verbatim}
-maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
-                         nrow(coordsGpu),type=theType)
-dim(maternCov)
-## [1] 24000  4800
-gpuBatchMatrix::maternBatch(var=maternCov, coords=coordsGpu, 
-                            param=paramsGpu,
+#maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
+#                         nrow(coordsGpu),type=theType)
+maternCov = gpuBatchMatrix::maternBatch(
+  params, myRaster,          
   Nglobal=c(128,64), Nlocal=c(16,4))
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(param): argument "{}param"{} is missing, with no default}}\begin{verbatim}
+dim(maternCov)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'maternCov' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{verbatim}
-diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-gpuBatchMatrix::cholBatch(A=maternCov, 
-                          D=diagMat, 
-                          numbatchD=nrow(myParamsBatch),
-                          Nglobal= c(128, 8), Nlocal= c(32, 8),
-                          NlocalCache=1000)
+# diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
+diagMat = gpuBatchMatrix::cholBatch(
+  maternCov,
+#                          D=diagMat, 
+#                          numbatchD=nrow(params),
+                          Nglobal= c(128, 8), Nlocal= c(32, 8)
+                        )
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(A): object 'maternCov' not found}}\end{kframe}
 \end{knitrout}
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'maternCov' not found}}\end{kframe}
 \end{knitrout}
@@
                    type = theType)
+\end{verbatim}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'zmatGpu' not found}}\begin{verbatim}
 gpuBatchMatrix::multiplyLowerDiagonalBatch(
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuBatchMatrix:::multiplyLowerDiagonalBatchBackend(output, L, : object 'simMat' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{verbatim}
-library(raster)
-simRaster = brick(myRaster, nl = ncol(simMat)*nrow(paramsGpu))
+simRaster = raster::brick(myRaster, nl = ncol(simMat)*nrow(params))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\begin{verbatim}
 values(simRaster) = as.vector(as.matrix(simMat))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'as.matrix': object 'simMat' not found}}\begin{verbatim}
 names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
@@
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'paramsGpu' not found}}\end{kframe}
 \end{knitrout}
-\begin{figure}[H]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-par(mar=rep(0.1, 4))
-for(D in names(simRaster)) {
-  plot(extent(simRaster))
-  plot(simRaster[[D]], legend=FALSE, add=TRUE)
-}
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=0.45\textwidth]{figure/maternplot-1} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-2} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-3} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-4} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-5} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-6} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-7} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-8} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-9} 
-\includegraphics[width=0.45\textwidth]{figure/maternplot-10} 
-\end{knitrout}
-\caption{Simulated Gaussian surface\label{fig:5} }
-\end{figure}
-
-\section{Discussion}
-The package \pkg{clrng} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
-We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
-Most of these functions uses local memory on GPU.
-Comparison of performance between using \pkg{clrng} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
-
-By leveraging the \pkg{gpuR} package, \pkg{clrng} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.
-\pkg{clrng} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for other applications.
-
-%One of the main difficulties in developing the package is bug fixes and uncertain software behaviors.
-Error messages are often vague and it is time-consuming to locate the error within the backend \proglang{C} program.
-The usual steps that we take to debug is commenting out some parts of codes and then recompile many times until we find the problem, or printing out flag messages in several places in the code to find out where the program stops at due to error.
-
-\pkg{clrng} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
-
-Future work on \pkg{clrng} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs.
-(2) Now that the package can do Cholesky decomposition and matrix multiplication in batches on GPU, which is a motivation for us to work on parallel likelihood evaluations on GPU for Gaussian spatial models in the next step.
-(3) Create a ``sparse'' matrix class on GPU and use it for simulating Gaussian random fields with sparse correlation structures such as the Gaussian Markov random fields.
-
-\bibliography{paper1}
-
-\newpage
-
-\begin{appendix}
-
-%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
-The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
-
-%\begin{figure}[H]
-%\centering
-
-%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
-%\end{figure}
-
-%In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
-The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{clrng::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-
-% \begin{figure}[H]
-% \centering
-
-% \caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
-% \end{figure}
-
-\end{appendix}
-
-\end{document}

### 2021-11-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2999 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{document}
-%\SweaveOpts{concordance=FALSE}
+%\SweaveOpts{concordance=TRUE}
 
@@
 <<simStuff2, eval=TRUE,tidy=FALSE>>=
-as.vector(
-  clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
+as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
 @
@@
 # diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-diagMat = gpuBatchMatrix::cholBatch(
-  maternCov,
-#                          D=diagMat, 
-#                          numbatchD=nrow(params),
-                          Nglobal= c(128, 8), Nlocal= c(32, 8)
-                        )
+diagMat = gpuBatchMatrix::cholBatch(maternCov, # D=diagMat,  # numbatchD=nrow(params),
+                          Nglobal= c(128, 8), Nlocal= c(32, 8))
 @
@@
 values(simRaster) = as.vector(as.matrix(simMat))
-names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
+names(simRaster) = apply(expand.grid('par',1:nrow(params), 
                                      'sim', 1:ncol(simMat)), 1, paste, collapse='')
@@
 <<maternplot, echo=FALSE, eval=TRUE, cache=FALSE,dev='png', out.width = '0.45\\textwidth', fig.height=3, fig.width=6, fig.cap="Simulated Gaussian random fields", fig.subcap = names(simRaster)>>=
-myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=0)
+myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=2)
 for(D in names(simRaster)) {

### 2021-11-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2999 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-\documentclass[article,nojss]{jss}
+\documentclass[article,nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
   E-mail: \email{lecuyer@iro.umontreal.ca}}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-%\SweaveOpts{concordance=FALSE}
+%\SweaveOpts{concordance=TRUE}
 
@@
 ##    user  system elapsed 
-##   0.094   0.004   0.528
+##   0.078   0.000   0.078
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   6.951   0.444   7.390
+##   6.658   1.125   7.779
 \end{verbatim}
@@
 ##            [,1]      [,2]     [,3]       [,4]
-## [1,] 0.2627643 0.005952261 0.4535731 0.1863516
-## [2,] 0.5783051 0.241706604 1.2820227 0.4889530
+## [1,] 0.05469212 0.6969336 1.089906 0.78689966
+## [2,] 0.04188041 0.1686520 1.905295 0.03536933
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.593   0.020   2.148
+##   0.317   0.004   0.321
 result_month$threshold
@@
 result_month$counts
-## [1] 4.34485e-309
+## [1] 410825
 result_month$p.value
-## [1] 9.84437e-07
+## [1] 0.4044323
 \end{verbatim}
@@
 ##    user  system elapsed 
-##  10.081   0.008  10.084
+##  14.445   0.036  14.473
 result_monthcpu$p.value
-## [1] 0.4039026
+## [1] 0.4034616
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.512   0.009  10.241
+##   1.926   0.025   1.949
 result_week$threshold
@@
 result_week$counts
-## [1] 1.266831e-311
+## [1] 1193
 result_week$p.value
-## [1] 9.989386e-08
+## [1] 0.0001192733
 \end{verbatim}
@@
 ##    user  system elapsed 
-## 115.767   0.111 115.784
+##  89.981   0.158  90.071
 result_weekcpu$p.value
-## [1] 0.0001297621
+## [1] 0.0001243679
 \end{verbatim}
@@
   Nglobal=c(128,64), Nlocal=c(16,4))
+dim(maternCov)
+## [1] 20400  4080
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(param): argument "{}param"{} is missing, with no default}}\begin{verbatim}
-dim(maternCov)
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'maternCov' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 # diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-diagMat = gpuBatchMatrix::cholBatch(
-  maternCov,
-#                          D=diagMat, 
-#                          numbatchD=nrow(params),
-                          Nglobal= c(128, 8), Nlocal= c(32, 8)
-                        )
+diagMat = gpuBatchMatrix::cholBatch(maternCov, # D=diagMat,  # numbatchD=nrow(params),
+                          Nglobal= c(128, 8), Nlocal= c(32, 8))
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(A): object 'maternCov' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'maternCov' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
                    type = theType)
-\end{verbatim}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'zmatGpu' not found}}\begin{verbatim}
 gpuBatchMatrix::multiplyLowerDiagonalBatch(
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuBatchMatrix:::multiplyLowerDiagonalBatchBackend(output, L, : object 'simMat' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 simRaster = raster::brick(myRaster, nl = ncol(simMat)*nrow(params))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\begin{verbatim}
 values(simRaster) = as.vector(as.matrix(simMat))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'as.matrix': object 'simMat' not found}}\begin{verbatim}
 names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
@@
 \end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{figure}
+\subfloat[layer.1\label{fig:maternplot-1}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-1} }\subfloat[layer.2\label{fig:maternplot-2}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-2} }\subfloat[layer.3\label{fig:maternplot-3}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-3} }\subfloat[layer.4\label{fig:maternplot-4}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-4} }\subfloat[layer.5\label{fig:maternplot-5}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-5} }\subfloat[layer.6\label{fig:maternplot-6}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-6} }\subfloat[layer.7\label{fig:maternplot-7}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-7} }\subfloat[layer.8\label{fig:maternplot-8}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-8} }\subfloat[layer.9\label{fig:maternplot-9}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-9} }\subfloat[layer.10\label{fig:maternplot-10}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-10} }\caption[Simulated Gaussian random fields]{Simulated Gaussian random fields}\label{fig:maternplot}
+\end{figure}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=2)
+for(D in names(simRaster)) {
+  mapmisc::map.new(simRaster)
+  plot(simRaster[[D]], legend=FALSE, add=TRUE, col=myCol$col, breaks=myCol$breaks)
+  plot(swissBorder, add=TRUE)
+}
+mapmisc::legendBreaks("right", myCol)
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\section{Discussion}
+The package \pkg{clrng} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
+We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
+Most of these functions uses local memory on GPU.
+Comparison of performance between using \pkg{clrng} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.
+
+By leveraging the \pkg{gpuR} package, \pkg{clrng} provides a user-friendly interface that bridges \proglang{R} and \proglang{OpenCL}, users can use the facilities in our package without the need to know the complex \proglang{OpenCL} or even the \proglang{\texttt{C++}} code.
+\pkg{clrng} is portable as its backend \proglang{OpenCL} supports multiple types of processors, and it is also flexible as its kernels can be incorporated or reconstructed in other \proglang{R} packages for other applications.
+
+%One of the main difficulties in developing the package is bug fixes and uncertain software behaviors.
+Error messages are often vague and it is time-consuming to locate the error within the backend \proglang{C} program.
+The usual steps that we take to debug is commenting out some parts of codes and then recompile many times until we find the problem, or printing out flag messages in several places in the code to find out where the program stops at due to error.
+
+\pkg{clrng} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
+
+Future work on \pkg{clrng} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs.
+(2) Now that the package can do Cholesky decomposition and matrix multiplication in batches on GPU, which is a motivation for us to work on parallel likelihood evaluations on GPU for Gaussian spatial models in the next step.
+(3) Create a ``sparse'' matrix class on GPU and use it for simulating Gaussian random fields with sparse correlation structures such as the Gaussian Markov random fields.
+
+\bibliography{paper1}
+
+\newpage
+
+\begin{appendix}
+
+%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
+The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
+
+%\begin{figure}[H]
+%\centering
+
+%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
+%\end{figure}
+
+%In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
+The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
+We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{clrng::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
+
+% \begin{figure}[H]
+% \centering
+
+% \caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
+% \end{figure}
+
+\end{appendix}
+
+\end{document}

### 2021-11-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3000 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 params = 
-rbind(c(shape=1.25, range=0.5, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
-c(shape=2.15, range=0.25, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
-c(shape=0.55, range=1.5, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
-c(shape=2.15, range=0.5, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = -pi/7),
-c(shape=2.15, range=0.5, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/4))
+rbind(c(shape=1.25, range=0.5*1e04, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
+c(shape=2.15, range=0.25*1e04, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
+c(shape=0.55, range=1.5*1e04, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
+c(shape=2.15, range=0.5*1e04, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = -pi/7),
+c(shape=2.15, range=0.5*1e04, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/4))
 @

### 2021-11-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3001 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 local double part[2];
-int startvalue = (
-    get_global_id (0)∗ get_global_size (1) + get_global_id (1)
-  ) * NpadStreams;
+int startvalue = (get_global_id(0) * get_global_size(1) + 
+get_global_id(1)) * NpadStreams;
 
@@
 <<maternplot, echo=FALSE, eval=TRUE, cache=FALSE,dev='png', out.width = '0.45\\textwidth', fig.height=3, fig.width=6, fig.cap="Simulated Gaussian random fields", fig.subcap = names(simRaster)>>=
-myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=2)
+myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=1)
 for(D in names(simRaster)) {
@@
 mapmisc::legendBreaks("right", myCol)
-@
-
-<<maternplot, echo=TRUE, eval=FALSE>>=
 @
@@
 
-%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
-The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
-<<uniforms_random_vector,include=FALSE>>=
-streams <- createStreamsGpu(n = 512)
-random_u<-as.vector(runif(n=50000, streams=streams, Nglobal=c(64,8), type="double"))
-@
-%\begin{figure}[H]
-%\centering
-<<largenumberuniforms, include=FALSE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold', dev='png'>>=
-hist(random_u, prob = TRUE, main="Histogram")
-plot(ecdf(random_u), ylab="F(x)",main="Empirical cumulative distribution")
-curve(punif(x, 0, 1), lty=2, add=TRUE, col="red")
-@
-%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
-%\end{figure}
-
-%In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
-The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{clrng::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-<<histAndQQ_prepare,include=FALSE,cache=TRUE>>=
-streams <- createStreamsGpu(n =512*128)
-normal_matrix<-clrng::rnorm(c(1024,512), streams=streams, 
-                                Nglobal=c(512,128), type="float")
-avector<-as.vector(as.matrix(normal_matrix))
-@
-% \begin{figure}[H]
-% \centering
- <<histAndQQ, echo=TRUE, include=FALSE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold',dev='png'>>=
-hist(avector,breaks=40, main = "histogram")
-clrng::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
-@
-% \caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
-% \end{figure}
-
 \end{appendix}

### 2021-11-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3001 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{verbatim}
-as.vector(
-  clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
+as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
 ## [1] 0.7353245 0.8423426 0.6142074 0.2159195 0.1100781 0.8698300
@@
 ##    user  system elapsed 
-##   0.078   0.000   0.078
+##   0.068   0.004   0.072
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.317   0.004   0.321
+##   0.315   0.000   0.314
 result_month$threshold
@@
 ##    user  system elapsed 
-##   1.926   0.025   1.949
+##   1.932   0.015   1.947
 result_week$threshold
@@
 ##      shape range variance nugget anisoRatio anisoAngleRadians
-## [1,]  1.25  0.50      1.5      0          1         0.0000000
-## [2,]  2.15  0.25      2.0      0          4         0.4487990
-## [3,]  0.55  1.50      2.0      0          4         0.4487990
-## [4,]  2.15  0.50      2.0      0          4        -0.4487990
-## [5,]  2.15  0.50      2.0      0          2         0.7853982
+## [1,]  1.25  5000      1.5      0          1         0.0000000
+## [2,]  2.15  2500      2.0      0          4         0.4487990
+## [3,]  0.55 15000      2.0      0          4         0.4487990
+## [4,]  2.15  5000      2.0      0          4        -0.4487990
+## [5,]  2.15  5000      2.0      0          2         0.7853982
 #paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
@@
 values(simRaster) = as.vector(as.matrix(simMat))
-names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
+names(simRaster) = apply(expand.grid('par',1:nrow(params), 
                                      'sim', 1:ncol(simMat)), 1, paste, collapse='')
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'paramsGpu' not found}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 \begin{figure}
-\subfloat[layer.1\label{fig:maternplot-1}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-1} }\subfloat[layer.2\label{fig:maternplot-2}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-2} }\subfloat[layer.3\label{fig:maternplot-3}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-3} }\subfloat[layer.4\label{fig:maternplot-4}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-4} }\subfloat[layer.5\label{fig:maternplot-5}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-5} }\subfloat[layer.6\label{fig:maternplot-6}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-6} }\subfloat[layer.7\label{fig:maternplot-7}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-7} }\subfloat[layer.8\label{fig:maternplot-8}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-8} }\subfloat[layer.9\label{fig:maternplot-9}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-9} }\subfloat[layer.10\label{fig:maternplot-10}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-10} }\caption[Simulated Gaussian random fields]{Simulated Gaussian random fields}\label{fig:maternplot}
+\subfloat[par1sim1\label{fig:maternplot-1}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-1} }\subfloat[par2sim1\label{fig:maternplot-2}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-2} }\subfloat[par3sim1\label{fig:maternplot-3}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-3} }\subfloat[par4sim1\label{fig:maternplot-4}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-4} }\subfloat[par5sim1\label{fig:maternplot-5}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-5} }\subfloat[par1sim2\label{fig:maternplot-6}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-6} }\subfloat[par2sim2\label{fig:maternplot-7}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-7} }\subfloat[par3sim2\label{fig:maternplot-8}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-8} }\subfloat[par4sim2\label{fig:maternplot-9}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-9} }\subfloat[par5sim2\label{fig:maternplot-10}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-10} }\caption[Simulated Gaussian random fields]{Simulated Gaussian random fields}\label{fig:maternplot}
 \end{figure}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=2)
-for(D in names(simRaster)) {
-  mapmisc::map.new(simRaster)
-  plot(simRaster[[D]], legend=FALSE, add=TRUE, col=myCol$col, breaks=myCol$breaks)
-  plot(swissBorder, add=TRUE)
-}
-mapmisc::legendBreaks("right", myCol)
-\end{verbatim}
 \end{kframe}
@@
 
-%We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}).
-The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{runif} does come from a standard uniform distribution.
-
-%\begin{figure}[H]
-%\centering
-
-%\caption{Histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
-%\end{figure}
-
-%In the following code, We create a $1024 \times 512$ matrix of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}.
-The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution.
-We find doing a Q-Q plot can occupy a large amount of time in the process of producing an \proglang{R} markdown file, so as an incidental work, we adapted the \fct{stat::qnorm} to make the function \fct{clrng::qqnorm} that does the quantile computation for the normal distribution part on a GPU device, which can speed up the Q-Q plot.
-
-% \begin{figure}[H]
-% \centering
-
-% \caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
-% \end{figure}
-
 \end{appendix}

### 2021-11-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3002 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-<<maternplot, echo=FALSE, eval=TRUE, cache=FALSE,dev='png', out.width = '0.45\\textwidth', fig.height=3, fig.width=6, fig.cap="Simulated Gaussian random fields", fig.subcap = names(simRaster)>>=
+<<maternplot, echo=FALSE, eval=FALSE, cache=FALSE,dev='png', out.width = '0.45\\textwidth', fig.height=3, fig.width=6, fig.cap="Simulated Gaussian random fields", fig.subcap = names(simRaster)>>=
 myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=1)

### 2021-11-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3002 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.068   0.004   0.072
+##   0.073   0.000   0.072
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.315   0.000   0.314
+##   0.308   0.008   0.316
 result_month$threshold
@@
 ##    user  system elapsed 
-##   1.932   0.015   1.947
+##   1.937   0.011   1.947
 result_week$threshold
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{figure}
-\subfloat[par1sim1\label{fig:maternplot-1}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-1} }\subfloat[par2sim1\label{fig:maternplot-2}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-2} }\subfloat[par3sim1\label{fig:maternplot-3}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-3} }\subfloat[par4sim1\label{fig:maternplot-4}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-4} }\subfloat[par5sim1\label{fig:maternplot-5}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-5} }\subfloat[par1sim2\label{fig:maternplot-6}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-6} }\subfloat[par2sim2\label{fig:maternplot-7}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-7} }\subfloat[par3sim2\label{fig:maternplot-8}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-8} }\subfloat[par4sim2\label{fig:maternplot-9}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-9} }\subfloat[par5sim2\label{fig:maternplot-10}]{\includegraphics[width=0.45\textwidth]{figure/maternplot-10} }\caption[Simulated Gaussian random fields]{Simulated Gaussian random fields}\label{fig:maternplot}
-\end{figure}
-\end{kframe}
-\end{knitrout}
-
 \section{Discussion}

### 2021-11-26 pkg/gpuRandom/R/likfitLgmGpu.R
variance profile
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3003 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -396,6 +396,7 @@ likfitLgmGpu <- function(data,
                       Nparam = Nparam,
                       ssqY=ssqY,     
                       ssqBetahat = ssqBetahat,
+                      ssqResidual = ssqResidual,
                       detVar = detVar,   
                       detReml = detReml,   
                       jacobian = jacobian,

### 2021-11-26 pkg/gpuRandom/R/likLgm_variancePro.R
variance profile
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3003 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,63 @@
+#' @title Estimate Log-likelihood for Gaussian random fields when stderror are given
+#' @import data.table
+#' @useDynLib gpuRandom
+#' @export
+
+
+######## this function will be incorporated in the other function later i think
+
+likLgm_variancePro <- function(stderror, #a vector  given by the user 
+                               Nobs,  # number of observations.
+                               Nparam,
+                               Ndata,
+                               BoxCox,
+                               detVar, # vclVector
+                               ssqResidual,   # vclMatrix
+                               jacobian, # vclVector  #form = c("loglik", "profileforBeta"),
+                               minustwotimes=FALSE){
+  
+  
+  ssqResidual <- as.matrix(ssqResidual)
+  detVar <- as.vector(detVar)
+  detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
+  jacobian <- as.vector(jacobian)
+  jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))
+  m <- length(stderror)
+  LogLik_optimized = matrix(0, nrow=m, ncol=1)
+  
+  
+  
+  for (var in 1:m){
+    All_min2loglik_forthisvar <- ssqResidual/(stderror[var]^2) + Nobs*log(stderror[var]^2) + detVar + jacobian + Nobs*log(2*pi) 
+    LogLik_optimized[var,] <- -0.5*min(All_min2loglik_forthisvar)
+  }
+
+  result = cbind(stderror, LogLik_optimized)
+  colnames(result) <- c("stderror",'LogL')
+  MLE <- result[,'stderror'][which.max(result[,'LogL'])]
+  # plot(result[,'stderror']^2, result[,'LogL'])
+  maximum <- max(LogLik_optimized)
+  breaks95 = maximum - qchisq(0.95,  df = 1)/2
+  
+  
+  leftOfMax = result[,'stderror'] < MLE
+  if(length(which(leftOfMax)) <=2 | length(which(!leftOfMax)) <=2){
+    ci95 = c(NA,NA)
+    print("Not enough data for CI calculation")
+  }else{
+    afLeft <- approxfun(result[,'LogL'][leftOfMax], result[,'stderror'][leftOfMax])   
+    afRight <- approxfun(result[,'LogL'][!leftOfMax], result[,'stderror'][!leftOfMax])   
+  
+    ci95= c(afLeft(breaks95), afRight(breaks95))
+  }
+    
+  
[…]

### 2021-11-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3005 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\pkg{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators.
+\proglang{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators.
 These four RNGs use different types of constructions.
@@
 \code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively.
-\code{NpadCol} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of \code{streams} (these variables are defined in macros not shown here).
-Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, where \code{g1} and \code{g2} point to the first three and second three elements of stream states.
-\code{index} represents the position of the work-item when mapping it to a one-dimensional data structure.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer called \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type.
+\code{Npad} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of \code{streams} (these variables are defined in macros not shown here).
+\code{index} represents the position of a work-item when mapping it to a one-dimensional data structure.
+Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states, respectively, and \code{startvalue} tells the starting position of the particular stream to be copied to the work-item.
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer named \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type.
 If to generate random integers, \code{temp} is not scaled.
@@
 Now we use the streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d.
-$U(0,1)$ random numbers with the function \fct{runif}.
-To view the generated random numbers, we need to convert them to R vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
+$U(0,1)$ random numbers with \fct{clrng::runif}.
+To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
 <<simStuff2, eval=TRUE,tidy=FALSE>>=
@@
 \subsection{Exponential random number generation}
-The Exponential random variates are produced by applying the inverse transform method on i.i.d.
+The exponential random variates are produced by applying the inverse transform method on i.i.d.
 $U(0,1)$ random numbers.
 The random variable $X \sim \text{Exponential}(\lambda)$ has cumulative distribution function $F_{X}(x)=1-e^{-\lambda x}$ for $x \geq 0$ and $\lambda > 0$.
-The inverse of $F_{X}$ is $F^{-1}_{x}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
+The inverse of $F_{X}(\cdot)$ is $F^{-1}_{X}(y)= -(1/ \lambda) \log (1-y)$, for $ 0 \leq y <1$.
 The kernel is not shown because it is mostly same as the kernel for uniform and normal random numbers, except for the part that does the inverse transform.
@@
 
-Below is an example that we produce a matrix of exponential random number with expectation 1.
+Below is an example that produces a matrix of Exponential random numbers with expectation equal to 1.
 <<random exponential, eval=TRUE, cache=FALSE>>=
@@
 \begin{itemize}
-\itemsep0em 
-  \item \code{x}: a ``vclMatrix''.
+\setlength{\parskip}{0pt}
+  \item \code{x}: a contingency table, a ``vclMatrix'' object.
   \item \code{N}: number of simulation runs.
@@
 \end{itemize}
-Users request $N$ number of replicates, while the actual number of replicates to be executed on GPU is \code{ceiling(N/prod(Nglobal))*prod(Nglobal)}.
+Users request $N$ number of replicates, while the actual number of replicates to be executed on GPU is larger (\code{ceiling(N/prod(Nglobal))*prod(Nglobal)}).
 We show the advantage of \fct{clrng::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
 The \proglang{R} outputs for testing on computer 2 is shown in the following.
-<<deviceinfo, include=FALSE,cache=FALSE>>=
-library(kableExtra)
-a <- data.frame(
-  var1 = c(" ", "Computer1", "Computer2"),
-  var2 = c('Host', 'Intel Xenon W-2145 3.7Ghz',     'VCPU Intel Xenon Skylake 2.5Ghz'),
-  var3 = c('Device', 'AMD Radeon VII', 'VGPU Nvidia Tesla V100')
-)
-
-knitr::kable(a, col.names = NULL, align='c', caption = "Device information for run-time comparison.\\label{tab:device}") %>%
-  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position")
-@
 
@@
  streams <- createStreamsGpu(n =256*64, initial=666)
-# almost.1 <- 1 + 64 * .Machine$double.eps
-# observed_m= -logfactSum(month, c(16,16))/almost.1
-
  month_gpu<-vclMatrix(month,type="integer")
@@
  @
-We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
-\fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
+We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
+\fct{clrng::fisher.sim} takes about 0.31 seconds, \pkg{clrng} accounts for 2.17\% of the elapsed time on CPU.
  <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
@@
 <<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
-#observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
 week_GPU<-gpuR::vclMatrix(week,type="integer")
@@
 \caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''.
-The test statistic values of the observed tables are calculated by \fct{clrng::logfactSum} and are marked with a blue line on each plot.
+The test statistic values of the observed tables are marked with a blue line on each plot.
 \label{fig4}}

### 2021-11-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3005 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Now we use the streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d.
-$U(0,1)$ random numbers with the function \fct{runif}.
-To view the generated random numbers, we need to convert them to R vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
+$U(0,1)$ random numbers with \fct{clrng::runif}.
+To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
 \begin{knitrout}
@@
 ##    user  system elapsed 
-##   0.073   0.000   0.072
+##   0.071   0.003   0.073
 \end{verbatim}
@@
 
-Below is an example that we produce a matrix of exponential random number with expectation 1.
+Below is an example that produces a matrix of Exponential random numbers with expectation equal to 1.
 \begin{knitrout}
@@
 \end{itemize}
-Users request $N$ number of replicates, while the actual number of replicates to be executed on GPU is \code{ceiling(N/prod(Nglobal))*prod(Nglobal)}.
+Users request $N$ number of replicates, while the actual number of replicates to be executed on GPU is larger (\code{ceiling(N/prod(Nglobal))*prod(Nglobal)}).
 We show the advantage of \fct{clrng::fisher.sim} by computing the p-values for two real data examples: one with a relatively big p-value and another with a very small p-value, and compare the run-time with using \fct{stats::fisher.test} for each of the data sets on two computers: one with a very good CPU and an ordinary GPU, the other is equipped with an excellent GPU and an ordinary CPU.
@@
 streams <- createStreamsGpu(n =256*64, initial=666)
-# almost.1 <- 1 + 64 * .Machine$double.eps
-# observed_m= -logfactSum(month, c(16,16))/almost.1
-
 month_gpu<-vclMatrix(month,type="integer")
@@
 ##    user  system elapsed 
-##   0.308   0.008   0.316
+##   0.315   0.000   0.315
 result_month$threshold
@@
 \end{knitrout}
-We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
-\fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU.
+We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
+\fct{clrng::fisher.sim} takes about 0.31 seconds, \pkg{clrng} accounts for 2.17\% of the elapsed time on CPU.
 
@@
 \begin{verbatim}
-#observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
 week_GPU<-gpuR::vclMatrix(week,type="integer")
@@
 ##    user  system elapsed 
-##   1.937   0.011   1.947
+##   1.968   0.022   1.991
 result_week$threshold

### 2021-11-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3006 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.071   0.003   0.073
+##   0.072   0.000   0.072
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   0.315   0.000   0.315
+##   0.303   0.012   0.315
 result_month$threshold
@@
 ##    user  system elapsed 
-##   1.968   0.022   1.991
+##   1.950   0.001   1.952
 result_week$threshold
@@
 A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})$ is multivariate Gaussian.
-The expectation $\mu(x)$ and covariance function $\Sigma$ completely determine the distribution of $U$, where
+The expectation of $U$ and covariance function $\Sigma$ completely determine the distribution of $U$, where
 \begin{gather*} 
-[U(s_1) \ldots U(s_N)] \sim \text{MVN}(0, \Sigma) \\
-\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|s_i = s_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|s_i = s_j|}{\phi}\right)
+U = [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma) \\
+\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|x_i = x_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|x_i = x_j|}{\phi}\right).
 \end{gather*}
-
-The common choice for the covariance function $\Sigma$ is the Mat\'ern covariance function \citep{matern1960spatial}. The Mat\'ern covariance between $U(x)$ and $U(x')$ takes the form 
-$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
+The common choice for the covariance function $\Sigma$ is the Mat\'ern covariance function \citep{matern1960spatial}. The Mat\'ern covariance between $U(x_i)$ and $U(x_j)$ takes the above form.
+%$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
 $\sigma^2$ the variance of the random field $U$.
-$\Gamma(\cdot)$ is the standard gamma function,  $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as $d$ increases.
+$\Gamma(\cdot)$ is the standard gamma function,  $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as distance increases.
 $K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$, specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.

### 2021-11-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3007 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % \end{itemize}
-The matrix decomposition method is straightforward to implement, however, when a large number of locations ($n$) is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
-\pkg{clrng} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
-The following shows an example, in which we simulate 10 Gaussian random fields of mat\'ern covariance with 5 sets of parameters at one time.
-%and introduces several GPU functions used in the simulation process.
+The matrix decomposition method is straightforward to implement.
+However, when a large number of locations $n$ is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
+Our another \proglang{R} package \pkg{gpuBatchMatrix} is able to compute batches of Mat\'ern covariance matrices in parallel, and do Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
+The following shows an example, in which we simulate on GPU 10 Gaussian random fields of mat\'ern covariance at one time with 5 sets of parameters, by taking advantage of the GPU-capabilities provided by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
+%\pkg{clrng} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
 

### 2021-11-27 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3007 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.072   0.000   0.072
+##   0.070   0.002   0.073
 \end{verbatim}
@@
 ##    user  system elapsed 
-##   1.950   0.001   1.952
+##   1.934   0.016   1.949
 result_week$threshold

### 2021-11-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3008 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-\documentclass[article,nojss]{jss}
+\documentclass[article]{jss}
 
-\usepackage{float}
-\usepackage{framed}
+<<downloadJssClassFile, eval=FALSE, include=FALSE>>=
+zUrl = 'http://www.jstatsoft.org/public/journals/1/jss-style.zip'
+zFile = basename(zUrl)
+if(!file.exists(zFile))
+        download.file(zUrl, zFile)
+unzip(zFile, exdir='.',
+                files = grep("^jss", unzip(zFile, list=TRUE)$Name, value=TRUE)
+)
+@
+
+%\usepackage{float}
+%\usepackage{framed}
 \usepackage{subcaption}
+\renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }
 \usepackage{amsmath}
@@
 \usepackage{booktabs}
-\renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }
 
@@
 
-<<preliminaries, echo=FALSE, results="hide">>=
+<<setup, cache=FALSE,include=FALSE>>=
+library('knitr')
+
+options(continue="+  ",  prompt="R> ", digits=3, width=57)
+
+opts_chunk$set(echo=FALSE,fig.path='Figures/G', fig.height=3,
+                fig.width=4.5,dev='png',tidy=TRUE,fig.align='center',
+                tidy.opts=list(blank=FALSE, width.cutoff=57),
+                prompt=TRUE, highlight=FALSE,cache.stuff=1)
+
+hook_output <- function(x, options) {
+        if (knitr:::output_asis(x, options)) return(x)
+        paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
+}
+
+knit_hooks$set(output = hook_output)
+knit_hooks$set(message = hook_output)
+knit_hooks$set(warning = hook_output)
+
+knit_hooks$set(source  =function(x, options) {
+        x = gsub("\n", "\n+", x)
+        x = paste("R>",paste(x, collapse='\nR> '))
+#       paste0(c('\\begin{example*}', x, '\\end{example*}', ''),
+#                       collapse = '\n')
+x =  paste0(c('\\begin{CodeInput}', x, '\\end{CodeInput}', ''),
+                        collapse = '\n')
+                        x               
+})
+
+knit_hooks$set(plot = function (x, options) 
+                {
+                        paste( hook_plot_tex(x, options), "\n", 
+                                        sep = "")
+                }
+)
+
+knit_hooks$set(chunk  = function(x, options) {
+        if (knitr:::output_asis(x, options)) return(x)
+        theend = gregexpr("end\\{Code(Input|Output)\\}", x)
+        theend=theend[[1]]
+        if(theend[1]>0){
+                theend = theend[length(theend)] + attributes(theend)$match.length[length(theend)]
+                x = paste(substr(x,1,theend-1),"\n\\end{CodeChunk}", substr(x,theend, nchar(x)))
+                x=sub('begin\\{Code', 'begin{CodeChunk}\n\\\\begin{Code', x)
+        }
+        x
+})
+knitr::knit_hooks$set(margins = function(before, options, envir) {
+    if (!before) 
+        return()
+    graphics::par(mar = c(1.5 + 0.9 * options$margins, 1.5 + 
+        0.9 * options$margins, 0.2, 0.2), mgp = c(1.45, 0.45, 
+        0), cex = 1.25)
+})
+opts_knit$set(header='')
+
+@
+
+<<preliminaries, echo=FALSE, results="hide", eval=FALSE>>=
   knitr::knit_hooks$set(plot=knitr::hook_plot_tex)  
   options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-knitr::opts_chunk$set(highlight=FALSE, background='#FFFFFF', fig.height=4, fig.width=6, out.width='0.45\\textwidth')
+knitr::opts_chunk$set(highlight=FALSE, background='#FFFFFF00', fig.height=4, fig.width=6, out.width='0.45\\textwidth')
 library("MASS")
@@
 
-\begin{leftbar}
 % Why?
@@
 
-\end{leftbar}
 <<echo=FALSE, results="hide", message=FALSE>>=
@@
 At the end of generating random numbers, streams are transferred back to global memory through the function \fct{streamsFromPrivate}.
-\begin{framed}
+
 \begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel}]
@@
 \end{lstlisting}
-\end{framed}
+
 %Figure \ref{fig:1} illustrates how this kernel executes in practice with a simple example.
@@
 %For example, stream 0 which is allocated in work-item $(0,0)$ generates a random number in matrix $(0,0)$ in the first iteration, then moves on to matrix $(0,2)$ and generates a random number there in the second iteration, in the third iteration it generates a random number in matrix $(0,4)$ and so on until it goes out of the matrix range in the next iteration.
-% \begin{figure}[H]
+% \begin{figure}[ht]
 %     \centering
@@
 
-\begin{algorithm*}[H] \label{algorithm1}
+\begin{algorithm*}[ht] \label{algorithm1}
 \SetAlgoLined
@@
 %
-\begin{framed}
+
 \begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},breaklines=true, escapeinside={(*@}{@*)}]
@@
 \end{lstlisting}
-\end{framed}
+
 %We illustrate in Figure \ref{fig3} how this kernel executes in practice with the previous toy example: to create a $4 \times 6$ matrix of Gaussian random numbers using a $2 \times 4$ NDRange.
@@
 
-% \begin{figure}[H]%
+% \begin{figure}[ht]%
 %     \centering
@@
 
- <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
+ <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h'>>=
 library(knitr)
@@
   caption = "Monthly birth anomaly data\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
-  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
+  kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
 @
@@
 Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
-<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide">>=
+<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
 library(latex2exp)
@@
 knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"), caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
-  kable_styling(full_width = F, position = "center", latex_options = "HOLD_position")
+  kable_styling(full_width = F, position = "center")#, latex_options = "HOLD_position")
 @
@@
 We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
-<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE>>=
+<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE, fig.pos='h'>>=
 library(kableExtra)
@@
 Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
-  kable_styling(full_width = FALSE, position = "center",latex_options = "HOLD_position") %>%
+  kable_styling(full_width = FALSE, position = "center") %>%
   kableExtra::group_rows(index = c("P-value" = 2, "Run-time" = 2)) %>%
@@
 
-\begin{figure}[H]
-\centering
-<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE, dev='png'>>=
-hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
+<<fighistMonth, eval=TRUE, echo=FALSE, dev='pdf', fig.cap = "Approximate sampling distributions of the test statistics from the two examples Month and Week.
+The test statistic values of the observed tables are marked with a blue line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1>>=
+hist(month_stats, xlab="test statistic",#TeX('-$\\sum(\\log(n_{ij}!))$'),
+     breaks=40, ylab="proportion", main="", prob=TRUE)
 abline(v = result_month$threshold, col = "blue", lwd = 1.5)
 
-hist(result_week$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
+hist(result_week$cpu, xlab='test statistic', ylab="proportion", prob=TRUE, main="")
 abline(v = result_week$threshold, col = "blue", lwd = 1.5)
 @
-\caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''.
-The test statistic values of the observed tables are marked with a blue line on each plot.
-\label{fig4}}
-\end{figure}
 
@@
 
-\begin{algorithm*}[H] \label{algorithm2}
+\begin{algorithm*}[ht] \label{algorithm2}
 \SetAlgoLined
@@
 <<setupcoords, eval=TRUE>>=
-Ngrid = 80
 NlocalCache = 1000
@@
 data("swissRain", package="geostatsp")
-myRaster = geostatsp::squareRaster(swissBorder, Ngrid)
+myRaster = geostatsp::squareRaster(swissBorder, 60)
 myRaster
@@
 params = 
-rbind(c(shape=1.25, range=0.5*1e04, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
-c(shape=2.15, range=0.25*1e04, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
-c(shape=0.55, range=1.5*1e04, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
-c(shape=2.15, range=0.5*1e04, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = -pi/7),
-c(shape=2.15, range=0.5*1e04, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/4))
+rbind(c(shape=1.25, range=50*1000, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
+c(shape=2.15, range=60*1000, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
+c(shape=0.6, range=30*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/7),
+c(shape=3, range=30*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/7)
+)
+#c(shape=2.15, range=40*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/4))
 @
@@
 <<r cholBach, eval=TRUE>>=
-# diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-diagMat = gpuBatchMatrix::cholBatch(maternCov, # D=diagMat,  # numbatchD=nrow(params),
+diagMat = gpuBatchMatrix::cholBatch(maternCov, 
                           Nglobal= c(128, 8), Nlocal= c(32, 8))
@@
 simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
-                   type = theType)
+                   type = gpuR::typeof(zmatGpu))
 
@@
 values(simRaster) = as.vector(as.matrix(simMat))
+```
+```{r theSubcapName, include=FALSE}
 names(simRaster) = apply(expand.grid('par',1:nrow(params), 
                                      'sim', 1:ncol(simMat)), 1, paste, collapse='')
+theSubcap = gsub("par", "parameter ", names(simRaster))
+theSubcap = gsub("sim", ", simuation ", theSubcap)
 @
 
-<<maternplot, echo=FALSE, eval=FALSE, cache=FALSE,dev='png', out.width = '0.45\\textwidth', fig.height=3, fig.width=6, fig.cap="Simulated Gaussian random fields", fig.subcap = names(simRaster)>>=
-myCol = mapmisc::colourScale(simRaster, breaks=9, type='equal', dec=1)
+<<maternplot, echo=FALSE, eval=TRUE, cache=FALSE, dev='png', out.width = '0.47\\textwidth', fig.height=3, fig.width=4.5, fig.cap="Simulated Gaussian random fields", fig.subcap = theSubcap, fig.align='default'>>=
+myCol = mapmisc::colourScale(breaks=c(-10, -5, seq(-3, 3), 5, 10), style='fixed', col='Spectral')
 for(D in names(simRaster)) {
@@
 }
-mapmisc::legendBreaks("right", myCol)
+mapmisc::legendBreaks("right", myCol, inset=0)
 @

### 2021-11-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3008 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-\documentclass[article,nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
+\documentclass[article]{jss}\usepackage[]{graphicx}\usepackage[]{color}
 % maxwidth is the original width if it is less than linewidth
@@
 
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-\let\hlipl\hlkwb
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
 \usepackage{alltt}
 
-\usepackage{float}
-\usepackage{framed}
+%\usepackage{float}
+%\usepackage{framed}
 \usepackage{subcaption}
+\renewcommand{\subfloat}[2][need a sub-caption]{ \subcaptionbox{#1}{#2} }
 \usepackage{amsmath}
@@
 Finally, the paper concludes with a short summary and a discussion in Section 6.
-
-\end{leftbar}
 
@@
 Substreams are not created in \pkg{clrng} because currently we use one stream per work-item.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-# creating streams on CPU
-library("gpuR")
-library("clrng")
-myStreamsCpu <- createStreamsCpu(n=4, initial=12345)
-t(myStreamsCpu)
+\begin{CodeChunk}
+\begin{CodeOutput}
 ##               [,1]       [,2]       [,3]       [,4]
@@
 ## initial.g2.3 12345 2094976052 1462898381  795289868
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\end{CodeOutput} 
+\end{CodeChunk}
+
 \fct{createStreamsCpu} has two arguments:
@@
 Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-# creating streams on GPU
-myStreamsGpu = vclMatrix(myStreamsCpu)
-myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+
 \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
@@
 To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
-## [1] 0.7353245 0.8423426 0.6142074 0.2159195 0.1100781 0.8698300
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeOutput}
+## [1] 0.735 0.842 0.614 0.216 0.110 0.870
+\end{CodeOutput} 
+\end{CodeChunk}
+
 The arguments of the \fct{clrng::runif} are described as follows:
@@
 Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
-# Load the streams object as streams_saved
-streams_saved <- vclMatrix(readRDS("myStreams.rds"))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> saveRDS(as.matrix(createStreamsGpu(n = 4)), "myStreams.rds")
+R> # Load the streams object as streams_saved
+R> streams_saved <- vclMatrix(readRDS("myStreams.rds"))
+\end{CodeInput} 
+\end{CodeChunk}
 
@@
 The difference in elapsed time becomes larger when matrix size goes larger.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streams <- createStreamsGpu(n = 512 * 128)
-system.time(clrng::rnorm(c(10000,10000), streams=streams, 
-                             Nglobal=c(512,128), type="double"))
+\begin{CodeChunk}
+\begin{CodeOutput}
 ##    user  system elapsed 
-##   0.070   0.002   0.073
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-system.time(matrix(stats::rnorm(10000^2),10000,10000))
+##   0.074   0.000   0.075
+\end{CodeOutput} 
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeOutput}
 ##    user  system elapsed 
-##   6.658   1.125   7.779
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+##   7.030   0.712   7.738
+\end{CodeOutput} 
+\end{CodeChunk}
 
@@
 Below is an example that produces a matrix of Exponential random numbers with expectation equal to 1.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-r_matrix <- clrng::rexp(c(2,4), rate=1, Nglobal=c(2,2), type="double")
-as.matrix(r_matrix)
+\begin{CodeChunk}
+\begin{CodeOutput}
 ##       [,1]  [,2]  [,3]  [,4]
-## [1,] 0.05469212 0.6969336 1.089906 0.78689966
-## [2,] 0.04188041 0.1686520 1.905295 0.03536933
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+## [1,] 0.183 2.025 0.323 1.861
+## [2,] 0.356 0.288 0.213 0.475
+\end{CodeOutput} 
+\end{CodeChunk}
+
 The arguments of the \fct{clrng::rexp} are described as follows:
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+\begin{table}
 
@@
 
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeOutput}
+##    user  system elapsed 
+##   0.313   0.003   0.315
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] -47955
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 1015808
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 410825
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 0.404
+\end{CodeOutput} 
+\end{CodeChunk}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streams <- createStreamsGpu(n =256*64, initial=666)
-month_gpu<-vclMatrix(month,type="integer")
-system.time(result_month <- clrng::fisher.sim(month_gpu, 1e6, streams=streams,
-                   type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
-##    user  system elapsed 
-##   0.303   0.012   0.315
-result_month$threshold
-## [1] -47954.8
-result_month$simNum
-## [1] 1015808
-result_month$counts
-## [1] 410825
-result_month$p.value
-## [1] 0.4044323
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using CPU
-system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e6))
+\begin{CodeChunk}
+\begin{CodeOutput}
 ##    user  system elapsed 
-##  14.445   0.036  14.473
-result_monthcpu$p.value
-## [1] 0.4034616
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+##  14.405   0.014  14.417
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 0.403
+\end{CodeOutput} 
+\end{CodeChunk}
 
@@
 
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeOutput}
+##    user  system elapsed 
+##   1.927   0.015   1.942
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] -54990
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 1e+07
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 1193
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 0.000119
+\end{CodeOutput} 
+\end{CodeChunk}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-week_GPU<-gpuR::vclMatrix(week,type="integer")
-system.time(result_week<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
-                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
-##    user  system elapsed 
-##   1.934   0.016   1.949
-result_week$threshold
-## [1] -54989.56
-result_week$simNum
-## [1] 10010624
-result_week$counts
-## [1] 1193
-result_week$p.value
-## [1] 0.0001192733
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value.
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#using CPU
-system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
+\begin{CodeChunk}
+\begin{CodeOutput}
 ##    user  system elapsed 
-##  89.981   0.158  90.071
-result_weekcpu$p.value
-## [1] 0.0001243679
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+##  89.719   0.116  89.817
+\end{CodeOutput}
+\begin{CodeOutput}
+## [1] 0.000127
+\end{CodeOutput} 
+\end{CodeChunk}
 
@@
 We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+\begin{table}
 
@@
 
-\end{knitrout}
+\begin{figure}
 
-\begin{figure}[H]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+{\centering \subfloat[Month data\label{fig:fighistMonth-1}]{\includegraphics[width=0.47\textwidth]{Figures/GfighistMonth-1} }
+\subfloat[week data\label{fig:fighistMonth-2}]{\includegraphics[width=0.47\textwidth]{Figures/GfighistMonth-2} }
 
-\includegraphics[width=0.45\textwidth]{figure/fighistMonth-1} \includegraphics[width=0.45\textwidth]{figure/fighistMonth-2} \end{kframe}
-\end{knitrout}
-\caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''.
-The test statistic values of the observed tables are marked with a blue line on each plot.
-\label{fig4}}
+}
+
+\caption[Approximate sampling distributions of the test statistics from the two examples Month and Week]{Approximate sampling distributions of the test statistics from the two examples Month and Week.
+The test statistic values of the observed tables are marked with a blue line on each plot.}\label{fig:fighistMonth}
 \end{figure}
@@
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-library("gpuR")
-library("clrng")
-library("gpuBatchMatrix")
-library('geostatsp')
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library("gpuR")
+R> library("clrng")
+R> library("gpuBatchMatrix")
+R> library("geostatsp")
+\end{CodeInput} 
+\end{CodeChunk}
+
 Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-Ngrid = 80
-NlocalCache = 1000
-Nglobal = c(128, 64, 2)
-Nlocal = c(4, 2, 2)
-theType = "double"
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-data("swissRain", package="geostatsp")
-myRaster = geostatsp::squareRaster(swissBorder, Ngrid)
-myRaster
+\begin{CodeChunk}
+\begin{CodeInput}
+R> data("swissRain", package = "geostatsp")
+R> myRaster = geostatsp::squareRaster(swissBorder, 40)
+R> myRaster
+\end{CodeInput}
+\begin{CodeOutput}
 ## class      : RasterLayer 
-## dimensions : 51, 80, 4080  (nrow, ncol, ncell)
-## resolution : 4358.065, 4358.065  (x, y)
-## extent     : 2485351, 2833996, 1075013, 1297275  (xmin, xmax, ymin, ymax)
+## dimensions : 26, 40, 1040  (nrow, ncol, ncell)
+## resolution : 8716, 8716  (x, y)
+## extent     : 2485351, 2833996, 1075013, 1301633  (xmin, xmax, ymin, ymax)
 ## crs        : +proj=somerc +lat_0=46.9524055555556 +lon_0=7.43958333333333 +k_0=1 +x_0=2600000 +y_0=1200000 +ellps=GRS80 +units=m +no_defs
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\end{CodeOutput} 
+\end{CodeChunk}
+
 Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets.
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-params
-##      shape range variance nugget anisoRatio anisoAngleRadians
-## [1,]  1.25  5000      1.5      0          1         0.0000000
-## [2,]  2.15  2500      2.0      0          4         0.4487990
-## [3,]  0.55 15000      2.0      0          4         0.4487990
-## [4,]  2.15  5000      2.0      0          4        -0.4487990
-## [5,]  2.15  5000      2.0      0          2         0.7853982
-#paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeOutput}
+##      shape range variance nugget anisoRatio
+## [1,]  1.25 50000      1.5      0          1
+## [2,]  2.15 60000      2.0      0          4
+## [3,]  0.60 30000      2.0      0          2
+## [4,]  3.00 30000      2.0      0          2
+##      anisoAngleRadians
+## [1,]             0.000
+## [2,]             0.449
+## [3,]             0.449
+## [4,]             0.449
+\end{CodeOutput} 
+\end{CodeChunk}
+
 Step 3, compute Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
 are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-#maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
-#                         nrow(coordsGpu),type=theType)
-maternCov = gpuBatchMatrix::maternBatch(
-  params, myRaster,          
-  Nglobal=c(128,64), Nlocal=c(16,4))
-dim(maternCov)
-## [1] 20400  4080
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeOutput}
+## [1] 4160 1040
+\end{CodeOutput} 
+\end{CodeChunk}
+
 Step 4, the first argument \code{A} in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
@@
 Each row of \code{diagMat} stores the diagonal values of each $D_i$, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-# diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
-diagMat = gpuBatchMatrix::cholBatch(maternCov, # D=diagMat,  # numbatchD=nrow(params),
-                          Nglobal= c(128, 8), Nlocal= c(32, 8))
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+
 \begin{gather}
@@
 Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-streamsGpu <- createStreamsGpu(n=128*64)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-zmatGpu = rnorm(c(nrow(maternCov),2), streams=streamsGpu, Nglobal=c(128,64),
-                 type = theType)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+
 Step 6, use \fct{gpuBatchMatrix::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
 \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
-                   type = theType)
 
-gpuBatchMatrix::multiplyLowerDiagonalBatch(
-  output=simMat, L=maternCov, 
-  D=diagMat, B=zmatGpu,
-  diagIsOne = 1L, # diagonal of L is one
-  transformD = "sqrt", # take the square root of each element of D
-  Nglobal,
-  Nlocal,
-  NlocalCache)
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 \begin{gather}
@@
 Step 7, Finally, plot the 10 realizations.
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-simRaster = raster::brick(myRaster, nl = ncol(simMat)*nrow(params))
-values(simRaster) = as.vector(as.matrix(simMat))
-names(simRaster) = apply(expand.grid('par',1:nrow(params), 
-                                     'sim', 1:ncol(simMat)), 1, paste, collapse='')
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> simRaster = raster::brick(myRaster, nl = ncol(simMat) * nrow(params))
+R> values(simRaster) = as.vector(as.matrix(simMat))
+R> names(simRaster) = apply(expand.grid("par", 1:nrow(params),
++    "sim", 1:ncol(simMat)), 1, paste, collapse = "")
+R> theSubcap = gsub("par", "parameter ", names(simRaster))
+R> theSubcap = gsub("sim", ", simuation ", theSubcap)
+\end{CodeInput} 
+\end{CodeChunk}
+
+\begin{figure}
+\subfloat[parameter 1, simuation 1\label{fig:maternplot-1}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-1} }
+\subfloat[parameter 2, simuation 1\label{fig:maternplot-2}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-2} }
+\subfloat[parameter 3, simuation 1\label{fig:maternplot-3}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-3} }
+\subfloat[parameter 4, simuation 1\label{fig:maternplot-4}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-4} }
+\subfloat[parameter 1, simuation 2\label{fig:maternplot-5}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-5} }
+\subfloat[parameter 2, simuation 2\label{fig:maternplot-6}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-6} }
+\subfloat[parameter 3, simuation 2\label{fig:maternplot-7}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-7} }
+\subfloat[parameter 4, simuation 2\label{fig:maternplot-8}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-8} }\caption[Simulated Gaussian random fields]{Simulated Gaussian random fields}\label{fig:maternplot}
+\end{figure}
 

### 2021-11-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper 1 stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3009 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  #using CPU
- system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e6))
+ system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e4))#B=1e6))
  result_monthcpu$p.value
@@
 #using CPU
-system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
+system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=1000))#B=10010624))
 result_weekcpu$p.value
@@
 values(simRaster) = as.vector(as.matrix(simMat))
-```
-```{r theSubcapName, include=FALSE}
+@
+
+<<r theSubcapName, include=FALSE>>=
 names(simRaster) = apply(expand.grid('par',1:nrow(params), 

### 2021-11-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1 stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3009 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.074   0.000   0.075
+##   0.077   0.000   0.076
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   7.030   0.712   7.738
+##   7.103   0.651   7.750
 \end{CodeOutput} 
@@
 ##       [,1]  [,2]  [,3]  [,4]
-## [1,] 0.183 2.025 0.323 1.861
-## [2,] 0.356 0.288 0.213 0.475
+## [1,] 0.343 0.364 0.542 0.054
+## [2,] 1.113 3.642 0.200 2.770
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.313   0.003   0.315
+##   0.316   0.000   0.316
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  14.405   0.014  14.417
+##   0.137   0.000   0.137
 \end{CodeOutput}
 \begin{CodeOutput}
-## [1] 0.403
+## [1] 0.4
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.927   0.015   1.942
+##   3.854   0.011   3.864
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  89.719   0.116  89.817
+##   0.009   0.000   0.009
 \end{CodeOutput}
 \begin{CodeOutput}
-## [1] 0.000127
+## [1] 0.000999
 \end{CodeOutput} 
@@
 R> data("swissRain", package = "geostatsp")
-R> myRaster = geostatsp::squareRaster(swissBorder, 40)
+R> myRaster = geostatsp::squareRaster(swissBorder, 60)
 R> myRaster
@@
 ## class      : RasterLayer 
-## dimensions : 26, 40, 1040  (nrow, ncol, ncell)
-## resolution : 8716, 8716  (x, y)
-## extent     : 2485351, 2833996, 1075013, 1301633  (xmin, xmax, ymin, ymax)
+## dimensions : 38, 60, 2280  (nrow, ncol, ncell)
+## resolution : 5811, 5811  (x, y)
+## extent     : 2485351, 2833996, 1075013, 1295822  (xmin, xmax, ymin, ymax)
 ## crs        : +proj=somerc +lat_0=46.9524055555556 +lon_0=7.43958333333333 +k_0=1 +x_0=2600000 +y_0=1200000 +ellps=GRS80 +units=m +no_defs
@@
 \begin{CodeOutput}
-## [1] 4160 1040
+## [1] 9120 2280
 \end{CodeOutput} 
@@
 R> values(simRaster) = as.vector(as.matrix(simMat))
-R> names(simRaster) = apply(expand.grid("par", 1:nrow(params),
-+    "sim", 1:ncol(simMat)), 1, paste, collapse = "")
-R> theSubcap = gsub("par", "parameter ", names(simRaster))
-R> theSubcap = gsub("sim", ", simuation ", theSubcap)
 \end{CodeInput} 

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
spatial edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3010 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  #using CPU
- system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e4))#B=1e6))
+ system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e6))
  result_monthcpu$p.value
@@
 #using CPU
-system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=1000))#B=10010624))
+system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
 result_weekcpu$p.value
@@
 data("swissRain", package="geostatsp")
-myRaster = geostatsp::squareRaster(swissBorder, 60)
+myRaster = geostatsp::squareRaster(swissBorder, 80)
 myRaster
@@
 <<r randomNormalsSim2, eval=TRUE, cache=FALSE>>=
-zmatGpu = rnorm(c(nrow(maternCov),2), streams=streamsGpu, Nglobal=c(128,64),
+zmatGpu = clrng::rnorm(
+  c(nrow(maternCov),2), streams=streamsGpu, 
+                Nglobal=c(128,2),
                  type = theType)

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3011 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), 
-  caption = "Monthly birth anomaly data\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
+  caption = "\\label{tab:month}Monthly birth anomaly data", booktabs=TRUE,linesep = "") %>%
   kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
@@
 \subsection{A summary of the results}
-We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
+We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure %\ref{fig4}.
 <<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE, fig.pos='h'>>=
@@
   c(nrow(maternCov),2), streams=streamsGpu, 
-                Nglobal=c(128,2),
+                Nglobal=c(128,64),
                 type = theType)

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3011 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{booktabs}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{pdfpages}
 
@@
 ##    user  system elapsed 
-##   0.077   0.000   0.076
+##   0.060   0.009   0.069
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   7.103   0.651   7.750
+##    6.61    0.80    7.41
 \end{CodeOutput} 
@@
 ##       [,1]  [,2]  [,3]  [,4]
-## [1,] 0.343 0.364 0.542 0.054
-## [2,] 1.113 3.642 0.200 2.770
+## [1,] 0.095 0.132 0.679 0.458
+## [2,] 0.275 0.132 1.118 0.111
 \end{CodeOutput} 
@@
 
-\begin{table}
-
-\caption{\label{tab:monthdata}Monthly birth anomaly data\label{tab:month}}
-\centering
-\begin{tabular}[t]{lrrrrrrrrrrrr}
-\toprule
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Dow & Chr & Hyp\\
-\midrule
-Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
-Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
-Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
-Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
-May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
-Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
-Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
-Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
-Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
-Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
-Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
-Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
-\bottomrule
-\end{tabular}
-\end{table}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in knitr::kable(month, format = "{}latex"{}, align = c("{}rrrrrrrrrrrr"{}), : could not find function "{}\%>\%"{}}}
 
@@
 ##    user  system elapsed 
-##   0.316   0.000   0.316
+##   0.342   0.005   0.348
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   0.137   0.000   0.137
+##  14.346   0.012  14.351
 \end{CodeOutput}
 \begin{CodeOutput}
-## [1] 0.4
+## [1] 0.404
 \end{CodeOutput} 
@@
 
-\begin{table}
-
-\caption{\label{tab:weekdata}Day-of-week birth anomaly data\label{tab:week}}
-\centering
-\begin{tabular}[t]{lrrrrrrrrrrrr}
-\toprule
-  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Dow & Chr & Hyp\\
-\midrule
-Mon & 30 & 34 & 173 & 37 & 23 & 80 & 49 & 191 & 83 & 122 & 109 & 216\\
-Tue & 60 & 121 & 383 & 80 & 83 & 131 & 71 & 349 & 146 & 164 & 168 & 352\\
-Wed & 51 & 106 & 417 & 92 & 73 & 145 & 72 & 333 & 136 & 179 & 196 & 351\\
-Thu & 60 & 86 & 362 & 69 & 74 & 120 & 85 & 326 & 132 & 220 & 187 & 359\\
-Fri & 52 & 94 & 347 & 87 & 59 & 123 & 68 & 323 & 145 & 170 & 166 & 345\\
-Sat & 52 & 63 & 323 & 67 & 64 & 135 & 73 & 316 & 170 & 189 & 188 & 357\\
-Sun & 49 & 51 & 211 & 40 & 32 & 96 & 69 & 216 & 108 & 143 & 130 & 258\\
-\bottomrule
-\end{tabular}
-\end{table}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in knitr::kable(week, format = "{}latex"{}, align = c("{}rrrrrrrrrrrr"{}), : could not find function "{}\%>\%"{}}}
 
@@
 ##    user  system elapsed 
-##   3.854   0.011   3.864
+##   1.941   0.009   1.949
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   0.009   0.000   0.009
+##  89.759   0.096  89.810
 \end{CodeOutput}
 \begin{CodeOutput}
-## [1] 0.000999
+## [1] 0.000117
 \end{CodeOutput} 
@@
 \subsection{A summary of the results}
-We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
+We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure %\ref{fig4}.
 \begin{table}
@@
 R> data("swissRain", package = "geostatsp")
-R> myRaster = geostatsp::squareRaster(swissBorder, 60)
+R> myRaster = geostatsp::squareRaster(swissBorder, 80)
 R> myRaster
@@
 ## class      : RasterLayer 
-## dimensions : 38, 60, 2280  (nrow, ncol, ncell)
-## resolution : 5811, 5811  (x, y)
-## extent     : 2485351, 2833996, 1075013, 1295822  (xmin, xmax, ymin, ymax)
+## dimensions : 51, 80, 4080  (nrow, ncol, ncell)
+## resolution : 4358, 4358  (x, y)
+## extent     : 2485351, 2833996, 1075013, 1297275  (xmin, xmax, ymin, ymax)
 ## crs        : +proj=somerc +lat_0=46.9524055555556 +lon_0=7.43958333333333 +k_0=1 +x_0=2600000 +y_0=1200000 +ellps=GRS80 +units=m +no_defs
@@
 \begin{CodeOutput}
-## [1] 9120 2280
+## [1] 16320  4080
 \end{CodeOutput} 

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3012 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 library('knitr')
-
 options(continue="+  ",  prompt="R> ", digits=3, width=57)
@@
 
- <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h'>>=
+ <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h', message=FALSE>>=
 library(knitr)
-knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), 
-  caption = "\\label{tab:month}Monthly birth anomaly data", booktabs=TRUE,linesep = "") %>%
+library(tidyverse)
+library(kableExtra)
+knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), #label="tab:month"
+  caption = "Monthly birth anomaly data\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
   kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
@@
 <<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
-knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"), caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
+knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"),  #label = "tab:week",
+             caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
 kable_styling(full_width = F, position = "center")#, latex_options = "HOLD_position")

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3012 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2021-11-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
@@
 ##    user  system elapsed 
-##   0.077   0.000   0.076
+##   0.050   0.000   0.051
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   7.103   0.651   7.750
+##   6.599   0.762   7.358
 \end{CodeOutput} 
@@
 ##       [,1] [,2]   [,3]  [,4]
-## [1,] 0.343 0.364 0.542 0.054
-## [2,] 1.113 3.642 0.200 2.770
+## [1,] 0.707 1.58 0.1815 0.387
+## [2,] 1.131 6.38 0.0157 2.768
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.316   0.000   0.316
+##    0.32    0.00    0.32
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   0.137   0.000   0.137
+##  14.448   0.018  14.457
 \end{CodeOutput}
 \begin{CodeOutput}
-## [1] 0.4
+## [1] 0.404
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   3.854   0.011   3.864
+##   1.936   0.005   1.940
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   0.009   0.000   0.009
+##  89.796   0.107  89.865
 \end{CodeOutput}
 \begin{CodeOutput}
-## [1] 0.000999
+## [1] 0.000125
 \end{CodeOutput} 

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3013 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Accomplishing this is complicated to do because each process must produce an independent sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of numbers at any point.
-Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \pkg{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
+Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \proglang{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
 
@@
 In Section 2, we introduce streams and the use of streams in work-items on a GPU device for generating uniform random numbers, and the usage of \fct{clrng::runif}.
-In Section 3, we introduce two non-uniform RNGs in \pkg{clrng}, as examples for users to develop other RNGs of interest on GPUs in R.
+In Section 3, we introduce two non-uniform RNGs in \pkg{clrng}, as examples for users to develop other RNGs of interest on GPUs in \proglang{R}.
 In Section 4, we apply GPU-generated uniform random numbers in Monte Carlo simulation for Fisher’s exact test, we introduce how the random numbers are used and how the algorithm is parallelized and implemented on GPU.
-Then we provide two real data examples to demonstrate the function usage and its R performance.
+Then we provide two real data examples to demonstrate the function usage and its \proglang{R} performance.
 Section 5, we show an useful application of normal random numbers on GPU, using them to simulate batches of Gaussian random surfaces with Mat\'ern covariance matrices simultaneously, the simulation also uses GPU-enabled functions from our another package \pkg{gpuBatchMatrix}.

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3014 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 We move streams to the GPU by converting them to a `\code{vclMatrix}'.
+\fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
+However, any data is lost when the \proglang{R} session is terminated.
 %\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
@@
 @
-\fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-However, any data is lost when the \proglang{R} session is terminated.
+
 %also has the two arguments \code{n} and \code{initial}.
@@
 <<simStuff2, eval=TRUE,tidy=FALSE>>=
-as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2), type = "double"))
+as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2)))
 @

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3015 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
-<<simStuff2, eval=TRUE,tidy=FALSE>>=
+<<simStuff2, eval=TRUE,tidy=TRUE>>=
 as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2)))

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3013 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
add new r file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3016 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-However, any data is lost when the \proglang{R} session is terminated.
 %also has the two arguments \code{n} and \code{initial}.
@@
 ##    user  system elapsed 
-##   0.050   0.000   0.051
+##   0.046   0.003   0.050
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   6.599   0.762   7.358
+##   6.572   0.891   7.460
 \end{CodeOutput} 
@@
 ##       [,1]  [,2]   [,3]  [,4]
-## [1,] 0.707 1.58 0.1815 0.387
-## [2,] 1.131 6.38 0.0157 2.768
+## [1,] 1.108 0.279 0.1244 0.329
+## [2,] 0.296 1.456 0.0504 3.717
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##    0.32    0.00    0.32
+##   0.319   0.000   0.318
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  14.448   0.018  14.457
+##  14.308   0.013  14.319
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.936   0.005   1.940
+##    1.94    0.01    1.95
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  89.796   0.107  89.865
+##  89.629   0.085  89.713
 \end{CodeOutput}
 \begin{CodeOutput}
-## [1] 0.000125
+## [1] 0.000124
 \end{CodeOutput} 

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
change scale in colour plot
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3017 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-\documentclass[article]{jss}
+\documentclass[article,nojss]{jss}
 
@@
 <<maternplot, echo=FALSE, eval=TRUE, cache=FALSE, dev='png', out.width = '0.47\\textwidth', fig.height=3, fig.width=4.5, fig.cap="Simulated Gaussian random fields", fig.subcap = theSubcap, fig.align='default'>>=
-myCol = mapmisc::colourScale(breaks=c(-10, -5, seq(-3, 3), 5, 10), style='fixed', col='Spectral')
+myCol = mapmisc::colourScale(breaks=sort(unique(c(-6, -4, seq(-2, 2), 4, 6))), style='fixed', col='Spectral')
 for(D in names(simRaster)) {

### 2021-11-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
echo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3018 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-opts_chunk$set(echo=FALSE,fig.path='Figures/G', fig.height=3,
+opts_chunk$set(echo=TRUE,fig.path='Figures/G', fig.height=3,
                 fig.width=4.5,dev='png',tidy=TRUE,fig.align='center',
@@
 To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
-<<simStuff2, eval=TRUE,tidy=TRUE>>=
+<<simStuff2, eval=TRUE,tidy=TRUE, echo=TRUE>>=
 as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2)))

### 2021-12-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
R file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3022 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-\documentclass[article]{jss}\usepackage[]{graphicx}\usepackage[]{color}
+\documentclass[article,nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
 % maxwidth is the original width if it is less than linewidth
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> # creating streams on CPU
+R> library("gpuR")
+R> library("clrng")
+R> myStreamsCpu <- createStreamsCpu(n = 4, initial = 12345)
+R> t(myStreamsCpu)
+\end{CodeInput}
 \begin{CodeOutput}
@@
 Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # creating streams on GPU
+R> myStreamsGpu = vclMatrix(myStreamsCpu)
+R> myStreamsGpu2 = createStreamsGpu(n = 4, initial = 12345)
+\end{CodeInput} 
+\end{CodeChunk}
 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> as.vector(clrng::runif(n = 6, streams = myStreamsGpu, Nglobal = c(2,
++    2)))
+\end{CodeInput}
 \begin{CodeOutput}
@@
 
-\begin{algorithm*}[ht] \label{algorithm1}
+\begin{algorithm}[ht] 
 \SetAlgoLined
@@
  \caption{Box-Muller algorithm.}
-\end{algorithm*}
+ \label{algorithm1}
+\end{algorithm}
 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> streams <- createStreamsGpu(n = 512 * 128)
+R> system.time(clrng::rnorm(c(10000, 10000), streams = streams,
++    Nglobal = c(512, 128), type = "double"))
+\end{CodeInput}
 \begin{CodeOutput}
 ##    user  system elapsed 
-##   0.046   0.003   0.050
+##   0.043   0.004   0.047
 \end{CodeOutput} 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> system.time(matrix(stats::rnorm(10000^2), 10000, 10000))
+\end{CodeInput}
 \begin{CodeOutput}
 ##    user  system elapsed 
-##   6.572   0.891   7.460
+##   6.700   0.828   7.526
 \end{CodeOutput} 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> r_matrix <- clrng::rexp(c(2, 4), rate = 1, Nglobal = c(2,
++    2), type = "double")
+R> as.matrix(r_matrix)
+\end{CodeInput}
 \begin{CodeOutput}
 ##      [,1]  [,2]   [,3]  [,4]
-## [1,] 1.108 0.279 0.1244 0.329
-## [2,] 0.296 1.456 0.0504 3.717
+## [1,] 1.48 1.916 1.0358 0.469
+## [2,] 1.90 0.297 0.0868 0.161
 \end{CodeOutput} 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> streams <- createStreamsGpu(n = 256 * 64, initial = 666)
+R> month_gpu <- vclMatrix(month, type = "integer")
+R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1e+06,
++    streams = streams, type = "double", returnStatistics = TRUE,
++    Nglobal = c(256, 64)))
+\end{CodeInput}
 \begin{CodeOutput}
 ##    user  system elapsed 
-##   0.319   0.000   0.318
+##   0.322   0.011   0.333
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_month$threshold
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_month$simNum
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_month$counts
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_month$p.value
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> # using CPU
+R> system.time(result_monthcpu <- stats::fisher.test(month, simulate.p.value = TRUE,
++    B = 1e+06))
+\end{CodeInput}
 \begin{CodeOutput}
 ##    user  system elapsed 
-##  14.308   0.013  14.319
+##  14.455   0.028  14.478
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_monthcpu$p.value
+\end{CodeInput}
 \begin{CodeOutput}
-## [1] 0.404
+## [1] 0.403
 \end{CodeOutput} 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
+R> system.time(result_week <- clrng::fisher.sim(week_GPU, 1e+07,
++    streams = streams, type = "double", returnStatistics = TRUE,
++    Nglobal = c(256, 64)))
+\end{CodeInput}
 \begin{CodeOutput}
 ##    user  system elapsed 
-##    1.94    0.01    1.95
+##   1.930   0.021   1.950
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_week$threshold
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_week$simNum
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_week$counts
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_week$p.value
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> # using CPU
+R> system.time(result_weekcpu <- fisher.test(week, simulate.p.value = TRUE,
++    B = 10010624))
+\end{CodeInput}
 \begin{CodeOutput}
 ##    user  system elapsed 
-##  89.629   0.085  89.713
+##  89.968   0.096  90.031
 \end{CodeOutput}
+\begin{CodeInput}
+R> result_weekcpu$p.value
+\end{CodeInput}
 \begin{CodeOutput}
-## [1] 0.000124
+## [1] 0.000132
 \end{CodeOutput} 
@@
 
-\begin{algorithm*}[ht] \label{algorithm2}
+\begin{algorithm}[ht]
 \SetAlgoLined
@@
  \caption{Gaussian random fields simulation using covariance matrix decomposition method.}
-\end{algorithm*}
+ \label{algorithm2}
+\end{algorithm}
 
@@
 Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> NlocalCache = 1000
+R> Nglobal = c(128, 64, 2)
+R> Nlocal = c(4, 2, 2)
+R> theType = "double"
+\end{CodeInput} 
+\end{CodeChunk}
 
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> params
+\end{CodeInput}
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+\begin{CodeInput}
+R> # paramsGpu =
+R> # gpuBatchMatrix::maternGpuParam(myParamsBatch,
+R> # type=theType)
+\end{CodeInput} 
 \end{CodeChunk}
@@
 \begin{CodeChunk}
+\begin{CodeInput}
+R> # maternCov = vclMatrix(0,
+R> # nrow(paramsGpu)*nrow(coordsGpu),
+R> # nrow(coordsGpu),type=theType)
+R> maternCov = gpuBatchMatrix::maternBatch(params, myRaster,
++    Nglobal = c(128, 64), Nlocal = c(16, 4))
+R> dim(maternCov)
+\end{CodeInput}
 \begin{CodeOutput}
@@
 Each row of \code{diagMat} stores the diagonal values of each $D_i$, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> diagMat = gpuBatchMatrix::cholBatch(maternCov, Nglobal = c(128,
++    8), Nlocal = c(32, 8))
+\end{CodeInput} 
+\end{CodeChunk}
 
@@
 Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> streamsGpu <- createStreamsGpu(n = 128 * 64)
+\end{CodeInput} 
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> zmatGpu = clrng::rnorm(c(nrow(maternCov), 2), streams = streamsGpu,
++    Nglobal = c(128, 64), type = theType)
+\end{CodeInput} 
+\end{CodeChunk}
 
@@
 \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), type = gpuR::typeof(zmatGpu))
+R> gpuBatchMatrix::multiplyLowerDiagonalBatch(output = simMat,
++    L = maternCov, D = diagMat, B = zmatGpu, diagIsOne = TRUE,
++    transformD = "sqrt", Nglobal, Nlocal, NlocalCache)
+\end{CodeInput} 
+\end{CodeChunk}
 

### 2021-12-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.R

@@ -1,0 +1,338 @@
+## check if needee packages are installed
+packagesList <- c("knitr", "tidyverse", "kableExtra", "devtools")
+
+for (i in 1:length(packagesList)){
+  if(packagesList[i] %in% rownames(installed.packages()) == FALSE) 
+   {install.packages(packagesList[i])}
+}
+
+if("gpuR" %in% rownames(installed.packages()) == FALSE) 
+{devtools::install_github("cdeterman/gpuR")}    #{install.packages("gpuR", repos='https://github.com/cdeterman/gpuR')}
+
+if("geostatsp" %in% rownames(installed.packages()) == FALSE) 
+{install.packages("geostatsp", repos='http://r-forge.r-project.org')}
+
+if("clrng" %in% rownames(installed.packages()) == FALSE) 
+{devtools::install_github("Ruoyong/clrng")}
+
+if("gpuBatchMatrix" %in% rownames(installed.packages()) == FALSE) 
+{devtools::install_github("Ruoyong/gpuBatchMatrix")}
+
+
+
+
+## ----downloadJssClassFile, eval=FALSE, include=FALSE-------------------------------------------------------------------------------------------------------
+## zUrl = 'http://www.jstatsoft.org/public/journals/1/jss-style.zip'
+## zFile = basename(zUrl)
+## if(!file.exists(zFile))
+##         download.file(zUrl, zFile)
+## unzip(zFile, exdir='.',
+##                 files = grep("^jss", unzip(zFile, list=TRUE)$Name, value=TRUE)
+## )
+
+
+## ----echo=FALSE,results="hide"-----------------------------------------------------------------------------------------------------------------------------
+options(width=65)
+
+
+## ----setup, cache=FALSE,include=FALSE----------------------------------------------------------------------------------------------------------------------
+library('knitr')
+options(continue="+  ",  prompt="R> ", digits=3, width=57) 
+
+opts_chunk$set(echo=TRUE,fig.path='Figures/G', fig.height=3,
+                fig.width=4.5,dev='png',tidy=TRUE,fig.align='center',
+                tidy.opts=list(blank=FALSE, width.cutoff=57),
+                prompt=TRUE, highlight=FALSE,cache.stuff=1)
+
+hook_output <- function(x, options) {
+        if (knitr:::output_asis(x, options)) return(x)
+        paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
+}
+
+knit_hooks$set(output = hook_output)
[…]

### 2021-12-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
R file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3022 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
-library(latex2exp)
 dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
@@
   D=diagMat, B=zmatGpu,
-  diagIsOne = 1L, # diagonal of L is one
-  transformD = "sqrt", # take the square root of each element of D
-  Nglobal,
-  Nlocal,
-  NlocalCache)
+  diagIsOne = TRUE,   
+  transformD = "sqrt", 
+  Nglobal, Nlocal, NlocalCache)
 @

### 2021-12-03 pkg/gpuRandom/R/betascalarPro.R
r functions update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3024 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -4,19 +4,25 @@
 
 
 
-likLgm_betascalarProfile <- function(Betas, #a m x 1 R vector  given by the user 
-                                     a,     # which beta_i?
-                                     Nobs,  # number of observations.
-                                     Ndata,
-                                     Nparam,
-                                     Ncov,
-                                     detVar, # vclVector
-                                     ssqY,   # vclMatrix
-                                     XVYXVX,   # vclMatrix
-                                     jacobian, # vclVector  #form = c("loglik", "profileforBeta"),
-                                     minustwotimes=TRUE){ 
+  betascalarProfile <- function(Betas, #a m x 1 R vector  given by the user 
+                                cilevel=0.95,
+                                a,     # which beta_i?
+                                Nobs,  # number of observations.
+                                Ndata,
+                                Nparam,
+                                Ncov,
+                                detVar, # vclVector
+                                ssqY,   # vclMatrix
+                                XVYXVX,   # vclMatrix
+                                jacobian # vclVector  #form = c("loglik", "profileforBeta"),
+                                ){ 
   
-
+  m <- length(Betas)
+  if(m < 5){
+    warning("need more values for accurate estimate")
+  }
+  
+  
   ssqY <- as.matrix(ssqY)
   detVar <- as.vector(detVar)
   detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
@@ -24,8 +30,6 @@ likLgm_betascalarProfile <- function(Betas, #a m x 1 R vector  given by the user
   jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))
   #dim(XVYXVX)
   Ucov <- Ncov-1
-  m <- length(Betas)
-  
   XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
   XVY <- XVYXVX[ , 1:Ndata]
   
@@ -51,7 +55,7 @@ likLgm_betascalarProfile <- function(Betas, #a m x 1 R vector  given by the user
     partC = matrix(0, nrow=Nparam, ncol=Ndata)
[…]

### 2021-12-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
R file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3023 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update r codes paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3029 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 library("gpuR")
-library("clrng")
 setContext(   grep('gpu', listContexts()$device_type) [1]    )
@@
 Substreams are not created in \pkg{clrng} because currently we use one stream per work-item.
-<<createStreams_CPU>>=
+<<createStreams_CPU, message=FALSE>>=
 # creating streams on CPU
-library("gpuR")
 library("clrng")
@@
 streams <- createStreamsGpu(n = 512 * 128)
-system.time(clrng::rnorm(c(10000,10000), streams=streams, 
-                             Nglobal=c(512,128), type="double"))
+system.time(clrng::rnorm(c(10000,10000), streams=streams, Nglobal=c(512,128), type="double"))
 @
@@
  <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
- datamonth<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv") 
- month<-as.matrix(datamonth[,-1])
- rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
- colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
+ # datamonth<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv")
+ # month<-as.matrix(datamonth[,-1])
+ # rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
+ # colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
+ month <- as.matrix(readRDS(system.file("data/month.Rds", package = "clrng")))
  @
@@
 <<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
-dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv") 
-week<-as.matrix(dataweek[,-1])
-rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
-colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
+# dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv")
+# week<-as.matrix(dataweek[,-1])
+# rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
+# colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
+week <- as.matrix(readRDS(system.file("data", "week.Rds", package = "clrng")))
 @

### 2021-12-05 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
R file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3023 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update r codes paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3029 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 R> # creating streams on CPU
-R> library("gpuR")
 R> library("clrng")
@@
 ##    user  system elapsed 
-##   0.043   0.004   0.047
+##   0.045   0.001   0.047
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   6.700   0.828   7.526
+##   6.629   0.804   7.428
 \end{CodeOutput} 
@@
 ##       [,1] [,2] [,3]  [,4]
-## [1,] 1.48 1.916 1.0358 0.469
-## [2,] 1.90 0.297 0.0868 0.161
+## [1,] 0.942 1.68 2.65 1.372
+## [2,] 0.132 1.03 1.31 0.478
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.322   0.011   0.333
+##   0.327   0.003   0.330
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  14.455   0.028  14.478
+##  14.413   0.021  14.428
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.930   0.021   1.950
+##   1.945   0.009   1.953
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  89.968   0.096  90.031
+##  89.721   0.103  89.802
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.000132
+## [1] 0.000119
 \end{CodeOutput} 

### 2021-12-07 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
r functions
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3031 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
- <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h', message=FALSE>>=
+ <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h', message=FALSE,purl=FALSE>>=
 library(knitr)
@@
   kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
+@
+<<monthDataForScript, include=FALSE, purl=TRUE>>-
+knitr::kable(month)
 @

### 2021-12-07 pkg/gpuRandom/R/variancePro.R

@@ -50,15 +50,18 @@
   ci <- rootSolve::uniroot.all(f2, lower = lower, upper = upper)
   
   if(length(ci)==1){
-    if((abs(ci-lower)) > (abs(ci-upper))){
+    if( ci > MLE){
       ci <- c(lower, ci)
     }else{
       ci <- c(ci, upper)}
   }
-  
+  if(length(ci)==0){
+    warning("require a better param matrix")
+    ci <- c(NA, NA)
+  }
   if(length(ci)>2){
+    warning("need wider range of beta to search ci's")
     ci <- ci[3:4]
-    warning("spline approximation is not good")
   }
   
   ############### output #####################################

### 2021-12-08 pkg/gpuRandom/R/likfitLgmGpu.R
gemm
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3032 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -101,7 +101,7 @@ likfitLgmGpu <- function(data,
     as.integer(Nglobal),  #12
     as.integer(Nlocal),  #16
     NlocalCache,  #14
-    verbose,  #15
+    verbose=2,  #15
     ssqYX, #
     ssqYXcopy,  #new
     LinvYX,  #18
@@ -113,7 +113,8 @@ likfitLgmGpu <- function(data,
   # resid^T V^(-1) resid, resid = Y - X betahat = ssqResidual
   ssqResidual <- ssqY - ssqBetahat
   # any(is.na(as.matrix(log(ssqResidual/Nobs))))
-  any(is.nan(as.matrix(log(ssqResidual/Nobs))))
+   as.matrix(cholDiagMat)[]
+   any(is.nan(as.matrix(log(ssqResidual/Nobs))))
    any(is.nan(as.vector(detVar)))
    any(is.na(as.matrix(varMat)))
    any(is.nan(as.matrix(varMat)))
@@ -124,9 +125,8 @@ likfitLgmGpu <- function(data,
    any(is.na(as.vector(detReml)))
    any(is.na(as.matrix(cholXVXdiag)))
    any(is.na(as.matrix(ssqResidual)))
-  #  
-  # any(is.nan(as.matrix(ssqResidual/Nobs)))
-  # 
+   any(is.nan(as.matrix(ssqResidual/Nobs)))
+   as.matrix(ssqY)[which(is.na(as.vector(detVar))),]
   # which(is.na(as.vector(detVar)))
   # debug <- as.matrix(log(ssqResidual/Nobs))
   # any(is.na(debug))
@@ -137,7 +137,7 @@ likfitLgmGpu <- function(data,
   # ssqY[which(is.na(debug),arr.ind = TRUE)[,1],]
   # ssqBetahat[which(is.na(debug),arr.ind = TRUE)[,1],]    # problem arises from here! 4.530057e+20
   # 
-  # 
+  # params0[which(is.na(as.vector(detVar))),]
   # 
   # if(fixVariance >0){
   #   temp <- vclMatrix(0, Nparam, Ndata, type=type)

### 2021-12-08 pkg/gpuRandom/src/cholBatch.cpp

@@ -234,7 +234,7 @@ template <typename T> std::string cholBatchKernelString(
     "} // Dcol loop\n\n";
   
   if(logDet){
-    result +=  "if(localIndexIsZero & (Drow < N)){\n"
+    result +=  "if(localIndexIsZero){\n"   //& (Drow < N)
     " logDet[logDetIndex + Dmatrix] = logDetHere;\n"
     "}\n";
   }

### 2021-12-09 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3035 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  # colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
- month <- as.matrix(readRDS(system.file("data/month.Rds", package = "clrng")))
+ month <- as.matrix(readRDS(system.file("extdata", "month.Rds", package = "clrng")))
  @
@@
 # colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
-week <- as.matrix(readRDS(system.file("data", "week.Rds", package = "clrng")))
+week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
 @

### 2021-12-09 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1_2.R

@@ -58,7 +58,7 @@ as.matrix(r_matrix)
 
 
 ## Section 4.1
-month <- as.matrix(readRDS(system.file("data/month.Rds", package = "clrng")))
+month <- as.matrix(readRDS(system.file("extdata", "month.Rds", package = "clrng")))
 library(knitr)
 kable(month, format = "markdown", caption = "Monthly birth anomaly data")
 
@@ -80,12 +80,13 @@ result_month$p.value
 month_stats <- as.vector(result_month$sim)
 
 # using CPU
+# notice this can take about more than 1 minute
 system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e6))
 result_monthcpu$p.value
 
 
 ## Section 4.2
-week <- as.matrix(readRDS(system.file("data", "week.Rds", package = "clrng")))
+week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
 kable(week, format = "markdown", caption = "Day-of-week birth anomaly data") 
 
 # using GPU
@@ -102,6 +103,8 @@ result_week$cpu = as.vector(result_week$sim)
 
 
 # using CPU
+# notice that running in rstudio can take about 10 minutes 
+# set a smaller number for B if you think that's too long time
 system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
 result_weekcpu$p.value
 

### 2021-12-13 pkg/gpuRandom/src/gemmbatchLocal.cpp
update r fun
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3037 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -537,4 +537,3 @@ SEXP gemmBatch2backend(
 
 
 
-

### 2021-12-13 pkg/gpuRandom/src/maternBatch.cpp

@@ -484,7 +484,7 @@ void maternBatchVcl(
     Npad = vclVar.internal_size2(),
     NpadBetweenMatrices = Npad*N; // change to Npad*(Nmat+k) to insert extra rows between matrices
   
-    const int Ncell = N * (N - 1)/2, maxIter = 4000;
+    const int Ncell = N * (N - 1)/2, maxIter = 5000;
     int NgroupsOfParameters = numWorkItems[1]/numLocalItems[1];
     
     int NparametersPerGroup = std::max(

### 2021-12-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3038 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 options(width=65)
+options(scipen=999)
 @
@@
 
-knit_hooks$set(source  =function(x, options) {
+knit_hooks$set(source  = function(x, options) {
         x = gsub("\n", "\n+", x)
@@
 We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{clrng::rnorm}.
-The best performance we have seen using \fct{clrng::rnorm} is around 120 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
+The best performance we have seen using \fct{clrng::rnorm} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size goes larger.
@@
   var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.28,    10.82),
-  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 15.00,  91.58),
-  var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.72,   4.18),
+  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.48,  91.00),
+  var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.32,   1.95),
   var6 = c('Data','month','week','month','week')

### 2021-12-13 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3038 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.045   0.001   0.047
+##   0.047   0.000   0.047
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   6.629   0.804   7.428
+##   6.567   0.818   7.378
 \end{CodeOutput} 
@@
 ##       [,1]   [,2]  [,3]  [,4]
-## [1,] 0.942 1.68 2.65 1.372
-## [2,] 0.132 1.03 1.31 0.478
+## [1,] 0.367 0.0488 0.490 0.308
+## [2,] 0.671 3.1532 0.831 0.281
 \end{CodeOutput} 
@@
 
+<<monthDataForScript, include=FALSE, purl=TRUE>>-
+knitr::kable(month)
+@
+
 \begin{CodeChunk}
@@
 R> month_gpu <- vclMatrix(month, type = "integer")
-R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1e+06,
+R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1000000,
 +    streams = streams, type = "double", returnStatistics = TRUE,
@@
 ##    user  system elapsed 
-##   0.327   0.003   0.330
+##   0.319   0.000   0.318
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  14.413   0.021  14.428
+##  14.478   0.012  14.480
 \end{CodeOutput}
@@
 R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
-R> system.time(result_week <- clrng::fisher.sim(week_GPU, 1e+07,
+R> system.time(result_week <- clrng::fisher.sim(week_GPU, 10000000,
 +    streams = streams, type = "double", returnStatistics = TRUE,
@@
 ##    user  system elapsed 
-##   1.945   0.009   1.953
+##   1.941   0.015   1.954
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 1e+07
+## [1] 10010624
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  89.721   0.103  89.802
+##   90.70    0.12   90.82
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.000119
+## [1] 0.000121
 \end{CodeOutput} 
@@
 \hline
-\hspace{1em}1M & 10.74 & 2.28 & 15 & 0.72 & month\\
+\hspace{1em}1M & 10.74 & 2.28 & 14.48 & 0.34 & month\\
 \hline
-\hspace{1em}10M & 63.24 & 10.82 & 91.58 & 4.18 & week\\
+\hspace{1em}10M & 63.24 & 10.82 & 91 & 1.95 & week\\
 \hline

### 2021-12-21 pkg/gpuRandom/R/Prof1dBetas.R
organize r functions
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3046 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -6,32 +6,28 @@
          m[upper.tri(m)] <- t(m)[upper.tri(m)]
          return(m)
      }
-     
-     
+  
 #' @export
-  betascalarProfile <- function(Betas, #a m x 1 R vector  given by the user 
+
+        Prof1dBetas <- function(Betas, #a m x 1 R vector  given by the user 
                                 cilevel=0.95,
                                 a,     # which beta_i?
                                 Nobs,  # number of observations.
                                 Ndata,
                                 Nparam,
                                 Ncov,
-                                detVar, # vclVector
-                                ssqY,   # vclMatrix
-                                XVYXVX,   # vclMatrix
-                                jacobian # vclVector  #form = c("loglik", "profileforBeta"),
-                                ){ 
+                                detVar, # 
+                                ssqY,   # 
+                                XVYXVX,   # 
+                                jacobian){ 
   
   m <- length(Betas)
   if(m < 5){
     warning("need more values for accurate estimate")
   }
   
-  
-  ssqY <- as.matrix(ssqY)
-  detVar <- as.vector(detVar)
+
   detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
-  jacobian <- as.vector(jacobian)
   jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))
   #dim(XVYXVX)
   Ucov <- Ncov-1

### 2021-12-21 pkg/gpuRandom/R/likfitLgmGpu.R

@@ -19,8 +19,8 @@ likfitLgmGpu <- function(data,
                          type = c("float", "double")[1+gpuInfo()$double_support],
                          reml=FALSE, 
                          NparamPerIter,
+                         df=1,
                          minustwotimes=FALSE,
-                         twoDbreak = FALSE,
                          Nglobal,
                          Nlocal,
                          NlocalCache,
@@ -115,17 +115,17 @@ likfitLgmGpu <- function(data,
   ssqResidual <- ssqY - ssqBetahat
   # any(is.na(as.matrix(log(ssqResidual/Nobs))))
    # as.matrix(cholDiagMat)[]
-   any(is.nan(as.matrix(log(ssqResidual/Nobs))))
-   any(is.nan(as.vector(detVar)))
-   any(is.na(as.matrix(varMat)))
-   any(is.nan(as.matrix(varMat)))
-   any(is.na(as.matrix(ssqYX)))
-   any(is.na(as.matrix(ssqY)))
-   any(is.na(as.matrix(ssqYXcopy)))
-   any(is.na(as.matrix(ssqBetahat)))
-   any(is.na(as.vector(detReml)))
-   any(is.na(as.matrix(cholXVXdiag)))
-   any(is.na(as.matrix(ssqResidual)))
+   # any(is.nan(as.matrix(log(ssqResidual/Nobs))))
+   # any(is.nan(as.vector(detVar)))
+   # any(is.na(as.matrix(varMat)))
+   # any(is.nan(as.matrix(varMat)))
+   # any(is.na(as.matrix(ssqYX)))
+   # any(is.na(as.matrix(ssqY)))
+   # any(is.na(as.matrix(ssqYXcopy)))
+   # any(is.na(as.matrix(ssqBetahat)))
+   # any(is.na(as.vector(detReml)))
+   # any(is.na(as.matrix(cholXVXdiag)))
+   # any(is.na(as.matrix(ssqResidual)))
    # any(is.nan(as.matrix(ssqResidual/Nobs)))
    # as.matrix(ssqY)[which(is.na(as.vector(detVar))),]
   # which(is.na(as.vector(detVar)))
@@ -185,11 +185,11 @@ likfitLgmGpu <- function(data,
   
   LogLikcpu <- as.matrix(-0.5*minusTwoLogLik)
   maximum <- max(LogLikcpu)
-  breaks = maximum - qchisq(cilevel,  df = 1)/2
[…]

### 2021-12-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3056 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{gather*}
-The common choice for the covariance function $\Sigma$ is the Mat\'ern covariance function \citep{matern1960spatial}. The Mat\'ern covariance between $U(x_i)$ and $U(x_j)$ takes the above form.
+A common choice for the covariance function $\Sigma$ is the Mat\'ern covariance \citep{matern1960spatial}. The Mat\'ern covariance between $U(x_i)$ and $U(x_j)$ takes the above form,
 %$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
-$\sigma^2$ the variance of the random field $U$.
-$\Gamma(\cdot)$ is the standard gamma function,  $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as distance increases.
+where $\sigma^2$ the variance of the random field $U$.
+$\Gamma(\cdot)$ is the standard gamma function.  
+$\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as distance increases.
 $K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$, specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.
@@
 The matrix decomposition method is exact, it works for all covariance functions and can generate random field on all types of grids.
-Other \proglang{R} packages that offer simulation of Gaussian random fields like \pkg{geoR} \citep{geoR2001}, does not work for large number of locations; the \pkg{RandomFields} \citep{RandomFields2015,RandomFields2020} package can use different methods for simulation of Gaussian fields, among which the (modified) circulant embedding method \citep{Dietrich1997FastAE} for covariance matrix decomposition is also an exact method (need to confirm???), however, it works only for isotropic Gaussian fields and on rectangular grids.
-\pkg{clrng} does exact simulation of exact Gaussian fields on the GPU as it relies on the matrix decomposition method.
+Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations; the \pkg{RandomFields} \citep{RandomFields2015,RandomFields2020} package can use different methods for simulation of Gaussian fields, among which the modified circulant embedding method \citep{Dietrich1997FastAE} for covariance matrix decomposition is also an exact method, however, it works only for isotropic Gaussian fields and on rectangular grids.
+\pkg{clrng} does simulation of exact Gaussian fields on the GPU as it relies on the matrix decomposition method.
 The implementation of this method is as follows.
@@
 The matrix decomposition method is straightforward to implement.
-However, when a large number of locations $n$ is involved, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
+However, when the number of locations $n$ is large, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
 Our another \proglang{R} package \pkg{gpuBatchMatrix} is able to compute batches of Mat\'ern covariance matrices in parallel, and do Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
-The following shows an example, in which we simulate on GPU 10 Gaussian random fields of mat\'ern covariance at one time with 5 sets of parameters, by taking advantage of the GPU-capabilities provided by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
+The following shows an example, in which we simulate on GPU 10 Gaussian random fields of mat\'ern covariance at one time with 5 sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
 %\pkg{clrng} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
@@
 @
-Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
+Step 1, set up a $51 \times 80$ raster using the 'swissRain' data from \pkg{geostatsp}.
 <<setupcoords, eval=TRUE>>=
@@
 @
-Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets.
-Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuBatchMatrix::maternGpuParam}.
+Step 2, create 4 parameter sets as a small example. 
 <<setup paramsBatch, eval=TRUE,echo=FALSE>>=
@@
 @
-Step 3, compute Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
-are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
+Step 3, compute the Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
+are each of size $4080 \times 4080$ and are stacked by row in the output \code{maternCov}. 
 <<r outputBatchCreate, eval=TRUE>>=
@@
 @
-Step 4, the first argument \code{A} in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
+Step 4, perform Cholesky decomposition of \code{maternCov}.
+The first argument in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
 Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}.
-Each row of \code{diagMat} stores the diagonal values of each $D_i$, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
+The diagonal values of each $D_i$ are returned and stored in each row of \code{diagMat}.
+For example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $D_i$ is $1 \times n$.
 <<r cholBach, eval=TRUE>>=
-diagMat = gpuBatchMatrix::cholBatch(maternCov, 
-                          Nglobal= c(128, 8), Nlocal= c(32, 8))
+diagMat = gpuBatchMatrix::cholBatch(maternCov, Nglobal= c(128, 8), Nlocal= c(32, 8))
 @
@@
 \end{gather}
-Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU to generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 <<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
@@
 @
-Step 6, use \fct{gpuBatchMatrix::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
-\code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
+Step 6, compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration for the matrix operation that \fct{multiplyLowerDiagonalBatch} does.
+\code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
 <<r multLowerdiag_LDZ, eval=TRUE, cache=FALSE, results='hide'>>=
-simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
-                   type = gpuR::typeof(zmatGpu))
-
-gpuBatchMatrix::multiplyLowerDiagonalBatch(
-  output=simMat, L=maternCov, 
-  D=diagMat, B=zmatGpu,
+simMat = gpuBatchMatrix::multiplyLowerDiagonalBatch(maternCov, 
+  diagMat, zmatGpu,
   diagIsOne = TRUE,   
   transformD = "sqrt", 
-  Nglobal, Nlocal, NlocalCache)
+  Nglobal=c(128, 64, 2), 
+  Nlocal= c(8, 2, 1), 
+  NlocalCache=1000)
 @

### 2021-12-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3057 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\begin{algorithm}[ht]
+\begin{algorithm}[H]
 \SetAlgoLined
@@
 )
-#c(shape=2.15, range=40*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/4))
 @
@@
 params
-#paramsGpu = gpuBatchMatrix::maternGpuParam(myParamsBatch, type=theType)
 @
@@
 <<r outputBatchCreate, eval=TRUE>>=
-#maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
-#                         nrow(coordsGpu),type=theType)
 maternCov = gpuBatchMatrix::maternBatch(

### 2021-12-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3056 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3057 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.047   0.000   0.047
+##   0.051   0.000   0.052
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   6.567   0.818   7.378
+##   6.978   0.847   7.821
 \end{CodeOutput} 
@@
 ##       [,1]  [,2]  [,3]   [,4]
-## [1,] 0.367 0.0488 0.490 0.308
-## [2,] 0.671 3.1532 0.831 0.281
+## [1,] 0.565 0.333 0.395 0.0932
+## [2,] 1.535 0.866 0.124 0.0315
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.319   0.000   0.318
+##   0.319   0.003   0.321
 \end{CodeOutput}
@@
 R> system.time(result_monthcpu <- stats::fisher.test(month, simulate.p.value = TRUE,
-+    B = 1e+06))
++    B = 1000000))
 \end{CodeInput}
@@
 ##    user  system elapsed 
-##  14.478   0.012  14.480
+##  14.876   0.016  14.879
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.403
+## [1] 0.405
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.941   0.015   1.954
+##    1.93    0.02    1.95
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   90.70    0.12   90.82
+##  92.369   0.143  92.455
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.000121
+## [1] 0.000127
 \end{CodeOutput} 
@@
 \hline
-\hspace{1em}1M & 10.74 & 2.28 & 14.48 & 0.34 & month\\
+\hspace{1em}1M & 10.74 & 2.28 & 14.48 & 0.32 & month\\
 \hline
@@
 
-Step 1, set up a $60 \times 80$ raster and coordinates on the raster, convert the matrix of coordinates \code{coordsSp@coords} to a ``vclMatrix'' object for later use.
+Step 1, set up a $51 \times 80$ raster using the 'swissRain' data from \pkg{geostatsp}.
 \begin{CodeChunk}
@@
 
-Step 2, create 5 parameter sets as a small example, in practical studies, there can be hundreds or thousands of parameter sets.
-Then, convert the matrix of parameters to a ``vclMatrix'' for later use by \fct{gpuBatchMatrix::maternGpuParam}.
+Step 2, create 4 parameter sets as a small example.
 
@@
 \end{CodeOutput} 
-\begin{CodeInput}
-R> # paramsGpu =
-R> # gpuBatchMatrix::maternGpuParam(myParamsBatch,
-R> # type=theType)
-\end{CodeInput} 
 \end{CodeChunk}
 
-Step 3, compute Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
-are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
+Step 3, compute the Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
+are each of size $4080 \times 4080$ and are stacked by row in the output \code{maternCov}. 
 \begin{CodeChunk}
 \begin{CodeInput}
-R> # maternCov = vclMatrix(0,
-R> # nrow(paramsGpu)*nrow(coordsGpu),
-R> # nrow(coordsGpu),type=theType)
 R> maternCov = gpuBatchMatrix::maternBatch(params, myRaster,
@@
 
-Step 4, the first argument \code{A} in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
+Step 4, perform Cholesky decomposition of \code{maternCov}.
+The first argument in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
 Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}.
-Each row of \code{diagMat} stores the diagonal values of each $D_i$, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
+The diagonal values of each $D_i$ are returned and stored in each row of \code{diagMat}.
+For example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $D_i$ is $1 \times n$.
 \begin{CodeChunk}
@@
 \end{gather}
-Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU to generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 \begin{CodeChunk}
@@
 
-Step 6, use \fct{gpuBatchMatrix::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration.
-\code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
+Step 6, compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration for the matrix operation that \fct{multiplyLowerDiagonalBatch} does.
+\code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
 \begin{CodeChunk}
 \begin{CodeInput}
-R> simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), type = gpuR::typeof(zmatGpu))
-R> gpuBatchMatrix::multiplyLowerDiagonalBatch(output = simMat,
-+    L = maternCov, D = diagMat, B = zmatGpu, diagIsOne = TRUE,
-+    transformD = "sqrt", Nglobal, Nlocal, NlocalCache)
+R> simMat = gpuBatchMatrix::multiplyLowerDiagonalBatch(maternCov,
++    diagMat, zmatGpu, diagIsOne = TRUE, transformD = "sqrt",
++    Nglobal = c(128, 64, 2), Nlocal = c(8, 2, 1), NlocalCache = 1000)
 \end{CodeInput} 

### 2021-12-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3058 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.28,    10.82),
-  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.48,  91.00),
+  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.80,  92.00),
   var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.32,   1.95),

### 2021-12-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3058 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3059 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3064 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Parallel processing by graphics processing unit (GPU) can be used to speed up computationally intensive tasks, which when combined with \proglang{R}, it can largely improve \proglang{R}’s downsides in terms of slow speed, memory usage and computation mode.
-There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling.
+%There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling.
 \pkg{clrng} enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
@@
 
-Graphics Processing Units (GPUs) have the potential to make important contribution to parallel computing with \proglang{R}.
-GPUs can perform thousands of tests simultaneously, which makes them powerful for doing massively parallel computing, and they are relatively cheap compared to multicore CPU's.
+Graphics Processing Units (GPUs) have the potential to make an important contribution to parallel computing with \proglang{R}.
+GPUs can perform thousands of computations simultaneously, which makes them powerful for doing massively parallel computing, and they are relatively cheap compared to multicore CPU's.
 Although there have been a number of \proglang{R} packages developed which provide some GPU capability, they inevitably come with some limitations.
-Packages such as \pkg{gputools}, \pkg{gpumatrix}, \pkg{cudaBayesreg}, \pkg{rpud} (available on github), are now no longer maintained, the popular \pkg{tensorflow} \citep{tensorflow1} package uses GPU via \proglang{Python}, which makes it difficult to include as a dependency for new \proglang{R} packages.
+Packages such as \pkg{gputools}, \pkg{gpumatrix}, \pkg{cudaBayesreg}, \pkg{rpud} (available on github), are no longer maintained, the popular \pkg{tensorflow} \citep{tensorflow1} package uses GPU via \proglang{Python}, which makes it difficult to include as a dependency for new \proglang{R} packages.
 All of these mentioned packages are restricted to \proglang{R} users with NVIDIA GPUs.
@@
 The \pkg{clrng} package described here is currently the only \proglang{R} package that is able to generate random numbers on GPUs.
-Accomplishing this is complicated to do because each process must produce an independent sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of numbers at any point.
+Accomplishing this is complicated because each process must produce an independent sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of numbers at any point.
 Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \proglang{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
@@
 Then we provide two real data examples to demonstrate the function usage and its \proglang{R} performance.
-Section 5, we show an useful application of normal random numbers on GPU, using them to simulate batches of Gaussian random surfaces with Mat\'ern covariance matrices simultaneously, the simulation also uses GPU-enabled functions from our another package \pkg{gpuBatchMatrix}.
+In Section 5, we show a useful application of normal random numbers on GPU, using them to simulate batches of Gaussian random surfaces with Mat\'ern covariance matrices simultaneously, the simulation also uses GPU-enabled functions from our other package \pkg{gpuBatchMatrix}.
 Finally, the paper concludes with a short summary and a discussion in Section 6.
@@
 These four RNGs use different types of constructions.
-\pkg{clrng} package uses the MRG31k3p RNG, making it able to generate random numbers on GPUs.
+The \pkg{clrng} package uses the MRG31k3p RNG, making it able to generate random numbers on GPUs.
 We choose MRG31k3p as the base generator for the following reasons:  the original \pkg{RNGStreams} package \citep{l2002object} was built with MRG32k3a, which was designed to be implemented in double precision, and not with 32-bit integers.
 The MRG31k3p generator was designed later, specifically for 32-bit integer arithmetic, so it runs faster on the 32-bit GPUs.
-It is also faster than Philox-4×32-10 \citep{rLEC00b}.
+The MRG31k3p generator was introduced in the paper \citep{rLEC00b}.
+It is also faster than Philox-4×32-10.
 MRG31k3p was also statistically tested extensively and successfully, \cite[see][]{rLEC07b}.
@@
 Creating a new stream amounts to computing its starting point.
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, although this is not currently implemented in \pkg{clrng}..
 In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
 Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
-Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}.
 
-In \proglang{clRNG} library, the MRG31k3p RNG's entire period of length approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
-Each stream can be further partitioned into substreams of length $2^{72}$, although this is not currently implemented in \pkg{clrng}.
+In the \proglang{clRNG} library, the MRG31k3p RNG's entire period of length approximately $2^{185}$ is divided into approximately $2^{51}$ non-overlapping streams of length $Z = 2^{134}$.
+%Each stream can be further partitioned into substreams of length $2^{72}$.
 The state (and seed) of each stream is a vector of six 31-bit integers.
@@
 Listing \ref{lst:normal} shows a fragment of the kernel that generates standard Gaussian random numbers.
-Aside from using local memory and the part that does the transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}.
+Aside from using local memory and the part that does the Box-Muller transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}.
 If developers want to use other Gaussian transform methods, this kernel is easy to be adapted and incorporated in other \proglang{R} packages.
-The kernel executes over a 2-dimensional index space with $1 \times 2$ work-groups.
+This kernel executes over a 2-dimensional index space with $1 \times 2$ work-groups.
 \code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
 \code{PI_2} is defined to be $\pi/2$.
-As the work-items in a work-group proceed at differing rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
+As the work-items in a work-group proceed at different rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
 %
@@
 ##    user  system elapsed 
-##   0.051   0.000   0.052
+##   0.047   0.000   0.048
 \end{CodeOutput} 
@@
 
-Below is an example that produces a matrix of Exponential random numbers with expectation equal to 1.
+Below is an example that produces a $2 \times 4$ matrix of Exponential random numbers with expectation equal to 1.
 \begin{CodeChunk}
@@
 ##    user  system elapsed 
-##   0.319   0.003   0.321
+##   0.311   0.004   0.315
 \end{CodeOutput}
@@
 R> system.time(result_monthcpu <- stats::fisher.test(month, simulate.p.value = TRUE,
-+    B = 1000000))
++    B = 1015808))
 \end{CodeInput}
@@
 ##    user  system elapsed 
-##  14.876   0.016  14.879
+##  14.791   0.027  14.826
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##    1.93    0.02    1.95
+##   1.944   0.021   1.963
 \end{CodeOutput}
@@
 
-The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value.
+The ``week'' table has a much smaller p-value: around 0.000125, which should require a larger number of simulations to get a more accurate p-value.
 With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}.
-\fct{stats::fisher.test} takes 91.6 seconds, while \fct{fisher.sim} takes about 4.282 seconds, the elapsed time is decreased to about 4.67\% of the time taken on CPU.
+\fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU.
 
@@
 \subsection{A summary of the results}
-We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure %\ref{fig4}.
+We summarized the comparison results in Table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
 \begin{table}
@@
 \hline
-\hspace{1em}1M & 10.74 & 2.28 & 14.48 & 0.32 & month\\
+\hspace{1em}1M & 10.74 & 2.28 & 14.8 & 0.32 & month\\
 \hline
-\hspace{1em}10M & 63.24 & 10.82 & 91 & 1.95 & week\\
+\hspace{1em}10M & 63.24 & 10.82 & 92 & 1.95 & week\\
 \hline
@@
 
-\caption[Approximate sampling distributions of the test statistics from the two examples Month and Week]{Approximate sampling distributions of the test statistics from the two examples Month and Week.
-The test statistic values of the observed tables are marked with a blue line on each plot.}\label{fig:fighistMonth}
+\caption[\label{fig4}Approximate sampling distributions of the test statistics from the two examples Month and Week]{\label{fig4}Approximate sampling distributions of the test statistics from the two examples Month and Week.
+The test statistic values of the observed tables are marked with a blue vertical line on each plot.}\label{fig:fighistMonth}
 \end{figure}
@@
 \begin{gather*} 
-U = [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma) \\
-\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|x_i = x_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|x_i = x_j|}{\phi}\right).
+U = [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
+\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|x_i = x_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|x_i = x_j|}{\phi}\right),
 \end{gather*}
-A common choice for the covariance function $\Sigma$ is the Mat\'ern covariance \citep{matern1960spatial}. The Mat\'ern covariance between $U(x_i)$ and $U(x_j)$ takes the above form,
+where \text{MVN} standards for the multivariate normal distribution, and $\Sigma$ is the covariance matrix.
+A common choice for $\Sigma$ is the Mat\'ern covariance \citep{matern1960spatial}.
+The Mat\'ern covariance between $U(x_i)$ and $U(x_j)$ takes the above form,
 %$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
@@
 %(For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.)
-There are several alternative parameterisations of Mat\'ern covariance functions in literatures  \citep[see][]{haskard2007anisotropic}.
+There are several alternative parameterisations of Mat\'ern covariance functions in literature  \citep[see][]{haskard2007anisotropic}.
 In \pkg{clrng}, we use the above form for computing Mat\'ern covariance.
@@
 %Due to the importance of Gaussian random field, there are a large amount of studies devoted to Gaussian random field generation techniques. 
-\cite{LiuandLi2019} gives a comprehensive review on 7 popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
+\cite{LiuandLi2019} gives a comprehensive review on seven popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
 all of these methods except matrix decomposition method, are approximations and has specific requirements on the type of grid or covariance functions.
@@
 
-Suppose $U=(U_1, U_2, \dots, U_n)$ is a Gaussian random field with mean $\mu$ and covariance matrix $\Sigma$, without loss of generality, we assume mean value zero $(\mu = 0)$.
+Suppose $U=(U_1, U_2, \dots, U_n)^\top$ is a Gaussian random field with mean $\mu$ and covariance matrix $\Sigma$, without loss of generality, we assume mean value zero $(\mu = 0)$.
 Since covariance matrix $\Sigma$ is symmetric and positive-definite, we can take Cholesky decomposition of $\Sigma$, then we can simulate random samples from $U$ using Algorithm \ref{algorithm2},
@@
  1, Calculate the covariance matrix $\Sigma$ between locations\;
- 2, Compute the Cholesky decomposition of $\Sigma = L \cdot D \cdot L^\top$, $L$ is the lower unit triangular matrix, $D$ is a diagonal matrix\;
+ 2, Compute the Cholesky decomposition of $\Sigma = L \cdot D \cdot L^\top$, where $L$ is a lower unit triangular matrix, and $D$ is a diagonal matrix\;
  3, Generate on GPU a random matrix $Z=(Z_1, Z_2, \dots, Z_n) \sim \text{MVN}(0,I_n)$, where $I_n$ is a $n \times n$ identity matrix\;
@@
 However, when the number of locations $n$ is large, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
-Our another \proglang{R} package \pkg{gpuBatchMatrix} is able to compute batches of Mat\'ern covariance matrices in parallel, and do Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
-The following shows an example, in which we simulate on GPU 10 Gaussian random fields of mat\'ern covariance at one time with 5 sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
+Our other \proglang{R} package \pkg{gpuBatchMatrix} is able to compute batches of Mat\'ern covariance matrices in parallel, and do Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
+The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
 %\pkg{clrng} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
@@
 Step 4, perform Cholesky decomposition of \code{maternCov}.
-The first argument in \fct{gpuBatchMatrix::cholBatch} specifies the object to take Cholesky decomposition.
+The first argument in \fct{cholBatch} specifies the object to take Cholesky decomposition.
 Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}.
@@
 The package \pkg{clrng} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
-We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
+We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute Mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
 Most of these functions uses local memory on GPU.
@@
 
-\pkg{clrng} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package, if developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
+\pkg{clrng} is limited by the number and type of \proglang{OpenCL} RNGs used and the features of the \pkg{gpuR} package, as it depends upon the \pkg{gpuR} package.
+If developers wants to use our package to do something that's not supported in \pkg{gpuR}, for example ``sparse'' class objects, they would have to write \proglang{OpenCL} code to implement it.
 

### 2021-12-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3059 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3064 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Below is an example that produces a matrix of Exponential random numbers with expectation equal to 1.
+Below is an example that produces a $2 \times 4$ matrix of Exponential random numbers with expectation equal to 1.
 <<random exponential, eval=TRUE, cache=FALSE>>=
@@
  #using CPU
- system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1e6))
+ system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1015808))
  result_monthcpu$p.value
@@
 @
-The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value.
+The ``week'' table has a much smaller p-value: around 0.000125, which should require a larger number of simulations to get a more accurate p-value.
 With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}.
-\fct{stats::fisher.test} takes 91.6 seconds, while \fct{fisher.sim} takes about 4.282 seconds, the elapsed time is decreased to about 4.67\% of the time taken on CPU.
+\fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU.
 <<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
@@
 \subsection{A summary of the results}
-We summarized the comparison results in the table \ref{tab:summary} and plot the test statistics in Figure %\ref{fig4}.
+We summarized the comparison results in Table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
 <<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE, fig.pos='h'>>=
@@
 
-<<fighistMonth, eval=TRUE, echo=FALSE, dev='pdf', fig.cap = "Approximate sampling distributions of the test statistics from the two examples Month and Week.
-The test statistic values of the observed tables are marked with a blue line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1>>=
+<<fighistMonth, eval=TRUE, echo=FALSE, dev='pdf', fig.cap = "\\label{fig4}Approximate sampling distributions of the test statistics from the two examples Month and Week.
+The test statistic values of the observed tables are marked with a blue vertical line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1>>=
 hist(month_stats, xlab="test statistic",#TeX('-$\\sum(\\log(n_{ij}!))$'),

### 2021-12-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3065 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   University of Montr\'eal\\
-  Pavillon André-Aisenstadt 2920, chemin de la Tour Montréal QCH3T 1J4\\
+  Pavillon André-Aisenstadt, 
+   2920 chemin de la Tour, 
+   Montréal, QC,  Canada,  H3T 1J4\\
   E-mail: \email{lecuyer@iro.umontreal.ca}}

### 2021-12-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3065 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.047   0.000   0.048
+##   0.038   0.010   0.049
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.311   0.004   0.315
+##   0.323   0.003   0.325
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.944   0.021   1.963
+##   1.932   0.025   1.955
 \end{CodeOutput}

### 2021-12-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper 1 bib
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3067 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The \pkg{clrng} package described here is currently the only \proglang{R} package that is able to generate random numbers on GPUs.
-Accomplishing this is complicated because each process must produce an independent sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of numbers at any point.
+Accomplishing this is complicated because each process must produce an independent (non-overlapping) sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of random numbers at any point.
 Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \proglang{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
@@
 \subsection{Creating streams}\label{createstreams}
-Random number generation is more complicated when the RNG algorithm is run in parallel processes in \proglang{R}, there is a risk that the generated sequences of random numbers have correlation.
+Random number generation is more complicated when RNG is used in parallel processes in \proglang{R}, there is a risk that the generated sequences of random numbers have correlation.
 \pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
@@
 Creating a new stream amounts to computing its starting point.
-Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, although this is not currently implemented in \pkg{clrng}..
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, although this is not currently implemented in \pkg{clrng}.
 In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
@@
 The streams created sequentially are then used by the GPU's work-items to generate random numbers.
-In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be greater than the number of total work-items in use.
+In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be at least the number of total work-items in use.
 The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.

### 2021-12-25 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1 bib
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3066 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
streamscpu
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3068 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The MRG31k3p generator was designed later, specifically for 32-bit integer arithmetic, so it runs faster on the 32-bit GPUs.
+It is also faster than Philox-4×32-10.
 The MRG31k3p generator was introduced in the paper \citep{rLEC00b}.
-It is also faster than Philox-4×32-10.
 MRG31k3p was also statistically tested extensively and successfully, \cite[see][]{rLEC07b}.
@@
 ##    user  system elapsed 
-##   0.038   0.010   0.049
+##   0.048   0.004   0.051
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.323   0.003   0.325
+##   0.311   0.005   0.315
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.932   0.025   1.955
+##   1.931   0.023   1.953
 \end{CodeOutput}

### 2021-12-26 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
streamscpu
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3068 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3069 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-%\usepackage{float}
+\usepackage{float}
 %\usepackage{framed}
@@
 \subsection{Creating streams}\label{createstreams}
-Random number generation is more complicated when RNG is used in parallel processes in \proglang{R}, there is a risk that the generated sequences of random numbers have correlation.
+When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation would be more complicated because there is no guarantee that the streams (or similarly \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers is possible to have correlation.
 \pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
@@
 In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
-Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by Z steps to find its initial state, and the three states in the stream object are set to it.
+Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by $Z$ steps to find its initial state, and the three states in the stream object are set to it.
 
@@
 This size of state is appropriate for having streams running in work-items on GPU cards, while providing a sufficient period length for most applications.
-The initial state of the first stream (also called ``initial seed for the package'' in \proglang{clRNG} and ``initial'' in \pkg{clrng}) for the MRG31k3p is by default $(12345, 12345, 12345, 12345, 12345, 12345)$.
+The initial state of the first stream (also called ``initial seed of the package'' or ``initial seed of the stream creator'' in \proglang{clRNG}) for the MRG31k3p RNG is by default $(12345, 12345, 12345, 12345, 12345, 12345)$.
 
@@
   \item \code{n}: number of streams to create.
-  \item \code{initial}: vector of length 6, recycled if shorter.
+  \item \code{initial}: a vector of six 31-bit integers, recycled if length is shorter. Default is set to \code{c(12345, 12345, 12345, 12345, 12345, 12345)}.
 \end{itemize}
@@
 \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-However, any data is lost when the \proglang{R} session is terminated.
+Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} and stick with that choice for their entire program, as we want to avoid the very small possibility of overlap between the streams with different initial seeds.
+The `\code{initial}' argument should be set only once at the very beginning of the program in each \proglang{R} session if users want to supply their own initial seed of the stream creator, or never, in which case users would go with the package default initial seed.
+After the `\code{initial}' argument is set, we can ignore the `\code{initial}' argument and call \fct{createStreamsCpu} or \fct{createStreamsGpu} once or several times to create various matrices of streams in the same program.
+
+%However, any data is lost when the \proglang{R} session is terminated.
 %\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
@@
 Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
-<<createStreams_Gpu>>=
+<<createStreams_Gpu, eval=FALSE>>=
 # creating streams on GPU
 myStreamsGpu = vclMatrix(myStreamsCpu)
+@
+<<createStreams_Gpu2>>=
 myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
@@
 \code{index} represents the position of a work-item when mapping it to a one-dimensional data structure.
-Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states, respectively, and \code{startvalue} tells the starting position of the particular stream to be copied to the work-item.
+Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states, respectively, and \code{startvalue} tells the starting position of the particular stream to be copied to its corresponding work-item.
 The function \fct{clrngMrg31k3pNextState} generates an uniform random integer named \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type.
@@
 
-Now we use the streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d.
+Now we use the 4 streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d.
 $U(0,1)$ random numbers with \fct{clrng::runif}.
@@
 <<simStuff2, eval=TRUE,tidy=TRUE, echo=TRUE>>=
-as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2)))
+as.vector(clrng::runif(n=6, streams=myStreamsGpu2, Nglobal=c(2,2)))
 @
@@
 %Below shows the \code{myStreamsGpu} after it being used.
-<<eval=TRUE,tidy=TRUE, include=FALSE>>=
+<<eval=FALSE,tidy=TRUE, include=FALSE>>=
 t(matrix(as.matrix(myStreamsGpu), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
@@
 Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
-However, we can save the streams on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, in this way we can reproduce the exact same sequence of random numbers in simulations.
+However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will restart from the current states after they are reloaded.
+In this way we can reproduce the exact same sequence of random numbers in simulations.
 Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
-<<save streams,echo=TRUE,eval=TRUE>>=
+<<save streams,echo=TRUE,eval=FALSE>>=
 saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
@@
   \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: streams. If missing, a stream with random initial state is supplied.
+  \item \code{streams}: streams. If missing, streams created on GPU with package default initial seed are supplied.
   \item \code{Nglobal}: global index space. Default is $(64,8)$.
@@
  <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE>>=
- streams <- createStreamsGpu(n =256*64, initial=666)
+ streams <- createStreamsGpu(n =256*64)
  month_gpu<-vclMatrix(month,type="integer")
@@
 We summarized the comparison results in Table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
-<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE, fig.pos='h'>>=
+<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE,fig.pos='H'>>=
 library(kableExtra)
@@
 Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
-  kable_styling(full_width = FALSE, position = "center") %>%
+  kable_styling(full_width = FALSE, position = "center", latex_options = "HOLD_position") %>%
   kableExtra::group_rows(index = c("P-value" = 2, "Run-time" = 2)) %>%
@@
 <<fighistMonth, eval=TRUE, echo=FALSE, dev='pdf', fig.cap = "\\label{fig4}Approximate sampling distributions of the test statistics from the two examples Month and Week.
-The test statistic values of the observed tables are marked with a blue vertical line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1>>=
+The test statistic values of the observed tables are marked with a blue vertical line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1,fig.pos='H'>>=
 hist(month_stats, xlab="test statistic",#TeX('-$\\sum(\\log(n_{ij}!))$'),

### 2021-12-26 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3069 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 R> myStreamsGpu = vclMatrix(myStreamsCpu)
+\end{CodeInput} 
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
 R> myStreamsGpu2 = createStreamsGpu(n = 4, initial = 12345)
@@
 \begin{CodeInput}
-R> as.vector(clrng::runif(n = 6, streams = myStreamsGpu, Nglobal = c(2,
+R> as.vector(clrng::runif(n = 6, streams = myStreamsGpu2, Nglobal = c(2,
 +    2)))
@@
 Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
-However, we can save the streams on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, in this way we can reproduce the exact same sequence of random numbers in simulations.
+However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will restart from the current states after they are reloaded.
+In this way we can reproduce the exact same sequence of random numbers in simulations.
 Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
@@
 ##    user  system elapsed 
-##   0.048   0.004   0.051
+##   0.043   0.007   0.050
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   6.978   0.847   7.821
+##   6.947   0.736   7.686
 \end{CodeOutput} 
@@
 ##        [,1]  [,2]  [,3]  [,4]
-## [1,] 0.565 0.333 0.395 0.0932
-## [2,] 1.535 0.866 0.124 0.0315
+## [1,] 0.1187 0.187 1.079 0.112
+## [2,] 0.0501 0.369 0.411 1.880
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> streams <- createStreamsGpu(n = 256 * 64, initial = 666)
+R> streams <- createStreamsGpu(n = 256 * 64)
 R> month_gpu <- vclMatrix(month, type = "integer")
@@
 ##    user  system elapsed 
-##   0.311   0.005   0.315
+##   0.318   0.005   0.322
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 410825
+## [1] 409435
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.404
+## [1] 0.403
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##  14.791   0.027  14.826
+##  14.859   0.014  14.874
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.931   0.023   1.953
+##    1.93    0.02    1.95
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 1193
+## [1] 1351
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.000119
+## [1] 0.000135
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##  92.369   0.143  92.455
+##  91.311   0.186  91.432
 \end{CodeOutput}
@@
 We summarized the comparison results in Table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
-\begin{table}
+\begin{table}[H]
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 

### 2021-12-26 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3070 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} and stick with that choice for their entire program, as we want to avoid the very small possibility of overlap between the streams with different initial seeds.
+Users may choose to use either \fct{createStreamsCpu} or \fct{createStreamsGpu} and stick with that choice for their entire program, as we want to avoid the very small possibility of overlap between the streams with different initial seeds.
 The `\code{initial}' argument should be set only once at the very beginning of the program in each \proglang{R} session if users want to supply their own initial seed of the stream creator, or never, in which case users would go with the package default initial seed.
@@
   \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: streams for random number generation.
+  \item \code{streams}: streams for random number generation. If missing, streams created on GPU with package default initial seed are supplied.
   \item \code{Nglobal}: global index space. Default is set as $(64,8)$.

### 2021-12-26 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3070 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.043   0.007   0.050
+##   0.052   0.000   0.051
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.318   0.005   0.322
+##   0.321   0.003   0.324
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##    1.93    0.02    1.95
+##   1.934   0.021   1.953
 \end{CodeOutput}

### 2021-12-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3072 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Given a parameter space $X$, a random field $U$ on $X$ is a collection of random variables $\{U_x :x\in X\}$.
-A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})$ is multivariate Gaussian.
-The expectation of $U$ and covariance function $\Sigma$ completely determine the distribution of $U$, where
+A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Gaussian.
+The expectation and covariance function of $U$ completely determine the distribution of $U$,
 \begin{gather*} 
@@
 \end{gather*}
-where \text{MVN} standards for the multivariate normal distribution, and $\Sigma$ is the covariance matrix.
+where \text{MVN} standards for the multivariate normal distribution, and $\Sigma$ standards for the covariance matrix of $U$.
 A common choice for $\Sigma$ is the Mat\'ern covariance \citep{matern1960spatial}.
@@
 $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as distance increases.
-$K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$, specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.
+$K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$.
+%specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.
 %(For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.)
@@
 \cite{LiuandLi2019} gives a comprehensive review on seven popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
-all of these methods except matrix decomposition method, are approximations and has specific requirements on the type of grid or covariance functions.
-The matrix decomposition method is exact, it works for all covariance functions and can generate random field on all types of grids.
-Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations; the \pkg{RandomFields} \citep{RandomFields2015,RandomFields2020} package can use different methods for simulation of Gaussian fields, among which the modified circulant embedding method \citep{Dietrich1997FastAE} for covariance matrix decomposition is also an exact method, however, it works only for isotropic Gaussian fields and on rectangular grids.
-\pkg{clrng} does simulation of exact Gaussian fields on the GPU as it relies on the matrix decomposition method.
+all of these methods except the matrix decomposition method, are approximations to the random field and have specific requirements on the type of grid or covariance functions.
+The matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \cite{LiuandLi2019}.
+Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations.
+The \pkg{RandomFields} \citep{RandomFields2015} package has several different methods for simulation of Gaussian fields, among which the circulant embedding, which is an improved method on direct matrix decomposition, and some variants of the method like the cut-off embedding \citep{gneiting2006fast} are also exact and fast for isotropic Gaussian random fields, however, they work only on rectangular grids.
+\pkg{clrng} is able to simulate on the GPU exact Gaussian fields on arbitrary grids as it relies on the matrix decomposition method.
 The implementation of this method is as follows.
+
+%it can generate random fields on arbitrary position Circulant embedding is a fast simulation method for stationary (possibly anisotropic) Gaussian ran- dom fields on regular grids based on Fourier transformations.
+Cut-off embedding is a fast simulation method for stationary, isotropic Gaussian random fields on square lattices based on the standard RPcirculant method, so that exact simulation is garantueed for further covariance models
 
@@
 % \end{itemize}
-The matrix decomposition method is straightforward to implement.
-However, when the number of locations $n$ is large, the cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and the cost of matrix-vector multiplication $L*Z$ to generate the random field $U$ is $\mathcal{O}(n^2)$ \citep{LiuandLi2019}.
-Our other \proglang{R} package \pkg{gpuBatchMatrix} is able to compute batches of Mat\'ern covariance matrices in parallel, and do Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
+The downside of this method is that it's computationally expensive and is thus very slow.
+When the number of locations $n$ is large, the time cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and is $\mathcal{O}(n^2)$ for the matrix-vector multiplication $L*Z$ \citep{LiuandLi2019}.
+Our other \proglang{R} package \pkg{gpuBatchMatrix} is able to address this issue since it computes batches of Mat\'ern covariance matrices in parallel, and does Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
 The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
@@
 @
-Step 1, set up a $51 \times 80$ raster using the 'swissRain' data from \pkg{geostatsp}.
+Step 1, set up a $51 \times 80$ raster using the `swissBorder' coordinates from \pkg{geostatsp}.
 <<setupcoords, eval=TRUE>>=
@@
 @
-Step 2, create 4 parameter sets as a small example. 
+Step 2, create four parameter sets as a small example. 
 <<setup paramsBatch, eval=TRUE,echo=FALSE>>=
@@
 @
-Step 4, perform Cholesky decomposition of \code{maternCov}.
+Step 4, perform Cholesky decomposition on \code{maternCov}.
 The first argument in \fct{cholBatch} specifies the object to take Cholesky decomposition.
-Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}.
+Computed unit lower triangular matrices $L_i$'s are stacked by row in \code{maternCov}.
 The diagonal values of each $D_i$ are returned and stored in each row of \code{diagMat}.
-For example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $D_i$ is $1 \times n$.
+So if each batch $\Sigma_i$ is of size $n \times n$, then each batch $D_i$ is $1 \times n$.
 <<r cholBach, eval=TRUE>>=
@@
 \end{gather}
-Step 5, create some streams on GPU to generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 <<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
@@
 \end{gather}
-Step 7, Finally, plot the 10 realizations.
+Step 7, Finally, plot the ten realizations.
 <<r plotSimSetup, echo=TRUE, eval=TRUE, cache=FALSE>>=
@@
 \section{Discussion}
-The package \pkg{clrng} has been created to make GPU-generated uniform and normal random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting seeds in streams on GPU.
-We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and (exact) Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, compute Mat\'ern covariance matrices, do Cholesky decomposition and matrix multiplication in batches on GPU.
-Most of these functions uses local memory on GPU.
+The package \pkg{clrng} has been created to make GPU-generated uniform and some non-uniform random numbers accessible for \proglang{R} users, it enables reproducible research in simulations by setting initial seeds in streams and saving and reloading the current states of streams.
+We further applied the GPU-generated random numbers in suitable statistical simulations, such as the Monte Carlo simulation for Fisher’s exact test, and exact Gaussian spatial surfaces simulation, for which we are able to calculate quantiles for the normal distribution, our package \pkg{gpuBatchMatrix} adds additional help in the second application by taking the computational work for batches of Mat\'ern covariance matrices, Cholesky decomposition and matrix multiplication on GPU.
+%Most of these functions uses local memory on GPU.
 Comparison of performance between using \pkg{clrng} and using classic \proglang{R} on CPU for some real data examples has demonstrated significant improvement in execution time.

### 2021-12-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3072 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.052   0.000   0.051
+##   0.051   0.001   0.052
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.321   0.003   0.324
+##   0.315   0.008   0.322
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.934   0.021   1.953
+##   1.927   0.019   1.945
 \end{CodeOutput}
@@
 
-Step 1, set up a $51 \times 80$ raster using the 'swissRain' data from \pkg{geostatsp}.
+Step 1, set up a $51 \times 80$ raster using the `swissBorder' coordinates from \pkg{geostatsp}.
 \begin{CodeChunk}
@@
 
-Step 2, create 4 parameter sets as a small example.
+Step 2, create four parameter sets as a small example.
 
@@
 
-Step 4, perform Cholesky decomposition of \code{maternCov}.
+Step 4, perform Cholesky decomposition on \code{maternCov}.
 The first argument in \fct{cholBatch} specifies the object to take Cholesky decomposition.
-Computed unit lower triangular matrices $L_i$'s are stacked by row and stored in \code{maternCov}.
+Computed unit lower triangular matrices $L_i$'s are stacked by row in \code{maternCov}.
 The diagonal values of each $D_i$ are returned and stored in each row of \code{diagMat}.
-For example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $D_i$ is $1 \times n$.
+So if each batch $\Sigma_i$ is of size $n \times n$, then each batch $D_i$ is $1 \times n$.
 \begin{CodeChunk}
@@
 \end{gather}
-Step 5, create some streams on GPU to generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 \begin{CodeChunk}
@@
 \end{gather}
-Step 7, Finally, plot the 10 realizations.
+Step 7, Finally, plot the ten realizations.
 \begin{CodeChunk}

### 2021-12-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3073 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   700 University Ave., Toronto, ON M5G 1Z5, Canada\\
-  E-mail: \email{ruoyong.xu@mail.utoronto.ca}, \email{patrick.brown@utoronto.ca}}
-
-\Address{
+  E-mail: \email{ruoyong.xu@mail.utoronto.ca}, \email{patrick.brown@utoronto.ca}\\\\ 
   Pierre L’Ecuyer\\

### 2021-12-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3073 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.051   0.001   0.052
+##   0.048   0.000   0.049
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.315   0.008   0.322
+##   0.308   0.011   0.319
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.927   0.019   1.945
+##   1.944   0.006   1.949
 \end{CodeOutput}

### 2021-12-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3076 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 all of these methods except the matrix decomposition method, are approximations to the random field and have specific requirements on the type of grid or covariance functions.
-The matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \cite{LiuandLi2019}.
+The matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \citep{LiuandLi2019}.
 Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations.

### 2021-12-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3076 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.048   0.000   0.049
+##   0.044   0.004   0.047
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.308   0.011   0.319
+##   0.312   0.006   0.319
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.944   0.006   1.949
+##    1.94    0.01    1.95
 \end{CodeOutput}

### 2021-12-29 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3077 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Accomplishing this is complicated because each process must produce an independent (non-overlapping) sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of random numbers at any point.
-Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \proglang{clRNG} \citep{l2015clrng} library, and can thus accelerate several types of statistical simulation and modelling.
+Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \proglang{clRNG} \citep{l2015clrng} library in \proglang{OpenCL}, and can thus accelerate several types of statistical simulation and modelling.
 
@@
 \subsection{Creating streams}\label{createstreams}
-When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation would be more complicated because there is no guarantee that the streams (or similarly \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers is possible to have correlation.
+When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation would be more complicated because there is no guarantee that the streams (analogous to \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers is possible to have correlation.
 \pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
@@
 ##    user  system elapsed 
-##   0.044   0.004   0.047
+##   0.045   0.004   0.048
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.312   0.006   0.319
+##   0.316   0.001   0.318
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##    1.94    0.01    1.95
+##    1.95    0.01    1.95
 \end{CodeOutput}

### 2021-12-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
patrick edits starting
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3078 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \Abstract{
-We introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by utilizing the \pkg{clRNG} (OpenCL) library.
-Parallel processing by graphics processing unit (GPU) can be used to speed up computationally intensive tasks, which when combined with \proglang{R}, it can largely improve \proglang{R}’s downsides in terms of slow speed, memory usage and computation mode.
+We introduce the \proglang{R} package \pkg{clrng} which leverages the \pkg{gpuR} package to generate random numbers on a Graphics Processing Unit (GPU) with the \pkg{clRNG} (OpenCL) library.
+Parallel processing with GPU's can speed up computationally intensive tasks, which when combined with \proglang{R}, it can largely improve \proglang{R}’s downsides in terms of slow speed, memory usage and computation mode.
 %There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling.
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE}
@@
 When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation would be more complicated because there is no guarantee that the streams (analogous to \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers is possible to have correlation.
+\pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
+A popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long disjoint pieces of equal length $Z$, and use each piece as a separate stream.
+Creating a new stream amounts to computing its starting point.
+Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, although this is not currently implemented in \pkg{clrng}.
+In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by $Z$ steps to find its initial state, and the three states in the stream object are set to it.
+When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation becomes be more complicated because there is no guarantee that the streams (or similarly \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers is possible to have correlation.
 \pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.

### 2021-12-30 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
a few edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3079 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 library("clrng")
-myStreamsCpu <- createStreamsCpu(n=4, initial=12345)
+myStreamsCpu <- createStreamsCpu(n=4, initial=rep(12345, 6))
 t(myStreamsCpu)
@@
 We move streams to the GPU by converting them to a `\code{vclMatrix}'.
+<<createStreams_Gpu, eval=FALSE>>=
+myStreamsGpu = vclMatrix(myStreamsCpu)
+@
+Equivalently, 
  \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-Users may choose to use either \fct{createStreamsCpu} or \fct{createStreamsGpu} and stick with that choice for their entire program, as we want to avoid the very small possibility of overlap between the streams with different initial seeds.
-The `\code{initial}' argument should be set only once at the very beginning of the program in each \proglang{R} session if users want to supply their own initial seed of the stream creator, or never, in which case users would go with the package default initial seed.
-After the `\code{initial}' argument is set, we can ignore the `\code{initial}' argument and call \fct{createStreamsCpu} or \fct{createStreamsGpu} once or several times to create various matrices of streams in the same program.
+<<createStreams_Gpu2>>=
+myStreamsGpu2 = createStreamsGpu(n=4, initial=rep(12345, 6)))
+@
+
+Users should run either \fct{createStreamsCpu} or \fct{createStreamsGpu} at the beginning of a task and retain these streams for all of the random number generation the task requires.
+Creating new streams with different initial states introduces the very small possibility of overlap between the original and new streams.
+If an \proglang{R} session is restarted during a task, perhaps because a preliminary analysis is done following an initial set of simulations and the full set of simulations continued at a later date, the state of the streams should be transferred to the CPU and saved to disk.
+The `\code{initial}' argument to the stream creator should be used with caution and we recommend using the default.
+It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
 
@@
 Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
-<<createStreams_Gpu, eval=FALSE>>=
-# creating streams on GPU
-myStreamsGpu = vclMatrix(myStreamsCpu)
-@
-<<createStreams_Gpu2>>=
-myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
-@
 
@@
 
-\subsection{Using streams to generate uniform random numbers}
-The streams created sequentially are then used by the GPU's work-items to generate random numbers.
+\subsection{Generating uniform random numbers}
+
+The streams created sequentially are used by work-items on a GPU (the GPU analogy of CPU cores) to generate random numbers.
 In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be at least the number of total work-items in use.
-The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
-Users can set \code{verbose=2} in \fct{runif} to print out the kernels.
-Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
-Here the pointers to \code{streams} and to the output matrix \code{out} on the global memory are passed to the kernel as arguments.
+The main part of the kernel (functions for execution on the GPU) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
+Kernels are written in the \proglang{C}-like \proglang{OpenCL} language, in which \code{__kernel} declares a function as a kernel, and the _\code{__global} prefix to the pointer kernel arguments specifies that they point to global memory space accessible to all work items.
+Users can set the argument \code{verbose=2} in the random number generator to print out the kernel, which is slightly more complex than the code in Listing \ref{lst:uniformkernel}.
+
+Here the pointers \code{streams} and \code{out} refer to the streams and the  output matrix respectively, which are stored in global memory.
 \code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively.
-\code{Npad} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of \code{streams} (these variables are defined in macros not shown here).
-\code{index} represents the position of a work-item when mapping it to a one-dimensional data structure.
-Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states, respectively, and \code{startvalue} tells the starting position of the particular stream to be copied to its corresponding work-item.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer named \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type.
+\code{Npad} is the "internal" number of columns, and \code{out} is an \code{Nrow} by \code{Ncol} submatrix of a larger \code{Nrow} by \code{Npad} matrix.
+
+Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states.
+The object \code{startvalue} gives the position of the work item's stream in the \code{streams} object and is computed with the
+\code{NpadStreams} object defined in a macro not shown here.
+
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer between 1 and 2147483647,  and then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl} (defined to be $1/2147483648$).
 If to generate random integers, \code{temp} is not scaled.
@@
 uint g1[3], g2[3];
-float temp, fact = mrg31k3p_NORM_cl;
+double temp, fact = mrg31k3p_NORM_cl;
+const int startvalue = (get_global_id(0) + get_global_size(0) * get_global_id(1)) * NpadStreams;
 
@@
 
-for(Drow=get_global_id(0); Drow < Nrow;
+for(Drow = get_global_id(0); Drow < Nrow;
     Drow += get_global_size(0)){
 
-    for(Dcol=get_global_id(1); DcolBlock < Ncol; 
+    for(Dcol = get_global_id(1); Dcol < Ncol; 
         Dcol += get_global_size(1)){

### 2022-01-01 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3080 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 R> library("clrng")
-R> myStreamsCpu <- createStreamsCpu(n = 4, initial = 12345)
+R> setBaseCreator(c(111, 222, 333, 444, 555, 666))
+R> myStreamsCpu <- createStreamsCpu(4)
 R> t(myStreamsCpu)
@@
 ##              [,1]       [,2]       [,3]       [,4]
-## current.g1.1 12345  336690377  502033783  739421137
-## current.g1.2 12345  597094797 1322587635 1475938232
-## current.g1.3 12345 1245771585 1964121530  730262207
-## current.g2.1 12345   85196284 1949818481 1630192198
-## current.g2.2 12345  523477687 1607232546  324551134
-## current.g2.3 12345 2094976052 1462898381  795289868
-## initial.g1.1 12345  336690377  502033783  739421137
-## initial.g1.2 12345  597094797 1322587635 1475938232
-## initial.g1.3 12345 1245771585 1964121530  730262207
-## initial.g2.1 12345   85196284 1949818481 1630192198
-## initial.g2.2 12345  523477687 1607232546  324551134
-## initial.g2.3 12345 2094976052 1462898381  795289868
+## current.g1.1  111  272931565 1097663848  628119774
+## current.g1.2  222  750526866  712444468  211319804
+## current.g1.3  333 1340941094  522805545  147140763
+## current.g2.1  444 1101709871  362320065 1665200176
+## current.g2.2  555 1485736734 1950619531  892386138
+## current.g2.3  666  375100068 1839101248  734766081
+## initial.g1.1  111  272931565 1097663848  628119774
+## initial.g1.2  222  750526866  712444468  211319804
+## initial.g1.3  333 1340941094  522805545  147140763
+## initial.g2.1  444 1101709871  362320065 1665200176
+## initial.g2.2  555 1485736734 1950619531  892386138
+## initial.g2.3  666  375100068 1839101248  734766081
 \end{CodeOutput} 
@@
 
-\fct{createStreamsCpu} has two arguments:
+\fct{setBaseCreator} has one argument:
+\begin{itemize}
+\itemsep0em 
+  \item \code{initial}: a vector of six 31-bit integers, recycled if length is shorter. Default is set to \code{c(12345, 12345, 12345, 12345, 12345, 12345)}.
+\end{itemize}
+\fct{createStreamsCpu} has one argument:
 \begin{itemize}
@@
   \item \code{n}: number of streams to create.
-  \item \code{initial}: a vector of six 31-bit integers, recycled if length is shorter. Default is set to \code{c(12345, 12345, 12345, 12345, 12345, 12345)}.
 \end{itemize}
 We move streams to the GPU by converting them to a `\code{vclMatrix}'.
-\fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-Users may choose to use either \fct{createStreamsCpu} or \fct{createStreamsGpu} and stick with that choice for their entire program, as we want to avoid the very small possibility of overlap between the streams with different initial seeds.
-The `\code{initial}' argument should be set only once at the very beginning of the program in each \proglang{R} session if users want to supply their own initial seed of the stream creator, or never, in which case users would go with the package default initial seed.
-After the `\code{initial}' argument is set, we can ignore the `\code{initial}' argument and call \fct{createStreamsCpu} or \fct{createStreamsGpu} once or several times to create various matrices of streams in the same program.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> myStreamsGpu = vclMatrix(myStreamsCpu)
+\end{CodeInput} 
+\end{CodeChunk}
+
+Equivalently, \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> myStreamsGpu2 = createStreamsGpu(4)
+\end{CodeInput} 
+\end{CodeChunk}
+
+%Users should run either \fct{createStreamsCpu} or \fct{createStreamsGpu} at the beginning of a task and retain these streams for all of the random number generation the task requires.
+Creating new streams with different initial states introduces the very small possibility of overlap between the original and new streams.
+If an \proglang{R} session is restarted during a task, perhaps because a preliminary analysis is done following an initial set of simulations and the full set of simulations continued at a later date, the state of the streams should be transferred to the CPU and saved to disk.
+The `\code{initial}' argument to the stream creator should be used with caution and we recommend using the default.
+It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
+
+There are a few important notes on the usage of these functions.
+(1), Users should run the function \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
+If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams are created using the package's default initial seed.
+(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to run and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program will create some of the streams twice.
+When you create a new stream, it should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
+Streams should be created only once, and moved or copied if needed.
 
@@
 Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # creating streams on GPU
-R> myStreamsGpu = vclMatrix(myStreamsCpu)
-\end{CodeInput} 
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> myStreamsGpu2 = createStreamsGpu(n = 4, initial = 12345)
-\end{CodeInput} 
-\end{CodeChunk}
 
@@
 
-\subsection{Using streams to generate uniform random numbers}
-The streams created sequentially are then used by the GPU's work-items to generate random numbers.
+\subsection{Generating uniform random numbers}
+
+The streams created sequentially are used by work-items on a GPU (the GPU analogy of CPU cores) to generate random numbers.
 In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be at least the number of total work-items in use.
-The main part of the kernel (\proglang{OpenCL} functions for execution on the device) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
-Users can set \code{verbose=2} in \fct{runif} to print out the kernels.
-Kernels are written in the \proglang{OpenCL C} language, in which \code{__kernel} declares a function as a kernel, pointer kernel arguments must be declared with an address space qualifier, for example \code{__global} and \code{__local}.
-Here the pointers to \code{streams} and to the output matrix \code{out} on the global memory are passed to the kernel as arguments.
+The main part of the kernel (functions for execution on the GPU) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
+Kernels are written in the \proglang{C}-like \proglang{OpenCL} language, in which \code{__kernel} declares a function as a kernel, and the \code{__global} prefix to the pointer kernel arguments specifies that they point to global memory space accessible to all work items.
+Users can set the argument \code{verbose=2} in the random number generator to print out the kernel, which is slightly more complex than the code in Listing \ref{lst:uniformkernel}.
+
+Here the pointers \code{streams} and \code{out} refer to the streams and the  output matrix respectively, which are stored in global memory.
 \code{Nrow} and \code{Ncol} represent row and column number of the matrix \code{out} respectively.
-\code{Npad} is the internal number of columns of \code{out}, \code{NpadStreams} is the internal number of columns of \code{streams} (these variables are defined in macros not shown here).
-\code{index} represents the position of a work-item when mapping it to a one-dimensional data structure.
-Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states, respectively, and \code{startvalue} tells the starting position of the particular stream to be copied to its corresponding work-item.
-The function \fct{clrngMrg31k3pNextState} generates an uniform random integer named \code{temp} between 1 and 2147483647,  the value of \code{temp} is then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl}, which is defined to be $1/2147483648$ depending on the precision type.
+\code{Npad} is the "internal" number of columns, and \code{out} is an \code{Nrow} by \code{Ncol} submatrix of a larger \code{Nrow} by \code{Npad} matrix.
+
+Each stream's current state is copied to the private memory of each work-item by the function \fct{streamsToPrivate}, in which \code{g1} and \code{g2} point to the first three and second three elements of stream states.
+The object \code{startvalue} gives the position of the work item's stream in the \code{streams} object and is computed with the
+\code{NpadStreams} object defined in a macro not shown here.
+
+The function \fct{clrngMrg31k3pNextState} generates an uniform random integer between 1 and 2147483647,  and then scaled to be in the interval $(0,1)$ by multiplying it by a constant \code{mrg31k3p\_NORM\_cl} (defined to be $1/2147483648$).
 If to generate random integers, \code{temp} is not scaled.
@@
 \begin{CodeOutput}
-## [1] 0.735 0.842 0.614 0.216 0.110 0.870
+## [1] 0.417 0.298 0.193 0.778 0.953 0.856
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.045   0.004   0.048
+##   0.048   0.004   0.052
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> r_matrix <- clrng::rexp(c(2, 4), rate = 1, Nglobal = c(2,
+R> streams <- createStreamsGpu(4)
+R> r_matrix <- clrng::rexp(c(2, 4), rate = 1, streams, Nglobal = c(2,
 +    2), type = "double")
@@
 ##       [,1]   [,2] [,3]    [,4]
-## [1,] 0.1187 0.187 1.079 0.112
-## [2,] 0.0501 0.369 0.411 1.880
+## [1,] 0.479 0.6886  1.2 0.62493
+## [2,] 0.721 0.0265  2.3 0.00598
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.316   0.001   0.318
+##   0.313   0.008   0.320
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 409435
+## [1] 410423
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.403
+## [1] 0.404
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##    1.95    0.01    1.95
+##   1.936   0.013   1.949
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 1351
+## [1] 1303
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.000135
+## [1] 0.00013
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> streamsGpu <- createStreamsGpu(n = 128 * 64)
+R> streamsGpu <- createStreamsGpu(128 * 64)
 \end{CodeInput} 

### 2022-01-01 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3080 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE}
@@
 Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by $Z$ steps to find its initial state, and the three states in the stream object are set to it.
-When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation becomes be more complicated because there is no guarantee that the streams (or similarly \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers is possible to have correlation.
-\pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
-A popular way of obtaining multiple streams is to take an RNG with a long period and cut the RNG's sequence into very long disjoint pieces of equal length $Z$, and use each piece as a separate stream.
-Creating a new stream amounts to computing its starting point.
-Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, although this is not currently implemented in \pkg{clrng}.
-In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
-Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by $Z$ steps to find its initial state, and the three states in the stream object are set to it.
 
@@
 library("clrng")
-myStreamsCpu <- createStreamsCpu(n=4, initial=rep(12345, 6))
+setBaseCreator(c(111,222,333,444,555,666))
+myStreamsCpu <- createStreamsCpu(4)
 t(myStreamsCpu)
 @
-\fct{createStreamsCpu} has two arguments:
+\fct{setBaseCreator} has one argument:
+\begin{itemize}
+\itemsep0em 
+  \item \code{initial}: a vector of six 31-bit integers, recycled if length is shorter. Default is set to \code{c(12345, 12345, 12345, 12345, 12345, 12345)}.
+\end{itemize}
+\fct{createStreamsCpu} has one argument:
 \begin{itemize}
@@
   \item \code{n}: number of streams to create.
-  \item \code{initial}: a vector of six 31-bit integers, recycled if length is shorter. Default is set to \code{c(12345, 12345, 12345, 12345, 12345, 12345)}.
 \end{itemize}
@@
 @
-Equivalently, 
- \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
+Equivalently, \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
 <<createStreams_Gpu2>>=
-myStreamsGpu2 = createStreamsGpu(n=4, initial=rep(12345, 6)))
+myStreamsGpu2 = createStreamsGpu(4)
 @
 
-Users should run either \fct{createStreamsCpu} or \fct{createStreamsGpu} at the beginning of a task and retain these streams for all of the random number generation the task requires.
+%Users should run either \fct{createStreamsCpu} or \fct{createStreamsGpu} at the beginning of a task and retain these streams for all of the random number generation the task requires.
 Creating new streams with different initial states introduces the very small possibility of overlap between the original and new streams.
@@
 It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
+
+There are a few important notes on the usage of these functions.
+(1), Users should run the function \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
+If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams are created using the package's default initial seed.
+(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to run and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program will create some of the streams twice.
+When you create a new stream, it should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
+Streams should be created only once, and moved or copied if needed.
 
@@
 The main part of the kernel (functions for execution on the GPU) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
-Kernels are written in the \proglang{C}-like \proglang{OpenCL} language, in which \code{__kernel} declares a function as a kernel, and the _\code{__global} prefix to the pointer kernel arguments specifies that they point to global memory space accessible to all work items.
+Kernels are written in the \proglang{C}-like \proglang{OpenCL} language, in which \code{__kernel} declares a function as a kernel, and the \code{__global} prefix to the pointer kernel arguments specifies that they point to global memory space accessible to all work items.
 Users can set the argument \code{verbose=2} in the random number generator to print out the kernel, which is slightly more complex than the code in Listing \ref{lst:uniformkernel}.
@@
 <<random exponential, eval=TRUE, cache=FALSE>>=
-r_matrix <- clrng::rexp(c(2,4), rate=1, Nglobal=c(2,2), type="double")
+streams <- createStreamsGpu(4)
+r_matrix <- clrng::rexp(c(2,4), rate=1, streams, Nglobal=c(2,2), type="double")
 as.matrix(r_matrix)
@@
 <<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
-streamsGpu <- createStreamsGpu(n=128*64)
+streamsGpu <- createStreamsGpu(128*64)
 @

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3081 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 (1), Users should run the function \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
-If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams are created using the package's default initial seed.
-(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to run and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program will create some of the streams twice.
-When you create a new stream, it should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
+If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams take the package's default initial seed.
+(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to run and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program results in creating some of the streams twice.
+A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
 Streams should be created only once, and moved or copied if needed.

### 2022-01-02 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3082 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{article}

\usepackage{arxiv}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography

\usepackage{float}
\usepackage{framed}
\usepackage{subcaption}
\usepackage{amsmath, amsthm,amssymb}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage[english]{babel}

\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\prog}[1]{\textsf{#1}}
\def\bigr{{\mathbb{R}}}
\newcommand{\pkg}[1]{\textbf{\emph{#1}}}
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\textit{#1()}}
\def\reml{{\text{reml}}}
\def\N{{\mathrm{N}}}
\def\MVN{{\mathrm{MVN}}}
\def\var{{\mathrm{Var}}}
\def\cov{{\mathrm{COV}}}
\def\Pr{{\mathbb{P}}}
\def\T{{\footnotesize {^{_{\sf T}}}}}
\def\E{\mathbb{E}} % expectation
\newcommand{\mathleft}{\@fleqntrue\@mathmargin0pt}

\newcommand\sectionprelude{%
  \vspace{1.5em}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem*{theorem*}{Theorem}

<<preliminaries, echo=FALSE, include=TRUE>>=
options(continue="+  ",  prompt="R> ", digits=3, width=66)

opts_chunk$set(fig.height=3,
		fig.width=4.5,dev='png',tidy=TRUE,fig.align='center',background='#FFFFFF00', prompt=TRUE, highlight=FALSE)

@

\title{A title}
\date{}

\author{
  Ruoyong Xu\\
  Department of Statistical Sciences\\
  University of Toronto\\
  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
  \texttt{ruoyong.xu@mail.utoronto.ca} \\
  %% examples of more authors
   \And
  Patrick Brown\\
  Department of Statistical Sciences\\
  University of Toronto\\
  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
  \texttt{patrick.brown@utoronto.ca} \\
}

\begin{document}
\maketitle
\setstretch{1.1}

\begin{abstract}
some keypoints to mention:\\
1, anisotropic spatial models are important but diffcult to evaluate because it consists of many covariance parameters.\\
2, standard approximations of spatial models (GMRF) don’t allow for anisotropy (at least as implemented). \\
3, gpuBatchMatrix is able to compute the profile LogL for each of the parameters in linear anisotropic spatial models, for dense matrix, use no approximation methods. allow for uncertainty in covariance parameters\\
4, investigated likelihood-based CI’s for variance parameters, wider than using information matrix based methods (gepstatsp)\\
\end{abstract}

% keywords can be removed
\keywords{First keyword \and Second keyword \and More}

\section{Introduction}
1- cI for the betas for the geostapsp package\\
2- Ci’s using profile likelihoods from gpuBatchMatrix
… second CI’s will be wider\\
… simulation study, profile CI’s have better coverage\\

emphasis for paper 2?
- anisotropic spatial models are important\\
- standard approximations of spatial models (GMRF) don’t allow for anisotropy (at least as implemented).  \\
- We do dense matrix, no approximations\\
- also, uncertainty in covariance parameters is important in anisotropic models because there are more of them.\\

- we should always allow for uncertainty in covariance parameters theta = sd, nugget, range, ratio, angle, shape, BoxCox\\
- … although most frequentist spatial software doesn’t, assumes theta = theta hat\\
- anisotropic models + boxcox + matern shape means many covariance parameters (7?)\\
- … important to understand uncertainty in covariance parameters, 2D profile L interesting\\
- … allow for uncertainty in covariance parameters to be reflected in CI’s for betas\\

- The easy way:  %L_easy(beta) = L(theta hat, beta), where theta hat = argmax_theta, beta L(theta, beta)， numerically optimize over nu, range, shape,….., then plug in and estimate beta sigma again）\\
- Hard way:\\  %L_p(beta) =  L( theta(beta), beta)\\
Likelihood-based CI’s for variance parameters\\

We can estimate shape, range, variance\\
… likelihood might be fairly flat, CI’s are wide\\
… but CI’s have good coverage

\section{The linear geostatistical model}
\label{section1}
We start with a review of the linear geostatistical model:
%
\begin{gather*}
Y_{i}|U(s_i) \overset{ind}\sim \N(\lambda(s_{i}),\tau^2 ),\\
\lambda(s_{i})=X(s_i)\beta+U(s_{i}),\\
(U(s_1), \dots,U(s_n))^\T \sim \MVN (0,\Sigma),\\
\Sigma_{ij}=\cov \left[U(s_i), U(s_j)\right]=\begin{cases} 
\sigma^2 R(\|s_i-s_j\|/\phi;\kappa),& i\neq j .\\
\sigma^2, & i=j.
\end{cases}
\end{gather*}
%
where
$Y_i: i=1,\dots,n$ is the observation obtained at location $s_i$. 
Given $U(s_i)$, $Y_i$'s are mutually independently distributed with Gaussian distribution with observation variance $\tau^2$.  
$X(s_i)$ is a $1 \times p$ vector of explanatory variables at location $s_i$, 
$\beta=(\beta_1,\dots,\beta_p)^\T$ is the corresponding vector of regression parameters. 
$U=(U(s_1), \dots,U(s_n))^\T $ is a zero-mean Gaussian random field with covariance matrix $\Sigma$. The covariance between locations $(s_i)$ and $(s_j)$ takes the above form, where $R(\cdot)$ is the correlation function, and $\sigma^2$ is the spatial variance or variability in residual variation.

Let $Y=(Y_1,\dots,Y_n)^\T$ be a vector of observations, and $X = (X(s_1), \dots, X(s_n))$ be an $n \times p$ design matrix, we rewrite the model in matrix format
%
\begin{equation*} 
Y \sim \MVN (X\beta, \sigma^2 R(\phi,\kappa)+\tau^2I).
\end{equation*}
% 
Write $V=R(\phi,\kappa)+\nu^2I$, and $\nu^2={\tau^2}/{\sigma^2}$, then
\begin{equation} \label{eq:1}
Y \sim \MVN (X\beta, \sigma^2 V).
\end{equation}

\subsection{Mat\'ern correlation}
The powered exponential family, the Mat\'ern family and the spherical family are the three most commonly used families of correlation functions in geostatistics. We favour the Mat\'ern family because of its flexibility, and it is made available in our \prog{R} package \pkg{gpuBatchMatrix}. The form of Mat\'ern correlation used in \pkg{gpuBatchMatrix} is 
\begin{equation*}
R(d;\phi,\kappa)=\frac{2^{\kappa-1}}{\Gamma(\kappa)} (\sqrt{8\kappa} \frac{d}{\phi})^\kappa  K_\kappa(\sqrt{8\kappa}  \frac{d}{\phi}),  \quad \text{where} \quad \phi>0,  \kappa > 0,
\end{equation*} 
$K_\kappa(\cdot)$ denotes the modified Bessel function of the second kind of order $\kappa$, 
$\kappa$ is the shape parameter which determines the smoothness of $U(x)$. For $\kappa=0.5$, the Mat\'ern correlation function coincides with the exponential correlation $\exp(-d/\phi)$. When $\kappa \rightarrow \infty$, it tends to the Gaussian correlation $\exp{\{-2(\|d\|/\phi)^2\}}$.
$\phi$ is called the range or scale parameter, it controls the rate of decay of the correlation as $d$ increases. 
%$d =\|x-x'\|$ is the Euclidean distance between two spatial points, $\Gamma(\cdot)$ is the standard gamma function,  $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as $d$ increases. $K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$, specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.  Estimation of $\kappa$ is difficult, our experience is to choose the values of $\kappa$ from a discrete set, for example $\{0.5,1.5,2.5,\dots \}$(ref), %and then maximize the log-likelihood function \eqref{profile} or \eqref{remlpro} over $\phi$ and $\nu^2$. When the number of location points is large, the computation for the Mat\'ern function ...... more stuff ....

\subsection{Box-Cox transformation}
% What are the Box-Cox power transformations? The inference on the transformations parameter?
The fit of the linear geostatistical model can often be improved by applying a transformation to the response variable $Y$. Skewed random fields that mildly deviate from the Gaussian distribution can be modeled by means of the Box-Cox transformation (BCT) \cite{box1964analysis}.
%\cite{cressie2015statistics} introduced the term trans-Gaussian model to refer to the linear geostatistical model with a Box-Cox transformed  response variable. 
For positive-valued response variable $Y$, the Box-Cox transformation has the form
\begin{equation*}
    Y'=
    \begin{cases}
      (Y^\lambda-1)/\lambda: & \lambda \in \bigr \quad \text{and} \quad \lambda \neq  0, \\
      \log Y: & \lambda =0,
    \end{cases}
\end{equation*}
where $Y'$ denotes the transformed response. The aim of the BCT is to make data closely resemble the Gaussian distribution, so that the usual assumptions for linear models hold. The BCT cannot guarantee that any probability distribution can be normalized. However, even in cases when an exact mapping to the normal distribution is not possible, BCT may still provide a useful approximation \cite{hristopulos2020random}.

%The transformation parameter can be estimated by using maximum likelihood estimation. The log-likelihood including $\lambda$ is
% \begin{equation*}
% -2\ell^*(\beta,\sigma^2,V,\lambda)= -2*(\lambda-1)\sum_{i=1}^{n}\log y_i + n\log(2\pi)+\log |\sigma^2 V|+(y'-X\beta)^\T(\sigma^2 V)^{-1}(y'-X\beta),
% \end{equation*}

\subsection{Geometric anisotroy}
Geometrical anisotropy is defined by two additional parameters in the correlation function. Let $(x_1, x_2)$ represent the original coordinates, which is rotated counterclockwise by $\varphi$, then the new coordinates $(x'_1, x'_2)$ is given by
%
  \begin{equation*}\label{eq:matrixeqn}
    \begin{pmatrix}
      x'_{1} \\
      x'_{2}  
    \end{pmatrix}
    =
    \begin{pmatrix}
      1/\phi_{1} & 0 \\
      0 & 1/ \phi_{2} 
    \end{pmatrix}
    \cdot
    \begin{pmatrix}
      \cos \varphi & -\sin \varphi \\
      \sin \varphi & \cos \varphi 
    \end{pmatrix}
    \cdot
      \begin{pmatrix}
      x_1 \\
      x_2 
    \end{pmatrix},
  \end{equation*}
%
where $\varphi \in [-\pi/2, \pi/2]$ is called the anisotropy angle, $\psi = \phi_1 / \phi_2 \in [0,1]$ is the anisotropy ratio. %Anisotropy in correlation functions Statistical anisotropy can also be present in the correlation functions, where it is more difficult to estimate accurately.

\section{Inference based on the likelihood function}
Let $y'^\T=(y'_1,\dots,y'_n)$ be the Box-Cox transformed data of the observations $y^\T=(y_1,\dots,y_n)$. The most general log-likelihood function for the linear geostatistical model parameters given $y'$ would be
%
\begin{equation}
-2\ell(\beta,\sigma^2, \lambda, \omega; y')=(y'-X\beta)^\T(\sigma^2 V)^{-1}(y'-X\beta)+\log |\sigma^2 V|-2*(\lambda-1)\sum_{i=1}^{n}\log y_i+n\log(2\pi).\label{eq:2}
\end{equation}
%
the third term on the right-hand side of \eqref{eq:2} arises from the Jacobian of the transformation. This log-likelihood function is complex to evaluate as we have several types of parameters to estimate, including the regression parameters $\beta$, the transformation parameter $\lambda$, and the covariance parameters $(\sigma^2, \phi,\kappa,\nu^2, \varphi, \psi)$. For simplicity we use $\omega=(\phi,\kappa,\nu^2, \varphi, \psi)$ to denote the correlation parameters in the following.

\subsection{Profile log-likelihood functions}
%ADD gpuBatchMatrix and MORE STUFF\\
%"a gradient-based optimization algorithm can be used to reduce the computational cost of the optimization\\
%Another way to reduce the computational effort involved in the maximisation of l is to reduce the dimensionality of the optimization problem as follows."\\
The profile log-likelihood for parameter $\varphi$ is defined as $\ell_p(\varphi;y) = \underset{\psi}{\text{sup}} \ell (\varphi, \psi; y ).$ Given $\omega$ and $\lambda$, The log-likelihood function \eqref{eq:2} is maximized at
%
\begin{gather}
\hat{\beta}_{\omega,\lambda}=(X^\T V^{-1}X)^{-1} X^\T V^{-1}y', \quad \text{and} \label{betahat}\\
\hat{{\sigma}}^2_{\omega,\lambda}= \frac{1}{n}(y'-X\hat{\beta}_{\omega,\lambda})^\T V^{-1}(y'-X\hat{\beta}_{\omega,\lambda}) \label{eq:4}.
\end{gather}
%
%For simplicity, $\hat{\beta}_{w,\lambda}$ is written as $\hat{\beta}$ in \eqref{eq:4}, it denotes the maximum likelihood estimate of $\beta$ for a given $(w, \lambda)$. 
Substitute \eqref{betahat} and \eqref{eq:4} back into \eqref{eq:2}, we have the profile log-likelihood for $\omega$ and $\lambda$
%
\begin{equation} \label{profile}
-2\ell_p(\omega, \lambda;y')=n \log\frac{(y'-X\hat{\beta}_{\omega,\lambda})^\T V^{-1}(y'-X\hat{\beta}_{\omega,\lambda})}{n}+\log|V| -2(\lambda-1)\sum_{i=1}^{n}\log y_i + n\log(2\pi) + n,
\end{equation}
%

%\subsection{Restricted maximum likelihood estimation}
The main problem of maximum likelihood estimation is that variance is underestimated. Think of the simple case when the covariance matrix of $Y$ is $\sigma^2 I$, then 
\begin{align*}
&\hat{\sigma}^2 = \frac{1}{n}(y-X\hat{\beta})^\T (y-X\hat{\beta}),\\
&\E[\hat{\sigma}^2]=\frac{n-p}{n} \sigma^2 < \sigma^2,
\end{align*}
where $p= \text{rank}(X)$, is the number of elements in $\beta$. The estimation bias comes from not taking into account the degree of freedom used for estimating $\beta$. Restricted maximum likelihood (REML) developed by \cite{patterson1971recovery} is an approach that produces unbiased estimators or less biased estimators than MLE in general. REML eliminates the influence of $X$ on $\hat{\sigma}^2$ and $\hat{\omega}$. In this method, we find a matrix $A$ of full rank such that $AX=0$. The $n \times n$ projection matrix $S= I-X(X^\T X)^{-1}X^\T$ satisfies $SX=0$, however, $S$ is degenerate as $\text{rank}(S)=n-p$, so we take just $n-p$ linearly independent rows from $S$ to make the matrix $A_{(n-p)\times n}$. Transform the data linearly to $Y^*=AY$, then $\E(Y^*)=AX\beta=0$. Recall in \eqref{eq:1} we obtained $Y \sim \MVN (X\beta, \sigma^2 V),$ thus
%
\begin{equation*}
Y^* \sim \MVN(0,\sigma^2 AVA^\T).
\end{equation*}
%
The principle of the REML method is to estimate the variance parameter $\sigma^2$ by maximizing the restricted log-likelihood function below, where $y^*$ represents a realization of $Y^*$,
%the distribution of $Y^*$ does not depend on $\beta$,
\begin{equation} \label{eq:6}
-2\ell^*(\sigma^2, \omega ;y^*)=y^*\T(\sigma^2 AVA^\T)^{-1}y^*+\log|\sigma^2 AVA^\T|+(n-p)\log(2\pi),
\end{equation}
%&=(n-p)\log(2\pi)+ (n-p)\log \sigma^2 + \log|AVA^\T|+ \frac{1}{\sigma^2}y^*\T(AVA^\T)^{-1}y^*.
Differentiating \eqref{eq:6} with regard to $\sigma^2$ and setting it equal to zero gives the unbiased estimate for $\sigma^2$
%
\begin{gather*} %\label{sigmahat_reml}
\hat{\sigma}_{reml}^2(\omega)= \frac{y^*\T (AVA^\T)^{-1} y^*}{n-p}.
\end{gather*}
%
The REML criterion is based on the likelihood of $Y^*$, in which $X$ does not appear.  REML bypasses estimating $\beta$ and can therefore, produce unbiased estimates for $\sigma^2$.

%\textit{reml} \\
We can write \eqref{eq:6} in terms of the original observed data $y$ using the two identities derived by \cite{searle1978notebook} in Section 5.1:
$$|A (\sigma^2V)A^\T|=|\sigma^2V||X^\T (\sigma^2V)^{-1}X|$$ and $$y^*(\sigma^2 AVA^\T)^{-1}y^*=(y-X\hat{\beta})^\T (\sigma^2 V)^{-1}(y-X\hat{\beta}),$$ where $\hat{\beta}=(X^\T V^{-1}X)^{-1} X^\T V^{-1}y$.  Then the restricted log-likelihood of $y$ is
\begin{align*}
-2\ell^*(\sigma^2, \omega;y)=(y-X\hat{\beta})^\T(\sigma^2V)^{-1}(y-X\hat{\beta})+(n-p)\log \sigma^2+\log |V|+\log|X^\T V^{-1}X|+n\log(2\pi).
\end{align*}
Applying the BCT on $y$ we have the restricted log-likelihood in terms of the transformed response data $y'$
\begin{align}
-2\ell^*(\sigma^2, \omega, \lambda;y')&=(y'-X\hat{\beta})^\T(\sigma^2V)^{-1}(y'-X\hat{\beta})+(n-p)\log \sigma^2+\log |V|+\log|X^\T V^{-1}X|+n\log(2\pi)-\notag\\
&2(\lambda-1)\sum_{i=1}^{n}\log y_i,\label{eq:7}
\end{align}
which does not depend on the choice of matrix $A$.

%\textit{reml profile} \\
Differentiate \eqref{eq:7} with respect to $\sigma^2$ and set it equal to zero, we have the expression for $\hat{\sigma}_{reml}^2$ in terms of the Box-Cox transformed data given $\omega$ and $\lambda$,
\begin{equation}\label{sigmahat_reml_y}
\hat{\sigma}_{reml}^2(\omega, \lambda)=\frac{(y'-X\hat{\beta})^\T  V^{-1}(y'-X\hat{\beta})}{n-p}.
\end{equation}
Substitute \eqref{sigmahat_reml_y} into \eqref{eq:7}, leading to the profile restricted log-likelihood for $(\omega, \lambda)$
\begin{equation}\label{remlpro}
-2\ell^*_p(\omega, \lambda;y')=(n-p)\log \hat{\sigma}^2+\log |V|+\log |X^\T V^{-1}X|-2(\lambda-1)\sum_{i=1}^{n}\log y_i+n\log(2\pi)+n-p.
\end{equation}

Compared to the ``ml'' profile log-likelihood given in \eqref{profile}, the effective dimension is reduced from $n$ of $\ell(\cdot)$  to $(n-p)$ of $\ell^*_p(\cdot)$ , this distinction is important if $p$ is large.

\subsection{Parameter estimation}
\textit{Geostatsp's methods}\vspace{0.5cm}\\
The following are two important theorems in likelihood-based inference, which lead to two ways of calculating likelihood-based confidence intervals (CIs). %We write θˆ(y) for an observed value of θˆ and θˆ(Y ) for the corresponding random variable. 
We write $I(\theta)$ for the expected Fisher's information matrix, and $\theta$ is the true parameter value. 
\begin{theorem}
\label{R1}
$\hat{\theta}(Y) \sim \MVN(\theta, I(\theta)^{-1})$.
\end{theorem}
%
\begin{theorem}
\label{R2}
$2\{\ell(\hat\theta(Y); Y)-\ell(\theta;Y) \} \sim \chi^2_p ,$ where $p$ is the dimension of $\theta$.
\end{theorem}
The \fct{lgm} function in \pkg{geostatsp} package estimates the correlation parameters $\omega=(\phi,\kappa,\nu^2, \varphi, \psi)$ and the transformation parameter $\lambda$ by maximization of the likelihood. Specifically, \fct{lgm} uses the \textit{optim} function to optimize the equation \eqref{profile} (if \fct{reml}=FALSE)  numerically over $\omega$ and $\lambda$ to find $(\hat{\omega},\hat{\lambda}) = \argmax_{\omega,\lambda} \ell_p(\omega, \lambda; y')$.  \pkg{geostatsp} treats $\hat{\omega}$ and $\hat{\lambda}$ as fixed and back substitute them into \eqref{betahat} and \eqref{eq:4} to obtain the maximum likelihood estimates (MLEs) $\hat{\beta}_{\omega,\lambda}$ and $\hat{{\sigma}}^2_{\omega,\lambda}$. The \fct{optim} function returns the Hessian evaluated at $\hat{\omega}$ and $\hat{\lambda}$.  Conventionally, \fct{lgm} calculates CIs based on Theorem \ref{R1}. The 95\% confidence intervals for each of the parameters are calculated by $\hat{\theta_i} \pm 1.96 \sqrt{\var(\hat{\theta_i})}$, where $\theta_i$ represents any of the parameter in $(\beta,\sigma, \omega,\lambda)$, and  $\var(\hat{\theta}_i)$ is approximated by the diagonals of the inverse of the negative Hessian. This way of lcnaelcac does not take into account of uncertainty in correlation parameters.

There is the other function \fct{profLlgm} in \pkg{geostatsp} that calculates confidence intervals or joint confidence regions based on the likelihood-ratio method \ref{R2}. If $c$ represents the $\beta$-quantile of $\chi^2_p$, i.e., $\Pr(\chi^2_p \leq c)=\beta$, then the set $\{\theta: \ell(\theta) \geq \ell(\hat{\theta}) - c/2 \}$ is asymptotically, a $\beta$-level confidence set for $\theta$. For a fixed $\phi$ value, \fct{optim} evaluates \eqref{profile} for different combinations of $(\kappa,\nu^2, \varphi, \psi)$ and finds the optim Log-likelihood value, repeat this process for a number of $\phi$ values to finally get the profile log-likelihood and MLE of $\phi$.  \fct{profLlgm} runs \fct{optim} a great many of times and thus is very slow.

%We can estimate $\lambda$ by maximization of the likelihood, since $\lambda$ is usually a nuisance parameter, an alternative and more pragmatic approach is to restrict $\lambda$ to a discrete set of values, for example $\{-1, 0, 1/2, 1\}$ \cite{laga2020model}. $\lambda = -1$ (reciprocal), $\lambda = 0$ (logarithm), $\lambda = 1/2$ (square-root) and $\lambda = 1$ (no transformation).

\textit{gpuBatchMatrix's method}\vspace{0.5cm}\\
The hard way is by computing the profile log-likelihood $\ell_p(\beta) =\ell(\beta, \hat{\omega}_{\beta},\hat{\sigma}_{\beta}, \hat{\lambda}_{\beta})$. (QUESTION: cannot geostatsp uses optim for beta? is it too slow so geostatsp does not do that?)

SHOW A PROFILE LOG-LIKELIHOOD PLOT FOR BETA'S. (MEUSE DATA OR SWISSRAIN DATA?)\\
SHOW ALL LIKELIHOODS PLOT VS BETA'S

"The results R1 and R2 are invoked routinely in modern statistical practice, but they rely on assumptions that are not always satisfied. In a geostatistical context, the two most common situations in which these assumptions do not hold are boundary problems and dimensional ambiguities. "

% Given $\omega$, $\lambda$ and $\sigma$, the log-likelihood \eqref{eq:2} would still be maximized at
% %
% \begin{gather*}
% \hat{\beta}_{V,\lambda, \sigma}=(X^\T V^{-1}X)^{-1} X^\T V^{-1}y,
% \end{gather*}
% substitute $\hat{\beta}_{V,\lambda, \sigma}$ into \eqref{eq:2}, we obtain The profile log-likelihood for $(V, \lambda,\sigma)$ as
%
% \begin{equation} \label{profileBeta}
% -2\ell_p(\sigma^2,\omega, \lambda;y)=(y-X\hat{\beta})^\T (\sigma^2 V)^{-1}(y-X\hat\beta)+\log|V|-2(\lambda-1)\sum_{i=1}^{n}\log y_i + n\log(2\pi)+ n \log\sigma^2.
% \end{equation}
%

%The easy way for estimating the regression parameter $\beta$ is through $\ell(\beta) = \ell(\beta, \hat{\sigma}^2, \hat{\omega}, \hat{\lambda})$, where $(\hat{\omega},\hat{\sigma}, \hat{\lambda}) = \argmax_{\omega,\lambda,\sigma} \ell(\beta,\omega,\sigma, \lambda)$. The \prog{R} package \pkg{geostatsp} uses the \textit{optim} function to optimize the equation \eqref{profile} numerically over $\omega$ and $\lambda$ to find

% 
% 3, Given $V$, $\lambda$ and $\beta$, the log-likelihood \eqref{eq:2} is maximized at 
%  %
%  \begin{gather}
%  \hat{\sigma}^2_{V,\lambda, \beta}= \frac{1}{n}(y-X \beta)^\T V^{-1}(y-X\beta) \label{sigmahat},
%  \end{gather}
%  %
%  substitute $\hat{\sigma}^2_{V,\lambda, \beta}$ into \eqref{eq:2}, we have the following profile log-likelihood function for $(V,\lambda, \beta)$,
%  %
%  \begin{equation} \label{profileSigma}
%  -2\ell_p(\beta,\phi,\kappa,\nu^2, \varphi, \psi, \lambda;y)=n \log\frac{(y-X \beta)^\T V^{-1}(y-X\beta)}{n}+\log|V|+\text{Jacobian}(\lambda)+n\log(2\pi)+n.
%  \end{equation}

\subsection{CI’s comparison for model parameters}
%TWO WAYS OF CALCULATING CI IN LIKELIHOOD METHODS CI USING INFORMATION MATRIX: BOUNDRY PROBLEMS AND dimensional ambiguities.\\
%The meaning of a confidence interval is a random set of values, constructed in such a way that, over repeated sampling of the data from the underlying model, that will contain the true, fixed but unknown value of the parameter with some specified probability. \cite{pd2007}
%

%
%Conventionally, \pkg{geostatsp} uses the function \fct{lgm} to return CIs for parameters, which is based on the \ref{R1} method. Confidence intervals for individual parameter $\varphi_i$ would be $\hat{\varphi_i} \pm 1.96*\sqrt{\var(\hat{\varphi_i})}$, where $\var(\hat{\varphi}_i)$ can be approximated by the diagonals of inverted observed information matrix $\frac{\partial^2 l(\varphi)}{\partial\varphi \partial\varphi^\top}\big |_{\varphi=\hat{\varphi}_i}$.

SHOW A TABLE OF CI compare between \pkg{geostatsp} and \pkg{gpuBatchMatrix}
%"Calculating a likelihood-based confidence region is straightforward in principle but difficult in practice, rapidly becoming impossible as the number of elements of $\varphi$ grows. However, a more general version of the result also holds, and is extremely useful in practice." The results R1 and R2 are invoked routinely in modern statistical practice, but they rely on assumptions that are not always satisfied. In a geostatistical context, the two most common situations in which these assumptions do not hold are boundary problems and dimensional ambiguities.

<<packages, results='hide', message=FALSE>>=
library(gpuRandom)
library(gpuR)
library(data.table)
library('geostatsp')
@

% swissrain matplot over betas
<<params, echo=FALSE, results='hide',eval=TRUE>>=
data('swissRain')
swissRain$elevation = extract(swissAltitude, swissRain)

params = rbind(
  c(shape=1.826358, range=20000, variance = 1, nugget = 0, anisoRatio = 10, anisoAngleRadians = 0.08),  
  c(shape=4, range=3.861873e+04, variance = 1, nugget = 15, anisoRatio = 16, anisoAngleRadians = 0.7),  
  c(shape=3, range=90*1000, variance = 1, nugget = 1.369707e-01, anisoRatio = 25, anisoAngleRadians = 0.3),
  c(shape=2, range=250*1000, variance = 1, nugget = 20, anisoRatio = 8.09, anisoAngleRadians = 0.5), 
  c(shape=0.2, range=15*1000, variance = 1, nugget = 0.02, anisoRatio = 65, anisoAngleRadians = 6.5),
  c(shape=0.3, range=20*1000, variance = 1, nugget = 0.08, anisoRatio = 45, anisoAngleRadians = 12.5)
)
colnames(params) <- c('shape','range', 'variance', 'nugget', 'anisoRatio', 'anisoAngleRadians')
@
<<showparams, eval=TRUE>>=
params
@

<<compute, echo=FALSE, results='hide', eval=TRUE>>=
result5<-gpuRandom::likfitLgmCov2d(
  data=swissRain,
  formula=rain~ elevation, 
  coordinates=swissRain@coords,
  params=params, 
  boxcox = c(1,0,4.962398e-01),
  cilevel=0.9,
  type = "double",
  reml=FALSE, 
  NparamPerIter=400,
  Nglobal=c(128,128),
  Nlocal=c(16,16),
  NlocalCache=2800,
  verbose=FALSE)

Betas <- seq(-12, 12, len=40)*1e-03

m <- length(Betas)
makeSymm <- function(m) {
  m[upper.tri(m)] <- t(m)[upper.tri(m)]
  return(m)
}
a=2 # which beta_i?
Nobs=result5$Nobs  # number of observations.
Ndata=result5$Ndata
Nparam=result5$Nparam
Ncov=result5$Ncov
XVYXVX=result5$XVYXVX
ssqY <- as.matrix(result5$ssqY)
detVar <- as.vector(result5$detVar)
detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
jacobian <- as.vector(result5$jacobian)
jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
Ucov <- Ncov-1
XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
XVY <- XVYXVX[ , 1:Ndata]

for (i in 1:Nparam){
  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
}

selectedrows <- (seq_len(Nparam)-1) * Ncov + a
XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
XVY_deleted <- XVY[-selectedrows, ]
X_aVY <- XVY[selectedrows, ]
X_aVX_a <- XTVinvX[selectedrows, a]

partA = matrix(0, nrow=Nparam, ncol=Ndata)
partB = matrix(0, nrow=Nparam, ncol=Ndata)
partC = matrix(0, nrow=Nparam, ncol=Ndata)
partD = matrix(0, nrow=Nparam, ncol=Ndata)
partE = matrix(0, nrow=Nparam, ncol=Ndata)

for (i in 1:Nparam){
  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
  temp <- solve(XTVinvX_deleted[interval, ])

  # part (A) have 2 data sets
  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
  # part (B) have 2 data sets.   has beta
  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
  # part (C) no data sets.  has beta
  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
  # part (D) have 2 data sets.    has beta
  partD[i,] = -2* X_aVY[i, ]
  # part (E)
  partE[i,] = X_aVX_a[i]
}

All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
for (bet in 1:m){
  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
}
colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5','6')
@

<<matplot_beta, eval=TRUE,echo=FALSE, fig.height =4.5, fig.width=6, cache=TRUE, fig.cap='profile log-Likelihoods of slope parameter for six covariance parameter sets',fig.pos='H'>>=
matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)
@
% swissrain matplot over betas

%meuse data
<<compare_mle, eval=TRUE,echo=FALSE>>=
rm(list=ls(all=TRUE))
library("sp")
data(meuse)
class(meuse)
coordinates(meuse) <- c("x", "y")
class(meuse)
library('geostatsp')
meuse$Y = log(meuse$zinc)
meuse$sqrtdist = sqrt(meuse$dist)

Res1 =  geostatsp::lgm( formula=Y~sqrtdist, 
                        data=meuse, 
                        grid=20,
                        reml = FALSE,
                        fixBoxcox=FALSE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
                        aniso=TRUE )
Res1$summary[,c('estimate','ci0.005', 'ci0.995')]
Res1$optim$mle
@

\pkg{gpuBatchMatrix} give wider ci and more accurate MLE's, expecially for the shape parameter.
<<compare_gpurandom, cache=FALSE, eval=FALSE,echo=FALSE>>=
library('gpuRandom')
library('gpuR')
  ParamList = list(
    range = c(exp(seq(log(100), log(173.1643928), len=6)), exp(seq(log(185), log(330), len=8))),
    shape = c(seq(0.1,3.5, len=4), seq(4, 6, len=4)),
    anisoRatio =c( seq(1, 2.301060332, len=3), seq(2.8, 7, len=4)),
    anisoAngleRadians = c( seq(0, 0.274896758, len=4), seq(0.4, 0.8, len=4)),
    nugget = c(seq(0,0.3946587, len=3), seq(0.5, 3, len=5))
  ) 
  params = do.call(expand.grid, ParamList)

result3<- gpuRandom::likfitLgmCov(spatialmodel=Res1,
                           params=params, # CPU matrix for now, users need to provide proper parameters given their specific need
                           boxcox=c(seq(-3, -0.903901505, len=4), seq(-0.5,1,len=4)),  
                           paramToEstimate = c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'), 
                           cilevel=0.95, 
                           type = "double",
                           reml=FALSE, 
                           NparamPerIter=400,
                           Nglobal=c(128,64),
                           Nlocal=c(16,16),
                           NlocalCache=2800,
                           verbose=FALSE)

cbind(Res1$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),c('estimate','ci0.005', 'ci0.995')], result3$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),])
@

<<matplotmeuse, echo=FALSE, eval=TRUE>>=
result1<-gpuRandom::getProfLogL(data=meuse,
                                formula=Y~sqrtdist, 
                                coordinates=meuse@coords,
                                params=params, 
                                boxcox = -0.903901505,
                                type = "double",
                                NparamPerIter=400,
                                gpuElementsOnly=FALSE,
                                reml=FALSE, 
                                Nglobal=c(128,64),
                                Nlocal=c(16,16),
                                NlocalCache=2800)

LogLik = result1$LogLik[seq(1,50176,400),3]
length(LogLik) 
Betas <- seq(-12, 12, len=40)*1e-03
m <- length(Betas)
makeSymm <- function(m) {
  m[upper.tri(m)] <- t(m)[upper.tri(m)]
  return(m)
}
a=2 # which beta_i?
Nobs=result1$Nobs  # number of observations.
Ndata=result1$Ndata
Nparam=result1$Nparam[seq(1,50176,400),]
Ncov=result1$Ncov
XVYXVX=result1$XVYXVX
ssqY <- as.matrix(result1$ssqY)
detVar <- as.vector(result1$detVar)
detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
jacobian <- as.vector(result1$jacobian)
jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
Ucov <- Ncov-1
XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
XVY <- XVYXVX[ , 1:Ndata]

for (i in 1:Nparam){
  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
}

selectedrows <- (seq_len(Nparam)-1) * Ncov + a
XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
#XTVinvX_a <- matrix(XTVinvX[-selectedrows, a], nrow=Nparam*Ucov, ncol=1)[1:10,]
XVY_deleted <- XVY[-selectedrows, ]
X_aVY <- XVY[selectedrows, ]
X_aVX_a <- XTVinvX[selectedrows, a]

partA = matrix(0, nrow=Nparam, ncol=Ndata)
partB = matrix(0, nrow=Nparam, ncol=Ndata)
partC = matrix(0, nrow=Nparam, ncol=Ndata)
partD = matrix(0, nrow=Nparam, ncol=Ndata)
partE = matrix(0, nrow=Nparam, ncol=Ndata)

for (i in 1:Nparam){
  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
  temp <- solve(XTVinvX_deleted[interval, ])

  # part (A) have 2 data sets
  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
  # part (B) have 2 data sets.   has beta
  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
  # part (C) no data sets.  has beta
  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
  # part (D) have 2 data sets.    has beta
  partD[i,] = -2* X_aVY[i, ]
  # part (E)
  partE[i,] = X_aVX_a[i]
}

All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
for (bet in 1:m){
  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
}
colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5', '6')
matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)

@
%meuse data

\section{A simulation study}

\section{Discussion}

\bibliographystyle{unsrt}  
\bibliography{paper2}

\end{document}

### 2022-01-02 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.tex
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3082 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2022-01-02 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\makeatletter
+\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}}}
+\makeatother
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\makeatletter
+\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+}}
+\makeatother
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 
-<<preliminaries, echo=FALSE, include=TRUE>>=
-options(continue="+  ",  prompt="R> ", digits=3, width=66)
-
-opts_chunk$set(fig.height=3,
-		fig.width=4.5,dev='png',tidy=TRUE,fig.align='center',background='#FFFFFF00', prompt=TRUE, highlight=FALSE)
-
-@
-
 \title{A title}
@@
 }
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-<<packages, results='hide', message=FALSE>>=
-library(gpuRandom)
-library(gpuR)
-library(data.table)
-library('geostatsp')
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+R> library(gpuRandom)
+R> library(gpuR)
+R> library(data.table)
+R> library("geostatsp")
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 % swissrain matplot over betas
-<<params, echo=FALSE, results='hide',eval=TRUE>>=
-data('swissRain')
-swissRain$elevation = extract(swissAltitude, swissRain)
 
-params = rbind(
-  c(shape=1.826358, range=20000, variance = 1, nugget = 0, anisoRatio = 10, anisoAngleRadians = 0.08),  
-  c(shape=4, range=3.861873e+04, variance = 1, nugget = 15, anisoRatio = 16, anisoAngleRadians = 0.7),  
-  c(shape=3, range=90*1000, variance = 1, nugget = 1.369707e-01, anisoRatio = 25, anisoAngleRadians = 0.3),
-  c(shape=2, range=250*1000, variance = 1, nugget = 20, anisoRatio = 8.09, anisoAngleRadians = 0.5), 
-  c(shape=0.2, range=15*1000, variance = 1, nugget = 0.02, anisoRatio = 65, anisoAngleRadians = 6.5),
-  c(shape=0.3, range=20*1000, variance = 1, nugget = 0.08, anisoRatio = 45, anisoAngleRadians = 12.5)
-)
-colnames(params) <- c('shape','range', 'variance', 'nugget', 'anisoRatio', 'anisoAngleRadians')
-@
-<<showparams, eval=TRUE>>=
-params
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+R> params
+##      shape  range variance nugget anisoRatio anisoAngleRadians
+## [1,]  1.83  20000        1  0.000      10.00              0.08
+## [2,]  4.00  38619        1 15.000      16.00              0.70
+## [3,]  3.00  90000        1  0.137      25.00              0.30
+## [4,]  2.00 250000        1 20.000       8.09              0.50
+## [5,]  0.20  15000        1  0.020      65.00              6.50
+## [6,]  0.30  20000        1  0.080      45.00             12.50
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
-<<compute, echo=FALSE, results='hide', eval=TRUE>>=
-result5<-gpuRandom::likfitLgmCov2d(
-  data=swissRain,
-  formula=rain~ elevation, 
-  coordinates=swissRain@coords,
-  params=params, 
-  boxcox = c(1,0,4.962398e-01),
-  cilevel=0.9,
-  type = "double",
-  reml=FALSE, 
-  NparamPerIter=400,
-  Nglobal=c(128,128),
-  Nlocal=c(16,16),
-  NlocalCache=2800,
-  verbose=FALSE)
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{figure}[H]
 
-Betas <- seq(-12, 12, len=40)*1e-03
+{\centering \includegraphics[width=\maxwidth]{figure/matplot_beta-1}
 
-m <- length(Betas)
-makeSymm <- function(m) {
-  m[upper.tri(m)] <- t(m)[upper.tri(m)]
-  return(m)
-}
-a=2 # which beta_i?
-Nobs=result5$Nobs  # number of observations.
-Ndata=result5$Ndata
-Nparam=result5$Nparam
-Ncov=result5$Ncov
-XVYXVX=result5$XVYXVX
-ssqY <- as.matrix(result5$ssqY)
-detVar <- as.vector(result5$detVar)
-detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
-jacobian <- as.vector(result5$jacobian)
-jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
-Ucov <- Ncov-1
-XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
-XVY <- XVYXVX[ , 1:Ndata]
-
-for (i in 1:Nparam){
-  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
 }
 
-selectedrows <- (seq_len(Nparam)-1) * Ncov + a
-XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
-XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
-XVY_deleted <- XVY[-selectedrows, ]
-X_aVY <- XVY[selectedrows, ]
-X_aVX_a <- XTVinvX[selectedrows, a]
+\caption[profile log-Likelihoods of slope parameter for six covariance parameter sets]{profile log-Likelihoods of slope parameter for six covariance parameter sets}\label{fig:matplot_beta}
+\end{figure}
 
-partA = matrix(0, nrow=Nparam, ncol=Ndata)
-partB = matrix(0, nrow=Nparam, ncol=Ndata)
-partC = matrix(0, nrow=Nparam, ncol=Ndata)
-partD = matrix(0, nrow=Nparam, ncol=Ndata)
-partE = matrix(0, nrow=Nparam, ncol=Ndata)
-
-for (i in 1:Nparam){
-  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
-  temp <- solve(XTVinvX_deleted[interval, ])
-
-  # part (A) have 2 data sets
-  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
-  # part (B) have 2 data sets.   has beta
-  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
-  # part (C) no data sets.  has beta
-  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
-  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
-  # part (D) have 2 data sets.    has beta
-  partD[i,] = -2* X_aVY[i, ]
-  # part (E)
-  partE[i,] = X_aVX_a[i]
-}
-
-All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
-for (bet in 1:m){
-  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
-    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
-}
-colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5','6')
-@
-
-<<matplot_beta, eval=TRUE,echo=FALSE, fig.height =4.5, fig.width=6, cache=TRUE, fig.cap='profile log-Likelihoods of slope parameter for six covariance parameter sets',fig.pos='H'>>=
-matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
-legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)
-@
+\end{knitrout}
 % swissrain matplot over betas
@@
 %meuse data
-<<compare_mle, eval=TRUE,echo=FALSE>>=
-rm(list=ls(all=TRUE))
-library("sp")
-data(meuse)
-class(meuse)
-coordinates(meuse) <- c("x", "y")
-class(meuse)
-library('geostatsp')
-meuse$Y = log(meuse$zinc)
-meuse$sqrtdist = sqrt(meuse$dist)
-
-Res1 =  geostatsp::lgm( formula=Y~sqrtdist, 
-                        data=meuse, 
-                        grid=20,
-                        reml = FALSE,
-                        fixBoxcox=FALSE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
-                        aniso=TRUE )
-Res1$summary[,c('estimate','ci0.005', 'ci0.995')]
-Res1$optim$mle
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+## [1] "data.frame"
+## [1] "SpatialPointsDataFrame"
+## attr(,"package")
+## [1] "sp"
+##                    estimate   ci0.005 ci0.995
+## (Intercept)         0.91634  0.121205   1.711
+## sqrtdist           -0.08445 -1.623508   1.455
+## sdNugget            0.00785  0.000379   0.163
+## sdSpatial           0.01250  0.000690   0.227
+## range             173.16439 92.775636 323.209
+## shape               4.00000 -9.880685  17.881
+## anisoRatio          2.30106  1.138810   4.649
+## anisoAngleRadians   0.27490 -0.167582   0.717
+## anisoAngleDegrees  15.75042 -9.601751  41.103
+## boxcox             -0.90390 -2.558961   0.751
+##            nugget          variance             range 
+##             0.395             1.000           173.164 
+##             shape        anisoRatio anisoAngleRadians 
+##             4.000             2.301             0.275 
+##            boxcox 
+##            -0.904
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 \pkg{gpuBatchMatrix} give wider ci and more accurate MLE's, expecially for the shape parameter.
-<<compare_gpurandom, cache=FALSE, eval=FALSE,echo=FALSE>>=
-library('gpuRandom')
-library('gpuR')
-  ParamList = list(
-    range = c(exp(seq(log(100), log(173.1643928), len=6)), exp(seq(log(185), log(330), len=8))),
-    shape = c(seq(0.1,3.5, len=4), seq(4, 6, len=4)),
-    anisoRatio =c( seq(1, 2.301060332, len=3), seq(2.8, 7, len=4)),
-    anisoAngleRadians = c( seq(0, 0.274896758, len=4), seq(0.4, 0.8, len=4)),
-    nugget = c(seq(0,0.3946587, len=3), seq(0.5, 3, len=5))
-  ) 
-  params = do.call(expand.grid, ParamList)
 
-result3<- gpuRandom::likfitLgmCov(spatialmodel=Res1,
-                           params=params, # CPU matrix for now, users need to provide proper parameters given their specific need
-                           boxcox=c(seq(-3, -0.903901505, len=4), seq(-0.5,1,len=4)),  
-                           paramToEstimate = c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'), 
-                           cilevel=0.95, 
-                           type = "double",
-                           reml=FALSE, 
-                           NparamPerIter=400,
-                           Nglobal=c(128,64),
-                           Nlocal=c(16,16),
-                           NlocalCache=2800,
-                           verbose=FALSE)
-
-cbind(Res1$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),c('estimate','ci0.005', 'ci0.995')], result3$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),])
-@
-
-<<matplotmeuse, echo=FALSE, eval=TRUE>>=
-result1<-gpuRandom::getProfLogL(data=meuse,
-                                formula=Y~sqrtdist, 
-                                coordinates=meuse@coords,
-                                params=params, 
-                                boxcox = -0.903901505,
-                                type = "double",
-                                NparamPerIter=400,
-                                gpuElementsOnly=FALSE,
-                                reml=FALSE, 
-                                Nglobal=c(128,64),
-                                Nlocal=c(16,16),
-                                NlocalCache=2800)
-
-LogLik = result1$LogLik[seq(1,50176,400),3]
-length(LogLik) 
-Betas <- seq(-12, 12, len=40)*1e-03
-m <- length(Betas)
-makeSymm <- function(m) {
-  m[upper.tri(m)] <- t(m)[upper.tri(m)]
-  return(m)
-}
-a=2 # which beta_i?
-Nobs=result1$Nobs  # number of observations.
-Ndata=result1$Ndata
-Nparam=result1$Nparam[seq(1,50176,400),]
-Ncov=result1$Ncov
-XVYXVX=result1$XVYXVX
-ssqY <- as.matrix(result1$ssqY)
-detVar <- as.vector(result1$detVar)
-detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
-jacobian <- as.vector(result1$jacobian)
-jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
-Ucov <- Ncov-1
-XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
-XVY <- XVYXVX[ , 1:Ndata]
-
-for (i in 1:Nparam){
-  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
-}
-
-selectedrows <- (seq_len(Nparam)-1) * Ncov + a
-XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
-XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
-#XTVinvX_a <- matrix(XTVinvX[-selectedrows, a], nrow=Nparam*Ucov, ncol=1)[1:10,]
-XVY_deleted <- XVY[-selectedrows, ]
-X_aVY <- XVY[selectedrows, ]
-X_aVX_a <- XTVinvX[selectedrows, a]
-
-partA = matrix(0, nrow=Nparam, ncol=Ndata)
-partB = matrix(0, nrow=Nparam, ncol=Ndata)
-partC = matrix(0, nrow=Nparam, ncol=Ndata)
-partD = matrix(0, nrow=Nparam, ncol=Ndata)
-partE = matrix(0, nrow=Nparam, ncol=Ndata)
-
-for (i in 1:Nparam){
-  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
-  temp <- solve(XTVinvX_deleted[interval, ])
-
-  # part (A) have 2 data sets
-  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
-  # part (B) have 2 data sets.   has beta
-  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
-  # part (C) no data sets.  has beta
-  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
-  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
-  # part (D) have 2 data sets.    has beta
-  partD[i,] = -2* X_aVY[i, ]
-  # part (E)
-  partE[i,] = X_aVX_a[i]
-}
-
-All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
-for (bet in 1:m){
-  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
-    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
-}
-colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5', '6')
-matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
-legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)
-
-@
 %meuse data

### 2022-01-02 pkg/oldgpuRandom/inst/doc/L2021_11/tempfile.tex
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3082 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2022-01-02 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.tex
@@
 \definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\makeatletter
-\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}}}
-\makeatother
 \newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
@@
 \definecolor{errorcolor}{rgb}{1, 0, 0}
-\makeatletter
-\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-}}
-\makeatother
 \newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
@@
 \usepackage{alltt}
-
-\usepackage{arxiv}
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-%\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-%\usepackage{microtype}      % microtypography
 
@@
 \usepackage{enumitem}
-\usepackage{setspace}
-\usepackage[english]{babel}
 
-\DeclareMathOperator*{\argmax}{argmax}
-\newcommand{\prog}[1]{\textsf{#1}}
-\def\bigr{{\mathbb{R}}}
-\newcommand{\pkg}[1]{\textbf{\emph{#1}}}
+\def\bigr{{\mathcal{R}}}
 \newcommand{\class}[1]{`\code{#1}'}
-\newcommand{\fct}[1]{\textit{#1()}}
+\newcommand{\fct}[1]{\code{#1()}}
 \def\reml{{\text{reml}}}
-\def\N{{\mathrm{N}}}
-\def\MVN{{\mathrm{MVN}}}
-\def\var{{\mathrm{Var}}}
-\def\cov{{\mathrm{COV}}}
-\def\Pr{{\mathbb{P}}}
+\def\N{{\text{N}}}
+\def\MVN{{\text{MVN}}}
+\def\cov{{\text{COV}}}
 \def\T{{\footnotesize {^{_{\sf T}}}}}
@@
 
-\newcommand\sectionprelude{%
-  \vspace{1.5em}
-}
+\author{Ruoyong Xu}
 
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem*{theorem*}{Theorem}
-
-\title{A title}
-\date{}
-
-\author{
-  Ruoyong Xu\\
-  Department of Statistical Sciences\\
-  University of Toronto\\
-  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
-  \texttt{ruoyong.xu@mail.utoronto.ca} \\
-  %% examples of more authors
-   \And
-  Patrick Brown\\
-  Department of Statistical Sciences\\
-  University of Toronto\\
-  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
-  \texttt{patrick.brown@utoronto.ca} \\
-}
+\date{preliminary draft}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
+
+\title{a title}
+
 \maketitle
-\setstretch{1.1}
+\begin{abstract}
+abstract
 
-\begin{abstract}
-some keypoints to mention:\\
-1, anisotropic spatial models are important but diffcult to evaluate because it consists of many covariance parameters.\\
-2, standard approximations of spatial models (GMRF) don’t allow for anisotropy (at least as implemented). \\
-3, gpuBatchMatrix is able to compute the profile LogL for each of the parameters in linear anisotropic spatial models, for dense matrix, use no approximation methods.
-allow for uncertainty in covariance parameters\\
-4, investigated likelihood-based CI’s for variance parameters, wider than using information matrix based methods (gepstatsp)\\
 \end{abstract}
 
-% keywords can be removed
-\keywords{First keyword \and Second keyword \and More}
+\bibliographystyle{plain}
 
 \section{Introduction} 
-1- cI for the betas for the geostapsp package\\
-2- Ci’s using profile likelihoods from gpuBatchMatrix
-… second CI’s will be wider\\
-… simulation study, profile CI’s have better coverage\\
-
-emphasis for paper 2?
-- anisotropic spatial models are important\\
-- standard approximations of spatial models (GMRF) don’t allow for anisotropy (at least as implemented).  \\
-- We do dense matrix, no approximations\\
-- also, uncertainty in covariance parameters is important in anisotropic models because there are more of them.\\
-
-- we should always allow for uncertainty in covariance parameters theta = sd, nugget, range, ratio, angle, shape, BoxCox\\
-- … although most frequentist spatial software doesn’t, assumes theta = theta hat\\
-- anisotropic models + boxcox + matern shape means many covariance parameters (7?)\\
-- … important to understand uncertainty in covariance parameters, 2D profile L interesting\\
-- … allow for uncertainty in covariance parameters to be reflected in CI’s for betas\\
-
-- The easy way:  %L_easy(beta) = L(theta hat, beta), where theta hat = argmax_theta, beta L(theta, beta)， numerically optimize over nu, range, shape,….., then plug in and estimate beta sigma again）\\
-- Hard way:\\  %L_p(beta) =  L( theta(beta), beta)\\
-Likelihood-based CI’s for variance parameters\\
-
-We can estimate shape, range, variance\\
-… likelihood might be fairly flat, CI’s are wide\\
-… but CI’s have good coverage
+introduction
 
 \section{The linear geostatistical model}
-\label{section1}
-We start with a review of the linear geostatistical model:
+We start with the linear geostatistical model:
 %
@@
 %
-where
+The basic ingredients are the following:
 $Y_i: i=1,\dots,n$ is the observation obtained at location $s_i$. 
-Given $U(s_i)$, $Y_i$'s are mutually independently distributed with Gaussian distribution with observation variance $\tau^2$.  
+Given $U(s_i)$, $Y_i$'s are mutually independently distributed with Gaussian distribution with observation variance or nugget effect $\tau^2$.  
 $X(s_i)$ is a $1 \times p$ vector of explanatory variables at location $s_i$, 
 $\beta=(\beta_1,\dots,\beta_p)^\T$ is the corresponding vector of regression parameters. 
-$U=(U(s_1), \dots,U(s_n))^\T $ is a zero-mean Gaussian random field with covariance matrix $\Sigma$.
-The covariance between locations $(s_i)$ and $(s_j)$ takes the above form, where $R(\cdot)$ is the correlation function, and $\sigma^2$ is the spatial variance or variability in residual variation.
+$U=(U(s_1), \dots,U(s_M))^\T $ is an (isotropic) zero-mean Gaussian random field with covariance matrix $\Sigma$ and correlation function $R(\cdot)$. 
+$\sigma^2$ is the spatial variance or variability in residual variation.
 
-Let $Y=(Y_1,\dots,Y_n)^\T$ be a vector of observations, and $X = (X(s_1), \dots, X(s_n))$ be an $n \times p$ design matrix, we rewrite the model in matrix format
+Let $Y=(Y_1,\dots,Y_n)^\T$ be the vector of observations, and $X$ be an $n \times p$ design matrix, we rewrite the model in matrix format:
 %
+\begin{equation} \label{eq:1}
+Y \sim \MVN (X\beta, \sigma^2 R(\phi,\kappa)+\tau^2I), 
+\end{equation}
+% 
+write $V=R(\phi,\kappa)+\nu^2I$, and $\nu^2={\tau^2}/{\sigma^2}$, then
 \begin{equation*}
-Y \sim \MVN (X\beta, \sigma^2 R(\phi,\kappa)+\tau^2I).
+Y \sim \MVN (X\beta, \sigma^2 V).
 \end{equation*}
-% 
-Write $V=R(\phi,\kappa)+\nu^2I$, and $\nu^2={\tau^2}/{\sigma^2}$, then
-\begin{equation} \label{eq:1}
-Y \sim \MVN (X\beta, \sigma^2 V).
-\end{equation}
 
-\subsection{Mat\'ern correlation}
-The powered exponential family, the Mat\'ern family and the spherical family are the three most commonly used families of correlation functions in geostatistics.
-We favour the Mat\'ern family because of its flexibility, and it is made available in our \prog{R} package \pkg{gpuBatchMatrix}.
-The form of Mat\'ern correlation used in \pkg{gpuBatchMatrix} is
+\subsection{Mat\'ern correlation function}
+There are three widely used classes of correlation functions in geostatistics: the Mat\'ern family, the powered exponential family and the spherical family.
+In general, we favour the Mat\'ern family because of its flexibility, this family has correlation function
 \begin{equation*}
-R(d;\phi,\kappa)=\frac{2^{\kappa-1}}{\Gamma(\kappa)} (\sqrt{8\kappa} \frac{d}{\phi})^\kappa  K_\kappa(\sqrt{8\kappa}  \frac{d}{\phi}),  \quad \text{where} \quad \phi>0,  \kappa > 0,
+R(d;\phi,\kappa)=\frac{2^{\kappa-1}}{\Gamma(\kappa)} (\sqrt{8\kappa} \frac{d}{\phi})^\kappa  K_\kappa(\sqrt{8\kappa}  \frac{d}{\phi}),  \quad \text{where} \quad \phi    \geq 0,  \kappa     \geq 0,
 \end{equation*} 
-$K_\kappa(\cdot)$ denotes the modified Bessel function of the second kind of order $\kappa$, 
-$\kappa$ is the shape parameter which determines the smoothness of $U(x)$.
-For $\kappa=0.5$, the Mat\'ern correlation function coincides with the exponential correlation $\exp(-d/\phi)$.
-When $\kappa \rightarrow \infty$, it tends to the Gaussian correlation $\exp{\{-2(\|d\|/\phi)^2\}}$.
-$\phi$ is called the range or scale parameter, it controls the rate of decay of the correlation as $d$ increases. 
-%$d =\|x-x'\|$ is the Euclidean distance between two spatial points, $\Gamma(\cdot)$ is the standard gamma function,  $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as $d$ increases.
+$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
+$\sigma^2$ the variance of the random field $U$.
+$\Gamma(\cdot)$ is the standard gamma function,  $\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as $d$ increases.
 $K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$, specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.
-Estimation of $\kappa$ is difficult, our experience is to choose the values of $\kappa$ from a discrete set, for example $\{0.5,1.5,2.5,\dots \}$(ref), %and then maximize the log-likelihood function \eqref{profile} or \eqref{remlpro} over $\phi$ and $\nu^2$.
+For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.
+Estimation of $\kappa$ is difficult, our experience is to choose the values of $\kappa$ from a discrete set, for example $\{0.5,1.5,2.5,\dots \}$(ref), and then maximize the log-likelihood function \eqref{profile} or \eqref{remlpro} over $\phi$ and $\nu^2$.
 When the number of location points is large, the computation for the Mat\'ern function ......
@@
 The fit of the linear geostatistical model can often be improved by applying a transformation to the response variable $Y$.
-Skewed random fields that mildly deviate from the Gaussian distribution can be modeled by means of the Box-Cox transformation (BCT) \cite{box1964analysis}.
+Asymmetric (skewed) random fields that mildly deviate from the Gaussian pdf can be modeled by means of the Box-Cox transformation \citep{box1964analysis}.
 %\cite{cressie2015statistics} introduced the term trans-Gaussian model to refer to the linear geostatistical model with a Box-Cox transformed  response variable. 
-For positive-valued response variable $Y$, the Box-Cox transformation has the form
-\begin{equation*}
-    Y'=
-    \begin{cases}
-      (Y^\lambda-1)/\lambda: & \lambda \in \bigr \quad \text{and} \quad \lambda \neq  0, \\
-      \log Y: & \lambda =0,
-    \end{cases}
-\end{equation*}
-where $Y'$ denotes the transformed response.
-The aim of the BCT is to make data closely resemble the Gaussian distribution, so that the usual assumptions for linear models hold.
-The BCT cannot guarantee that any probability distribution can be normalized.
-However, even in cases when an exact mapping to the normal distribution is not possible, BCT may still provide a useful approximation \cite{hristopulos2020random}.
-
-%The transformation parameter can be estimated by using maximum likelihood estimation. The log-likelihood including $\lambda$ is
-% \begin{equation*}
-% -2\ell^*(\beta,\sigma^2,V,\lambda)= -2*(\lambda-1)\sum_{i=1}^{n}\log y_i + n\log(2\pi)+\log |\sigma^2 V|+(y'-X\beta)^\T(\sigma^2 V)^{-1}(y'-X\beta),
-% \end{equation*}
+For positive-valued response variable $Y$, the Box-Cox transformation has the form:
 
 \subsection{Geometric anisotroy}
-Geometrical anisotropy is defined by two additional parameters in the correlation function.
-Let $(x_1, x_2)$ represent the original coordinates, which is rotated counterclockwise by $\varphi$, then the new coordinates $(x'_1, x'_2)$ is given by
-%
-  \begin{equation*}\label{eq:matrixeqn}
-    \begin{pmatrix}
-      x'_{1} \\
-      x'_{2}  
-    \end{pmatrix}
-    =
-    \begin{pmatrix}
-      1/\phi_{1} & 0 \\
-      0 & 1/ \phi_{2} 
-    \end{pmatrix}
-    \cdot
-    \begin{pmatrix}
-      \cos \varphi & -\sin \varphi \\
-      \sin \varphi & \cos \varphi 
-    \end{pmatrix}
-    \cdot
-      \begin{pmatrix}
-      x_1 \\
-      x_2 
-    \end{pmatrix},
-  \end{equation*}
-%
-where $\varphi \in [-\pi/2, \pi/2]$ is called the anisotropy angle, $\psi = \phi_1 / \phi_2 \in [0,1]$ is the anisotropy ratio.
-%Anisotropy in correlation functions Statistical anisotropy can also be present in the correlation functions, where it is more difficult to estimate accurately.
-
-\section{Inference based on the likelihood function}
-Let $y'^\T=(y'_1,\dots,y'_n)$ be the Box-Cox transformed data of the observations $y^\T=(y_1,\dots,y_n)$.
-The most general log-likelihood function for the linear geostatistical model parameters given $y'$ would be
-%
-\begin{equation}
--2\ell(\beta,\sigma^2, \lambda, \omega; y')=(y'-X\beta)^\T(\sigma^2 V)^{-1}(y'-X\beta)+\log |\sigma^2 V|-2*(\lambda-1)\sum_{i=1}^{n}\log y_i+n\log(2\pi).\label{eq:2}
-\end{equation}
-%
-the third term on the right-hand side of \eqref{eq:2} arises from the Jacobian of the transformation.
-This log-likelihood function is complex to evaluate as we have several types of parameters to estimate, including the regression parameters $\beta$, the transformation parameter $\lambda$, and the covariance parameters $(\sigma^2, \phi,\kappa,\nu^2, \varphi, \psi)$.
-For simplicity we use $\omega=(\phi,\kappa,\nu^2, \varphi, \psi)$ to denote the correlation parameters in the following.
-
-\subsection{Profile log-likelihood functions}
-%ADD gpuBatchMatrix and MORE STUFF\\
-%"a gradient-based optimization algorithm can be used to reduce the computational cost of the optimization\\
-%Another way to reduce the computational effort involved in the maximisation of l is to reduce the dimensionality of the optimization problem as follows."\\
-The profile log-likelihood for parameter $\varphi$ is defined as $\ell_p(\varphi;y) = \underset{\psi}{\text{sup}} \ell (\varphi, \psi; y ).$ Given $\omega$ and $\lambda$, The log-likelihood function \eqref{eq:2} is maximized at
-%
-\begin{gather}
-\hat{\beta}_{\omega,\lambda}=(X^\T V^{-1}X)^{-1} X^\T V^{-1}y', \quad \text{and} \label{betahat}\\
-\hat{{\sigma}}^2_{\omega,\lambda}= \frac{1}{n}(y'-X\hat{\beta}_{\omega,\lambda})^\T V^{-1}(y'-X\hat{\beta}_{\omega,\lambda}) \label{eq:4}.
-\end{gather}
-%
-%For simplicity, $\hat{\beta}_{w,\lambda}$ is written as $\hat{\beta}$ in \eqref{eq:4}, it denotes the maximum likelihood estimate of $\beta$ for a given $(w, \lambda)$. 
-Substitute \eqref{betahat} and \eqref{eq:4} back into \eqref{eq:2}, we have the profile log-likelihood for $\omega$ and $\lambda$
-%
-\begin{equation} \label{profile}
--2\ell_p(\omega, \lambda;y')=n \log\frac{(y'-X\hat{\beta}_{\omega,\lambda})^\T V^{-1}(y'-X\hat{\beta}_{\omega,\lambda})}{n}+\log|V| -2(\lambda-1)\sum_{i=1}^{n}\log y_i + n\log(2\pi) + n,
-\end{equation}
-%
-
-%\subsection{Restricted maximum likelihood estimation}
-The main problem of maximum likelihood estimation is that variance is underestimated. Think of the simple case when the covariance matrix of $Y$ is $\sigma^2 I$, then 
-\begin{align*}
-&\hat{\sigma}^2 = \frac{1}{n}(y-X\hat{\beta})^\T (y-X\hat{\beta}),\\
-&\E[\hat{\sigma}^2]=\frac{n-p}{n} \sigma^2 < \sigma^2,
-\end{align*}
-where $p= \text{rank}(X)$, is the number of elements in $\beta$.
-The estimation bias comes from not taking into account the degree of freedom used for estimating $\beta$.
-Restricted maximum likelihood (REML) developed by \cite{patterson1971recovery} is an approach that produces unbiased estimators or less biased estimators than MLE in general.
-REML eliminates the influence of $X$ on $\hat{\sigma}^2$ and $\hat{\omega}$.
-In this method, we find a matrix $A$ of full rank such that $AX=0$.
-The $n \times n$ projection matrix $S= I-X(X^\T X)^{-1}X^\T$ satisfies $SX=0$, however, $S$ is degenerate as $\text{rank}(S)=n-p$, so we take just $n-p$ linearly independent rows from $S$ to make the matrix $A_{(n-p)\times n}$.
-Transform the data linearly to $Y^*=AY$, then $\E(Y^*)=AX\beta=0$.
-Recall in \eqref{eq:1} we obtained $Y \sim \MVN (X\beta, \sigma^2 V),$ thus
-%
-\begin{equation*}
-Y^* \sim \MVN(0,\sigma^2 AVA^\T).
-\end{equation*}
-%
-The principle of the REML method is to estimate the variance parameter $\sigma^2$ by maximizing the restricted log-likelihood function below, where $y^*$ represents a realization of $Y^*$,
-%the distribution of $Y^*$ does not depend on $\beta$,
-\begin{equation} \label{eq:6}
--2\ell^*(\sigma^2, \omega ;y^*)=y^*\T(\sigma^2 AVA^\T)^{-1}y^*+\log|\sigma^2 AVA^\T|+(n-p)\log(2\pi),
-\end{equation}
-%&=(n-p)\log(2\pi)+ (n-p)\log \sigma^2 + \log|AVA^\T|+ \frac{1}{\sigma^2}y^*\T(AVA^\T)^{-1}y^*.
-Differentiating \eqref{eq:6} with regard to $\sigma^2$ and setting it equal to zero gives the unbiased estimate for $\sigma^2$
-%
-\begin{gather*} %\label{sigmahat_reml}
-\hat{\sigma}_{reml}^2(\omega)= \frac{y^*\T (AVA^\T)^{-1} y^*}{n-p}.
-\end{gather*}
-%
-The REML criterion is based on the likelihood of $Y^*$, in which $X$ does not appear.  REML bypasses estimating $\beta$ and can therefore, produce unbiased estimates for $\sigma^2$.
-
-%\textit{reml} \\
-We can write \eqref{eq:6} in terms of the original observed data $y$ using the two identities derived by \cite{searle1978notebook} in Section 5.1:
-$$|A (\sigma^2V)A^\T|=|\sigma^2V||X^\T (\sigma^2V)^{-1}X|$$ and $$y^*(\sigma^2 AVA^\T)^{-1}y^*=(y-X\hat{\beta})^\T (\sigma^2 V)^{-1}(y-X\hat{\beta}),$$ where $\hat{\beta}=(X^\T V^{-1}X)^{-1} X^\T V^{-1}y$.
-Then the restricted log-likelihood of $y$ is
-\begin{align*}
--2\ell^*(\sigma^2, \omega;y)=(y-X\hat{\beta})^\T(\sigma^2V)^{-1}(y-X\hat{\beta})+(n-p)\log \sigma^2+\log |V|+\log|X^\T V^{-1}X|+n\log(2\pi).
-\end{align*}
-Applying the BCT on $y$ we have the restricted log-likelihood in terms of the transformed response data $y'$
-\begin{align}
--2\ell^*(\sigma^2, \omega, \lambda;y')&=(y'-X\hat{\beta})^\T(\sigma^2V)^{-1}(y'-X\hat{\beta})+(n-p)\log \sigma^2+\log |V|+\log|X^\T V^{-1}X|+n\log(2\pi)-\notag\\
-&2(\lambda-1)\sum_{i=1}^{n}\log y_i,\label{eq:7}
-\end{align}
-which does not depend on the choice of matrix $A$.
-
-%\textit{reml profile} \\
-Differentiate \eqref{eq:7} with respect to $\sigma^2$ and set it equal to zero, we have the expression for $\hat{\sigma}_{reml}^2$ in terms of the Box-Cox transformed data given $\omega$ and $\lambda$,
-\begin{equation}\label{sigmahat_reml_y}
-\hat{\sigma}_{reml}^2(\omega, \lambda)=\frac{(y'-X\hat{\beta})^\T  V^{-1}(y'-X\hat{\beta})}{n-p}.
-\end{equation}
-Substitute \eqref{sigmahat_reml_y} into \eqref{eq:7}, leading to the profile restricted log-likelihood for $(\omega, \lambda)$
-\begin{equation}\label{remlpro}
--2\ell^*_p(\omega, \lambda;y')=(n-p)\log \hat{\sigma}^2+\log |V|+\log |X^\T V^{-1}X|-2(\lambda-1)\sum_{i=1}^{n}\log y_i+n\log(2\pi)+n-p.
-\end{equation}
-
-Compared to the ``ml'' profile log-likelihood given in \eqref{profile}, the effective dimension is reduced from $n$ of $\ell(\cdot)$  to $(n-p)$ of $\ell^*_p(\cdot)$ , this distinction is important if $p$ is large.
-
-\subsection{Parameter estimation}
-\textit{Geostatsp's methods}\vspace{0.5cm}\\
-The following are two important theorems in likelihood-based inference, which lead to two ways of calculating likelihood-based confidence intervals (CIs).
-%We write θˆ(y) for an observed value of θˆ and θˆ(Y ) for the corresponding random variable.
-We write $I(\theta)$ for the expected Fisher's information matrix, and $\theta$ is the true parameter value. 
-\begin{theorem}
-\label{R1}
-$\hat{\theta}(Y) \sim \MVN(\theta, I(\theta)^{-1})$.
-\end{theorem}
-%
-\begin{theorem}
-\label{R2}
-$2\{\ell(\hat\theta(Y); Y)-\ell(\theta;Y) \} \sim \chi^2_p ,$ where $p$ is the dimension of $\theta$.
-\end{theorem}
-The \fct{lgm} function in \pkg{geostatsp} package estimates the correlation parameters $\omega=(\phi,\kappa,\nu^2, \varphi, \psi)$ and the transformation parameter $\lambda$ by maximization of the likelihood.
-Specifically, \fct{lgm} uses the \textit{optim} function to optimize the equation \eqref{profile} (if \fct{reml}=FALSE)  numerically over $\omega$ and $\lambda$ to find $(\hat{\omega},\hat{\lambda}) = \argmax_{\omega,\lambda} \ell_p(\omega, \lambda; y')$.
-\pkg{geostatsp} treats $\hat{\omega}$ and $\hat{\lambda}$ as fixed and back substitute them into \eqref{betahat} and \eqref{eq:4} to obtain the maximum likelihood estimates (MLEs) $\hat{\beta}_{\omega,\lambda}$ and $\hat{{\sigma}}^2_{\omega,\lambda}$.
-The \fct{optim} function returns the Hessian evaluated at $\hat{\omega}$ and $\hat{\lambda}$.
-Conventionally, \fct{lgm} calculates CIs based on Theorem \ref{R1}.
-The 95\% confidence intervals for each of the parameters are calculated by $\hat{\theta_i} \pm 1.96 \sqrt{\var(\hat{\theta_i})}$, where $\theta_i$ represents any of the parameter in $(\beta,\sigma, \omega,\lambda)$, and  $\var(\hat{\theta}_i)$ is approximated by the diagonals of the inverse of the negative Hessian.
-This way of lcnaelcac does not take into account of uncertainty in correlation parameters.
-
-There is the other function \fct{profLlgm} in \pkg{geostatsp} that calculates confidence intervals or joint confidence regions based on the likelihood-ratio method \ref{R2}.
-If $c$ represents the $\beta$-quantile of $\chi^2_p$, i.e., $\Pr(\chi^2_p \leq c)=\beta$, then the set $\{\theta: \ell(\theta) \geq \ell(\hat{\theta}) - c/2 \}$ is asymptotically, a $\beta$-level confidence set for $\theta$.
-For a fixed $\phi$ value, \fct{optim} evaluates \eqref{profile} for different combinations of $(\kappa,\nu^2, \varphi, \psi)$ and finds the optim Log-likelihood value, repeat this process for a number of $\phi$ values to finally get the profile log-likelihood and MLE of $\phi$.
-\fct{profLlgm} runs \fct{optim} a great many of times and thus is very slow.
-
-%We can estimate $\lambda$ by maximization of the likelihood, since $\lambda$ is usually a nuisance parameter, an alternative and more pragmatic approach is to restrict $\lambda$ to a discrete set of values, for example $\{-1, 0, 1/2, 1\}$ \cite{laga2020model}.
-$\lambda = -1$ (reciprocal), $\lambda = 0$ (logarithm), $\lambda = 1/2$ (square-root) and $\lambda = 1$ (no transformation).
-
-\textit{gpuBatchMatrix's method}\vspace{0.5cm}\\
-The hard way is by computing the profile log-likelihood $\ell_p(\beta) =\ell(\beta, \hat{\omega}_{\beta},\hat{\sigma}_{\beta}, \hat{\lambda}_{\beta})$.
-(QUESTION: cannot geostatsp uses optim for beta?
-is it too slow so geostatsp does not do that?)
-
-SHOW A PROFILE LOG-LIKELIHOOD PLOT FOR BETA'S. (MEUSE DATA OR SWISSRAIN DATA?)\\
-SHOW ALL LIKELIHOODS PLOT VS BETA'S
-
-"The results R1 and R2 are invoked routinely in modern statistical practice, but they rely on assumptions that are not always satisfied.
-In a geostatistical context, the two most common situations in which these assumptions do not hold are boundary problems and dimensional ambiguities.
-"
-
-% Given $\omega$, $\lambda$ and $\sigma$, the log-likelihood \eqref{eq:2} would still be maximized at
-% %
-% \begin{gather*}
-% \hat{\beta}_{V,\lambda, \sigma}=(X^\T V^{-1}X)^{-1} X^\T V^{-1}y,
-% \end{gather*}
-% substitute $\hat{\beta}_{V,\lambda, \sigma}$ into \eqref{eq:2}, we obtain The profile log-likelihood for $(V, \lambda,\sigma)$ as
-%
-% \begin{equation} \label{profileBeta}
-% -2\ell_p(\sigma^2,\omega, \lambda;y)=(y-X\hat{\beta})^\T (\sigma^2 V)^{-1}(y-X\hat\beta)+\log|V|-2(\lambda-1)\sum_{i=1}^{n}\log y_i + n\log(2\pi)+ n \log\sigma^2.
-% \end{equation}
-%
-
-%The easy way for estimating the regression parameter $\beta$ is through $\ell(\beta) = \ell(\beta, \hat{\sigma}^2, \hat{\omega}, \hat{\lambda})$, where $(\hat{\omega},\hat{\sigma}, \hat{\lambda}) = \argmax_{\omega,\lambda,\sigma} \ell(\beta,\omega,\sigma, \lambda)$.
-The \prog{R} package \pkg{geostatsp} uses the \textit{optim} function to optimize the equation \eqref{profile} numerically over $\omega$ and $\lambda$ to find
-
-% 
-% 3, Given $V$, $\lambda$ and $\beta$, the log-likelihood \eqref{eq:2} is maximized at 
-%  %
-%  \begin{gather}
-%  \hat{\sigma}^2_{V,\lambda, \beta}= \frac{1}{n}(y-X \beta)^\T V^{-1}(y-X\beta) \label{sigmahat},
-%  \end{gather}
-%  %
-%  substitute $\hat{\sigma}^2_{V,\lambda, \beta}$ into \eqref{eq:2}, we have the following profile log-likelihood function for $(V,\lambda, \beta)$,
-%  %
-%  \begin{equation} \label{profileSigma}
-%  -2\ell_p(\beta,\phi,\kappa,\nu^2, \varphi, \psi, \lambda;y)=n \log\frac{(y-X \beta)^\T V^{-1}(y-X\beta)}{n}+\log|V|+\text{Jacobian}(\lambda)+n\log(2\pi)+n.
-%  \end{equation}
-
-\subsection{CI’s comparison for model parameters}
-%TWO WAYS OF CALCULATING CI IN LIKELIHOOD METHODS CI USING INFORMATION MATRIX: BOUNDRY PROBLEMS AND dimensional ambiguities.\\
-%The meaning of a confidence interval is a random set of values, constructed in such a way that, over repeated sampling of the data from the underlying model, that will contain the true, fixed but unknown value of the parameter with some specified probability.
-\cite{pd2007}
-%
-
-%
-%Conventionally, \pkg{geostatsp} uses the function \fct{lgm} to return CIs for parameters, which is based on the \ref{R1} method.
-Confidence intervals for individual parameter $\varphi_i$ would be $\hat{\varphi_i} \pm 1.96*\sqrt{\var(\hat{\varphi_i})}$, where $\var(\hat{\varphi}_i)$ can be approximated by the diagonals of inverted observed information matrix $\frac{\partial^2 l(\varphi)}{\partial\varphi \partial\varphi^\top}\big |_{\varphi=\hat{\varphi}_i}$.
-
-SHOW A TABLE OF CI compare between \pkg{geostatsp} and \pkg{gpuBatchMatrix}
-%"Calculating a likelihood-based confidence region is straightforward in principle but difficult in practice, rapidly becoming impossible as the number of elements of $\varphi$ grows.
-However, a more general version of the result also holds, and is extremely useful in practice." The results R1 and R2 are invoked routinely in modern statistical practice, but they rely on assumptions that are not always satisfied.
-In a geostatistical context, the two most common situations in which these assumptions do not hold are boundary problems and dimensional ambiguities.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-R> library(gpuRandom)
-R> library(gpuR)
-R> library(data.table)
-R> library("geostatsp")
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-% swissrain matplot over betas
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-R> params
-##      shape  range variance nugget anisoRatio anisoAngleRadians
-## [1,]  1.83  20000        1  0.000      10.00              0.08
-## [2,]  4.00  38619        1 15.000      16.00              0.70
-## [3,]  3.00  90000        1  0.137      25.00              0.30
-## [4,]  2.00 250000        1 20.000       8.09              0.50
-## [5,]  0.20  15000        1  0.020      65.00              6.50
-## [6,]  0.30  20000        1  0.080      45.00             12.50
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{figure}[H]
-
-{\centering \includegraphics[width=\maxwidth]{figure/matplot_beta-1}
-
-}
-
-\caption[profile log-Likelihoods of slope parameter for six covariance parameter sets]{profile log-Likelihoods of slope parameter for six covariance parameter sets}\label{fig:matplot_beta}
-\end{figure}
-
-\end{knitrout}
-% swissrain matplot over betas
-
-%meuse data
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## [1] "data.frame"
-## [1] "SpatialPointsDataFrame"
-## attr(,"package")
-## [1] "sp"
-##                    estimate   ci0.005 ci0.995
-## (Intercept)         0.91634  0.121205   1.711
-## sqrtdist           -0.08445 -1.623508   1.455
-## sdNugget            0.00785  0.000379   0.163
-## sdSpatial           0.01250  0.000690   0.227
-## range             173.16439 92.775636 323.209
-## shape               4.00000 -9.880685  17.881
-## anisoRatio          2.30106  1.138810   4.649
-## anisoAngleRadians   0.27490 -0.167582   0.717
-## anisoAngleDegrees  15.75042 -9.601751  41.103
-## boxcox             -0.90390 -2.558961   0.751
-##            nugget          variance             range 
-##             0.395             1.000           173.164 
-##             shape        anisoRatio anisoAngleRadians 
-##             4.000             2.301             0.275 
-##            boxcox 
-##            -0.904
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\pkg{gpuBatchMatrix} give wider ci and more accurate MLE's, expecially for the shape parameter.
-
-%meuse data
-
-\section{A simulation study}
-
-\section{Discussion}
-
-\bibliographystyle{unsrt}  
-\bibliography{paper2}
 

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3083 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<random exponential, eval=TRUE, cache=FALSE>>=
-streams <- createStreamsGpu(4)
-r_matrix <- clrng::rexp(c(2,4), rate=1, streams, Nglobal=c(2,2), type="double")
+r_matrix <- clrng::rexp(c(2,4), rate=1, myStreamsGpu2, Nglobal=c(2,2), type="double")
 as.matrix(r_matrix)
@@
 <<setupData, eval=TRUE,echo=TRUE, message=FALSE>>=
-library("gpuR")
-library("clrng")
 library("gpuBatchMatrix")

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1_2.R

@@ -28,21 +28,22 @@ setContext(   grep('gpu', listContexts()$device_type) [1]    )
 ## Section 2.1
 # Creating streams on CPU
 library("clrng")
-myStreamsCpu <- createStreamsCpu(n=4, initial=12345)
+setBaseCreator(c(111,222,333,444,555,666))
+myStreamsCpu <- createStreamsCpu(4)
 t(myStreamsCpu)
 
 # Creating streams on GPU
-myStreamsGpu = vclMatrix(myStreamsCpu)
-myStreamsGpu2 = createStreamsGpu(n=4, initial=12345)
+#myStreamsGpu = vclMatrix(myStreamsCpu)
+myStreamsGpu2 = createStreamsGpu(4)
 
 ## Section 2.2
 # Generate 6 i.i.d. U (0,1) random numbers
-as.vector(clrng::runif(n=6, streams=myStreamsGpu, Nglobal=c(2,2)))
+as.vector(clrng::runif(n=6, streams=myStreamsGpu2, Nglobal=c(2,2)))
 
 # Save streams on CPU
-saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
+# saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
 # Load the streams object and transfer it to GPU
-streams_saved <- vclMatrix(readRDS("myStreams.rds"))
+# streams_saved <- vclMatrix(readRDS("myStreams.rds"))
 
 ## Section 3.1
 # Generate a large matrix of normal random numbers, test the run time
@@ -53,7 +54,7 @@ system.time(matrix(stats::rnorm(10000^2),10000,10000))
 
 ## Section 3.2
 # Generate exponential random numbers
-r_matrix <- clrng::rexp(c(2,4), rate=1, Nglobal=c(2,2), type="double")
+r_matrix <- clrng::rexp(c(2,4), rate=1, myStreamsGpu2, Nglobal=c(2,2), type="double")
 as.matrix(r_matrix)
 
 
@@ -66,7 +67,7 @@ kable(month, format = "markdown", caption = "Monthly birth anomaly data")
 
 
 # using GPU
-streams <- createStreamsGpu(n =256*64, initial=666)
+streams <- createStreamsGpu(n =256*64)
 month_gpu<-vclMatrix(month,type="integer")
 system.time(result_month <- clrng::fisher.sim(month_gpu, 1e6, streams=streams,
[…]

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3084 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
-\pkg{clrng} is able to create streams both on the host and on the GPU device. The \fct{createStreamsCpu} function does the former. The \proglang{R} output below creates 4 streams on the host. 
+\fct{setBaseCreator} sets the initial seed of the package, it plays a role in \pkg{clrng} like the function \fct{SetPackageSeed} in \pkg{RngStreams} and \fct{clrngSetBaseCreatorState} in \pkg{clRNG}.
+\pkg{clrng} is able to create streams both on the host and on the GPU device.
+The \fct{createStreamsCpu} function does the former.
+The \proglang{R} output below creates 4 streams on the host.
 %is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
@@
 library("clrng")
-setBaseCreator(c(111,222,333,444,555,666))
+setBaseCreator(rep(12345,6))
 myStreamsCpu <- createStreamsCpu(4)
@@
 @
-
 %Users should run either \fct{createStreamsCpu} or \fct{createStreamsGpu} at the beginning of a task and retain these streams for all of the random number generation the task requires.
@@
 It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
-
-There are a few important notes on the usage of these functions.
-(1), Users should run the function \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
+There are a few important notes on the usage of these above functions.
+(1), Users should run \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
 If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams take the package's default initial seed.
-(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to run and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program results in creating some of the streams twice.
+(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to create streams and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program is very likely to create some streams twice.
 A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
 Streams should be created only once, and moved or copied if needed.
+And also it causes unnecessary data transfer between host and device if users switch often between \fct{createStreamsCpu} and \fct{createStreamsGpu}.
 

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3084 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 R> library("clrng")
-R> setBaseCreator(c(111, 222, 333, 444, 555, 666))
+R> setBaseCreator(rep(12345, 6))
 R> myStreamsCpu <- createStreamsCpu(4)
@@
 ##               [,1]       [,2]       [,3]       [,4]
-## current.g1.1  111  272931565 1097663848  628119774
-## current.g1.2  222  750526866  712444468  211319804
-## current.g1.3  333 1340941094  522805545  147140763
-## current.g2.1  444 1101709871  362320065 1665200176
-## current.g2.2  555 1485736734 1950619531  892386138
-## current.g2.3  666  375100068 1839101248  734766081
-## initial.g1.1  111  272931565 1097663848  628119774
-## initial.g1.2  222  750526866  712444468  211319804
-## initial.g1.3  333 1340941094  522805545  147140763
-## initial.g2.1  444 1101709871  362320065 1665200176
-## initial.g2.2  555 1485736734 1950619531  892386138
-## initial.g2.3  666  375100068 1839101248  734766081
+## current.g1.1 12345  336690377  502033783  739421137
+## current.g1.2 12345  597094797 1322587635 1475938232
+## current.g1.3 12345 1245771585 1964121530  730262207
+## current.g2.1 12345   85196284 1949818481 1630192198
+## current.g2.2 12345  523477687 1607232546  324551134
+## current.g2.3 12345 2094976052 1462898381  795289868
+## initial.g1.1 12345  336690377  502033783  739421137
+## initial.g1.2 12345  597094797 1322587635 1475938232
+## initial.g1.3 12345 1245771585 1964121530  730262207
+## initial.g2.1 12345   85196284 1949818481 1630192198
+## initial.g2.2 12345  523477687 1607232546  324551134
+## initial.g2.3 12345 2094976052 1462898381  795289868
 \end{CodeOutput} 
@@
 It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
-
-There are a few important notes on the usage of these functions.
-(1), Users should run the function \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
-If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams are created using the package's default initial seed.
-(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to run and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program will create some of the streams twice.
-When you create a new stream, it should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
+There are a few important notes on the usage of these above functions.
+(1), Users should run \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
+If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams take the package's default initial seed.
+(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to create streams and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program is very likely to create some streams twice.
+A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
 Streams should be created only once, and moved or copied if needed.
+And also it causes unnecessary data transfer between host and device if users switch often between \fct{createStreamsCpu} and \fct{createStreamsGpu}.
 
@@
 \begin{CodeOutput}
-## [1] 0.417 0.298 0.193 0.778 0.953 0.856
+## [1] 0.735 0.842 0.614 0.216 0.110 0.870
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.048   0.004   0.052
+##   0.050   0.000   0.051
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> streams <- createStreamsGpu(4)
-R> r_matrix <- clrng::rexp(c(2, 4), rate = 1, streams, Nglobal = c(2,
-+    2), type = "double")
+R> r_matrix <- clrng::rexp(c(2, 4), rate = 1, myStreamsGpu2,
++    Nglobal = c(2, 2), type = "double")
 R> as.matrix(r_matrix)
@@
 ##       [,1] [,2] [,3]  [,4]
-## [1,] 0.479 0.6886  1.2 0.62493
-## [2,] 0.721 0.0265  2.3 0.00598
+## [1,] 0.433 2.20 1.00 0.689
+## [2,] 1.770 1.13 1.48 1.205
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.313   0.008   0.320
+##   0.311   0.006   0.317
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 410423
+## [1] 409429
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.936   0.013   1.949
+##   1.940   0.013   1.951
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 1303
+## [1] 1350
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.00013
+## [1] 0.000135
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> library("gpuR")
-R> library("clrng")
 R> library("gpuBatchMatrix")

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3085 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<createStreams_Gpu2>>=
+rm(myStreamsCpu)
 myStreamsGpu2 = createStreamsGpu(4)
@@
 
-The streams created sequentially are used by work-items on a GPU (the GPU analogy of CPU cores) to generate random numbers.
+The streams created sequentially are used by work-items (the GPU analogy of CPU cores) on a GPU to generate random numbers.
 In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be at least the number of total work-items in use.
@@
 double temp, fact = mrg31k3p_NORM_cl;
-const int startvalue = (get_global_id(0) + get_global_size(0) * get_global_id(1)) * NpadStreams;
+const int startvalue = (get_global_id(0) + 
+get_global_size(0) * get_global_id(1)) * NpadStreams;
 
@@
 Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
-However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will restart from the current states after they are reloaded.
+However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will start from the current states after they are loaded.
 In this way we can reproduce the exact same sequence of random numbers in simulations.
@@
 <<save streams,echo=TRUE,eval=FALSE>>=
-saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
+saveRDS(as.matrix(myStreamsGpu2), "myStreams.rds")
 # Load the streams object as streams_saved

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3085 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{CodeInput}
+R> rm(myStreamsCpu)
 R> myStreamsGpu2 = createStreamsGpu(4)
@@
 \begin{CodeInput}
-R> saveRDS(as.matrix(createStreamsGpu(n = 4)), "myStreams.rds")
+R> saveRDS(as.matrix(myStreamsGpu2), "myStreams.rds")
 R> # Load the streams object as streams_saved
@@
 ##    user  system elapsed 
-##   0.050   0.000   0.051
+##   0.044   0.008   0.053
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.311   0.006   0.317
+##   0.302   0.016   0.318
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.940   0.013   1.951
+##   1.937   0.017   1.952
 \end{CodeOutput}

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1_2.R

@@ -28,12 +28,13 @@ setContext(   grep('gpu', listContexts()$device_type) [1]    )
 ## Section 2.1
 # Creating streams on CPU
 library("clrng")
-setBaseCreator(c(111,222,333,444,555,666))
+setBaseCreator(rep(12345,6))
 myStreamsCpu <- createStreamsCpu(4)
 t(myStreamsCpu)
 
 # Creating streams on GPU
 #myStreamsGpu = vclMatrix(myStreamsCpu)
+rm(myStreamsCpu)
 myStreamsGpu2 = createStreamsGpu(4)
 
 ## Section 2.2
@@ -41,7 +42,7 @@ myStreamsGpu2 = createStreamsGpu(4)
 as.vector(clrng::runif(n=6, streams=myStreamsGpu2, Nglobal=c(2,2)))
 
 # Save streams on CPU
-# saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
+# saveRDS(as.matrix(myStreamsGpu2), "myStreams.rds")
 # Load the streams object and transfer it to GPU
 # streams_saved <- vclMatrix(readRDS("myStreams.rds"))
 
@@ -188,12 +189,12 @@ zmatGpu = clrng::rnorm(
 
 # L*D*Z
 simMat = gpuBatchMatrix::multiplyLowerDiagonalBatch(maternCov, 
-                diagMat, zmatGpu,
-                diagIsOne = TRUE,   
-                transformD = "sqrt", 
-                Nglobal=c(128, 64, 2), 
-                Nlocal= c(8, 2, 1), 
-                NlocalCache=1000)
+                                                    diagMat, zmatGpu,
+                                                    diagIsOne = TRUE,   
+                                                    transformD = "sqrt", 
+                                                    Nglobal=c(128, 64, 2), 
+                                                    Nlocal= c(8, 2, 1), 
+                                                    NlocalCache=1000)
 
 
 # plot Setup
@@ -213,8 +214,6 @@ for(D in names(simRaster)) {
   plot(swissBorder, add=TRUE)
 }
 mapmisc::legendBreaks("right", myCol, inset=0)
-
-
 
 
 

### 2022-01-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3086 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
@@
 If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
-\pkg{clrng} is able to create streams both on the host and on the GPU device. The \fct{createStreamsCpu} function does the former. The \proglang{R} output below creates 4 streams on the host. 
+\fct{setBaseCreator} sets the initial state of the first stream that is created, it plays a role in \pkg{clrng} like the function \fct{SetPackageSeed} in \pkg{RngStreams} and \fct{clrngSetBaseCreatorState} in \pkg{clRNG}.
+\pkg{clrng} is able to create streams both on the host and on the GPU device.
+The \fct{createStreamsCpu} function does the former.
+The \proglang{R} output below creates 4 streams on the host.
 %is an interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on host as \proglang{R} matrices, see the \proglang{R} output below, where we created 4 streams on the host using \fct{createStreamsCpu}, after transposing \code{myStreamsCpu}, each column of the matrix is a stream, creating streams on host makes it easy for us to view and check the streams behavior.
@@
 It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
+There are a few important notes on the usage of these above functions.
+(1), We allow users to call \fct{setBaseCreator}  at any time.
+\fct{setBaseCreator} should be called once before \fct{createStreamsCpu} or \fct{createStreamsGpu} if they would like to select their own initial seeds, or never, in which case streams take the package's default initial seed.
+(2), Users may switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program, however, this is not recommended as it causes unnecessary data transfer between host and device.
+(3), If an \proglang{R} session is restarted during a task, as long as users saved the \proglang{R} environment, the initial states (seeds) of the newly created streams will carry on from the last streams' seed from the previous \proglang{R} session.
 
-There are a few important notes on the usage of these functions.
-(1), Users should run the function \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
-If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams take the package's default initial seed.
-(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to run and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program results in creating some of the streams twice.
-A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
+%A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
 Streams should be created only once, and moved or copied if needed.
+And also it causes unnecessary data transfer between host and device if users switch often between \fct{createStreamsCpu} and \fct{createStreamsGpu}.
 
@@
 \itemsep0em 
-  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: streams for random number generation. If missing, streams created on GPU with package default initial seed are supplied.
-  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
+  \item \code{n}: A vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
+  \item \code{streams}: Streams for random number generation, streams cannot be missing.
+  \item \code{Nglobal}: Global index space. Default is set as $(64,8)$.
   \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
-  \item \code{verbose}: print extra information if $\text{verbose} > 1$. Default is FALSE.
+  \item \code{verbose}: Print extra information if $\text{verbose} > 1$. Default is FALSE.
 \end{itemize}
@@
 Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
-However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will restart from the current states after they are reloaded.
-In this way we can reproduce the exact same sequence of random numbers in simulations.
-Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
+There are two ways to reproduce results, depending on how the program calls streams creator.
+The simpler way to reproduce is just to restart the program with the same initial seed for the creator.
+But it is important that the program always creates the streams in exactly the same order, for example, like the toy \proglang{R} program for demonstration below, we can reproduce the random matrix \code{sim_mat} just by keeping a record of the initial seed \code{c(11,22,33,44,55,66)}.
+<<showreproduce1, echo=TRUE, eval=FALSE>>=
+library('clrng')
+setBaseCreator(c(11,22,33,44,55,66))
+sim_mat <- matrix(0, nrow=10, ncol= 6)
+for (i in 1:10){
+  if( i %% 2 == 0){
+    streams1 <- createStreamsGpu(4)
+    sim_mat[i,] <- as.vector(clrng::rnorm(n=6, streams=streams1, Nglobal=c(2,2)))
+  }else if(i==3){
+    streams2 <- createStreamsGpu(4)
+    sim_mat[i,] <- as.vector(clrng::runif(n=6, streams=streams2, Nglobal=c(2,2)))
+  }else{
+    streams3 <- createStreamsGpu(8)
+    sim_mat[i,] <- as.vector(clrng::rexp(n=6, streams=streams3, Nglobal=c(4,2)))
+}
+}
+sim_mat
+@
+For more complicated applications, we recommend users to save the matrix of streams (current states and initial states) on the CPU in a file as a \code{.rds} object.
+And later recall the saved streams for regenerating results, the streams will start from their current states after they are loaded.
+This is a safer way than the previous one for reproducing results in simulations.
+Below shows the code that saves streams to a data file called \code{myStreams.rds} on CPU and then load it back and transfer it to a `vclMatrix' \code{streams_saved} on GPU, and using it to (re)generate some Normal random numbers.
 <<save streams,echo=TRUE,eval=FALSE>>=
-saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
+saveRDS(as.matrix(myStreamsGpu2), "myStreams.rds")
 # Load the streams object as streams_saved
 streams_saved <- vclMatrix(readRDS("myStreams.rds"))
+clrng::rnorm(n=6, streams=streams_saved, Nglobal=c(2,2))
 @
@@
 \itemsep0em 
-  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: streams. If missing, streams created on GPU with package default initial seed are supplied.
-  \item \code{Nglobal}: global index space. Default is $(64,8)$.
+  \item \code{n}: A vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
+  \item \code{streams}: Streams for random number generation, streams cannot be missing.
+  \item \code{Nglobal}: Global index space. Default is $(64,8)$.
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.
-  \item \code{verbose}: print extra information if $\text{verbose} > 1$. Default is FALSE.
+  \item \code{verbose}: Print extra information if $\text{verbose} > 1$. Default is FALSE.
 \end{itemize}
@@
 \setlength{\parskip}{0pt}
-  \item \code{x}: a contingency table, a ``vclMatrix'' object.
-  \item \code{N}: number of simulation runs.
-  \item \code{streams}: streams.
+  \item \code{x}: A contingency table, a ``vclMatrix'' object.
+  \item \code{N}: Number of simulation runs.
+  \item \code{streams}: Streams for random number generation, streams cannot be missing.
   \item \code{type}: ``double'' or ``float'' of returned test statistics.
-  \item \code{returnStatistics}: if TRUE, return all test statistics. 
-  \item \code{Nglobal}: global index space. Default is set as $(64,16)$.
+  \item \code{returnStatistics}: If TRUE, return all test statistics. 
+  \item \code{Nglobal}: Global index space. Default is set as $(64,16)$.
 \end{itemize}

### 2022-01-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3086 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 If we want to use the streams in work-items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work-item copies the current states only of the streams to its private memory.
-\fct{setBaseCreator} sets the initial seed of the package, it plays a role in \pkg{clrng} like the function \fct{SetPackageSeed} in \pkg{RngStreams} and \fct{clrngSetBaseCreatorState} in \pkg{clRNG}.
+\fct{setBaseCreator} sets the initial state of the first stream that is created, it plays a role in \pkg{clrng} like the function \fct{SetPackageSeed} in \pkg{RngStreams} and \fct{clrngSetBaseCreatorState} in \pkg{clRNG}.
 \pkg{clrng} is able to create streams both on the host and on the GPU device.
@@
 ##        [,1]       [,2]       [,3]       [,4]
-## current.g1.1 12345  336690377  502033783  739421137
-## current.g1.2 12345  597094797 1322587635 1475938232
-## current.g1.3 12345 1245771585 1964121530  730262207
-## current.g2.1 12345   85196284 1949818481 1630192198
-## current.g2.2 12345  523477687 1607232546  324551134
-## current.g2.3 12345 2094976052 1462898381  795289868
-## initial.g1.1 12345  336690377  502033783  739421137
-## initial.g1.2 12345  597094797 1322587635 1475938232
-## initial.g1.3 12345 1245771585 1964121530  730262207
-## initial.g2.1 12345   85196284 1949818481 1630192198
-## initial.g2.2 12345  523477687 1607232546  324551134
-## initial.g2.3 12345 2094976052 1462898381  795289868
+##  [1,] 12345  336690377  502033783  739421137
+##  [2,] 12345  597094797 1322587635 1475938232
+##  [3,] 12345 1245771585 1964121530  730262207
+##  [4,] 12345   85196284 1949818481 1630192198
+##  [5,] 12345  523477687 1607232546  324551134
+##  [6,] 12345 2094976052 1462898381  795289868
+##  [7,] 12345  336690377  502033783  739421137
+##  [8,] 12345  597094797 1322587635 1475938232
+##  [9,] 12345 1245771585 1964121530  730262207
+## [10,] 12345   85196284 1949818481 1630192198
+## [11,] 12345  523477687 1607232546  324551134
+## [12,] 12345 2094976052 1462898381  795289868
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> rm(myStreamsCpu)
 R> myStreamsGpu2 = createStreamsGpu(4)
@@
 There are a few important notes on the usage of these above functions.
-(1), Users should run \fct{setBaseCreator} once at the very beginning of a \proglang{R} session if they would like to choose their own initial seeds.
-If \fct{setBaseCreator} is not called before \fct{createStreamsCpu} or \fct{createStreamsGpu}, then streams take the package's default initial seed.
-(2), Users may choose either \fct{createStreamsCpu} or \fct{createStreamsGpu} to create streams and must stick with that choice for their entire program, switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program is very likely to create some streams twice.
-A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
+(1), We allow users to call \fct{setBaseCreator}  at any time.
+\fct{setBaseCreator} should be called once before \fct{createStreamsCpu} or \fct{createStreamsGpu} if they would like to select their own initial seeds, or never, in which case streams take the package's default initial seed.
+(2), Users may switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program, however, this is not recommended as it causes unnecessary data transfer between host and device.
+(3), If an \proglang{R} session is restarted during a task, as long as users saved the \proglang{R} environment, the initial states (seeds) of the newly created streams will carry on from the last streams' seed from the previous \proglang{R} session.
+
+%A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
 Streams should be created only once, and moved or copied if needed.
@@
 \begin{CodeOutput}
-## [1] 0.735 0.842 0.614 0.216 0.110 0.870
+## [1] 0.993 0.279 0.401 0.468 0.510 0.290
 \end{CodeOutput} 
@@
 Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
-However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will start from the current states after they are loaded.
-In this way we can reproduce the exact same sequence of random numbers in simulations.
-Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
+There are two ways to reproduce results, depending on how the program calls streams creator.
+The simpler way to reproduce is just to restart the program with the same initial seed for the creator.
+But it is important that the program always creates the streams in exactly the same order, for example, like the toy \proglang{R} program for demonstration below, we can reproduce the random matrix \code{sim_mat} just by keeping a record of the initial seed \code{c(11,22,33,44,55,66)}.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> library("clrng")
+R> setBaseCreator(c(11, 22, 33, 44, 55, 66))
+R> sim_mat <- matrix(0, nrow = 10, ncol = 6)
+R> for (i in 1:10) {
++    if (i%%2 == 0) {
++        streams1 <- createStreamsGpu(4)
++        sim_mat[i, ] <- as.vector(clrng::rnorm(n = 6, streams = streams1,
++            Nglobal = c(2, 2)))
++    } else if (i == 3) {
++        streams2 <- createStreamsGpu(4)
++        sim_mat[i, ] <- as.vector(clrng::runif(n = 6, streams = streams2,
++            Nglobal = c(2, 2)))
++    } else {
++        streams3 <- createStreamsGpu(8)
++        sim_mat[i, ] <- as.vector(clrng::rexp(n = 6, streams = streams3,
++            Nglobal = c(4, 2)))
++    }
++}
+R> sim_mat
+\end{CodeInput} 
+\end{CodeChunk}
+
+For more complicated applications, we recommend users to save the matrix of streams (current states and initial states) on the CPU in a file as a \code{.rds} object.
+And later recall the saved streams for regenerating results, the streams will start from their current states after they are loaded.
+This is a safer way than the previous one for reproducing results in simulations.
+Below shows the code that saves streams to a data file called \code{myStreams.rds} on CPU and then load it back and transfer it to a `vclMatrix' \code{streams_saved} on GPU, and using it to (re)generate some Normal random numbers.
 \begin{CodeChunk}
@@
 R> streams_saved <- vclMatrix(readRDS("myStreams.rds"))
+R> clrng::rnorm(n = 6, streams = streams_saved, Nglobal = c(2,
++    2))
 \end{CodeInput} 
@@
 ##    user  system elapsed 
-##   0.044   0.008   0.053
+##   0.047   0.004   0.051
 \end{CodeOutput} 
@@
 ##      [,1]  [,2] [,3] [,4]
-## [1,] 0.433 2.20 1.00 0.689
-## [2,] 1.770 1.13 1.48 1.205
+## [1,] 3.97 7.068 1.54 0.23
+## [2,] 1.11 0.247 2.05 1.33
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.302   0.016   0.318
+##   0.318   0.003   0.322
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 409429
+## [1] 409435
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.937   0.017   1.952
+##    1.93    0.02    1.95
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 1350
+## [1] 1351
 \end{CodeOutput}

### 2022-01-04 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
sdnugget
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3088 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\title{A title}
+\title{Profile likelihoods for geostatsistical models using GPU's}
 \date{}
@@
 \section{Introduction}
+Profile likelihoods are rarely used in geostatistical models due to the computational burden imposed by repeated likelihood evaluations.
+Accounting for uncertainty in covariance parameters can be highly consequential in geostatistical models as some covariance parameters are poorly identified, 
+the problem is severe enough that the differentiability parameter of the Matern correlation function is
+typically treated as fixed.  
+The problem is compounded with anisotropic spatial models as there are two additional parameters to consider.  
+This paper makes the following three contributions:
+
+1- a methodology and software implementation for highly parallel
+computing profile likelihoods on GPU's is presented;
+2- as expected, the profile-based confidence intervals have superior
+coverage to the more standard Wald-type intervals; and
+3- a simulation study shows that despite the inconsistency inherent in
+maximum likelihood estimation of geostatistical models, MLE's of covariance parameters (including the shape parameter) are unbiased and
+credible regions have the desired coverage.
+
 1- cI for the betas for the geostapsp package\\
@@
 \subsection{Parameter estimation}
+%The following are two important theorems in likelihood-based inference, which lead to two ways of calculating likelihood-based confidence intervals (CIs). 
+We write $I(\theta)$ for the expected Fisher information, here $\theta$ represents the true parameter value. $I(\hat{\theta})$ is the observed Fisher information evaluated at the MLE of $\theta$.
+\begin{description}[style=unboxed,leftmargin=0cm]
+\item[Wald confidence interval] $\{\hat{\theta} \pm z_{1-\alpha/2} * I(\hat{\theta})^{-1/2}\}$ is a $100(1 - \alpha)\%$ confidence interval for $\theta$, where $z_{1-\alpha/2}$ is the $1-\alpha/2$ percentile of the standard normal distribution.
+
+\item[Likelihood-based confidence interval] $\{\theta: \ell(\theta) \geq \ell(\hat{\theta}) - c_p(\beta)/2 \}$ is asymptotically, a $\beta$-level confidence set for $\theta$, where $c_p(\beta)$ is the $\beta$-quantile of $\chi^2_p$, i.e., $\Pr(\chi^2_p \leq c_p(\beta))=\beta$.
+\end{description}
+Wald intervals are always symmetric, and computationally much easier than likelihood-based confidence intervals, as quantities need to be calculated are all available from the algorithm used to find the MLEs of the model parameters.
+However, Wald intervals may include values which are not valid for the parameter in question.
+And if the likelihood function is not regular (cannot be well approximated by a quadratic function), then the curvature of the log-likelihood at the MLE ($I(\hat{\theta})$) or the standard error is not meaningful, which is often the case for spatial models where the parameter estimate is near a boundary, and the likelihood function is quite flat.
+In this case, Wald interval is deficient, likelihood ratio based confidence interval is a better and safer choice \cite{pawitan2001all}.
+For that reason, plotting the profile likelihood function in unfamiliar problems is generally advised.
+
+% \begin{theorem}
+% \label{R1}
+% $\hat{\theta}(Y) \sim \MVN(\theta, I(\theta)^{-1})$.
+% \end{theorem}
+% %
+% \begin{theorem}
+% \label{R2}
+% $2\{\ell(\hat\theta(Y); Y)-\ell(\theta;Y) \} \sim \chi^2_p ,$ where $p$ is the dimension of $\theta$.
+% \end{theorem}
 \textit{Geostatsp's methods}\vspace{0.5cm}\\
-The following are two important theorems in likelihood-based inference, which lead to two ways of calculating likelihood-based confidence intervals (CIs).
-%We write θˆ(y) for an observed value of θˆ and θˆ(Y ) for the corresponding random variable.
-We write $I(\theta)$ for the expected Fisher's information matrix, and $\theta$ is the true parameter value. 
-\begin{theorem}
-\label{R1}
-$\hat{\theta}(Y) \sim \MVN(\theta, I(\theta)^{-1})$.
-\end{theorem}
-%
-\begin{theorem}
-\label{R2}
-$2\{\ell(\hat\theta(Y); Y)-\ell(\theta;Y) \} \sim \chi^2_p ,$ where $p$ is the dimension of $\theta$.
-\end{theorem}
 The \fct{lgm} function in \pkg{geostatsp} package estimates the correlation parameters $\omega=(\phi,\kappa,\nu^2, \varphi, \psi)$ and the transformation parameter $\lambda$ by maximization of the likelihood.
-Specifically, \fct{lgm} uses the \textit{optim} function to optimize the equation \eqref{profile} (if \fct{reml}=FALSE)  numerically over $\omega$ and $\lambda$ to find $(\hat{\omega},\hat{\lambda}) = \argmax_{\omega,\lambda} \ell_p(\omega, \lambda; y')$.
-\pkg{geostatsp} treats $\hat{\omega}$ and $\hat{\lambda}$ as fixed and back substitute them into \eqref{betahat} and \eqref{eq:4} to obtain the maximum likelihood estimates (MLEs) $\hat{\beta}_{\omega,\lambda}$ and $\hat{{\sigma}}^2_{\omega,\lambda}$.
-The \fct{optim} function returns the Hessian evaluated at $\hat{\omega}$ and $\hat{\lambda}$.
-Conventionally, \fct{lgm} calculates CIs based on Theorem \ref{R1}.
-The 95\% confidence intervals for each of the parameters are calculated by $\hat{\theta_i} \pm 1.96 \sqrt{\var(\hat{\theta_i})}$, where $\theta_i$ represents any of the parameter in $(\beta,\sigma, \omega,\lambda)$, and  $\var(\hat{\theta}_i)$ is approximated by the diagonals of the inverse of the negative Hessian.
-This way of lcnaelcac does not take into account of uncertainty in correlation parameters.
+Specifically, \fct{lgm} uses the \textit{optim} function to optimize the equation \eqref{profile} (if \fct{reml}=FALSE)  numerically over many different sets of $(\omega, \lambda)$ to find $(\hat{\omega},\hat{\lambda}) = \argmax_{\omega,\lambda} \ell_p(\omega, \lambda; y')$.
+\pkg{geostatsp} then treats $\hat{\omega}$ and $\hat{\lambda}$ as fixed known parameters and back substitute them into \eqref{betahat} and \eqref{eq:4} to obtain the MLEs $\hat{\beta}_{\omega,\lambda}$ and $\hat{\sigma}^2_{\omega,\lambda}$.
+\pkg{geostatsp} calculates Wald confidence intervals for model parameters in \fct{lgm}.
+The \fct{numderiv} function inside \fct{lgm} returns the Hessian matrix evaluated at $\hat{\omega}$ and $\hat{\lambda}$, the diagonals of the inverse of the negative Hessian is used for approximating the variance of the parameters.This way does not take into account of uncertainty in correlation parameters.
+\pkg{geostatsp} does the above computation process in \prog{C}, and (ADD LATER!) which has achieved a great perfomance in terms of speed.
 
-There is the other function \fct{profLlgm} in \pkg{geostatsp} that calculates confidence intervals or joint confidence regions based on the likelihood-ratio method \ref{R2}.
-If $c$ represents the $\beta$-quantile of $\chi^2_p$, i.e., $\Pr(\chi^2_p \leq c)=\beta$, then the set $\{\theta: \ell(\theta) \geq \ell(\hat{\theta}) - c/2 \}$ is asymptotically, a $\beta$-level confidence set for $\theta$.
-For a fixed $\phi$ value, \fct{optim} evaluates \eqref{profile} for different combinations of $(\kappa,\nu^2, \varphi, \psi)$ and finds the optim Log-likelihood value, repeat this process for a number of $\phi$ values to finally get the profile log-likelihood and MLE of $\phi$.
-\fct{profLlgm} runs \fct{optim} a great many of times and thus is very slow.
+The other function \fct{profLlgm} in \pkg{geostatsp} calculates Likelihood-based confidence intervals.
+\fct{profLlgm} calculates the profile log-likelihood for each of the correlation parameter $\omega$ and $\lambda$, specifically, for estimating for example the range parameter $\phi$, for a fixed value of $\phi$, it uses \fct{optim} to maximize \eqref{profile} over a number of different combinations of $(\kappa,\nu^2, \varphi, \psi)$, and repeat this process for a number of $\phi$ values, and use \fct{approxfun} to get curve of the profile log-likelihood and MLE of $\phi$, then computes the confidence interval based on the profile log-likelihood.
+\fct{profLlgm} runs \fct{optim} a great many of times and can be time-consuming.
 
@@
 \subsection{CI’s comparison for model parameters}
-%TWO WAYS OF CALCULATING CI IN LIKELIHOOD METHODS CI USING INFORMATION MATRIX: BOUNDRY PROBLEMS AND dimensional ambiguities.\\
-%The meaning of a confidence interval is a random set of values, constructed in such a way that, over repeated sampling of the data from the underlying model, that will contain the true, fixed but unknown value of the parameter with some specified probability.
-\cite{pd2007}
-%
-
-%
-%Conventionally, \pkg{geostatsp} uses the function \fct{lgm} to return CIs for parameters, which is based on the \ref{R1} method.
-Confidence intervals for individual parameter $\varphi_i$ would be $\hat{\varphi_i} \pm 1.96*\sqrt{\var(\hat{\varphi_i})}$, where $\var(\hat{\varphi}_i)$ can be approximated by the diagonals of inverted observed information matrix $\frac{\partial^2 l(\varphi)}{\partial\varphi \partial\varphi^\top}\big |_{\varphi=\hat{\varphi}_i}$.
 
@@
  library(data.table)
-library('geostatsp')
+ #library(geostatsp)
 @
@@
 % swissrain matplot over betas
-<<params, echo=FALSE, results='hide',eval=TRUE>>=
-data('swissRain')
-swissRain$elevation = extract(swissAltitude, swissRain)
 
-params = rbind(
-  c(shape=1.826358, range=20000, variance = 1, nugget = 0, anisoRatio = 10, anisoAngleRadians = 0.08),  
-  c(shape=4, range=3.861873e+04, variance = 1, nugget = 15, anisoRatio = 16, anisoAngleRadians = 0.7),  
-  c(shape=3, range=90*1000, variance = 1, nugget = 1.369707e-01, anisoRatio = 25, anisoAngleRadians = 0.3),
-  c(shape=2, range=250*1000, variance = 1, nugget = 20, anisoRatio = 8.09, anisoAngleRadians = 0.5), 
-  c(shape=0.2, range=15*1000, variance = 1, nugget = 0.02, anisoRatio = 65, anisoAngleRadians = 6.5),
-  c(shape=0.3, range=20*1000, variance = 1, nugget = 0.08, anisoRatio = 45, anisoAngleRadians = 12.5)
-)
-colnames(params) <- c('shape','range', 'variance', 'nugget', 'anisoRatio', 'anisoAngleRadians')
-@
-<<showparams, eval=TRUE>>=
-params
-@
-
-<<compute, echo=FALSE, results='hide', eval=TRUE>>=
-result5<-gpuRandom::likfitLgmCov2d(
-  data=swissRain,
-  formula=rain~ elevation, 
-  coordinates=swissRain@coords,
-  params=params, 
-  boxcox = c(1,0,4.962398e-01),
-  cilevel=0.9,
-  type = "double",
-  reml=FALSE, 
-  NparamPerIter=400,
-  Nglobal=c(128,128),
-  Nlocal=c(16,16),
-  NlocalCache=2800,
-  verbose=FALSE)
-
-Betas <- seq(-12, 12, len=40)*1e-03
-
-m <- length(Betas)
-makeSymm <- function(m) {
-  m[upper.tri(m)] <- t(m)[upper.tri(m)]
-  return(m)
-}
-a=2 # which beta_i?
-Nobs=result5$Nobs  # number of observations.
-Ndata=result5$Ndata
-Nparam=result5$Nparam
-Ncov=result5$Ncov
-XVYXVX=result5$XVYXVX
-ssqY <- as.matrix(result5$ssqY)
-detVar <- as.vector(result5$detVar)
-detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
-jacobian <- as.vector(result5$jacobian)
-jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
-Ucov <- Ncov-1
-XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
-XVY <- XVYXVX[ , 1:Ndata]
-
-for (i in 1:Nparam){
-  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
-}
-
-selectedrows <- (seq_len(Nparam)-1) * Ncov + a
-XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
-XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
-XVY_deleted <- XVY[-selectedrows, ]
-X_aVY <- XVY[selectedrows, ]
-X_aVX_a <- XTVinvX[selectedrows, a]
-
-partA = matrix(0, nrow=Nparam, ncol=Ndata)
-partB = matrix(0, nrow=Nparam, ncol=Ndata)
-partC = matrix(0, nrow=Nparam, ncol=Ndata)
-partD = matrix(0, nrow=Nparam, ncol=Ndata)
-partE = matrix(0, nrow=Nparam, ncol=Ndata)
-
-for (i in 1:Nparam){
-  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
-  temp <- solve(XTVinvX_deleted[interval, ])
-
-  # part (A) have 2 data sets
-  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
-  # part (B) have 2 data sets.   has beta
-  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
-  # part (C) no data sets.  has beta
-  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
-  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
-  # part (D) have 2 data sets.    has beta
-  partD[i,] = -2* X_aVY[i, ]
-  # part (E)
-  partE[i,] = X_aVX_a[i]
-}
-
-All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
-for (bet in 1:m){
-  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
-    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
-}
-colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5','6')
-@
-
-<<matplot_beta, eval=TRUE,echo=FALSE, fig.height =4.5, fig.width=6, cache=TRUE, fig.cap='profile log-Likelihoods of slope parameter for six covariance parameter sets',fig.pos='H'>>=
-matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
-legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)
-@
-% swissrain matplot over betas
-
-%meuse data
-<<compare_mle, eval=TRUE,echo=FALSE>>=
-rm(list=ls(all=TRUE))
-library("sp")
-data(meuse)
-class(meuse)
-coordinates(meuse) <- c("x", "y")
-class(meuse)
-library('geostatsp')
-meuse$Y = log(meuse$zinc)
-meuse$sqrtdist = sqrt(meuse$dist)
-
-Res1 =  geostatsp::lgm( formula=Y~sqrtdist, 
-                        data=meuse, 
-                        grid=20,
-                        reml = FALSE,
-                        fixBoxcox=FALSE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
-                        aniso=TRUE )
-Res1$summary[,c('estimate','ci0.005', 'ci0.995')]
-Res1$optim$mle
-@
-
-\pkg{gpuBatchMatrix} give wider ci and more accurate MLE's, expecially for the shape parameter.
-<<compare_gpurandom, cache=FALSE, eval=FALSE,echo=FALSE>>=
-library('gpuRandom')
-library('gpuR')
-  ParamList = list(
-    range = c(exp(seq(log(100), log(173.1643928), len=6)), exp(seq(log(185), log(330), len=8))),
-    shape = c(seq(0.1,3.5, len=4), seq(4, 6, len=4)),
-    anisoRatio =c( seq(1, 2.301060332, len=3), seq(2.8, 7, len=4)),
-    anisoAngleRadians = c( seq(0, 0.274896758, len=4), seq(0.4, 0.8, len=4)),
-    nugget = c(seq(0,0.3946587, len=3), seq(0.5, 3, len=5))
-  ) 
-  params = do.call(expand.grid, ParamList)
-
-result3<- gpuRandom::likfitLgmCov(spatialmodel=Res1,
-                           params=params, # CPU matrix for now, users need to provide proper parameters given their specific need
-                           boxcox=c(seq(-3, -0.903901505, len=4), seq(-0.5,1,len=4)),  
-                           paramToEstimate = c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'), 
-                           cilevel=0.95, 
-                           type = "double",
-                           reml=FALSE, 
-                           NparamPerIter=400,
-                           Nglobal=c(128,64),
-                           Nlocal=c(16,16),
-                           NlocalCache=2800,
-                           verbose=FALSE)
-
-cbind(Res1$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),c('estimate','ci0.005', 'ci0.995')], result3$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),])
-@
-
-<<matplotmeuse, echo=FALSE, eval=TRUE>>=
-result1<-gpuRandom::getProfLogL(data=meuse,
-                                formula=Y~sqrtdist, 
-                                coordinates=meuse@coords,
-                                params=params, 
-                                boxcox = -0.903901505,
-                                type = "double",
-                                NparamPerIter=400,
-                                gpuElementsOnly=FALSE,
-                                reml=FALSE, 
-                                Nglobal=c(128,64),
-                                Nlocal=c(16,16),
-                                NlocalCache=2800)
-
-LogLik = result1$LogLik[seq(1,50176,400),3]
-length(LogLik) 
-Betas <- seq(-12, 12, len=40)*1e-03
-m <- length(Betas)
-makeSymm <- function(m) {
-  m[upper.tri(m)] <- t(m)[upper.tri(m)]
-  return(m)
-}
-a=2 # which beta_i?
-Nobs=result1$Nobs  # number of observations.
-Ndata=result1$Ndata
-Nparam=result1$Nparam[seq(1,50176,400),]
-Ncov=result1$Ncov
-XVYXVX=result1$XVYXVX
-ssqY <- as.matrix(result1$ssqY)
-detVar <- as.vector(result1$detVar)
-detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
-jacobian <- as.vector(result1$jacobian)
-jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
-Ucov <- Ncov-1
-XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
-XVY <- XVYXVX[ , 1:Ndata]
-
-for (i in 1:Nparam){
-  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
-}
-
-selectedrows <- (seq_len(Nparam)-1) * Ncov + a
-XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
-XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
-#XTVinvX_a <- matrix(XTVinvX[-selectedrows, a], nrow=Nparam*Ucov, ncol=1)[1:10,]
-XVY_deleted <- XVY[-selectedrows, ]
-X_aVY <- XVY[selectedrows, ]
-X_aVX_a <- XTVinvX[selectedrows, a]
-
-partA = matrix(0, nrow=Nparam, ncol=Ndata)
-partB = matrix(0, nrow=Nparam, ncol=Ndata)
-partC = matrix(0, nrow=Nparam, ncol=Ndata)
-partD = matrix(0, nrow=Nparam, ncol=Ndata)
-partE = matrix(0, nrow=Nparam, ncol=Ndata)
-
-for (i in 1:Nparam){
-  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
-  temp <- solve(XTVinvX_deleted[interval, ])
-
-  # part (A) have 2 data sets
-  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
-  # part (B) have 2 data sets.   has beta
-  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
-  # part (C) no data sets.  has beta
-  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
-  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
-  # part (D) have 2 data sets.    has beta
-  partD[i,] = -2* X_aVY[i, ]
-  # part (E)
-  partE[i,] = X_aVX_a[i]
-}
-
-All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
-for (bet in 1:m){
-  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
-    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
-}
-colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5', '6')
-matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
-legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)
-
-@
 %meuse data

### 2022-01-04 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.tex
sdnugget
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3088 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 R> library(data.table)
-R> library("geostatsp")
+R> # library(geostatsp)
 \end{verbatim}
@@
 % swissrain matplot over betas
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-R> params
-##      shape  range variance nugget anisoRatio anisoAngleRadians
-## [1,]  1.83  20000        1  0.000      10.00              0.08
-## [2,]  4.00  38619        1 15.000      16.00              0.70
-## [3,]  3.00  90000        1  0.137      25.00              0.30
-## [4,]  2.00 250000        1 20.000       8.09              0.50
-## [5,]  0.20  15000        1  0.020      65.00              6.50
-## [6,]  0.30  20000        1  0.080      45.00             12.50
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{figure}[H]
-
-{\centering \includegraphics[width=\maxwidth]{figure/matplot_beta-1}
-
-}
-
-\caption[profile log-Likelihoods of slope parameter for six covariance parameter sets]{profile log-Likelihoods of slope parameter for six covariance parameter sets}\label{fig:matplot_beta}
-\end{figure}
-
-\end{knitrout}
-% swissrain matplot over betas
-
-%meuse data
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## [1] "data.frame"
-## [1] "SpatialPointsDataFrame"
-## attr(,"package")
-## [1] "sp"
-##                    estimate   ci0.005 ci0.995
-## (Intercept)         0.91634  0.121205   1.711
-## sqrtdist           -0.08445 -1.623508   1.455
-## sdNugget            0.00785  0.000379   0.163
-## sdSpatial           0.01250  0.000690   0.227
-## range             173.16439 92.775636 323.209
-## shape               4.00000 -9.880685  17.881
-## anisoRatio          2.30106  1.138810   4.649
-## anisoAngleRadians   0.27490 -0.167582   0.717
-## anisoAngleDegrees  15.75042 -9.601751  41.103
-## boxcox             -0.90390 -2.558961   0.751
-##            nugget          variance             range 
-##             0.395             1.000           173.164 
-##             shape        anisoRatio anisoAngleRadians 
-##             4.000             2.301             0.275 
-##            boxcox 
-##            -0.904
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\pkg{gpuBatchMatrix} give wider ci and more accurate MLE's, expecially for the shape parameter.
 

### 2022-01-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3087 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3089 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 library('knitr')
-options(continue="+  ",  prompt="R> ", digits=3, width=57)
+options(continue="+  ",  prompt="R> ", digits=3, width=58)
 
@@
 \Abstract{
-We introduce the \proglang{R} package \pkg{clrng} which leverages the \pkg{gpuR} package to generate random numbers on a Graphics Processing Unit (GPU) with the \pkg{clRNG} (OpenCL) library.
+We introduce the \proglang{R} package \pkg{clrng} which leverages the \pkg{gpuR} package and is able to generate random numbers in parallel on a Graphics Processing Unit (GPU) with the \pkg{clRNG} (OpenCL) library.
 Parallel processing with GPU's can speed up computationally intensive tasks, which when combined with \proglang{R}, it can largely improve \proglang{R}’s downsides in terms of slow speed, memory usage and computation mode.
 %There is currently no \proglang{R} package that does random number generation on GPU, while random number generation is critical in simulation-based statistical inference and modeling. 
-\pkg{clrng} enables reproducible research by setting random streams on GPU and can thus accelerate several types of simulation and modelling.
+\pkg{clrng} enables reproducible research by setting random initial seeds for streams on GPU and CPU, and can thus accelerate several types of statistical simulation and modelling. 
+The random number generator in \pkg{clrng} guarantees independent parallel samples even when \proglang{R} is used interactively in an ad-hoc manner, with sessions being interrupted and restored.
 This package is portable and flexible, developers can use its random number generation kernel for various other purposes and applications.
@@
 
-The \pkg{clrng} package described here is currently the only \proglang{R} package that is able to generate random numbers on GPUs.
+The \pkg{clrng} package described here is currently the only \proglang{R} package that provides facilities for generating random numbers in parallel on GPU.
 Accomplishing this is complicated because each process must produce an independent (non-overlapping) sequence of random numbers, and in order to ensure reproducibility it should be possible to save and restore the current state of each stream of random numbers at any point.
-Here we introduce the \proglang{R} package \pkg{clrng} that leverages the \pkg{gpuR} package, and can generate random numbers on GPU by using the \proglang{clRNG} \citep{l2015clrng} library in \proglang{OpenCL}, and can thus accelerate several types of statistical simulation and modelling.
+By leveraging the \pkg{gpuR} package, \pkg{clrng} provides an extensible framework for \proglang{R} package developers to make use of these facilities in \proglang{C} or \proglang{OpenCL} code, and the random number generators guarantee independent parallel samples even when \proglang{R} is used interactively in an ad-hoc manner, with sessions being interrupted and restored.
+%and can generate random numbers on GPU by using the \proglang{clRNG} \citep{l2015clrng} library in \proglang{OpenCL}, and can thus accelerate several types of statistical simulation and modelling.
 
@@
 We move streams to the GPU by converting them to a `\code{vclMatrix}'.
-<<createStreams_Gpu, eval=FALSE>>=
+<<createStreams_Gpu, eval=TRUE>>=
 myStreamsGpu = vclMatrix(myStreamsCpu)
@@
 <<eval=FALSE,tidy=TRUE, include=FALSE>>=
-t(matrix(as.matrix(myStreamsGpu), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
+t(matrix(as.matrix(myStreamsGpu2), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
 @
@@
 <<random normal time compare, eval=TRUE, cache=FALSE>>=
-streams <- createStreamsGpu(n = 512 * 128)
+streams <- createStreamsGpu(512 * 128)
 system.time(clrng::rnorm(c(10000,10000), streams=streams, Nglobal=c(512,128), type="double"))
@@
 @
-<<monthDataForScript, include=FALSE, purl=TRUE>>-
+<<monthDataForScript, include=FALSE, purl=TRUE>>=
 knitr::kable(month)

### 2022-01-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3087 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3089 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.047   0.004   0.051
+##   0.045   0.004   0.049
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.318   0.003   0.322
+##   0.320   0.002   0.322
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##    1.93    0.02    1.95
+##   1.949   0.008   1.960
 \end{CodeOutput}

### 2022-01-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3090 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{CodeInput}
-R> streams <- createStreamsGpu(n = 512 * 128)
+R> streams <- createStreamsGpu(512 * 128)
 R> system.time(clrng::rnorm(c(10000, 10000), streams = streams,
@@
 ##    user  system elapsed 
-##   0.045   0.004   0.049
+##   0.049   0.000   0.048
 \end{CodeOutput} 
@@
 
-<<monthDataForScript, include=FALSE, purl=TRUE>>-
-knitr::kable(month)
-@
-
 \begin{CodeChunk}
@@
 ##    user  system elapsed 
-##   0.320   0.002   0.322
+##   0.319   0.002   0.320
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.949   0.008   1.960
+##   1.942   0.007   1.948
 \end{CodeOutput}

### 2022-01-05 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.tex
boxcox
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3095 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{abstract}
-some keypoints to mention:\\
-1, anisotropic spatial models are important but diffcult to evaluate because it consists of many covariance parameters.\\
-2, standard approximations of spatial models (GMRF) don’t allow for anisotropy (at least as implemented). \\
-3, gpuBatchMatrix is able to compute the profile LogL for each of the parameters in linear anisotropic spatial models, for dense matrix, use no approximation methods.
-allow for uncertainty in covariance parameters\\
-4, investigated likelihood-based CI’s for variance parameters, wider than using information matrix based methods (gepstatsp)\\
-\end{abstract}
-
-% keywords can be removed
-\keywords{First keyword \and Second keyword \and More}
-
-\section{Introduction}
 Profile likelihoods are rarely used in geostatistical models due to the computational burden imposed by repeated likelihood evaluations.
 Accounting for uncertainty in covariance parameters can be highly consequential in geostatistical models as some covariance parameters are poorly identified, 
-the problem is severe enough that the differentiability parameter of the Matern correlation function is
-typically treated as fixed.  
+the problem is severe enough that the differentiability parameter of the Matern correlation function is typically treated as fixed.  
 The problem is compounded with anisotropic spatial models as there are two additional parameters to consider.  
@@
 credible regions have the desired coverage.
+
+\end{abstract}
+
+% keywords can be removed
+\keywords{First keyword \and Second keyword \and More}
+
+\section{Introduction}
+some keypoints to mention:\\
+1, anisotropic spatial models are important but diffcult to evaluate because it consists of many covariance parameters.\\
+2, standard approximations of spatial models (GMRF) don’t allow for anisotropy (at least as implemented). \\
+3, gpuBatchMatrix is able to compute the profile LogL for each of the parameters in linear anisotropic spatial models, for dense matrix, use no approximation methods.
+allow for uncertainty in covariance parameters\\
+4, investigated likelihood-based CI’s for variance parameters, wider than using information matrix based methods (gepstatsp)\\
 
@@
 For that reason, plotting the profile likelihood function in unfamiliar problems is generally advised.
+"Further, a situation in which the Wald approach completely fails while the likelihood ratio approach is still (often) reasonable is when testing whether a parameter lies on the boundary of its parameter space"
 

### 2022-01-05 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
paper 2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3097 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  library(data.table)
- #library(geostatsp)
+ library(geostatsp)
 @
@@
 \section{A simulation study}
+<<simulate_datasets, echo=FALSE, cache=TRUE, results='hide', message=FALSE>>=
+############### A SIMULATION STUDY IN THE PAPER 2 ####################################
+Nsim=25
+set.seed(88)
+mydat = SpatialPointsDataFrame(cbind(seq(100*1e3,140*1e3, len=500), seq(100*1e3,120*1e3, len=500)), 
+                               data=data.frame(cov1 =stats::rnorm(500, 120, 20), cov2 = stats::rpois(500, 50))
+)
+
+## simulate a random field
+trueParam = c(range=1000, shape=2, variance=0.5^2, tauSq=1^2, anisoRatio=1, anisoAngleRadians=0)
+nugget = trueParam['tauSq']/trueParam['variance']
+names(nugget) <- "nugget"
+nugget   #4
+boxcox = c(boxcox=1)
+options("useRandomFields" = FALSE)
+for (i in 1:Nsim){
+  set.seed(i*6-1)
+  U <- geostatsp::RFsimulate(model=trueParam[setdiff(names(trueParam),'tauSq')],x=mydat)@data
+  # simulate response
+  Y <- 3 + 1*mydat$cov1 + 0.5*mydat$cov2 + U + stats::rnorm(length(mydat), 0, sd=sqrt(trueParam["tauSq"]))
+  mydat@data <- cbind(mydat@data,Y)
+}
+colnames(mydat@data) <- c("cov1", "cov2", paste("Y", c(1:Nsim), sep=""))
+
+## geostatsp's estimates
+swissRes =  lgm( formula=Y10~ cov1 + cov2,
+                 data=mydat,
+                 grid=20,
+                 reml = FALSE,
+                 boxcox=1,
+                 fixBoxcox=TRUE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
+                 aniso=FALSE )
+swissRes$summary[,c('estimate','ci0.005', 'ci0.995')]
+swissRes$optim$mle
+@
+
+<<set params, echo=FALSE, cache=TRUE>>=
+newParamList = list(
+  range = c(exp(seq(log(0.3), log(1), len=10))*1000, exp(seq(log(1.2), log(2.9), len=12))*1000),  
+  nugget = c(seq(0, 4, len=8), seq(4.5, 20, len=12)),  
+  shape = 2,
+  anisoRatio =1,
+  anisoAngleRadians = 0.0
+) 
+params = do.call(expand.grid, newParamList)
+@
+
+<<run_forloop, echo=FALSE, results='hide', message=FALSE, warning=FALSE, cache=TRUE>>=
+count = 0
+breaksf = rep(0, 10)
+makedata <- rep(0, 4)
+estimates <- rep(0, 6)
+
+library(gpuRandom)
+library(gpuR)
+count = 0
+breaksf = rep(0, 10)
+makedata <- rep(0, 4)
+estimates <- rep(0, 6)
+## calculate logl
+for (i in 1:Nsim){
+  result<-gpuRandom::likfitLgmCov2d(
+    data= mydat,
+    formula= as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")), 
+    coordinates=mydat@coords,
+    params=params, 
+    paramToEstimate = c('range','nugget'),
+    boxcox = 1,
+    cilevel=0.8,
+    type = "double",
+    reml=FALSE, 
+    NparamPerIter=400,
+    Nglobal=c(128,64),
+    Nlocal=c(16,16),
+    NlocalCache=2800,
+    verbose=FALSE)
+
+  a <- geostatsp::loglikLgm(
+    c(trueParam['range'],
+      trueParam['shape'],
+      nugget,
+      trueParam[c('anisoRatio')],
+      trueParam[c('anisoAngleRadians')],
+      boxcox),
+    data = mydat,
+    formula = as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")),
+    reml = FALSE,
+    minustwotimes=FALSE)[['logLik']]
+
+  if (a  >= result$breaks){
+    count = count +1
+  }
+  breaksf[i] = result$breaks
+  makedata0 <- cbind(params[,1:2], result$LogLik, D=i)
+  makedata <- rbind(makedata, makedata0)
+  estimates <- cbind(estimates, result$summary)
+}
+colnames(makedata) <- c("range",'nugget', 'LogLik', 'D')
+makedatahey <- makedata[-1,]
+estimates_summary <- estimates[,-1]
+mean_estimates <- apply(estimates_summary, 1, mean)
+@
+
+<<results, cache=TRUE>>=
+count/Nsim
+mean_estimates
+@
+
+<<graphs1, echo = FALSE, eval=TRUE, cache=TRUE>>=
+library(RColorBrewer)
+library('ggplot2')
+forigin <- ggplot(subset(makedatahey, D==1), aes(x = range, y = nugget, z=LogLik)) + theme_bw() + 
+  scale_y_continuous(limits=c(0,20))+
+  scale_x_continuous(limits=c(300,3000) )+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  geom_point(size = 0.5, colour='grey') + 
+  stat_contour(breaks=breaksf[1])  + 
+  geom_vline(xintercept = trueParam['range'])+
+  geom_hline(yintercept=nugget)
+cols <- rainbow(Nsim/2)
+for (i in 2:Nsim/2){
+  forigin <- forigin + geom_contour(data = subset(makedatahey, D==i), aes(x = range, y = nugget, z=LogLik), breaks=breaksf[i], col= cols[i])
+}
+@
+<<showgraphs1, echo = FALSE, eval=TRUE, cache=TRUE, fig.height =3, fig.width =4, fig.show ="hold",  fig.cap = "title", dev='png'>>=
+forigin
+@
 

### 2022-01-06 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3098 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The remaining sections are organized as follows:
-In Section 2, we introduce streams and the use of streams in work-items on a GPU device for generating uniform random numbers, and the usage of \fct{clrng::runif}.
+In Section 2, we introduce streams and the use of streams in work-items on a GPU device for generating uniform random numbers, and the usage of \fct{runifGpu}.
 In Section 3, we introduce two non-uniform RNGs in \pkg{clrng}, as examples for users to develop other RNGs of interest on GPUs in \proglang{R}.
@@
 (1), We allow users to call \fct{setBaseCreator}  at any time.
-\fct{setBaseCreator} should be called once before \fct{createStreamsCpu} or \fct{createStreamsGpu} if they would like to select their own initial seeds, or never, in which case streams take the package's default initial seed.
+\fct{setBaseCreator} should be called once before \fct{createStreamsCpu} and \fct{createStreamsGpu} if they would like to select their own initial seeds, or never, in which case streams take the package's default initial seed.
 (2), Users may switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program, however, this is not recommended as it causes unnecessary data transfer between host and device.
@@
 Now we use the 4 streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d.
-$U(0,1)$ random numbers with \fct{clrng::runif}.
+$U(0,1)$ random numbers with \fct{runifGpu}.
 To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
@@
 \begin{CodeInput}
-R> as.vector(clrng::runif(n = 6, streams = myStreamsGpu2, Nglobal = c(2,
+R> as.vector(runifGpu(n = 6, streams = myStreamsGpu2, Nglobal = c(2,
 +    2)))
@@
 
-The arguments of the \fct{clrng::runif} are described as follows:
+The arguments of the \fct{runifGpu} are described as follows:
 \begin{itemize}
@@
 \end{itemize}
-%\fct{runif} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here. 
+%\fct{runifGpu} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here. 
 Reusing \code{myStreamsGpu} will produce a vector different from \code{sim_1}.
@@
 +        streams1 <- createStreamsGpu(4)
-+        sim_mat[i, ] <- as.vector(clrng::rnorm(n = 6, streams = streams1,
++        sim_mat[i, ] <- as.vector(clrng::rnormGpu(n = 6, streams = streams1,
 +            Nglobal = c(2, 2)))
@@
 +        streams2 <- createStreamsGpu(4)
-+        sim_mat[i, ] <- as.vector(clrng::runif(n = 6, streams = streams2,
++        sim_mat[i, ] <- as.vector(runifGpu(n = 6, streams = streams2,
 +            Nglobal = c(2, 2)))
@@
 +        streams3 <- createStreamsGpu(8)
-+        sim_mat[i, ] <- as.vector(clrng::rexp(n = 6, streams = streams3,
++        sim_mat[i, ] <- as.vector(rexpGpu(n = 6, streams = streams3,
 +            Nglobal = c(4, 2)))
@@
 R> streams_saved <- vclMatrix(readRDS("myStreams.rds"))
-R> clrng::rnorm(n = 6, streams = streams_saved, Nglobal = c(2,
+R> clrng::rnormGpu(n = 6, streams = streams_saved, Nglobal = c(2,
 +    2))
@@
 
-We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{clrng::rnorm}.
-The best performance we have seen using \fct{clrng::rnorm} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
+We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}.
+The best performance we have seen using \fct{rnormGpu} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size goes larger.
@@
 R> streams <- createStreamsGpu(512 * 128)
-R> system.time(clrng::rnorm(c(10000, 10000), streams = streams,
-+    Nglobal = c(512, 128), type = "double"))
+R> system.time(rnormGpu(c(10000, 10000), streams = streams, Nglobal = c(512,
++    128), type = "double"))
 \end{CodeInput}
@@
 ##    user  system elapsed 
-##   0.049   0.000   0.048
-\end{CodeOutput} 
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> system.time(matrix(stats::rnorm(10000^2), 10000, 10000))
+##   0.051   0.000   0.050
+\end{CodeOutput} 
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> system.time(matrix(rnorm(10000^2), 10000, 10000))
 \end{CodeInput}
@@
 ##    user  system elapsed 
-##   6.947   0.736   7.686
+##   7.006   0.777   7.782
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> r_matrix <- clrng::rexp(c(2, 4), rate = 1, myStreamsGpu2,
-+    Nglobal = c(2, 2), type = "double")
+R> r_matrix <- rexpGpu(c(2, 4), rate = 1, myStreamsGpu2, Nglobal = c(2,
++    2), type = "double")
 R> as.matrix(r_matrix)
@@
 
-The arguments of the \fct{clrng::rexp} are described as follows:
+The arguments of the \fct{rexpGpu} are described as follows:
 \begin{itemize}
@@
 ##    user  system elapsed 
-##   0.319   0.002   0.320
+##   0.323   0.000   0.324
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##   1.942   0.007   1.948
+##   1.942   0.014   1.954
 \end{CodeOutput}
@@
 \end{gather}
-Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnormGpu}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 \begin{CodeChunk}
@@
 \begin{CodeInput}
-R> zmatGpu = clrng::rnorm(c(nrow(maternCov), 2), streams = streamsGpu,
+R> zmatGpu = clrng::rnormGpu(c(nrow(maternCov), 2), streams = streamsGpu,
 +    Nglobal = c(128, 64), type = theType)

### 2022-01-06 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3098 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Now we use the 4 streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d.
-$U(0,1)$ random numbers with \fct{clrng::runif}.
+$U(0,1)$ random numbers with \fct{runifGpu}.
 To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
 <<simStuff2, eval=TRUE,tidy=TRUE, echo=TRUE>>=
-as.vector(clrng::runif(n=6, streams=myStreamsGpu2, Nglobal=c(2,2)))
+as.vector(runifGpu(n=6, streams=myStreamsGpu2, Nglobal=c(2,2)))
 @
-The arguments of the \fct{clrng::runif} are described as follows:
+The arguments of the \fct{runifGpu} are described as follows:
 \begin{itemize}
@@
     streams1 <- createStreamsGpu(4)
-    sim_mat[i,] <- as.vector(clrng::rnorm(n=6, streams=streams1, Nglobal=c(2,2)))
+    sim_mat[i,] <- as.vector(clrng::rnormGpu(n=6, streams=streams1, Nglobal=c(2,2)))
   }else if(i==3){
     streams2 <- createStreamsGpu(4)
-    sim_mat[i,] <- as.vector(clrng::runif(n=6, streams=streams2, Nglobal=c(2,2)))
+    sim_mat[i,] <- as.vector(runifGpu(n=6, streams=streams2, Nglobal=c(2,2)))
   }else{
     streams3 <- createStreamsGpu(8)
-    sim_mat[i,] <- as.vector(clrng::rexp(n=6, streams=streams3, Nglobal=c(4,2)))
+    sim_mat[i,] <- as.vector(rexpGpu(n=6, streams=streams3, Nglobal=c(4,2)))
 }
@@
 streams_saved <- vclMatrix(readRDS("myStreams.rds"))
-clrng::rnorm(n=6, streams=streams_saved, Nglobal=c(2,2))
+clrng::rnormGpu(n=6, streams=streams_saved, Nglobal=c(2,2))
 @
@@
 streams <- createStreamsGpu(512 * 128)
-system.time(clrng::rnorm(c(10000,10000), streams=streams, Nglobal=c(512,128), type="double"))
+system.time(rnormGpu(c(10000,10000), streams=streams, Nglobal=c(512,128), type="double"))
 @
 <<random normal time compare2, eval=TRUE, cache=TRUE>>=
-system.time(matrix(stats::rnorm(10000^2),10000,10000))
+system.time(matrix(rnorm(10000^2),10000,10000))
 @
@@
 <<random exponential, eval=TRUE, cache=FALSE>>=
-r_matrix <- clrng::rexp(c(2,4), rate=1, myStreamsGpu2, Nglobal=c(2,2), type="double")
+r_matrix <- rexpGpu(c(2,4), rate=1, myStreamsGpu2, Nglobal=c(2,2), type="double")
 as.matrix(r_matrix)
 @
-The arguments of the \fct{clrng::rexp} are described as follows:
+The arguments of the \fct{rexpGpu} are described as follows:
 \begin{itemize}
@@
 \end{gather}
-Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnormGpu}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 <<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
@@
 <<r randomNormalsSim2, eval=TRUE, cache=FALSE>>=
-zmatGpu = clrng::rnorm(
+zmatGpu = clrng::rnormGpu(
                 c(nrow(maternCov),2), streams=streamsGpu, 

### 2022-01-07 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3100 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 @
-\fct{setBaseCreator} has one argument:
-\begin{itemize}
-\itemsep0em 
-  \item \code{initial}: a vector of six 31-bit integers, recycled if length is shorter. Default is set to \code{c(12345, 12345, 12345, 12345, 12345, 12345)}.
-\end{itemize}
-\fct{createStreamsCpu} has one argument:
-\begin{itemize}
-\itemsep0em 
-  \item \code{n}: number of streams to create.
-\end{itemize}
+\fct{setBaseCreator} has a single argument, the state of the first stream which defaults to a vector of 12345's.  
+The argument of \fct{createStreamsCpu} is the number of streams to create.
+Normally users would create many more than the four presented here, as most GPU's have thousands of work items.
+The default number of streams is 1024.
+
 We move streams to the GPU by converting them to a `\code{vclMatrix}'.
@@
 Equivalently, \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-<<createStreams_Gpu2>>=
-myStreamsGpu2 = createStreamsGpu(4)
-@
-%Users should run either \fct{createStreamsCpu} or \fct{createStreamsGpu} at the beginning of a task and retain these streams for all of the random number generation the task requires.
-Creating new streams with different initial states introduces the very small possibility of overlap between the original and new streams.
-If an \proglang{R} session is restarted during a task, perhaps because a preliminary analysis is done following an initial set of simulations and the full set of simulations continued at a later date, the state of the streams should be transferred to the CPU and saved to disk.
-The `\code{initial}' argument to the stream creator should be used with caution and we recommend using the default.
-It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
+
 There are a few important notes on the usage of these above functions.
-(1), We allow users to call \fct{setBaseCreator}  at any time.
+\begin{itemize}
+\item Calling \fct{setBaseCreator} creates a hidden object \code{.Random.seed.clrng}, which (unlike the standard \code{.Random.seed}) controls the creation of new streams (not random number).
+We allow users to call \fct{setBaseCreator}  at any time.
 \fct{setBaseCreator} should be called once before \fct{createStreamsCpu} and \fct{createStreamsGpu} if they would like to select their own initial seeds, or never, in which case streams take the package's default initial seed.
-(2), Users may switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program, however, this is not recommended as it causes unnecessary data transfer between host and device.
-(3), If an \proglang{R} session is restarted during a task, as long as users saved the \proglang{R} environment, the initial states (seeds) of the newly created streams will carry on from the last streams' seed from the previous \proglang{R} session.
+Otherwise, there is a small probability of producing streams which overlap.
+\item Users may switch between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program, calling either will read and update \code{.Random.seed.clrng}.
+This is not recommended as it causes unnecessary data transfer between host and device.
+\item If an \proglang{R} session is restarted during a task, as long as users saved the \proglang{R} environment (which Rstudio does automatically), the initial states (seeds) of the newly created streams will carry on from the last streams' seed from the previous \proglang{R} session.
+\item Objects on the GPU are lost when an R session is restarted, even if the workspace is saved.
+The object \code{myStreamsGpu} will remain in the workspace if the session is restarted, but the GPU memory containing \code{myStreamsGpu}'s data will no longer be accessible and an error will occur if the object is accessed.
+To retain streams when a session is restarted the streams should be copied to the CPU with \code{as.matrix}.
+\end{itemize}
 
-%A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
-Streams should be created only once, and moved or copied if needed.
-And also it causes unnecessary data transfer between host and device if users switch often between \fct{createStreamsCpu} and \fct{createStreamsGpu}.
-
-%However, any data is lost when the \proglang{R} session is terminated.
-%\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
-Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we produce on device exactly the same streams as those created on host as follows.
-Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
-
-%also has the two arguments \code{n} and \code{initial}.
-\fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
-By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{initial} argument, we make the sequence of streams repeatable.
+Unlike \proglang{R}'s standard random number generators, the streams need to be explicitly specified when generating random numbers with \code{clrng}.
+It would be possible to hide streams from the user by storing them as a hidden object, which the \pkg{rlecuyer} package does to some extent, we have chosen not to do so for two reasons.
+First, the streams would need to be stored on the CPU and transferred to and from the GPU every time they were used.
+Second, computations might be distributed across multiple GPU devices or multiple computers, and being explicit about which streams are used where ensures reproducibility and avoids the possibility that the same streams might be used on different devices without the user being aware.
 
@@
 The streams created sequentially are used by work-items (the GPU analogy of CPU cores) on a GPU to generate random numbers.
-In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams should always be at least the number of total work-items in use.
+In \pkg{clrng} each work-item takes one distinct stream to generate random numbers, so the number of streams created should always equal (or exceed) the maximum number of work items likely to be used.
 The main part of the kernel (functions for execution on the GPU) for generating uniform random numbers is shown in Listing \ref{lst:uniformkernel}.
@@
 <<simStuff2, eval=TRUE,tidy=TRUE, echo=TRUE>>=
-as.vector(runifGpu(n=6, streams=myStreamsGpu2, Nglobal=c(2,2)))
+sim_1 =  runifGpu(n = 8, streams = myStreamsGpu, Nglobal = c(2,2)))
+as.vector(sim_1)
 @
@@
 \itemsep0em 
-  \item \code{n}: A vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: Streams for random number generation, streams cannot be missing.
-  \item \code{Nglobal}: Global index space. Default is set as $(64,8)$.
-  \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
-  \item \code{verbose}: Print extra information if $\text{verbose} > 1$. Default is FALSE.
+  \item \code{n}: Either a scalar giving the number of samples to generate or a vector of length 2 specifying the size of a matrix to be output.
+  \item \code{streams}: Streams for random number generation, which must be stored on the GPU as a \code{vclMatrix} object.
+  \item \code{Nglobal}: Number of work items, by default it is set as $(64,8)$.
+  \item \code{type}: \code{double}, \code{float} or \code{integer}, the format of generated random numbers, defaults to \code{double}.
+  \item \code{verbose}: Print extra information if $\text{verbose} > 1$.
 \end{itemize}
-%\fct{runifGpu} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here. 
-Reusing \code{myStreamsGpu} will produce a vector different from \code{sim_1}.
-Because each time a random number is generated, the current state of the stream advances by one position.
-%Below shows the \code{myStreamsGpu} after it being used.
-<<eval=FALSE,tidy=TRUE, include=FALSE>>=
-t(matrix(as.matrix(myStreamsGpu2), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
-@
-Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart.
-Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
+
+The object returned is either a \code{vclMatrix} or \code{vclVector}, depending on whether \code{n} is two- or one-dimensional.
+Reusing \code{myStreamsGpu} will produce a vector different form \code{sim_1}, as the current states of the streams in \code{myStreamsGpu} have advanced by two positions (each of the 4 work item generated 2 random numbers).
+
+As mentioned earlier, streams on the GPU do not remain in the memory when an \proglang{R} session is restarted.
 There are two ways to reproduce results, depending on how the program calls streams creator.
-The simpler way to reproduce is just to restart the program with the same initial seed for the creator.
+The simpler way to reproduce is just to have a single program in an \proglang{R} script file with the initial seed for the creator set at the beginning.
 But it is important that the program always creates the streams in exactly the same order, for example, like the toy \proglang{R} program for demonstration below, we can reproduce the random matrix \code{sim_mat} just by keeping a record of the initial seed \code{c(11,22,33,44,55,66)}.
@@
 @
+
+Were we to replace the \code{for} loop above with a parallel equivalent (i.e.\ \code{mcmapply}), the order of stream creation could change with each program execution and the result would not be reproducible.
 For more complicated applications, we recommend users to save the matrix of streams (current states and initial states) on the CPU in a file as a \code{.rds} object.
@@
 <<save streams,echo=TRUE,eval=FALSE>>=
-saveRDS(as.matrix(myStreamsGpu2), "myStreams.rds")
+saveRDS(as.matrix(myStreamsGpu), "myStreams.rds")
 # Load the streams object as streams_saved
@@
 @
-
-%\subsection{Plot the uniform random numbers (now deleted?)}
 
@@
 %The derivation basically uses transformation from Cartesian coordinates to polar coordinates to represent two independent, Normally distributed random variables.
-The Box-Muller algorithm is a very good choice for Gaussian transform on GPU compared to other transform methods \citep{howes2007efficient}, because this algorithm has no branching or looping, which are the things GPU is not good at.
-%but GPU is good at the high computational load of sine and cosine functions in the algorithm.
+The Box-Muller algorithm is a very good choice for Gaussian transform on GPU's compared to other transform methods \citep{howes2007efficient}, because this algorithm has no branching or looping, which are the things GPU's are poor at.
 
@@
 Listing \ref{lst:normal} shows a fragment of the kernel that generates standard Gaussian random numbers.
-Aside from using local memory and the part that does the Box-Muller transformation, this kernel is the same with the kernel for generating uniform random numbers shown in Listing \ref{lst:uniformkernel}.
-If developers want to use other Gaussian transform methods, this kernel is easy to be adapted and incorporated in other \proglang{R} packages.
-This kernel executes over a 2-dimensional index space with $1 \times 2$ work-groups.
-\code{part[0]} and \code{part[1]} correspond to $R$ and $\Theta$ in the formulas respectively.
-\code{PI_2} is defined to be $\pi/2$.
+The kernel has work items operating in pairs with shared local memory, the $U_1$ and $U_2$ are generated in parallel and stored in the two-dimenstional vector \code{part} below.
+The \code{get_local_id(1)} command will return either zero or one, for the first and second item in the pair respectively.
 As the work-items in a work-group proceed at different rates, \code{barrier(CLK_LOCAL_MEM_FENCE)} ensures correct ordering of memory operations to local memory, so that Gaussian random numbers $(X_1,Y_1), \dots, (X_n,Y_n)$ are generated in pairs correctly, errors such as $(X_n, Y_{(n-1)})$ or $(X_{(n-1)}, Y_{(n)})$ are avoided.
-%
 
@@
 uint g1[3], g2[3];
-local double part[2];
 int startvalue = (get_global_id(0) * get_global_size(1) + 
 get_global_id(1)) * NpadStreams;
+double temp;
 
-double sinOrCosPart1, addForSine = - get_local_id(1) * PI_2;
-double temp, fact = mrg31k3p_NORM_cl;
-if(get_local_id(1)){
-  fact = TWOPI_mrg31k3p_NORM_cl;
-}
+const double fact[2] = { mrg31k3p_NORM_cl, TWOPI * mrg31k3p_NORM_cl };
+const double addForSine[2] = { 0.0, - PI_2 };
+local double part[2];
 
@@
 
-    for(Dcol=get_global_id(1); DcolBlock < Ncol; 
+    for(Dcol=get_global_id(1); Dcol < Ncol; 
         Dcol += get_global_size(1)){
 
-      temp = fact * clrngMrg31k3pNextState(g1, g2);
-      part[get_local_id(1)] = temp;
-
-      if(!get_local_id(1)) {
-        part[0] = sqrt(-2.0*log(part[0]));
-      }
+      part[get_local_id(1)] = fact[get_local_id(1)] * clrngMrg31k3pNextState(g1, g2);
       barrier(CLK_LOCAL_MEM_FENCE);
-
-      // sine for local[1], cos for local[0]
-      sinOrCosPart1 = cos(part[1] + addForSine);
-      out[Drow * Npad + Dcol] = part[0]*sinOrCosPart1;
-
+      temp = sqrt( -2.0*log(part[0]) ) * cos(part[1] + addForSine[get_local_id(1)] );
+      out[Dentry] = temp;
       barrier(CLK_LOCAL_MEM_FENCE);
@@
 
-We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}.
-The best performance we have seen using \fct{rnormGpu} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
-The difference in elapsed time becomes larger when matrix size goes larger.
+Below we generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}.
+The best performance we have seen using \fct{rnormGpu} is above over 100 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}.
+The difference in elapsed time becomes larger when matrix size increases.
 <<random normal time compare, eval=TRUE, cache=FALSE>>=
@@
 \section{An application of uniform RNG: Fisher's simulation}
-The GPU-generated random numbers can be applied in suitable statistical simulations to accelerate the performance.
+The GPU-generated random numbers can be applied in suitable statistical simulations to accelerate computations.
 One application of GPU-generated random numbers in \pkg{clrng} is Monte Carlo simulation for Fisher's exact test.
@@
   \item Calculate the test statistic for the observed table.
-    \item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table using the \fct{rcont2} algorithm, compute and sometimes save the test statistic from the random table.
+    \item In each iteration, simulate a random table with the same dimensions and marginal totals as the observed table using the \fct{rcont2} algorithm, compute and optionally save the test statistic from the random table.
     \item Count the number of iterations (\emph{Counts}) that have test statistics less or equal to the one from the observed table.

### 2022-01-07 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
edits to paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3101 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 uint g1[3], g2[3];
-double temp, fact = mrg31k3p_NORM_cl;
+double temp;
 const int startvalue = (get_global_id(0) + 
@@
 <<simStuff2, eval=TRUE,tidy=TRUE, echo=TRUE>>=
-sim_1 =  runifGpu(n = 8, streams = myStreamsGpu, Nglobal = c(2,2)))
+sim_1 =  runifGpu(n = 8, streams = myStreamsGpu, Nglobal = c(2,2))
 as.vector(sim_1)
@@
   get_global_id(1)) * NpadStreams;
-double temp;
 
@@
 
-for(Drow=get_global_id(0); Drow < Nrow;
-    Drow += get_global_size(0)){
+for(Drow=get_global_id(0); Drow < Nrow; Drow += get_global_size(0)){
+  for(Dcol=get_global_id(1); Dcol < Ncol; Dcol += get_global_size(1)){
 
-    for(Dcol=get_global_id(1); Dcol < Ncol; 
-        Dcol += get_global_size(1)){
+    part[get_local_id(1)] = fact[get_local_id(1)] * 
+      clrngMrg31k3pNextState(g1, g2);
+    barrier(CLK_LOCAL_MEM_FENCE);
+    out[Drow * Npad + Dcol] = sqrt( -2.0*log(part[0]) ) * 
+      cos(part[1] + addForSine[get_local_id(1)] );
+    barrier(CLK_LOCAL_MEM_FENCE);
 
-      part[get_local_id(1)] = fact[get_local_id(1)] * clrngMrg31k3pNextState(g1, g2);
-      barrier(CLK_LOCAL_MEM_FENCE);
-      temp = sqrt( -2.0*log(part[0]) ) * cos(part[1] + addForSine[get_local_id(1)] );
-      out[Dentry] = temp;
-      barrier(CLK_LOCAL_MEM_FENCE);
   }//Dcol
@@
 <<random exponential, eval=TRUE, cache=FALSE>>=
-r_matrix <- rexpGpu(c(2,4), rate=1, myStreamsGpu2, Nglobal=c(2,2), type="double")
+r_matrix <- rexpGpu(c(2,4), rate=1, myStreamsGpu, Nglobal=c(2,2), type="double")
 as.matrix(r_matrix)
@@
 
-\subsection{Comparing run-time: Month data example}\label{fisher_month}
-The 2-way contingency tables consist of selected data from the 2018 Natality public use data \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics (NCHS), the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
+\subsection{Comparing run-time: Monthly birth anomolies}\label{fisher_month}
+
+The 2-way contingency table in Table \ref{tab:monthdata} comes from the 2018 Natality public use data from the Centers for Disease Control and Prevention’s National Center for Health Statistics \citep{National2018}.
+The 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
 Table \ref{tab:month} is a $12 \times 12$ table that shows frequencies for congenital anomalies of the newborn by birth month in 2018 within the United States.
 The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
-The first 5 rows of the table are displayed below.
+
  <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
@@
 knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), #label="tab:month"
-  caption = "Monthly birth anomaly data\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
+  caption = "Monthly counts of birth anomalies.\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
   kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
 @
+
 <<monthDataForScript, include=FALSE, purl=TRUE>>=
@@
 
- <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE>>=
- streams <- createStreamsGpu(n =256*64)
- month_gpu<-vclMatrix(month,type="integer")
- system.time(result_month <- clrng::fisher.sim(month_gpu, 1e6, streams=streams,
-                    type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
- result_month$threshold
- result_month$simNum
- result_month$counts
+ <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE, tidy=FALSE>>=
+ streams <- createStreamsGpu(n = 256*64)
+ month_gpu <- vclMatrix(month, type = "integer")
+ system.time(result_month <- 
+   clrng::fisher.sim(month_gpu, 1e6, 
+      streams=streams, type="double", 
+      returnStatistics=TRUE,  Nglobal = c(256,64)))
+ @
+<<timecomparemonth1Gpuresult>>=
+unlist(result_month[c('threshold','simNum','counts')])
 result_month$p.value
@@
 
-\subsection{Comparing run-time: Week data example}\label{fisher_week}
+\subsection{Comparing run-time: Day-of-week birth anomolies}\label{fisher_week}
+
 Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
@@
 week_GPU<-gpuR::vclMatrix(week,type="integer")
-system.time(result_week<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
+system.time(resultWeek<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                 type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
-result_week$threshold
-result_week$simNum
-result_week$counts
-result_week$p.value
+unlist(resultWeek[c('threshold','simNum','counts')])
+resultWeek$p.value
 @
 The ``week'' table has a much smaller p-value: around 0.000125, which should require a larger number of simulations to get a more accurate p-value.
-With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}.
-\fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU.
+With more than ten million simulations, we get 1205 cases and a p-value around
+\Sexpr{resultWeek$p.value}
+with \fct{fisher.sim}. \fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU. 
 <<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
-result_week$cpu = as.vector(result_week$sim)
+resultWeek$cpu = as.vector(resultWeek$sim)
 @
@@
 
-hist(result_week$cpu, xlab='test statistic', ylab="proportion", prob=TRUE, main="")
-abline(v = result_week$threshold, col = "blue", lwd = 1.5)
+hist(resultWeek$cpu, xlab='test statistic', ylab="proportion", prob=TRUE, main="")
+abline(v = resultWeek$threshold, col = "blue", lwd = 1.5)
 @
 
-\section{An application of normal RNG: Gaussian surface simulation} 
-Given a parameter space $X$, a random field $U$ on $X$ is a collection of random variables $\{U_x :x\in X\}$.
-A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Gaussian.
+\section{Simulating Gaussian Random Fields}
+
+Simulating Gaussian random fields (GRF's) is computationally intensive due the high dimensionality of the problem, a 100 by 100 grid has 10,000 cells and a 10,000 by 10,000 variance matrix.
+Writing $x_1,\dots, x_n$ as point locations (i.e.\ centres of grid cells), the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Gaussian.
+
+GEOMETRIC ANISOTROPY
+
 The expectation and covariance function of $U$ completely determine the distribution of $U$, 
@@
 U = [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
-\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|x_i = x_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|x_i = x_j|}{\phi}\right),
+\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|x_i - x_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|x_i - x_j|}{\phi}\right),
 \end{gather*}
@@
 Nglobal = c(128, 64, 2)
-Nlocal = c(4, 2, 2)
+Nlocal = c(16, 8, 2)
 theType = "double"
@@
 data("swissRain", package="geostatsp")
-myRaster = geostatsp::squareRaster(swissBorder, 80)
-myRaster
+myRaster = geostatsp::squareRaster(swissBorder, 90)
+dim(myRaster)
 @
@@
 @
-<<show parameterBatch,eval=TRUE>>=
+<<show parameterBatch,eval=TRUE,R.options=list(width=100)>>=
 params
@@
 <<maternplot, echo=FALSE, eval=TRUE, cache=FALSE, dev='png', out.width = '0.47\\textwidth', fig.height=3, fig.width=4.5, fig.cap="Simulated Gaussian random fields", fig.subcap = theSubcap, fig.align='default'>>=
-myCol = mapmisc::colourScale(breaks=sort(unique(c(-6, -4, seq(-2, 2), 4, 6))), style='fixed', col='Spectral')
+myCol = mapmisc::colourScale(breaks=sort(unique(c(-7, -4, seq(-2, 2), 4, 7))), style='fixed', col='Spectral')
 for(D in names(simRaster)) {

### 2022-01-08 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1, ref timenumber anistr
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3102 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\fct{setBaseCreator} has one argument:
+\fct{setBaseCreator} has a single argument, the state of the first stream which defaults to a vector of 12345's.  
+The argument of \fct{createStreamsCpu} is the number of streams to create.
+Normally users would create many more than the four presented here, as most GPU's have thousands of work items.
+The default number of streams is 1024.
+
+We move streams to the GPU by converting them to a `\code{vclMatrix}'.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> myStreamsGpu = vclMatrix(myStreamsCpu)
+\end{CodeInput} 
+\end{CodeChunk}
+
+Equivalently, \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
+
+There are a few important notes on the usage of these above functions.
 \begin{itemize}
-\itemsep0em 
-  \item \code{initial}: a vector of six 31-bit integers, recycled if length is shorter. Default is set to \code{c(12345, 12345, 12345, 12345, 12345, 12345)}.
+\item Calling \fct{setBaseCreator} creates a hidden object \code{.Random.seed.clrng}, which (unlike the standard \code{.Random.seed}) controls the creation of new streams (not random number).
+We allow users to call \fct{setBaseCreator}  at any time.
+\fct{setBaseCreator} should be called once before \fct{createStreamsCpu} and \fct{createStreamsGpu} if they would like to select their own initial seeds, or never, in which case streams take the package's default initial seed.
+Otherwise, there is a small probability of producing streams which overlap.
+\item Users may switch between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program, calling either will read and update \code{.Random.seed.clrng}.
+This is not recommended as it causes unnecessary data transfer between host and device.
+\item If an \proglang{R} session is restarted during a task, as long as users saved the \proglang{R} environment (which Rstudio does automatically), the initial states (seeds) of the newly created streams will carry on from the last streams' seed from the previous \proglang{R} session.
+\item Objects on the GPU are lost when an R session is restarted, even if the workspace is saved.
+The object \code{myStreamsGpu} will remain in the workspace if the session is restarted, but the GPU memory containing \code{myStreamsGpu}'s data will no longer be accessible and an error will occur if the object is accessed.
+To retain streams when a session is restarted the streams should be copied to the CPU with \code{as.matrix}.
 \end{itemize}
-\fct{createStreamsCpu} has one argument:
-\begin{itemize}
-\itemsep0em 
-  \item \code{n}: number of streams to create.
-\end{itemize}
-We move streams to the GPU by converting them to a `\code{vclMatrix}'.
-\begin{CodeChunk}
-\begin{CodeInput}
-R> myStreamsGpu = vclMatrix(myStreamsCpu)
-\end{CodeInput} 
-\end{CodeChunk}
-
-Equivalently, \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
-\begin{CodeChunk}
-\begin{CodeInput}
-R> myStreamsGpu2 = createStreamsGpu(4)
-\end{CodeInput} 
-\end{CodeChunk}
-
-%Users should run either \fct{createStreamsCpu} or \fct{createStreamsGpu} at the beginning of a task and retain these streams for all of the random number generation the task requires.
-Creating new streams with different initial states introduces the very small possibility of overlap between the original and new streams.
-If an \proglang{R} session is restarted during a task, perhaps because a preliminary analysis is done following an initial set of simulations and the full set of simulations continued at a later date, the state of the streams should be transferred to the CPU and saved to disk.
-The `\code{initial}' argument to the stream creator should be used with caution and we recommend using the default.
-It bears repeating that two sets of streams with different initial states may produce overlapping numbers, the probability is small but increases when the number of streams is large.
-There are a few important notes on the usage of these above functions.
-(1), We allow users to call \fct{setBaseCreator}  at any time.
-\fct{setBaseCreator} should be called once before \fct{createStreamsCpu} and \fct{createStreamsGpu} if they would like to select their own initial seeds, or never, in which case streams take the package's default initial seed.
-(2), Users may switching between \fct{createStreamsCpu} and \fct{createStreamsGpu} in the same program, however, this is not recommended as it causes unnecessary data transfer between host and device.
-(3), If an \proglang{R} session is restarted during a task, as long as users saved the \proglang{R} environment, the initial states (seeds) of the newly created streams will carry on from the last streams' seed from the previous \proglang{R} session.
-
-%A newly created stream should not have the same seed as another existing stream, just like when you create a new car, it must have a distinct serial number.
-Streams should be created only once, and moved or copied if needed.
-And also it causes unnecessary data transfer between host and device if users switch often between \fct{createStreamsCpu} and \fct{createStreamsGpu}.
-
-%However, any data is lost when the \proglang{R} session is terminated.
-%\fct{createStreamsGpu} is adapted from the clrngMrg31k3pCreateStreams() and creates streams directly on a GPU device.
-Setting a seed same as the MRG31k3p default seed in \fct{createStreamsGpu}, we produce on device exactly the same streams as those created on host as follows.
-Notice that the code is just for showing \fct{createStreamsGpu} and \fct{createStreamsCpu} are \proglang{R} interface functions for \fct{clrngMrg31k3pCreateStreams} on GPU and CPU respectively, in practice it is not recommended to create a new stream that has the same initial seed as another existing stream in one simulation program.
-
-%also has the two arguments \code{n} and \code{initial}.
-\fct{createStreamsGpu} is similar to \code{.Random.seed} in \proglang{R}.
-By setting the RNG status with the \proglang{R} function \fct{set.seed}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs.
-Similarly in \fct{createStreamsGpu} and \fct{createStreamsCpu}, with the \code{initial} argument, we make the sequence of streams repeatable.
+
+Unlike \proglang{R}'s standard random number generators, the streams need to be explicitly specified when generating random numbers with \code{clrng}.
+It would be possible to hide streams from the user by storing them as a hidden object, which the \pkg{rlecuyer} package does to some extent, we have chosen not to do so for two reasons.
+First, the streams would need to be stored on the CPU and transferred to and from the GPU every time they were used.
+Second, computations might be distributed across multiple GPU devices or multiple computers, and being explicit about which streams are used where ensures reproducibility and avoids the possibility that the same streams might be used on different devices without the user being aware.
 
@@
 \begin{CodeInput}
-R> as.vector(runifGpu(n = 6, streams = myStreamsGpu2, Nglobal = c(2,
-+    2)))
-\end{CodeInput}
-\begin{CodeOutput}
-## [1] 0.993 0.279 0.401 0.468 0.510 0.290
+R> sim_1 = runifGpu(n = 8, streams = myStreamsGpu, Nglobal = c(2,
++    2))
+R> as.vector(sim_1)
+\end{CodeInput}
+\begin{CodeOutput}
+## [1] 0.735 0.842 0.614 0.216 0.110 0.870 0.649 0.170
 \end{CodeOutput} 
@@
 \itemsep0em 
-  \item \code{n}: A vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
-  \item \code{streams}: Streams for random number generation, streams cannot be missing.
-  \item \code{Nglobal}: Global index space. Default is set as $(64,8)$.
-  \item \code{type}: ``double'' or ``float'' or ``integer'' format of generated random numbers.
-  \item \code{verbose}: Print extra information if $\text{verbose} > 1$. Default is FALSE.
+  \item \code{n}: Either a scalar giving the number of samples to generate or a vector of length 2 specifying the size of a matrix to be output.
+  \item \code{streams}: Streams for random number generation, which must be stored on the GPU as a \code{vclMatrix} object.
+  \item \code{Nglobal}: Number of work items, by default it is set as $(64,8)$.
+  \item \code{type}: \code{double}, \code{float} or \code{integer}, the format of generated random numbers, defaults to \code{double}.
+  \item \code{verbose}: Print extra information if $\text{verbose} > 1$.
 \end{itemize}
-%\fct{runifGpu} returns the same sequence of random numbers because it uses the same streams \code{myStreamsGpu} and \code{myStreamsGpu2} here. 
-Reusing \code{myStreamsGpu} will produce a vector different from \code{sim_1}.
-Because each time a random number is generated, the current state of the stream advances by one position.
-%Below shows the \code{myStreamsGpu} after it being used.
-
-Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart.
-Streams created on GPU does not remain in the memory when a new \proglang{R} session begins.
+
+The object returned is either a \code{vclMatrix} or \code{vclVector}, depending on whether \code{n} is two- or one-dimensional.
+Reusing \code{myStreamsGpu} will produce a vector different form \code{sim_1}, as the current states of the streams in \code{myStreamsGpu} have advanced by two positions (each of the 4 work item generated 2 random numbers).
+
+As mentioned earlier, streams on the GPU do not remain in the memory when an \proglang{R} session is restarted.
 There are two ways to reproduce results, depending on how the program calls streams creator.
-The simpler way to reproduce is just to restart the program with the same initial seed for the creator.
+The simpler way to reproduce is just to have a single program in an \proglang{R} script file with the initial seed for the creator set at the beginning.
 But it is important that the program always creates the streams in exactly the same order, for example, like the toy \proglang{R} program for demonstration below, we can reproduce the random matrix \code{sim_mat} just by keeping a record of the initial seed \code{c(11,22,33,44,55,66)}.
@@
 
+Were we to replace the \code{for} loop above with a parallel equivalent (i.e.\ \code{mcmapply}), the order of stream creation could change with each program execution and the result would not be reproducible.
 For more complicated applications, we recommend users to save the matrix of streams (current states and initial states) on the CPU in a file as a \code{.rds} object.
@@
 \begin{CodeInput}
-R> saveRDS(as.matrix(myStreamsGpu2), "myStreams.rds")
+R> saveRDS(as.matrix(myStreamsGpu), "myStreams.rds")
 R> # Load the streams object as streams_saved
@@
 \end{CodeChunk}
-
-%\subsection{Plot the uniform random numbers (now deleted?)}
 
@@
 uint g1[3], g2[3];
-local double part[2];
 int startvalue = (get_global_id(0) * get_global_size(1) + 
@@
 
-double sinOrCosPart1, addForSine = - get_local_id(1) * PI_2;
-double temp, fact = mrg31k3p_NORM_cl;
-if(get_local_id(1)){
-  fact = TWOPI_mrg31k3p_NORM_cl;
-}
+const double fact[2] = { mrg31k3p_NORM_cl, TWOPI * mrg31k3p_NORM_cl };
+const double addForSine[2] = { 0.0, - PI_2 };
+local double part[2];
 
@@
 
-for(Drow=get_global_id(0); Drow < Nrow;
-    Drow += get_global_size(0)){
-
-    for(Dcol=get_global_id(1); DcolBlock < Ncol; 
-        Dcol += get_global_size(1)){
-
-      temp = fact * clrngMrg31k3pNextState(g1, g2);
-      part[get_local_id(1)] = temp;
-
-      if(!get_local_id(1)) {
-        part[0] = sqrt(-2.0*log(part[0]));
-      }
+for(Drow=get_global_id(0); Drow < Nrow; Drow += get_global_size(0)){
+  for(Dcol=get_global_id(1); Dcol < Ncol; Dcol += get_global_size(1)){
+
+    part[get_local_id(1)] = fact[get_local_id(1)] * 
+      clrngMrg31k3pNextState(g1, g2);
     barrier(CLK_LOCAL_MEM_FENCE);
-
-      // sine for local[1], cos for local[0]
-      sinOrCosPart1 = cos(part[1] + addForSine);
-      out[Drow * Npad + Dcol] = part[0]*sinOrCosPart1;
-
+    out[Drow * Npad + Dcol] = sqrt( -2.0*log(part[0]) ) * 
+      cos(part[1] + addForSine[get_local_id(1)] );
     barrier(CLK_LOCAL_MEM_FENCE);
+
   }//Dcol
@@
 
-We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}.
-The best performance we have seen using \fct{rnormGpu} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}.
-The difference in elapsed time becomes larger when matrix size goes larger.
+Below we generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}.
+The best performance we have seen using \fct{rnormGpu} is above over 100 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}.
+The difference in elapsed time becomes larger when matrix size increases.
 \begin{CodeChunk}
@@
 ##    user  system elapsed 
-##   0.051   0.000   0.050
+##   0.046   0.004   0.050
 \end{CodeOutput} 
@@
 \begin{CodeInput}
-R> r_matrix <- rexpGpu(c(2, 4), rate = 1, myStreamsGpu2, Nglobal = c(2,
+R> r_matrix <- rexpGpu(c(2, 4), rate = 1, myStreamsGpu, Nglobal = c(2,
 +    2), type = "double")
@@
 ##      [,1]  [,2]  [,3]  [,4]
-## [1,] 3.97 7.068 1.54 0.23
-## [2,] 1.11 0.247 2.05 1.33
+## [1,] 1.00 0.689 2.218 1.167
+## [2,] 1.48 1.205 0.406 0.241
 \end{CodeOutput} 
@@
 
-\subsection{Comparing run-time: Month data example}\label{fisher_month}
-The 2-way contingency tables consist of selected data from the 2018 Natality public use data \citep{National2018} from the Centers for Disease Control and Prevention’s National Center for Health Statistics (NCHS), the 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
+\subsection{Comparing run-time: Monthly birth anomolies}\label{fisher_month}
+
+The 2-way contingency table in Table \ref{tab:monthdata} comes from the 2018 Natality public use data from the Centers for Disease Control and Prevention’s National Center for Health Statistics \citep{National2018}.
+The 2018 natality data file may be downloaded at \url{https://www.cdc.gov/nchs/data_access/VitalStatsOnline.htm}.
 Table \ref{tab:month} is a $12 \times 12$ table that shows frequencies for congenital anomalies of the newborn by birth month in 2018 within the United States.
 The column variables of these two tables represent the twelve categories of congenital anomalies of the newborn: 1) Anencephaly; 2) Meningomyelocele/Spina bifida; 3) Cyanotic congenital heart disease; 4) Congenital diaphragmatic hernia; 5) Omphalocele; 6) Gastrochisis; 7) Limb reduction defect; 8) Cleft lip with or without cleft palate; 9) Cleft palate alone; 10) Down syndrome; 11) Suspected chromosomal disorder; and 12) Hypospadias.
-The first 5 rows of the table are displayed below.
 
@@
 
-\caption{\label{tab:monthdata}Monthly birth anomaly data\label{tab:month}}
+\caption{\label{tab:monthdata}Monthly counts of birth anomalies.\label{tab:month}}
 \centering
@@
 \begin{CodeInput}
-R> streams <- createStreamsGpu(n = 256 * 64)
+R> streams <- createStreamsGpu(n = 256*64)
 R> month_gpu <- vclMatrix(month, type = "integer")
-R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1000000,
-+    streams = streams, type = "double", returnStatistics = TRUE,
-+    Nglobal = c(256, 64)))
+R> system.time(result_month <- 
++  clrng::fisher.sim(month_gpu, 1e6, 
++     streams=streams, type="double", 
++     returnStatistics=TRUE,  Nglobal = c(256,64)))
 \end{CodeInput}
@@
 ##    user  system elapsed 
-##   0.323   0.000   0.324
-\end{CodeOutput}
-\begin{CodeInput}
-R> result_month$threshold
-\end{CodeInput}
-\begin{CodeOutput}
-## [1] -47955
-\end{CodeOutput}
-\begin{CodeInput}
-R> result_month$simNum
-\end{CodeInput}
-\begin{CodeOutput}
-## [1] 1015808
-\end{CodeOutput}
-\begin{CodeInput}
-R> result_month$counts
-\end{CodeInput}
-\begin{CodeOutput}
-## [1] 409435
+##   0.316   0.008   0.323
+\end{CodeOutput} 
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> unlist(result_month[c("threshold", "simNum", "counts")])
+\end{CodeInput}
+\begin{CodeOutput}
+## threshold    simNum    counts 
+##    -47955   1015808    409429
 \end{CodeOutput}
@@
 
-\subsection{Comparing run-time: Week data example}\label{fisher_week}
+\subsection{Comparing run-time: Day-of-week birth anomolies}\label{fisher_week}
+
 Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States.
@@
 R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
-R> system.time(result_week <- clrng::fisher.sim(week_GPU, 10000000,
+R> system.time(resultWeek <- clrng::fisher.sim(week_GPU, 10000000,
 +    streams = streams, type = "double", returnStatistics = TRUE,
@@
 ##    user  system elapsed 
-##   1.942   0.014   1.954
-\end{CodeOutput}
-\begin{CodeInput}
-R> result_week$threshold
-\end{CodeInput}
-\begin{CodeOutput}
-## [1] -54990
-\end{CodeOutput}
-\begin{CodeInput}
-R> result_week$simNum
-\end{CodeInput}
-\begin{CodeOutput}
-## [1] 10010624
-\end{CodeOutput}
-\begin{CodeInput}
-R> result_week$counts
-\end{CodeInput}
-\begin{CodeOutput}
-## [1] 1351
-\end{CodeOutput}
-\begin{CodeInput}
-R> result_week$p.value
+##    1.94    0.01    1.95
+\end{CodeOutput}
+\begin{CodeInput}
+R> unlist(resultWeek[c("threshold", "simNum", "counts")])
+\end{CodeInput}
+\begin{CodeOutput}
+## threshold    simNum    counts 
+##    -54990  10010624      1350
+\end{CodeOutput}
+\begin{CodeInput}
+R> resultWeek$p.value
 \end{CodeInput}
@@
 The ``week'' table has a much smaller p-value: around 0.000125, which should require a larger number of simulations to get a more accurate p-value.
-With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}.
-\fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU.
+With more than ten million simulations, we get 1205 cases and a p-value around
+0
+with \fct{fisher.sim}. \fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU.
 
@@
 \hline
-\hspace{1em}1M & 10.74 & 2.28 & 14.8 & 0.32 & month\\
+\hspace{1em}1M & 49.33 & 2.15 & 14.8 & 0.32 & month\\
 \hline
-\hspace{1em}10M & 63.24 & 10.82 & 92 & 1.95 & week\\
+\hspace{1em}10M & 327.27 & 10.23 & 92 & 1.95 & week\\
 \hline
@@
 
-\section{An application of normal RNG: Gaussian surface simulation} 
-Given a parameter space $X$, a random field $U$ on $X$ is a collection of random variables $\{U_x :x\in X\}$.
-A Gaussian random field is a random field with the property that for any positive integer $n$ and any set of locations $x_1,\dots,x_n \in X$, the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Gaussian.
+\section{Simulating Gaussian Random Fields}
+
+Simulating Gaussian random fields (GRF's) is computationally intensive due the high dimensionality of the problem, a 100 by 100 grid has 10,000 cells and a 10,000 by 10,000 variance matrix.
+Writing $x_1,\dots, x_n$ as point locations (i.e.\ centres of grid cells), the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Gaussian.
+
+GEOMETRIC ANISOTROPY
+Geometrical anisotropy is defined by two additional parameters in the correlation function.
+Let $(x_1, x_2)$ represent the original coordinates, which is rotated counterclockwise by $\varphi$, then the new coordinates $(x'_1, x'_2)$ is given by
+%
+  \begin{equation*}\label{eq:matrixeqn}
+    \begin{pmatrix}
+      x'_{1} \\
+      x'_{2}  
+    \end{pmatrix}
+    =
+    \begin{pmatrix}
+      1/\phi_{1} & 0 \\
+      0 & 1/ \phi_{2} 
+    \end{pmatrix}
+    \cdot
+    \begin{pmatrix}
+      \cos \varphi & -\sin \varphi \\
+      \sin \varphi & \cos \varphi 
+    \end{pmatrix}
+    \cdot
+      \begin{pmatrix}
+      x_1 \\
+      x_2 
+    \end{pmatrix},
+  \end{equation*}
+
 The expectation and covariance function of $U$ completely determine the distribution of $U$, 
@@
 R> Nglobal = c(128, 64, 2)
-R> Nlocal = c(4, 2, 2)
+R> Nlocal = c(16, 8, 2)
 R> theType = "double"
@@
 R> data("swissRain", package = "geostatsp")
-R> myRaster = geostatsp::squareRaster(swissBorder, 80)
-R> myRaster
-\end{CodeInput}
-\begin{CodeOutput}
-## class      : RasterLayer 
-## dimensions : 51, 80, 4080  (nrow, ncol, ncell)
-## resolution : 4358, 4358  (x, y)
-## extent     : 2485351, 2833996, 1075013, 1297275  (xmin, xmax, ymin, ymax)
-## crs        : +proj=somerc +lat_0=46.9524055555556 +lon_0=7.43958333333333 +k_0=1 +x_0=2600000 +y_0=1200000 +ellps=GRS80 +units=m +no_defs
+R> myRaster = geostatsp::squareRaster(swissBorder, 90)
+R> dim(myRaster)
+\end{CodeInput}
+\begin{CodeOutput}
+## [1] 57 90  1
 \end{CodeOutput} 
@@
 \begin{CodeOutput}
-##      shape range variance nugget anisoRatio
-## [1,]  1.25 50000      1.5      0          1
-## [2,]  2.15 60000      2.0      0          4
-## [3,]  0.60 30000      2.0      0          2
-## [4,]  3.00 30000      2.0      0          2
-##      anisoAngleRadians
-## [1,]             0.000
-## [2,]             0.449
-## [3,]             0.449
-## [4,]             0.449
+##      shape range variance nugget anisoRatio anisoAngleRadians
+## [1,]  1.25 50000      1.5      0          1             0.000
+## [2,]  2.15 60000      2.0      0          4             0.449
+## [3,]  0.60 30000      2.0      0          2             0.449
+## [4,]  3.00 30000      2.0      0          2             0.449
 \end{CodeOutput} 
@@
 \begin{CodeOutput}
-## [1] 16320  4080
+## [1] 20520  5130
 \end{CodeOutput} 

### 2022-01-08 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1, ref timenumber anistr
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3102 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   var1 = c("B", "1M", "10M", "1M", "10M"),
-  var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 10.74,  63.24),
-  var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.28,    10.82),
+  var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 49.33,  327.27),
+  var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.15,    10.23),
   var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.80,  92.00),
@@
 GEOMETRIC ANISOTROPY
+Geometrical anisotropy is defined by two additional parameters in the correlation function.
+Let $(x_1, x_2)$ represent the original coordinates, which is rotated counterclockwise by $\varphi$, then the new coordinates $(x'_1, x'_2)$ is given by
+%
+  \begin{equation*}\label{eq:matrixeqn}
+    \begin{pmatrix}
+      x'_{1} \\
+      x'_{2}  
+    \end{pmatrix}
+    =
+    \begin{pmatrix}
+      1/\phi_{1} & 0 \\
+      0 & 1/ \phi_{2} 
+    \end{pmatrix}
+    \cdot
+    \begin{pmatrix}
+      \cos \varphi & -\sin \varphi \\
+      \sin \varphi & \cos \varphi 
+    \end{pmatrix}
+    \cdot
+      \begin{pmatrix}
+      x_1 \\
+      x_2 
+    \end{pmatrix},
+  \end{equation*}
 

### 2022-01-08 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3103 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2022-01-05 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
@@
 options(continue="+  ",  prompt="R> ", digits=3, width=66) 
-
+options(scipen=999)
 opts_chunk$set(fig.height=3,
@@
 \section{Introduction}
+
 some keypoints to mention:\\
@@
 \pkg{geostatsp} calculates Wald confidence intervals for model parameters in \fct{lgm}.
-The \fct{numderiv} function inside \fct{lgm} returns the Hessian matrix evaluated at $\hat{\omega}$ and $\hat{\lambda}$, the diagonals of the inverse of the negative Hessian is used for approximating the variance of the parameters.This way does not take into account of uncertainty in correlation parameters.
+The \fct{numderiv} function inside \fct{lgm} returns the Hessian matrix evaluated at $\hat{\omega}$ and $\hat{\lambda}$, the diagonals of the inverse of the negative Hessian is used for approximating the variance of the parameters.
+In this procedure, the uncertainty of the (estimated) correlation parameters is ignored .
 \pkg{geostatsp} does the above computation process in \prog{C}, and (ADD LATER!) which has achieved a great perfomance in terms of speed.
@@
 
-\subsection{CI’s comparison for model parameters}
+<<packages, echo=FALSE,results='hide', message=FALSE>>=
+ library(gpuRandom)
+ library(gpuR)
+ library(data.table)
+ library(geostatsp)
+ library("sp")
+library(knitr)
+library(kableExtra)
+@
+
+\subsection{swissRain example}
+<<swissgeostatsp, echo=FALSE, cache=TRUE>>=
+data("swissRain")
+swissRain$elevation = extract(swissAltitude, swissRain)
+swissRes =  lgm( formula=rain~ elevation, 
+                     data=swissRain, 
+                     grid=20,
+                     covariates=swissAltitude, 
+                     reml = FALSE,
+                     fixBoxcox=FALSE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
+                     aniso=TRUE )
+A<-swissRes$summary[,c('estimate','ci0.1', 'ci0.9')]
+@
+
+<<swisssummarycompare, echo=FALSE>>=
+table1 <- read.csv("Book1.csv")
+colnames(table1) <- c('Parameter','Estimate','95% CI', 'Estimate','95% CI')
+knitr::kable(table1, format = "latex", align = c("rcccc"),booktabs = T, linesep = "", caption= "Summary comparison", digits=3)%>%
+ kable_styling(full_width = FALSE, position = "center", latex_options = "HOLD_position") %>%
+ add_header_above(c(" ", "geostatsp" = 2, "gpuBatchMatrix" = 2)) 
+@
+
+<<swiss_profile_many,echo = FALSE, eval=TRUE, cache=TRUE, fig.width=4,fig.height=3, fig.show ="hold", fig.align='center', fig.cap = "profile LogL of slope parameter", out.width = "50%",dev='pdf',fig.pos="H">>=
+setwd("/home/ruoyong/diseasemapping/pkg/oldgpuRandom/inst/doc/L2021_11")
+knitr::include_graphics(c('profile_logl_swiss.png'))
+@
+
+<<swiss_profile_sole_prep, out.width = "100%", fig.align = "center",echo=FALSE>>=
+knitr::include_graphics(c('swiss_pro_4.png'))
+@
+
+<<range>>=
+
+@
+
+\subsection{meuse example}
 
@@
 
-<<packages, results='hide', message=FALSE>>=
- library(gpuRandom)
- library(gpuR)
- library(data.table)
- #library(geostatsp)
+Estimates by \pkg{geostatsp}
+<<meusedata, echo=FALSE, cache=TRUE>>=
+data(meuse)
+coordinates(meuse) <- c("x", "y")
+meuse$Y = log(meuse$zinc)
+meuse$sqrtdist = sqrt(meuse$dist)
+
+Res1 =  geostatsp::lgm( formula=Y~sqrtdist, 
+                        data=meuse, 
+                        grid=20,
+                        reml = FALSE,
+                        fixBoxcox=FALSE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
+                        aniso=TRUE )
+Res1$summary[,c('estimate','ci0.005', 'ci0.995')]
+Res1$optim$logL
 @
 
-% swissrain matplot over betas
+<<meuse_params, cache=TRUE, echo=FALSE>>=
+  ParamList = list(
+    range = c(exp(seq(log(100), log(173.1643928), len=6)), exp(seq(log(185), log(330), len=8))),
+    shape = c(seq(0.1,3.5, len=4), seq(4, 6, len=4)),
+    anisoRatio =c( seq(1, 2.301060332, len=3), seq(2.8, 7, len=4)),
+    anisoAngleRadians = c( seq(0, 0.274896758, len=4), seq(0.4, 0.8, len=4)),
+    nugget = c(seq(0,0.3946587, len=3), seq(0.5, 3, len=5))
+  ) 
+  params = do.call(expand.grid, ParamList)
+@
 
-%meuse data
+<<meuse_logl, cache=TRUE, message=FALSE>>=
+result3<- gpuRandom::likfitLgmCov(spatialmodel=Res1,
+                  params=params, 
+                  boxcox=c(1,0,-0.903901505),  
+                  paramToEstimate = c('range','shape','nugget','sdNugget', 'anisoRatio','anisoAngleRadians'), 
+                  cilevel=0.95, 
+                  type = "double",
+                  reml=FALSE, 
+                  NparamPerIter=400,
+                  Nglobal=c(128,64),
+                  Nlocal=c(16,16),
+                  NlocalCache=2800,
+                  verbose=FALSE)
+result3$summary
+@
 
 \section{A simulation study}
+<<simulate_datasets, echo=FALSE, cache=TRUE, results='hide', message=FALSE>>=
+############### A SIMULATION STUDY IN THE PAPER 2 ####################################
+Nsim=100
+set.seed(88)
+mydat = SpatialPointsDataFrame(cbind(seq(100*1e3,140*1e3, len=500), seq(100*1e3,120*1e3, len=500)), 
+                               data=data.frame(cov1 =stats::rnorm(500, 120, 20), cov2 = stats::rpois(500, 50))
+)
+
+## simulate a random field
+trueParam = c(range=1000, shape=2, variance=0.5^2, tauSq=1^2, anisoRatio=1, anisoAngleRadians=0)
+nugget = trueParam['tauSq']/trueParam['variance']
+names(nugget) <- "nugget"
+nugget   #4
+boxcox = c(boxcox=1)
+options("useRandomFields" = FALSE)
+for (i in 1:Nsim){
+  set.seed(i*6-1)
+  U <- geostatsp::RFsimulate(model=trueParam[setdiff(names(trueParam),'tauSq')],x=mydat)@data
+  # simulate response
+  Y <- 3 + 1*mydat$cov1 + 0.5*mydat$cov2 + U + stats::rnorm(length(mydat), 0, sd=sqrt(trueParam["tauSq"]))
+  mydat@data <- cbind(mydat@data,Y)
+}
+colnames(mydat@data) <- c("cov1", "cov2", paste("Y", c(1:Nsim), sep=""))
+
+## geostatsp's estimates
+swissRes =  lgm( formula=Y10~ cov1 + cov2,
+                 data=mydat,
+                 grid=20,
+                 reml = FALSE,
+                 boxcox=1,
+                 fixBoxcox=TRUE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
+                 aniso=FALSE )
+swissRes$summary[,c('estimate','ci0.005', 'ci0.995')]
+swissRes$optim$mle
+@
+
+<<ranuparams, echo=FALSE, cache=TRUE>>=
+newParamListrn = list(
+  range = c(exp(seq(log(0.3), log(1), len=10))*1000, exp(seq(log(1.2), log(2.9), len=12))*1000),  
+  nugget = c(seq(0, 4, len=8), seq(4.5, 20, len=12)),  
+  shape = 2,
+  anisoRatio =1,
+  anisoAngleRadians = 0.0
+) 
+paramsrn = do.call(expand.grid, newParamListrn)
+@
+
+<<run_forloop, echo=FALSE, results='hide', message=FALSE, warning=FALSE, cache=TRUE>>=
+count = 0
+breaksf = rep(0, 10)
+makedata <- rep(0, 4)
+estimates <- rep(0, 6)
+## calculate logl
+for (i in 1:Nsim){
+  result<-gpuRandom::likfitLgmCov2d(
+    data= mydat,
+    formula= as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")), 
+    coordinates=mydat@coords,
+    params=paramsrn, 
+    paramToEstimate = c('range','nugget'),
+    boxcox = 1,
+    cilevel=0.8,
+    type = "double",
+    reml=FALSE, 
+    NparamPerIter=400,
+    Nglobal=c(128,64),
+    Nlocal=c(16,16),
+    NlocalCache=2800,
+    verbose=FALSE)
+
+  a <- geostatsp::loglikLgm(
+    c(trueParam['range'],
+      trueParam['shape'],
+      nugget,
+      trueParam[c('anisoRatio')],
+      trueParam[c('anisoAngleRadians')],
+      boxcox),
+    data = mydat,
+    formula = as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")),
+    reml = FALSE,
+    minustwotimes=FALSE)[['logLik']]
+
+  if (a  >= result$breaks){
+    count = count +1
+  }
+  breaksf[i] = result$breaks
+  makedata0 <- cbind(paramsrn[,1:2], result$LogLik, D=i)
+  makedata <- rbind(makedata, makedata0)
+  estimates <- cbind(estimates, result$summary)
+}
+colnames(makedata) <- c("range",'nugget', 'LogLik', 'D')
+makedatahey <- makedata[-1,]
+estimates_summary <- estimates[,-1]
+mean_estimates <- apply(estimates_summary, 1, mean)
+@
+
+<<results, cache=TRUE>>=
+count/Nsim
+mean_estimates
+@
+
+<<graphs, echo = FALSE, eval=TRUE, cache=TRUE>>=
+library(RColorBrewer)
+library('ggplot2')
+forigin <- ggplot(subset(makedatahey, D==1), aes(x = range, y = nugget, z=LogLik)) + theme_bw() + 
+  scale_y_continuous(limits=c(0,20))+
+  scale_x_continuous(limits=c(300,3000) )+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  geom_point(size = 0.5, colour='grey') + 
+  stat_contour(breaks=breaksf[1])  + 
+  geom_vline(xintercept = trueParam['range'])+
+  geom_hline(yintercept=nugget)
+cols <- rainbow(Nsim)
+for (i in 2:Nsim){
+  forigin <- forigin + geom_contour(data = subset(makedatahey, D==i), aes(x = range, y = nugget, z=LogLik), breaks=breaksf[i], col= cols[i])
+}
+@
+
+<<showgraphs, echo = FALSE, eval=TRUE, cache=TRUE, fig.width=4,fig.height=3, fig.show ="hold", fig.align='center', fig.cap = "range and nugget contours", out.width = "50%",dev='pdf'>>=
+#par(mfrow = c(1, 2))
+#forigin
+knitr::include_graphics(c('p2Rplot_100.png','rsplot.png'))
+@
 

### 2022-01-08 pkg/oldgpuRandom/inst/doc/L2021_11/simulation_ratio_radians.r

@@ -1,0 +1,85 @@
+########################################################################################################################################
+## ---gpuRandom------ratio vs radians contour--------------------------------------------------------------
+## set params
+newParamListsn = list(
+  anisoRatio =c(seq(0.1, 1, len=2), seq(2,9, len=8)),
+  anisoAngleRadians = seq(-1,1,len=9),
+  nugget = 4,  
+  shape = 2,
+  range = 1000
+) 
+paramssn = do.call(expand.grid, newParamListsn)
+
+count = 0
+breaksf = rep(0, 10)
+makedata <- rep(0, 4)
+estimates <- rep(0, 6)
+## calculate logl
+for (i in 1:15){
+  result<-gpuRandom::likfitLgmCov2d(
+    data= mydat,
+    formula= as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")), 
+    coordinates=mydat@coords,
+    params=paramssn, 
+    paramToEstimate = c('anisoRatio','anisoAngleRadians'),
+    boxcox = 1,
+    cilevel=0.8,
+    type = "double",
+    reml=FALSE, 
+    NparamPerIter=400,
+    Nglobal=c(128,64),
+    Nlocal=c(16,16),
+    NlocalCache=2800,
+    verbose=FALSE)
+  
+  a <- geostatsp::loglikLgm(
+    c(trueParam['range'],
+      trueParam['shape'],
+      nugget,
+      trueParam[c('anisoRatio')],
+      trueParam[c('anisoAngleRadians')],
+      boxcox),
+    data = mydat,
+    formula = as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")),
+    reml = FALSE,
+    minustwotimes=FALSE)[['logLik']]
+  
+  if (a  >= result$breaks){
+    count = count +1
+  }
+  breaksf[i] = result$breaks
+  makedata0 <- cbind(paramssn[,1:2], result$LogLik, D=i)
+  makedata <- rbind(makedata, makedata0)
+  estimates <- cbind(estimates, result$summary)
+}
+colnames(makedata) <- c('anisoRatio','anisoAngleRadians', 'LogLik', 'D')
+makedatahey <- makedata[-1,]
+estimates_summary <- estimates[,-1]
+mean_estimates <- apply(estimates_summary, 1, mean)
+count/15
+mean_estimates
+
+
[…]

### 2022-01-08 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.tex
paper 2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3097 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3103 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\subsection{CI’s comparison for model parameters}
+\subsection{swissRain example}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
+
+\caption{\label{tab:swisssummarycompare}Summary comparison}
+\centering
+\begin{tabular}[t]{rcccc}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{geostatsp} & \multicolumn{2}{c}{gpuBatchMatrix} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+Parameter & Estimate & 95\% CI & Estimate & 95\% CI\\
+\midrule
+range & 38618.730 & (21805.93, 68394.54) & 38618.730 & NA\\
+sdNugget & 0.992 & (0.684,1.439) & 0.992 & NA\\
+shape & 1.826 & (-0.574,4.227) & 1.826 & NA\\
+anisoRatio & 8.092 & (5.321,12.305) & 8.092 & NA\\
+anisoAngleRadians & 0.652 & (0.595, 0.709) & 0.652 & NA\\
+boxcox & 0.496 & (0.373,0.620) & 0.496 & NA\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{figure}[H]
+
+{\centering \includegraphics[width=0.5\linewidth]{profile_logl_swiss}
+
+}
+
+\caption[profile LogL of slope parameter]{profile LogL of slope parameter}\label{fig:swiss_profile_many}
+\end{figure}
+
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
+
+{\centering \includegraphics[width=1\linewidth]{swiss_pro_4}
+
+}
+
+\end{knitrout}
+
+\subsection{meuse example}
 
@@
 
+Estimates by \pkg{geostatsp}
 \begin{knitrout}
@@
 \begin{verbatim}
-R> library(gpuRandom)
-R> library(gpuR)
-R> library(data.table)
-R> # library(geostatsp)
+##                    estimate   ci0.005 ci0.995
+## (Intercept)         0.91634  0.121205   1.711
+## sqrtdist           -0.08445 -1.623508   1.455
+## sdNugget            0.00785  0.000379   0.163
+## sdSpatial           0.01250  0.000690   0.227
+## range             173.16439 92.775636 323.209
+## shape               4.00000 -9.880685  17.881
+## anisoRatio          2.30106  1.138810   4.649
+## anisoAngleRadians   0.27490 -0.167582   0.717
+## anisoAngleDegrees  15.75042 -9.601751  41.103
+## boxcox             -0.90390 -2.558961   0.751
+## m2logL.ml   logL.ml 
+##     133.4     -66.7
 \end{verbatim}
@@
 
-% swissrain matplot over betas
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+R> result3 <- gpuRandom::likfitLgmCov(spatialmodel = Res1, params = params,
++      boxcox = c(1, 0, -0.903901505), paramToEstimate = c("range", "shape",
++          "nugget", "sdNugget", "anisoRatio", "anisoAngleRadians"), cilevel = 0.95,
++      type = "double", reml = FALSE, NparamPerIter = 400, Nglobal = c(128,
++          64), Nlocal = c(16, 16), NlocalCache = 2800, verbose = FALSE)
+## workgroupSize
+## 128 64
+## localSize
+## 16 16
+## 
+##  Nparams 50176 NparamsPerIter 400 Niter 126 Ncovariates 2 Ndatasets 3 NlocalCacheD 1272
+\end{verbatim}
+\end{kframe}
 
-%meuse data
+{\centering \includegraphics[width=\maxwidth]{figure/meuse_logl-1}
+
+}
+
+\begin{kframe}\begin{verbatim}
+R> result3$summary
+##                   estimate lower95ci upper95ci
+## (Intercept)         0.9162        NA        NA
+## sqrtdist           -0.0843        NA        NA
+## sdSpatial           0.0126        NA        NA
+## range             173.1644   100.000  277.7688
+## shape               6.0000     0.784    6.0000
+## nugget              0.3947     0.000    1.1582
+## sdNugget            0.0079     0.000    0.0135
+## anisoRatio          2.3011     1.469    5.5220
+## anisoAngleRadians   0.2749     0.000    0.5604
+## boxcox             -0.9039        NA        NA
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 \section{A simulation study}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+R> count/Nsim
+## [1] 0.81
+R> mean_estimates
+## (Intercept)        cov1        cov2   sdSpatial       range 
+##       2.992       1.000       0.500       0.492     986.372 
+##      nugget 
+##       4.427
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{p2Rplot_100} 
+\includegraphics[width=0.5\linewidth]{rsplot}
+
+}
+
+\caption[range and nugget contours]{range and nugget contours}\label{fig:showgraphs}
+\end{figure}
+
+\end{knitrout}
 

### 2022-01-08 pkg/oldgpuRandom/inst/doc/L2021_11/shape_radians.r

@@ -1,0 +1,78 @@
+########################################################################################################################################
+## ---gpuRandom------shape vs radians contour--------------------------------------------------------------
+## set params
+newParamListsn = list(
+  shape = c(0.1, 0.15, 0.2, 0.23, 0.3, 0.4, 0.45, 0.5, 0.75, 0.9, 1, 2, 3, 5, 6),
+  anisoAngleRadians = seq(-1,1,len=9),
+  anisoRatio =1,
+  nugget = 4,  
+  range = 1000
+) 
+paramssn = do.call(expand.grid, newParamListsn)
+
+count = 0
+breaksf = rep(0, 10)
+makedata <- rep(0, 4)
+estimates <- rep(0, 6)
+## calculate logl
+for (i in 1:15){
+  result<-gpuRandom::likfitLgmCov2d(
+    data= mydat,
+    formula= as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")), 
+    coordinates=mydat@coords,
+    params=paramssn, 
+    paramToEstimate = c('shape','anisoAngleRadians'),
+    boxcox = 1,
+    cilevel=0.8,
+    type = "double",
+    reml=FALSE, 
+    NparamPerIter=400,
+    Nglobal=c(128,64),
+    Nlocal=c(16,16),
+    NlocalCache=2800,
+    verbose=FALSE)
+  
+  a <- geostatsp::loglikLgm(
+    c(trueParam['range'],
+      trueParam['shape'],
+      nugget,
+      trueParam[c('anisoRatio')],
+      trueParam[c('anisoAngleRadians')],
+      boxcox),
+    data = mydat,
+    formula = as.formula(paste(colnames(mydat@data)[i+2], "~", "cov1 + cov2")),
+    reml = FALSE,
+    minustwotimes=FALSE)[['logLik']]
+  
+  if (a  >= result$breaks){
+    count = count +1
+  }
+  breaksf[i] = result$breaks
+  makedata0 <- cbind(paramssn[,1:2], result$LogLik, D=i)
+  makedata <- rbind(makedata, makedata0)
+  estimates <- cbind(estimates, result$summary)
+}
+colnames(makedata) <- c('shape',"anisoAngleRadians", 'LogLik', 'D')
+makedatahey <- makedata[-1,]
+estimates_summary <- estimates[,-1]
[…]

### 2022-01-09 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3105 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 table1 <- read.csv("Book1.csv")
-colnames(table1) <- c('Parameter','Estimate','95% CI', 'Estimate','95% CI')
+colnames(table1) <- c('Parameter','Estimate','90% CI', 'Estimate','90% CI')
 knitr::kable(table1, format = "latex", align = c("rcccc"),booktabs = T, linesep = "", caption= "Summary comparison", digits=3)%>%
@@
 setwd("/home/ruoyong/diseasemapping/pkg/oldgpuRandom/inst/doc/L2021_11")
-knitr::include_graphics(c('profile_logl_swiss.png'))
+knitr::include_graphics(c('profile_logl_swiss.png','betaprofile.png'))
 @
 
-<<swiss_profile_sole_prep, out.width = "100%", fig.align = "center",echo=FALSE>>=
-knitr::include_graphics(c('swiss_pro_4.png'))
+<<swiss_profile_sole_prep, out.width = "50%", fig.align = "center",fig.pos="H",echo=FALSE>>=
+knitr::include_graphics(c('swiss_pro_4.png','swiss_pro_2.png'))
 @

### 2022-01-09 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.tex
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3105 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-Parameter & Estimate & 95\% CI & Estimate & 95\% CI\\
+Parameter & Estimate & 90\% CI & Estimate & 90\% CI\\
 \midrule
-range & 38618.730 & (21805.93, 68394.54) & 38618.730 & NA\\
-sdNugget & 0.992 & (0.684,1.439) & 0.992 & NA\\
-shape & 1.826 & (-0.574,4.227) & 1.826 & NA\\
-anisoRatio & 8.092 & (5.321,12.305) & 8.092 & NA\\
-anisoAngleRadians & 0.652 & (0.595, 0.709) & 0.652 & NA\\
-boxcox & 0.496 & (0.373,0.620) & 0.496 & NA\\
+range & 38618.730 & (18544.14 ,80424.68) & 38618.730 & (24202.12, 130428.60)\\
+sdNugget & 0.992 & (0.615,1.599) & 0.992 & (0, 1.236)\\
+shape & 1.826 & (-1.255,4.907) & 1.826 & (4.540, 6.000)\\
+anisoRatio & 8.092 & (4.725,13.386) & 8.092 & (4.721, 14.124)\\
+anisoAngleRadians & 0.652 & (0.578, 0.725) & 0.652 & (0.567,0.723)\\
+boxcox & 0.496 & (0.375,0.655) & 0.496 & (0.403, 0.628)\\
 \bottomrule
@@
 {\centering \includegraphics[width=0.5\linewidth]{profile_logl_swiss} 
+\includegraphics[width=0.5\linewidth]{betaprofile}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{swiss_pro_4}
+{\centering \includegraphics[width=0.5\linewidth]{swiss_pro_4}
+
+}
+
+{\centering \includegraphics[width=0.5\linewidth]{swiss_pro_2}
 

### 2022-01-10 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.Rnw
paper 2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3106 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-<<swiss_profile_many,echo = FALSE, eval=TRUE, cache=TRUE, fig.width=4,fig.height=3, fig.show ="hold", fig.align='center', fig.cap = "profile LogL of slope parameter", out.width = "50%",dev='pdf',fig.pos="H">>=
-setwd("/home/ruoyong/diseasemapping/pkg/oldgpuRandom/inst/doc/L2021_11")
-knitr::include_graphics(c('profile_logl_swiss.png','betaprofile.png'))
+<<swiss_profile_many,echo = FALSE, eval=TRUE, cache=TRUE, fig.width=4,fig.height=3, fig.show ="hold", fig.align='center', fig.cap = "profile LogL of slope parameter", out.width = "33%",dev='pdf',fig.pos="H">>=
+#setwd("/home/ruoyong/diseasemapping/pkg/oldgpuRandom/inst/doc/L2021_11")
+setwd("/home/ruoyong/L2021_11")
+knitr::include_graphics(c('profile_logl_swiss.png','betaprofile.png','varprofile.png'))
 @

### 2022-01-10 pkg/oldgpuRandom/inst/doc/L2021_11/simulation_code_big.r

@@ -1,13 +1,13 @@
 ############### A SIMULATION STUDY IN THE PAPER 2 ####################################
 library('geostatsp')
 Nsim=15
-Ngrid=40
+Ngrid=20
 set.seed(88)
-pointsGrid = expand.grid(seq(100*1e3,140*1e3, len=Ngrid), seq(100*1e3,120*1e3, len=Ngrid))
+pointsGrid = expand.grid(seq(100*1e4,120*1e4, len=Ngrid), seq(100*1e4,120*1e4, len=Ngrid))
 #plot(pointsGrid,cex=0.3)
 pointsGrid = pointsGrid[,1] + 1i*pointsGrid[,2]
-pointsRandom = pointsGrid + stats::runif(length(pointsGrid), 50, 500) * exp(stats::runif(length(pointsGrid), 0, 2*pi)*1i)
-#plot(pointsRandom,cex=0.3)
+pointsRandom = pointsGrid + stats::runif(length(pointsGrid), 500, 600) * exp(stats::runif(length(pointsGrid), 0, 2*pi)*1i)
+plot(pointsRandom,cex=0.3)
 #plot(Re(pointsRandom),Im(pointsRandom),cex=0.3)
 mydat = SpatialPointsDataFrame(cbind(Re(pointsRandom),Im(pointsRandom)),
                                data=data.frame(cov1 =stats::rnorm(Ngrid^2, 120, 20), cov2 = stats::rpois(Ngrid^2, 50))
@@ -19,13 +19,19 @@ mydat = SpatialPointsDataFrame(cbind(Re(pointsRandom),Im(pointsRandom)),
 # )
 # plot(cbind(seq(100*1e3,140*1e3, len=500), seq(100*1e3,120*1e3, len=500)), xlab="",ylab="")
 
+par(mfrow = c(1, 2))
+library('geoR')
+sim1 <- grf(400, cov.pars = c(1, 0.25))
+points.geodata(sim1, main = "simulated locations and values")
+plot(sim1, max.dist = 1, main = "true and empirical variograms")
 
+pict
 
 #mydat@data
 #mydat@coords
 
 ## simulate a random field
-trueParam = c(range=1000, shape=2, variance=0.5^2, tauSq=1^2, anisoRatio=1, anisoAngleRadians=0)
+trueParam = c(range=1000, shape=2, variance=0.5^2, tauSq=1^2, anisoRatio=2, anisoAngleRadians=0.3)
[…]

### 2022-01-10 pkg/oldgpuRandom/inst/doc/L2021_11/swiss_plots.r

@@ -18,16 +18,12 @@ library(gpuRandom)
 
 ParamList = list(
   range = c(exp(seq(log(15), log(swissRes$parameters['range']/1000), len=5))*1000, exp(seq(log(swissRes$parameters['range']/1000+20), log(240), len=5))*1000),
-  shape = c(0.1,0.25, 0.4, seq(0.8, swissRes$parameters['shape'], len=3), seq(swissRes$parameters['shape']+0.6, 4.5, len=3)),
+  shape = c(0.1,0.25, 0.4, seq(0.8, swissRes$parameters['shape'], len=3), seq(swissRes$parameters['shape']+0.5, 4.5, len=3)),
   anisoRatio =c( 1,1.5,2,3,seq(4, swissRes$parameters['anisoRatio'], len=4), seq(swissRes$parameters['anisoRatio']+1, 16, len=4) ),
   anisoAngleDegrees = c( seq(25,swissRes$parameters['anisoAngleDegrees'], len=5),  seq(swissRes$parameters['anisoAngleDegrees']+2, 45, len=5)),
   nugget = c(seq(0, swissRes$optim$mle['nugget'], len=4), seq(swissRes$optim$mle['nugget']+0.5, 8, len=5))
 ) 
 params = do.call(expand.grid, ParamList)
-
-
-
-
 
 
 output <- gpuRandom::likfitLgmCov(swissRes,

### 2022-01-10 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1, ref timenumber anistr
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3107 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.046   0.004   0.050
+##   0.044   0.004   0.047
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.316   0.008   0.323
+##   0.314   0.000   0.316
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##    1.94    0.01    1.95
+##   1.926   0.016   1.947
 \end{CodeOutput}
@@
 Simulating Gaussian random fields (GRF's) is computationally intensive due the high dimensionality of the problem, a 100 by 100 grid has 10,000 cells and a 10,000 by 10,000 variance matrix.
-Writing $x_1,\dots, x_n$ as point locations (i.e.\ centres of grid cells), the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Gaussian.
+Writing $x_1,\dots, x_n$ as point locations (i.e.\ centres of grid cells), the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Normal (MVN).
 
-GEOMETRIC ANISOTROPY
-Geometrical anisotropy is defined by two additional parameters in the correlation function.
-Let $(x_1, x_2)$ represent the original coordinates, which is rotated counterclockwise by $\varphi$, then the new coordinates $(x'_1, x'_2)$ is given by
-%
-  \begin{equation*}\label{eq:matrixeqn}
-    \begin{pmatrix}
-      x'_{1} \\
-      x'_{2}  
-    \end{pmatrix}
-    =
-    \begin{pmatrix}
-      1/\phi_{1} & 0 \\
-      0 & 1/ \phi_{2} 
-    \end{pmatrix}
-    \cdot
-    \begin{pmatrix}
-      \cos \varphi & -\sin \varphi \\
-      \sin \varphi & \cos \varphi 
-    \end{pmatrix}
-    \cdot
-      \begin{pmatrix}
-      x_1 \\
-      x_2 
-    \end{pmatrix},
-  \end{equation*}
-
-The expectation and covariance function of $U$ completely determine the distribution of $U$, 
 \begin{gather*} 
@@
 \end{gather*}
-where \text{MVN} standards for the multivariate normal distribution, and $\Sigma$ standards for the covariance matrix of $U$.
+where $\Sigma$ standards for the covariance matrix of $U$.
 A common choice for $\Sigma$ is the Mat\'ern covariance \citep{matern1960spatial}.
@@
 $\Gamma(\cdot)$ is the standard gamma function.  
-$\phi$ is the range parameter or scale parameter with the dimension of distance, it controls the rate of decay of the correlation as distance increases.
+$\phi$ is the range parameter or scale parameter, it controls the rate of decay of the correlation as distance increases.
 $K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$.
@@
 There are several alternative parameterisations of Mat\'ern covariance functions in literature  \citep[see][]{haskard2007anisotropic}.
-In \pkg{clrng}, we use the above form for computing Mat\'ern covariance.
+In \pkg{gpuBatchMatrix}, we use the above form for computing Mat\'ern covariance.
+GEOMETRIC ANISOTROPY
+Geometrical anisotropy is defined by two additional parameters (`anisoRatio' and `anisoAngle') in the correlation function, which makes the resulting multidimensional probability surface even more complex to evaluate.
 
@@
 \cite{LiuandLi2019} gives a comprehensive review on seven popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
-all of these methods except the matrix decomposition method, are approximations to the random field and have specific requirements on the type of grid or covariance functions.
-The matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \citep{LiuandLi2019}.
+all of these methods except the direct matrix decomposition method, are approximations to the random field and have specific requirements on the type of grid or covariance functions.
+The direct matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \citep{LiuandLi2019}.
 Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations.
 The \pkg{RandomFields} \citep{RandomFields2015} package has several different methods for simulation of Gaussian fields, among which the circulant embedding, which is an improved method on direct matrix decomposition, and some variants of the method like the cut-off embedding \citep{gneiting2006fast} are also exact and fast for isotropic Gaussian random fields, however, they work only on rectangular grids.
-\pkg{clrng} is able to simulate on the GPU exact Gaussian fields on arbitrary grids as it relies on the matrix decomposition method.
+We are able to simulate on GPU exact Gaussian fields on arbitrary grids by using the direct matrix decomposition method.
 The implementation of this method is as follows.

### 2022-01-10 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3108 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.044   0.004   0.047
+##   0.045   0.004   0.049
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.314   0.000   0.316
+##   0.319   0.000   0.319
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.926   0.016   1.947
+##   1.923   0.023   1.949
 \end{CodeOutput}
@@
 The direct matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \citep{LiuandLi2019}.
-Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations.
-The \pkg{RandomFields} \citep{RandomFields2015} package has several different methods for simulation of Gaussian fields, among which the circulant embedding, which is an improved method on direct matrix decomposition, and some variants of the method like the cut-off embedding \citep{gneiting2006fast} are also exact and fast for isotropic Gaussian random fields, however, they work only on rectangular grids.
-We are able to simulate on GPU exact Gaussian fields on arbitrary grids by using the direct matrix decomposition method.
+By using this method, we are able to simulate on GPU exact Gaussian fields on arbitrary grids.
 The implementation of this method is as follows.
-
-%it can generate random fields on arbitrary position Circulant embedding is a fast simulation method for stationary (possibly anisotropic) Gaussian ran- dom fields on regular grids based on Fourier transformations.
-Cut-off embedding is a fast simulation method for stationary, isotropic Gaussian random fields on square lattices based on the standard RPcirculant method, so that exact simulation is garantueed for further covariance models
 
@@
 
+The downside of this method is that it's computationally expensive and is thus very slow.
+When the number of locations $n$ is large, the time cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and is $\mathcal{O}(n^2)$ for the matrix-vector multiplication $L*Z$ \citep{LiuandLi2019}.
+Methods employing approximations gain benefits in computations while sacrificing exactness, we preserve the exactness and also largely reduce the computation burden by utilizing our other \proglang{R} package \pkg{gpuBatchMatrix}.
+\pkg{gpuBatchMatrix} address this issue since it computes batches of Mat\'ern covariance matrices in parallel, and does Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
+
+Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations.
+The \pkg{RandomFields} \citep{RandomFields2015} package has several different methods for simulation of Gaussian fields, among which the circulant embedding, which is an improved method on direct matrix decomposition, and some variants of the method like the cut-off embedding \citep{gneiting2006fast} are also exact and fast for isotropic Gaussian random fields, however, they work only on rectangular grids.
+
+%it can generate random fields on arbitrary position Circulant embedding is a fast simulation method for stationary (possibly anisotropic) Gaussian ran- dom fields on regular grids based on Fourier transformations.
+Cut-off embedding is a fast simulation method for stationary, isotropic Gaussian random fields on square lattices based on the standard RPcirculant method, so that exact simulation is garantueed for further covariance models
+
 % $U=L*D^{1/2}*Z$, where 
@@
 % \end{itemize}
-The downside of this method is that it's computationally expensive and is thus very slow.
-When the number of locations $n$ is large, the time cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and is $\mathcal{O}(n^2)$ for the matrix-vector multiplication $L*Z$ \citep{LiuandLi2019}.
-Our other \proglang{R} package \pkg{gpuBatchMatrix} is able to address this issue since it computes batches of Mat\'ern covariance matrices in parallel, and does Cholesky decomposition and matrix-matrix multiplication in batches in parallel on GPU.
-The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
 %\pkg{clrng} package not only does the matrices decomposition (Cholesky decomposition and matrix-vector multiplication) in parallel, but also computes the covariance matrices in parallel on GPU.
@@
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
+The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
 \begin{CodeChunk}

### 2022-01-10 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update paper 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3108 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
+The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
 <<setupData, eval=TRUE,echo=TRUE, message=FALSE>>=

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3110 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   get_global_id(1)) * NpadStreams;
-
 const double fact[2] = { mrg31k3p_NORM_cl, TWOPI * mrg31k3p_NORM_cl };

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
simulation section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3111 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Writing $x_1,\dots, x_n$ as point locations (i.e.\ centres of grid cells), the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Normal (MVN).
+An isotropic random field has a variance matrix  with entries which depend on the distances between point locations, and a GRF with geometric anisotoropy has entries depending on the direction as well as the distance.
 
-\begin{gather*} 
-U = [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
-\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|x_i - x_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|x_i - x_j|}{\phi}\right),
+A geometrically anisotropic GRF with a \citet{matern1960spatial} correlation function is defined as
+\begin{aligned*} 
+U = &[U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
+\Sigma_{ij} = & \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} ||d_{ij}||/\phi \right)^\kappa  K_\kappa\left(\sqrt{8\kappa} ||d_{ij}||/\phi \right),\\
+d_{ij} = &
+        \left[
+          \begin{array}{rr}
+            1 & 0 \\
+            0 & \omega
+          \end{array}
+        \right]
+        \left[
+          \begin{array}{rr}
+            \cos(\theta) & -\sin(\theta) \\
+            \sin(\theta) & \cos(\theta)
+          \end{array}
+        \right]
+(x_i - x_j)
 \end{gather*}
-where $\Sigma$ standards for the covariance matrix of $U$.
-A common choice for $\Sigma$ is the Mat\'ern covariance \citep{matern1960spatial}.
-The Mat\'ern covariance between $U(x_i)$ and $U(x_j)$ takes the above form,
-%$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
-where $\sigma^2$ the variance of the random field $U$.
-$\Gamma(\cdot)$ is the standard gamma function.  
-$\phi$ is the range parameter or scale parameter, it controls the rate of decay of the correlation as distance increases.
-$K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$.
-%specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.
-%(For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.)
+where $K_\kappa(\cdot)$  is a modified Bessel function of the second kind with order $\kappa$ and $\Gamma(\cdot)$ is the standard gamma function.  The model paramaters are:
+\begin{itemize}
+\item $\sigma$, the (marginal) variance  $\var(U_i) = \sigma$;
+\item $\kappa$, the shape parameter controlling the differentiability of $U(\cdot)$;
+\item $\phi$, the range parameter controlling how quickly correlation decays with distance in the direction where correlation is strongest;
+\item $\omega \geq 1$, a parameter controlling how much faster correlation decays in the direction where correlation is weakest; and 
+\item $\theta$, the angle of rotation  required to put the direction of maximum correlation on the x-axis.
+\end{itemize}
+The standard isotropic Mat\'ern model is obtained by setting $\omega = 1$.
 There are several alternative parameterisations of Mat\'ern covariance functions in literature  \citep[see][]{haskard2007anisotropic}.
-In \pkg{gpuBatchMatrix}, we use the above form for computing Mat\'ern covariance.
-GEOMETRIC ANISOTROPY
-Geometrical anisotropy is defined by two additional parameters (`anisoRatio' and `anisoAngle') in the correlation function, which makes the resulting multidimensional probability surface even more complex to evaluate.
+In \pkg{gpuBatchMatrix}, we use the above form with the $\sqrt{8\kappa}$ term as it allows $\phi$ to be interpretable as the distance beyond which correlation is less than 0.14.
 
-%Due to the importance of Gaussian random field, there are a large amount of studies devoted to Gaussian random field generation techniques. 
+The conceptually simplest way of simulating a GRF is to multiply a vector of independent standard Normals by the Cholesky decomposition of $\Sigma$.  This is the direct decomposition method.  
 \cite{LiuandLi2019} gives a comprehensive review on seven popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
-all of these methods except the direct matrix decomposition method, are approximations to the random field and have specific requirements on the type of grid or covariance functions.
-The direct matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \citep{LiuandLi2019}.
-By using this method, we are able to simulate on GPU exact Gaussian fields on arbitrary grids.
-The implementation of this method is as follows.
-
-Suppose $U=(U_1, U_2, \dots, U_n)^\top$ is a Gaussian random field with mean $\mu$ and covariance matrix $\Sigma$, without loss of generality, we assume mean value zero $(\mu = 0)$.
-Since covariance matrix $\Sigma$ is symmetric and positive-definite, we can take Cholesky decomposition of $\Sigma$, then we can simulate random samples from $U$ using Algorithm \ref{algorithm2},
+all of these methods other than the direct matrix decomposition method, involve approximations to the random field and have specific requirements on the type of grid or covariance functions.
+The direct matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary point locations, and is straightforward to implement.
 
@@
  4, Compute the random samples from $U$ in batches using $U = L \cdot D^{\frac{1}{2}} \cdot Z$ \;
- \caption{Gaussian random fields simulation using covariance matrix decomposition method.}
+ \caption{Gaussian random fields simulation using the direct matrix decomposition method.}
  \label{algorithm2}
@@
 
-The downside of this method is that it's computationally expensive and is thus very slow.
-When the number of locations $n$ is large, the time cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and is $\mathcal{O}(n^2)$ for the matrix-vector multiplication $L*Z$ \citep{LiuandLi2019}.
+ The direct method is detailed in Algorithm \ref{algorithm2}.
+It is computationally demanding for two reasons.
+First, evaluating the Bessel function is an iterative procedure that must be done for each of the $n^2/2$ entries of $\Sigma$.
+Second, the time cost of Cholesky decomposition of the covariance matrix is $\mathcal{O}(n^3)$,  and is $\mathcal{O}(n^2)$ for the matrix-vector multiplication $L*Z$ \citep{LiuandLi2019}.
 Methods employing approximations gain benefits in computations while sacrificing exactness, we preserve the exactness and also largely reduce the computation burden by utilizing our other \proglang{R} package \pkg{gpuBatchMatrix}.
@@
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
+
 The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
-<<setupData, eval=TRUE,echo=TRUE, message=FALSE>>=
+Motivated by the classic Swiss rain dataset, we simluate random fields on a grid of points covering Switzerland.
+
+<<setupData, eval=TRUE, include=FALSE, message=FALSE>>=
 library("gpuBatchMatrix")
@@
 @
-Step 1, set up a $51 \times 80$ raster using the `swissBorder' coordinates from \pkg{geostatsp}.
-<<setupcoords, eval=TRUE>>=
+Step 1, we create a grid of points covering Switzerland using the the \code{swissBorder} object from \pkg{geostatsp}.  The grid is specified as having 90 cells in the horizontal direction.
+<<setupcoords, eval=TRUE, include=FALSE>>=
 NlocalCache = 1000
@@
 <<swissRainBoundary, echo=TRUE>>=
+library('geostatsp')
 data("swissRain", package="geostatsp")
@@
 @
-Step 2, create four parameter sets as a small example. 
-<<setup paramsBatch, eval=TRUE,echo=FALSE>>=
+Step 2, create a matrix with four different parameter sets.  The five parameters named are $\kappa$, $\phi$, $\sigma^2$, $\omega$ and $\theta$.  
+<<setup paramsBatch, eval=TRUE,echo=TRUE>>=
 params = 
-rbind(c(shape=1.25, range=50*1000, variance = 1.5, nugget = 0,anisoRatio = 1, anisoAngleRadians = 0), 
-c(shape=2.15, range=60*1000, variance = 2, nugget = 0, anisoRatio = 4, anisoAngleRadians = pi/7),
-c(shape=0.6, range=30*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/7),
-c(shape=3, range=30*1000, variance = 2, nugget = 0, anisoRatio = 2, anisoAngleRadians = pi/7)
+rbind(c(shape=1.25, range=50*1000, variance = 1.5, anisoRatio = 1, anisoAngleRadians = 0), 
+c(2.15, 60*1000, 2,  4, pi/7),
+c(0.6, 30*1000, 2,  2, pi/5),
+c(3, 30*1000, 2,   2, pi/7)
 )
@@
 Step 3, compute the Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
-are each of size $4080 \times 4080$ and are stacked by row in the output \code{maternCov}. 
+are each of size $\Sexpr{prod(dim(myRaster)[1:2])} \times \Sexpr{prod(dim(myRaster)[1:2])}$ and are stacked by row in the output \code{maternCov}. 
 <<r outputBatchCreate, eval=TRUE>>=
@@
 \end{gather}
-Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnormGpu}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU and generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnormGpu}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 <<r randomNormalsSim1, eval=TRUE, cache=FALSE>>=
-streamsGpu <- createStreamsGpu(128*64)
-@
-<<r randomNormalsSim2, eval=TRUE, cache=FALSE>>=
+streamsForGrf <- createStreamsGpu(128*64)
 zmatGpu = clrng::rnormGpu(
-                c(nrow(maternCov),2), streams=streamsGpu, 
+                c(nrow(maternCov),2), streams=streamsForGrf, 
                 Nglobal=c(128,64),
-                type = theType)
+                type = 'double')
 @
@@
 \end{gather}
-Step 7, Finally, plot the ten realizations.
+Step 7, Finally, convert the results to a spatial \code{raster} object and plot them.
 <<r plotSimSetup, echo=TRUE, eval=TRUE, cache=FALSE>>=
@@
 @
-
+A simple plot can be produced with
+<<plotSimplematern, eval=FALSE, echo=TRUE>>=
+plot(simRaster)
+@
+and
+Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} package.
 <<r theSubcapName, include=FALSE>>=
@@
 
-<<maternplot, echo=FALSE, eval=TRUE, cache=FALSE, dev='png', out.width = '0.47\\textwidth', fig.height=3, fig.width=4.5, fig.cap="Simulated Gaussian random fields", fig.subcap = theSubcap, fig.align='default'>>=
-myCol = mapmisc::colourScale(breaks=sort(unique(c(-7, -4, seq(-2, 2), 4, 7))), style='fixed', col='Spectral')
+<<maternplot, echo=FALSE, eval=TRUE, cache=FALSE, dev='png', out.width = '0.47\\textwidth', fig.height=3, fig.width=4.5, fig.cap="Simulated Gaussian random fields", fig.subcap = theSubcap, fig.align='default', fig.pos='p'>>=
+myCol = mapmisc::colourScale(breaks=c(-7, -4, seq(-2, 2), 4, 7), style='fixed', col='Spectral')
 for(D in names(simRaster)) {
@@
 }
-mapmisc::legendBreaks("right", myCol, inset=0)
+mapmisc::legendBreaks("topleft", myCol, inset=0)
 @

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
pkg cite & small fix
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3112 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \def\E{{\text{E}}}
+\def\var{{\text{var}}}
 \def\T{{\footnotesize{^{_{\sf T}}}}}
@@
 ##    user  system elapsed 
-##   0.045   0.004   0.049
+##   0.053   0.000   0.053
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.319   0.000   0.319
+##   0.335   0.000   0.335
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.923   0.023   1.949
+##   1.934   0.016   1.948
 \end{CodeOutput}
@@
 Writing $x_1,\dots, x_n$ as point locations (i.e.\ centres of grid cells), the joint distribution of $U=(U_{x_1},\dots,U_{x_n})^\top$ is multivariate Normal (MVN).
+An isotropic random field has a variance matrix  with entries which depend on the distances between point locations, and a GRF with geometric anisotoropy has entries depending on the direction as well as the distance.
+A geometrically anisotropic GRF with a \citet{matern1960spatial} correlation function is defined as
+\begin{align*}
+U &= [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
+\Sigma_{ij} &= \cov[ U(x_i),U(x_j) ]= \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} ||d_{ij}||/\phi \right)^\kappa  K_\kappa\left(\sqrt{8\kappa} ||d_{ij}||/\phi \right),\\
+d_{ij} &= 
+        \left[
+          \begin{array}{rr}
+            1 & 0 \\
+            0 & \omega
+          \end{array}
+        \right]
+        \left[
+          \begin{array}{rr}
+            \cos(\theta) & -\sin(\theta) \\
+            \sin(\theta) & \cos(\theta)
+          \end{array}
+        \right]
+(x_i - x_j)
+\end{align*}
+where $K_\kappa(\cdot)$  is a modified Bessel function of the second kind with order $\kappa$ and $\Gamma(\cdot)$ is the standard gamma function.  The model parameters are:
+\begin{itemize}
+\setlength{\parskip}{0pt}
+\item $\sigma$, the (marginal) variance  $\var(U_i) = \sigma$;
+\item $\kappa$, the shape parameter controlling the differentiability of $U(\cdot)$;
+\item $\phi$, the range parameter controlling how quickly correlation decays with distance in the direction where correlation is strongest;
+\item $\omega \geq 1$, a parameter controlling how much faster correlation decays in the direction where correlation is weakest; and 
+\item $\theta$, the angle of rotation  required to put the direction of maximum correlation on the x-axis.
+\end{itemize}
+The standard isotropic Mat\'ern model is obtained by setting $\omega = 1$.
+There are several alternative parameterisations of Mat\'ern covariance functions in literature  \citep[see][]{haskard2007anisotropic}.
+In \pkg{gpuBatchMatrix}, we use the above form with the $\sqrt{8\kappa}$ term as it allows $\phi$ to be interpretable as the distance beyond which correlation is less than 0.14.
 
-\begin{gather*} 
-U = [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
-\Sigma_{ij} = \cov[ U(x_i),U(x_j) ] = \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} \frac{|x_i - x_j|}{\phi}\right)^\kappa  K_\kappa\left(\sqrt{8\kappa}  \frac{|x_i - x_j|}{\phi}\right),
-\end{gather*}
-where $\Sigma$ standards for the covariance matrix of $U$.
-A common choice for $\Sigma$ is the Mat\'ern covariance \citep{matern1960spatial}.
-The Mat\'ern covariance between $U(x_i)$ and $U(x_j)$ takes the above form,
-%$d =\|x-x'\|$ is the Euclidean distance between two spatial points, 
-where $\sigma^2$ the variance of the random field $U$.
-$\Gamma(\cdot)$ is the standard gamma function.  
-$\phi$ is the range parameter or scale parameter, it controls the rate of decay of the correlation as distance increases.
-$K_\kappa(\cdot)$ is the modified Bessel function of the second kind with order $\kappa$, $\kappa$ is the shape parameter which determines the smoothness of $U(x)$.
-%specifically, $U(x)$ is $m$ times mean-square differentiable if and only if $\kappa > m$.
-%(For $\kappa=0.5$, the Mat\'ern covariance function reduces to the exponential covariance function $\sigma^2*\exp(-d/\phi)$, when $\kappa \rightarrow \infty$, $\Sigma(d) \rightarrow \sigma^2*\exp{-(\|d\|/\phi)^2}$, which is called the Gaussian covariance.)
-There are several alternative parameterisations of Mat\'ern covariance functions in literature  \citep[see][]{haskard2007anisotropic}.
-In \pkg{gpuBatchMatrix}, we use the above form for computing Mat\'ern covariance.
-GEOMETRIC ANISOTROPY
-Geometrical anisotropy is defined by two additional parameters (`anisoRatio' and `anisoAngle') in the correlation function, which makes the resulting multidimensional probability surface even more complex to evaluate.
-
-%Due to the importance of Gaussian random field, there are a large amount of studies devoted to Gaussian random field generation techniques. 
-\cite{LiuandLi2019} gives a comprehensive review on seven popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
-all of these methods except the direct matrix decomposition method, are approximations to the random field and have specific requirements on the type of grid or covariance functions.
-The direct matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary grids, and is straightforward to implement \citep{LiuandLi2019}.
-By using this method, we are able to simulate on GPU exact Gaussian fields on arbitrary grids.
-The implementation of this method is as follows.
-
-Suppose $U=(U_1, U_2, \dots, U_n)^\top$ is a Gaussian random field with mean $\mu$ and covariance matrix $\Sigma$, without loss of generality, we assume mean value zero $(\mu = 0)$.
-Since covariance matrix $\Sigma$ is symmetric and positive-definite, we can take Cholesky decomposition of $\Sigma$, then we can simulate random samples from $U$ using Algorithm \ref{algorithm2},
+The conceptually simplest way of simulating a GRF is to multiply a vector of independent standard Normals by the Cholesky decomposition of $\Sigma$.  This is the direct matrix decomposition method.  
+\cite{LiuandLi2019} gives a comprehensive review on seven popular methods for GRF generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
+all of these methods other than the matrix decomposition method, involve approximations to the random field and have specific requirements on the type of grid or covariance functions.
+The direct matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary point locations, and is straightforward to implement.
 
@@
 
-Other \proglang{R} packages that offer simulation of Gaussian random fields such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations.
-The \pkg{RandomFields} \citep{RandomFields2015} package has several different methods for simulation of Gaussian fields, among which the circulant embedding, which is an improved method on direct matrix decomposition, and some variants of the method like the cut-off embedding \citep{gneiting2006fast} are also exact and fast for isotropic Gaussian random fields, however, they work only on rectangular grids.
+Other \proglang{R} packages that offer simulation of GRF's such as \pkg{geoR} \citep{geoR2001}, does not work for large number of locations.
+The \pkg{RandomFields} \citep{RandomFields2015} package has several different methods for simulation of Gaussian fields, among which the circulant embedding, which is an improved method on direct matrix decomposition, and some variants of the method like the cut-off embedding \citep{gneiting2006fast} are also exact and fast for isotropic GRF's, however, they work only on rectangular grids.
 
@@
 \subsection{Simulating Gaussian random fields with Mat\'ern covariances}
-The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library("gpuBatchMatrix")
-R> library("geostatsp")
-\end{CodeInput} 
-\end{CodeChunk}
 
-Step 1, set up a $51 \times 80$ raster using the `swissBorder' coordinates from \pkg{geostatsp}.
-\begin{CodeChunk}
-\begin{CodeInput}
-R> NlocalCache = 1000
-R> Nglobal = c(128, 64, 2)
-R> Nlocal = c(16, 8, 2)
-R> theType = "double"
-\end{CodeInput} 
-\end{CodeChunk}
+The following shows an example, in which we simulate on GPU eight GRF's of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
+Motivated by the classic Swiss rain dataset, we simluate random fields on a grid of points covering Switzerland.
+
+Step 1, we create a grid of points covering Switzerland using the the \code{swissBorder} object from \pkg{geostatsp}.  The grid is specified as having 90 cells in the horizontal direction.
 
@@
 \begin{CodeInput}
+R> library("geostatsp")
 R> data("swissRain", package = "geostatsp")
@@
 
-Step 2, create four parameter sets as a small example.
+Step 2, create a matrix with four different parameter sets.  The five parameters named are $\kappa$, $\phi$, $\sigma^2$, $\omega$ and $\theta$.  
+\begin{CodeChunk}
+\begin{CodeInput}
+R> params = rbind(c(shape = 1.25, range = 50 * 1000, variance = 1.5,
++    anisoRatio = 1, anisoAngleRadians = 0), c(2.15, 60 * 1000,
++    2, 4, pi/7), c(0.6, 30 * 1000, 2, 2, pi/5), c(3, 30 *
++    1000, 2, 2, pi/7))
+\end{CodeInput} 
+\end{CodeChunk}
 
@@
 \begin{CodeOutput}
-##      shape range variance nugget anisoRatio anisoAngleRadians
-## [1,]  1.25 50000      1.5      0          1             0.000
-## [2,]  2.15 60000      2.0      0          4             0.449
-## [3,]  0.60 30000      2.0      0          2             0.449
-## [4,]  3.00 30000      2.0      0          2             0.449
+##      shape range variance anisoRatio anisoAngleRadians
+## [1,]  1.25 50000      1.5          1             0.000
+## [2,]  2.15 60000      2.0          4             0.449
+## [3,]  0.60 30000      2.0          2             0.628
+## [4,]  3.00 30000      2.0          2             0.449
 \end{CodeOutput} 
@@
 Step 3, compute the Mat\'ern covariance matrices using \fct{gpuBatchMatrix::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices
-are each of size $4080 \times 4080$ and are stacked by row in the output \code{maternCov}. 
+are each of size $5130 \times 5130$ and are stacked by row in the output \code{maternCov}. 
 \begin{CodeChunk}
@@
 \end{gather}
-Step 5, create some streams on GPU to generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnormGpu}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
+Step 5, create some streams on GPU and generate two standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{clrng::rnormGpu}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
 \begin{CodeChunk}
 \begin{CodeInput}
-R> streamsGpu <- createStreamsGpu(128 * 64)
-\end{CodeInput} 
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> zmatGpu = clrng::rnormGpu(c(nrow(maternCov), 2), streams = streamsGpu,
-+    Nglobal = c(128, 64), type = theType)
+R> streamsForGrf <- createStreamsGpu(128 * 64)
+R> zmatGpu = clrng::rnormGpu(c(nrow(maternCov), 2), streams = streamsForGrf,
++    Nglobal = c(128, 64), type = "double")
 \end{CodeInput} 
@@
 \end{gather}
-Step 7, Finally, plot the ten realizations.
+Step 7, Finally, convert the results to a spatial \code{raster} object and plot them.
 \begin{CodeChunk}
@@
 
-\begin{figure}
+A simple plot can be produced with
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plot(simRaster)
+\end{CodeInput} 
+\end{CodeChunk}
+
+and
+Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} \citep{mapmiscPackage} package.
+
+\begin{figure}[p]
 \subfloat[parameter 1, simuation 1\label{fig:maternplot-1}]{\includegraphics[width=0.47\textwidth]{Figures/Gmaternplot-1} }

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
pkg cite & small fix
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3112 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 An isotropic random field has a variance matrix  with entries which depend on the distances between point locations, and a GRF with geometric anisotoropy has entries depending on the direction as well as the distance.
-
 A geometrically anisotropic GRF with a \citet{matern1960spatial} correlation function is defined as
-\begin{aligned*} 
-U = &[U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
-\Sigma_{ij} = & \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} ||d_{ij}||/\phi \right)^\kappa  K_\kappa\left(\sqrt{8\kappa} ||d_{ij}||/\phi \right),\\
-d_{ij} = &
+\begin{align*}
+U &= [U(x_1), \ldots, U(x_n)]^\top \sim \text{MVN}(0, \Sigma), \\
+\Sigma_{ij} &= \cov[ U(x_i),U(x_j) ]= \sigma^2 \frac{2^{\kappa-1}}{\Gamma(\kappa)} \left(\sqrt{8\kappa} ||d_{ij}||/\phi \right)^\kappa  K_\kappa\left(\sqrt{8\kappa} ||d_{ij}||/\phi \right),\\
+d_{ij} &= 
         \left[
@@
 (x_i - x_j)
-\end{gather*}
-where $K_\kappa(\cdot)$  is a modified Bessel function of the second kind with order $\kappa$ and $\Gamma(\cdot)$ is the standard gamma function.  The model paramaters are:
+\end{align*}
+where $K_\kappa(\cdot)$  is a modified Bessel function of the second kind with order $\kappa$ and $\Gamma(\cdot)$ is the standard gamma function.  The model parameters are:
 \begin{itemize}
+\setlength{\parskip}{0pt}
 \item $\sigma$, the (marginal) variance  $\var(U_i) = \sigma$;
@@
 
-The conceptually simplest way of simulating a GRF is to multiply a vector of independent standard Normals by the Cholesky decomposition of $\Sigma$.  This is the direct decomposition method.  
-\cite{LiuandLi2019} gives a comprehensive review on seven popular methods for Gaussian random field generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
-all of these methods other than the direct matrix decomposition method, involve approximations to the random field and have specific requirements on the type of grid or covariance functions.
+The conceptually simplest way of simulating a GRF is to multiply a vector of independent standard Normals by the Cholesky decomposition of $\Sigma$.  This is the direct matrix decomposition method.  
+\cite{LiuandLi2019} gives a comprehensive review on seven popular methods for GRF generation, %which are the turning bands method \citep{matheron_1973}, spectral method \citep{Meja1974OnTS,1972JSV}, matrix decomposition method \citep{davis1987}, Karhunen-Lo\'eve expansion \citep{LoeveM:1978}, moving average method \citep{journel1974, oliver1995moving}, sequential simulation \citep{johnson1987multivariate, gomez1993joint, pebesma2004multivariable}, and local average subdivision \citep{fenton1990simulation}.
+all of these methods other than the matrix decomposition method, involve approximations to the random field and have specific requirements on the type of grid or covariance functions.
 The direct matrix decomposition method is exact, it works for all covariance functions and can generate random fields on arbitrary point locations, and is straightforward to implement.
@@
 
-The following shows an example, in which we simulate on GPU eight Gaussian random fields of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
+The following shows an example, in which we simulate on GPU eight GRF's of Mat\'ern covariance at one time with four sets of parameters, by taking advantage of the GPU-capabilities offered by \pkg{clrng} and \pkg{gpuBatchMatrix} together.
 Motivated by the classic Swiss rain dataset, we simluate random fields on a grid of points covering Switzerland.
@@
 and
-Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} package.
+Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} \citep{mapmiscPackage} package.
 <<r theSubcapName, include=FALSE>>=

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
item sep
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3113 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{booktabs}
+\usepackage{enumitem}
 
@@
 where $K_\kappa(\cdot)$  is a modified Bessel function of the second kind with order $\kappa$ and $\Gamma(\cdot)$ is the standard gamma function.  The model parameters are:
+\vspace{-\topsep}
 \begin{itemize}
 \setlength{\parskip}{0pt}
+\itemsep0em 
+%\setlength{\itemsep}{0.5pt plus 1pt}
 \item $\sigma$, the (marginal) variance  $\var(U_i) = \sigma$;
@@
 \end{itemize}
+\vspace{-\topsep}
 The standard isotropic Mat\'ern model is obtained by setting $\omega = 1$.
@@
 @
-and
-Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} \citep{mapmiscPackage} package.
+and Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} \citep{mapmiscPackage} package.
 <<r theSubcapName, include=FALSE>>=
@@
 @
-
 <<maternplot, echo=FALSE, eval=TRUE, cache=FALSE, dev='png', out.width = '0.47\\textwidth', fig.height=3, fig.width=4.5, fig.cap="Simulated Gaussian random fields", fig.subcap = theSubcap, fig.align='default', fig.pos='p'>>=

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
item sep
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3114 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   var1 = c("B", "1M", "10M", "1M", "10M"),
-  var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 49.33,  327.27),
-  var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.15,    10.23),
-  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.80,  92.00),
-  var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.32,   1.95),
+  var2 = c('Intel 2.5ghz', 0.403804, 0.0001251, 49.3,  327.3),
+  var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.2,    10.2),
+  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.8,  91.5),
+  var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.3,   1.9),
   var6 = c('Data','month','week','month','week')
@@
 Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII.
-Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}") %>%
+Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}, run-time varies a little bit for each execution.") %>%
   kable_styling(full_width = FALSE, position = "center", latex_options = "HOLD_position") %>%

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
item sep
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3113 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
item sep
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3114 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.053   0.000   0.053
+##   0.045   0.007   0.052
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.335   0.000   0.335
+##   0.317   0.000   0.317
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.934   0.016   1.948
+##   1.962   0.002   1.966
 \end{CodeOutput}
@@
 Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII.
-Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\label{tab:summary}}
+Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\label{tab:summary}, run-time varies a little bit for each execution.}
 \centering
@@
 \hline
-\hspace{1em}1M & 49.33 & 2.15 & 14.8 & 0.32 & month\\
+\hspace{1em}1M & 49.3 & 2.2 & 14.8 & 0.3 & month\\
 \hline
-\hspace{1em}10M & 327.27 & 10.23 & 92 & 1.95 & week\\
+\hspace{1em}10M & 327.3 & 10.2 & 91.5 & 1.9 & week\\
 \hline
@@
 
-and
-Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} \citep{mapmiscPackage} package.
+and Figure \ref{fig:maternplot} uses some facilities from the \pkg{mapmisc} \citep{mapmiscPackage} package.
 

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
item sep
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3115 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.045   0.007   0.052
+##   0.060   0.011   0.071
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.317   0.000   0.317
+##   0.318   0.000   0.318
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.962   0.002   1.966
+##   1.925   0.027   1.951
 \end{CodeOutput}
@@
 Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII.
-Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\label{tab:summary}, run-time varies a little bit for each execution.}
+Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100 \label{tab:summary}.
+Run-time varies a little bit for each execution.}
 \centering

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
item sep
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3115 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII.
-Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\\label{tab:summary}, run-time varies a little bit for each execution.") %>%
+Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100 \\label{tab:summary}.
+Run-time varies a little bit for each execution.") %>%
   kable_styling(full_width = FALSE, position = "center", latex_options = "HOLD_position") %>%

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3116 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Future work on \pkg{clrng} package could be: (1) Explore other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators in \proglang{clRNG} library, and compare the \proglang{R} performance between using different GPU RNGs.
-(2) Now that the package can do Cholesky decomposition and matrix multiplication in batches on GPU, which is a motivation for us to work on parallel likelihood evaluations on GPU for Gaussian spatial models in the next step.
-(3) Create a ``sparse'' matrix class on GPU and use it for simulating Gaussian random fields with sparse correlation structures such as the Gaussian Markov random fields.
+Some possible future work on \pkg{clrng} package could be: (1) Incorporate other RNGs, for example the MRG32k3a, LFSR113, and Philox-4×32-10 generators from \proglang{clRNG} library, and compare the \proglang{R} performance between using different RNGs on GPU.
+(2) Create a ``sparse'' matrix class on GPU and use it for simulating Gaussian random fields with sparse correlation structures such as the Gaussian Markov random fields.
+(3) Apply GPU-generated random numbers in suitable Monte Carlo simulations to speed up the performance.
 

### 2022-01-17 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
p1 replication r codes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3117 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
last update on paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3118 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 options(scipen=999)
+options(digits = 6)
 @
@@
 
-\proglang{clRNG} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators.
+\proglang{clRNG} \citep{l2015clrng} is an \proglang{OpenCL} library for uniform random number generation, it provides four different RNGs: the MRG31k3p, MRG32k3a, LFSR113, and Philox-4×32-10 generators.
 These four RNGs use different types of constructions.
@@
 It is also faster than Philox-4×32-10.
-The MRG31k3p generator was introduced in the paper \citep{rLEC00b}.
+The MRG31k3p generator was introduced by \cite{rLEC00b}.
 MRG31k3p was also statistically tested extensively and successfully, \cite[see][]{rLEC07b}.
@@
 \subsection{Creating streams}\label{createstreams}
-When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation would be more complicated because there is no guarantee that the streams (analogous to \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers is possible to have correlation.
+When a RNG is called in parallel processes or successively called at several places in a program in \proglang{R}, random number generation would be more complicated because there is no guarantee that the streams (analogous to \code{.Random.seed} in \proglang{R}) do not overlap, and thus the generated sequences of random numbers may have dependence.
 \pkg{clrng} uses multiple distinct streams that are used in work-items that executes in parallel on a GPU device.
@@
 Each of these streams can also be partitioned into substreams with equally-spaced starting points \citep{l2002object, rLEC15a}, although this is not currently implemented in \pkg{clrng}.
-In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream (by default it is equal to the seed).
+In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed).
 Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by $Z$ steps to find its initial state, and the three states in the stream object are set to it.
@@
 \begin{itemize}
-\item Calling \fct{setBaseCreator} creates a hidden object \code{.Random.seed.clrng}, which (unlike the standard \code{.Random.seed}) controls the creation of new streams (not random number).
+\item Calling \fct{setBaseCreator} creates a hidden object \code{.Random.seed.clrng}, which (unlike the standard \code{.Random.seed}) controls the creation of new streams (not random numbers).
 We allow users to call \fct{setBaseCreator}  at any time.
@@
 
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:uniformkernel},caption= Uniform random numbers generation kernel.]
 __kernel void mrg31k3pMatrix(
@@
 This is a safer way than the previous one for reproducing results in simulations.
-Below shows the code that saves streams to a data file called \code{myStreams.rds} on CPU and then load it back and transfer it to a `vclMatrix' \code{streams_saved} on GPU, and using it to (re)generate some Normal random numbers.
+Below we show the code that saves streams to a data file called \code{myStreams.rds} on CPU and then load it back and transfer it to a `vclMatrix' \code{streams_saved} on GPU, and using it to (re)generate some Normal random numbers.
 <<save streams,echo=TRUE,eval=FALSE>>=
@@
 
-\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},breaklines=true, escapeinside={(*@}{@*)}]
+\begin{lstlisting}[language=C,basicstyle=\small,label={lst:normal},breaklines=true, escapeinside={(*@}{@*)}, caption = Normal random numbers generation kernel.]
 __kernel void mrg31k3pMatrix(
@@
 Below we generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}.
-The best performance we have seen using \fct{rnormGpu} is above over 100 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}.
+The best performance we have seen using \fct{rnormGpu} is above 170 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}.
 The difference in elapsed time becomes larger when matrix size increases.
@@
   \item \code{type}: ``double'' or ``float'' format of generated random numbers.
-  \item \code{verbose}: Print extra information if $\text{verbose} > 1$. Default is FALSE.
+  \item \code{verbose}: Print extra information if $\text{verbose} > 1$. Default is set to 0.
 \end{itemize}
@@
  @
-
+ <<echo=FALSE,eval=TRUE>>= 
+sn <- function(x,digits)
+{
+  if (x==0) return("0")
+  ord <- floor(log(abs(x),10))
+  x <- x / 10^ord
+  if (!missing(digits)) x <- format(x,digits=digits)
+  if (ord==0) return(as.character(x))
+  return(paste(x,"\\times 10^{",ord,"}",sep=""))
+}
+@
  <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h', message=FALSE,purl=FALSE>>=
@@
 @
-We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
-\fct{clrng::fisher.sim} takes about 0.31 seconds, \pkg{clrng} accounts for 2.17\% of the elapsed time on CPU.
+We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $\Sexpr{sn(result_month$p.value,3)}$, which is close enough to the p-value from \fct{stats::fisher.test}.
+\fct{clrng::fisher.sim} takes about 0.3 seconds, \pkg{clrng} accounts for 2.1\% of the elapsed time on CPU.
  <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
@@
 @
-
 <<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
@@
 @
-The ``week'' table has a much smaller p-value: around 0.000125, which should require a larger number of simulations to get a more accurate p-value.
-With more than ten million simulations, we get 1205 cases and a p-value around
-\Sexpr{resultWeek$p.value}
-with \fct{fisher.sim}. \fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU. 
 <<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
@@
 @
-
 <<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
@@
 @
+The ``week'' table has a much smaller p-value: around $\Sexpr{sn(result_weekcpu$p.value,3)}$, which should require a larger number of simulations to get a more accurate p-value.
+With more than ten million simulations, we get 1350 cases and a p-value around $\Sexpr{sn(resultWeek$p.value,3)}$ with \fct{fisher.sim}.
+\fct{stats::fisher.test} takes about 91.4 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
 
@@
   var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.2,    10.2),
-  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.8,  91.5),
+  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.8,  91.4),
   var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.3,   1.9),
@@
 where $K_\kappa(\cdot)$  is a modified Bessel function of the second kind with order $\kappa$ and $\Gamma(\cdot)$ is the standard gamma function.  The model parameters are:
-\vspace{-\topsep}
+%\vspace{-\topsep}
 \begin{itemize}
-\setlength{\parskip}{0pt}
-\itemsep0em 
+%\setlength{\parskip}{0pt}
+%\itemsep0em 
 %\setlength{\itemsep}{0.5pt plus 1pt}
@@
 \end{itemize}
-\vspace{-\topsep}
+%\vspace{-\topsep}
 The standard isotropic Mat\'ern model is obtained by setting $\omega = 1$.
@@
 <<swissRainBoundary, echo=TRUE>>=
-library('geostatsp')
 data("swissRain", package="geostatsp")

### 2022-01-17 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
last update on paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3118 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 This is a safer way than the previous one for reproducing results in simulations.
-Below shows the code that saves streams to a data file called \code{myStreams.rds} on CPU and then load it back and transfer it to a `vclMatrix' \code{streams_saved} on GPU, and using it to (re)generate some Normal random numbers.
+Below we show the code that saves streams to a data file called \code{myStreams.rds} on CPU and then load it back and transfer it to a `vclMatrix' \code{streams_saved} on GPU, and using it to (re)generate some Normal random numbers.
 \begin{CodeChunk}
@@
 ##    user  system elapsed 
-##   0.060   0.011   0.071
+##   0.045   0.004   0.049
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   7.006   0.777   7.782
+##   6.615   0.828   7.437
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.318   0.000   0.318
+##   0.311   0.010   0.320
 \end{CodeOutput} 
@@
 
-We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}.
-\fct{clrng::fisher.sim} takes about 0.31 seconds, \pkg{clrng} accounts for 2.17\% of the elapsed time on CPU.
+We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $4.03\times 10^{-1}$, which is close enough to the p-value from \fct{stats::fisher.test}.
+\fct{clrng::fisher.sim} takes about 0.3 seconds, \pkg{clrng} accounts for 2.1\% of the elapsed time on CPU.
 
@@
 ##    user  system elapsed 
-##  14.859   0.014  14.874
+##  14.852   0.031  14.873
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.405
+## [1] 0.403
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.925   0.027   1.951
+##   1.934   0.015   1.947
 \end{CodeOutput}
@@
 
-The ``week'' table has a much smaller p-value: around 0.000125, which should require a larger number of simulations to get a more accurate p-value.
-With more than ten million simulations, we get 1205 cases and a p-value around
-0
-with \fct{fisher.sim}. \fct{stats::fisher.test} takes 92.45 seconds, while \fct{fisher.sim} takes about 1.95 seconds, the GPU elapsed time is decreased to about 2.10\% of the time taken on CPU.
-
 \begin{CodeChunk}
@@
 ##    user  system elapsed 
-##  91.311   0.186  91.432
+##  91.312   0.139  91.367
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.000127
+## [1] 0.000121
 \end{CodeOutput} 
 \end{CodeChunk}
+
+The ``week'' table has a much smaller p-value: around $1.21\times 10^{-4}$, which should require a larger number of simulations to get a more accurate p-value.
+With more than ten million simulations, we get 1350 cases and a p-value around $1.35\times 10^{-4}$ with \fct{fisher.sim}.
+\fct{stats::fisher.test} takes about 91.4 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
 
@@
 \hline
-\hspace{1em}10M & 327.3 & 10.2 & 91.5 & 1.9 & week\\
+\hspace{1em}10M & 327.3 & 10.2 & 91.4 & 1.9 & week\\
 \hline
@@
 \begin{CodeInput}
-R> library("geostatsp")
 R> data("swissRain", package = "geostatsp")

### 2022-01-17 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update format
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3119 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE, tidy=FALSE>>=
+# using GPU
  streams <- createStreamsGpu(n = 256*64)
@@
 @
-We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $\Sexpr{sn(result_month$p.value,3)}$, which is close enough to the p-value from \fct{stats::fisher.test}.
+We got 409429 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $\Sexpr{sn(result_month$p.value,3)}$, which is close enough to the p-value from \fct{stats::fisher.test}.
 \fct{clrng::fisher.sim} takes about 0.3 seconds, \pkg{clrng} accounts for 2.1\% of the elapsed time on CPU.
@@
  <<timecomparemonth4Cpu, cache=TRUE, eval=TRUE>>=
- #using CPU
+ # using CPU
  system.time(result_monthcpu<-stats::fisher.test(month,simulate.p.value = TRUE, B=1015808))
@@
 <<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
+# using GPU
 week_GPU<-gpuR::vclMatrix(week,type="integer")
@@
 <<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
-#using CPU
+# using CPU
 system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
@@
 With more than ten million simulations, we get 1350 cases and a p-value around $\Sexpr{sn(resultWeek$p.value,3)}$ with \fct{fisher.sim}.
-\fct{stats::fisher.test} takes about 91.4 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
+\fct{stats::fisher.test} takes about 91.8 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
 
@@
   var3 = c('AMD Radeon', 0.403507, 0.0001274,   2.2,    10.2),
-  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 14.8,  91.4),
+  var4 = c('Intel 3.7ghz', 0.4035606,0.0001202, 15.0,  91.8),
   var5 = c('NVIDIA V100', 0.403507,0.0001274,   0.3,   1.9),
@@
 <<fighistMonth, eval=TRUE, echo=FALSE, dev='pdf', fig.cap = "\\label{fig4}Approximate sampling distributions of the test statistics from the two examples Month and Week.
-The test statistic values of the observed tables are marked with a blue vertical line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1,fig.pos='H'>>=
+The test statistic values of the observed tables are marked with a blue vertical line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1,fig.pos='t'>>=
 hist(month_stats, xlab="test statistic",#TeX('-$\\sum(\\log(n_{ij}!))$'),

### 2022-01-17 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update format
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3119 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##    user  system elapsed 
-##   0.045   0.004   0.049
+##   0.044   0.004   0.049
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   6.615   0.828   7.437
+##   6.686   0.799   7.478
 \end{CodeOutput} 
@@
 \begin{CodeInput}
+R> # using GPU
 R> streams <- createStreamsGpu(n = 256*64)
@@
 ##    user  system elapsed 
-##   0.311   0.010   0.320
+##   0.335   0.014   0.348
 \end{CodeOutput} 
@@
 
-We got 410825 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $4.03\times 10^{-1}$, which is close enough to the p-value from \fct{stats::fisher.test}.
+We got 409429 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $4.03\times 10^{-1}$, which is close enough to the p-value from \fct{stats::fisher.test}.
 \fct{clrng::fisher.sim} takes about 0.3 seconds, \pkg{clrng} accounts for 2.1\% of the elapsed time on CPU.
@@
 ##    user  system elapsed 
-##  14.852   0.031  14.873
+##  15.019   0.016  15.024
 \end{CodeOutput}
@@
 \begin{CodeInput}
+R> # using GPU
 R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
@@
 ##    user  system elapsed 
-##   1.934   0.015   1.947
+##   1.934   0.014   1.947
 \end{CodeOutput}
@@
 ##    user  system elapsed 
-##  91.312   0.139  91.367
+##  91.768   0.108  91.828
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-## [1] 0.000121
+## [1] 0.000129
 \end{CodeOutput} 
@@
 
-The ``week'' table has a much smaller p-value: around $1.21\times 10^{-4}$, which should require a larger number of simulations to get a more accurate p-value.
+The ``week'' table has a much smaller p-value: around $1.29\times 10^{-4}$, which should require a larger number of simulations to get a more accurate p-value.
 With more than ten million simulations, we get 1350 cases and a p-value around $1.35\times 10^{-4}$ with \fct{fisher.sim}.
-\fct{stats::fisher.test} takes about 91.4 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
+\fct{stats::fisher.test} takes about 91.8 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
 
@@
 \hline
-\hspace{1em}1M & 49.3 & 2.2 & 14.8 & 0.3 & month\\
+\hspace{1em}1M & 49.3 & 2.2 & 15 & 0.3 & month\\
 \hline
-\hspace{1em}10M & 327.3 & 10.2 & 91.4 & 1.9 & week\\
+\hspace{1em}10M & 327.3 & 10.2 & 91.8 & 1.9 & week\\
 \hline
@@
 
-\begin{figure}[H]
+\begin{figure}[t]
 

### 2022-01-17 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update format
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3120 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 )
-
-knitr::kable(dt, col.names = NULL, caption = "Summary of comparions of Fisher's test simulation on different devices.
+knitr::kable(dt, col.names = NULL, digits = c(7, 7, 1, 1), caption = "Summary of comparions of Fisher's test simulation on different devices.
 Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII.

### 2022-01-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
typo p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3122 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2022-01-17 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
@@
 Then we provide two real data examples to demonstrate the function usage and its \proglang{R} performance.
-In Section 5, we show a useful application of normal random numbers on GPU, using them to simulate batches of Gaussian random surfaces with Mat\'ern covariance matrices simultaneously, the simulation also uses GPU-enabled functions from our other package \pkg{gpuBatchMatrix}.
+In Section 5, we show a useful application of Normal random numbers on GPU, using them to simulate batches of Gaussian random surfaces with Mat\'ern covariance matrices simultaneously, the simulation also uses GPU-enabled functions from our other package \pkg{gpuBatchMatrix}.
 Finally, the paper concludes with a short summary and a discussion in Section 6.
@@
 In general, a stream object contains three elements: the current state of the stream, the initial state of the stream (or seed).
-Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by $Z$ steps to find its initial state, and the three states in the stream object are set to it.
+Streams are created sequentially in the way that whenever the user creates a new stream, the software automatically jumps ahead by $Z$ steps to find its initial state, and the two states in the stream object are set to it.
 
@@
 The object returned is either a \code{vclMatrix} or \code{vclVector}, depending on whether \code{n} is two- or one-dimensional.
-Reusing \code{myStreamsGpu} will produce a vector different form \code{sim_1}, as the current states of the streams in \code{myStreamsGpu} have advanced by two positions (each of the 4 work item generated 2 random numbers).
+Reusing \code{myStreamsGpu} will produce a vector different from \code{sim_1}, as the current states of the streams in \code{myStreamsGpu} have advanced by two positions (each of the 4 work item generated 2 random numbers).
 
@@
 We summarized the comparison results in Table \ref{tab:summary} and plot the test statistics in Figure \ref{fig4}.
-<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE,fig.pos='H'>>=
+<<summarycompare, echo=FALSE, eval=TRUE, cache=TRUE,fig.pos='h'>>=
 library(kableExtra)
@@
 <<fighistMonth, eval=TRUE, echo=FALSE, dev='pdf', fig.cap = "\\label{fig4}Approximate sampling distributions of the test statistics from the two examples Month and Week.
-The test statistic values of the observed tables are marked with a blue vertical line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1,fig.pos='t'>>=
+The test statistic values of the observed tables are marked with a blue vertical line on each plot.", fig.subcap = c('Month data', 'week data'), out.width="0.47\\textwidth", margins=1,fig.pos='h'>>=
 hist(month_stats, xlab="test statistic",#TeX('-$\\sum(\\log(n_{ij}!))$'),

### 2022-01-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
update format
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3120 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
typo p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3122 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ##               [,1]       [,2]       [,3]       [,4]
-##  [1,] 12345  336690377  502033783  739421137
-##  [2,] 12345  597094797 1322587635 1475938232
-##  [3,] 12345 1245771585 1964121530  730262207
-##  [4,] 12345   85196284 1949818481 1630192198
-##  [5,] 12345  523477687 1607232546  324551134
-##  [6,] 12345 2094976052 1462898381  795289868
-##  [7,] 12345  336690377  502033783  739421137
-##  [8,] 12345  597094797 1322587635 1475938232
-##  [9,] 12345 1245771585 1964121530  730262207
-## [10,] 12345   85196284 1949818481 1630192198
-## [11,] 12345  523477687 1607232546  324551134
-## [12,] 12345 2094976052 1462898381  795289868
+## current.g1.1 12345  336690377  502033783  739421137
+## current.g1.2 12345  597094797 1322587635 1475938232
+## current.g1.3 12345 1245771585 1964121530  730262207
+## current.g2.1 12345   85196284 1949818481 1630192198
+## current.g2.2 12345  523477687 1607232546  324551134
+## current.g2.3 12345 2094976052 1462898381  795289868
+## initial.g1.1 12345  336690377  502033783  739421137
+## initial.g1.2 12345  597094797 1322587635 1475938232
+## initial.g1.3 12345 1245771585 1964121530  730262207
+## initial.g2.1 12345   85196284 1949818481 1630192198
+## initial.g2.2 12345  523477687 1607232546  324551134
+## initial.g2.3 12345 2094976052 1462898381  795289868
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.044   0.004   0.049
+##   0.044   0.005   0.049
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   0.335   0.014   0.348
+##   0.317   0.000   0.318
 \end{CodeOutput} 
@@
 ##    user  system elapsed 
-##   1.934   0.014   1.947
+##   1.942   0.015   1.955
 \end{CodeOutput}
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
 

### 2022-07-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
typo p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3123 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3181 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   \item \code{type}: \code{double}, \code{float} or \code{integer}, the format of generated random numbers, defaults to \code{double}.
-  \item \code{verbose}: Print extra information if $\text{verbose} > 1$.
+  \item \code{verbose}: Print extra information if $\text{verbose} > 1$, default is set to 0.
 \end{itemize}
@@
  # colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
- month <- as.matrix(readRDS(system.file("extdata", "month.Rds", package = "clrng")))
+ #month <- as.matrix(readRDS(system.file("extdata", "month.Rds", package = "clrng")))
+ data("month")
+ month <- as.matrix(month)
  @
@@
 # colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
-week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
+#week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
+data("week")
+as.matrix(week)
 @

### 2022-07-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3182 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<timecomparemonth1Gpuresult>>=
-unlist(result_month[c('threshold','simNum','counts')])
-result_month$p.value
+unlist(result_month[c('p.value','threshold','simNum','counts')])
 @
@@
                 type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
-unlist(resultWeek[c('threshold','simNum','counts')])
-resultWeek$p.value
+unlist(resultWeek[c('p.value','threshold','simNum','counts')])
 @

### 2023-07-12 pkg/mapmiscTerra/R/gridlinesWrap.R
tpeqd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3228 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -35,19 +35,20 @@ gridlinesWrap = function(crs,
 	glines[glines$type == 'east' & glines$neg, 'direction'] = 'W'
 	terra::values(glines)$ID = paste0(glines$degrees, glines$direction)
 
-	glinesT = wrapPoly(x=glines, crsT)
-
 	if(!all(c('ellipse','crop') %in% names(attributes(crsT)))) {
 		ellipseAndCrop = llCropBox(crsT)
 		attributes(crsT)$ellipse = ellipseAndCrop$ellipse
 		attributes(crsT)$crop = ellipseAndCrop$crop
 	} 
 
+	glinesT = wrapPoly(x=glines, crsT)
+
+
 	
-	ellipseSmall =  attributes(crsT)$ellipse
-	if(!is.null(ellipseSmall)) {
-		glinesT = terra::crop(glinesT, ellipseSmall)
-	}
+#	ellipseSmall =  attributes(crsT)$ellipse
+#	if(!is.null(ellipseSmall)) {
+#		glinesT = terra::crop(glinesT, ellipseSmall)
+#	}
 
 	
 	legendPoints = terra::centroids(glinesT, inside=TRUE)

### 2023-07-21 pkg/mapmiscTerra/R/moll.R
vignettes work
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3230 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -114,7 +114,7 @@ moll = function(x=0, angle=NULL, flip=FALSE) {
 
 
 		fromOptim = optim(
-			c(lon_p=0,lat_p=90, lon_0minuslon_p=0),
+			c(lon_p=midX,lat_p=midY, lon_0minuslon_p=0),
 			testFunCrs,
 			lower = c(-180, -90, -100), upper = c(180, 90, 100),
 			method = 'L-BFGS-B',

### 2023-07-21 pkg/mapmiscTerra/R/openmap.R

@@ -243,7 +243,6 @@ openmap = function(
 
     newNumberOfCells = fact*NtestCols*sqrt(areaRatio)
 
-
     outraster = rast(outExtent, res = (terra::xmax(outExtent) - terra::xmin(outExtent))/newNumberOfCells, crs = crsOut)
   } # end not merc
 

### 2023-07-21 pkg/mapmiscTerra/R/scaleBar.R

@@ -58,7 +58,7 @@ scaleBar = function(crs,
 	
 # how long (in m) is our target dash
 	dashTargetDist = as.matrix(terra::distance(xll))[1,2]
-	desiredLen = strwidth("x")*forLegend$seg.len 
+	desiredLen = strwidth("m")*forLegend$seg.len 
 	
 	theb = log10(dashTargetDist)
 	candidates = 10^c(floor(theb), ceiling(theb))

### 2023-07-25 pkg/mapmiscTerra/R/gridlinesWrap.R
perspective
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3232 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -35,22 +35,16 @@ gridlinesWrap = function(crs,
 	glines[glines$type == 'east' & glines$neg, 'direction'] = 'W'
 	terra::values(glines)$ID = paste0(glines$degrees, glines$direction)
 
-	if(!all(c('ellipse','crop') %in% names(attributes(crsT)))) {
-		ellipseAndCrop = llCropBox(crsT)
-		attributes(crsT)$ellipse = ellipseAndCrop$ellipse
-		attributes(crsT)$crop = ellipseAndCrop$crop
+	if(!all(c('ellipse','crop') %in% names(attributes(crsOrig)))) {
+		ellipseAndCrop = llCropBox(crsOrig)
+		attributes(crsOrig)$ellipse = ellipseAndCrop$ellipse
+		attributes(crsOrig)$crop = ellipseAndCrop$crop
 	} 
 
-	glinesT = wrapPoly(x=glines, crsT)
+	glinesT = wrapPoly(x=glines, crsOrig)
 
+	glinesT = glinesT[terra::perim(glinesT)>0]
 
-	
-#	ellipseSmall =  attributes(crsT)$ellipse
-#	if(!is.null(ellipseSmall)) {
-#		glinesT = terra::crop(glinesT, ellipseSmall)
-#	}
-
-	
 	legendPoints = terra::centroids(glinesT, inside=TRUE)
 
 	# at most three labels per line
@@ -63,7 +57,6 @@ gridlinesWrap = function(crs,
 
 	okPoints = c(okPoints, unlist(lapply(split(toTrim$index, toTrim$ID), function(xx) xx[seq(from=1, len=3, by=pmax(1,floor(length(xx)/3)))] )))
 
-	legendPoints[okPoints]
 	legendPoints$minDist = apply(terra::distance(legendPoints, legendPoints), 2, function(xx) min(xx[xx>0]))
 
 

### 2023-07-25 pkg/mapmiscTerra/R/tpers.R

@@ -1,31 +1,42 @@
 # http://proj4.org/projections/tpers.html
 tpers = function(x,
   hKm = 100*1000, tilt = -10,
+  azi,
   offset=c(0,0), 
   axis='enu') {
+
+  if(is.vector(x)) x = vect(matrix(x, ncol=2), crs=crsLL)
   
   if(!terra::is.lonlat(x)) x = project(x, crsLL)
+
+
+  if(missing(azi)) {
+  if(length(x)==1) {
+    azi = 0
+  } else {
+    azi = geosphere::bearing(terra::crds(x)[1,],terra::crds(x)[2,])[1]    
+  }
+  }
 
   myCrs = crs(paste(
       "+proj=tpers +h=",
       hKm*1000,
       " +lat_0=",
-      x@coords[1,2],
+      terra::crds(x)[1,2],
       " +lon_0=",
-      x@coords[1,1],
-      " +azi=",
-      geosphere::bearing(x@coords[1,],x@coords[2,])[1],
+      terra::crds(x)[1,1],
+      " +azi=", azi,
       " +tilt=", tilt,
       " +ellps=WGS84 +axis=", axis,
       " +x_0=", offset[1],
       " +y_0=", offset[2],
       sep=""))
   
-  cropBox = llCropBox(crs=myCrs)
+  cropBox = llCropBox(crs=myCrs, crop.poles=FALSE, remove.holes=FALSE,  buffer.width=10*1000, densify.interval = 2000)
   
   attributes(myCrs)$ellipse = cropBox$ellipse
   attributes(myCrs)$crop = cropBox$crop
-   attributes(myCrs)$regionLL = cropBox$regionLL
+   attributes(myCrs)$regionLL = terra::fillHoles(cropBox$crop)
 
   
   myCrs

### 2023-07-25 pkg/mapmiscTerra/R/moll.R
perspective
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3235 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -95,7 +95,7 @@ moll = function(x=0, angle=NULL, flip=FALSE) {
 
 		testFunCrs = function(param, testPoints) {
 			crsHere = crsFromNumeric(param)
-			xxx = try(crds(project(testPoints, crsHere)), silent=TRUE)
+			xxx = try(terra::crds(project(testPoints, crsHere)), silent=TRUE)
 			if(all(class(xxx) == 'try-error')) xxx = matrix(NA, 2,2)
 				xxx = xxx/1e6
 	# x coords the same, x point close to origin, dest point above x
@@ -144,8 +144,8 @@ moll = function(x=0, angle=NULL, flip=FALSE) {
 		angleOpposite = Arg(exp(1i*2*pi*(midX + 180)/360))*(360/(2*pi))
 		densify.interval = 10*1000
 		theLine = terra::densify(vect(rbind(cbind(angleOpposite,-89.9), cbind(angleOpposite, 89.9)), type='lines', crs=crsLL), densify.interval)
-		theBox$crop = terra::buffer(theLine, width=buffer.width)
-		theBox$regionLL = terra::erase(terra::as.polygons(ext(-100,180,-90,90), crsLL), theBox$crop)
+		theBox$crop = terra::buffer(theLine, width=50*1000)
+		theBox$regionLL = terra::erase(terra::as.polygons(terra::ext(-100,180,-90,90), crsLL), theBox$crop)
 	} else {
 		theBox = llCropBox(result, ellipse = theBox$ellipse, 
 			buffer.width=75*1000, densify.interval = 20*1000, 

### 2023-07-26 pkg/mapmiscTerra/R/omerc.R
omerc
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3236 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -60,7 +60,8 @@ omercProj4string = function(
 omerc = function(
     x, angle=0, 
     post=c('none', 'north', 'wide','tall'),
-    preserve=NULL
+    preserve=NULL,
+    ellipse=TRUE
 ) {
   
 
@@ -370,10 +371,33 @@ omerc = function(
     attributes(rotatedCRS)$obj = objectiveResult
   }
   
-  theBox = llCropBox(rotatedCRS, ellipse = NULL, 
-      buffer.width=75*1000, densify.interval = 20*1000, 
+
+  equatorMeetsGreatCircle = geosphere::gcIntersectBearing(theCentre, angle, c(0,0), 90)
+  if(any(is.na(equatorMeetsGreatCircle)))
+  equatorMeetsGreatCircle = geosphere::gcIntersectBearing(theCentre, angle, c(0,0), -90)
+  if(any(is.na(equatorMeetsGreatCircle))) {
+    theY = 0
+  } else {
+  equatorMeetsGreatCircle = vect(matrix(equatorMeetsGreatCircle, ncol=2, byrow=TRUE), crs=crsLL)
+  equatorMeetsGreatCircleProj = project(equatorMeetsGreatCircle, rotatedCRS)
+  theY = terra::crds(equatorMeetsGreatCircleProj)[,2]
+  theY = theY[which.min(abs(theY))]
+  }
+
+
+#  theEllipse = terra::as.polygons(ext(c(-1.6e7, 1.6e7, -2e7, 2e7) + theY*c(0,0,1,1)), rotatedCRS)
+  theEllipse = terra::as.polygons(terra::ext(as.vector(extentMerc) + theY*c(0,0,1,1)), rotatedCRS)
+
+
+  if(ellipse) {
+  theBox = llCropBox(crs=rotatedCRS, ellipse = theEllipse, 
+      buffer.width=40*1000, densify.interval = 20*1000, 
         crop.distance = Inf, crop.poles = FALSE, crop.leftright=TRUE,
       remove.holes=TRUE)
+} else {
+  theBox = list(ellipse=theEllipse)
+}
+#  plot(project(worldMap, crsLL));plot(theBox$crop, col='red', add=TRUE)   
 
   attributes(rotatedCRS)[names(theBox)] = theBox
 

### 2023-07-28 pkg/mapmiscTerra/tests/openmap.R
package names changed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3253 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,6 +1,6 @@
 #+ setup
 
-library('mapmiscTerra')
+library('mapmisc')
 
 #if(!exists('mapmiscCachePath'))
 #  mapmiscCachePath = system.file('extdata', package='mapmisc')

### 2023-07-28 pkg/diseasemapping/tests/bym.R
working
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3254 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -77,8 +77,10 @@ colExc = colourScale(kBYM$data$exc1 ,
 # and try passing a data frame and adjacency matrix
 
 	
-adjMat = spdep::poly2nb(kentucky, row.names =as.character(kentucky$County) )
-kBYM = bym(data=kentucky@data, formula=observed ~ offset(logExpected) + poverty,
+adjMat = terra::adjacent(kentucky)
+attributes(adjMat)$region.id = kentucky$County
+
+kBYM = bym(data=terra::values(kentucky), formula=observed ~ offset(logExpected) + poverty,
 		adjMat = adjMat, region.id="County",
 		priorCI = list(sdSpatial=c(0.1, 5), sdIndep=c(0.1, 5)))
 

### 2023-07-28 pkg/diseasemapping/R/bym.R
fixes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3257 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -98,8 +98,6 @@ bym.needAdjmat = function(
  	adjMatNB = terra::adjacent(data)
  	attributes(adjMatNB)$region.id = unname(unlist(data[[region.id]]))
 
- 	stuff <<- adjMatNB
-
 	methods::callGeneric(
 			formula=formula, data=data,
 			adjMat=adjMatNB, region.id=region.id,

### 2023-07-28 pkg/diseasemapping/tests/bym.R

@@ -133,7 +133,7 @@ kBYM = bym(
 		formula = observed ~ offset(logExpected) + povertyFac,
 		data = kentucky, 
 		region.id="County",
-		priorCI = list(sdSpatial=c(0.1, 5), sdIndep=c(0.1, 5))
+		prior = list(sd=c(0.1, 0.5), propSpatial=c(0.1, 0.5))
 )
 
 

### 2023-08-17 pkg/geostatsp/R/0gm.R
@data gone
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3307 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -391,7 +391,7 @@ gm.dataSpatial = function(
     }
 
     # ensure row names are identical
-    rownames(values(data)) = rownames(crds(data)) = 
+    rownames(terra::values(data)) = rownames(terra::crds(data)) = 
         1:length(data)
   # loop through covariates which aren't in data, extract it from `covariates`
     for(D in setdiff(alltermsPlain, names(data))){

### 2023-08-17 pkg/geostatsp/R/krige.R
errors in lgm fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3308 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -201,7 +201,7 @@ krigeLgm = function(
 	 } # end data is spdf	
   
 	 
-	 if(!length(observations) | is.null(meanRaster)) {
+	 if(!length(observations) | is.null(meanRaster)) { # old code, not called from lgm
 		  # the above didn't create observations and meanRaster
 		  # use the old code, probably not being called from lgm
  	  
@@ -642,8 +642,8 @@ krigeLgm = function(
 		  names(result)[names(result)=="predict"] = "predict.boxcox"
 		  
     bcpred = meanBoxCox(
-      pred=values(result[['predict.boxcox']]), 
-      sd=values(result[['krigeSd']]),
+      pred=values(result[['predict.boxcox']], mat=FALSE, dataframe=FALSE), 
+      sd= values(result[['krigeSd']], mat=FALSE, dataframe=FALSE),
       boxcox=param['boxcox']
     )
     

### 2023-09-15 pkg/mapmiscTerra/R/crsRegion.R
tpers stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3368 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -3,17 +3,22 @@ crsRegionEllipse = function(x, offset) {
   # need this here because of unknown global variable warning in package checks
   isoSp = vect(mapmisc::isohedron, crs=crsLL)
   isoT = project(isoSp, x)
+  if(missing(offset)) {
+    offset = apply(matrix(ext(isoT), ncol=2), 2, mean)  
+  }
   isoT = crds(isoT) 
   isoT = isoT - matrix(offset, nrow(isoT), 2, byrow=TRUE)
+
+
   isoTc = isoT[,1] + 1i*isoT[,2]
-  isoTc = cbind(abs(Arg(isoTc)), Mod(isoTc))
-  isoTc = isoTc[isoTc[,2] > quantile(isoTc[,2], 0.8),]
+  isoTc = cbind(Arg(isoTc), Mod(isoTc))
+#  isoTc = isoTc[isoTc[,2] > quantile(isoTc[,2], 0.1),]
 
   axisX = max(isoTc[,2])
 
   rfun = function(theta, a, b) sqrt((b*cos(theta))^2 + (a*sin(theta))^2)
   objFun = function(param, x=isoTc) {
-    sum(pmax(0, x[,2] -rfun(x[,1], param[1], param[2]))^2) + sum(log(param))
+    100*sum(pmax(0, x[,2] -rfun(x[,1], param[1], param[2]))^2) + sum(log(param))
   }
   oo = stats::optim(c(axisX, axisX), objFun, x=isoTc)
 
@@ -21,5 +26,5 @@ crsRegionEllipse = function(x, offset) {
   angleSeq = seq(0, 2*pi, len=1001)
   ellipseP = cbind(cos(angleSeq)*oo$par[2], sin(angleSeq)*oo$par[1])
 
-  ellipseP
+  ellipseP + matrix(offset, nrow(ellipseP), 2, byrow=TRUE)
 }

### 2023-09-18 pkg/diseasemapping/tests/bym.R
blas threads
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3371 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -4,6 +4,14 @@ havePackages = c(
 
 print(havePackages)
 
+if(havePackages) {
+	INLA::inla.setOption(num.threads=2)
+	INLA::inla.getOption(blas.num.threads=2)
+	print(INLA::inla.getOption('num.threads'))
+	print(INLA::inla.getOption('blas.num.threads'))
+}
+
+
 library('diseasemapping')
 data('kentucky')
 kentucky = terra::unwrap(kentucky)
@@ -26,8 +34,7 @@ if(all(havePackages)){
 			adjMat = kentuckyAdjMat,
       prior = list(sd=c(0.1, 0.5), propSpatial=c(0.5, 0.5)),
 			region.id='County',
-			control.predictor=list(compute=TRUE),
-			num.threads=2
+			control.predictor=list(compute=TRUE)
   )
 	kBYM$parameters$summary
 
@@ -46,8 +53,7 @@ if(all(havePackages)){
 kBYM = bym(
 		formula = observed ~ offset(logExpected),
 		prior = list(sd = 0.5, propSpatial = 0.5),
-		data=kentucky,
-			num.threads=2)
+		data=kentucky)
 
 
 
@@ -81,8 +87,7 @@ kBYM = bym(
     formula=observed ~ offset(logExpected) + poverty,
     data=terra::values(kentucky)[-(1:4),],  
  	  adjMat = kentuckyAdjMat, region.id="County",
-		prior = list(sd=0.1, propSpatial = 0.1),
-			num.threads=2)
+		prior = list(sd=0.1, propSpatial = 0.1))
  
 
 kBYM$par$summary
@@ -91,8 +96,7 @@ kBYM$par$summary
 
 
 kBYM = bym(data=kentucky,  formula=observed ~ 1,
-		prior = list(sd=1, propSpatial=0.9),
-			num.threads=2)
+		prior = list(sd=1, propSpatial=0.9))
 
 kBYM$par$summary
 
@@ -116,8 +120,7 @@ kentucky[1:2,-grep("County", names(kentucky))] = NA
 kBYM = bym(observed ~ offset(logExpected) + poverty,
 		kentucky, 
 		region.id="County",
-		prior = list(sd=0.1, propSpatial = 0.5),
-			num.threads=2)
+		prior = list(sd=0.1, propSpatial = 0.5))
 
  
 kBYM$par$summary
@@ -135,8 +138,7 @@ kBYM = bym(
[…]

### 2023-12-06 pkg/mapmiscTerra/inst/extdoc/mapLayers.Rmd
maplayers list
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3448 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Examples of map layers}
-->

# Examples of map layers

This document will be incomplete if there is no
internet connection when this document is compiled.  The full document is at
[diseasemapping.r-forge.r-project.org](http://diseasemapping.r-forge.r-project.org).

See [mc.bbbike.org/mc](http://mc.bbbike.org/mc/) for examples more map tiles 
(not all of which are compatible with `mapmisc`).

```{r setup, include=FALSE}
library('knitr')
library('mapmisc')

knitr::knit_hooks$set(plot=Pmisc::hook_plot_htmlsubfig)
knitr::opts_chunk$set(dev='png')

fact = 0.5
allTiles = FALSE

# uncomment for nicer images
# fact = 1 
# reRun = TRUE
# or run in sandbox	
# R --slave -e "knitr::knit('../pkg/mapmisc/vignettes/mapLayers.Rmd')" fact=1 allTiles=TRUE
# pandoc --to=html5 --output=mapLayers.html mapLayers.md
# cp -r figure ../www
# cp -r mapLayers.html ../www
theArgs = grep("^(fact|allTiles)=", commandArgs(), value=TRUE)
for(D in theArgs)
  eval(parse(text=D))

opts_chunk$set(dpi=50*fact, 
  dev.args=list(bg='white'))

cachePath = "/tmp/cacheLayers/"
dir.create(cachePath, showWarnings=FALSE)

theNames = names(osmTiles())

```

```{r theNames}
theNames
```

```{r theExtent, include=FALSE}
africaExtent = c(25, 25)
trafalgarSquare=c(-0.128056,51.508056)
stjohns=c(-52.683320, 47.566285)
```

```{r downloadTiles,include=FALSE}

  theNames = grep("sputnik", theNames, invert=TRUE, value=TRUE)
  print(theNames)
  print(cachePath)

  africaTiles = mapply(openmap, path=theNames,
    MoreArgs = list(x=africaExtent, zoom=2,
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath))

  londonTiles = mapply(openmap, path=theNames,
    MoreArgs = list(x=trafalgarSquare,zoom=12, 
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath))

  nfTiles = mapply(openmap, path=theNames,
    MoreArgs = list(x=stjohns, zoom=14,
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath))

```

The following are some examples of maps produced with `mymap=openmap(x, path="something")`

```{r plotSetup, include=FALSE}
mytiles = list(africa=africaTiles, london=londonTiles, newfoundland = nfTiles)
theCap = rep(theNames, each=length(mytiles))
```

```{r someTiles, echo=FALSE, fig.ncol=3, out.width='32%', fig.cap = 'some tiles', fig.height=4, fig.width=4, dpi=120, fig.subcap = theCap}

for(D in theNames){

# cat("\n\n# ", D, "\n\n")

      for(Dplot in names(mytiles)) {
        map.new(mytiles[[Dplot]][[D]])
        plot(mytiles[[Dplot]][[D]],add=TRUE, maxcells = 1e12)
    }
}
```

# Attributions

```{r attributionsExample, echo=TRUE}
openmapAttribution('osm', type='text', short=TRUE)
openmapAttribution('osm', type='text', short=FALSE)
```

```{r theAttributions, echo=FALSE, results='asis'}
for(D in theNames) {
  cat("## ", D, '\n\n')
    cat(openmapAttribution(D, 'html', short=TRUE), '\n\n')
    cat(openmapAttribution(D, 'html', short=FALSE), '\n\n')
}

```

```{r someTilesOld, eval=FALSE, echo=FALSE, results='asis', fig.height=5, fig.width=5}

for(D in theNames){
  Dsub = gsub("-",".", D)

  threeTiles = length(
    grep(
      paste("^", Dsub, "([rR]ed|[gG]reen|[bB]lue)$", sep=""), 
      names(mytiles[[1]])
    )
  ) == 3

  cat("\n\n# ", D, "\n\n")
  cat("`plot", "RGB"[threeTiles], 
    "(openmap(x,path=\"", D, "\"))`\n\n", 
    sep='')
  cat(openmapAttribution(D, 'html'), '\n\n')

  cat("\n|     |     |     |\n|-----|-----|-----|\n")
  for(Dplot in 1:length(mytiles)) {
    cat("|")

    oneTileName = grep(paste("^", Dsub, "$", sep=""), 
      names(mytiles[[Dplot]]), value=TRUE)
    threeTilesName = grep(paste("^", Dsub, 
        "([rR]ed|[gG]reen|[bB]lue)$", sep=""), 
      names(mytiles[[Dplot]]),value=TRUE)

    if(length(threeTilesName)==3) { # plotRGB
      map.new(mytiles[[Dplot]])

      plotRGB(mytiles[[Dplot]][[
          paste(Dsub, c("Red","Green","Blue"),sep="")
        ]],add=TRUE)

    } else if (length(oneTileName)==1){

      map.new(mytiles[[Dplot]])
      plot(mytiles[[Dplot]][[oneTileName]],add=TRUE)

    } else {

      plot(0.5, 0.8)
      text(0.5, 0.8, paste(D, "is\n not currently\navailable"))

    }

  }
  cat("|")

}

```

### 2023-12-06 pkg/mapmiscTerra/R/openmapAttribution.R

@@ -7,10 +7,13 @@ esriAttribution = function(name) {
       'esri'='Esri, HERE, DeLorme, USGS, Intermap, increment P Corp., NRCAN, Esri Japan, METI, Esri China (Hong Kong), Esri (Thailand), TomTom, MapmyIndia, copyright OpenStreetMap contributors, and the GIS User Community',
       'esri-grey' = 'Esri, HERE, DeLorme, MapmyIndia, copyright OpenStreetMap contributors, and the GIS user community',
       'esri-topo' = 'Esri, HERE, DeLorme, TomTom, Intermap, increment P Corp., GEBCO, USGS, FAO, NPS, NRCAN, GeoBase, IGN, Kadaster NL, Ordnance Survey, Esri Japan, METI, Esri China (Hong Kong), swisstopo, MapmyIndia, copyright OpenStreetMap contributors, and the GIS User Community',
-      'esri-transport' = 'Esri, HERE, DeLorme, MapmyIndia, copyright OpenStreetMap contributors'
+      'esri-transport' = 'Esri, HERE, DeLorme, MapmyIndia, copyright OpenStreetMap contributors',
+      'esri-satellite' = 'Esri, Earthstar Geographics and others',
+      'esri-natgeo'=  'Esri, National Geographic and others',
+      'esri-overlay' = 'Esri'
   )
   
-  short = 'Esri, OpenStreetMap, and others'
+  short = 'Esri and others'
   
   if(!length(grep("^http", name))){
     name = gsub("esri\\.", "esri-", name)
@@ -112,6 +115,7 @@ odl = list(
     ),
     text='Data by OpenStreetMap.org available under the Open Database License (opendatacommons.org/licenses/odbl)'
 )
+
 
 # openstreetmap.org
 osm = list(long=list(
@@ -241,6 +245,17 @@ osmLandscape$long$text = gsub("cartography[[:print:]]+$",
 osmLandscape$long$html = gsub("cartography[[:print:]]+$",
     "cartography by <a href=\"http://www.thunderforest.com/\">Thunderforest</a>",
     osmLandscape$long$html)
+ 
+ f4map = list(
+    short = list(text = "f4map.com", 
+        latex = "\\href{http://www.f4map.com}{f4map}",
+        html='<a href="http://f4map">f4map</a>'),
+    long = list(test = "f4map.com",
+           latex = "\\href{http://www.f4map.com}{f4map}",
+        html='<a href="http://f4map">f4map</a>',
+        markdown='[f4map](http://www.f4map.com)'
[…]

### 2023-12-06 pkg/mapmiscTerra/inst/extsrc/landsat.R

@@ -1,0 +1,227 @@
+#+ getLandsatTilesFunction
+
+getLandsatTiles = function(x,
+  tiles = getLandsatCoverage(),
+  buffer = 1000,
+  crsOut = raster::projection(x)) {
+  
+  xExtent = 
+    raster(extend(extent(x), buffer), 
+      ncol=10, nrow=10, crs=projection(x))
+  
+  xExtentProj = projectExtent(xExtent, projection(tiles))
+  
+  tilesCrop = raster::crop(
+    tiles, xExtentProj
+  )
+  tilesCrop2 = spTransform(tilesCrop, projection(xExtent))
+  tilesCrop3 = raster::crop(tilesCrop2, xExtent)
+  
+  tiles = tiles[tiles$PR %in% tilesCrop3$PR, ]
+  
+  if(identical(crsOut, 'landsat'))
+    crsOut = paste(
+      '+proj=lsat +path=',
+      tiles$PATH[1],
+      ' +lsat=5 +ellps=WGS84 +datum=NAD83', sep='')
+  
+  if(identical(crsOut, 'omerc'))
+    crsOut = mapmisc::omerc(
+      tiles, 
+      angle=seq(-25,25,by=0.1)
+    )
+  
+  
+  tilesProjCrop = spTransform(tiles, crsOut)
+  
+  tilesProjCrop$pPath = paste('p',  sprintf("%03d", tilesProjCrop$PATH), sep='')
+  tilesProjCrop$rRow = paste('r',  sprintf("%03d", tilesProjCrop$ROW), sep='')
+  
+  tilesProjCrop$pathString = sprintf("%03d", tilesProjCrop$PATH)
+  tilesProjCrop$rowString = sprintf("%03d", tilesProjCrop$ROW)
+  
+  tilesProjCrop$tileString = paste(
+    tilesProjCrop$pathString,
+    tilesProjCrop$rowString, sep='')
+  
+  tilesProjCrop$tTile = paste(
+    tilesProjCrop$pPath,
+    tilesProjCrop$rRow, sep=' ')
+  
+  attributes(tilesProjCrop)$landsat = xExtent
+  tilesProjCrop
+}
+
+getLandsatScenes = function(
+  timeRange,
+  tiles,
+  sceneList = getLandsatSceneList()
+) {
+  
+  if(any(names(tiles) == 'tileString')) {
+    tiles = tiles$tileString
+  }
+  
[…]

### 2024-02-09 pkg/mapmiscTerra/tests/openmap.R
files again
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3483 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -62,11 +62,8 @@ points(myPoints)
   # raster, result will be in project of the raster (long-lat)
   mytiles = openmap(
     x=extend(myraster,1),
-    zoom=thezoom, filename=NULL)
-  mytiles2 = openmap(
-    x=extend(myraster,1),
-    zoom=thezoom, filename=tempfile(fileext='.tif'))
-#		mycities = GNcities(extend(myraster,1),max=5)
+    zoom=thezoom)
+
   myplot(myraster, myPoints)
 
 print(2)

### 2024-02-13 pkg/Pmisc/R/markdownFigureArray.R
Pmisc new stuff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3498 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,4 +1,6 @@
- hook_plot_array  = function (x, options) {
+#' @rdname hook_plot_htmlsubfig
+#' @export
+hook_plot_array  = function (x, options) {
     # fig.tabset is a vector
     base = knitr::opts_knit$get("base.url")
     if (is.null(base)) 
@@ -18,7 +20,7 @@
 
     fig.names= options$fig.names
 
-        fig.num = options$fig.num
+    fig.num = options$fig.num
     if (is.null(fig.num)) 
         fig.num = 1
         

### 2024-02-14 pkg/Pmisc/inst/extdoc/document.Rmd
rename folders
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3502 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2019-02-12 pkg/Pmisc/inst/docs/document.Rmd
@@
 
-```{r aplot, fig.ncol=2, fig.cap='Uninteresting plots', fig.subcap=c('stuff','qq','eeee','www')}
+```{r aplot, fig.ncol=2, fig.cap='Uninteresting plots', fig.subcap=c('stuff','qq','eeee','www'), out.width=Pmisc::out.width(0.45)}
 plot(1:10)
@@
 
-```{r anotherplot, fig.ncol=3, fig.cap='Other uninteresting plots', fig.subcap=month.name[1:6], out.width=Pmisc::out.width(0.32)}
+```{r anotherplot, fig.ncol=3, fig.cap='Other uninteresting plots', fig.subcap=month.name[1:6], out.width=Pmisc::out.width(0.31)}
 

### 2024-04-04 pkg/mapmiscTerra/R/openmap.R
mapmisc now passes checks without internet connection
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3511 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -372,10 +372,9 @@ openmap = function(
       cachePath = cachePath),
     silent=!verbose)
 
-  if(any(class(result)=="try-error")){
+  if(all(is.na(values(result)))){
     message(paste(Durl, "not accessible"))
     # create an empty raster
-    result = outraster
     attributes(result)$openmap = list(
       tiles=NA,
       message=result,
